### 2018-05-22 chapters/01-introduction/intro.tex
Init commit with original master thesis

A typical workflow of a scientist who writes a scientific publication is to use a \gls{wp} to write the publication and one or more \gls{cas} for Verification, Analysis and Visualization, among other tasks. his formulae. Especially in \gls{stem}, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}. However, since \LaTeX{} (an extension of \TeX) is the standard to write in \gls{stem} and we only focus on mathematical writings in this thesis, we categorize \LaTeX{} as a \gls{wp} as well.} has become the de facto standard for writing scientific publications over the last 30 years~\cites{LATEX:Standard}[559]{DigitalTypo}{Knuth}. \LaTeX{} enables printing of mathematical formulae in a structure similar to handwritten style. For example, consider a Jacobi polynomial~\cite[(18.3), table 1]{NIST:DLMF}
\begin{equation}\label{eq:P}
P_n^{(\alpha , \beta)}(\cos(a\Theta)).
\end{equation}
This formula is written in \LaTeX{} as
\begin{equation}\label{eq:P-tex}
\verb|P_n^{(\alpha,\beta)}(\cos(a\Theta))|.
\end{equation}

While \LaTeX{} focuses on displaying mathematics, \gls{cas} focus on computations and user friendly syntax. Therefore, each system (such as \LaTeX{}, \Maple{} or \Mathematica) uses its own representation and syntax. Hence, a writer needs to constantly translate mathematical expressions from one representation to another and back again. Table~\ref{tab:JacobiP-usecase} shows four different representations for the same formula~(\ref{eq:P}).

\begin{table}[ht]
	\centering
	\begin{tabular}{ll}
		\hline
		Systems & Representations \\
		\hline
		\hline
		Generic \LaTeX\ & \verb|P_n^{(\alpha,\beta)}(\cos(a\Theta))| \\ 
		Semantic \LaTeX\ & \verb|\JacobiP{\alpha}{\beta}{n}@{\cos@{a\Theta}}| \\
		\Maple & \verb|JacobiP(n,alpha,beta,cos(a*Theta))| \\ 
		\Mathematica & \verb|JacobiP[n,\[Alpha],\[Beta],Cos[a \[CapitalTheta]]]|\\
		\hline
	\end{tabular}
	\caption{Different representations for the same mathematical formula~(\ref{eq:P}).}
	\label{tab:JacobiP-usecase}
\end{table}

Such translations can become complex, if there are non-obvious differences between the used systems. For example, the \gls{cas} could use a different order of the arguments, define a normalization of the function for better performances or even use different domains for the variables. Multivalued functions are particularly difficult~\cite{AISC:MultivaluedFunctions}. A \gls{cas} usually defines so called \textit{branch cuts} to compute principal values of multivalued functions~\cite{Maple:Cuts}. Thereby, implementations of multivalued functions are discontinuous. In general, positioning branch cuts follow some conventions. They are mostly defined conveniently, such as a straight line. However, since a branch cut is a curve, it also can be defined as a spiral or it can be defined for a different position in the complex plane. The position of branch cuts can therefore varies from \gls{cas} to \gls{cas}~\cite{Branches:acot}. Figure~\ref{fig:acot-cut-compare} illustrates two examples of different branch cut positioning for the inverse trigonometric arccotangent function. While \Maple{} defines the branch cut at ${[-\infty\iunit, -\iunit]}$, ${[\iunit,\infty\iunit]}$ (figure~\ref{fig:acot-cut1}), \Mathematica{} defines the branch cut at ${[-\iunit, \iunit]}$ (figure~\ref{fig:acot-cut2}).

\begin{figure}[ht]
    \centering
    \subfloat[The real part of arccotangent with a branch cut at ${[-\infty\iunit, -\iunit]}$, ${[\iunit,\infty\iunit]}$.]{%
        \includegraphics[width=0.45\textwidth]{acotCut1.png}
        \label{fig:acot-cut1}%
    }
    \hspace{0.5cm}
    \subfloat[The real part of arccotangent with a branch cut at ${[-\iunit, \iunit]}$.]{%
        \includegraphics[width=0.45\textwidth]{acotCut2.png}
        \label{fig:acot-cut2}%
    }
    \caption{Two plots for the real part of the arccotangent function with a branch cut at $[-\infty\iunit, -\iunit]$, $[\iunit,\infty\iunit]$ in figure~\protect\subref{fig:acot-cut1} and at $[-\iunit, \iunit]$ in figure~\protect\subref{fig:acot-cut2}, respectively. (Plotted with \Maple{} 2016)}
    \label{fig:acot-cut-compare}
\end{figure}

Hence, a \gls{cas} user needs to fully understand the properties and special definitions (such as the position of branch cuts) in the \gls{cas} to avoid mistakes during a translation~\cite{Maple:Cuts}. In consequence, a manual translation process is not only laborious, but also prone to errors. Note that this general problem has been named to automatic \gls{p2c} conversion~\cite{POM-Tagger}.

An automatic translation process is desirable. Providing translations for mathematical \LaTeX{} expressions is difficult, because the semantic information is absent. Semantics is the meaning of an expression. Consider the example of the Jacobi polynomial in~(\ref{eq:P}). A scientific reader might be able to conclude information like $P$ indicating the Jacobi polynomial and $n$ is being a non-negative integer. However, for a computer it is difficult to understand the expression~(\ref{eq:P-tex}). Since \LaTeX{} is the de facto standard for mathematical expressions, most \gls{cas} provide import and export functions for \LaTeX{} expressions~\cite{Maple:ImportExport,Mathematica:ImportExport,Matlab:ImportExport,Sage:ImportExport}. However, because of the explained difficulties, those functions are very limited. In fact, they only work for expressions with unique semantic information, but fail for more complex expressions, such as our Jacobi polynomial example.

The \gls{nist} has therefore developed a set of semantic \LaTeX{} macros~\cite{DLMF:Macros}. Each of these macros is tied with a unique well-defined mathematical object and linked to the corresponding definition in the \gls{dlmf}~\cite{NIST:DLMF,NIST:DLMF:Paper,NIST:Handbook}. The semantic macros are used to semantically enhance the \gls{dlmf}. An outgrowth of the \gls{dlmf} project is the \gls{drmf}~\cite{DRMF:14,DRMF:15}, which defines more semantic macros, especially for \gls{opsf}. In table~\ref{tab:JacobiP-usecase} we have already seen an example of a semantic macro. We call \LaTeX{} expressions with semantic macros \textit{semantic \LaTeX{}} and without \textit{generic \LaTeX}, respectively. With semantic \LaTeX, it is possible to provide translations to \gls{cas}.

Note that partial results of this Master's thesis have already been published at the 10$^\text{th}$ Conference on Intelligent Computer Mathematics 2017~\cite{CICM:Paper}. Techniques, approaches and results have been roughly introduced in the article and are discussed in more detail in this thesis. Pictures and ideas that have already been presented there are not explicitly referenced in the following.

Furthermore, note that the translator we will present in this thesis is based on not published code from the \gls{mlp} project and the \gls{pom}-tagger. We will also use the mentioned set of \Macro s, which are not published yet as well. Therefore, the current version of the translator is not online available yet. However, we are planning to publish the translator soon on the \gls{drmf} website\footnote{\url{http://drmf.wmflabs.org}}.

Moreover, I will use 'we' rather than 'I' in the subsequent chapters of this thesis, since I published and discussed my ideas with others including my advisors Dr.~Moritz Schubotz and Professor Abdou Youssef, the hosts during research visits Howard S.~Cohl and J\"urgen Gerhard, and the students Joon Bang and Kevin Chen.\footnote{Special thanks to M.~Schubotz to give me the permission for using this phrase from his doctoral thesis \cite{MORITZ:Dis}.}

\section{Goals of the project}
In this thesis, we will present an automatic tool for translations between semantic \LaTeX{} and the \gls{cas} \Maple{} and \Mathematica. The translation tool should be lite, in a sense that a user does not need a whole software package to use the program. It is also a part of the \gls{drmf} project and should provide interactive translations in the \gls{drmf}. Furthermore, it should provide additional information about the translation process and present solutions if an appropriate translation is not possible.

Another goal is the translator's extensibility in order to provide the basis for future work.

\section{Structure}
In the following chapters, we will discuss our translator and verification techniques. Prior to explaining implementation details, we need to introduce some basic background information in chapter~\ref{ch:background}. We start this chapter with an overview of related work (section~\ref{sec:related-work}). Multivalued functions and branch cuts, which were already mentioned, will be introduced in section~\ref{sec:math-background}. Besides the mathematical background, we will discuss semantics in \LaTeX{} in section~\ref{sec:semantics}. A brief introduction to grammar in languages and how we will use it to understand mathematical expressions will be given in section~\ref{sec:mlp}. The chapter is closed by some definitions for the following chapters.

The next chapter~\ref{ch:translator} is the main chapter of the thesis. We will explain implementations and discuss problems and solutions of and for the translator. It starts with an outline of the goals and our approaches in section~\ref{sec:gen-appr}. While we will present implementation details for the forward translations in section~\ref{sec:forward-translation}, we will present the developed backward translation in the following section~\ref{sec:backward-translation}.

In chapter~\ref{ch:evaluation} we will discuss, how we can verify or validate a translated expression. The chapter presents some approaches, such as round trip tests (section~\ref{sec:round-trip}) and numerical tests (section~\ref{sec:numerical-tests}). Extracted formulae from the \gls{dlmf} and \gls{drmf} created a comprehensive test suite for these approaches. Section~\ref{sec:test-summary} gives a summary of the results.

The results and the current status of the translation tool will be discussed in chapter~\ref{ch:results}. In conclusion, chapter~\ref{ch:conc-future-work} explains some approaches to further improve the explained techniques in future work.

### 2018-05-22 chapters/02-background/cas.tex
Init commit with original master thesis

\section{Computer Algebra Systems}\label{sec:cas}
\glsreset{cas}
A \gls{cas} is a mathematical software, which is, for example, able to analyze, simplify and compute mathematical expressions. It is based on the theory of computer algebra. Computer Algebra is a part of computer science that designs, analyzes, implements and applies algebraic algorithms~\cite{ComputerAlgebra}. In 1963, \textit{Schoonschip} was one of the first \gls{cas} developed by Martinus J. G. Veltman~\cite{Schoonschip}. It was mainly used for high-energy physics.

Using \gls{cas} was highly desirable, because it was a faster and easier alternative to calculation tables~\cite{Tables}. Another \gls{cas} \textit{MATLAB}\footnote{\textit{MATLAB} is an abbreviation for \textit{\textbf{mat}rix \textbf{lab}oratory}.} was initially developed by Cleve Moler in the late 1970s to allow students to use Fortran\footnote{Originally \textit{FORTRAN}, was the first widely used high-level programming language for general purposes and is still very popular for high-performance computing.} libraries without learning Fortran itself~\cite{MATLAB}. \textit{MATLAB} was constantly updated\footnote{Newest version is \textit{MATLAB 9.2} from march 2017.} and is still widely used for education and in image processing, because of its performance in linear algebra and numerical analysis.

Nowadays, there are several different \gls{cas} available. From proprietary solutions, such as \textit{MATLAB}, \Maple, \Mathematica, to freely available software packages, such as \textit{SageMath}. Most \gls{cas} are more or less specialized for fields in mathematics, such as \textit{MATLAB} for linear algebra and numerical computations. However, nowadays the most popular \gls{cas} are \Mathematica{} and \Maple.

\subsection{Maple}\label{subsec:maple}
\Maple{} was initially released in 1982 and is written in \textit{C}, \textit{Java} and its internal programming language, also called \textit{Maple}. The main purposes was to create an \gls{cas}, which would be able to run on lower cost computers. \Maple{} bases on a small kernel, which is written in \textit{C}, that provides the \Maple{} language. Additional libraries are written in the \Maple{} language, while the \gls{gui} is written in \textit{Java}.

In this thesis, we refer to version \textit{Maple 2016} and all information about internal data structures and the usage of \Maple{} is described in the manual~\cite{MAPLE:ProgrammingGuide}.

Internally, symbolic expressions are stored in the memory as \gls{dag} data structures. This avoids multiple blocks in the memory for the same object in an expression.

Furthermore, \Maple{} allows different input notations: the \texttt{1D} input and the \texttt{2D} input. Consider the infinite integral
\begin{equation}\label{eq:maple-input}
\int\displaylimits_0^\infty \frac{\pi+\sin(2x)}{x^2} dx.
\end{equation}
The corresponding \texttt{1D} input notation in \Maple{} is
\begin{equation}\label{eq:1dmaple-input}
\texttt{int((Pi+sin(2*x))/x\^{}2, x=0..infinity)}.
\end{equation}

Historically, the \texttt{1D} format was used for inputting commands and expressions into \Maple, while the \texttt{2D} input is more used for interaction with human users. In fact, the internal \texttt{2D} representation is similar to \gls{mathML}~\cite{MathML} and its display is similar to \eqref{eq:maple-input}.

We consider \Maple's \texttt{1D} input as syntactically correct and our translation process always refers to the \texttt{1D} input representation rather than the \texttt{2D} input representation.

Another interesting characteristic, which becomes import for our translator, is determinism.

\begin{definition}[Deterministic Algorithm]
An algorithm is called \textbf{deterministic} if it produces always the same results on two identical inputs.
\end{definition}

\Maple{} is in general non-deterministic. That means, that two identical inputs possibly produce two different outputs depending on the run time environment. The reason is, that for example the order of the elements of a set\footnote{Mathematically, the elements in a set are not ordered. This is technically different in computer algebra, because each element is somewhere located in the memory. In a scope of a low-level programming environment, the elements of sets always have an order.} depends on the addresses of the elements in the memory. Mostly the background operation systems determine the positions of variables in the memory rather than the application \Maple{} does.

However, \Maple{} works in sessions. When a user starts a worksheet a new \Maple{-session} will be created. It is also possible to start a new session in a worksheet with the command \texttt{restart}. Within such sessions \Maple{} is deterministic as long as the input doesn't use non-deterministic algorithms, such as probabilistic algorithms\footnote{Probabilistic algorithms use randomness to find approximate solutions. Such algorithms are obviously non-deterministic.}.

\subsection{Mathematica}\label{subsec:mathematica}
\textit{Wolfram Mathematica} was conceived by Stephen Wolfram and initially released in 1988~\cite{Mathematica}. It is also written in \textit{C/C++}, \textit{Java} and its internal programming language \textit{Wolfram Language}. It is one of the most comprehensive \gls{cas} nowadays and an ever growing software. The previously mentioned interactive document \gls{cdf} uses \Mathematica. Also the knowledge engine\footnote{Also known as answer engine, is a computer science discipline which tries to automatically answer questions posed by humans in natural language.} \textit{Wolfram|Alpha}\footnote{\url{https://www.wolframalpha.com/}, seen 08/2017} is based on \Mathematica.

In 2008 an error~\cite{Mathematica:Bug} in \Mathematica{} shows a problem of confidence in \gls{cas}. As we will explain later, our translator uses symbolic simplification\cite[section 2]{ComputerAlgebra} techniques in \Maple. Implemented symbolic simplification algorithms are rarely\footnote{In fact, there is no proof of correctness for symbolic simplifications in any \gls{cas} known.} proven and may therefore produce errors. We can only presume that the simplifications are correct.

### 2018-05-22 chapters/02-background/definitions.tex
Init commit with original master thesis

\section{Translator Definitions}\label{sec:definition}
To explain translations mathematically, we need to define some techniques first. Mainly, a translation process is a function that maps an input expression to an output expression. The input and output expressions are written in two different formal languages. For a translation, the exact types of input and output language are insignificant. Therefore, we consider mathematical \LaTeX, semantic \LaTeX, \Maple's \texttt{1D} input and \Mathematica{} input as formal languages.

\begin{definition}[Translation]
Let $\mathcal{L}$ be a set of formal languages and $L_1, L_2 \in \mathcal{L}$ be two formal languages. A \textbf{translation} $\mathrm{tr}$ is a mapping from an input language $L_1$ to an output language $L_2$, denoted as
\begin{equation*}
\mathrm{tr}^{L_1}_{L_2}: L_1 \rightarrow L_2.
\end{equation*}
\end{definition}

\begin{definition}[Forward \& Backward Translation]
Let $\mathcal{L}$ be a set of formal languages. Let $L_1 \in \mathcal{L}$ be the language of semantic \LaTeX{} and $L_2 \in \mathcal{L}$ the language of any \gls{cas}. The function $\mathrm{tr}^{L_1}_{L_2}$ is called \textbf{forward translation to $L_2$} and $\mathrm{tr}^{L_2}_{L_1}$ is called \textbf{backward translation from $L_2$} respectively.
\end{definition}

Therefore, we define $\langTex$ as the formal language of semantic \LaTeX, $\langMaple$ as the formal language of \Maple{} and $\langMathe$ as the formal language of \Mathematica. Since we will only translate expressions between $\langTex$ and a \gls{cas}, we will use a more convenient notation for forward and backward translations. Let $t \in \langTex$ be an arbitrary expression in semantic \LaTeX{}. A forward translation to \Maple{} is denoted as
\begin{equation}
t \overset{\langMaple}{\mapsto} c, \quad \text{with } c \in \langMaple,
\end{equation}
and a backward translation from \Maple{} as
\begin{equation}
t \overset{\langMaple}{\mapsfrom} c, \quad \text{with } c \in \langMaple.
\end{equation}

Note that there is an identity translation between two formal languages $L_1, L_2$ defined if there is a word $w \in L_1 \cap L_2$. We specially denote the identity translation with 
\begin{align}
&\mathrm{id}^{L_1}_{L_2}:\ L_1 \rightarrow L_2,\notag\\
&\mathrm{id}^{L_1}_{L_2}(w) := w,\quad \text{with } w \in L_1 \cap L_2.
\end{align}

In this thesis, we will translate expressions from one system to another. There is no formal definition to compare mathematical expressions in two different systems. As an example, consider the trigonometric cosine function $\cos@{z}$. There is a formal definition in the \gls{dlmf} given~\cite[(4.14.2)]{NIST:DLMF}
\begin{equation}
\cos@{z} = \frac{\expe^{\iunit z} + \expe^{-\iunit z}}{2}.
\end{equation}
The internal definition of the cosine function in \Maple{} is given by
\begin{equation}
\cos@{z} = \frac{1}{2}\expe^{\iunit z} + \frac{1}{2\expe^{\iunit z}}.
\end{equation}
Technically, we would agree that both definitions are equivalent. However, \Maple{} is limited by the possibilities of a computer system. For example, the mathematical constant $\expe$ is irrational and can therefore never be \textit{absolutely equivalent} to the abstract mathematical construct of Euler's number. Instead of defining a new equivalence term for our purpose, we conveniently look for an \textbf{appropriate} translations between two formal languages. Obviously, this is not a formal definition, but should prevent us from translating one mathematical object\footnote{We are aware that it could be even more difficult to formally define what a \textit{mathematical object} is. We consider a mathematical object to be a function or symbol with a formal mathematical definition behind it.} into a completely different object in another language.

Therefore, we would label a translation such as
\begin{equation}\label{eq:cos-def}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|cos(z)|
\end{equation}
as \textit{appropriate}, while a translation such as
\begin{equation}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|sin(z)|
\end{equation}
would be \textit{inappropriate}. However, note that it is not always as easy as in this example to decide if a translation is appropriate or not. Therefore, we developed validation techniques to automatically check if a translation can be considered appropriate. In addition to this terminology, we introduce \textit{direct translations}. Later in the thesis, we will explain that a translation from one specific mathematical object to its \textit{appropriate} counterpart in the other system is not possible. We call a translation to the \textit{appropriate} counterpart \textbf{direct}. For example, the translation~(\ref{eq:cos-def}) is \textit{direct}, while a translation to the definition of the cosine function
\begin{equation}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|(exp(I*z)+exp(-I*z))/2|
\end{equation}
is not a \textit{direct} translation.

### 2018-05-22 chapters/02-background/dlmf.tex
Init commit with original master thesis

\section{Digital Library of Mathematical Functions}\label{sec:dlmf}
\glsreset{dlmf}\glsreset{drmf}\glsreset{opsf}
The \gls{nist} \gls{dlmf}~\cite{NIST:DLMF:Paper} is the digital version of the \textit{NIST Handbook of Mathematical Function}. Internally, the \gls{dlmf} websites are \LaTeX{} files and all equations are numbered as usual. Therefore, we reference to a formula in the \gls{dlmf} with the given equation number.

The \gls{nist} \gls{drmf}~\cite{DRMF:14} is an outgrowth of the \gls{dlmf} and wants to create a digital compendium of mathematical formulae for \gls{opsf}. One of the main purposes is to facilitate interaction among a community of mathematicians and scientists~\cite{DRMF:15}. The \gls{drmf} also aims at expandability and interactivity. That was also a reason for this translation project, because a goal was to provide a one-click translation on the formulae in the \gls{drmf}.

Together, the \gls{dlmf} and \gls{drmf} build a comprehensive compendium for \gls{opsf} and they are therefore our main references in this thesis. A definition of a function in any \gls{cas} is therefore called different, if it differs from the corresponding \gls{dlmf}/\gls{drmf} definition.

### 2018-05-22 chapters/02-background/intro.tex
Init commit with original master thesis

Before we start to explain the translation process, we need to introduce some basic information. We will see that it is not as easy as it seems to explain what semantic information is, that there can be hidden complexity even behind simple mathematical expressions, and why we need grammar to understand mathematics. All this requires a comprehensive introduction.

In this chapter we will lay the foundation to understand the complexity and our solutions in this project. The first section~\ref{sec:related-work} talks about the ideas and problems of related work. In the following section~\ref{sec:math-background}, we will give an introduction to complex analysis, branch cuts and special functions in general. Sections~\ref{sec:dlmf} and \ref{sec:semantics} will introduce the \gls{dlmf} and semantic \LaTeX, which is the most important part for our translations to \gls{cas}. Section~\ref{sec:cas} will give a brief introduction to \gls{cas} and mainly discusses \Maple{} and \Mathematica. The \gls{mlp} is another important part of the translation process, because it gives us the opportunity to parse mathematical expressions. Section~\ref{sec:mlp} explains how this works and why this is so important for a translator. This chapter finishes then with some extra definitions for the rest of the thesis.

Later on we will focus on the translation process between mathematical expressions in semantic \LaTeX{} and in \gls{cas}, where the source of semantic \LaTeX{} is the \gls{dlmf} and the source for \gls{cas} expressions is the corresponding \gls{cas}. Because there is no umbrella term for those representations from different sources, we hereafter call them \textit{systems}. A translation between semantic \LaTeX{} and \Maple{} is therefore in general a translation between two different systems.

### 2018-05-22 chapters/02-background/math-background.tex
Init commit with original master thesis

\section{Mathematical Background}\label{sec:math-background}
To understand the following chapters in the thesis, we need a brief introduction in complex analysis and multivalued functions. The following mathematical introduction is based on the book \textit{Introduction to Complex Analysis} by H.~A.~Priestley~\cite{ComplexAna:Intro}.

A complex number $z \in \Complex$ is specified by a two real numbers $x$ and $y$ as $z = x+y\iunit$, with $\iunit$ as the imaginary unit, which is defined as a solution of $\iunit^2 = -1$. The two real numbers are called the real part $x = \Re(z)$ and the imaginary part $y = \Im(z)$ of $z$. It is convenient to represent complex numbers as points in a $\Real^2$ plane, which is called \textbf{complex plane}. We identify $z = x+y\iunit$ as $(x,y) \in \Real^2$ in the complex plane. It is sometimes useful to represent complex numbers by polar coordinates and write $x = r\cos@@{\phi}$ and $y = r\sin@@{\phi}$, with $r \geq 0$ and $\phi \in \Real$. Since $\cos@@{\phi}+\sin@@{\phi}\iunit$ can be written as $\expe^{\iunit\phi}$, we write $z = x+y\iunit$ in polar coordinates as $z = r\expe^{\iunit\phi}$. The angle $\phi$ is often called the \textbf{phase} (or argument) of $z$ and $r$ is the absolute value of $z$. Similar to $\Re(z)$ for the real part and $\Im(z)$ for the imaginary part of $z$, we can access $r = |z|$ and $\phi = \ph@{z}$.\footnote{Note that the phase of $z$ is often called the \textbf{argument} of $z$. Therefore, in CAS the phase function is often the \texttt{argument} function.}

\begin{figure}[ht]
    \centering
    \subfloat[Four color phase mapping.]{%
        \includegraphics[width=0.28\textwidth]{dlmfColorMap.png}
        \label{fig:four-color}%
    }
    \hspace{0.4cm}
    \subfloat[Continues phase mapping in \DLMF.]{%
        \includegraphics[width=0.28\textwidth]{dlmfColorMapCont.png}
        \label{fig:cont-color}%
    }
    \hspace{0.4cm}
    \subfloat[Continues phase mapping in \Maple.]{%
        \includegraphics[width=0.28\textwidth]{mapleColorMap.png}
        \label{fig:maple-color}%
    }
    \caption{Three different domain coloring methods. Mappings~\protect\subref{fig:four-color} and \protect\subref{fig:cont-color} used in the \DLMF{} (Source: \DLMF{~}\cite{NIST:DLMF} in help page, \textit{About Color Map}). Mapping~\protect\subref{fig:maple-color} is the default coloring map in \Maple.}
    \label{fig:domain-coloring}
\end{figure}

Since the same basic arithmetical rules apply in $\Complex$ as in $\Real$, we will not introduce all arithmetical rules for complex numbers here. However, another interesting part are plots of complex functions. Consider a function $f: \Complex \mapsto \Complex$. This can be considered as $f: \Real^2 \mapsto \Real^2$, which shows us four real dimensions that need to be displayed in a three-dimensional space. The most common solution to plot those four dimensions is called domain coloring. With this technique, each value in the complex plane is represented by a color. There is no standard method to map complex values to colors. A simple method is to map a unique color for each of the quadrants in the complex plane (used in the DLMF). Another variation are continues phase mappings, which use color wheels. In those cases, $\phi$ represents the hue and optionally $r$ defines the intensity of the hue. The second variation is mostly used in \gls{cas}, but there is no standard for the orientation of the color wheel. Figure~\ref{fig:domain-coloring} draws three different examples.

This thesis mostly focus on the \gls{cas} \Maple, therefore all of the following complex plots use the default color map in \Maple{~}\ref{fig:maple-color}.

A three-dimensional plot of a complex function, such as $f$ from above, usually maps the real and imaginary part of the variable to the $x$ and $y$ axes, while the height is defined by the absolute value of the complex variable $|z|$. Figure~\ref{fig:cylinderU} demonstrates a complex plot of the parabolic cylinder function $\ParabolicU@{a}{z}$~\cite[(12.2i)]{NIST:DLMF} in \Maple.

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.8\textwidth]{cylinderUPlot.png}
	\caption{A 3-dimensional complex plot of the parabolic cylinder function $\ParabolicU@{a}{z}$ at $a=0$ plotted by \Maple.~\cite[(12.2i)]{NIST:DLMF}}
	\label{fig:cylinderU}
\end{figure}

Another way to plot such complex functions is to plot the real and imaginary parts of the solution separately. Figure~\ref{fig:cylinderUSplit} illustrates this plotting method for the parabolic cylinder function again.

Note that for real function plots, such separate plots for the imaginary and real values of a complex function, will use a color map as well. In that case, the color map is just an adornment and has nothing to do with complex color maps. Figure~\ref{fig:cylinderUSplit} illustrates the separate plots for real and imaginary values for the parabolic cylinder function again.

\begin{figure}[!htp]
    \centering
    \subfloat[Plot of the real value of $\ParabolicU@{0}{z}$.]{%
        \includegraphics[width=0.8\textwidth]{cylinderURe.png}
        \label{fig:cylinderURe}%
    }
    \vspace{-0.1cm}
    \subfloat[Plot of the imaginary value of $\ParabolicU@{0}{z}$.]{%
        \includegraphics[width=0.8\textwidth]{cylinderUIm.png}
        \label{fig:cylinderUIm}%
    }
    \caption{Separately plot of the real~\protect\subref{fig:cylinderURe} and imaginary~\protect\subref{fig:cylinderUIm} part of the parabolic cylinder function $\ParabolicU@{a}{z}$ at $a=0$.}
    \label{fig:cylinderUSplit}
\end{figure}

\subsection{Multivalued Functions}
The following definitions for multivalued functions are based on \textit{Complex Variables}~\cite{ComplexVariables} by M.~J.~Ablowitz and A.~S.~Fokas.

Unlike a typical\footnote{In lack of a better word to distinguish single-valued and multivalued functions, we call single-valued functions here typical functions, because these functions are best known.} function, multivalued functions associate not only one output to any particular input, but a whole set of output values. A real-valued example is the square root function, which associates for each positive real number two real square roots. For example\footnote{Note that the radical sign $\sqrt{}$\ is defined for the principal square roots, which are the unique non-negative square roots. Therefore, $\sqrt{4}$ is always $2$, and not $-2$. However, the solutions of the square root of a positive number $a$ are $\left\{+\sqrt{a},-\sqrt{a}\right\}$.}, the solution of $4^{\frac{1}{2}}$ is $\{+2,-2\}$.

Consider now the natural logarithm with complex arguments and $z = r\expe^{\iunit\phi}$ in polar coordinates. Since it is the natural logarithm, $\ln@{z} = \ln@{r} + \iunit\phi$ and the imaginary part is just $\phi$. Let $r = 1$, so that $z$ is on the unit circle for all $\phi$. Let be $z_1$ the complex number for $\phi = 0$ and $z_2$ the complex number for $\phi = 2\cpi$. Obviously $z_1 = z_2$ but $\ln@{z_1} \neq \ln@{z_2}$. Therefore, the natural logarithm is also multivalued and has an infinite number of solutions for $\ln@{z}$.

The center of the circle we choose (in that case the origin) is called branch point.

\begin{definition}[Branch Point]
A point $p$ is a \textbf{branch point} if the multivalued function $w(z)$ is discontinuous upon traversing an arbitrarily small circuit around $p$.
\end{definition}

The complex square root function has also a branch point in the origin. Consider
\begin{equation}
\sqrt{z} = \sqrt{r} \expe^{\frac{\iunit\phi}{2}},\quad \text{with } r \geq 0, \phi \in \Real.
\end{equation}
Indeed, at $\phi = 2\cpi$, the square root of $z$ is $\sqrt{r}\expe^{\iunit\cpi} = -\sqrt{r}$, but at $\phi = 0$ it is $\sqrt{r}$. Because $r$ was arbitrary, the origin is also a branch point of the complex square root function. Another branch point is $z = \infty$.

There exist two approaches to analytically study multivalued functions. One approach is to express the multivalued function as a single-valued function.

\subsubsection{Branch Cuts}\label{subsec:branch-cuts}
Consider a multivalued function in a restricted region and choose for every argument a value, such that the resulting function in the region is continues and single-valued. This regional function of the multivalued function is called \textbf{branch} of the multivalued function.

Consider the complex square root function again. If we cut out the region between both branch points, namely $(0, \infty)$, the function is continues and single-valued. This region is one branch of the function and the cut region $(0, \infty)$ is called \textbf{branch cut}. Another branch of the function can be reached by further increasing $\phi$, such that $\phi \in [2\cpi, 4\cpi)$. At $\phi = 4\cpi$ the square root function returns the same result as for $\phi=0$. Therefore, $\phi \in [4\cpi, 6\cpi)$ is again on the first branch. We can define a branch cut for the complex logarithm as well. As already mentioned, the complex logarithm has an infinite number of branches. Each of these branches can be reached by increasing or decreasing $\phi$ with a multiple of $2\cpi$. Consider $\phi+n2\cpi$ with an integer number $n$. The branch for $n=0$ is called the \textbf{principal branch}.

Defining branch cuts to allow analytic computations on the principal branches is a common task for \gls{cas}~\cite{Maple:Cuts}. Unfortunately, the position of the branch cuts is not well-defined. Indeed, consider $\phi \in [-\cpi,\cpi)$ instead of $\phi \in [0,2\cpi)$ and define the branch cut at $(-\infty, 0)$; than the branch is also continues and single-valued.

In general a branch cut is a curve which ends can be possibly open, half-open or closed. A branch cut can be also defined in a more complicated shape (e.g. spirals). Most of the multivalued functions have standard, convenient positions for the branch cuts, which are commonly accepted. The branch cut at $(-\infty, 0]$ for the complex square root function is one of those standard positions. However, the positions can vary from system to system~\cite{Branches:acot}.

Figure~\ref{fig:sqrtBranch} draws the principal branch of the imaginary part of the complex square root function in \Maple.

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.65\textwidth]{sqrtBranch.png}
	\caption{The imaginary part of the complex square root function with a branch cut at $(-\infty, 0)$. The resulted function is continues and single-valued for $\phi \in (-\cpi, \cpi)$.}
	\label{fig:sqrtBranch}
\end{figure}

The plot of the principal branch for the complex logarithm looks similar. The branch cut for the logarithm is in \Maple{} defined at $(-\infty,0)$. Figure~\ref{fig:lnBranch} illustrates the principal branch of the imaginary part of the logarithm.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.75\textwidth]{lnBranch.png}
	\caption{The imaginary part of the complex logarithm with a branch cut at $(-\infty, 0]$. The resulted function is continues and single-valued for $\phi \in (-\cpi, \cpi)$.}
	\label{fig:lnBranch}
\end{figure}

Note that in complex plots, a branch cut can also be located in the color map, we were talking about above. In the next example, we use the notation $\pm 0$ to come arbitrarily close to the branch cut from the \textit{upper} and \textit{lower} side. We presume the one-side limits is a well known notation and define
\begin{align}
-0 &:= \underset{x\nearrow 0}{\lim}x,\\
+0 &:= \underset{x\searrow 0}{\lim}x.
\end{align}

The natural logarithm $\ln@{z}$ has a branch cut at $(-\infty,0]$. We have the following values for the \textit{upper side} and \textit{lower side} of the branch cut at $\ln@{-1}$
\begin{align}
\ln@{-1+\iunit 0} &=\hphantom{-}\iunit \cpi,\label{eq:ln-up}\\
\ln@{-1-\iunit 0} &= -\iunit \cpi.\label{eq:ln-down}
\end{align}

Figure~\ref{fig:ln-compl-comp} shows two plots of the natural logarithm. The value of (\ref{eq:ln-up}) is represented by a shade of purple and (\ref{eq:ln-down}) is represented by a shade of green, with \Maple's continues phase color map from figure~\ref{fig:maple-color}. The branch cut can be seen by the changing of the color from purple directly to green on $(-\infty,0]$. The colors are counterparts of each other in the used color map.

\begin{figure}[!htp]
    \centering
    \subfloat[The complex plot for the natural logarithm $\ln@{z}$. The values of $\ln@{z}$ are visualized by colors. The coloring method is the continues phase mapping from \Maple.]{%
        \includegraphics[width=0.4\textwidth]{lnComp1.png}
        \label{fig:ln-color}%
    }
    \hspace{0.5cm}
    \subfloat[The complex plot for the natural logarithm $\ln@{z}$. The height is the absolute value. The coloring method is the continues phase mapping from \Maple.]{%
        \includegraphics[width=0.47\textwidth]{lnComp2.png}
        \label{fig:ln-compl}%
    }
    \caption{The complex plot of the natural logarithm $\ln@{z}$. Figure~\protect\subref{fig:ln-color} plots only the colors of $\ln@{z}$. The coloring method is \Maple's continues phase mapping from figure~\protect\ref{fig:maple-color}. The branch cut of $\ln@{z}$ is at $(-\infty,0)$ and the values $\ln@{1+\iunit 0} = \iunit \cpi$ (which is mapped to purple), while $\ln@{1-\iunit 0} = -\iunit \cpi$ (which is represented by green).}
    \label{fig:ln-compl-comp}
\end{figure}

An alternative representation for multivalued functions to branch cuts are Riemann surfaces.

\subsubsection{Riemann Surfaces}
Assume we would not cut the multivalued functions and allow multiple values for a particular input. For example, if we allow $\phi$ to constantly increase in the complex logarithm function, we end up in helix structure as plotted in figure~\ref{fig:lnRiemann}. This representation of a multivalued function is called the Riemann surface. Technically, a Riemann surface is a one-dimensional complex manifold.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=0.8\textwidth]{lnRiemann.png}
	\caption{The Riemann surface of the imaginary part of the natural logarithm.}
	\label{fig:lnRiemann}
\end{figure}

In the Riemann surface of logarithm in figure~\ref{fig:lnRiemann} we can see the multiple branches of the function. If we use the same method for the square root function, we get figure~\ref{fig:sqrtRiemann}, where we can see that the square root function has only two branches.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=0.9\textwidth]{sqrtRiemann.png}
	\caption{The Riemann surface of the imaginary part of the square root function.}
	\label{fig:sqrtRiemann}
\end{figure}

\subsection{Elementary \& Special Functions}\label{subsec:special-functions}
A function is called \textbf{elementary} if it can be constructed using only a finite combination of constant functions, field operations, and algebraic, trigonometric, exponential, and logarithmic functions together with their inverses~\cites{Elementary:1}[145]{Elementary:2}.

Unlike elementary functions, special functions cannot be listed easily. They are mostly solutions of differential equations or integrals of elementary functions. Almost all of them have established names and notations. Many of them are complex and some are also multivalued. Besides that, special functions are highly related to each other. Therefore, most special functions can be expressed by other special functions. However, there is no general formal definition, but there are lists and mathematical compendia collecting functions which are commonly accepted as special.

For example, in 1964 Milton Abramowitz and Irene A. Stegun published the \textit{Handbook of Mathematical Functions with Formulas, Graphs, and Mathematical Tables}~\cite{AbramowitzStegun} at \gls{nist}\footnote{In 1988 the \textit{National Bureau of Standards} (NBS) was renamed to the \textit{National Instutute of Standards and Technology} (NIST).}. The handbook became the most comprehensive source of information on special functions. Therefore, the \gls{nist} has published a new version called \textit{NIST Handbook of Mathematical Function}~\cite{NIST:Handbook} in 2010 to replace the predecessor.

Because there is no formal definition for special functions, some problems appear for our translator project. For example the \textit{Encyclop\ae{dia} Britannica} wrote:
\begin{myQuote}{Encyclop\ae{dia} Britannica - Special function~\cite{SpecialFunctions:britannica}}
	Special function, any of a class of mathematical functions that arise in the solution of various classical problems of physics. [...] Different scientists might not completely agree on which functions are to be included among the special functions, although there would certainly be very substantial overlap.
\end{myQuote}

In \textit{Special Functions}~\cite{SF:Book}, G.~E.~Andrews~et~al. wrote in the preface:
\begin{myQuote}{Special Functions - Preface~\cite{SF:Book}}
Paul Tur\'an once remarked that special functions would be more appropriately labeled "useful functions".
\end{myQuote}

Our reference for special functions in this thesis will be the \textit{NIST Handbook of Mathematical Function} and its digital counterpart the \gls{dlmf}~\cite{NIST:DLMF}. See the later section~\ref{sec:dlmf} for a more detailed explanation.

Since most of the special functions represent solutions for other mathematical expressions, they were intensively studied - mostly in applied mathematics. In the past, huge tables of values were used for calculations~\cite{Tables}. Therefore, the book by Abramowitz and Stegun also contains tables with values for several special functions~\cite{AbramowitzStegun}. Obviously, computations of special functions with \gls{cas} were desired. Therefore, \gls{cas} support numerous of special functions. However, all \gls{cas} contain its own ever growing lists of them. Sometimes they use the same definitions, domains and position of branch cuts as defined the handbooks, sometimes not. A translator needs to pay attention to those differences as well. Otherwise the translation process could produce unexpected effects. One of these effects can be produced by different positions of branch cuts in the systems. The next section will give some examples and explanations of these problems.

\subsection{Problems of Branch Cuts}\label{subsec:branch-cut-issues}
Obviously, different positions of branch cuts can cause problems, if we transfer a function from one system to the other. Those differences can even appear in well-known elementary functions. For example, the \gls{dlmf} defines the branch cut for the inverse cotangent function~\cite[(4.23.9)]{NIST:DLMF} at $[-\iunit, \iunit]$, while \Maple{} reasonably~\cite{Branches:acot} defines the branch cut at $(-\infty\iunit, -\iunit]$ and $[\iunit, \infty\iunit)$. Figure~\ref{fig:acot} illustrates the problem. A scientist, who is familiar with the definition for the branch cut by the \gls{dlmf} would assume an output such as in figure~\ref{fig:acotCont}, but get an output such as in figure~\ref{fig:acotJump} in \Maple.

\begin{figure}[!ht]
    \centering
    \subfloat[Plot of the arccotangent function with the branch cut defined by \DLMF.]{%
        \includegraphics[width=0.45\textwidth]{acotCont.png}
        \label{fig:acotCont}%
    }
    \hspace{0.2cm}
    \subfloat[Plot of the arccotangent function with the branch cut defined by \Maple.]{%
        \includegraphics[width=0.45\textwidth]{acotJump.png}
        \label{fig:acotJump}%
    }
    \caption{Plots for the arccotangent function $\acot@{z}$ with $z = \frac{3}{2}\expe^{\iunit\phi}$, where $\phi \in [0,2\cpi]$. Figure~\protect\subref{fig:acotCont} illustrates a branch cut defined at $[-\iunit, \iunit]$ and figure~\protect\subref{fig:acotJump} illustrates the branch cuts defined at $(-\infty\iunit, -\iunit]$ and $[\iunit, \infty\iunit)$. The blue and red dots represents the positions, where the function jumps over the branch cuts in~\protect\subref{fig:acotJump}.}
    \label{fig:acot}
\end{figure}

We provide the following solution for such problems. Instead of translating the function itself, the translator uses other equivalent representations for the function. In case of the arccotangent function, we can use the following alternative representations, based on the suggestions in~\cite{Branches:acot}.

Note that the following notation uses our standard notation for a translation from semantic \LaTeX{} to \Maple. We will introduce this notation in a formal way later in section~\ref{sec:definition}.

\begin{eqnarray}
\verb|\acot@{z}| & \overset{\langMaple}{\mapsto} & \verb|arccot(z)|\label{eq:acot-alternatives}\\
& \overset{\langMaple}{\mapsto} & \verb|arctan(1/z)|\label{eq:acot-alternatives-1}\\
& \overset{\langMaple}{\mapsto} & \verb|I/2*ln((z-I)/(z+I))|\label{eq:acot-alternatives-2}
\end{eqnarray}

As illustrated above, translation~(\ref{eq:acot-alternatives}) contains the branch cut issue. Translation~(\ref{eq:acot-alternatives-1}) is an alternative, because the arctangent function in \Maple{} has the same position for the branch cut of the arctangent function in the \gls{dlmf}. However, in this case, the arccotangent function would be no longer defined at $z=0$. Translation~(\ref{eq:acot-alternatives-2}) is another alternative and defined at $z=0$.

A more complicated example might be the equation of the already introduced parabolic cylinder function $\ParabolicU@{a}{z}$ and the modified Bessel function of the second kind $\BesselK{\nu}@{z}$. The parabolic cylinder function is a solution of a second order differential equation~\cite[(12.2i)]{NIST:DLMF} and can be represented by $\BesselK{\nu}@@{z}$ for $a=0$. The relation is defined in~(\ref{eq:branch-cut-issue}).

\begin{equation}\label{eq:branch-cut-issue}
\displaystyle \ParabolicU@{0}{z} = \sqrt{\frac{z}{2\cpi}}\BesselK{\frac{1}{4}}@{\frac{1}{4}z^2}
\end{equation}

Following the same approach as for the arccotangent function discovers results in a branch cut problem. Figure~\ref{fig:u-bessel} illustrates that problem for equation~(\ref{eq:branch-cut-issue}) with $z=2.5\expe^{\iunit\phi}$ and letting $\phi$ increase from $0$ to $2\cpi$. While the left hand side is an entire function of $z$ (can be seen in figure~\ref{fig:u-cont}), the right hand side has two functions with a branch cut along the negative real axis each. As already discussed, the branch cut of the square root function is at $(-\infty,0]$. Therefore, at $\phi = \cpi$ the right hand side jumps over the branch cut (green dots) back to the principal branch rather than continue on the other branch. Furthermore, the principal branch of $K_\nu$ is defined for $\phi \in (-\cpi,\cpi]$. With $z^2 = r^2\expe^{2\iunit\phi}$, the branch cut of $K_\nu$ is reached at $\phi = \left\{ \frac{\cpi}{2}, \frac{3\cpi}{2} \right\}$ (red dots and blue dots). The branch cut of $K_\nu$ "moves" to $(-\infty\iunit,0]$ and $[0,\infty\iunit)$. Hence, we end up with three branch cuts for the right hand side of (\ref{eq:branch-cut-issue}).

\begin{figure}[!ht]
    \centering
    \subfloat[The left hand side of (\ref{eq:branch-cut-issue}).]{%
        \includegraphics[width=0.45\textwidth]{CylinderU.png}
        \label{fig:u-cont}%
    }
    \hspace{1cm}
    \subfloat[The right hand side of (\ref{eq:branch-cut-issue}).]{%
        \includegraphics[width=0.45\textwidth]{WithCuts.png}
        \label{fig:bessel-cut}%
    }
    \caption{Two polar plots with $z=2.5\expe^{\iunit\phi}$ for $\phi \in [0,2\cpi]$ of left and right hand side of (\ref{eq:branch-cut-issue}) in \Maple. $U$ is an entire function, while $K_\nu$ and the square root function have a branch cut along $(-\infty,0]$. The colored dots represent the jumps over the branch cuts.}
    \label{fig:u-bessel}
\end{figure}

Since \gls{cas} do not provide Riemann surfaces, we use another approach to solve this problem, called \textbf{analytic continuation}. Analytic continuation is the process of extending the range of validity of a representation or more generally extending the region of definition of an analytic function~\cite[152]{ComplexVariables}. For the modified Bessel function of the second kind the analytic continuation is defined in the \gls{dlmf}~\cite[(10.34.4)]{NIST:DLMF}. With analytic continuation, the right hand side of equation~(\ref{eq:branch-cut-issue}) is illustrated in figure~\ref{fig:analytic-cont}.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.6\textwidth]{WithAnalyticCont.png}
    \caption{The right hand side of (\ref{eq:branch-cut-issue}) using analytic continuation, when the function jumps over the branch cut at $\phi = \left\{\frac{\cpi}{2},\cpi\right\}$. At $\phi = \frac{3\cpi}{2}$, the function jumps back to the principal branch.}
    \label{fig:analytic-cont}
\end{figure}

\subsection{Injective, Surjective \& Bijective Mappings}
Later in the thesis, we will define our translation as a function. Since our main goal is to provide bijectivity for our translation, we will briefly introduce the terminology.

Consider a function $f: X \mapsto Y$.

\begin{definition}[Injective Function]
The function $f$ is \textbf{injective} if
\begin{equation}
\forall x,x' \in X: \quad f(x) = f(x') \Rightarrow x = x'.
\end{equation}
\end{definition}

\begin{definition}[Surjective Function]
The function $f$ is \textbf{surjective} if
\begin{equation}
\forall y \in Y, \exists x \in X: \quad y = f(x).
\end{equation}
\end{definition}

\begin{definition}[Bijection]
The function $f$ is \textbf{bijective} if $f$ is injective and surjective.
\end{definition}

The most useful part about bijective mappings is that every bijective function $f$ has an inverse function $f^{-1}$. Consider a translation between two systems as a function that is bijective. Such a translation has an inverse function, or in other words backward translation. This would allow us to translate each expression in one system to the other and back again.

Obviously, such a translation is desirable. However, we will see that this goal is not achievable for all possible expressions. Even worse, we will see that our translation is neither injective nor surjective.

\subsection{Graphs and Trees}
\glsreset{dag}
A typical way to formalize the structure of mathematical expressions are trees. We will briefly introduce what trees are and what an expression tree is.

\begin{definition}[Graph]
A \textbf{graph} is an ordered pair $G=(V,E)$, where $V$ is the set of \textbf{vertices} (or \textbf{nodes}) and $E$ the set of \textbf{edges}, which are a 2-element subset of $V$.
\end{definition}

\begin{definition}[Directed Graph]
A \textbf{directed graph} (sometimes also referred as \textbf{digraph}) $G = (V,E)$ is a graph with oriented edges. That means $E$ is a set of ordered pairs of vertices. Therefore, a graph is \textbf{undirected} if all edges $e \in E$ have no orientation.
\end{definition}

\begin{definition}[Paths, Cycles and connected Graphs]
Let $G = (V,E)$ be a directed or undirected graph with at least three vertices ($|V| \geq 3$). A \textbf{path} from one vertex $i\in V$ to another $j\in V$ is a finite sequence of vertices and edges
\begin{equation}
i = i_1, (i_1, i_2), i_2, \ldots , i_{k-1}, (i_{k-1}, i_k), i_k = j.
\end{equation}
A path is a \textbf{cycle} if $i = j$. If a graph $G = (V,E)$ has no cycles, it is called \textbf{acyclic}. Furthermore, if there is a path for every pair of vertices, $G$ is called \textbf{connected}.
\end{definition}

A \gls{dag}, for example, is used internally to represent mathematical expressions. However, another way to represent mathematical expressions are trees.

\begin{definition}[Tree]
A \textbf{tree} $T = (V,E)$ is an acyclic and connected graph.
\end{definition}

Let $T = (V,E)$ be a tree and $r \in V$. We call $r$ \textbf{root} of $T$ and order the other nodes in an ascending ordering according to its length of the path to $r$. Each node $v \in V$, which is directly connected with $r$ is called \textit{child} of $r$ and $r$ is called \textit{parent} of $v$. This can be defined for all nodes in the tree which created a hierarchical structure in the tree. A node is in a higher hierarchy, when it is closer to root node. A node can have multiple \textit{children}. Consider a node $a \in V$ as a node with three children $u,v,w$. We also define a order for the children, so that $u < v < w$. \textbf{Siblings} are nodes that share the same parent node. With the order we call $u$ the previous sibling of $v$ and $w$ the following sibling of $v$. With this additional terminology, we can move through trees easily. Trees with an ordering for all nodes and a root node are also called \textbf{ordered trees}.

Mathematical expressions can be represented by ordered trees consisting of terminal symbols, such as identifiers or numbers (leaf nodes), and functions or operators (non-leaf nodes)~\cite{VMEXT}. Consider the mathematical expression
\begin{equation}\label{eq:simple-expr}
\expe^{x} + \frac{x^2}{\cpi}.
\end{equation}

The tool \gls{vmext}~\cite{VMEXT} visualizes expression trees. Figure~\ref{fig:simple-expr-tree} draws the expression tree for (\ref{eq:simple-expr}).

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.7\textwidth]{expressionTreeSimple.png}
    \caption{The expression tree of (\ref{eq:simple-expr}) visualized with VMEXT~\cite{VMEXT}.}
    \label{fig:simple-expr-tree}
\end{figure}

### 2018-05-22 chapters/02-background/mlp.tex
Init commit with original master thesis

\section{Mathematical Language Processor}\label{sec:mlp}
\glsreset{mlp}\glsreset{pom}\glsreset{bnf}
The \gls{mlp}\footnote{Note that the official publication refers \gls{mlp} to \textit{Mathematical Language \textbf{Processor}}, rather than to the \textbf{parser}. However, the \textit{Mathematical Language Processor} project~\cite{MLP:Project} is slightly different and trying to analyize the natural language context of mathematical expressions. However, both projects are highly coupled and uses the same abbreviation. Since we focus on the parser, we use \gls{mlp} to refer to the \textit{Mathematical Language Parser}, even the official publication understood it as a \textbf{processor}.} project was developed to parse any mathematical expressions~\cite{POM-Tagger}.

According to \gls{nlp}~\cite{NLP}, the \gls{mlp} understood mathematical expressions as a word of a formal language and parses expressions to so called parse trees. To understand the \gls{mlp}, we need to understand the concept of formal languages first. In the following, we will briefly introduce formal languages~\cite{Math:Encyclopedia:FormalLanguage}.

\begin{definition}[Alphabet]
An \textbf{Alphabet} $\Sigma$ is a finite, non-empty set. The elements of $\Sigma$ are called \textbf{letters} (or \textbf{symbols}).
\end{definition}

For example, the alphabet of a computer is simply the binary alphabet $\Sigma = \{1,0\}$.

\begin{definition}[Word]
A \textbf{word} (or string) $w$ over an alphabet $\Sigma$ is a finite sequence of letters of $\Sigma$. The set of all words over $\Sigma$ is denoted as $\Sigma^\ast$ (known as the Kleene star of $\Sigma$). The empty word $\epsilon$ (sequence of length 0) is always an element of $\Sigma^\ast$.
\end{definition}

\begin{definition}[Formal Language]
A subset $L \subseteq \Sigma^\ast$ is called \textbf{formal language} over an alphabet $\Sigma$.
\end{definition}

Hereafter we use the term \textit{formal language} and \textit{language} interchangeably. Usually, a formal language is constructed by a set of rules. In our case we construct a formal language by a context-free grammar~\cites[79]{Grammar}{IntroComp}.

\begin{definition}[Context-Free Grammar]
A context-free grammar $G = (V, \Sigma, R, S)$ is a $4$-tuple, where
\begin{enumerate}
\item $V$ is a finite set. The elements of $V$ are called \textbf{non-terminal symbols}.
\item $\Sigma$ is a finite set with $\Sigma \cap V = \varnothing$. The elements of $\Sigma$ are called \textbf{terminal symbols}.
\item $R$ is a finite relation from $V$ to $(\Sigma \cup V)^\ast$, thus $R \subseteq V \times (\Sigma \cup V)^\ast$. The members of $R$ are called \textbf{rules}.
\item $S \in V$ is called the \textbf{start symbol}.
\end{enumerate}
\end{definition}

A rule $(\alpha,\beta) \in R$ can be understood as a substitution. We denote the rule also as $\alpha \longrightarrow \beta$.

\begin{definition}[Derived Words \& Applied Rules]
Let $G = (V, \Sigma, R, S)$ be a context-free grammar. Let $A \in V$ and $u,v,w \in (\Sigma \cup V)^\ast$, such that $(A,w) \in R$. Than we say a word '$uwv$' can be \textbf{derived} in one step from the word '$uAv$' by \textbf{applying} the rule $(A,w)$.
\begin{equation}
uAv \Rightarrow uwv.
\end{equation}
\end{definition}

We can also apply multiple rules sequentially.

\begin{definition}[Chain Derivation]
Let $G = (V, \Sigma, R, S)$ be a context-free grammar and $u,v \in (\Sigma \cup V)^\ast$, with $u \neq v$. We say $v$ can be derived from $u$, if
\begin{equation}
\exists\ k \geq 2\ \ \exists u_1, u_2, \ldots, u_k \in (\Sigma \cup V)^\ast: \quad u = u_1 \Rightarrow u_2 \Rightarrow \ldots \Rightarrow u_k = v,
\end{equation}
and write
\begin{equation}
u \overset{\ast}{\Rightarrow} v.
\end{equation}
\end{definition}

With a context-free grammar, we can define a context-free language over the grammar.

\begin{definition}[Context-Free Language]
Let $G = (V, \Sigma, R, S)$ be a context-free grammar. $G$ defines a language $L(G)$ by
\begin{equation}
L(G) = \left\{ w \in \Sigma^\ast:\ S \overset{\ast}{\Rightarrow} w \right\}.
\end{equation}
\end{definition}

A language $L$ is called \textbf{context-free language} if there is a grammar $G$, such that $L = L(G)$. In other words, a context-free language $L(G)$ is a set of all strings that can be derived from the start variable $S$ of $G$.

\begin{example}\label{ex:cfl}
A simple example for a context-free grammar are well-formed parentheses. Let $V = \{S\},\ \Sigma = \{(,\ )\}$ and $S$ be the start symbol. Furthermore, let $R$ contain the rules
\begin{eqnarray}
&&S \longrightarrow SS\\
&&S \longrightarrow (S)\\
&&S \longrightarrow ()
\end{eqnarray}
Let $G = (V, \Sigma, R, S)$. Than, for example
\begin{equation}\label{eq:example-proof}
\text{'}(()(()))\text{'} \in L(G).
\end{equation}
\end{example}

\begin{proof}[Proof of~\ref{eq:example-proof}]
We can produce the following chain of rules from $R$.
\begin{eqnarray}
S &\Rightarrow & (S)\label{eq:chain}\\
&\Rightarrow & (SS)\notag\\
&\Rightarrow & (()S)\notag\\
&\Rightarrow & (()(S))\notag\\
&\Rightarrow & (()(())).\notag
\end{eqnarray}
Therefore
\begin{equation}
S \overset{\ast}{\Rightarrow} \text{'}(()(()))\text{'} 
\end{equation}
and '$(()(()))$' $\in L(G)$.
\end{proof}

On the other hand
\begin{equation}
\text{'}(()\text{'} \not\in L(G).
\end{equation}

The chain of rules can also be represented in a tree structure. Such trees are called \textbf{parse trees}. The tree for~(\ref{eq:chain}) is visualized in figure~\ref{fig:parseTree}. Note that a parse tree is different to an expression tree. However, a parse tree can become an expression tree if the appropriated mathematical rules are defined.

\begin{figure}[ht]
	\centering
	\includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm,width=0.6\textwidth]{ParseTree.pdf}
	\caption{The parse tree of the expression '$(()(()))$' for the well-formed parenthesis grammar $G = (V, \Sigma, R, S)$. Produced by the chain of rules~\ref{eq:chain}.}
	\label{fig:parseTree}
\end{figure}

There are several more definitions around formal languages, which we will not mention here. Note, for example, that a parse tree such as in figure~\ref{fig:parseTree} is not necessarily unique and can depend on the choice of a rule where more than one is \textit{applicable}. Consider a context-free grammar for simple arithmetic expressions with sums and subtractions of three variables $x,y$ and $z$. Let $V = \{S\},\ \Sigma = \{x,y,z,+,-\}$, $S$ be the start symbol and define the rules
\begin{eqnarray}
&&S \longrightarrow x\label{eq:chain-1}\\
&&S \longrightarrow y\label{eq:chain-2}\\
&&S \longrightarrow z\label{eq:chain-3}\\
&&S \longrightarrow S + S\label{eq:chain-4}\\
&&S \longrightarrow S - S\label{eq:chain-5}.
\end{eqnarray}
Now an expression such as $x+y-z$ is an element of the context-free language of $G = (V, \Sigma, R, S)$. However, there are two possible parse trees for this expression, tree~\ref{fig:parse1} and \ref{fig:parse2}, depending on the order of which rule gets applied first. Such grammars are also called ambiguous.

\begin{figure}[!ht]
    \centering
    \subfloat[The parse tree of $x+y-z$ if rule~(\ref{eq:chain-4}) gets applied first.]{%
        \includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm,width=0.4\textwidth]{ParseTree2a.pdf}
        \label{fig:parse1}%
    }
    \hspace{2cm}
    \subfloat[The parse tree of $x+y-z$ if rule~(\ref{eq:chain-5}) gets applied first.]{%
        \includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm,width=0.4\textwidth]{ParseTree2b.pdf}
        \label{fig:parse2}%
    }
    \caption{Two different parse trees for $x+y-z$. \protect\subref{fig:parse1} applies rule~(\ref{eq:chain-4}) before (\ref{eq:chain-5}), while \protect\subref{fig:parse2} applies the rules vice versa.}
    \label{fig:parseMulti}
\end{figure}

A parser can define now in which order the rules get applied. For example, an \textbf{LL}-Parser parses the input from \textbf{L}eft to right and performing the \textbf{L}eftmost derivations. Leftmost in this terminology means, the parser takes the first applicable rule. In consequence, the internal ordering of rules is no longer irrelevant. The \gls{mlp} is such an \textbf{LL}-Parser. In case of the expression $x+y-z$ from above with the rules~(\ref{eq:chain-1}~-~\ref{eq:chain-5}), an \textbf{LL}-Parser would generates tree~\ref{fig:parse1}.

\subsection{BNF Grammar}
The \gls{bnf} is a notation technique for context-free grammars in computers. For example, a \gls{bnf} specification is a set of rules, written as
\begin{equation}
\verb|<symbol> ::= exp|,
\end{equation}
where \verb|<symbol>| is a non-terminal element and \verb|exp| are one or more sequences of symbols. Symbols that never appear on the left side are terminals.

The \gls{bnf} is often used to describe the syntax of programming languages. There are several tools that automatically generates a parser from a given grammar in \gls{bnf}. One of those tools is \textit{JavaCC} (for Java Compiler Compiler).

\subsection{POM-Tagger \& MLP}\label{subsec:pom-mlp}
The \gls{mlp} defines a \gls{bnf} grammar to parse mathematical \LaTeX{} expressions. The \gls{pom}-tagger (according to Part-of-Speech tagger in \gls{nlp}) categorizes math tokens. The main purpose of the \gls{pom}-tagger is the extraction of semantic information from mathematical \LaTeX. The \gls{pom} tagger is designed to perform multiple \textit{scans}. Given an input \LaTeX{} math document, the first scan of the tagger examines \textit{terms}, and groups them into sub-expressions where suitable. For instance $\verb|\frac{1}{2}|$ consists of two sub-expression of numerator and denominator. A \textit{term} is a non-terminal symbol defined in the \gls{bnf}. In the definition of a \LaTeX{} grammar this can be \LaTeX{} macros, environments, reserved symbols (such as the \LaTeX{} line break command \textbackslash\textbackslash), or numerical or alphanumerical expressions. The first scan of the \gls{pom}-tagger adds information to each term. The information is defined in lexicon files, which provide meanings, descriptions, categories and many more for a large variety of symbols. For example, table~\ref{tab:plus-lex-example} gives an example of the lexicon entry for the plus symbol.

\begin{table}[ht]
	\centering
	\begin{tabular}{lll}
	\hline
	\multicolumn{3}{l}{Symbol: \texttt{+}} \\
	\! & \multicolumn{2}{l}{Feature Set: universal} \\
	\! & \! & Category: operation\\
	\! & \! & Description: plus\\
	\! & \! & Source: unicode-math\\
	\! & \! & \TeX-equivalent: \verb|\mathplus|\\
	\! & \! & Unicode: u+0002b\\
	\hline
	\end{tabular}
	\caption{The lexicon entry for the $+$ symbol.}
	\label{tab:plus-lex-example}
\end{table}

A symbol can be tagged with multiple feature sets according to its ambiguous semantics. Tagged terms are called math terms. Math terms are rarely distinct at this stage and often have multiple features. All information are centralized in the \verb|global-lexicon.txt|.

Further scans will try to reduce the number of alternative feature sets by scanning the context of the math expression and concluding possible feature sets from the structure of the expression. See section~\ref{sec:resolution} for more information about this approach.

In its current state, the \gls{pom}-tagger only performs the first scan. Note that, technically, the \gls{mlp} parses mathematical expressions while the \gls{pom}-tagger categorizes elements of the parse tree. However, the \gls{mlp} also acts as an interface to interact with the parse tree and the \gls{pom}-tagger is a kind of an intermediate step of the whole parsing process. Therefore, we will refer to the \gls{mlp} in following chapters rather than to the \gls{pom}-tagger.

\subsection{MLP-Parse Tree}\label{subsec:mlp-syntax-tree}
Hereafter we consider the \gls{mlp-pt} as the parse tree, which is produced by the \gls{mlp}. The \gls{bnf} of the \gls{mlp} has no mathematical rules implemented in its current state. For example, there is no rule, similar to~(\ref{eq:chain-4}) for the binary operator $+$. Therefore, the \gls{mlp-pt} is not an expression tree. The \gls{pom} project was designed to extract semantic information from mathematical expressions. Certainly, the semantics of mathematical symbols is rarely unique. Consider the $-$ symbol as an example. As a binary operator it indicates the subtraction of two arguments. However, it is also a unary operator as a leading sign.

In its current stage, it is futile to implement arithmetical rules into the \gls{mlp} grammar. The next scan of the \gls{mlp} will reduce ambiguous meanings from the \gls{mlp-pt}. It can be anticipated that the \gls{mlp-pt} will become an expression tree when the planned next scans are implemented.

\begin{figure}[!ht]
    \centering
    \subfloat[The MLP-PT for $a+b$. The POM-Tagger adds information to all terminals. F-Set are feature sets. The terms $a$ and $b$ have multiple feature sets. An F-set represents another semantic meaning. In the figure are not all F-sets listed.]{%
        \includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm,width=0.65\textwidth]{ParseTreeSimple.pdf}
        \label{fig:tree-parse}%
    }
    \hspace{0.5cm}
    \subfloat[The expression tree for $a+b$.]{%
        \includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm,width=0.25\textwidth]{ExpressionTreeSimple.pdf}
        \label{fig:tree-expr}%
    }
    \caption{Comparing the MLP-Parse tree with an expression tree for $a+b$.}
    \label{fig:tree-compare}
\end{figure}

### 2018-05-22 chapters/02-background/related-work.tex
Init commit with original master thesis

\section{Related Work}\label{sec:related-work}
Since \LaTeX{} became the de facto standard for writing papers in mathematics, most of the \gls{cas} nowadays are able to import and export mathematical \LaTeX{} expressions~\cite{Maple:ImportExport,Mathematica:ImportExport,Matlab:ImportExport,Sage:ImportExport}. Those tools have two essential problems. First of all, they cannot import mathematical expressions where the semantic information is absent, which is the usual case for generic \LaTeX{} expressions. Therefore, they are only able to import simple mathematical expressions, where the semantics are unique. For example, the internal \LaTeX{} macro \verb|\frac| always indicates a fraction. These import tools fail for more complex expressions, for example the Jacobi polynomial in \cref{tab:JacobiP-usecase}. The second problem appears in the export tools. Mathematical expressions in \gls{cas} are fully semantic, otherwise the \gls{cas} wouldn't be able to compute or evaluate the expressions. During the export process, the semantic information gets lost, because generic \LaTeX{} is not able to carry semantic information. In consequence of these two problems, an exported expression cannot be imported to the same system again in most cases (except those simple expressions described above). Our tool should solve these problems and provide round trip translations between \LaTeX{} and \gls{cas}.

Obviously, the semantics must be well known before an expression can be translated. There are two approaches to solve that problem: someone could specify the semantic information during the writing process (pre-define semantics) or the translator can find out the right semantic information in general mathematical expressions before it translates the expression. The second approach is unreasonably difficult compared to the pre-definition approach. We will talk about those difficulties in the following sections. We will also introduce a new approach in \cref{ch:conc-future-work} to achieve the second idea.

However, the pre-definition approach is simple and easy to realize, as long as the system can carry semantic information. A typical, not mathematical example are Wikipedia articles~\cite{Wiki}. An author of such an article can specify further information about symbols, words and sentences by creating a hyperlink to a more detailed explanation. Usually, the author defines the hyperlink, when he is writing the article or he needs to manually add the hyperlink later on. Similar to this approach are \textit{interactive documents}\footnote{There is no adequate definition what interactive documents are. However, this name is widely used to describe electronic document formats that allow interactivity to change the content in real time.}, such as the \gls{cdf}~\cite{CDF:Wolfram} by Wolfram Research or the \textit{worksheets} of \Maple{}, which are also sometimes referred to as interactive documents. Those documents pre-define the semantics to allow computations. Those complex document formats require specialized tools to show and work with the documents (Wolfram CDF Player, or \Maple{} for the \textit{worksheets}). The \JOBAD{} architecture~\cite{JOBAD:orig} is able to create web based interactive documents and uses \gls{omdoc}~\cite{OMDoc} to carry semantics. The documents can be viewed and edited in the browser. Those \JOBAD{-documents} also allow to perform computations during \gls{cas}. This gives the opportunity to calculate, compute and change mathematical expressions directly in the document. The translation performs in the background, invisible for the user. Similar to the \JOBAD{} architecture other interactive web documents exist, such as \textit{MathDox}~\cite{MathDox} and \textit{The Planetary System}~\cite{Planetary}. All of them demonstrate the potential for the education system.

All of these systems carry semantic information. Most of them create their own way to make this possible. The web based documents mostly use standardized ways, such as the content \gls{mathML}, a specialization of \gls{mathML}~\cite{MathML}, that focuses on the semantics of an expression rather than any particular rendering for the expression. Since \LaTeX{} is the de facto standard to write mathematical expressions in documents there are projects to extend \LaTeX{} in a way that enable to carry the semantic information directly in \LaTeX. There are two known projects that created a semantic version of \LaTeX: \sTeX{}~\cite{sTeX} developed by Michael Kohlhase and the \Macro s developed by Bruce Miller. Our translator uses the \Macro s rather than \sTeX. A full explanation of this decision will be given in \cref{sec:semantics}.

However, \textit{interactive documents}, web based or not, have an essential problem: all of them create an entire new document format. We want to create an independent lite translation tool for \LaTeX. Once we achieve this goal, an integration into \textit{interactive documents} might be possible, besides many other possible applications.

Another approach that tries to avoid the translation problem are \LaTeX{} packages that allow computations directly through the \LaTeX{} compiler (for example \textit{LaTeXCalc}~\cite{LatexCalc}) or allow \gls{cas} commands for computations directly in \TeX -files. For example, \textit{sagetex}~\cite{Sagetex} is an interface \LaTeX{} package for the open source \gls{cas} \textit{sage}\footnote{An abbreviation for \textit{SageMath}.}. This package allows \textit{sage} commands in \TeX -files and uses \textit{sage} in the background to compute the commands. Obviously, the first example is not as powerful as a \gls{cas} would be. However, \textit{sagetex} does not really solve our problem of the scientific workflow. The writer still needs to translate expressions manually. The only difference is that the author can write computations directly into the \TeX -file. However, these approaches have the potential for a powerful combination with our translation tool.

In summary, today's translation tools are either specialized for one specific \gls{cas} (such as the import and export functions of the \gls{cas}) or they completely avoid the translation process by creating a new document format (such as \textit{interactive documents}). Our goal is to provide a lite translation tool for mathematical \LaTeX{} expressions and multiple \gls{cas}.

### 2018-05-22 chapters/02-background/semantic-latex.tex
Init commit with original master thesis

\section{Semantic \& Generic \LaTeX}\label{sec:semantics}
Donald E. Knuth developed the typesetting system \TeX{} in 1977~\cite[559]{DigitalTypo}, because he was unsatisfied with the typography of his book \textit{The Art of Computer Programming}, Volume 2~\cites[5]{DigitalTypo}{Knuth}. Around 1980, Leslie Lamport starts to extend \TeX{} by a set of macros, which became a new version of \TeX. According to his name \textit{Lamport}, this new version was called \LaTeX{~}\cite{LATEX}. Nowadays, \LaTeX{} has become the de facto standard for scientists to write their publications~\cite{LATEX:Standard}. This thesis, for example, is written in \LaTeXe, the current version of \LaTeX.

It is possible to customize and extend \LaTeX, for example through style files, where an editor can define new \LaTeX{} macros, environments and document styles.

\subsection{Mathematical Generic \LaTeX}
\LaTeX{} and \TeX{} enables printing of mathematical formulae in a structure similar to handwritten style. For example, the \LaTeX{} expression~\cite[(4.26.13)]{NIST:DLMF}
\begin{equation}\label{eq:latex-ex}
\verb|\int_{0}^{\infty} \sin(t^2) dt = \frac{1}{2} \sqrt{\frac{\pi}{2}}|
\end{equation}
will be rendered as
\begin{equation}\label{eq:latex-ex-r}
\int_{0}^{\infty} \sin(t^2) dt = \frac{1}{2} \sqrt{\frac{\pi}{2}}.
\end{equation}

We call the syntax of such mathematical expressions in \LaTeX{} \textit{mathematical \LaTeX{}} and hereafter use the terms \textit{\LaTeX{} expressions}, \textit{mathematical \LaTeX{} expressions} and \textit{mathematical \LaTeX{}} interchangeably.

Since \TeX{} and \LaTeX{} are typesetting systems, there are several ways to change the rendering of such mathematical expressions. However, for a \gls{cas} the information about the exact rendering of a mathematical expression is not as important as the meaning of the expression, the semantics.

\subsection{Semantic Information}\label{subsec:semantic-latex}
\textit{Semantics} in general is the study of meanings. Semantic information, for example, can be in words, whole sentences, mathematical expressions, and signs.

Consider the equation~(\ref{eq:latex-ex-r}). The meaning behind this expression is an equation. The left-hand side uses an infinite integral over the trigonometric sine function, which can be defined with~\cite[(4.14.1)]{NIST:DLMF}
\begin{equation}
\sin@{z} := \frac{\expe^{\iunit z}-\expe^{-\iunit z}}{2\iunit}.
\end{equation}
The right-hand side of~(\ref{eq:latex-ex-r}) contains a Greek letter $\cpi$, which is known to indicate the mathematical constant for the ratio of a circle's circumference to its diameter. However, $\pi$ can also indicate the prime-counting function. A human reader might conclude that $\pi$ cannot be the prime-counting function, because there is no variable given. In (\ref{eq:latex-ex-r}), $\cpi$ is much more likely the mathematical constant, also because the left-hand side contains the sine function and the mathematical constant $\cpi$ belonging to trigonometric functions.

As demonstrated, there are a lot of meanings behind~(\ref{eq:latex-ex-r}). Some of them are clear, others can be questionable. For a \gls{cas}, the exact meaning of a given input must be unique, otherwise the \gls{cas} cannot compute the input. Therefore, mathematical expressions in \gls{cas} contain accessible semantic information. This is realized to connect character sequences with unique internal procedures and classes, which represent specific mathematical objects.

Mathematical \LaTeX{} usually only contains information about the rendering of the expression and not about the meaning of the expression. However, there is semantic information in mathematical \LaTeX, but it's often hidden. This was shown in the example discussion of the meaning of equation~(\ref{eq:latex-ex-r}). In the plain text of the Greek letter $\pi$ (\verb|\pi|) is no exact meaning. Or, to be more precisely, there are multiple possible meanings, but there is no way to find the correct meaning. However, together with the left-hand side we can conclude that $\pi$ is the mathematical constant in that case. As demonstrated, there is semantic information in the expression, but a reader needs to analyze the whole expression and make conclusions to find it. Hence, for a computer it is difficult to understand mathematical \LaTeX. To provide a translation from mathematical \LaTeX{} to a \gls{cas}, the semantics must be clear and accessible.

We will introduce in the next subsection a semantic version of mathematical \LaTeX. To distinguish this version with classical mathematical \LaTeX, we call classical \LaTeX \textit{generic \LaTeX}.

Obviously, there must be different levels of semantics in expressions. We say an expression in generic \LaTeX{} such as 
\begin{equation}\label{eq:jac-latex}
\verb|P_n^{(\alpha,\beta)}(\cos(a\Theta))|
\end{equation}
contains less semantic information (more semantics are absent) than an appropriate representation in \Maple{}
\begin{equation}\label{eq:jac-maple}
\verb|JacobiP(n,alpha,beta,cos(a*Theta))|.
\end{equation}
This is because the meaning of (\ref{eq:jac-maple}) is unique and accessible, while the exact meaning of (\ref{eq:jac-latex}) must be concluded from the structure and context of the formula. Unfortunately, there is no formal definition for grading semantics or a \textbf{fully semantic} expression. However, we use this phrase, when there is sufficient semantic information accessible to provide an appropriate translation.

\subsection{DLMF/DRMF \LaTeX{} Macro Set}\label{subsec:macros}
Bruce Miller at \gls{nist} has created a set of semantic \LaTeX{} macros~\cite{DLMF:Macros}. Each macro ties specific character sequences to a well-defined mathematical object and is linked with the corresponding definition in the \gls{dlmf} or \gls{drmf}. Therefore, we call these semantic macros \Macro s. These semantic macros are internally used in the \gls{dlmf}, \gls{drmf} and also in this thesis. They are defined in multiple style files and allow optional parameters and multiple renderings. Table~\ref{tab:macro-def} show how a semantic macro is defined in a style file.

\begin{table}[ht]
\centering
\begin{tabular}{lp{7cm}}
	\hline
	\verb|\defSpecFun{function}| & Define a macro \verb|\function|.\\\hline
	\hspace*{0.5cm}\verb|[nparams]| & Number of parameters of the macro.\\\hline
	\hspace*{0.5cm}\verb|[optparams]| & The number of optional parameters of the macro.\\\hline
	\hspace*{0.5cm}\verb|{format}| & Defines the format of the macro. (How it will be displayed).\\\hline
	\hspace*{0.5cm}\verb|[keys]| & Attributes of the function. For example, \textit{meaning} or \textit{role}.\\\hline
	\hspace*{0.5cm}\verb|{arity}| & Number of arguments. Defines how many \{$\bullet$\} blocks come after the $@$ symbol.\\\hline
	\hspace*{0.5cm}\verb|[argformat]| & Format of arguments.\\\hline
	\hspace*{0.5cm}\verb|[altargformat]| & Alternative formats depends on the number of $@$ symbol.\\\hline
\end{tabular}
\caption{Overview for the definition of a semantic macro.}
\label{tab:macro-def}
\end{table}

The \verb|\defSpecFun| is a defined macro to easily define new macros for special functions. It takes a unique character sequence for the function name. The first optional argument \verb|nparams| defines the number of parameters of the special function. Some special functions, for example the associated Legendre functions, can define optional parameters. The general structure of a semantic macro has two parts to display the formula. The first part defines the display for the function symbol (including parameters) in \verb|{format}|. This format is fixed and cannot be changed. In contrast, the display for the arguments of the function can be controlled by the number of $@$ symbols and is defined in multiple optional \verb|[argformat]| blocks.

Consider the \textit{hypergeometric function}~\cite[(15.2.1)]{NIST:DLMF}. The semantic macro 
\begin{equation}
	\verb|\HypergeoF@{a}{b}{c}{z}| 
\end{equation}
is linked to the definition (15.2.1) in the \gls{dlmf}. The \textit{hypergeometric function} provides three different representations. Table~\ref{tab:hypergeo} gives an overview for all of them.

\begin{table}[ht]
\centering
\begin{tabular}{lc}
	\hline
	Semantic macro & Display\\
	\hline
	\tableRowSpace \verb|\HypergeoF@{a}{b}{c}{z}| & $\HypergeoF@{a}{b}{c}{z}$\\
	\tableRowSpace \verb|\HypergeoF@@{a}{b}{c}{z}| & $F\!\left({\genfrac{}{}{0pt}{}{a, b}{c}}; z \right)$\\
	\tableRowSpace \verb|\HypergeoF@@@{a}{b}{c}{z}| & $\HypergeoF@@@{a}{b}{c}{z}$\\
	\hline
\end{tabular}
\caption{The semantic macro for the \textit{hypergeometric function}~\cite[(15.2.1)]{NIST:DLMF} has three different ways to display the function. The different representations are controlled by the number of $@$ symbols.}
\label{tab:hypergeo}
\end{table}

The display of the function name $F$ is fixed, but the number of $@$ symbols control the display of the arguments. The third representation displays only the variable of the function, but hide the given parameters. Therefore, the semantic \LaTeX{} source in the background provide more semantic information than the displayed version. However, all of them represent the \textit{hypergeometric function} defined in the \gls{dlmf}.

The sets of macros from the \gls{dlmf} and the \gls{drmf} are mostly different. The \gls{drmf} uses the macros from the \gls{dlmf} and developed new macros as well. However, it happens that there are two different macros for the same mathematical object defined. One example is the \textit{Gegenbauer polynomial}\footnote{Also known as the \textit{ultraspherical polynomial}.}
\begin{eqnarray}
	&&\verb|\Ultra{\lambda}{n}@{x}|,\\
	&&\verb|\Ultraspherical{\lambda}{n}@{x}|.
\end{eqnarray}

There are currently 685 semantic macros defined. Note that the number of macros is constantly changing, because the project of semantic macros is still work in progress.

\subsection{Weakness of DLMF/DRMF Macros}\label{subsec:semanticVSmacro}
The set of \Macro s gives the opportunity to semantically enhance mathematical expressions. However, there are mostly macros for \gls{opsf}. It is difficult to foresee if a semantic macro provides sufficient semantic information for an appropriate translation, or, as we formulated above, if a semantic macro is really fully semantic.

One example is the \textit{Wronskian}. For example, for two differentiable functions the \textit{Wronskian} is defined as~\cite[(1.13.4)]{NIST:DLMF}
\begin{equation}
	\Wronskian@{w_1(z), w_2(z)} = w_1(z)w_2'(z) - w_2(z)w_1'(z).
\end{equation}
In semantic \LaTeX{} it is used with
\begin{equation}
	\verb|\Wronskian@{w_1(z), w_2(z)}|.
\end{equation}
It turned out, that for an appropriate translation to a \gls{cas}, we need access to the variables of $w_1(z)$ and $w_2(z)$. Therefore, we improved the macro and replaced it to a new version
\begin{equation}
	\verb|\Wron{z}@{w_1(z)}{w_2(z)}|.
\end{equation}
This new macro is displayed in the same way as the old version was displayed, but it provides more information.

Another version of semantic \LaTeX{} is \sTeX, developed by Michael Kohlhase~\cite{sTeX}. This version of semantically enhanced \LaTeX{} is more comprehensive for general mathematics. It is not particularly useful for \gls{opsf}, and it doesn't link macros to unique mathematical definitions. Therefore, we did not use \sTeX{} in this project, although it provides more semantic information for simple arithmetic expressions. This part is almost completely missing in the \Macro s in its current state. One example is the question of white spaces. Scientists frequently uses white spaces to indicate multiplications. Most \gls{cas} are familiar with this convention and are able to understand such inputs. However, in \Maple's 1-D input notation, an asterisk is mandatory for multiplications. Therefore, we also added a macro \verb|\idot| (for \textit{invisible dot}), to provide a multiplication symbol which is not displayed in \LaTeX, but semantically enhances the \LaTeX{} expressions.

### 2018-05-22 chapters/03-translator/all-types-table.tex
Init commit with original master thesis

\begin{table}[H]
\vspace{-10pt}
\centering
\sloppy
\newcolumntype{L}{>{\raggedright\arraybackslash}X}
\newcolumntype{Y}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\begin{tabularx}{\textwidth}{Y{1.55cm} Y{2.5cm} L L}
	\hline
	& \textbf{Node type} & \textbf{Explanation} & \textbf{Example} \\
	\hline
	\hline
	\multicolumn{1}{Y{1.5cm}|}{\textbf{$r$ has children}} & Sequence & Contains a list of expressions. & $a+b$ is a sequence with three children ($a$, $+$ and $b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Balanced Expression & Similar to a sequence. But in this case the sequence is wrapped by \texttt{\tbs left} and \texttt{\tbs right} delimiters. & \texttt{\tbs left(}$a+b$\texttt{\tbs right)} is a balanced expression with three children ($a$, $+$ and $b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Fraction & All kinds of fractions, such as \texttt{\tbs frac}, \texttt{\tbs ifrac}, etc. & \texttt{\tbs ifrac\{a\}\{b\}} is a fraction with two children ($a$ and $b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Binomial & Binomials & \texttt{\tbs binom\{a\}\{b\}} has two children ($a$ and $b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Square Root & The square root with one child. & \texttt{\tbs sqrt\{a\}} has one child ($a$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Radical with a specified index & $n$-th root with two children. & \texttt{\tbs sqrt[a]\{b\}} has two children ($a$ and $b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Underscore & The underscore '\_' for subscripts. & The sequence $a$\_$b$ has two children ($a$ and '\_'). The underscore itself '\_' has one child ($b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Caret & The caret '\^{}' to for superscripts or exponents. Similar to the underscore. & The sequence $a$\^{}$b$ has two children ($a$ and '\^{}'). The caret itself '\^{}' has one child ($b$).\\
	\hline
	\hline
	\multicolumn{1}{Y{1.5cm}|}{\textbf{$r$ is a leaf}} & \Macro{} & A semantic \LaTeX{} macro & \texttt{\tbs JacobiP}, etc.\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Generic \LaTeX{} macro & All kinds of \LaTeX{} macros & \texttt{\tbs Rightarrow}, \texttt{\tbs alpha}, etc.\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Alphanumerical Expressions & Letters, numbers and general strings. & Depends on the order of symbols. $ab3$ is alphanumerical, while $4b$ are two nodes ($4$ and $b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Symbols & All kind of symbols & '$@$', '$*$', '$+$', '$!$', etc.\\
	\hline
\end{tabularx}
\caption{A table of all kinds of nodes in a MLP syntax tree. Note that this table groups some kinds for a better overview. For a complete list and a more detailed version see~\cite{POM-Tagger}.}
\label{tab:allTypesTable}
\end{table}

### 2018-05-22 chapters/03-translator/backward-trans.tex
Init commit with original master thesis

\section{Maple to Semantic \LaTeX{} Translator}\label{sec:backward-translation}
The backward translation is based on the internal data structure of \Maple. Therefore, a license of \Maple{} is mandatory to perform backward translations. The following subsections will explain why this approach was necessary and how \Maple{} works internally. Using the internal data structure also requires to handle the internal properties of \Maple. This can causes some problems. Those problems and our solutions will be explained in subsection~\ref{subsec:maple-probs}. In the last subsection~\ref{subsec:maple-impl}, we will focus on the implementation of the backward translation.

Note that all explanations in this section is based on \Maple's programming guide~\cite{MAPLE:ProgrammingGuide} for \Maple{} 2016.

\subsection{Internal Data Structure}
While our forward translation is realized for the \texttt{1D} input representation in \Maple, this representation is not suitable for a backward translation. Since \Maple{} uses it's own syntax and programming language, we would need a new parser for \Maple{} expressions to realize a backward translation tool. Of course, \Maple{} has this parser implemented. But it is not possible to use the parser without using \Maple{} itself. Therefore, an installed version of \Maple{} is mandatory for our backward translation.

As already mentioned, \Maple{} has several ways to represent an expression. The \texttt{1D} and \texttt{2D} representations are used as input representations. Besides those, there are internal data structures to perform all kinds of computations on the expression. Internally, each expression is handled and stored as a \gls{dag}. The \Maple{} \gls{dag} is very similar to an expression tree. But for efficiency, it does not store copies of an object. Consider the integral from (\ref{eq:maple-input})
\begin{equation}\label{eq:maple-input-2}
\int\displaylimits_0^\infty \frac{\pi+\sin(2x)}{x^2} dx.
\end{equation}

As already mentioned, the \Maple{} \gls{dag} has no copies of objects. Therefore, the '$x$' in \eqref{eq:maple-input-2} only appears once in the \gls{dag}.
\begin{figure}[ht]
\centering
\includegraphics[scale=0.7]{DAG.pdf}
\caption{The \Maple{} DAG of equation~\eqref{eq:maple-input}}
\label{fig:maple-dag}
\end{figure}

The \Maple{} \gls{dag} can be visualized in a graph, such as in figure~\ref{fig:maple-dag}. But the internal representation looks a bit different. Each node in the \gls{dag} stores its children and has a header, which defines the type and the length of the node. Consider the polynomial $x^2+x$. Figure~\ref{fig:internal-maple-dag} illustrates the internal \gls{dag} representation with headers and arguments.

\begin{figure}[ht]
	\centering
	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.7]{DAGreal.pdf}
	\caption{The internal \Maple{} DAG representation of $x^2+x$.}
	\label{fig:internal-maple-dag}
\end{figure}

\Maple{} has also another representation, similar to the \gls{dag}, that allows multiple copies of the same object in its representation. This representation is called \inertF{} and is more similar to an expression tree than \Maple's \gls{dag}. The \inertF{} represents the \gls{dag} as a tree by splitting object copies and displays the tree in a list structure. Each node (or element in the list) has the prefix \texttt{\_Inert\_XXX}, that indicates the node is in the \inertF. The suffix \texttt{XXX} is the type of the node. Some of the important types are specified in table~\ref{tab:maple-types}. Note that this is only a subset of all possible types.

\begin{table}[ht]
\centering
\begin{tabular}{cp{10cm}}
	\hline
	Type & Explanation\\
	\hline
	SUM & Sums.\\
	PROD & Products.\\
	EXPSEQ & Expression sequence is a kind of list. The arguments of functions are stored in such sequences.\\
	INTPOS & Positive integers.\\
	INTNEG & Negative integers.\\
	COMPLEX & Complex numbers with real and imaginary part.\\
	FLOAT & Float numbers are stored in the scientific notation with integer values for the exponent $n$ and the significand $m$ in $m \cdot 10^n$.\\
	RATIONAL & Rational numbers are fractions stored in integer values for the numerator and positive integers for the denominator.\\
	POWER & Exponentiation with expressions as base and exponent.\\
	FUNCTION & Function invocation with the name, arguments and attributes of the function.\\
	\hline
\end{tabular}
\caption{A subset of important internal \Maple{} structures. See~\cite{MAPLE:ProgrammingGuide} for a complete list.}
\label{tab:maple-types}
\end{table}

\subsection{Maple's Open Maple API}
\Maple{} provides an \gls{api}, called \texttt{OpenMaple}~\cite[\S 14.3]{MAPLE:ProgrammingGuide} for interaction with the \Maple{} kernel. The \texttt{OpenMaple} \gls{api} is implemented for different programming languages such as \texttt{C}, \texttt{Java} and \texttt{Visual Basic}. Since our forward translation and the \gls{mlp} is implemented in \texttt{Java}, we use the \texttt{Java} \gls{api} for the backward translation. The \texttt{Java} \gls{api} has some limitations to interact with \Maple's \gls{dag}.

For the backward translation, we use the \inertF. This representation can be converted to a nested list version and the \texttt{OpenMaple} Java \gls{api} can handle lists. Therefore, the backward translation firstly converts the \inertF{} to a nested list structure before any translation process starts.

The following figure~\ref{fig:inertform-list} shows the \inertF{} and the nested list of the polynomial $x^2+x$.

\begin{figure}[ht]
\centering
\begin{minipage}{.45\linewidth}
\centering
\begin{tabular}{lll}
\multicolumn{3}{l}{\_Inert\_SUM(}\\
& \multicolumn{2}{l}{\_Inert\_POWER(}\\
& & \_Inert\_NAME("x"),\\
& & \_Inert\_INTPOS(2)\\
& ),\\ 
& \multicolumn{2}{l}{\_Inert\_NAME("x")}\\
)
\end{tabular}
\end{minipage}
\begin{minipage}{.45\linewidth}
\centering
\begin{tabular}{lll}
\multicolumn{3}{l}{[\_Inert\_SUM,}\\
& \multicolumn{2}{l}{[\_Inert\_POWER, }\\
& & [\_Inert\_NAME, "x"],\\
& & [\_Inert\_INTPOS, 2]\\
& ],\\ 
& \multicolumn{2}{l}{[\_Inert\_NAME, "x"]}\\
]
\end{tabular}
\end{minipage}
\caption{\inertF{} and the nested list representation of $x^2+x$.}
\label{fig:inertform-list}
\end{figure}

In the nested list, the first element specifies the type of the current list while the following arguments store the arguments and optional attributes.

\subsection{Workarounds for Problems}\label{subsec:maple-probs}
Note that the backward translations usually do not have problems with ambiguous expressions. Every \gls{cas} needs to solve ambiguities in its own representation, otherwise it could not compute the expression. Our backward translation from \Maple{} back to semantic \LaTeX{} is therefore based on already disambiguated expressions.

The general walkthrough of our backward translation starts with a string of a \texttt{1D} \Maple{} input expression. As described above, this string is entered in \Maple{} to parse the expression and work on the nested list version of the \inertF{} afterwards. This causes some problems. Mostly because \Maple{} automatically evaluates input expressions immediately - even before we have a chance to take a look at the internal data structure of the input. For example, an input such as $\sin@{\frac{\cpi}{2}}$ is immediately evaluated to $1$ and the internal data structure just shows us the positive integer value $1$ instead of the trigonometric function.

Furthermore, \Maple{} also changes input expressions slightly, mostly because of missing data types for a specific expression. For example, the internal data structure has no type to represent fractions except the \texttt{RATIONAL} type, which only allows integer values for the numerator and denominator. Thereby, an expression like $\frac{x}{2}$ is automatically changed to $x\frac{1}{2}$. There is no way no avoid such simple arithmetic changes in the expression, but we can avoid the automatic evaluations.

Firstly, we summarize all known changes and evaluations \Maple{} automatically performs on input expressions.

\begin{itemize}
\item \Maple{} evaluates input expressions immediately.
\item There is no data type to represents square roots such as $\sqrt{x}$ (or $n$-th roots). Therefore, \Maple{} stores roots as an exponentiation with a fractional exponent. For example, $\sqrt{x}$ is stored as $x^{\frac{1}{2}}$.
\item There is no data type for subtractions, only for sums. Negative terms are changed to absolute values times '$-1$'. For example, $x-y$ is stored as $x + y \cdot (-1)$. 
\item Floating point numbers are stored in the scientific notation with a mantissa and an exponent in the base $10$. For example, $3.1$ is internally represented as $31 \cdot 10^{-1}$.
\item There is only a data type for rational numbers (fractions with integer numerator and positive denominator), but not for general fractions, such as $\frac{x+y}{z}$. This will be automatically changed to $(x+y)\cdot z^{-1}$.
\end{itemize}

There are unevaluation quotes implemented to avoid evaluations on input expression. Table~\ref{tab:unevaluation-quotes} gives an example how those unevaluation quotes work.

\begin{table}[ht]
\centering
\begin{tabular}{lcc}
\hline
& Without unevaluation quotes & With unevalation quotes\\
\hline
Input expression: & \texttt{sin(Pi)+2-1} & \texttt{'sin(Pi)+2-1'}\\
Stored expression: & 1 & \texttt{sin(Pi)+1}\\
\hline
\end{tabular}
\caption{Example of unevaluation quotes for \texttt{1D} \Maple{} input expressions.}
\label{tab:unevaluation-quotes}
\end{table}

Since we want to keep a translated expression similar to the input expression, we use unevaluation quotes during the whole translation process. Unevaluation quotes also solve the problems with square roots and $n$-th roots. In \Maple's \texttt{1D} input representation, a square root is a function call (\texttt{sqrt(x)}) and the unevaluation quotes prevent evaluations on functions. Therefore, a backward translation simply needs to translate a square root as a normal function. Hence, we do not have problems with the internal representation of roots in \Maple.

Because of the absent data type for subtractions, a long sum with negative terms could be difficult to read. We change the order of constants (such as an integer value) to be in front of products. Therefore, an expression such as '$-y$' is not represented by '$y\cdot (-1)$', but by '$(-1) \cdot y$'. The translator now needs to check if there is a leading '$-1$' and can translate it to '$-y$' rather than to '$(-1) \cdot y$'.

Floating point numbers in scientific notation also causes a problem. Consider the input expression $0.41$ as the nested list \inertF{} representation
\begin{equation}
\texttt{[\_Inert\_FLOAT, [\_Inert\_INTPOS, 41], [\_Inert\_INTNEG, 2]]}.
\end{equation}

The backward translator is implemented in \texttt{Java}. To translate such an expression back to $0.41$, computations become necessary. Of course, it makes sense to perform those computations in the \gls{cas} (in that case, in \Maple) rather than in the translation program. Therefore, we implemented a procedure\footnote{\textit{Procedures} in \Maple{} are small programs similar to methods and functions in \gls{oop} languages.} to automatically convert \texttt{FLOAT} nodes to a string representation of the floating point number. We created a new type, called \texttt{MYFLOAT} to represent such numbers. With this approach, we avoided computations in the translation process.

For fractions, we use a similar approach and introduce a new internal data type \texttt{DIVIDE}. The implementation of \texttt{DIVIDE} was difficult. On the one hand, some expressions with negative exponent should be displayed as a fraction while others should should be displayed with the negative exponent. We follow the same rule as \Maple{} follows internally to display fractions. When the exponent is a negative numeric element, the expression should be displayed as a fraction. If the exponent is something else, we uses the negative exponent representation. Table~\ref{tab:maple-fracs} shows two examples with negative exponents. One of them is displayed as a fraction (exponent '$-2$') and the other is not (exponent '$-2y$'). We implement the same rule for our newly developed \texttt{DIVIDE} data type. Note that with the new created data type, we are still not able to avoid changes from $\frac{x}{2}$ to $x\frac{1}{2}$, but we are able to work with more general fraction expressions.

\begin{table}[ht]
\centering
\begin{tabular}{p{2.6cm}cc}
	\hline
	& Negative numeric exponent & General negative exponent\\
	\hline
	\texttt{1D} \Maple{} input expression & $\texttt{(x-1)\^{}(-2)}$ & \rule{0pt}{0.9\normalbaselineskip}$\texttt{(x-1)\^{}(-2*y)}$\\
	How \Maple{} displays the input & $\displaystyle\frac{1}{(x-1)^2}$ & $\displaystyle(x-1)^{-2y}$\\
	\hline
\end{tabular}
\caption{Different styles to display negative exponents depending on the type of the exponent. \Maple{} displays expressions with negative numeric exponents as fractions, while other negative exponents are not displayed as fractions.}
\label{tab:maple-fracs}
\end{table}

We implement all of these changes before the actual translation process starts. This helps keeping the translated expression similar to the input expression.

\subsection{Implementation Details}\label{subsec:maple-impl}
\begin{wrapfigure}{l}{0.485\textwidth}
	\vspace{-12pt}
	\centering
	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.39]{BackwardTranslator.pdf}
	\caption{A scheme of the backward translator and its specialized subtranslators.}
	\label{fig:BackwardDel}
	\vspace{-13pt}
\end{wrapfigure}

The general translation process runs through different states. Each translation process starts with the string representation of a \texttt{1D} \Maple{} input. This string will be converted to the internal nested list representation of our custom \inertF{} (see the previous subsection~\ref{subsec:maple-probs}). Since the \inertF{} is similar to an expression tree, it can be translated as the \gls{mlp-pt} in the forward translation. Hence, we use the same approach of multiple subtranslators for the backward translation process.

Figure~\ref{fig:BackwardDel} shows the subtranslators of the backward translation process. The structure is the same as for the forward translation, seen in figure~\ref{fig:ForwardDel}. The \verb|MapleInterface| represents the entry point of a translation process. Before the translation starts, the \Maple{} kernel needs to be initialized and prepared for the translation process by loading all custom procedures. Once everything is loaded, an expression is delegated to specialized subtranslators depending on the internal data type, defined by custom procedures and \Maple's \inertF.

The following figure~\ref{fig:backward-trans} illustrates the backward translation process for the Jacobi polynomial example $\JacobiP{\alpha}{\beta}{n}@{\cos@{a\Theta}}$ from table~\ref{tab:JacobiP-usecase}. The input expression is converted to the nested list representation of the \texttt{MyInertForm} (the customized \inertF{} of \Maple). Afterwards, each node of the tree is translated separately by specialized subtranslators (visualized by blue arrows). The translation of functions (bold blue arrows) is again realized by translation patterns, which define the correct position for each argument. The last step is used to put all translated arguments to the right position (visualized by red arrows).

\begin{figure}[!htp]
	\centering
	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.7]{MapleTranslation.pdf}
	\caption{A scheme of the backward translation process from \Maple{} for the Jacobi polynomial $\JacobiP{\alpha}{\beta}{n}@{\cos@{a\Theta}}$. The input string is converted by the \Maple{} kernel into the nested list representation. This list is translated by subtranslators (blue and red arrows). A function translation (bold blue arrows) is again realized by translation patterns to define the position of the arguments (red arrows).}
	\label{fig:backward-trans}
\end{figure}

\cleardoublepage

### 2018-05-22 chapters/03-translator/forward-trans.tex
Init commit with original master thesis

\section{The Forward Translation}\label{sec:forward-translation}
\glsreset{teo}
The translation process starts with semantic \LaTeX{} expressions. These expressions get analyzed and parsed by the \gls{mlp}, which produces a \gls{mlp-pt} of the input expression with additional information about each symbol. An abstract translator class analyzes each node and delegates them to specialized subtranslators. In this process an object called \gls{teo} is build. This \gls{teo} is needed to rebuild a string representation of the mathematical expression given by the \gls{mlp-pt}.

The overall forward translation process is explained in figure~\ref{fig:forward-trans}. All translation patterns and related information are stored in the DLMF/DRMF tables. These tables are converted by the \verb|lexicon-creator.jar| to the \verb|DLMF-macros-lexicon.txt| lexicon file. Together with the \verb|global-lexicon.txt| file, the \gls{mlp-pt} will be created by the \gls{mlp}. The \verb|latex-converter.jar| takes a string representation of a semantic \LaTeX{} expression and uses the \gls{mlp} as well as our \verb|Translator| to create an appropriate string representation for a specified \gls{cas}. The following \cref{subsec:analyze-mlp,subsec:delegation,subsec:translated-expr,subsec:add-info} will focus on the \verb|Translator|.

\begin{figure}[ht]
	\vspace{-10pt}
	\centering
	\includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm, scale=0.72]{ForwardTranslationProcess.pdf}
	\caption{Process diagram of a forward translation process. The MLP generates the PT based on lexicon and JSON files. The PT will be translated to different CAS.}
	\label{fig:forward-trans}
	\vspace{-10pt}
\end{figure}

\subsection{Analyzing the MLP-Parse Tree}\label{subsec:analyze-mlp}
The main task is to analyze the \gls{mlp-pt}. Remember that the \gls{mlp-pt} is different to expression trees as described in section~\ref{subsec:pom-mlp}.

Since the \gls{bnf} does not define rules for semantic macros, each argument of the semantic macro and each $@$ symbol are following siblings of the semantic macro node. That is the reason, why we stored the number of parameters, variables and $@$ symbols in the lexicon files. Otherwise, the translator could not find the end of a semantic macro in the \gls{mlp-pt}.

Figure~\ref{fig:syntax-tree-usecase} visualizes the \gls{mlp-pt} of the Jacobi polynomial example from table~\ref{tab:JacobiP-usecase}. Because of these differences to expression trees, a backward conversion of the \gls{mlp-pt} to a string representation can be difficult, especially for finding necessary or unnecessary parentheses. Therefore we create the \gls{teo}. Of course, the \gls{teo} is deeply incorporated into the translation process, but in this subsection we will focus on the general idea and process of the translation. We will focus on the usage of \gls{teo} in subsection~\ref{subsec:translated-expr}.

\begin{figure}[ht]
	\centering
	\includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm, scale=0.75]{SyntaxTreeUseCase.pdf}
	\caption{MLP-PT for a Jacobi polynomial using the DLMF/DRMF \LaTeX{} macro. Each leaf contains information from the lexicon files.}
	\label{fig:syntax-tree-usecase}
\end{figure}

The general idea of a translation of a \gls{mlp-pt} is a simple recursive algorithm, such as algorithm~\ref{alg:simple-translation}. Whenever the algorithm finds a leaf, it can translate this single term. If the node is not a leaf, it starts to translate all children of the node recursively.

\begin{wrapfigure}{r}{0.5\textwidth}
\vspace{-20pt}
\begin{minipage}{0.5\textwidth}
\begin{algorithm}[H]
\caption{Simple translation algorithm for the MLP-Parse Tree}\label{alg:simple-translation}
	\begin{algorithmic}[1]
	\Require Root $r$ of the parse tree $T$
	\Procedure{translate}{$r$}
	\If{$r$ is leaf}
		\State {\scriptsize TRANSLATE\_LEAF}($r$);
	\Else
		\ForAll{children $v_n$ of $r$}
			\State {\scriptsize TRANSLATE}($v_n$);
		\EndFor
	\EndIf
	\EndProcedure
	\end{algorithmic}
\end{algorithm}
\end{minipage}
\vspace{-14pt}
\end{wrapfigure}

But this approach is not completely feasible, because sometimes the algorithm needs to look ahead and check the following siblings for a valid translation, for example in case of a semantic macro with arguments (see figure~\ref{fig:syntax-tree-usecase}). Therefore, the {\footnotesize \verb|TRANSLATE_LEAF|} function also needs to know the following siblings of the current node. To avoid multiple translations of the same node, a list representation would be a better solution. Algorithm~\ref{alg:translation} explains an abstract version of the final algorithm for our translation tool.

If the root $r$ is a leaf, it can be translated as a leaf. Eventually, some of the following siblings are needed in order to translate $r$ correctly. If $r$ is not a leaf, it contains one or more children. Therefore, we can call the {\footnotesize \verb|ABSTRACT_TRANSLATOR|} recursively for the children. Once we have translated $r$, we can go a step further and translate the next node. Line~\ref{line:ifNotEmpty} checks if there are following siblings left and calls the {\footnotesize \verb|ABSTRACT_TRANSLATOR|} recursively in that case.

\begin{algorithm}[ht]
\caption{Abstract translation algorithm to translate MLP-Parse trees.}\label{alg:translation}
	\begin{algorithmic}[1]
	\Require Root $r$ of a MLP-Parse tree $T$. List \textit{following\_siblings} with the following siblings of $r$. The list can be empty.
	\Procedure{abstract\_translator}{$r$, \textit{following\_siblings}}
	\If{$r$ is leaf}
		\State {\scriptsize TRANSLATE\_LEAF}($r$, \textit{following\_siblings\_of\_r});\label{line:transLeaf}
	\Else
		\State \textit{siblings} $ = r$.getChildren(); \Comment{\textit{siblings} is a list of children}
		\State {\scriptsize ABSTRACT\_TRANSLATOR}(\textit{siblings}.removeFirst(), \textit{siblings});\label{line:transSeq}
	\EndIf
	\If{\textit{following\_siblings} is not empty}\label{line:ifNotEmpty}
		\State $r =$ \textit{following\_siblings}.removeFirst();
		\State {\scriptsize ABSTRACT\_TRANSLATOR}($r$, \textit{following\_siblings});
	\EndIf
	\EndProcedure
	\end{algorithmic}
\end{algorithm}

With this approach we are able to translate most of the semantic \LaTeX{} expressions. To take a look at the problems with this approach, we need to think about the operators and their standard notations in mathematics. There are many types of notation used to represent formulae. For example, the \gls{pol}\footnote{Also known as \textit{Warsaw Notation} or \textit{prefix notation}} (hereafter called prefix notation) places the operator to the left of/before its operands. The \gls{rpol}\footnote{Also known as \textit{postfix notation}} (hereafter called postfix notation) does the opposite and places the operator to the right of/after its operands. The infix notation is commonly used in arithmetic and places the operator between their operands, which only makes sense as long as the operator is a binary operator. For example, table~\ref{tab:notations} shows an expression in different notations.

\begin{table}[!ht]
\centering
\begin{tabular}{lc}
	\hline
	Notation & Expression \\
	\hline
	Infix & $(a+b) \cdot x$\\
	Prefix & $\cdot + a\ b\ x$\\
	Postfix & $a\ b + x\ \cdot$\\
	Functional & $\cdot(+(a, b), x)$\\
	\hline
\end{tabular}
\caption{The mathematical expression '$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
\label{tab:notations}
\end{table}

In mathematical expressions, notations are mostly mixed, depending on the case and number of operands. For example, infix notation is common for binary operators ($+$, $-$, $\cdot$, $\mod$ etc.), while functional notations are obviously used for any kind of functions ($sin$, $cos$, etc.). Sometimes the same symbol is used in different notations to identify a different meaning. For example, the '$-$' as an unary operator is used in prefix notation to indicate the negative value of its operand, such as in '$-2$'. Of course, $-$ can also be the binary operator for subtraction, which is commonly used in infix notation. An example for the postfix notation is factorial, such as '$2!$'.

Most programming languages (and \gls{cas} as well) internally use prefix or postfix notation and do not mix the notations in one expression, because it is easier to parse those notations. However, the common practice in science is to use mixed notations in expressions. Since the \gls{mlp} has barely implemented mathematical grammatical rules, it takes the input as it is and does not build an expression tree. Therefore, it parses all four examples from table~\ref{tab:notations} to four different \gls{mlp-pt}s rather than to one unique expression tree. In general, this is not a problem for our translation process because most \gls{cas} are familiar with the most common notations. Therefore, the translator does not need to know that $a$ and $b$ are the operands of the binary operator '$+$' in '$a+b$'. We could just translate the symbols in '$a+b$' in the same order as they appear in the expression and the \gls{cas} would understand it. However, this simple approach generates two problems.
\begin{enumerate}%[label=\arabic*)]
\item \label{prob:1} The translated expression is only syntactically correct if the input expression was syntactically correct.
\item \label{prob:2} We cannot translate expressions to a \gls{cas} which uses a different notation.
\end{enumerate}

Problem~\ref{prob:1} should be obvious. Since we want to develop a translation tool and not a verification tool for mathematical \LaTeX{} expressions, we can assume syntactically correct input expressions and produce errors otherwise. Problem~\ref{prob:2} is more difficult to solve. If a user wants to support a \gls{cas} that uses prefix or postfix notation by default, the whole translator would fail in its current state. Supporting \gls{cas} with another notation would be a part of future work and will be mentioned in chapter~\ref{ch:conc-future-work}.

Nonetheless, changing a notation could also solve ambiguities in some situations. Consider the two ambiguous examples in table~\ref{tab:amb_ex}. While a scientist would probably just ask for the right interpretation of the first example, \Maple{} automatically computes the first interpretation. On the other hand, \LaTeX{} automatically disambiguate the first example by only recognizing the very next element (single symbols or sequence in curly brackets) for the superscript and therefore displays the second interpretation. The second example is already interpreted as the double factorial function of $n$, since this notation is the standard interpretation in science. We wrote the second interpretation as the standard way in science to make it even more obvious. However, surprisingly, \Maple{} computes the first interpretation again rather than the common standard interpretation.
\begin{table}[ht]
\centering
\begin{tabular}{lccc}
	\hline
	& Text Format Expression & First Interpretation & Second Interpretation\\
	\hline
	1: & \rule{0pt}{0.9\normalbaselineskip} $4\ \hat{\ }\ 2!$ & $4^{2!}$ & $4^2!$ \\
	2: & $n!!$ & $(n!)!$ & $(n)!!$\\
	\hline
\end{tabular}
\caption{Ambiguous examples of the factorial and double factorial function. One expression in a text format can be interpreted in different ways.}
\label{tab:amb_ex}
\end{table}

In most cases, parentheses can be used to disambiguate expressions. We used them in table~\ref{tab:amb_ex} to clarify the different interpretations in example 2. But sometimes, even parentheses cannot solve a mistaken computation. For example, there is no way to add parentheses to force \Maple{} to compute $n!!$ as the double factorial function. Even $(n)!!$ will be interpreted as $(n!)!$. Rather than using the exclamation mark in \Maple, one could also use the functional notation. For example, the interpretations $(2!)!$ and $(2)!!$ can be distinguished in \Maple{} by using \verb|factorial(factorial(2))| and \verb|doublefactorial(2)| respectively.

As already mentioned, the structure of the \gls{mlp-pt} makes it difficult to change the notation for all kind of operators. Therefore, and especially because of the examples above, we change the notation during the forward translation process to the functional notation only for the factorial and double factorial function. Thus,
\begin{eqnarray*}
\verb|n!| &\overset{\langMaple}{\mapsto}& \verb|factorial(n)|,\\
\verb|n!!| &\overset{\langMaple}{\mapsto}& \verb|doublefactorial(n)|.
\end{eqnarray*}
The translator needs to presume some properties for the functions similar to \LaTeX, which only recognizes the very next element right after the caret for the superscript.

The biggest problem about translating the factorial (or double factorial) function is the common postfix notation for them. In this stage, algorithm~\ref{alg:translation} only translates the current node and if necessary the following siblings, but not the preceding siblings. Consider an expression such as '$4!$'. When the variable $r$ reaches the exclamation mark, the argument '$4$' of the function has already been translated. The current version of the algorithm has no access to this previously translated expression. Even more complicated, would be the case of the double factorial '$4!!$'. The following two approaches would solve this problem.

\begin{enumerate}
\item\label{appr:1} The translator always checks if the next two following siblings are one or two exclamation marks.
\item\label{appr:2} The current state of program has access to the previously translated expressions and can modify them retrospectively.
\end{enumerate}

Approach~\ref{appr:1} would be unnecessarily inefficient. Therefore we implement approach~\ref{appr:2}. That is another reason why we implement an extra object to organize translated expressions, the \gls{teo}. This object stores all parts of the previously translated expressions. The translator has access to these parts and is able to modify them afterwards. A more detailed explanation about the \gls{teo} follows in subsection~\ref{subsec:translated-expr}.
\newpage

\input{chapters/03-translator/all-types-table.tex}
\newpage

\subsection{Delegation of Translations}\label{subsec:delegation}

\begin{wrapfigure}{l}{0.45\textwidth}
	\vspace{-20pt}
	\centering
	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.4]{ForwardTranslator.pdf}
	\caption{A scheme of the forward translator and its specialized subtranslators.}
	\label{fig:ForwardDel}
	\vspace{-25pt}
\end{wrapfigure}

Algorithm~\ref{alg:translation} is a simplified version of the translator process. The main progress happens in \cref{line:transLeaf,line:transSeq}. There are several cases what $r$ can be. Table~\ref{tab:allTypesTable} on page \pageref{tab:allTypesTable} gives an overview of all different types in a \gls{mlp-pt}. A more detailed explanation of the types can be found in~\cite{POM-Tagger}.

The \gls{bnf} grammar defines some basic grammatical rules for generic \LaTeX{} macros, such as for \verb|\frac|, \verb|\sqrt|. Therefore, there is a hierarchical structure for those symbols, similar to the structure in expression trees. As already mentioned, some of these types can be translated directly, such as Greek letters, while others are more complex, such as semantic \LaTeX{} macros. Therefore, the translators delegates the translation to specialized subtranslators. This delegation process is implemented in \cref{line:transLeaf,line:transSeq} of algorithm~\ref{alg:translation}.

Figure~\ref{fig:ForwardDel} is a scheme of the forward translator and its subtranslators. The entry point for each translation process of the program is the \verb|SemanticLatexTranslator|. It delegates the translations for the expression through the \verb|AbstractTranslator| to the other subtranslators. The \verb|AbstractTranslator| is the interface of all other translators and coordinates the translation process by delegating translations for specific nodes to specialized classes. For example, the \verb|MacroTranslator| only translates semantic \LaTeX{} macros. The \verb|EmptyExpressionTranslator| is used to delegate the so called \textit{empty expressions} to other classes. An \textit{empty expression} is a node with children, but without a symbol itself, such as \textit{sequences} and \textit{balanced expressions}.

\subsubsection{Subtranslators}
The most interesting translators are the \verb|MacroTranslator| (translating semantic \LaTeX{} macros) and the \verb|SequenceTranslator| (translating all kinds of \textit{sequences}).

The \verb|SequenceTranslator| translates the \textit{sequence} and \textit{balanced expressions} in the \gls{mlp-pt}. If a node $n$ is a leaf and the represented symbol an open bracket (parentheses, square brackets and so on) the following nodes are also taken as a \textit{sequence}. Hence, combined with the recursive translation approach, the \verb|SequenceTranslator| also checks balances of parentheses in expressions. An expression such as '$(a]$' is producing a mismatched parentheses error. On the other hand, this is a problem for interval expressions such as '$[a,b)$'. In the current version, the program cannot distinguish between mismatched parentheses and an half-opened, half-closed interval. Whether an expression is an interval or another expression is difficult to decide and can depend on the context. In addition, the parenthesis checker could simply be deactivated to allow mismatched parentheses in an expression. But this functionality is not implemented yet.

Another problem that the \verb|SequenceTranslator| solves is the position of multiplication signs in an expression. There are a couple of obvious choices to translate multiplication signs. For example, using \verb|\cdot| or \verb|\idot| will be obviously translated as a multiplication symbol. However, \verb|\idot| is fairly new and therefore not frequently used yet. Furthermore, the most common symbol for multiplications is still the white space or none symbol at all, as explained above. Consider the simple expression $2n\pi$. The \gls{mlp-pt} generates a sequence node with three children, namely $2$, $n$ and $\pi$. This sequence should be interpreted as a multiplication of the three elements. Therefore, the \verb|SequenceTranslator| checks the types of the current and next nodes in the tree to decide if there should be a multiplication symbol or not. For example, if the current or next node is an operator, a relation symbol or an ellipsis there will be no multiplication symbol added. However, this approach implies an important property. The translator interprets all sequences of nodes as multiplications as long as it is not defined otherwise. This potentially produces strange effects. Consider an expression such as $f(x)$. Translate this to \Maple{} will be \verb|f*(x)|. But we do not consider this translation to be wrong, because there is a semantic macro to represent functions. In this case, the user should use \verb|\f{f}@{x}| instead of \verb|f(x)|.

Besides the translation of sequences, the translation process of the \Macro s can be complex too; not only because of the structure of the \gls{mlp-pt}, but also because of the complex definition of the semantic macros. As explained in \ref{subsec:macros}, they can contain optional parameters, exponents are allowed before and after the arguments and the number of $@$ symbols varies. All these different cases will be solved by the \verb|MacroTranslator|.

Algorithm~\ref{alg:macro-translation} is the translation function of the \verb|MacroTranslator| without error handling. It analyzes the next element right after the macro first. As mentioned, this can be one of the following:
\begin{itemize}
\item an exponent such as '\verb|^2|'.
\item an optional parameter in square brackets.
\item a parameter in curly brackets (a \textit{sequence} node in the \gls{mlp-pt}) if none of the above.
\item an $@$ symbols if none of the above.
\item a variable in curly brackets (a \textit{sequence} node) if none of the above.
\end{itemize}
\begin{algorithm}[!ht]
\caption{The translate function of the MacroTranslator. This code ignores error handling.}\label{alg:macro-translation}
	\begin{algorithmic}[1]
	\Require 
		\Statex \textit{macro} - node of the semantic macro. 
		\Statex \textit{args} - list of the following siblings of \textit{macro}. 
		\Statex \textit{lexicon} - lexicon file
	\Ensure 
		\Statex Translated semantic macro.

	\Procedure{translate\_macro}{\textit{macro}, \textit{args}, \textit{lexicon}}
	\State \textit{info} = \textit{lexicon}.getInfo(\textit{macro});
	\State \textit{argList} = new List(); \Comment{create a sorted list for the translated arguments.}
	\State \textit{next} = \textit{args}.getNextElement();
	\If{\textit{next} is caret}\label{line:next_caret}
		\State \textit{power} = translateCaret(\textit{next});
		\State \textit{next} = \textit{args}.getNextElement();
	\EndIf

	\While{\textit{next} is $[$}\label{line:next_optional} \Comment{square brackets indicate optional arguments.}
		\State \textit{optional} = {\scriptsize TRANSLATE\_UNTIL\_CLOSED\_BRACKET}(\textit{args});
		\State \textit{argList}.add(\textit{optional});
		\State \textit{next} = \textit{args}.getNextElement();
	\EndWhile

	\State \textit{argList}.add( {\scriptsize TRANSLATE\_PARAMETERS}(\textit{args}, \textit{info}) ); \label{line:trans_paras}
	\State {\scriptsize SKIP\_AT\_SIGNS}( \textit{args}, \textit{info} ); \label{line:skip_ats}
	\State \textit{argList}.add( {\scriptsize TRANSLATE\_VARIABLES}(\textit{args}, \textit{info}) ); \label{line:trans_vars}

	\State \textit{pattern} = \textit{info}.getTranslationPattern();
	\State \textit{translatedMacro} = \textit{pattern}.fillPlaceHolders(\textit{argList});
	\If{\textit{power} is not \NULL}
		\State \textit{translatedMacro}.add(\textit{power});
	\EndIf
	\State \Return\ \textit{translatedMacro};
	\EndProcedure
	\end{algorithmic}
\end{algorithm}

An exponent will be translated and shifted to the end of the translated semantic macro, since it is more common to write the exponent after the arguments of a function in \gls{cas}. Therefore, the function translates and stores the exponent in line~\ref{line:next_caret}. One could ask what happens when there is an exponent given before and after the arguments. The \verb|MacroTranslator| only translates the following siblings until each argument is translated. The first exponent will be shifted to the end. If right after the translated macro (with all arguments) follows another exponent, we interpret it as another exponent for the whole previous expression. In that case, it would be the macro with the first translated exponent. Table~\ref{tab:multi-expo} shows an example for the trigonometric cosine function with multiple exponents. 
\begin{table}[ht]
\centering
\begin{tabular}{lc}
\hline
Displayed As & \rule{0pt}{0.9\normalbaselineskip} $\cos^2(x)^2$ \\
Semantic \LaTeX & \verb|\cos^2@{x}^2| \\
Translated \Maple{} Expression & \verb|((cos(x))^(2))^2|\\
\hline
\end{tabular}
\caption{A trigonometric cosine function example with exponents before and after the argument.}
\label{tab:multi-expo}
\end{table}

As you can see, the input expression in semantic \LaTeX{} is interpreted as~\eqref{eq:multi-expo}.
\begin{equation}\label{eq:multi-expo}
\left( \cos(x)^2 \right)^2
\end{equation}

An open square bracket right after the semantic macro (after the exponent respectively) is an optional parameter. Since expressions in square brackets are not considered to be a \textit{sequence} in the \gls{mlp-pt}, the optional parameter is not grouped in a single node. Therefore, every node after the opening square bracket forms an expression for just one optional parameter until the algorithm reaches the closing square bracket. Such expressions in square brackets will be also translated by the \verb|SequenceTranslator| as described above. A semantic macro could have multiple optional parameters. Therefore, the function translates all optional parameters and stores them in line~\ref{line:next_optional}.

The \cref{line:trans_paras,line:skip_ats,line:trans_vars} translates the parameters and variables. The $@$ symbols will be skipped. Nevertheless, the current program has an error handling implemented in \cref{line:skip_ats} if the maximum number of $@$ symbols is exceeded. In the following lines, the function takes the translation pattern from the lexicon and removes each placeholder by the corresponding argument. In addition, if there is an exponent right after the semantic macro, it will be attached to the translated semantic macro (see table~\ref{tab:multi-expo}). The parameters and variables are always grouped by curly brackets in \textit{sequence}-nodes in the \gls{mlp-pt}. Therefore, every argument after the optional parameters and potential exponent cannot be longer than one node in the \gls{mlp-pt}.

\subsubsection{Handle Ambiguities}\label{subsec:ambiguities}
Our semantic \LaTeX{} expressions are not full semantic with respect to possible ambiguities in the expression. We already mentioned some techniques to solve ambiguities in semantic \LaTeX{} expressions, for example how to handle double factorials. In table~\ref{tab:amb-latex} are four examples of ambiguous expressions. \LaTeX{} implemented the rules that the superscript or subscript is only one symbol (or a balanced expression in curly brackets). Therefore, each of these input expressions are not ambiguous in \LaTeX. Since we talking about the forward translation, we should follow the rules of \LaTeX{} to interpret ambiguous expressions.

\begin{table}[ht]
\centering
\begin{tabular}{cc}
	\hline
	Ambiguous Input & \LaTeX{} Output\\
	\hline
	\verb|n^m!| & $n^m!$\\
	\verb|a^bc^d| & $a^bc^d$\\
	\verb|x^y^z| & Double superscript error\\
	\verb|x_y_z| & Double subscript error\\
	\hline
\end{tabular}
\caption{Ambiguous \LaTeX{} expressions and how \LaTeX{} displays them.}
\label{tab:amb-latex}
\end{table}

Another more questionable translation decision are alphanumerical expressions. As explained in table~\ref{tab:allTypesTable}, the \gls{mlp} handles strings of letters and numbers differently, depending on the order of the symbols. The reason is, that an expression such as '$4b$' is usually considered to be a multiplication of $4$ and '$b$', while '$b4$' looks like indexing '$b$' by $4$. While the first example produces two nodes, namely $4$ and '$b$', the second example '$b4$' produces just a single alphanumerical node in the \gls{mlp-pt}. The first expression would be translated as a multiplication of $4$ and '$b$'. Because '$4b$' looks like '$4 \cdot b$' and the multiplication is commutative, we would assume that '$4b$' and '$b4$' are equivalent. This is one reason why we interpret such alphanumerical expressions as multiplications of the symbols. On the other hand, an expression such as '$energy$', would be an alphanumerical node as well and could be interpreted as a single variable named 'energy'. In that case, someone could wonder why the variable '$energy$' was translated as '$e \cdot n \cdot e \cdot r \cdot g \cdot y$'. While in physics or engineering variables possibly appear with longer names, such as the 'energy' example, it is more common to use single symbols for variables in mathematics~\cite{Notation:History}. Therefore, we interpret such alphanumerical expressions as multiplications of variables rather than one variable with an alphanumerical name.

Another, more elegant way to solve this problem would be to use the newly created semantic macro \verb|\idot| (see subsection~\ref{subsec:semanticVSmacro}) to define an invisible multiplication symbol. However, because \verb|\idot| is such a new semantic macro, it is not used in the \gls{dlmf} or \gls{drmf} quite often yet. Therefore, our program still interprets alphanumerical expressions as multiplications.

Other ambiguities can appear, when the input expression contains a symbol, which is typically used to represent a specific mathematical object, but the user did not use the semantic macro to represent it. For example, the input expressions '$3i$' contains an '$i$', which is usually associated with the imaginary unit. Since there is a semantic macro for the imaginary unit, namely \verb|\iunit|, we do not translate the letter to the imaginary symbol in \Maple{} and use the identity translation instead. Furthermore, we inform the user about the potential misusage of \verb|i| (see~\ref{subsec:add-info}).

In general the translator is drafted to solve ambiguous expressions or automatically find a workaround to disambiguate the expression. Only if there is no way to solve the ambiguity with the defined rules, the translation process stops.

\subsection{Translated Expression Objects}\label{subsec:translated-expr}
\glsreset{teo}
\begin{wrapfigure}{r}{0.5\textwidth}
	\vspace{-20pt}
	\centering
	\includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm, scale=1]{SyntaxTreeGroup.pdf}
	\caption{The MLP-PT for '$\cos(2n\pi)!$' and the grouped argument of the factorial function.}
	\label{fig:syntax-group}
	\vspace{-15pt}
\end{wrapfigure}

It is sometimes necessary to provide access to already translated symbols or subexpressions. A typical example is the (double) factorial function, where the argument was translated before the function was registered by the translator (postfix notation). To realize this, the \gls{teo} is implemented. It is mainly a list object of translated expressions. It also groups subexpressions when it becomes clear that multiple symbols belong together. This typically happens when something is wrapped in parentheses but also when a function is completely translated.

Figure~\ref{fig:syntax-group} shows an example for the expression '$\cos(2n\pi)!$'. If we just analyze the \gls{mlp-pt}, the previous node of the exclamation mark (which indicates the factorial function) is the sequence. Now it could be possible to misinterpret the sequence as the argument of the factorial function. But the translator will first translate the cosine function, which has one argument. In that case the argument is the sequence node. Once the cosine function has been completely translated (which means the whole argument has been translated as well) the \gls{teo} groups the translated cosine and the sequence. Therefore, when the translator reaches the exclamation mark and asks for the last translated object in the list, it is not the translated sequence but the whole cosine function.

Table~\ref{tab:teo-list} shows some examples of the \gls{teo} list after the translation process was finished. Note that a translation of '$a+b$' contains three translated objects, while '$(a+b)$' contains just one.

\begin{table}[ht]
\centering
\begin{tabular}{cc}
	\hline
	Input Expression & TEO List\\
	\hline
	$a+b$ & \verb|[a, +, b]|\\
	$(a+b)$ & \verb|[(a+b)]|\\
	$\frac{a}{b}-2$ & \verb|[(a)/(b), -, 2]|\\
	\hline
\end{tabular}
\caption{How the TEO-list groups subexpressions.}
\label{tab:teo-list}
\end{table}

\subsubsection{Local \& Global Translated Expressions}
During the translation process multiple local \gls{teo} are created by the subtranslators. Besides the local \gls{teo} there is just one global \gls{teo}. The reason to create local objects is originally motivated by the recursive structure of the translator. A subtranslator should only handle the current subexpression. However, the global \gls{teo} becomes necessary because of the functions in postfix notation such as the factorial function. For example, when a subtranslator tries to translate the exclamation mark as the factorial function, the local \gls{teo} is empty, because the argument of the function has been previously translated by another subtranslator and stored in another local \gls{teo}. Therefore, we implement a global \gls{teo}, which contains all of the previous translated subexpressions in each state of the translation process. With this implementation the local \gls{teo} becomes more and more obsolete. In the current version, the program stores redundant translated expressions. For debugging reasons, a translation process still handles multiple local and one global \gls{teo}. Starting the forward translation with the flag \verb|--debug| (for the debugging mode) shows the elements of the global \gls{teo} after the translation has been finished.

\subsection{Additional Information}\label{subsec:add-info}
As already mentioned, a translation is not always straight forward. In those cases, the user should get informed about the translation and the reason for it. Consider the translation
\begin{equation}\label{eq:i-trans}
\verb|3i| \overset{\langMaple}{\mapsto} \verb|3*i|,
\end{equation}
Since the imaginary unit in \Maple{} is associated with '\raw{I}' instead of '\raw{i}', a user might be confused by~\ref{eq:i-trans}. Therefore, we created the \texttt{InformationLogger} class to organize further information about the translation process. In the case of '\raw{i}', the user is informed about the semantic macro \verb|\iunit| and the potential misusage of '\raw{i}'.

Essentially, the \raw{InformationLogger} provides further information about a translation process. This information is defined in the lexicon files. Consider the example of the arccotangent function, which is defined with a different branch cut in \gls{dlmf} and in \Maple. As already mentioned, we provide alternative translation patterns to solve these problems. These alternative translations, explanations, branch cuts, domains etc. will be automatically stored in the \raw{InformationLogger} when an expression contains a mathematical object with further information.

### 2018-05-22 chapters/03-translator/gen-app.tex
Init commit with original master thesis

The overall goal of the whole project was to provide an interactive one click-translation for each formula in the \gls{drmf} to a specified \gls{cas}. The program mainly supports translations from and to the \gls{cas} \Maple. However, the program is designed to support other \gls{cas}. In the current stage it also supports translations to \Mathematica{}, but the backend library needs to get extended for a full support.

In this chapter, the functionality and the design of the translator will be explained. The structure follows the development process and explains our decisions, problems and solutions. It starts with an overview of our approaches in~\ref{sec:gen-appr} and solutions for the problems. We will also focus on the theory of our backend libraries in~\ref{subsec:libs} and how to extend them in~\ref{subsec:extension}. The sections~\ref{sec:forward-translation} and \ref{sec:backward-translation} explains the forward and backward translator separately.

\section{General Approach and Goals}\label{sec:gen-appr}
Initially this project aimed for translations not only for semantic \LaTeX{} expressions but for generic \LaTeX{} as well. However, as discussed before, the semantic information in generic \LaTeX{} is hardly accessible (as seen in \ref{subsec:semantic-latex}). Supporting generic \LaTeX{} translations would go beyond the scope of a Master's thesis. Hence, we focused on translations for semantic \LaTeX{} expressions. Furthermore, the translator should support multiple \gls{cas} and because of the work in progress state of the \Macro{} it should be easily extendable and changeable. Due to the time limitations of the project we focused our implementations on the support for \Maple{} and \Mathematica.

Another strive for a goal was to keep a translation as straight forward as possible. In other words, the structure of the translated expression should be similar to the structure of the input expression. This makes it easier for a user to follow the translation process and debug the program. This intention should not to be confused with the task for \textit{appropriate} translations. We will see that a translation process (especially the backward translation) can easily change expressions without changing the semantics. We want to avoid that behavior in our translations as often as possible.

Furthermore, the translation process should solve problems of existing technologies, presented in section~\ref{sec:related-work}. In the following subsection, we will focus on known problems and present solutions before we start to explain implementation details.

\subsection{Translation Problems}
Most of the existing translation tools cannot preserve the semantic information during the translation process (see section~\ref{sec:related-work}). Therefore, our program should not lose semantic information during the process. However, the semantics can be changed during a translation. This happens when a definition or property of a mathematical object varies from system to system. Our reference systems for all mathematical \LaTeX{} expressions are the \gls{dlmf} and the \gls{drmf}.

Let us focus on translations of mathematical objects that are not defined in one of the languages, first. Obviously, we cannot translate these objects \textit{directly}. Our workaround is to translate not the mathematical object itself, but the definition of the object. For example, the \textit{Gudermannian}~\cite[(4.23.10)]{NIST:DLMF} $\Gudermannian@{x}$ function can be defined by
\begin{equation}\label{eq:gudermannian}
\Gudermannian@{x} := \atan@{\sinh@{x}} \quad x \in \Real.
\end{equation}
This function is not implemented in \Maple. Therefore, we translate its definition~\eqref{eq:gudermannian}
\begin{equation}\label{eq:guder-trans}
\verb|\Gudermannian@{x}| \overset{\langMaple}{\mapsto} \texttt{arctan(sinh(x))}.
\end{equation}

We want to avoid such translations, because they are not intuitive and therefore can confuse a user. However, providing such translations is still better than providing no translation.

Instead of none possible translation, how can we handle multiple alternative translations? We already discussed the problem of the inverse trigonometric function $\acot@{x}$. Because of a different branch cut in \Maple{} (compared to the \gls{dlmf} definition), there are alternative translations available (see~\eqref{eq:acot-alternatives}). In such cases, we define the most intuitive, the direct translation. In addition, we inform the user about the available alternative translations and the branch cut issues. Hence,
\begin{equation}
\verb|\acot@{x}| \overset{\langMaple}{\mapsto} \texttt{arccot(x)}.
\end{equation}

Since we have these alternative translations, we also use the same technique for the previous Gudermannian example. We provide one direct translation and possible alternative translations for a mathematical object.

These two problems already explained, why our translation process is in general not bijective. However, we will explain later in a more detailed way that the translation is only bijective for a subset of possible expressions. Thereby, a bigger goal is to increase this subset as much as possible.

\subsection{Libraries}\label{subsec:libs}
The first step for providing an automatic translation tool is to defining the translation for each function. We manually create a table of translations for each \Macro. Translations for generic \LaTeX{} functions, mathematical symbols, Greek letters and constants are organized in \gls{json} files for a better performance.

For each translation, we design a translation pattern with placeholders to define the correct position of each argument. The placeholders are defined as \verb|$i|, where \verb|i| is a non-negative integer specifying the position, starting at $0$. Table~\ref{tab:placeholder_ex1} illustrates the translation patterns for the trigonometric sine function in semantic \LaTeX, \Maple{} and \Mathematica.
\begin{table}[ht]
	\centering
	\begin{tabular}{lc}
		\hline
		Semantic \LaTeX & \verb|\sin@{$0}| \\
		\Maple & \verb|sin($0)| \\
		\Mathematica & \verb|Sin[$0]|\\
		\hline
	\end{tabular}
	\caption{Translation patterns for the sine function in the \DLMF, \Maple{} and \Mathematica.}
	\label{tab:placeholder_ex1}
\end{table}

Placeholders gives us the opportunity to define the order of the arguments and handle nested function calls. The order of arguments can differ between the systems. For example, if a function possesses parameters written in sub- and superscripts, \Maple{} and \Mathematica{} usually put the parameters in the subscript before the parameters in the superscripts. The notation in the \Macro s uses the reversed order. Table~\ref{tab:JacobiP-usecase} in the introduction shows an example of such different orders. The translation patterns for the Jacobi polynomial are illustrated in table~\ref{tab:placeholder_ex2}.
\begin{table}[ht]
	\centering
	\begin{tabular}{lc}
		\hline
		\multicolumn{2}{l}{\textit{Forward Translation:}} \\
		\Maple & \verb|JacobiP($2, $0, $1, $3)| \\
		\Mathematica & \verb|JacobiP[$2, $0, $1, $3]|\\
		\hline
		\multicolumn{2}{l}{\textit{Backward Translation from \Maple/\Mathematica:}} \\
		Semantic \LaTeX & \verb|\JacobiP{$1}{$2}{$0}@{$3}|\\
		\hline
	\end{tabular}
	\caption{Forward and backward translation patterns of the Jacobi polynomial. The pattern for the backward translation is the same for \Maple{} and \Mathematica.}
	\label{tab:placeholder_ex2}
\end{table}

Note that \Maple{} also uses the symbol \verb|$| in few functions. This can cause trouble in the replacement process. However, we can handle all of the few cases by adding addition parenthesis. Consider the differentiation function in \Maple
\begin{equation}
\verb|diff(f, [x$n])|,
\end{equation}
with $f$ as an algebraic expression or an equation, $x$ the name of the differentiation variable and $n$ for the $n$-th order differentiation. Therefore
\begin{equation}
\deriv[2]{x^2}{x} = \verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [x$2])|.
\end{equation}
But the translation pattern in \Maple{} would be
\begin{equation}
\verb|diff($1, [$2$$0])|.
\end{equation}
Replace the placeholders sequentially by the arguments ends in
\begin{eqnarray*}
&&\verb|diff($1, [$2$$0])|\\
\text{Replace index 0 by $2$\hphantom{$^2$}}& \rightarrow &\verb|diff($1, [$2$2])|\\
\text{Replace index 1 by $x^2$}& \rightarrow &\verb|diff(x^2, [$2$2])|\\
\text{Replace index 2 by $x$\hphantom{$^2$}}& \rightarrow &\verb|diff(x^2, [xx])|.
\end{eqnarray*}
We use parenthesis to solve this issue and define the translation pattern with
\begin{equation}
\verb|diff($1, [$2$($0)])|.
\end{equation}

A special problem can be binary operators. A good example is the multiplication sign. In \LaTeX{} most scientist would use white spaces or \verb|\cdot| to indicate a multiplication (using none of them can cause ambiguities, see \ref{subsec:ambiguities}). Similar to that approach, \Mathematica{} also understands white spaces and asterisks as multiplications. When we translate semantic \LaTeX{} expressions to \Maple, we use the {\sf 1D \Maple{}} representation (see subsection~\ref{subsec:maple}), where an asterisk becomes mandatory for multiplications. White spaces would produce a syntactical error in this input format. On the other hand, white spaces in \LaTeX{} do not necessarily indicate a multiplication. It could also be used to improve the readability. Whether a white space indicates a multiplication or not is a semantic information in the expression. We introduce a new \DLMF{} macro that indicates a multiplication, but not gets rendered. This macro is \verb|\idot|, for an invisible dot.

As already mentioned, the information is stored in three different file types. Together they form the backend library.

\subsubsection{CSV Tables}
The \gls{csv} file represents the translation tables generated by MS Excel. Although \gls{csv} is an abbreviation for \textit{comma-seperated}, one can specify the split symbol. Most common are semicolons. In our case we use semicolons, because commas are frequently used in string representations of mathematical functions to separate multiple parameters and variables. Note that semicolons are also sometimes used, but \gls{csv} files are able to use quotation marks to separate a value with semicolon from the split symbols.

Note that \gls{csv} files are databases. While we do not want to introduce the theory of databases in this thesis, note that our databases contain a \textit{primary key}. A primary key uniquely identifies and provides access to information in the database. The main \gls{csv} file is called \verb|DLMFMacro.csv|. It provides information about each macro. The primary key of this database is the macro itself. Further information consists of the name of the mathematical function, the hyperlink to the \DLMF{} definition, the numbers of parameters, variables and the maximum number of $@$ symbols. The maximum number of $@$ symbols is mostly used for backward translations. Additionally, the \gls{csv} file can contain information about the constraints, branch cuts and a role. The additional role can be used to ignore the macro in the translation process, or to specify that this macro is a mathematical constant. For example, mathematical constants are ignored, because they are organized in the \gls{json} files. This avoid multiple definitions in the database.

Since a translation is not necessarily bijective, the translations are defined in two separate \gls{csv} files per \gls{cas}. For \Maple{} those files are \verb|DLMF_Maple.csv| for the forward translation, and \verb|CAS_Maple.csv| for the backward translation. Both are organized in the same already described logic.

In~\ref{eq:guder-trans} we defined the definition for the Gudermannian function. The translated expression contains two function calls: the inverse tangent function and the hyperbolic sine function. Instead of providing a hyperlink to the definitions for both functions on the \Maple{} help page, we only provide one hyperlink. We specify the hyperlink by a special prefix in the translation pattern. To explain this prefix, we need to take a look to optional parameters before. While the primary key in the \gls{csv} files is only the macro name, we need to specify the number of optional parameters in front of the macro name. Otherwise, for example, the Legendre and the associated Legendre function of the first kind would be linked to the same definition. Our prefix notation solve both problems. It links to one specific function for the translation and specify the correct version of the macro, if there are optional parameters included. The prefix starts and ends with an $X$ and contains two information. The number of optional parameters (if none, than this number will be $0$) and the name of the used function. Table~\ref{tab:legendreP-lex} illustrates the lexicon entry of the Legendre and the associated Legendre function of the first kind.

\begin{table}
	\centering
	\begin{tabular}{ll}
		\hline
		Semantic Macro & Entry in the CSV file\\
		\hline
		\verb|\LegendreP{\nu}@{x}| & \verb|\LegendreP{\nu}@{x}| \\
		\verb|\LegendreP[\mu]{\nu}@{x}| & \verb|X1:\LegendrePX\LegendreP[\mu]{\nu}@{x}|\\
		\hline
	\end{tabular}
	\caption{CSV entry example of the Legendre and associated Legendre function of the first kind.}
	\label{tab:legendreP-lex}
\end{table}

In \Maple{} the number of optional arguments is not stored in the dictionaries. Therefore, we link to the total number of arguments. Consider Olver's associated Legendre function. A forward translation can be defined with 
\begin{equation}
\verb|\LegendreBlackQ{\nu}@{z}| \overset{\langMaple}{\mapsto} \verb|LegendreQ(\nu,z)/GAMMA(\nu+1)|.
\end{equation}
In this case, we would prefer to provide the hyperlink to the definition of \Maple's \verb|LegendreQ| function rather than to the definition of the \verb|GAMMA| function. The used function has two arguments in \Maple. Hence, the lexicon entry for the translation pattern is
\begin{equation}
\verb|X2:LegendreQXLegendreQ($0,$1)/GAMMA($1+1)|.
\end{equation}

All of the \gls{csv} files just store the information to provide an easy access for the developer. Nevertheless, the program itself works only with the \textit{lexicon files}. Therefore, each \gls{csv} file needs to get converted into a lexicon file. Subsection~\ref{subsec:extension} explains how to do that and how to update the database.

\subsubsection{JSON Libraries}
Some translations are defined in separate \gls{json} files. The reason is to separate translations for \gls{opsf} from translation for single symbols and generic \LaTeX{} macros. These files therefore define translations for the following types expressions.

\begin{itemize}
\item \textbf{Generic \LaTeX{} macros for functions}:\newline 
For example, the macros \verb|\frac{a}{b}|, \texttt{\tbs binom\{a\}\{b\}} and \verb|\sqrt{a}|. This section also defines translations for special symbols for mathematical functions such as the exclamation mark '\verb|!|' to indicate the factorial function or \verb|\mod| for the modulo function.
\item \textbf{Mathematical Symbols}:\newline 
For example, the symbol to indicate multiplications, sums, inequalities.
\item \textbf{Greek Letters}:\newline 
For example, \verb|\alpha|, \verb|\beta| or \verb|\Theta|.
\item \textbf{Mathematical Constants}:\newline 
For example \verb|\cpi| for the constant $\cpi$ or \verb|\EulerConstant| for the Euler-Mascheroni constant $\EulerConstant$.
\end{itemize}

\subsubsection{Lexicon Files}
The knowledge of the \gls{mlp} is based on lexicon files, as explained in subsection~\ref{subsec:pom-mlp}. Since the \gls{mlp} project was not a part of the \Macro s project, the \gls{bnf} grammar defines no rules about semantic macros. Furthermore, the knowledge database has no information about the semantic macros. Therefore, we created a new lexicon file that defines each semantic \LaTeX{} macro and provides further information about it. This information is stored in the previously mentioned \gls{csv} files.

Table~\ref{tab:sine-lex-example} presents the entry for the sine function in the \verb|DLMF-macros-lexicon.txt|. The \verb|global-lexicon.txt| file is the main dictionary of the \gls{mlp} and combines the whole knowledge of the parser in one file. This lexicon files also help to translate ambiguous expressions. We will discuss some of these decision in the following sections. However, it is important to mention that a successful translation strongly depends on the comprehensive database of the lexicon files.

\begin{table}
	\centering
	\begin{tabular}{lll}
	\hline
	\multicolumn{3}{l}{Symbol: \texttt{\textbackslash sin}} \\
	\! & \multicolumn{2}{l}{Feature Set: dlmf-macro} \\
	\! & \! & DLMF: \verb|\sin@@{z}|\\
	\! & \! & DLMF-Link: dlmf.nist.gov/4.14\# E1\\
	\! & \! & Meanings: Sine\\
	\! & \! & Number of Parameters: 0\\
	\! & \! & Number of Variables: 1\\
	\! & \! & Number of Ats: 2\\
	\! & \! & Maple: \verb|sin($0)|\\
	\! & \! & Maple-Link: www.maplesoft.com/support/\\
	\! & \! & \hspace{32pt} help/maple/view.aspx?path=sin\\
	\! & \! & help/maple/view.aspx?path=sin\\
	\! & \! & Mathematica: \verb|Sin[$0]|\\
	\! & \! & Mathematica-Link: reference.wolfram.com/\\
	\! & \! & \hspace{32pt} language/ref/Sin\\
	\hline
	\end{tabular}
	\caption{The entry of the sine function in the lexicon file.}
	\label{tab:sine-lex-example}
\end{table}

The concept of the lexicon files is also adapted to store the backward translations from \Maple. Since the \gls{mlp} project is developed to parse \LaTeX{} expressions, it could be confusing to use the original lexicon files to also store names of \Maple{} functions. Therefore, we create another type of lexicon files, which are very similar to the original lexicon files. In \Maple{} a function can have multiple entries differing in the number of arguments. Therefore, the primary key of the backward translation database is the combination of the function name and the number of arguments. For example, the name \verb|GAMMA| in \Maple{} is reserved for the Euler Gamma function in case of one argument, but also for the incomplete Gamma function in the case of two arguments. Therefore the lexicon file for the backward translation also contains the number of arguments to specify a translation.

\subsection{Extending the Forward Translator}\label{subsec:extension}
The whole project is still work in progress. Therefore, it is an important goal to keep it easy to update the backend libraries. This section will explain the typical workflow to add new translations for semantic \LaTeX{} macros and \Maple{} functions or how to support a forward translation to another \gls{cas}.

\subsubsection{Adding Translations to Existing CAS}
Since \gls{csv} files are not as easy to maintain as MS Excel\footnote{Microsoft Excel is widely used to work with tables and is able to import, export and manipulate \gls{csv} files.} worksheets, the workflow starts with MS Excel. Once a user updates the existing tables, the MS Excel file needs to get exported as a \gls{csv} file. The translator project contains the \verb|lexicon-creator.jar| file to convert the \gls{csv} files to the backend lexicon files. Of course, these steps are not necessary if the user just changes a translation in the \gls{json} files.

Once the \verb|lexicon-creator.jar| finished the conversion process, the translator is aware of all changes and ready to use.

\subsubsection{Supporting New CAS}
To support a complete new \gls{cas} for the forward translation, the user needs to add the \gls{cas} in the \gls{json} files and create two new \gls{csv} files. Even if there is no backward translation implemented yet, the \gls{csv} file for the backward translation is needed.

Once all of the files are updated, the \verb|lexicon-creator.jar| can be used to update the lexicon files.

### 2018-05-22 chapters/04-evaluation/evaluation.tex
Init commit with original master thesis

This project aims for a translation of mathematical expressions from \LaTeX{} sources to \gls{cas} and back again. Essentially, the translator translates only string representations. We need to verify that such translations are \textit{appropriate}. Therefore, we develop verification and validation techniques for translations and present them in this chapter.

First of all, each semantic \LaTeX{} macro ties a specific character sequence to a well-defined mathematical object. This definition can be found at the \gls{dlmf} or \gls{drmf}. Therefore, our reference source for semantic \LaTeX{} expressions is always the \gls{dlmf} or \gls{drmf}. This gives us the opportunity to compare more than the string representations of the mathematical objects in a translation process. For example, consider the simple translation of the cosine function
\begin{equation}\label{eq:simple-cosine}
\verb|\cos@{x}| \overset{\langMaple}{\mapsto} \texttt{cos(x)}.
\end{equation}
Our verification and validation techniques, to proof for example that~(\ref{eq:simple-cosine}) is an \textit{appropriate} translation, can be grouped into three parts. We present in the following sections these three parts and conclusively, we use all these techniques to actually verify translations for a test suite directly extracted from the \gls{dlmf}.

\section{Round Trip Tests}\label{sec:round-trip}
Round trip tests are our first approach to verify translations. The idea is to translate an expression back and forth again and check if the representation changes during this process. The translator is able to translate expressions from and to \Maple. Therefore, we implement the following round trip tests only for \Maple.

Furthermore, we take advantage of \Maple's symbolic simplification function to compare a \Maple{} expression in the beginning of an round trip test with the \Maple{} expression after the translations. If \Maple's \textit{simplify} function returns zero for the difference between two representations, both expressions are symbolically equivalent. Note that this relies on the correctness of the simplification, which is not proven (see section~\ref{subsec:mathematica}). On the other hand, the equivalence is not disproven if the simplification returns something different to zero.

A round trip test always starts with a valid expression either in semantic \LaTeX{} or in \Maple. A translation from one system to another is called \textbf{a step}. A complete round trip translation (two steps) is called \textbf{one cycle}.

\begin{definition}[Fixed Point Representation]
A \textbf{fixed point representation} (or short fixed point) in a round trip translation process is a string representation that is identical to all string representations in the following cycles.
\end{definition}

Obviously, there may appear two fixed point representations in round trip tests, one for each system. We will see that most of the round trip translations reach fixed points after a certain amount of steps.

\begin{theorem}[Fixed Point Properties]\label{theo:fixpoint}
Consider two languages $A,B \in \mathcal{L}$. A fixed point representation is reached if the current representation is just identical to the next representation in the same language (after one cycle). Furthermore, if $a \in A$ is a fixed point representation, the translation of $a$ to $b \in B$ is also a fixed point representation.
\end{theorem}
\begin{proof}
Let $A, B \in \mathcal{L}$ be languages. Consider a round trip translation test between the languages $A$ and $B$. Without loss of generality, let $a \in A$ be a fixed point representation and let $a$ be the \textit{first} fixed point representation that appears in this test. The \gls{mlp} as well as \Maple{} are deterministic, because a test case works in only one \Maple{-session} and we do not compute the inputs in \Maple. Therefore, the translation process is not able to call any probabilistic algorithms. Hence, our translator is also deterministic.

Let $b \in B$ with
\begin{equation}
\mathrm{tr}^A_B(a) = b.
\end{equation}

Since the translation process is deterministic, $a$ will be always translated to $b$. Consider $a' \in A$ as the backward translation from $b$
\begin{equation}
\mathrm{tr}^B_A(b) = a'.
\end{equation}
Now $a$ was defined to be a fixed point representation in $A$. Therefore, $a$ is identical to all string representation in the following cycles. Thus $a = a'$.

Furthermore, because the translations are deterministic, $b$ will be always backward translated to $a' = a$ and $a$ will be always forward translated to $b$. Therefore, $b$ is identical to the representation after one cycle. Per definition, $b$ is therefore a fixed point in $B$.

Remember, $a$ was chosen arbitrarily as the first fixed point representation that appears in the round trip test. Therefore, whenever a fixed point appears in a round trip test, the next translation to the respective other language will be a fixed point as well. Consequently, a fixed point is already reached, when just the next representation (after one cycle) is identical to the current representation.
\end{proof}

With theorem~\ref{theo:fixpoint}, we are able to find fixed points in round trip tests by only testing the equality to the next representation (after one cycle). Furthermore, we only need to find the first fixed point representation in either system, because the next step will be a fixed point representation for the respective other system. When a round trip test reaches such a fixed point in \Maple, the \textit{simplify} function tries to simplify the difference between the first expression in \Maple{} and the fixed point expression. It is not feasible to simplify or compute \LaTeX{} expressions yet (neither generic nor semantic \LaTeX) - actually, our translation tool just makes this possible. Therefore, we check the equivalence between a fixed point and its first expression in semantic \LaTeX{} manually.

Table~\ref{tab:fixpoint} illustrates an example of a round trip test which reaches a fixed point. The test formula is
\begin{equation}
\frac{\cos@{a\Theta}}{2}.
\end{equation}

\begin{table}[ht]
\centering
\begin{tabular}{cc}
	\hline 
	Steps & semantic \LaTeX{}/\Maple{} representations\\
	\hline
	\rule{0pt}{0.9\normalbaselineskip}0 & \verb|\frac{\cos@{a\Theta}}{2}|\\
	1 & \verb|(cos(a*Theta))/(2)| \\
	2 & \verb|\frac{1}{2}\idot\cos@{a\idot\Theta}| \\
	3 & \verb|(1)/(2)*cos(a*Theta)|\\
	4 & \verb|\frac{1}{2}\idot\cos@{a\idot\Theta}| \\
	\hline
\end{tabular}
\caption{A round trip test reaching a fixed point.}
\label{tab:fixpoint}
\end{table}

Step 4 is identical to step 2. Theorem~\ref{theo:fixpoint} implies that step 2 is the fixed point representation for semantic \LaTeX{} and step 3 is the fixed point representation for \Maple. The structure of the expression changes between the steps 1 and 2. The reason are internal changes made by \Maple. These problems have been discussed in the previous \cref{subsec:maple-probs}.

Almost all test cases are reaching a fixed point representation. Starting from semantic \LaTeX, the first fixed point is reached at step 2 in semantic \LaTeX, because of \Maple's internal changes. If the structure of the semantic \LaTeX{} input is already identical to the internal structure of \Maple's representation and the expression only contains functions and symbols with an available direct translation from and to each system, the input expression is already the fixed point representation. For example, consider the semantic \LaTeX{} expression in step 2 in table~\ref{tab:fixpoint} as the input expression. In that case the input is already a fixed point. Starting a round trip test from \Maple, the first fixed point representation is reached in step 1 in semantic \LaTeX, because \Maple's changes has already been performed after the first translation. Again, this is only true if the expression only contains functions and symbols with a direct translation from and to each system. Furthermore, if a test case reaches a fixed point, it is equivalent (checked by \Maple's \textit{simplify} function or manually) to the input expression.

There is currently only one exception known where a round trip test does not reach a fixed point representation at all: Legendre's incomplete elliptic integrals~\cite[(19.2.4-7)]{NIST:DLMF} are defined with the amplitude $\phi$ in the first argument in the \DLMF, while \Maple{} takes the trigonometric sine of the amplitude as the first argument. Therefore, the forward and backward translations are defined as
\begin{eqnarray}
\verb|\EllIntF@{\phi}{k}| & \overset{\langMaple}{\mapsto} & \verb|EllipticF(sin(phi),k)|,\\
\verb|\EllIntF@{\asin@{\phi}}{k}| & \overset{\langMaple}{\mapsfrom} & \verb|EllipticF(phi,k)|.
\end{eqnarray}

A round trip translation for this function produces an infinite chain of sine and inverse sine calls. The problem is that the round trip tests prevent simplifications in \Maple{} during the translation process (see \cref{subsec:maple-probs}). This problem could be resolved by allowing automatic evaluations in the round trip tests. However, currently we do not allow those evaluations because the changes can be serious, which makes it hard to follow possible errors.

The round trip tests are very successful, but they only detect errors in string representations. However, because of the simplification techniques of fixed points, we are able to at least detect logical errors in one system: \Maple. On the other hand, these tests cannot determine logical errors in the translations between the two systems - but this is the critical part we want to verify. Consider defining a forward and backward translation for the sine function obviously wrong as
\begin{eqnarray}
\verb|\sin@{\phi}| & \overset{\langMaple}{\leftrightarrow} & \verb|cos(phi)|,\label{eq:wrong-trans-1}\\
\verb|\cos@{\phi}| & \overset{\langMaple}{\leftrightarrow} & \verb|sin(phi)|.\label{eq:wrong-trans-2}
\end{eqnarray}
In that case the round trip test would not detect any errors and reach a fixed point representation, because the simplification techniques only simplify two representations in the same system but cannot compare the representation in one system to those in the other. Therefore, we implemented tests for equations rather than general formulae.

\section{Function Relation Tests}\label{sec:relation-tests}
In addition to the correctness of string representations, it is more important to validate whether a translation is \textit{appropriate}. Special functions are highly related to each other. The theory of special functions researches these relations and other properties of the functions. The \gls{dlmf} is a compendium of special functions and lists a lot of those relations. The translation of such relations to another system and the following analysis of the relations with the translated formulae would form a very effective test.

With this technique we can detect translation errors such as in (\ref{eq:wrong-trans-1} and \ref{eq:wrong-trans-2}). Consider the \gls{dlmf} equation for the sine and cosine function~\cite[(4.21.2)]{NIST:DLMF}
\begin{equation}
\sin@{u+v} = \sin@@{u}\cos@@{v} + \cos@@{u}\sin@@{v}.
\end{equation}
Assume the translator would forward translate the expression based on (\ref{eq:wrong-trans-1},~\ref{eq:wrong-trans-2}). Than
\begin{eqnarray}
\verb|\sin@{u + v}| & \overset{\langMaple}{\mapsto} & \verb|cos(u + v)|,\\
\verb|\sin@@{u}\cos@@{v}| & \overset{\langMaple}{\mapsto} & \verb|cos(u)*sin(v)|,\\
\verb|\cos@@{u}\sin@@{v}| & \overset{\langMaple}{\mapsto} & \verb|sin(u)*cos(v)|.
\end{eqnarray}
This produces the equation in \Maple
\begin{equation}
\cos@{u+v} = \cos@{u}\sin@{v} + \sin@{u}\cos@{v},
\end{equation}
which is wrong. Since the expression is correct before the translation, we conclude an error during the translation process.

However, there are two essential problems with this approach. Testing the mathematical equivalence of expressions is hard to solve and \gls{cas} have trouble to test even simple equations. Furthermore, this approach only checks forward translations because there is no way to check equivalence of expressions in \LaTeX{} automatically (again this could become feasible with our translator). We use \Maple's \textit{simplify} function to check if the difference of the left-hand side and the right-hand side of the equation is equal to zero. In addition, we use \textit{simplify} and check if the division of the right-hand side by the left-hand side returns a numerical value or not. This simplification function is the most powerful function to check the equivalence in \Maple. However, there are several cases where the simplification fails. We implement some tricks to help \textit{simplify}: pre-conversion of the functions of the expression to other special functions is able to help \textit{simplify}, because the algorithms in the background can use different simplification methods for different functions. For example, the user can use the \Maple{} function \textit{convert} to convert the expression
\begin{equation}
\sinh@{x} + \sin@{x}
\end{equation}
to an equivalent representation using the exponential function:
\begin{equation}
\frac{1}{2}\expe^x - \frac{1}{2}\expe^{-x} - \frac{1}{2} \iunit \left( \expe^{\iunit x}-\expe^{-\iunit x} \right).
\end{equation}
With such pre-conversions we are able to improve the simplification process in \Maple. However, the limitations of the \textit{simplify} function are still the weakest part of this verification approach. Consider the equation~(\ref{eq:branch-cut-issue}) from \cref{subsec:branch-cut-issues}. 
\begin{equation}\label{eq:branch-cut-near}
\displaystyle \ParabolicU@{0}{z} = \sqrt{\frac{z}{2\cpi}}\BesselK{\frac{1}{4}}@{\frac{1}{4}z^2}
\end{equation}
As already discussed, the modified Bessel function of the second kind~\cite[(12.7.10)]{NIST:DLMF} has its branch cut in \Maple{} at $z < 0$. However, the argument of $K$ contains $z^2$. For now, just focus on
\begin{equation}\label{eq:besselK}
\BesselK{\frac{1}{4}}@{\frac{1}{4}z^2}
\end{equation}
Because of $z^2$, if $|\ph@{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ the value of (\ref{eq:besselK}) would be no longer on the principal branch. However, \Maple{} will still compute (\ref{eq:besselK}) on the principal branch independently of the value of $z$. Hence, a translation
\begin{equation}
\verb|\BesselK{\frac{1}{4}}@{\frac{1}{4}z^2}| \overset{\langMaple}{\mapsto} \verb|BesselK(1/4,(1/4)*z^2)|
\end{equation}
is incorrect if $|\ph@{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ and one has to use analytic continuation for the right-hand side of equation (\ref{eq:branch-cut-near}). In subsection~\ref{subsec:branch-cut-issues} we also shown that the complete right-hand side of equation~(\ref{eq:branch-cut-near}) has three branch cuts, caused by the square root function and the $z^2$ argument of the modified Bessel function of the second kind. Such complex examples cannot be simplified yet.

However, this validation approach is powerful and runs automatically. We use this approach to verify a large set of test cases extracted directly from the \gls{dlmf}, see \cref{sec:test-summary}.

\section{Numerical Tests}\label{sec:numerical-tests}
Another approach of verification are numerical tests. In principle, we follow the same approach as the relation tests in \cref{sec:relation-tests}, but rather than using the simplification functions we compare the left-hand side and right-hand side for actual values.

For example, reconsider equation~(\ref{eq:branch-cut-near}) from \cite[(12.7.10)]{NIST:DLMF}. Define the difference of right-hand side and left-hand side
\begin{equation}\label{eq:difference}
D(z) := \ParabolicU@{0}{z} - \sqrt{\frac{z}{2\cpi}}\BesselK{\frac{1}{4}}@{\frac{1}{4}z^2},
\end{equation}
and compute $D(z)$ for some values of $z$. Table~\ref{tab:computations-for-difference} presents four computations for $D(z)$, one value for each quadrant in the complex plane.
\begin{table}[ht]
\centering
\begin{tabular}{rcc}
	\hline
	$z\ \ $ & & $D(z)$\\
	\hline
	\tableRowSpace{} $1+\iunit$ & & $2 \cdot 10^{-10} - 2 \cdot 10^{-10} \iunit$\\
	$-1+\iunit$& & $2.222121916 - 1.116719816 \iunit$\\
	$-1-\iunit$& & $2.222121916 + 1.116719816 \iunit$\\
	$1-\iunit$ & & $2 \cdot 10^{-10} + 2 \cdot 10^{-10} \iunit$\\
	\hline
\end{tabular}
\caption{Four computations of $D(z)$ in \Maple.}
\label{tab:computations-for-difference}
\end{table}

The values are computed with the default precision of $10$ significant digits. Caused by the machine accuracy, the first and last values are different from zero. In such cases, we can increase the precision to figure out if it is an accuracy issue or $D(z)$ is really different from zero at those points. Increasing the precision of the computation for the first value $z = 1+\iunit$ up to $100$ significant digits returns $3\cdot 10^{-100} \iunit$ and can therefore be considered as zero (we get the same results for $z = 1-\iunit$). Hence, table~\ref{tab:computations-for-difference} shows us what we already know. The second and third values are computed on the principal branch of the modified Bessel function of the second kind and therefore produces a difference between the left-hand side and right-hand side of equation~(\ref{eq:branch-cut-near}).

This approach of verification is very powerful, because it can verify not only our translations but also equations in a mathematical compendium. This has been demonstrated by discovering an overall sign error of the \gls{dlmf} equation~(\ref{eq:signerror}), see \cref{sec:test-summary}.

However, this approach has also a significant problem. It is easy to see that something went wrong, but very hard to proof equivalence. Obviously, only four tested values are not sufficient to proof the equivalence. How many and which values do we need for a comprehensive result? Reformulating this question to a general problem brings us to: 
\begin{quote}
\textit{How can we discretize equations down to a finite set of test points such that the probability of equivalence for the continues equation is high if the equation is valid for the finite set of test points?}
\end{quote}
As far as we know, there is no answer to this question yet. This is one reason why we do not implement an automatic process of numerical tests yet. Another reason are constraints. Constraints are not as easy to translate as general formulae. They often contain quantifiers or logical operators which are difficult to translate in the current state of the program. However constraints are necessary for numerical tests.

\section{DLMF Test Suite}\label{sec:test-summary}
To verify our translations with the approaches described above we create test suites by extracting equations and formulae from the sources of the \gls{dlmf}. We start with a small test suite of 110\footnote{The set of test cases is available at\newline \url{https://gist.github.com/AndreG-P/f48121d70fbd2a57b2cf6d88019ddadf}} equations and provide a more detailed look to those tests. We are able to translate 108 of those 110 cases.

Two of the test cases (lines 42 and 43) contain macros we are not able to translate yet. Those macros are \verb|\subplus| and \verb|\CFK|, which are used for continued fractions. They have been created for another translation project~\cite[§5]{CICM:Paper} for continued fractions defined in the \textit{Handbook of Continued Fractions for Special Functions} (2008) \cite{CF:Handbook}. The conversion process for the \gls{cfsf} \cite{CF:Maple} dataset from \Maple{} and the \gls{ecf} \cite{CF:eCF,eCF:FinalReport} dataset from Wolfram uses special file types. Therefore, our translator is not able to handle those macros yet.

However, we translate 108 test cases and for 84 of them the simplification function of \Maple{} returns zero as the difference of the left-hand side and right-hand side. The 24 remaining cases are translated, but the simplification function is not able to simplify the difference to zero. Therefore, we try to improve this part. Line 46~\cite[(18.5.10)]{NIST:DLMF} for example, can be simplified to zero, if $n$ is previously set to an arbitrary integer value. For all those 24 test cases, we perform also manual numerical tests, which reveal an overall sign error in equation~\cite[(14.5.14)]{NIST:DLMF}\footnote{The equation had originally been stated as shown in equation~(\ref{eq:signerror}). The \gls{dlmf} has already updated the wrong equation.}
\begin{equation}\label{eq:signerror}
\displaystyle \LegendreP[1/2]{\nu}@{\cosh@@{\xi}} = - \left( \frac{2}{\cpi \sinh@@{\xi}} \right)^{1/2} \cosh@{\left( \nu + \tfrac{1}{2} \right) \xi}.
\end{equation}
Besides this wrong equation, all of these manual numerical tests return zero for the difference between the left-hand side and right-hand side. Of course it is still not possible to conclude the equivalence based on our manual numerical tests, as explained above.

Those 110 test cases are not sufficiently comprehensive for all of the functions of the \gls{dlmf}. Therefore, we create a larger dataset\footnote{We are planning to publish this dataset at \url{http://drmf.wmflabs.org}.} of 4,165 semantic \LaTeX{} formulae, which is extracted from all chapters of the \gls{dlmf}.

We are able to translate 2,232 (approx. 53.59\%) of those test cases. The \textit{simplify} function of \Maple{} returns zero for the difference for 490 of the translated 2,232 test cases. We increase the number to 713 with automatic pre-conversions helping the simplification process. With simplification over the division of the left-hand side by the right-hand side (rather than simplify the difference) we have three more verified translations and 716 in total. Therefore, 1,516 test cases were translated but the translation cannot be verified automatically yet.

The translator cannot translate 1,933 test cases. First of all, 32 lines contain neither an equation nor a relation. Because we have created this dataset for our verification approaches, we do not translate those 32 lines at all. Most of the other failures are produced because of missing semantic information in the \LaTeX{} expression or the formulae contains a macro that cannot be translated in the current state of the program, such as in 978 lines that contain a macro that cannot be translated. For example, in 171 lines one of the symmetric elliptic integrals~\cite[(19.16.1-6)]{NIST:DLMF} appears and we do not provide translation patterns for those integrals, because there is no direct translation possible yet. A workaround would be to allow translations based on the definitions, see~(\ref{eq:gudermannian}).

Other errors appear in 914 cases, where the semantic information is missing or is not unique. An example are 284 lines that use a potentially ambigious single quote (or prime) notation '\aSingleQuote' to identify derivatives of functions.

\cleardoublepage

### 2018-05-22 chapters/05-results/results.tex
Init commit with original master thesis

The goal is to develop an independent, lite translation tool between mathematical \LaTeX{} expressions and \gls{cas}. The semantic \LaTeX{} macros are constantly changing. Therefore, the translator can easily be updated as well, without implementing new code. Furthermore, we have implemented forward translations to the \gls{cas} \Maple{} and \Mathematica{} and a backward translation from \Maple. The translator is therefore able to allow complete round trip translations between \Maple{} and semantic \LaTeX. We have summarized the results of those validation techniques in chapter~\ref{ch:evaluation}.

The translator is based on not published code from the \gls{mlp} and \gls{pom}-tagger. Furthermore, the \Macro s are not published yet as well. Therefore, the current version of the translator is not online available yet. However, we are planning to publish the translator soon on the \gls{drmf} website\footnote{\url{http://drmf.wmflabs.org}}.

Our translator knows 665 \Macro s. We have defined forward translations for 201 of those to \Maple{} and 195 functions for backward translation from \Maple, plus translations for all Greek letters, constants and generic \LaTeX{} macros. The small percentage of covered macros is the translator's most important weakness. The reason is that most of the functions which are represented by a \Macro{} are not defined in \Maple. For example, the multivalued versions of the inverse trigonometric functions~\cite[(4.23.1-6)]{NIST:DLMF} and the $q$-generalizations are not defined in \Maple. A possible workaround is to define alternative translations based on the definition or other representations of the function, see the equations~\ref{eq:acot-alternatives} as an example.

Nonetheless, the translator is able to handle a wide range of input expressions. This is shown by the test case of several formulae directly taken from the \DLMF. Out of 4165 test equations, 716 have been successfully forward translated to \Maple{} and proven to be correct by the internal simplification functions of \Maple. In addition, 1519 have been successfully forward translated to \Maple, but all validation methods failed. Thus, the translator is able to translate approximately $53.59\%$ of all test cases. Numerical test cases can advance the validation process and have been proven to be very powerful, but implementations of such automatic numerical tests are very complex and therefore not realized yet.
\newpage

\section{Bijection of Translations}
One of the main goals has been the provision of bijective mappings between semantic \LaTeX{} and \gls{cas}. This goal has not been achieved, because of the large number of functions that are only defined in one system, but not in the other. Our translations are in general neither injective nor surjective.

The forward translation is not injective, because there exist multiple macros for the same mathematical object. For example, the ultraspherical orthogonal polynomial (or Gegenbauer polynomial) has two macros and both are translated to the same function in the \gls{cas}. One of the macros is defined in the \gls{dlmf}, while the other is defined by the \gls{drmf}. For the background translation we choose the \gls{dlmf} representation.
\begin{eqnarray*}
	\verb|\Ultra{\lambda}{n}@{x}| & \overset{\langMaple}{\mapsto} & \verb|GegenbauerC(n, lambda, x)|,\\
	\verb|\Ultraspherical{\lambda}{n}@{x}| & \overset{\langMaple}{\mapsto} & \verb|GegenbauerC(n, lambda, x)|,\\
	\verb|\Ultraspherical{\lambda}{n}@{x}| & \overset{\langMaple}{\mapsfrom} & \verb|GegenbauerC(n, lambda, x)|.
\end{eqnarray*}

Deleting those multiple definitions from the set of macros would make a forward translation injective. Note that this is only true if two expressions are defined as identical as soon as their semantic meanings are identical. For example, formulae with a different number of white spaces are technically not identical, but will be translated to the same expression. However, those formulae can be defined as identical.

The forward translation is not surjective either, because obviously there are expressions in \gls{cas} which the translator would never produce. The same problems appear for the backward translation. For example, a backward translation from \Maple{} can never produce $\frac{x}{2}$, because of \Maple's internal changes it will only produce $x\frac{1}{2}$. Therefore, a backward translation is not surjective. Furthermore, it is not injective because of functions that cannot be translated yet.

However, for a subset of functions and expressions a bijective mapping is possible and has been proven by round trip tests. The main goal is therefore to enlarge these subsets. Mathematically, we looking for subsets $L_1 \subseteq \langTex$ and $L_2 \subseteq \langMaple$, such that $|L_1|$ and $|L_2|$ are maximal and $\mathrm{tr}^{L_1}_{L_2}$ is bijective and its inverse function is $\mathrm{tr}^{L_2}_{L_1}$.

\section{Supported CAS}
We support a forward translation to \Maple{} and \Mathematica. Nevertheless, the forward translation to \Mathematica{} is in an early state and only support a few translations for functions. Since the main work is already done for \Maple{}, one only needs to update the libraries for a better support of \Mathematica.

In theory it is possible to support even more \gls{cas}, but this has not been tested yet. Especially difficult are \gls{cas} that use different mathematical notations in their input format, such as the prefix notation. Those \gls{cas} cannot be supported as long as the \gls{mlp} does not produce expression trees.

A backward translation has only been implemented for \Maple{} yet and it needs to have a licensed version of \Maple{} installed on the system to work, because access to the internal expression tree of \Maple{} expressions is necessary for the translator. Therefore, possible backward translations from other \gls{cas} also need access to the internal data structure.

\section{Open Problems}
One still unsolved problem are different branch cuts in two systems for the same mathematical function. Our workaround only provides additional information to the user, who needs to handle this issue on his own. However, we have presented an approach to solve this problem. Instead of translating the function itself, the translator could use equivalent formulae for the translation that only contain functions that use the same positions for branch cuts in both systems.

On the other hand, this approach does not solve the problem of the relation~(\ref{eq:branch-cut-near}), between the modified Bessel function of the second kind and the parabolic cylinder function. 
\begin{equation*}
\displaystyle \ParabolicU@{0}{z} = \sqrt{\frac{z}{2\cpi}}\BesselK{\frac{1}{4}}@{\frac{1}{4}z^2}
\end{equation*}
In this case, the problem is not an inconsistent definition of the position of the branch cuts, but that the variable $z$ jumps over branch cuts. While this is generally an issue of \gls{cas}, the translator could solve it automatically by using analytic continuation if necessary. This presumes that the translator can compute semantic \LaTeX{} expressions to detect critical values of $z^2$. Together with a \gls{cas}, this approach is feasible.

Another open problem are the verification techniques for translated expressions. While the backward translations are mostly verified by round trip tests, almost all of the translations need a powerful equivalence checker in the \gls{cas}. The limitations of this functionality are an important open problem for our validations. Furthermore, we currently need to assume the correctness of the used simplification functions. A bug in the simplification process spuriously verifying or mistakenly refusing the translation must be taken into consideration.

An approach to solve this problems are the described numerical tests. However, the verification of equivalence with a finite set of discrete values in a sufficient way, is an open problem.

Besides all of this, the translator still has problems with unclear semantics. One example are the already explained prime symbols to indicate derivatives. Another example are mathematical conventions for notations. For example, an expression
\begin{equation}
\sin^{-1}@{x}
\end{equation}
is mostly interpreted as the arcsine function $\asin@{x}$. However, our translator translates it as
\begin{equation}
\verb|\sin^{-1}@{x}| \overset{\langMaple}{\mapsto} \verb|1/(sin(x))|.
\end{equation}

Another problem are mismatched parenthesis. An expression for an half-opened interval $(a,b]$ currently returns an error because of the mismatch of the parenthesis. Nonetheless, this expression is valid for the representation of an interval.

We have also discovered a bug in the grammatical rules for the square root \LaTeX{} macro. The expression
\begin{equation}\label{eq:mlp-bug}
\verb|\sqrt \frac{a}{b}|
\end{equation}
produces an error. The generic \LaTeX{} macros always taken a fixed number of following expression as arguments. It is possible to group an expression with curly brackets. However, if there are no curly brackets used, a \LaTeX{} compiler interprets just one symbol as the argument. Therefore, the expression '\verb|\frac{a}{b}|' and '\verb|\frac ab|' produces the same output: $\frac ab$. The next symbol can also be a \LaTeX{} macro, such as in expression~(\ref{eq:mlp-bug}). Because the \gls{bnf} rule of the \gls{mlp} defines curly brackets currently as mandatory around the argument of the square root function, the \gls{mlp} produces an error during the parsing process.

### 2018-05-22 chapters/06-future-work-and-conclusion/conclusion.tex
Init commit with original master thesis

The results and conclusions of the translations of the formula set from the \DLMF{} are diverse. Some test cases have shown us that the set of semantic macros is not perfect yet and needs to be improved, seen in the example of the macro definition for the Wronskian symbol. However, simply adding more and more macros to improve the set is not the right decision. In that case we would end up in something similar to \sTeX, which is overloaded and too complicated to use. Obviously, if the system is no longer handy, it will not be used.

On the other hand, the concept of the translator has been proven. For example, one test case has revealed an overall sign error in an equation in the \DLMF~\cite[(14.5.14)]{NIST:DLMF} and therefore the validation system has shown the potential to become a strong verification tool for mathematical compendia. The test cases have also shown how difficult it is to validate a translated expression and have uncovered the problems of translations between two systems with different sets of supported functions. Our validation techniques also assume the correctness of the simplification and computational algorithms in \gls{cas}. However, combining those techniques and automatically running translation checks can potentially not only discover errors in mathematical compendia, but also detect errors in simplifications or computations in \gls{cas}.

Finally we can conclude that the project has been successful and the translator has great potential to become a handy translation tool. Nonetheless, it is without any doubts an ever-growing project and needs to be improved constantly.

### 2018-05-22 chapters/06-future-work-and-conclusion/future-work.tex
Init commit with original master thesis

\section{Approaches for Further Improvements}\label{sec:resolution}
As already discussed, the percentage of defined translations is small. Therefore, extending the backend libraries with additional translations for more semantic macros is the first obvious task to improve the translator. Adding more \gls{cas} would also be worthwhile for forward and backward translations. Some \gls{cas} or other software use other notations than the infix notations. Therefore, another improvement for the translator would be the support for other types of notations. This could be achieved by using expression trees during the translation process rather than the \gls{mlp-pt} generated by the \gls{mlp}.

Another minor issue is the way of representing functions after a backward translation. The \Macro s allow different representations and control the different styles by the number of $@$ symbols. Currently the backward translation simply adds the largest possible number of $@$ symbols, which has undesirable effects. Categorizing the functions (Possible categories may be \textit{trigonometric functions} or \textbf{polynomials}) can be used to define the right typical number of $@$ symbols for the most common representations.

Besides the extensions for the backend libraries, we saw in \cref{sec:numerical-tests} that automatic numerical tests are not implemented yet. Certainly, automatic numerical tests are a powerful approach to verify translations and also to find errors in the mathematical sources. However, the realization of such powerful numerical tests is very difficult. As far as we know, there is no general theory for such methods developed yet. An approach could be to adapt error handling methods in telecommunications such as the cyclic redundancy checks~\cite{CRC} that use checksums to detect errors. Adapting such approaches for general mathematical functions could be helpful.

The most desirable improvement for the translator is the support for generic \LaTeX. Since the semantic \LaTeX{} macros are newly invented and generic \LaTeX{} is still the de facto standard for mathematical expressions, it is definitely worthwhile to support generic \LaTeX{} later on. However, we already explained the difficulties of this task. One approach we want to realize is the \textbf{multiple-scan approach}. This approach is inspired by the approach of humans. When a human reader reads mathematical expressions, he concludes the correct semantics from the structure and the context of the expression. His conclusions highly depend on his own knowledge. If he is not familiar with a mathematical symbol, he has difficulties to understand the whole expression.

The multiple-scan approach tries to adapt these techniques with the following three objectives:
\begin{enumerate}[label=(\arabic*)]
\item\label{multi-scan-1} narrow down possible meanings only from the expression itself, without referring to the context of the expression;
\item\label{multi-scan-2} refine the process with conclusions from the nearby context of the expressions and
\item\label{multi-scan-3} improve the previous process by analyzing not only the nearby context but the overall topic of the whole scientific paper or book, its references and other publications by the authors.
\end{enumerate}
Starting with the lexicon files from the \gls{pom}-tagger, we want to narrow down the possible meanings of an expression. It is planned to extend the \gls{pom}-tagger from \cite{POM-Tagger} to achieve objective~\ref{multi-scan-1}. Furthermore, a large-scale corpus study has shown that around 70 percent of the symbolic elements in scientific papers are denoted in the surrounded text \cite{SymbolDec}. With this awareness we can achieve objective~\ref{multi-scan-2} by just searching the close context \cite{MLP:Project,SIGIR:Semantification,MORITZ:Evaluating}. If the correct semantic information is still unsure, objective~\ref{multi-scan-3} is the last way to find a solution. Online compendia, such as arXiv, can be used to discover the overall topic of a scientific paper, the references and the area of research of the authors. The MCAT search engine developed by Kristianto, Topic and Aizawa \cite{MCAT,MCAT:MathSemanticSearch} is able to extract and score information from the document at this '\textit{document granularity level}'. Ideally, there is only one possible interpretation left after step~\ref{multi-scan-3}. Otherwise, even a human reader might have difficulties to understand the expression.

Figure~\ref{fig:multiple-scan} visualizes the process of conclusions for the Jacobi polynomial example from \cref{tab:JacobiP-usecase}.

\begin{figure}[htp]
	\centering
	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.7]{MultipleScanApproach.pdf}
	\caption{Visualization of the multiple-scan approach. Exemplified using the Jacobi polynomial example.}
	\label{fig:multiple-scan}
\end{figure}

With such improvements the translator possibly becomes multifunctional: from automatic translations between word processors and computer algebra systems over to a verification tool for mathematical compendia, then over to a tool that can automatically enrich mathematical expressions with semantic information. A combination of the automatic semantically enrichment process and \LaTeX{} extraction tools~\cite{MaxTract} would be even hypothetically able to semantically enrich already published articles in PDF format afterwards.

\cleardoublepage

### 2018-05-22 chapters/abstract.tex
Init commit with original master thesis

\cleardoublepage
\begin{abstract}
\setcounter{page}{3}
\glsresetall
\thispagestyle{plain}
This Master's thesis presents the development of a translation tool between representations of semantically enriched mathematical formulae in a \gls{wp} and its corresponding representations in a \gls{cas}. The representatives for the \gls{cas} are \Maple{} and \Mathematica{}, while the representative \gls{wp} is \LaTeX.

Semantic information of a mathematical formula in \LaTeX{} is rather given in the context than in the formula itself. Extracting the information is complicated and sometimes even impossible. However, a translation to a \gls{cas} is only achievable, if the semantic of the formula is unique. The \gls{nist} in the U.S. has developed a set of semantic \LaTeX{} macros for all \glslink{opsf}{orthogonal polynomials and special functions (OPSF)} defined in the \gls{dlmf}. Using these macros can disambiguate formulae and provide access to the semantic information.

Even if the semantics of a representation of one formula is unique, it does not need to match the semantics in another representation of the formula. For example, there are may be differences in the domains or a function in a \gls{cas} is normalized for a better performance. An important concept for complex and multivalued functions are \textit{branch cuts}. However, the positions of these cuts are not consistently defined and can vary from system to system. Therefore, a \gls{cas} user needs to fully understand the properties and definitions of the used functions in the \gls{cas}. Hence, a manual translation from one system to another is not only laborious, but also prone to errors.

This thesis explains the realization of the translator, and discusses and suggests the approaches to the problems mentioned above.
\end{abstract}

\newpage
% back side of title should be empty -> empty page here
\mbox{}
\begin{otherlanguage}{ngerman}
\begin{abstract}
\setcounter{page}{5}
\glsresetall
\thispagestyle{plain}
Im Rahmen der Masterarbeit wurde ein Programm entwickelt, welches semantisch angereicherte mathematische Formeln von einer Darstellung in einem \glslink{wp}{Textverarbeitungssystem (WP)} wie \LaTeX{} zu einer entsprechenden Darstellung in einem \glslink{cas}{Computeralgebrasystem (CAS)} wie \Maple{} oder \Mathematica{} vor- und zurückübersetzen kann.

Semantische Informationen von mathematischen Formeln in \LaTeX{} ergeben sich vor allem aus dem Kontext der Formeln. Diese Informationen zu extrahieren ist jedoch schwierig und in einigen Fällen bisher sogar unmöglich. Eine direkte Übersetzung zu einem \acrshort{cas} ist jedoch nur möglich, wenn die semantischen Informationen eindeutig sind. Das \gls{nist} in den USA hat dafür eine Liste von \LaTeX{} Makros erstellt, welche semantische Informationen für jedes \glslink{opsf}{orthogonale Polynom und jede spezielle Funktion (OPSF)} in der \gls{dlmf} bereitstellen. Mit Hilfe dieser Makros können Doppeldeutigkeiten vermieden und der Zugang zu den semantischen Informationen erleichtert werden.

Selbst wenn die Semantik in einer Darstellung eindeutig ist, muss diese nicht mit der Semantik in einer anderen Darstellung übereinstimmen. Es kann Unterschiede in den Definitionsmengen geben oder eine Normalisierung definiert sein, um eine schnellere Berechnung zu ermöglichen. Bei mehrwertigen, komplexen Funktionen spielen sogenannte \textit{Branch-Cuts} eine wichtige Rolle für \acrshort{cas}. Die Position dieser \textit{Branch-Cuts} folgt zwar allgemeinen Konventionen, diese müssen aber nicht zwangsläufig eingehalten werden. Der Umgang mit einem \acrshort{cas} erfordert daher gute Kenntnisse über die speziellen Eigenschaften und Definitionen. Deshalb ist eine Übersetzung von einer Darstellung (z.B. \LaTeX) zu einer Anderen (z.B. \Maple) per Hand nicht nur mühselig, sondern auch fehleranfällig.

In der vorliegenden Arbeit wird die Umsetzung des Übersetzers erläutert und Lösungsansätze für die oben genannten Probleme werden erörtert.
\end{abstract}
\end{otherlanguage}
\setcounter{page}{6}
\newpage
\mbox{}
\newpage

### 2018-05-22 del.sh

@@ -1,0 +1,15 @@
+rm -f Masterthesis.acn
+rm -f Masterthesis.acr
+rm -f Masterthesis.alg
+rm -f Masterthesis.aux
+rm -f Masterthesis.bbl
+rm -f Masterthesis.bcf
+rm -f Masterthesis.blg
+rm -f Masterthesis.ist
+rm -f Masterthesis.loa
+rm -f Masterthesis.lof
+rm -f Masterthesis.lot
+rm -f Masterthesis.toc
+rm -f Masterthesis.out
+rm -f Masterthesis.run.xml
+rm -f Masterthesis.synctex.gz

### 2018-05-22 del.sh
Update bash for deletions

@@ -1,3 +1,5 @@
+#!/bin/bash
+
 rm -f Masterthesis.acn
 rm -f Masterthesis.acr
 rm -f Masterthesis.alg
@@ -9,6 +11,7 @@ rm -f Masterthesis.ist
 rm -f Masterthesis.loa
 rm -f Masterthesis.lof
 rm -f Masterthesis.lot
+rm -f Masterthesis.log
 rm -f Masterthesis.toc
 rm -f Masterthesis.out
 rm -f Masterthesis.run.xml

### 2018-07-09 chapters/01-introduction/intro.tex
Apply article format and create a clean version

@@
-A typical workflow of a scientist who writes a scientific publication is to use a \gls{wp} to write the publication and one or more \gls{cas} for Verification, Analysis and Visualization, among other tasks.
-his formulae.
-Especially in \gls{stem}, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
-However, since \LaTeX{} (an extension of \TeX) is the standard to write in \gls{stem} and we only focus on mathematical writings in this thesis, we categorize \LaTeX{} as a \gls{wp} as well.} has become the de facto standard for writing scientific publications over the last 30 years~\cites{LATEX:Standard}[559]{DigitalTypo}{Knuth}.
+A typical workflow of a scientist who writes a scientific publication is to use a \gls*{wp} to write the publication and one or more \gls*{cas} for Verification, Analysis and Visualization, among other tasks.
+Especially in \gls*{stem}, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
+However, since \LaTeX{} (an extension of \TeX) is the standard to write in \gls*{stem} and we only focus on mathematical writings in this thesis, we categorize \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the last 30 years~\cites{LATEX:Standard}[559]{DigitalTypo}{Knuth}.
 \LaTeX{} enables printing of mathematical formulae in a structure similar to handwritten style.
@@
 
-While \LaTeX{} focuses on displaying mathematics, \gls{cas} focus on computations and user friendly syntax.
+While \LaTeX{} focuses on displaying mathematics, \gls*{cas} focus on computations and user friendly syntax.
 Therefore, each system (such as \LaTeX{}, \Maple{} or \Mathematica) uses its own representation and syntax.
@@
 Such translations can become complex, if there are non-obvious differences between the used systems.
-For example, the \gls{cas} could use a different order of the arguments, define a normalization of the function for better performances or even use different domains for the variables.
+For example, the \gls*{cas} could use a different order of the arguments, define a normalization of the function for better performances or even use different domains for the variables.
 Multivalued functions are particularly difficult~\cite{AISC:MultivaluedFunctions}.
-A \gls{cas} usually defines so called \textit{branch cuts} to compute principal values of multivalued functions~\cite{Maple:Cuts}.
+A \gls*{cas} usually defines so called \textit{branch cuts} to compute principal values of multivalued functions~\cite{Maple:Cuts}.
 Thereby, implementations of multivalued functions are discontinuous.
@@
 However, since a branch cut is a curve, it also can be defined as a spiral or it can be defined for a different position in the complex plane.
-The position of branch cuts can therefore varies from \gls{cas} to \gls{cas}~\cite{Branches:acot}.
+The position of branch cuts can therefore varies from \gls*{cas} to \gls*{cas}~\cite{Branches:acot}.
 Figure~\ref{fig:acot-cut-compare} illustrates two examples of different branch cut positioning for the inverse trigonometric arccotangent function.
@@
 
-Hence, a \gls{cas} user needs to fully understand the properties and special definitions (such as the position of branch cuts) in the \gls{cas} to avoid mistakes during a translation~\cite{Maple:Cuts}.
+Hence, a \gls*{cas} user needs to fully understand the properties and special definitions (such as the position of branch cuts) in the \gls*{cas} to avoid mistakes during a translation~\cite{Maple:Cuts}.
 In consequence, a manual translation process is not only laborious, but also prone to errors.
-Note that this general problem has been named to automatic \gls{p2c} conversion~\cite{POM-Tagger}.
+Note that this general problem has been named to automatic \gls*{p2c} conversion~\cite{POM-Tagger}.
 
@@
 However, for a computer it is difficult to understand the expression~(\ref{eq:P-tex}).
-Since \LaTeX{} is the de facto standard for mathematical expressions, most \gls{cas} provide import and export functions for \LaTeX{} expressions~\cite{Maple:ImportExport,Mathematica:ImportExport,Matlab:ImportExport,Sage:ImportExport}.
+Since \LaTeX{} is the de facto standard for mathematical expressions, most \gls*{cas} provide import and export functions for \LaTeX{} expressions~\cite{Maple:ImportExport,Mathematica:ImportExport,Matlab:ImportExport,Sage:ImportExport}.
 However, because of the explained difficulties, those functions are very limited.
@@
 
-The \gls{nist} has therefore developed a set of semantic \LaTeX{} macros~\cite{DLMF:Macros}.
-Each of these macros is tied with a unique well-defined mathematical object and linked to the corresponding definition in the \gls{dlmf}~\cite{NIST:DLMF,NIST:DLMF:Paper,NIST:Handbook}.
-The semantic macros are used to semantically enhance the \gls{dlmf}.
-An outgrowth of the \gls{dlmf} project is the \gls{drmf}~\cite{DRMF:14,DRMF:15}, which defines more semantic macros, especially for \gls{opsf}.
+The \gls*{nist} has therefore developed a set of semantic \LaTeX{} macros~\cite{DLMF:Macros}.
+Each of these macros is tied with a unique well-defined mathematical object and linked to the corresponding definition in the \gls*{dlmf}~\cite{NIST:DLMF,NIST:DLMF:Paper,NIST:Handbook}.
+The semantic macros are used to semantically enhance the \gls*{dlmf}.
+An outgrowth of the \gls*{dlmf} project is the \gls*{drmf}~\cite{DRMF:14,DRMF:15}, which defines more semantic macros, especially for \gls*{opsf}.
 In table~\ref{tab:JacobiP-usecase} we have already seen an example of a semantic macro.
 We call \LaTeX{} expressions with semantic macros \textit{semantic \LaTeX{}} and without \textit{generic \LaTeX}, respectively.
-With semantic \LaTeX, it is possible to provide translations to \gls{cas}.
+With semantic \LaTeX, it is possible to provide translations to \gls*{cas}.
 
@@
 
-Furthermore, note that the translator we will present in this thesis is based on not published code from the \gls{mlp} project and the \gls{pom}-tagger.
+Furthermore, note that the translator we will present in this thesis is based on not published code from the \gls*{mlp} project and the \gls*{pom}-tagger.
 We will also use the mentioned set of \Macro s, which are not published yet as well.
 Therefore, the current version of the translator is not online available yet.
-However, we are planning to publish the translator soon on the \gls{drmf} website\footnote{\url{http://drmf.wmflabs.org}}.
+However, we are planning to publish the translator soon on the \gls*{drmf} website\footnote{\url{http://drmf.wmflabs.org}}.
 
@@
 \section{Goals of the project}
-In this thesis, we will present an automatic tool for translations between semantic \LaTeX{} and the \gls{cas} \Maple{} and \Mathematica.
+In this thesis, we will present an automatic tool for translations between semantic \LaTeX{} and the \gls*{cas} \Maple{} and \Mathematica.
 The translation tool should be lite, in a sense that a user does not need a whole software package to use the program.
-It is also a part of the \gls{drmf} project and should provide interactive translations in the \gls{drmf}.
+It is also a part of the \gls*{drmf} project and should provide interactive translations in the \gls*{drmf}.
 Furthermore, it should provide additional information about the translation process and present solutions if an appropriate translation is not possible.
@@
 The chapter presents some approaches, such as round trip tests (section~\ref{sec:round-trip}) and numerical tests (section~\ref{sec:numerical-tests}).
-Extracted formulae from the \gls{dlmf} and \gls{drmf} created a comprehensive test suite for these approaches.
+Extracted formulae from the \gls*{dlmf} and \gls*{drmf} created a comprehensive test suite for these approaches.
 Section~\ref{sec:test-summary} gives a summary of the results.

### 2018-07-09 chapters/abstract.tex
Apply article format and create a clean version

@@
-\cleardoublepage
 \begin{abstract}
-\setcounter{page}{3}
 \glsresetall
-\thispagestyle{plain}
-This Master's thesis presents the development of a translation tool between representations of semantically enriched mathematical formulae in a \gls{wp} and its corresponding representations in a \gls{cas}.
-The representatives for the \gls{cas} are \Maple{} and \Mathematica{}, while the representative \gls{wp} is \LaTeX.
-
+\noindent
+This Master's thesis presents the development of a translation tool between representations of semantically enriched mathematical formulae in a \gls*{wp} and its corresponding representations in a \gls*{cas}.
+The representatives for the \gls*{cas} are \Maple{} and \Mathematica{}, while the representative \gls*{wp} is \LaTeX.
+%
 Semantic information of a mathematical formula in \LaTeX{} is rather given in the context than in the formula itself.
 Extracting the information is complicated and sometimes even impossible.
-However, a translation to a \gls{cas} is only achievable, if the semantic of the formula is unique.
-The \gls{nist} in the U.S.
-has developed a set of semantic \LaTeX{} macros for all \glslink{opsf}{orthogonal polynomials and special functions (OPSF)} defined in the \gls{dlmf}.
+However, a translation to a \gls*{cas} is only achievable, if the semantic of the formula is unique.
+The \gls*{nist} in the U.S.
+has developed a set of semantic \LaTeX{} macros for all \gls*{opsf} defined in the \gls*{dlmf}.
 Using these macros can disambiguate formulae and provide access to the semantic information.
-
+%
 Even if the semantics of a representation of one formula is unique, it does not need to match the semantics in another representation of the formula.
-For example, there are may be differences in the domains or a function in a \gls{cas} is normalized for a better performance.
+For example, there are may be differences in the domains or a function in a \gls*{cas} is normalized for a better performance.
 An important concept for complex and multivalued functions are \textit{branch cuts}.
 However, the positions of these cuts are not consistently defined and can vary from system to system.
-Therefore, a \gls{cas} user needs to fully understand the properties and definitions of the used functions in the \gls{cas}.
+Therefore, a \gls*{cas} user needs to fully understand the properties and definitions of the used functions in the \gls*{cas}.
 Hence, a manual translation from one system to another is not only laborious, but also prone to errors.
-
+%
 This thesis explains the realization of the translator, and discusses and suggests the approaches to the problems mentioned above.
 \end{abstract}
-
-\newpage
-% back side of title should be empty -> empty page here
-\mbox{}
-\begin{otherlanguage}{ngerman}
-\begin{abstract}
-\setcounter{page}{5}
-\glsresetall
-\thispagestyle{plain}
-Im Rahmen der Masterarbeit wurde ein Programm entwickelt, welches semantisch angereicherte mathematische Formeln von einer Darstellung in einem \glslink{wp}{Textverarbeitungssystem (WP)} wie \LaTeX{} zu einer entsprechenden Darstellung in einem \glslink{cas}{Computeralgebrasystem (CAS)} wie \Maple{} oder \Mathematica{} vor- und zurückübersetzen kann.
-
-Semantische Informationen von mathematischen Formeln in \LaTeX{} ergeben sich vor allem aus dem Kontext der Formeln.
-Diese Informationen zu extrahieren ist jedoch schwierig und in einigen Fällen bisher sogar unmöglich.
-Eine direkte Übersetzung zu einem \acrshort{cas} ist jedoch nur möglich, wenn die semantischen Informationen eindeutig sind.
-Das \gls{nist} in den USA hat dafür eine Liste von \LaTeX{} Makros erstellt, welche semantische Informationen für jedes \glslink{opsf}{orthogonale Polynom und jede spezielle Funktion (OPSF)} in der \gls{dlmf} bereitstellen.
-Mit Hilfe dieser Makros können Doppeldeutigkeiten vermieden und der Zugang zu den semantischen Informationen erleichtert werden.
-
-Selbst wenn die Semantik in einer Darstellung eindeutig ist, muss diese nicht mit der Semantik in einer anderen Darstellung übereinstimmen.
-Es kann Unterschiede in den Definitionsmengen geben oder eine Normalisierung definiert sein, um eine schnellere Berechnung zu ermöglichen.
-Bei mehrwertigen, komplexen Funktionen spielen sogenannte \textit{Branch-Cuts} eine wichtige Rolle für \acrshort{cas}.
-Die Position dieser \textit{Branch-Cuts} folgt zwar allgemeinen Konventionen, diese müssen aber nicht zwangsläufig eingehalten werden.
-Der Umgang mit einem \acrshort{cas} erfordert daher gute Kenntnisse über die speziellen Eigenschaften und Definitionen.
-Deshalb ist eine Übersetzung von einer Darstellung (z.B.
-\LaTeX) zu einer Anderen (z.B.
-\Maple) per Hand nicht nur mühselig, sondern auch fehleranfällig.
-
-In der vorliegenden Arbeit wird die Umsetzung des Übersetzers erläutert und Lösungsansätze für die oben genannten Probleme werden erörtert.
-\end{abstract}
-\end{otherlanguage}
-\setcounter{page}{6}
-\newpage
-\mbox{}
-\newpage

### 2018-07-12 chapters/01-introduction/intro.tex
Apply template rules

@@
 Especially in \gls*{stem}, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
-However, since \LaTeX{} (an extension of \TeX) is the standard to write in \gls*{stem} and we only focus on mathematical writings in this thesis, we categorize \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the last 30 years~\cites{LATEX:Standard}[559]{DigitalTypo}{Knuth}.
+However, since \LaTeX{} (an extension of \TeX) is the standard to write in \gls*{stem} and we only focus on mathematical writings in this thesis, we categorize \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the last 30 years~\parencites{LATEX:Standard}[559]{DigitalTypo}{Knuth}.
 \LaTeX{} enables printing of mathematical formulae in a structure similar to handwritten style.
-For example, consider a Jacobi polynomial~\cite[(18.3), table 1]{NIST:DLMF}
+For example, consider a Jacobi polynomial~\parencite[18.3 in table 1]{NIST:DLMF}
 \begin{equation}\label{eq:P}

### 2018-07-12 chapters/abstract.tex
Apply template rules

@@
 \noindent
-This Master's thesis presents the development of a translation tool between representations of semantically enriched mathematical formulae in a \gls*{wp} and its corresponding representations in a \gls*{cas}.
+This paper presents a translation tool between representations of semantically enriched mathematical formulae in a \gls*{wp} and its corresponding representations in a \gls*{cas}.
 The representatives for the \gls*{cas} are \Maple{} and \Mathematica{}, while the representative \gls*{wp} is \LaTeX.
 %
-Semantic information of a mathematical formula in \LaTeX{} is rather given in the context than in the formula itself.
-Extracting the information is complicated and sometimes even impossible.
-However, a translation to a \gls*{cas} is only achievable, if the semantic of the formula is unique.
-The \gls*{nist} in the U.S.
-has developed a set of semantic \LaTeX{} macros for all \gls*{opsf} defined in the \gls*{dlmf}.
-Using these macros can disambiguate formulae and provide access to the semantic information.
+The \gls*{nist} in the U.S.~has developed a set of semantic \LaTeX{} macros for \gls*{opsf}.
+The \gls*{dlmf} use these semantic macros to provide a semantically enhanced mathematical online compendium.
+This kind of semantic macros provides exclusive access to the semantic information of the functions.
+However, even if the semantics of a representation of one formula is unique, it does not need to match the semantics in another representation for the same formula.
+While some distinctions are obvious, such as syntactical characteristics, others are more difficult to examine, such as differences in definitions, which leads to error-prone manual translations.
+This paper presents an automatic translation system and discusses the problems and suggests solutions. 
 %
-Even if the semantics of a representation of one formula is unique, it does not need to match the semantics in another representation of the formula.
-For example, there are may be differences in the domains or a function in a \gls*{cas} is normalized for a better performance.
-An important concept for complex and multivalued functions are \textit{branch cuts}.
-However, the positions of these cuts are not consistently defined and can vary from system to system.
-Therefore, a \gls*{cas} user needs to fully understand the properties and definitions of the used functions in the \gls*{cas}.
-Hence, a manual translation from one system to another is not only laborious, but also prone to errors.
-%
-This thesis explains the realization of the translator, and discusses and suggests the approaches to the problems mentioned above.
+Furthermore, this paper introduces new evaluation approaches for the developed translation tool. 
+With the help of the automatic translation tool, the evaluation experiments were able to discover errors in the \gls*{dlmf} and in the \gls*{cas} Maple.
 \end{abstract}

### 2018-07-17 chapters/01-introduction/intro.tex
General structure (without 'background')

--- 2018-07-09 chapters/01-introduction/intro.tex
@@
-A typical workflow of a scientist who writes a scientific publication is to use a \gls*{wp} to write the publication and one or more \gls*{cas} for Verification, Analysis and Visualization, among other tasks.
-Especially in \gls*{stem}, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
-However, since \LaTeX{} (an extension of \TeX) is the standard to write in \gls*{stem} and we only focus on mathematical writings in this thesis, we categorize \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the last 30 years~\cites{LATEX:Standard}[559]{DigitalTypo}{Knuth}.
+A typical workflow of a scientist who writes a scientific publication is to use a \gls*{wp} to write the paper and one or more \gls*{cas} for verification, analysis and visualization, among other tasks.
+Especially in \gls*{stem} literature, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
+However, since \LaTeX{} (an extension of \TeX) is the standard to write articles in \gls*{stem} and we only focus on mathematical writings in this paper, we categorize \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the past 30 years~\parencites{LATEX:Standard}[559]{DigitalTypo}{Knuth}.
 \LaTeX{} enables printing of mathematical formulae in a structure similar to handwritten style.
-For example, consider a Jacobi polynomial~\cite[(18.3), table 1]{NIST:DLMF}
+For example, consider the Jacobi polynomial~\parencite[18.3 in table 1]{NIST:DLMF}
 \begin{equation}\label{eq:P}
-P_n^{(\alpha , \beta)}(\cos(a\Theta)).
+P_n^{(\alpha , \beta)}(\cos(a\Theta)),
 \end{equation}
+where $\alpha, \beta > -1$, and $n$ a nonnegative integer. 
 This formula is written in \LaTeX{} as
@@
 While \LaTeX{} focuses on displaying mathematics, \gls*{cas} focus on computations and user friendly syntax.
-Therefore, each system (such as \LaTeX{}, \Maple{} or \Mathematica) uses its own representation and syntax.
+Therefore, each system uses different representations and syntax.
 Hence, a writer needs to constantly translate mathematical expressions from one representation to another and back again.
@@
 
-Such translations can become complex, if there are non-obvious differences between the used systems.
-For example, the \gls*{cas} could use a different order of the arguments, define a normalization of the function for better performances or even use different domains for the variables.
-Multivalued functions are particularly difficult~\cite{AISC:MultivaluedFunctions}.
-A \gls*{cas} usually defines so called \textit{branch cuts} to compute principal values of multivalued functions~\cite{Maple:Cuts}.
+Such translations can become complex, if there are non-obvious distinctions between the used systems.
+For example, a \gls*{cas} could use a normalized form of the function or of their arguments for better performances.
+In this translation scenario, multivalued functions are particularly difficult~\parencite{AISC:MultivaluedFunctions}.
+A \gls*{cas} usually defines so called \textit{branch cuts} to compute principal values of multivalued functions~\parencite{Maple:Cuts}.
 Thereby, implementations of multivalued functions are discontinuous.
@@
 They are mostly defined conveniently, such as a straight line.
-However, since a branch cut is a curve, it also can be defined as a spiral or it can be defined for a different position in the complex plane.
-The position of branch cuts can therefore varies from \gls*{cas} to \gls*{cas}~\cite{Branches:acot}.
+Technically, branch cut are curves and therefore can be defined in more complex shapes, such as spirals.
+The position of branch cuts can vary from \gls*{cas} to \gls*{cas}~\parencite{Branches:acot}.
 Figure~\ref{fig:acot-cut-compare} illustrates two examples of different branch cut positioning for the inverse trigonometric arccotangent function.
@@
 
-Hence, a \gls*{cas} user needs to fully understand the properties and special definitions (such as the position of branch cuts) in the \gls*{cas} to avoid mistakes during a translation~\cite{Maple:Cuts}.
+Hence, a \gls*{cas} user needs to fully understand the properties and special definitions (such as the position of branch cuts) in the \gls*{cas} to avoid mistakes during a translation~\parencite{Maple:Cuts}.
 In consequence, a manual translation process is not only laborious, but also prone to errors.
-Note that this general problem has been named to automatic \gls*{p2c} conversion~\cite{POM-Tagger}.
+Note that this general problem has been named to automatic \gls*{p2c} conversion~\parencite{POM-Tagger}.
 
@@
 However, for a computer it is difficult to understand the expression~(\ref{eq:P-tex}).
-Since \LaTeX{} is the de facto standard for mathematical expressions, most \gls*{cas} provide import and export functions for \LaTeX{} expressions~\cite{Maple:ImportExport,Mathematica:ImportExport,Matlab:ImportExport,Sage:ImportExport}.
+Since \LaTeX{} is the de facto standard for mathematical expressions, most \gls*{cas} provide import and export functions for \LaTeX{} expressions\footnote{Maple since version Maple 18, \textit{Produce output suitable for LaTeX 2e}, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;\\
+Mathematica, \textit{Generating and Importing TeX}, \url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html} seen 06/2017;\\
+Matlab, \textit{\LaTeX{} form of symbolic expression in MATLAB}, \url{https://www.mathworks.com/help/symbolic/latex.html} seen 06/2017;\\
+SageMath, \textit{Sage Math}, \url{http://doc.sagemath.org/html/en/tutorial/latex.html} seen 06/2017}.
 However, because of the explained difficulties, those functions are very limited.
@@
 
-The \gls*{nist} has therefore developed a set of semantic \LaTeX{} macros~\cite{DLMF:Macros}.
-Each of these macros is tied with a unique well-defined mathematical object and linked to the corresponding definition in the \gls*{dlmf}~\cite{NIST:DLMF,NIST:DLMF:Paper,NIST:Handbook}.
+The \gls*{nist} has therefore developed a set of semantic \LaTeX{} macros~\parencite{DLMF:Macros}.
+Each of these macros is tied with a unique well-defined mathematical object and linked to the corresponding definition in the \gls*{dlmf}~\parencite{NIST:DLMF,NIST:DLMF:Paper,NIST:Handbook}.
 The semantic macros are used to semantically enhance the \gls*{dlmf}.
-An outgrowth of the \gls*{dlmf} project is the \gls*{drmf}~\cite{DRMF:14,DRMF:15}, which defines more semantic macros, especially for \gls*{opsf}.
+An outgrowth of the \gls*{dlmf} project is the \gls*{drmf}~\parencite{DRMF:14,DRMF:15}, which defines more semantic macros, especially for \gls*{opsf}.
 In table~\ref{tab:JacobiP-usecase} we have already seen an example of a semantic macro.
 We call \LaTeX{} expressions with semantic macros \textit{semantic \LaTeX{}} and without \textit{generic \LaTeX}, respectively.
-With semantic \LaTeX, it is possible to provide translations to \gls*{cas}.
 
-Note that partial results of this Master's thesis have already been published at the 10$^\text{th}$ Conference on Intelligent Computer Mathematics 2017~\cite{CICM:Paper}.
-Techniques, approaches and results have been roughly introduced in the article and are discussed in more detail in this thesis.
-Pictures and ideas that have already been presented there are not explicitly referenced in the following.
+Note that partial results of this paper have been published in~\parencite{CICM:Paper}.
 
-Furthermore, note that the translator we will present in this thesis is based on not published code from the \gls*{mlp} project and the \gls*{pom}-tagger.
+%Furthermore, note that the translator we will present in this thesis is based on not published code from the \gls*{mlp} project and the \gls*{pom}-tagger.
 We will also use the mentioned set of \Macro s, which are not published yet as well.
@@
 
-Moreover, I will use 'we' rather than 'I' in the subsequent chapters of this thesis, since I published and discussed my ideas with others including my advisors Dr.~Moritz Schubotz and Professor Abdou Youssef, the hosts during research visits Howard S.~Cohl and J\"urgen Gerhard, and the students Joon Bang and Kevin Chen.\footnote{Special thanks to M.~Schubotz to give me the permission for using this phrase from his doctoral thesis \cite{MORITZ:Dis}.}
+%Moreover, I will use 'we' rather than 'I' in the subsequent chapters of this thesis, since I published and discussed my ideas with others including my advisors Dr.~Moritz Schubotz and Professor Abdou Youssef, the hosts during research visits Howard S.~Cohl and J\"urgen Gerhard, and the students Joon Bang and Kevin Chen.\footnote{Special thanks to M.~Schubotz to give me the permission for using this phrase from his doctoral thesis \parencite{MORITZ:Dis}.}
 
-\section{Goals of the project}
-In this thesis, we will present an automatic tool for translations between semantic \LaTeX{} and the \gls*{cas} \Maple{} and \Mathematica.
+%\section{Goals of the project}
+%In this thesis, we will present an automatic tool for translations between semantic \LaTeX{} and the \gls*{cas} \Maple{} and \Mathematica.
 The translation tool should be lite, in a sense that a user does not need a whole software package to use the program.
@@
 
-Another goal is the translator's extensibility in order to provide the basis for future work.
+%Another goal is the translator's extensibility in order to provide the basis for future work.
 
-\section{Structure}
-In the following chapters, we will discuss our translator and verification techniques.
+%\section{Structure}
+%In the following chapters, we will discuss our translator and verification techniques.
 Prior to explaining implementation details, we need to introduce some basic background information in chapter~\ref{ch:background}.
@@
 
-The next chapter~\ref{ch:translator} is the main chapter of the thesis.
+%The next chapter~\ref{ch:translator} is the main chapter of the thesis.
 We will explain implementations and discuss problems and solutions of and for the translator.
@@
 
-In chapter~\ref{ch:evaluation} we will discuss, how we can verify or validate a translated expression.
+%In chapter~\ref{ch:evaluation} we will discuss, how we can verify or validate a translated expression.
 The chapter presents some approaches, such as round trip tests (section~\ref{sec:round-trip}) and numerical tests (section~\ref{sec:numerical-tests}).
@@
 
-The results and the current status of the translation tool will be discussed in chapter~\ref{ch:results}.
+%The results and the current status of the translation tool will be discussed in chapter~\ref{ch:results}.
 In conclusion, chapter~\ref{ch:conc-future-work} explains some approaches to further improve the explained techniques in future work.

### 2018-07-17 chapters/02-background/math-background.tex
General structure (without 'background')

@@
 A three-dimensional plot of a complex function, such as $f$ from above, usually maps the real and imaginary part of the variable to the $x$ and $y$ axes, while the height is defined by the absolute value of the complex variable $|z|$.
-Figure~\ref{fig:cylinderU} demonstrates a complex plot of the parabolic cylinder function $\ParabolicU@{a}{z}$~\cite[(12.2i)]{NIST:DLMF} in \Maple.
+Figure~\ref{fig:cylinderU} demonstrates a complex plot of the parabolic cylinder function $U(a,z)$~\cite[(12.2i)]{NIST:DLMF} in \Maple.
 
@@
 	\includegraphics[width=0.8\textwidth]{cylinderUPlot.png}
-	\caption{A 3-dimensional complex plot of the parabolic cylinder function $\ParabolicU@{a}{z}$ at $a=0$ plotted by \Maple.~\cite[(12.2i)]{NIST:DLMF}}
+	\caption{A 3-dimensional complex plot of the parabolic cylinder function $U(a,z)$ at $a=0$ plotted by \Maple.~\cite[(12.2i)]{NIST:DLMF}}
 	\label{fig:cylinderU}
@@
     }
-    \caption{Separately plot of the real~\protect\subref{fig:cylinderURe} and imaginary~\protect\subref{fig:cylinderUIm} part of the parabolic cylinder function $\ParabolicU@{a}{z}$ at $a=0$.}
+    \caption{Separately plot of the real~\protect\subref{fig:cylinderURe} and imaginary~\protect\subref{fig:cylinderUIm} part of the parabolic cylinder function $U(a,z)$ at $a=0$.}
     \label{fig:cylinderUSplit}
@@
 
-A more complicated example might be the equation of the already introduced parabolic cylinder function $\ParabolicU@{a}{z}$ and the modified Bessel function of the second kind $\BesselK{\nu}@{z}$.
+A more complicated example might be the equation of the already introduced parabolic cylinder function $U(a,z)$ and the modified Bessel function of the second kind $\BesselK{\nu}@{z}$.
 The parabolic cylinder function is a solution of a second order differential equation~\cite[(12.2i)]{NIST:DLMF} and can be represented by $\BesselK{\nu}@@{z}$ for $a=0$.
@@
 \begin{equation}\label{eq:branch-cut-issue}
-\displaystyle \ParabolicU@{0}{z} = \sqrt{\frac{z}{2\cpi}}\BesselK{\frac{1}{4}}@{\frac{1}{4}z^2}
+\displaystyle U(0,z) = \sqrt{\frac{z}{2\cpi}}\BesselK{\frac{1}{4}}@{\frac{1}{4}z^2}
 \end{equation}

### 2018-07-17 chapters/02-background/related-work.tex
General structure (without 'background')

@@
 \section{Related Work}\label{sec:related-work}
-Since \LaTeX{} became the de facto standard for writing papers in mathematics, most of the \gls{cas} nowadays are able to import and export mathematical \LaTeX{} expressions~\cite{Maple:ImportExport,Mathematica:ImportExport,Matlab:ImportExport,Sage:ImportExport}.
+Since \LaTeX{} became the de facto standard for writing papers in mathematics, most of the \gls{cas} nowadays are able to import and export mathematical \LaTeX{} expressions.
 Those tools have two essential problems.
@@
 However, the pre-definition approach is simple and easy to realize, as long as the system can carry semantic information.
-A typical, not mathematical example are Wikipedia articles~\cite{Wiki}.
+A typical, not mathematical example are Wikipedia articles~\parencite{Wiki}.
 An author of such an article can specify further information about symbols, words and sentences by creating a hyperlink to a more detailed explanation.
@@
 Similar to this approach are \textit{interactive documents}\footnote{There is no adequate definition what interactive documents are.
-However, this name is widely used to describe electronic document formats that allow interactivity to change the content in real time.}, such as the \gls{cdf}~\cite{CDF:Wolfram} by Wolfram Research or the \textit{worksheets} of \Maple{}, which are also sometimes referred to as interactive documents.
+However, this name is widely used to describe electronic document formats that allow interactivity to change the content in real time.}, such as the \gls{cdf}~\parencite{CDF:Wolfram} by Wolfram Research or the \textit{worksheets} of \Maple{}, which are also sometimes referred to as interactive documents.
 Those documents pre-define the semantics to allow computations.
 Those complex document formats require specialized tools to show and work with the documents (Wolfram CDF Player, or \Maple{} for the \textit{worksheets}).
-The \JOBAD{} architecture~\cite{JOBAD:orig} is able to create web based interactive documents and uses \gls{omdoc}~\cite{OMDoc} to carry semantics.
+The \JOBAD{} architecture~\parencite{JOBAD:orig} is able to create web based interactive documents and uses \gls{omdoc}~\parencite{OMDoc} to carry semantics.
 The documents can be viewed and edited in the browser.
@@
 The translation performs in the background, invisible for the user.
-Similar to the \JOBAD{} architecture other interactive web documents exist, such as \textit{MathDox}~\cite{MathDox} and \textit{The Planetary System}~\cite{Planetary}.
+Similar to the \JOBAD{} architecture other interactive web documents exist, such as \textit{MathDox}~\parencite{MathDox} and \textit{The Planetary System}~\parencite{Planetary}.
 All of them demonstrate the potential for the education system.
@@
 Most of them create their own way to make this possible.
-The web based documents mostly use standardized ways, such as the content \gls{mathML}, a specialization of \gls{mathML}~\cite{MathML}, that focuses on the semantics of an expression rather than any particular rendering for the expression.
+The web based documents mostly use standardized ways, such as the content \gls{mathML}, a specialization of \gls{mathML}~\parencite{MathML}, that focuses on the semantics of an expression rather than any particular rendering for the expression.
 Since \LaTeX{} is the de facto standard to write mathematical expressions in documents there are projects to extend \LaTeX{} in a way that enable to carry the semantic information directly in \LaTeX.
-There are two known projects that created a semantic version of \LaTeX: \sTeX{}~\cite{sTeX} developed by Michael Kohlhase and the \Macro s developed by Bruce Miller.
+There are two known projects that created a semantic version of \LaTeX: \sTeX{}~\parencite{sTeX} developed by Michael Kohlhase and the \Macro s developed by Bruce Miller.
 Our translator uses the \Macro s rather than \sTeX.
@@
 
-Another approach that tries to avoid the translation problem are \LaTeX{} packages that allow computations directly through the \LaTeX{} compiler (for example \textit{LaTeXCalc}~\cite{LatexCalc}) or allow \gls{cas} commands for computations directly in \TeX -files.
-For example, \textit{sagetex}~\cite{Sagetex} is an interface \LaTeX{} package for the open source \gls{cas} \textit{sage}\footnote{An abbreviation for \textit{SageMath}.}.
+Another approach that tries to avoid the translation problem are \LaTeX{} packages that allow computations directly through the \LaTeX{} compiler (for example \textit{LaTeXCalc}~\parencite{LatexCalc}) or allow \gls{cas} commands for computations directly in \TeX -files.
+For example, \textit{sagetex}~\parencite{Sagetex} is an interface \LaTeX{} package for the open source \gls{cas} \textit{sage}\footnote{An abbreviation for \textit{SageMath}.}.
 This package allows \textit{sage} commands in \TeX -files and uses \textit{sage} in the background to compute the commands.

### 2018-07-17 chapters/03-translator/backward-trans.tex
General structure (without 'background')

@@
 
-Note that all explanations in this section is based on \Maple's programming guide~\cite{MAPLE:ProgrammingGuide} for \Maple{} 2016.
+Note that all explanations in this section is based on \Maple's programming guide~\parencite{MAPLE:ProgrammingGuide} for \Maple{} 2016.
 
@@
 \end{tabular}
-\caption{A subset of important internal \Maple{} structures. See~\cite{MAPLE:ProgrammingGuide} for a complete list.}
+\caption{A subset of important internal \Maple{} structures. See~\parencite{MAPLE:ProgrammingGuide} for a complete list.}
 \label{tab:maple-types}
@@
 \subsection{Maple's Open Maple API}
-\Maple{} provides an \gls{api}, called \texttt{OpenMaple}~\cite[\S 14.3]{MAPLE:ProgrammingGuide} for interaction with the \Maple{} kernel.
+\Maple{} provides an \gls{api}, called \texttt{OpenMaple}~\parencite[\S 14.3]{MAPLE:ProgrammingGuide} for interaction with the \Maple{} kernel.
 The \texttt{OpenMaple} \gls{api} is implemented for different programming languages such as \texttt{C}, \texttt{Java} and \texttt{Visual Basic}.
@@
 
-The following figure~\ref{fig:backward-trans} illustrates the backward translation process for the Jacobi polynomial example $\JacobiP{\alpha}{\beta}{n}@{\cos@{a\Theta}}$ from table~\ref{tab:JacobiP-usecase}.
+The following figure~\ref{fig:backward-trans} illustrates the backward translation process for the Jacobi polynomial example $P_n^{(\alpha , \beta)}(\cos(a\Theta))$ from table~\ref{tab:JacobiP-usecase}.
 The input expression is converted to the nested list representation of the \texttt{MyInertForm} (the customized \inertF{} of \Maple).
@@
 	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.7]{MapleTranslation.pdf}
-	\caption{A scheme of the backward translation process from \Maple{} for the Jacobi polynomial $\JacobiP{\alpha}{\beta}{n}@{\cos@{a\Theta}}$.
+	\caption{A scheme of the backward translation process from \Maple{} for the Jacobi polynomial $P_n^{(\alpha , \beta)}(\cos(a\Theta))$.
 The input string is converted by the \Maple{} kernel into the nested list representation.

### 2018-07-17 chapters/03-translator/forward-trans.tex
General structure (without 'background')

@@
 \section{The Forward Translation}\label{sec:forward-translation}
-\glsreset{teo}
+%\glsreset{teo}
 The translation process starts with semantic \LaTeX{} expressions.
@@
 Table~\ref{tab:allTypesTable} on page \pageref{tab:allTypesTable} gives an overview of all different types in a \gls{mlp-pt}.
-A more detailed explanation of the types can be found in~\cite{POM-Tagger}.
+A more detailed explanation of the types can be found in~\parencite{POM-Tagger}.
 
@@
 In that case, someone could wonder why the variable '$energy$' was translated as '$e \cdot n \cdot e \cdot r \cdot g \cdot y$'.
-While in physics or engineering variables possibly appear with longer names, such as the 'energy' example, it is more common to use single symbols for variables in mathematics~\cite{Notation:History}.
+While in physics or engineering variables possibly appear with longer names, such as the 'energy' example, it is more common to use single symbols for variables in mathematics~\parencite{Notation:History}.
 Therefore, we interpret such alphanumerical expressions as multiplications of variables rather than one variable with an alphanumerical name.

### 2018-07-17 chapters/03-translator/gen-app.tex
General structure (without 'background')

@@
-The overall goal of the whole project was to provide an interactive one click-translation for each formula in the \gls{drmf} to a specified \gls{cas}.
+%The overall goal of the whole project was to provide an interactive one click-translation for each formula in the \gls{drmf} to a specified \gls{cas}.
 The program mainly supports translations from and to the \gls{cas} \Maple.
@@
 
-In this chapter, the functionality and the design of the translator will be explained.
+%In this chapter, the functionality and the design of the translator will be explained.
 The structure follows the development process and explains our decisions, problems and solutions.
@@
 
-\section{General Approach and Goals}\label{sec:gen-appr}
-Initially this project aimed for translations not only for semantic \LaTeX{} expressions but for generic \LaTeX{} as well.
+%\section{General Approach and Goals}\label{sec:gen-appr}
+%Initially this project aimed for translations not only for semantic \LaTeX{} expressions but for generic \LaTeX{} as well.
 However, as discussed before, the semantic information in generic \LaTeX{} is hardly accessible (as seen in \ref{subsec:semantic-latex}).
@@
 
-Another strive for a goal was to keep a translation as straight forward as possible.
+%Another strive for a goal was to keep a translation as straight forward as possible.
 In other words, the structure of the translated expression should be similar to the structure of the input expression.
@@
 
-Furthermore, the translation process should solve problems of existing technologies, presented in section~\ref{sec:related-work}.
+%Furthermore, the translation process should solve problems of existing technologies, presented in section~\ref{sec:related-work}.
 In the following subsection, we will focus on known problems and present solutions before we start to explain implementation details.
 
-\subsection{Translation Problems}
+\section{Translation Problems}
 Most of the existing translation tools cannot preserve the semantic information during the translation process (see section~\ref{sec:related-work}).
@@
 Our workaround is to translate not the mathematical object itself, but the definition of the object.
-For example, the \textit{Gudermannian}~\cite[(4.23.10)]{NIST:DLMF} $\Gudermannian@{x}$ function can be defined by
+For example, the \textit{Gudermannian}~\parencite[(4.23.10)]{NIST:DLMF} $\Gudermannian{x}$ function can be defined by
 \begin{equation}\label{eq:gudermannian}
-\Gudermannian@{x} := \atan@{\sinh@{x}} \quad x \in \Real.
+\Gudermannian{x} := \atan{\sinh{x}} \quad x \in \Real.
 \end{equation}
@@
 \begin{equation}\label{eq:guder-trans}
-\verb|\Gudermannian@{x}| \overset{\langMaple}{\mapsto} \texttt{arctan(sinh(x))}.
+\verb|\Gudermannian{x}| \overset{\langMaple}{\mapsto} \texttt{arctan(sinh(x))}.
 \end{equation}
@@
 \begin{equation}
-\verb|\acot@{x}| \overset{\langMaple}{\mapsto} \texttt{arccot(x)}.
+\verb|\acot{x}| \overset{\langMaple}{\mapsto} \texttt{arccot(x)}.
 \end{equation}
@@
 This can cause trouble in the replacement process.
-However, we can handle all of the few cases by adding addition parenthesis.
+However, we can handle all of the few cases by adding additional parenthesis.
 Consider the differentiation function in \Maple
@@
 \begin{equation}
-\deriv[2]{x^2}{x} = \verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [x$2])|.
+\frac{d^2x^2}{dx^2} = \verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [x$2])|.
 \end{equation}

### 2018-07-17 chapters/04-evaluation/evaluation.tex
General structure (without 'background')

@@
 \begin{equation}\label{eq:simple-cosine}
-\verb|\cos@{x}| \overset{\langMaple}{\mapsto} \texttt{cos(x)}.
+\verb|\cos{x}| \overset{\langMaple}{\mapsto} \texttt{cos(x)}.
 \end{equation}
@@
 
-There is currently only one exception known where a round trip test does not reach a fixed point representation at all: Legendre's incomplete elliptic integrals~\cite[(19.2.4-7)]{NIST:DLMF} are defined with the amplitude $\phi$ in the first argument in the \DLMF, while \Maple{} takes the trigonometric sine of the amplitude as the first argument.
+There is currently only one exception known where a round trip test does not reach a fixed point representation at all: Legendre's incomplete elliptic integrals~\parencite[(19.2.4-7)]{NIST:DLMF} are defined with the amplitude $\phi$ in the first argument in the \DLMF, while \Maple{} takes the trigonometric sine of the amplitude as the first argument.
 Therefore, the forward and backward translations are defined as
@@
 With this technique we can detect translation errors such as in (\ref{eq:wrong-trans-1} and \ref{eq:wrong-trans-2}).
-Consider the \gls{dlmf} equation for the sine and cosine function~\cite[(4.21.2)]{NIST:DLMF}
+Consider the \gls{dlmf} equation for the sine and cosine function~\parencite[(4.21.2)]{NIST:DLMF}
 \begin{equation}
-\sin@{u+v} = \sin@@{u}\cos@@{v} + \cos@@{u}\sin@@{v}.
+\sin \left(u+v\right) = \sin{u}\cos{v} + \cos{u}\sin{v}.
 \end{equation}
@@
 \begin{equation}
-\cos@{u+v} = \cos@{u}\sin@{v} + \sin@{u}\cos@{v},
+\cos\left(u+v\right) = \cos{u}\sin{v} + \sin{u}\cos{v},
 \end{equation}
@@
 \begin{equation}
-\sinh@{x} + \sin@{x}
+\sinh{x} + \sin{x}
 \end{equation}
@@
 \begin{equation}\label{eq:branch-cut-near}
-\displaystyle \ParabolicU@{0}{z} = \sqrt{\frac{z}{2\cpi}}\BesselK{\frac{1}{4}}@{\frac{1}{4}z^2}
+\displaystyle U(0,z) = \sqrt{\frac{z}{2\cpi}} K_{\frac{1}{4}}\left(\frac{1}{4}z^2\right)
 \end{equation}
-As already discussed, the modified Bessel function of the second kind~\cite[(12.7.10)]{NIST:DLMF} has its branch cut in \Maple{} at $z < 0$.
+As already discussed, the modified Bessel function of the second kind~\parencite[(12.7.10)]{NIST:DLMF} has its branch cut in \Maple{} at $z < 0$.
 However, the argument of $K$ contains $z^2$.
@@
 \begin{equation}\label{eq:besselK}
-\BesselK{\frac{1}{4}}@{\frac{1}{4}z^2}
+K_{\frac{1}{4}}\left(\frac{1}{4}z^2\right)
 \end{equation}
-Because of $z^2$, if $|\ph@{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ the value of (\ref{eq:besselK}) would be no longer on the principal branch.
+Because of $z^2$, if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ the value of (\ref{eq:besselK}) would be no longer on the principal branch.
 However, \Maple{} will still compute (\ref{eq:besselK}) on the principal branch independently of the value of $z$.
@@
 \end{equation}
-is incorrect if $|\ph@{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ and one has to use analytic continuation for the right-hand side of equation (\ref{eq:branch-cut-near}).
+is incorrect if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ and one has to use analytic continuation for the right-hand side of equation (\ref{eq:branch-cut-near}).
 In subsection~\ref{subsec:branch-cut-issues} we also shown that the complete right-hand side of equation~(\ref{eq:branch-cut-near}) has three branch cuts, caused by the square root function and the $z^2$ argument of the modified Bessel function of the second kind.
@@
 
-For example, reconsider equation~(\ref{eq:branch-cut-near}) from \cite[(12.7.10)]{NIST:DLMF}. Define the difference of right-hand side and left-hand side
+For example, reconsider equation~(\ref{eq:branch-cut-near}) from \parencite[(12.7.10)]{NIST:DLMF}. Define the difference of right-hand side and left-hand side
 \begin{equation}\label{eq:difference}
-D(z) := \ParabolicU@{0}{z} - \sqrt{\frac{z}{2\cpi}}\BesselK{\frac{1}{4}}@{\frac{1}{4}z^2},
+D(z) := U(0,z) - \sqrt{\frac{z}{2\cpi}} K_{\frac{1}{4}}\left(\frac{1}{4}z^2\right),
 \end{equation}
@@
 Those macros are \verb|\subplus| and \verb|\CFK|, which are used for continued fractions.
-They have been created for another translation project~\cite[§5]{CICM:Paper} for continued fractions defined in the \textit{Handbook of Continued Fractions for Special Functions} (2008) \cite{CF:Handbook}.
-The conversion process for the \gls{cfsf} \cite{CF:Maple} dataset from \Maple{} and the \gls{ecf} \cite{CF:eCF,eCF:FinalReport} dataset from Wolfram uses special file types.
+They have been created for another translation project~\parencite[§5]{CICM:Paper} for continued fractions defined in the \textit{Handbook of Continued Fractions for Special Functions} (2008) \parencite{CF:Handbook}.
+The conversion process for the \gls{cfsf} \parencite{CF:Maple} dataset from \Maple{} and the \gls{ecf} \parencite{CF:eCF,eCF:FinalReport} dataset from Wolfram uses special file types.
 Therefore, our translator is not able to handle those macros yet.
@@
 Therefore, we try to improve this part.
-Line 46~\cite[(18.5.10)]{NIST:DLMF} for example, can be simplified to zero, if $n$ is previously set to an arbitrary integer value.
-For all those 24 test cases, we perform also manual numerical tests, which reveal an overall sign error in equation~\cite[(14.5.14)]{NIST:DLMF}\footnote{The equation had originally been stated as shown in equation~(\ref{eq:signerror}).
+Line 46~\parencite[(18.5.10)]{NIST:DLMF} for example, can be simplified to zero, if $n$ is previously set to an arbitrary integer value.
+For all those 24 test cases, we perform also manual numerical tests, which reveal an overall sign error in equation~\parencite[(14.5.14)]{NIST:DLMF}\footnote{The equation had originally been stated as shown in equation~(\ref{eq:signerror}).
 The \gls{dlmf} has already updated the wrong equation.}
 \begin{equation}\label{eq:signerror}
-\displaystyle \LegendreP[1/2]{\nu}@{\cosh@@{\xi}} = - \left( \frac{2}{\cpi \sinh@@{\xi}} \right)^{1/2} \cosh@{\left( \nu + \tfrac{1}{2} \right) \xi}.
+\displaystyle P^{1/2}_{\nu}\left(\cosh{\xi}\right) = - \left( \frac{2}{\cpi \sinh{\xi}} \right)^{1/2} \cosh{\left( \nu + \tfrac{1}{2} \right) \xi}.
 \end{equation}
@@
 Most of the other failures are produced because of missing semantic information in the \LaTeX{} expression or the formulae contains a macro that cannot be translated in the current state of the program, such as in 978 lines that contain a macro that cannot be translated.
-For example, in 171 lines one of the symmetric elliptic integrals~\cite[(19.16.1-6)]{NIST:DLMF} appears and we do not provide translation patterns for those integrals, because there is no direct translation possible yet.
+For example, in 171 lines one of the symmetric elliptic integrals~\parencite[(19.16.1-6)]{NIST:DLMF} appears and we do not provide translation patterns for those integrals, because there is no direct translation possible yet.
 A workaround would be to allow translations based on the definitions, see~(\ref{eq:gudermannian}).

### 2018-07-20 chapters/01-introduction/intro.tex
Update introduction with more important information

@@
-A typical workflow of a scientist who writes a scientific publication is to use a \gls*{wp} to write the paper and one or more \gls*{cas} for verification, analysis and visualization, among other tasks.
+A typical workflow of a scientist who writes a scientific publication is to use a \gls*{wp} to write the paper and one or more \gls*{cas} for verification, analysis and visualization.
 Especially in \gls*{stem} literature, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
@@
 This formula is written in \LaTeX{} as
-\begin{equation}\label{eq:P-tex}
+\begin{equation*}
 \verb|P_n^{(\alpha,\beta)}(\cos(a\Theta))|.
-\end{equation}
+\end{equation*}
 
-While \LaTeX{} focuses on displaying mathematics, \gls*{cas} focus on computations and user friendly syntax.
-Therefore, each system uses different representations and syntax.
+While \LaTeX{} focuses on displaying mathematics, a \gls*{cas} concentrates on computations and user friendly syntax.
+Especially important for a CAS is to embed unambiguous semantic information within the input.
+Each system uses different representations and syntax in consequence.
 Hence, a writer needs to constantly translate mathematical expressions from one representation to another and back again.
@@
 	\end{tabular}
-	\caption{Different representations for the same mathematical formula~(\ref{eq:P}).}
+	\caption{Different representations for the same mathematical formula~(\ref{eq:P}).
+Generic \LaTeX{} is the default \LaTeX{} expression.
+Semantic \LaTeX{} uses special semantic macros to embed semantic information.}
 	\label{tab:JacobiP-usecase}
@@
 
-Such translations can become complex, if there are non-obvious distinctions between the used systems.
-For example, a \gls*{cas} could use a normalized form of the function or of their arguments for better performances.
-In this translation scenario, multivalued functions are particularly difficult~\parencite{AISC:MultivaluedFunctions}.
-A \gls*{cas} usually defines so called \textit{branch cuts} to compute principal values of multivalued functions~\parencite{Maple:Cuts}.
-Thereby, implementations of multivalued functions are discontinuous.
-In general, positioning branch cuts follow some conventions.
-They are mostly defined conveniently, such as a straight line.
-Technically, branch cut are curves and therefore can be defined in more complex shapes, such as spirals.
-The position of branch cuts can vary from \gls*{cas} to \gls*{cas}~\parencite{Branches:acot}.
+Translations from generic \LaTeX{} to \gls{cas} are difficult to realize since the semantic information are not explicitely given in the input.
+The \gls{nist} has created a set of semantic \LaTeX{} macros~\cite{DLMF:Macros}.
+Each macro ties specific character sequences to a well-defined mathematical object and is linked with the corresponding definition in the \gls{dlmf} or \gls{drmf}.
+These macros embeds necessary semantic information into \LaTeX{} expressions.
+One example of such a macro is given table~\ref{tab:JacobiP-usecase} for the semantic \LaTeX{} representation for the Jacobi polynomial.
+The macros provide isolated access to important parts of the mathematical function, such as the arguments.
+
+Even with embedded semantic information, a translation between the systems can be difficult.
+A typical example of complex problems occurs for multivalued functions~\parencite{AISC:MultivaluedFunctions}.
+A \gls*{cas} usually defines \textit{branch cuts} to compute principal values of multivalued functions~\parencite{Maple:Cuts}, which makes the implementation of a theoretically continues function to a discontinued presentation of it.
+In general, positioning branch cuts follows conventions, but can be positioned arbitrarily in many cases.
+Communicate and explain the decision of defined branch cuts is a critical point for \gls{cas} and can vary between the systems~\parencite{Branches:acot}.
 Figure~\ref{fig:acot-cut-compare} illustrates two examples of different branch cut positioning for the inverse trigonometric arccotangent function.
@@
 
-An automatic translation process is desirable.
-Providing translations for mathematical \LaTeX{} expressions is difficult, because the semantic information is absent.
-Semantics is the meaning of an expression.
-Consider the example of the Jacobi polynomial in~(\ref{eq:P}).
-A scientific reader might be able to conclude information like $P$ indicating the Jacobi polynomial and $n$ is being a non-negative integer.
-However, for a computer it is difficult to understand the expression~(\ref{eq:P-tex}).
-Since \LaTeX{} is the de facto standard for mathematical expressions, most \gls*{cas} provide import and export functions for \LaTeX{} expressions\footnote{Maple since version Maple 18, \textit{Produce output suitable for LaTeX 2e}, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;\\
-Mathematica, \textit{Generating and Importing TeX}, \url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html} seen 06/2017;\\
-Matlab, \textit{\LaTeX{} form of symbolic expression in MATLAB}, \url{https://www.mathworks.com/help/symbolic/latex.html} seen 06/2017;\\
-SageMath, \textit{Sage Math}, \url{http://doc.sagemath.org/html/en/tutorial/latex.html} seen 06/2017}.
-However, because of the explained difficulties, those functions are very limited.
-In fact, they only work for expressions with unique semantic information, but fail for more complex expressions, such as our Jacobi polynomial example.
-
-The \gls*{nist} has therefore developed a set of semantic \LaTeX{} macros~\parencite{DLMF:Macros}.
-Each of these macros is tied with a unique well-defined mathematical object and linked to the corresponding definition in the \gls*{dlmf}~\parencite{NIST:DLMF,NIST:DLMF:Paper,NIST:Handbook}.
-The semantic macros are used to semantically enhance the \gls*{dlmf}.
-An outgrowth of the \gls*{dlmf} project is the \gls*{drmf}~\parencite{DRMF:14,DRMF:15}, which defines more semantic macros, especially for \gls*{opsf}.
-In table~\ref{tab:JacobiP-usecase} we have already seen an example of a semantic macro.
-We call \LaTeX{} expressions with semantic macros \textit{semantic \LaTeX{}} and without \textit{generic \LaTeX}, respectively.
+This article presents a new approach for automatic \gls{p2c} and vice versa conversions.
+Translations from presentational to computational (computational to presentational) systems are called forward (backward) translations.
+A forward translation is denoted with an arrow with the target system language above the arrow.
+For example,
+\begin{equation*}
+t \overset{\langMaple}{\mapsto} c,
+\end{equation*}
+where $t$ is an expression in the \LaTeX{} language and $c$ is an element of the \Maple{} language $\langMaple$.
+As we will see later in this article, we need to compare mathematical concepts between systems.
+This is impossible from a mathematical point of view.
+Consider the irrational mathematical constant $e$, known as Euler's number.
+The theoretical construct for this symbol cannot be mathematically equivalent to the value \verb|exp(0)| in \Maple, caused by computational and implementational limitations.
+Instead of using the term \textit{equivalent}, we introduce a \textit{appropriate} and \textit{inappropriate} translations.
+We call a translation such as
+\begin{equation}\label{eq:cos-def}
+\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|cos(z)|
+\end{equation}
+as \textit{appropriate}, while a translation such as
+\begin{equation}
+\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|sin(z)|
+\end{equation}
+is called \textit{inappropriate}.
+Note that it is not always as easy as in this example to decide if a translation is appropriate or not.
+Therefore, this article also presents several validation techniques to automatically verify if a translation is appropriate or inappropriate.
+In addition to this terminology, we introduce \textit{direct translations}.
+Later in the paper, we will explain that a translation from one specific mathematical object to its \textit{appropriate} counterpart in the other system is not always possible.
+We call a translation to the \textit{appropriate} counterpart \textit{direct}.
+For example, the translation~(\ref{eq:cos-def}) is \textit{direct}, while a translation to the definition of the cosine function
+\begin{equation*}
+\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|(exp(I*z)+exp(-I*z))/2|
+\end{equation*}
+is not a \textit{direct} translation.
 
 Note that partial results of this paper have been published in~\parencite{CICM:Paper}.
-
-%Furthermore, note that the translator we will present in this thesis is based on not published code from the \gls*{mlp} project and the \gls*{pom}-tagger.
-We will also use the mentioned set of \Macro s, which are not published yet as well.
-Therefore, the current version of the translator is not online available yet.
-However, we are planning to publish the translator soon on the \gls*{drmf} website\footnote{\url{http://drmf.wmflabs.org}}.
-
-%Moreover, I will use 'we' rather than 'I' in the subsequent chapters of this thesis, since I published and discussed my ideas with others including my advisors Dr.~Moritz Schubotz and Professor Abdou Youssef, the hosts during research visits Howard S.~Cohl and J\"urgen Gerhard, and the students Joon Bang and Kevin Chen.\footnote{Special thanks to M.~Schubotz to give me the permission for using this phrase from his doctoral thesis \parencite{MORITZ:Dis}.}
-
-%\section{Goals of the project}
-%In this thesis, we will present an automatic tool for translations between semantic \LaTeX{} and the \gls*{cas} \Maple{} and \Mathematica.
-The translation tool should be lite, in a sense that a user does not need a whole software package to use the program.
-It is also a part of the \gls*{drmf} project and should provide interactive translations in the \gls*{drmf}.
-Furthermore, it should provide additional information about the translation process and present solutions if an appropriate translation is not possible.
-
-%Another goal is the translator's extensibility in order to provide the basis for future work.
-
-%\section{Structure}
-%In the following chapters, we will discuss our translator and verification techniques.
-Prior to explaining implementation details, we need to introduce some basic background information in chapter~\ref{ch:background}.
-We start this chapter with an overview of related work (section~\ref{sec:related-work}).
-Multivalued functions and branch cuts, which were already mentioned, will be introduced in section~\ref{sec:math-background}.
-Besides the mathematical background, we will discuss semantics in \LaTeX{} in section~\ref{sec:semantics}.
-A brief introduction to grammar in languages and how we will use it to understand mathematical expressions will be given in section~\ref{sec:mlp}.
-The chapter is closed by some definitions for the following chapters.
-
-%The next chapter~\ref{ch:translator} is the main chapter of the thesis.
-We will explain implementations and discuss problems and solutions of and for the translator.
-It starts with an outline of the goals and our approaches in section~\ref{sec:gen-appr}.
-While we will present implementation details for the forward translations in section~\ref{sec:forward-translation}, we will present the developed backward translation in the following section~\ref{sec:backward-translation}.
-
-%In chapter~\ref{ch:evaluation} we will discuss, how we can verify or validate a translated expression.
-The chapter presents some approaches, such as round trip tests (section~\ref{sec:round-trip}) and numerical tests (section~\ref{sec:numerical-tests}).
-Extracted formulae from the \gls*{dlmf} and \gls*{drmf} created a comprehensive test suite for these approaches.
-Section~\ref{sec:test-summary} gives a summary of the results.
-
-%The results and the current status of the translation tool will be discussed in chapter~\ref{ch:results}.
-In conclusion, chapter~\ref{ch:conc-future-work} explains some approaches to further improve the explained techniques in future work.

### 2018-07-20 chapters/02-background/math-background.tex
Update introduction with more important information

@@
-\section{Mathematical Background}\label{sec:math-background}
-To understand the following chapters in the thesis, we need a brief introduction in complex analysis and multivalued functions.
-The following mathematical introduction is based on the book \textit{Introduction to Complex Analysis} by H.~A.~Priestley~\cite{ComplexAna:Intro}.
+%\section{Mathematical Background}\label{sec:math-background}
+%To understand the following chapters in the thesis, we need a brief introduction in complex analysis and multivalued functions.
+The following mathematical introduction is based on the book \textit{Introduction to Complex Analysis} by H.~A.~Priestley~\parencite{ComplexAna:Intro}.
 
-A complex number $z \in \Complex$ is specified by a two real numbers $x$ and $y$ as $z = x+y\iunit$, with $\iunit$ as the imaginary unit, which is defined as a solution of $\iunit^2 = -1$.
+%A complex number $z \in \Complex$ is specified by a two real numbers $x$ and $y$ as $z = x+y\iunit$, with $\iunit$ as the imaginary unit, which is defined as a solution of $\iunit^2 = -1$.
 The two real numbers are called the real part $x = \Re(z)$ and the imaginary part $y = \Im(z)$ of $z$.
@@
 
-\begin{figure}[ht]
-    \centering
-    \subfloat[Four color phase mapping.]{%
-        \includegraphics[width=0.28\textwidth]{dlmfColorMap.png}
-        \label{fig:four-color}%
-    }
-    \hspace{0.4cm}
-    \subfloat[Continues phase mapping in \DLMF.]{%
-        \includegraphics[width=0.28\textwidth]{dlmfColorMapCont.png}
-        \label{fig:cont-color}%
-    }
-    \hspace{0.4cm}
-    \subfloat[Continues phase mapping in \Maple.]{%
-        \includegraphics[width=0.28\textwidth]{mapleColorMap.png}
-        \label{fig:maple-color}%
-    }
-    \caption{Three different domain coloring methods.
-Mappings~\protect\subref{fig:four-color} and \protect\subref{fig:cont-color} used in the \DLMF{} (Source: \DLMF{~}\cite{NIST:DLMF} in help page, \textit{About Color Map}).
+%\begin{figure}[ht]
+%    \centering
+%    \subfloat[Four color phase mapping.]{%
+%        \includegraphics[width=0.28\textwidth]{dlmfColorMap.png}
+%        \label{fig:four-color}%
+%    }
+%    \hspace{0.4cm}
+%    \subfloat[Continues phase mapping in \DLMF.]{%
+%        \includegraphics[width=0.28\textwidth]{dlmfColorMapCont.png}
+%        \label{fig:cont-color}%
+%    }
+%    \hspace{0.4cm}
+%    \subfloat[Continues phase mapping in \Maple.]{%
+%        \includegraphics[width=0.28\textwidth]{mapleColorMap.png}
+%        \label{fig:maple-color}%
+%    }
+%    \caption{Three different domain coloring methods.
+Mappings~\protect\subref{fig:four-color} and \protect\subref{fig:cont-color} used in the \DLMF{} (Source: \DLMF{~}\parencite{NIST:DLMF} in help page, \textit{About Color Map}).
 Mapping~\protect\subref{fig:maple-color} is the default coloring map in \Maple.}
-    \label{fig:domain-coloring}
-\end{figure}
+%    \label{fig:domain-coloring}
+%\end{figure}
 
-Since the same basic arithmetical rules apply in $\Complex$ as in $\Real$, we will not introduce all arithmetical rules for complex numbers here.
+%Since the same basic arithmetical rules apply in $\Complex$ as in $\Real$, we will not introduce all arithmetical rules for complex numbers here.
 However, another interesting part are plots of complex functions.
@@
 
-This thesis mostly focus on the \gls{cas} \Maple, therefore all of the following complex plots use the default color map in \Maple{~}\ref{fig:maple-color}.
+%This thesis mostly focus on the \gls{cas} \Maple, therefore all of the following complex plots use the default color map in \Maple{~}\ref{fig:maple-color}.
 
-A three-dimensional plot of a complex function, such as $f$ from above, usually maps the real and imaginary part of the variable to the $x$ and $y$ axes, while the height is defined by the absolute value of the complex variable $|z|$.
-Figure~\ref{fig:cylinderU} demonstrates a complex plot of the parabolic cylinder function $U(a,z)$~\cite[(12.2i)]{NIST:DLMF} in \Maple.
+%A three-dimensional plot of a complex function, such as $f$ from above, usually maps the real and imaginary part of the variable to the $x$ and $y$ axes, while the height is defined by the absolute value of the complex variable $|z|$.
+Figure~\ref{fig:cylinderU} demonstrates a complex plot of the parabolic cylinder function $U(a,z)$~\parencite[(12.2i)]{NIST:DLMF} in \Maple.
 
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width=0.8\textwidth]{cylinderUPlot.png}
-	\caption{A 3-dimensional complex plot of the parabolic cylinder function $U(a,z)$ at $a=0$ plotted by \Maple.~\cite[(12.2i)]{NIST:DLMF}}
-	\label{fig:cylinderU}
-\end{figure}
+%\begin{figure}[ht]
+%	\centering
+%	\includegraphics[width=0.8\textwidth]{cylinderUPlot.png}
+%	\caption{A 3-dimensional complex plot of the parabolic cylinder function $U(a,z)$ at $a=0$ plotted by \Maple.~\parencite[(12.2i)]{NIST:DLMF}}
+%	\label{fig:cylinderU}
+%\end{figure}
 
-Another way to plot such complex functions is to plot the real and imaginary parts of the solution separately.
+%Another way to plot such complex functions is to plot the real and imaginary parts of the solution separately.
 Figure~\ref{fig:cylinderUSplit} illustrates this plotting method for the parabolic cylinder function again.
 
-Note that for real function plots, such separate plots for the imaginary and real values of a complex function, will use a color map as well.
+%Note that for real function plots, such separate plots for the imaginary and real values of a complex function, will use a color map as well.
 In that case, the color map is just an adornment and has nothing to do with complex color maps.
@@
 
-\begin{figure}[!htp]
-    \centering
-    \subfloat[Plot of the real value of $\ParabolicU@{0}{z}$.]{%
-        \includegraphics[width=0.8\textwidth]{cylinderURe.png}
-        \label{fig:cylinderURe}%
-    }
-    \vspace{-0.1cm}
-    \subfloat[Plot of the imaginary value of $\ParabolicU@{0}{z}$.]{%
-        \includegraphics[width=0.8\textwidth]{cylinderUIm.png}
-        \label{fig:cylinderUIm}%
-    }
-    \caption{Separately plot of the real~\protect\subref{fig:cylinderURe} and imaginary~\protect\subref{fig:cylinderUIm} part of the parabolic cylinder function $U(a,z)$ at $a=0$.}
-    \label{fig:cylinderUSplit}
-\end{figure}
+%\begin{figure}[!htp]
+%    \centering
+%    \subfloat[Plot of the real value of $\ParabolicU@{0}{z}$.]{%
+%        \includegraphics[width=0.8\textwidth]{cylinderURe.png}
+%        \label{fig:cylinderURe}%
+%    }
+%    \vspace{-0.1cm}
+%    \subfloat[Plot of the imaginary value of $\ParabolicU@{0}{z}$.]{%
+%        \includegraphics[width=0.8\textwidth]{cylinderUIm.png}
+%        \label{fig:cylinderUIm}%
+%    }
+%    \caption{Separately plot of the real~\protect\subref{fig:cylinderURe} and imaginary~\protect\subref{fig:cylinderUIm} part of the parabolic cylinder function $U(a,z)$ at $a=0$.}
+%    \label{fig:cylinderUSplit}
+%\end{figure}
 
-\subsection{Multivalued Functions}
-The following definitions for multivalued functions are based on \textit{Complex Variables}~\cite{ComplexVariables} by M.~J.~Ablowitz and A.~S.~Fokas.
+%\subsection{Multivalued Functions}
+%The following definitions for multivalued functions are based on \textit{Complex Variables}~\parencite{ComplexVariables} by M.~J.~Ablowitz and A.~S.~Fokas.
 
-Unlike a typical\footnote{In lack of a better word to distinguish single-valued and multivalued functions, we call single-valued functions here typical functions, because these functions are best known.} function, multivalued functions associate not only one output to any particular input, but a whole set of output values.
+%Unlike a typical\footnote{In lack of a better word to distinguish single-valued and multivalued functions, we call single-valued functions here typical functions, because these functions are best known.} function, multivalued functions associate not only one output to any particular input, but a whole set of output values.
 A real-valued example is the square root function, which associates for each positive real number two real square roots.
@@
 
-Consider now the natural logarithm with complex arguments and $z = r\expe^{\iunit\phi}$ in polar coordinates.
+%Consider now the natural logarithm with complex arguments and $z = r\expe^{\iunit\phi}$ in polar coordinates.
 Since it is the natural logarithm, $\ln@{z} = \ln@{r} + \iunit\phi$ and the imaginary part is just $\phi$.
@@
 
-Defining branch cuts to allow analytic computations on the principal branches is a common task for \gls{cas}~\cite{Maple:Cuts}.
+Defining branch cuts to allow analytic computations on the principal branches is a common task for \gls{cas}~\parencite{Maple:Cuts}.
 Unfortunately, the position of the branch cuts is not well-defined.
@@
 The branch cut at $(-\infty, 0]$ for the complex square root function is one of those standard positions.
-However, the positions can vary from system to system~\cite{Branches:acot}.
+However, the positions can vary from system to system~\parencite{Branches:acot}.
 
@@
 \subsection{Elementary \& Special Functions}\label{subsec:special-functions}
-A function is called \textbf{elementary} if it can be constructed using only a finite combination of constant functions, field operations, and algebraic, trigonometric, exponential, and logarithmic functions together with their inverses~\cites{Elementary:1}[145]{Elementary:2}.
+A function is called \textbf{elementary} if it can be constructed using only a finite combination of constant functions, field operations, and algebraic, trigonometric, exponential, and logarithmic functions together with their inverses~\parencites{Elementary:1}[145]{Elementary:2}.
 
@@
 For example, in 1964 Milton Abramowitz and Irene A.
-Stegun published the \textit{Handbook of Mathematical Functions with Formulas, Graphs, and Mathematical Tables}~\cite{AbramowitzStegun} at \gls{nist}\footnote{In 1988 the \textit{National Bureau of Standards} (NBS) was renamed to the \textit{National Instutute of Standards and Technology} (NIST).}.
+Stegun published the \textit{Handbook of Mathematical Functions with Formulas, Graphs, and Mathematical Tables}~\parencite{AbramowitzStegun} at \gls{nist}\footnote{In 1988 the \textit{National Bureau of Standards} (NBS) was renamed to the \textit{National Instutute of Standards and Technology} (NIST).}.
 The handbook became the most comprehensive source of information on special functions.
-Therefore, the \gls{nist} has published a new version called \textit{NIST Handbook of Mathematical Function}~\cite{NIST:Handbook} in 2010 to replace the predecessor.
+Therefore, the \gls{nist} has published a new version called \textit{NIST Handbook of Mathematical Function}~\parencite{NIST:Handbook} in 2010 to replace the predecessor.
 
 Because there is no formal definition for special functions, some problems appear for our translator project. For example the \textit{Encyclop\ae{dia} Britannica} wrote:
-\begin{myQuote}{Encyclop\ae{dia} Britannica - Special function~\cite{SpecialFunctions:britannica}}
+\begin{myQuote}{Encyclop\ae{dia} Britannica - Special function~\parencite{SpecialFunctions:britannica}}
 	Special function, any of a class of mathematical functions that arise in the solution of various classical problems of physics.
@@
 
-In \textit{Special Functions}~\cite{SF:Book}, G.~E.~Andrews~et~al. wrote in the preface:
-\begin{myQuote}{Special Functions - Preface~\cite{SF:Book}}
+In \textit{Special Functions}~\parencite{SF:Book}, G.~E.~Andrews~et~al. wrote in the preface:
+\begin{myQuote}{Special Functions - Preface~\parencite{SF:Book}}
 Paul Tur\'an once remarked that special functions would be more appropriately labeled "useful functions".
@@
 
-Our reference for special functions in this thesis will be the \textit{NIST Handbook of Mathematical Function} and its digital counterpart the \gls{dlmf}~\cite{NIST:DLMF}.
+Our reference for special functions in this thesis will be the \textit{NIST Handbook of Mathematical Function} and its digital counterpart the \gls{dlmf}~\parencite{NIST:DLMF}.
 See the later section~\ref{sec:dlmf} for a more detailed explanation.
@@
 Since most of the special functions represent solutions for other mathematical expressions, they were intensively studied - mostly in applied mathematics.
-In the past, huge tables of values were used for calculations~\cite{Tables}.
-Therefore, the book by Abramowitz and Stegun also contains tables with values for several special functions~\cite{AbramowitzStegun}.
+In the past, huge tables of values were used for calculations~\parencite{Tables}.
+Therefore, the book by Abramowitz and Stegun also contains tables with values for several special functions~\parencite{AbramowitzStegun}.
 Obviously, computations of special functions with \gls{cas} were desired.
@@
 Those differences can even appear in well-known elementary functions.
-For example, the \gls{dlmf} defines the branch cut for the inverse cotangent function~\cite[(4.23.9)]{NIST:DLMF} at $[-\iunit, \iunit]$, while \Maple{} reasonably~\cite{Branches:acot} defines the branch cut at $(-\infty\iunit, -\iunit]$ and $[\iunit, \infty\iunit)$.
+For example, the \gls{dlmf} defines the branch cut for the inverse cotangent function~\parencite[(4.23.9)]{NIST:DLMF} at $[-\iunit, \iunit]$, while \Maple{} reasonably~\parencite{Branches:acot} defines the branch cut at $(-\infty\iunit, -\iunit]$ and $[\iunit, \infty\iunit)$.
 Figure~\ref{fig:acot} illustrates the problem.
@@
 Instead of translating the function itself, the translator uses other equivalent representations for the function.
-In case of the arccotangent function, we can use the following alternative representations, based on the suggestions in~\cite{Branches:acot}.
+In case of the arccotangent function, we can use the following alternative representations, based on the suggestions in~\parencite{Branches:acot}.
 
@@
 A more complicated example might be the equation of the already introduced parabolic cylinder function $U(a,z)$ and the modified Bessel function of the second kind $\BesselK{\nu}@{z}$.
-The parabolic cylinder function is a solution of a second order differential equation~\cite[(12.2i)]{NIST:DLMF} and can be represented by $\BesselK{\nu}@@{z}$ for $a=0$.
+The parabolic cylinder function is a solution of a second order differential equation~\parencite[(12.2i)]{NIST:DLMF} and can be represented by $\BesselK{\nu}@@{z}$ for $a=0$.
 The relation is defined in~(\ref{eq:branch-cut-issue}).
@@
 Since \gls{cas} do not provide Riemann surfaces, we use another approach to solve this problem, called \textbf{analytic continuation}.
-Analytic continuation is the process of extending the range of validity of a representation or more generally extending the region of definition of an analytic function~\cite[152]{ComplexVariables}.
-For the modified Bessel function of the second kind the analytic continuation is defined in the \gls{dlmf}~\cite[(10.34.4)]{NIST:DLMF}.
+Analytic continuation is the process of extending the range of validity of a representation or more generally extending the region of definition of an analytic function~\parencite[152]{ComplexVariables}.
+For the modified Bessel function of the second kind the analytic continuation is defined in the \gls{dlmf}~\parencite[(10.34.4)]{NIST:DLMF}.
 With analytic continuation, the right hand side of equation~(\ref{eq:branch-cut-issue}) is illustrated in figure~\ref{fig:analytic-cont}.
@@
 
-Mathematical expressions can be represented by ordered trees consisting of terminal symbols, such as identifiers or numbers (leaf nodes), and functions or operators (non-leaf nodes)~\cite{VMEXT}.
+Mathematical expressions can be represented by ordered trees consisting of terminal symbols, such as identifiers or numbers (leaf nodes), and functions or operators (non-leaf nodes)~\parencite{VMEXT}.
 Consider the mathematical expression
@@
 
-The tool \gls{vmext}~\cite{VMEXT} visualizes expression trees. Figure~\ref{fig:simple-expr-tree} draws the expression tree for (\ref{eq:simple-expr}).
+The tool \gls{vmext}~\parencite{VMEXT} visualizes expression trees. Figure~\ref{fig:simple-expr-tree} draws the expression tree for (\ref{eq:simple-expr}).
 
@@
     \includegraphics[width=0.7\textwidth]{expressionTreeSimple.png}
-    \caption{The expression tree of (\ref{eq:simple-expr}) visualized with VMEXT~\cite{VMEXT}.}
+    \caption{The expression tree of (\ref{eq:simple-expr}) visualized with VMEXT~\parencite{VMEXT}.}
     \label{fig:simple-expr-tree}

### 2018-07-20 chapters/02-background/related-work.tex
Update introduction with more important information

@@
 \section{Related Work}\label{sec:related-work}
-Since \LaTeX{} became the de facto standard for writing papers in mathematics, most of the \gls{cas} nowadays are able to import and export mathematical \LaTeX{} expressions.
+Since \LaTeX{} became the de facto standard for writing papers in mathematics, most of the \gls{cas} provide simple functions to import and export mathematical \LaTeX{} expressions%
+\footnote{The selected \gls{cas} Maple, Mathematica, Matlab, and SageMath provide import and/or export functions for \LaTeX:\\
+Maple, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;\\
+Mathematica, \url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html} seen 06/2017;\\
+Matlab, \url{https://www.mathworks.com/help/symbolic/latex.html} seen 06/2017;\\
+SageMath, \url{http://doc.sagemath.org/html/en/tutorial/latex.html} seen 06/2017}. %
 Those tools have two essential problems.

### 2018-07-20 chapters/03-translator/backward-trans.tex
Update introduction with more important information

@@
 Mostly because \Maple{} automatically evaluates input expressions immediately - even before we have a chance to take a look at the internal data structure of the input.
-For example, an input such as $\sin@{\frac{\cpi}{2}}$ is immediately evaluated to $1$ and the internal data structure just shows us the positive integer value $1$ instead of the trigonometric function.
+For example, an input such as $\sin\left(\frac{\cpi}{2}\right)$ is immediately evaluated to $1$ and the internal data structure just shows us the positive integer value $1$ instead of the trigonometric function.
 

### 2018-07-23 chapters/02-background/related-work.tex
Update until forward/backword translation

--- 2018-07-17 chapters/02-background/related-work.tex
@@
 \section{Related Work}\label{sec:related-work}
-Since \LaTeX{} became the de facto standard for writing papers in mathematics, most of the \gls{cas} nowadays are able to import and export mathematical \LaTeX{} expressions.
+Since \LaTeX{} became the de facto standard for writing papers in mathematics, most of the \gls*{cas} provide simple functions to import and export mathematical \LaTeX{} expressions%
+\footnote{The selected \gls*{cas} Maple, Mathematica, Matlab, and SageMath provide import and/or export functions for \LaTeX:\\
+Maple, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;\\
+Mathematica, \url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html} seen 06/2017;\\
+Matlab, \url{https://www.mathworks.com/help/symbolic/latex.html} seen 06/2017;\\
+SageMath, \url{http://doc.sagemath.org/html/en/tutorial/latex.html} seen 06/2017}. %
 Those tools have two essential problems.
-First of all, they cannot import mathematical expressions where the semantic information is absent, which is the usual case for generic \LaTeX{} expressions.
-Therefore, they are only able to import simple mathematical expressions, where the semantics are unique.
+They are only able to import simple mathematical expressions, where the semantics are unique.
 For example, the internal \LaTeX{} macro \verb|\frac| always indicates a fraction.
-These import tools fail for more complex expressions, for example the Jacobi polynomial in \cref{tab:JacobiP-usecase}.
+However, for more complex expressions, e.g., the Jacobi polynomial in \cref{tab:JacobiP-usecase}, the import functions fail.
 The second problem appears in the export tools.
-Mathematical expressions in \gls{cas} are fully semantic, otherwise the \gls{cas} wouldn't be able to compute or evaluate the expressions.
+Mathematical expressions in \gls*{cas} are fully semantic, otherwise the \gls*{cas} wouldn't be able to compute or evaluate the expressions.
 During the export process, the semantic information gets lost, because generic \LaTeX{} is not able to carry semantic information.
 In consequence of these two problems, an exported expression cannot be imported to the same system again in most cases (except those simple expressions described above).
-Our tool should solve these problems and provide round trip translations between \LaTeX{} and \gls{cas}.
+Our tool should solve these problems and provide round trip translations between \LaTeX{} and \gls*{cas}.
 
-Obviously, the semantics must be well known before an expression can be translated.
-There are two approaches to solve that problem: someone could specify the semantic information during the writing process (pre-define semantics) or the translator can find out the right semantic information in general mathematical expressions before it translates the expression.
-The second approach is unreasonably difficult compared to the pre-definition approach.
-We will talk about those difficulties in the following sections.
-We will also introduce a new approach in \cref{ch:conc-future-work} to achieve the second idea.
-
-However, the pre-definition approach is simple and easy to realize, as long as the system can carry semantic information.
-A typical, not mathematical example are Wikipedia articles~\parencite{Wiki}.
-An author of such an article can specify further information about symbols, words and sentences by creating a hyperlink to a more detailed explanation.
-Usually, the author defines the hyperlink, when he is writing the article or he needs to manually add the hyperlink later on.
-Similar to this approach are \textit{interactive documents}\footnote{There is no adequate definition what interactive documents are.
-However, this name is widely used to describe electronic document formats that allow interactivity to change the content in real time.}, such as the \gls{cdf}~\parencite{CDF:Wolfram} by Wolfram Research or the \textit{worksheets} of \Maple{}, which are also sometimes referred to as interactive documents.
-Those documents pre-define the semantics to allow computations.
+The semantics must be well known before an expression can be translated.
+There are two naiv approaches to solve that problem: (1) someone could specify the semantic information during the writing process (pre-defined semantics) or (2) the translator can find out the right semantic information in general mathematical expressions before it translates the expression.
+So called \textit{interactive documents}\footnote{There is no adequate definition what interactive documents are.
+However, this name is widely used to describe electronic document formats that allow interactivity to change the content in real time.}, such as the \gls*{cdf}~\parencite{CDF:Wolfram} by Wolfram Research or the \textit{worksheets} by \Maple{}, try to solve this problem with approach (2) and allow to embed semantic information into the input.
 Those complex document formats require specialized tools to show and work with the documents (Wolfram CDF Player, or \Maple{} for the \textit{worksheets}).
-The \JOBAD{} architecture~\parencite{JOBAD:orig} is able to create web based interactive documents and uses \gls{omdoc}~\parencite{OMDoc} to carry semantics.
+The \JOBAD{} architecture~\parencite{JOBAD:orig} is able to create web based interactive documents and uses \gls*{omdoc}~\parencite{OMDoc} to carry semantics.
 The documents can be viewed and edited in the browser.
-Those \JOBAD{-documents} also allow to perform computations during \gls{cas}.
+Those \JOBAD{-documents} also allow to perform computations during \gls*{cas}.
 This gives the opportunity to calculate, compute and change mathematical expressions directly in the document.
@@
 
-All of these systems carry semantic information.
+%All of these systems carry semantic information.
 Most of them create their own way to make this possible.
-The web based documents mostly use standardized ways, such as the content \gls{mathML}, a specialization of \gls{mathML}~\parencite{MathML}, that focuses on the semantics of an expression rather than any particular rendering for the expression.
+The web based documents mostly use standardized ways, such as the content \gls*{mathML}, a specialization of \gls*{mathML}~\parencite{MathML}, that focuses on the semantics of an expression rather than any particular rendering for the expression.
 Since \LaTeX{} is the de facto standard to write mathematical expressions in documents there are projects to extend \LaTeX{} in a way that enable to carry the semantic information directly in \LaTeX.
@@
 
-However, \textit{interactive documents}, web based or not, have an essential problem: all of them create an entire new document format.
+%However, \textit{interactive documents}, web based or not, have an essential problem: all of them create an entire new document format.
 We want to create an independent lite translation tool for \LaTeX.
@@
 
-Another approach that tries to avoid the translation problem are \LaTeX{} packages that allow computations directly through the \LaTeX{} compiler (for example \textit{LaTeXCalc}~\parencite{LatexCalc}) or allow \gls{cas} commands for computations directly in \TeX -files.
-For example, \textit{sagetex}~\parencite{Sagetex} is an interface \LaTeX{} package for the open source \gls{cas} \textit{sage}\footnote{An abbreviation for \textit{SageMath}.}.
-This package allows \textit{sage} commands in \TeX -files and uses \textit{sage} in the background to compute the commands.
-Obviously, the first example is not as powerful as a \gls{cas} would be.
+Another approach tries to avoid translation problems by allow computations directly via the \LaTeX{} compiler, e.g., \textit{LaTeXCalc}~\parencite{LatexCalc}.
+Those packages are limited to the possibilities of the compiler and therefore not as powerful as \gls*{cas}.
+A workaround for this case is \textit{sagetex}~\parencite{Sagetex}, which is a \LaTeX{} package interface for the open source \gls*{cas} \textit{sage}\footnote{An abbreviation for \textit{SageMath}.}.
+This package allows \textit{sage} commands in \TeX{-}files and uses \textit{sage} in the background to compute the commands.
+In this scenario, a writer still needs to manually translate expressions to the syntax of \textit{sage}.
+
+%Obviously, the first example is not as powerful as a \gls*{cas} would be.
 However, \textit{sagetex} does not really solve our problem of the scientific workflow.
@@
 
-In summary, today's translation tools are either specialized for one specific \gls{cas} (such as the import and export functions of the \gls{cas}) or they completely avoid the translation process by creating a new document format (such as \textit{interactive documents}).
-Our goal is to provide a lite translation tool for mathematical \LaTeX{} expressions and multiple \gls{cas}.
+%In summary, today's translation tools are either specialized for one specific \gls*{cas} (such as the import and export functions of the \gls*{cas}) or they completely avoid the translation process by creating a new document format (such as \textit{interactive documents}).
+Our goal is to provide a lite translation tool for mathematical \LaTeX{} expressions and multiple \gls*{cas}.
+
+There exists two approaches that allow to embed semantic information within \LaTeX{} expressions by using custom macros.
+Namely, \sTeX{}~\parencite{sTeX} developed by Kohlhase and the \Macro s developed by Miller~\parencite{DLMF:Macros}.
+This paper shows that it is possible to develop a context-free translation tool using the semantic macros introduced by these two projects.
+\sTeX{} aims to embed semantic information for general mathematics with a comprehensive set of macros.
+The macro set developed by Miller introduces new macros for special functions, orthogonal polynomials, and mathematical constants.
+Each of these macros ties specific character sequences to a well-defined mathematical object and is linked with the corresponding definition in the \gls*{dlmf} or \gls*{drmf}.
+Therefore, we call these semantic macros \Macro s.
+These semantic macros are internally used in the \gls*{dlmf} and the \gls*{drmf}.
+Because of the linked definitions to the \gls*{dlmf} and the special macros for functions and mathematical constants this paper using the \Macro s for performing translations to \gls*{cas} instead of using \sTeX.

### 2018-07-23 chapters/03-translator/gen-app.tex
Update introduction with more important information
Update until forward/backword translation

@@
-%The overall goal of the whole project was to provide an interactive one click-translation for each formula in the \gls{drmf} to a specified \gls{cas}.
-The program mainly supports translations from and to the \gls{cas} \Maple.
-However, the program is designed to support other \gls{cas}.
-In the current stage it also supports translations to \Mathematica{}, but the backend library needs to get extended for a full support.
+\section{Translation Problems}
+There are several potential problems to peform translations between systems that embed semantic information in the input.
+Those problems vary from simple cases, e.g., a function is not defined in system, to complex cases, e.g., different positioning of branch cuts for multivalued functions.
+This section will discuss the problems and workarounds.
 
-%In this chapter, the functionality and the design of the translator will be explained.
-The structure follows the development process and explains our decisions, problems and solutions.
-It starts with an overview of our approaches in~\ref{sec:gen-appr} and solutions for the problems.
-We will also focus on the theory of our backend libraries in~\ref{subsec:libs} and how to extend them in~\ref{subsec:extension}.
-The sections~\ref{sec:forward-translation} and \ref{sec:backward-translation} explains the forward and backward translator separately.
-
-%\section{General Approach and Goals}\label{sec:gen-appr}
-%Initially this project aimed for translations not only for semantic \LaTeX{} expressions but for generic \LaTeX{} as well.
-However, as discussed before, the semantic information in generic \LaTeX{} is hardly accessible (as seen in \ref{subsec:semantic-latex}).
-Supporting generic \LaTeX{} translations would go beyond the scope of a Master's thesis.
-Hence, we focused on translations for semantic \LaTeX{} expressions.
-Furthermore, the translator should support multiple \gls{cas} and because of the work in progress state of the \Macro{} it should be easily extendable and changeable.
-Due to the time limitations of the project we focused our implementations on the support for \Maple{} and \Mathematica.
-
-%Another strive for a goal was to keep a translation as straight forward as possible.
-In other words, the structure of the translated expression should be similar to the structure of the input expression.
-This makes it easier for a user to follow the translation process and debug the program.
-This intention should not to be confused with the task for \textit{appropriate} translations.
-We will see that a translation process (especially the backward translation) can easily change expressions without changing the semantics.
-We want to avoid that behavior in our translations as often as possible.
-
-%Furthermore, the translation process should solve problems of existing technologies, presented in section~\ref{sec:related-work}.
-In the following subsection, we will focus on known problems and present solutions before we start to explain implementation details.
-
-\section{Translation Problems}
-Most of the existing translation tools cannot preserve the semantic information during the translation process (see section~\ref{sec:related-work}).
-Therefore, our program should not lose semantic information during the process.
-However, the semantics can be changed during a translation.
-This happens when a definition or property of a mathematical object varies from system to system.
-Our reference systems for all mathematical \LaTeX{} expressions are the \gls{dlmf} and the \gls{drmf}.
-
-Let us focus on translations of mathematical objects that are not defined in one of the languages, first.
-Obviously, we cannot translate these objects \textit{directly}.
-Our workaround is to translate not the mathematical object itself, but the definition of the object.
-For example, the \textit{Gudermannian}~\parencite[(4.23.10)]{NIST:DLMF} $\Gudermannian{x}$ function can be defined by
+If a function is defined in one system but not in the other, we can translate the definition of the mathematical function.
+For example, the \textit{Gudermannian}~\parencite[eq.
+4.23.10]{NIST:DLMF} $\Gudermannian{x}$ function is defined by
 \begin{equation}\label{eq:gudermannian}
@@
 \end{equation}
-This function is not implemented in \Maple. Therefore, we translate its definition~\eqref{eq:gudermannian}
+and linked with the semantic macro \verb|\Gudermannian| in the \gls*{dlmf} but does not exist in \Maple. We can perform a translation for the definition~\ref{eq:gudermannian} instead of macro itself
 \begin{equation}\label{eq:guder-trans}
@@
 
-We want to avoid such translations, because they are not intuitive and therefore can confuse a user. However, providing such translations is still better than providing no translation.
+Since those translations are nonintuitive, describing explanations become necessary for the translation process.
+A special logging function take care about each translation and provide details after a successful translation process.
+Section~\ref{sec:forward-translation} explains this task further.
 
-Instead of none possible translation, how can we handle multiple alternative translations?
-We already discussed the problem of the inverse trigonometric function $\acot@{x}$.
-Because of a different branch cut in \Maple{} (compared to the \gls{dlmf} definition), there are alternative translations available (see~\eqref{eq:acot-alternatives}).
-In such cases, we define the most intuitive, the direct translation.
-In addition, we inform the user about the available alternative translations and the branch cut issues.
-Hence,
+Providing detailed information also solves the problem for multiple alternative translations.
+In some cases, a semantic macro has two alternative representations in the \gls*{cas} or vice versa.
+In such cases, the translator picks one of the alternatives and informs the user about the decision.
+
+In case of differences between defined branch cuts we can also use alternative translations to solve the problems.
+Consider the mentioned case of the arccotangent function~\parencite{Branches:acot} that has different positioned branch cuts in \Maple{} compared to the \gls*{dlmf} or \Mathematica{} definitions.
+Alternative but mathematically equivalent translations are
+\begin{eqnarray}
+\verb|\acot@{z}| & \overset{\langMaple}{\mapsto} & \verb|arccot(z)|,\label{eq:acot-alternatives}\\
+& \overset{\langMaple}{\mapsto} & \verb|arctan(1/z)|,\label{eq:acot-alternatives-1}\\
+& \overset{\langMaple}{\mapsto} & \verb|I/2*ln((z-I)/(z+I))|.\label{eq:acot-alternatives-2}
+\end{eqnarray}
+
+The direct translation~(\ref{eq:acot-alternatives}) has the branch cut issue, while the alternative translations~(\ref{eq:acot-alternatives-1}) and~(\ref{eq:acot-alternatives-2}) using other functions instead of the arccotangent function.
+The arctangent function~(\ref{eq:acot-alternatives-1}) and the natural logarithm~(\ref{eq:acot-alternatives-2}) have the same positioned branch cuts in the \gls*{dlmf} and in \Maple.
+In consequence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user do not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except for $z = -\iunit$.
+
+Other problematic cases for translations are the \Macro s itself.
+In some cases, they do not provide sufficient semantic information to perform translations.
+One example is the \textit{Wronskian} symbol.
+For two differentiable functions the \textit{Wronskian} is defined as~\parencite[eq.
+1.13.4]{NIST:DLMF}
+\begin{equation*}
+	\mathscr{W} \left\{ w_1(z), w_2(z) \right\} = w_1(z)w_2'(z) - w_2(z)w_1'(z).
+\end{equation*}
+In semantic \LaTeX{} it is used with
 \begin{equation}
-\verb|\acot{x}| \overset{\langMaple}{\mapsto} \texttt{arccot(x)}.
+	\verb|\Wronskian@{w_1(z), w_2(z)}|.
 \end{equation}
+A translation become infeasible, because the macro do not explicitely defined the variable of the functions $w_1$ and $w_2$.
+For a correct translation, the \gls*{cas} needs to be aware of the used variable $z$.
+We solved this issue by creating a new macro
+\begin{equation}
+	\verb|\Wron{z}@{w_1(z)}{w_2(z)}|.
+\end{equation}
+This example shows that the \Macro s are still work in progress and are getting constantly updated.
 
-Since we have these alternative translations, we also use the same technique for the previous Gudermannian example.
-We provide one direct translation and possible alternative translations for a mathematical object.
+A similar problem are multiplications, since they are rarely explicitely marked in \LaTeX{} expressions, e.g., scientists using whitespaces to indicate multiplications rather than using \verb|\cdot| or similar symbols.
+For such problems, we introduced a new macro \verb|\idot| for an invisible multiplication sign (this macro will not be rendered).
+Since this macro is newly introduced and an automatic conversion of existing equations is difficult, none of the equations in the \gls*{dlmf} uses this macro yet.
+As a consequence, the translator has some simple rules to perform translations without explicitely marked translations with \verb|\idot|.
 
-These two problems already explained, why our translation process is in general not bijective.
-However, we will explain later in a more detailed way that the translation is only bijective for a subset of possible expressions.
-Thereby, a bigger goal is to increase this subset as much as possible.
-
-\subsection{Libraries}\label{subsec:libs}
-The first step for providing an automatic translation tool is to defining the translation for each function.
+\section{The Translator}
+%The first step for providing an automatic translation tool is to defining the translation for each function.
 We manually create a table of translations for each \Macro.
@@
 
-For each translation, we design a translation pattern with placeholders to define the correct position of each argument.
+%For each translation, we design a translation pattern with placeholders to define the correct position of each argument.
 The placeholders are defined as \verb|$i|, where \verb|i| is a non-negative integer specifying the position, starting at $0$.
 Table~\ref{tab:placeholder_ex1} illustrates the translation patterns for the trigonometric sine function in semantic \LaTeX, \Maple{} and \Mathematica.
-\begin{table}[ht]
-	\centering
-	\begin{tabular}{lc}
-		\hline
-		Semantic \LaTeX & \verb|\sin@{$0}| \\
-		\Maple & \verb|sin($0)| \\
-		\Mathematica & \verb|Sin[$0]|\\
-		\hline
-	\end{tabular}
-	\caption{Translation patterns for the sine function in the \DLMF, \Maple{} and \Mathematica.}
-	\label{tab:placeholder_ex1}
-\end{table}
+%\begin{table}[ht]
+%	\centering
+%	\begin{tabular}{lc}
+%		\hline
+%		Semantic \LaTeX & \verb|\sin@{$0}| \\
+%		\Maple & \verb|sin($0)| \\
+%		\Mathematica & \verb|Sin[$0]|\\
+%		\hline
+%	\end{tabular}
+%	\caption{Translation patterns for the sine function in the \DLMF, \Maple{} and \Mathematica.}
+%	\label{tab:placeholder_ex1}
+%\end{table}
 
-Placeholders gives us the opportunity to define the order of the arguments and handle nested function calls.
+All translations are defined by a library that defines translation patterns for each function and symbol.
+The pattern uses \verb|$i| as placeholders to define the positions of the arguments.
+For example, the translation patterns for the Jacobi polynomial are illustrated in table~\ref{tab:placeholder_ex2}
+
+%Placeholders gives us the opportunity to define the order of the arguments and handle nested function calls.
 The order of arguments can differ between the systems.
@@
 
-Note that \Maple{} also uses the symbol \verb|$| in few functions.
-This can cause trouble in the replacement process.
-However, we can handle all of the few cases by adding additional parenthesis.
-Consider the differentiation function in \Maple
+In some special cases, these placeholders causes trouble when the \gls*{cas} uses the symbol \verb|$| for other reasons, e.g., the differentiation in \Maple{} is defined by
 \begin{equation}
@@
 \end{equation}
-with $f$ as an algebraic expression or an equation, $x$ the name of the differentiation variable and $n$ for the $n$-th order differentiation. Therefore
+where $f$ as an algebraic expression or an equation, $x$ the name of the differentiation variable and $n$ for the $n$-th order differentiation.
+A translation for $\frac{d^2x^2}{dx^2}$ should be like this
 \begin{equation}
-\frac{d^2x^2}{dx^2} = \verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [x$2])|.
+\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [x$2])|
 \end{equation}
-But the translation pattern in \Maple{} would be
+but would end up as
 \begin{equation}
-\verb|diff($1, [$2$$0])|.
+\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff($1, [$2$$0])|.
 \end{equation}
-Replace the placeholders sequentially by the arguments ends in
-\begin{eqnarray*}
-&&\verb|diff($1, [$2$$0])|\\
-\text{Replace index 0 by $2$\hphantom{$^2$}}& \rightarrow &\verb|diff($1, [$2$2])|\\
-\text{Replace index 1 by $x^2$}& \rightarrow &\verb|diff(x^2, [$2$2])|\\
-\text{Replace index 2 by $x$\hphantom{$^2$}}& \rightarrow &\verb|diff(x^2, [xx])|.
-\end{eqnarray*}
-We use parenthesis to solve this issue and define the translation pattern with
+
+%Replace the placeholders sequentially by the arguments ends in
+%\begin{eqnarray*}
+%&&\verb|diff($1, [$2$$0])|\\
+%\text{Replace index 0 by $2$\hphantom{$^2$}}& \rightarrow &\verb|diff($1, %[$2$2])|\\
+%\text{Replace index 1 by $x^2$}& \rightarrow &\verb|diff(x^2, [$2$2])|\\
+%\text{Replace index 2 by $x$\hphantom{$^2$}}& \rightarrow &\verb|diff(x^2, [xx])|.
+%\end{eqnarray*}
+
+We can solve this issue by using paranthesis in such cases, e.g., 
 \begin{equation}
@@
 
-A special problem can be binary operators.
+%A special problem can be binary operators.
 A good example is the multiplication sign.
@@
 
-As already mentioned, the information is stored in three different file types. Together they form the backend library.
+%As already mentioned, the information is stored in three different file types. Together they form the backend library.
 
-\subsubsection{CSV Tables}
-The \gls{csv} file represents the translation tables generated by MS Excel.
+%\subsubsection{CSV Tables}
+%The \gls{csv} file represents the translation tables generated by MS Excel.
 Although \gls{csv} is an abbreviation for \textit{comma-seperated}, one can specify the split symbol.
@@
 
-Note that \gls{csv} files are databases.
+%Note that \gls{csv} files are databases.
 While we do not want to introduce the theory of databases in this thesis, note that our databases contain a \textit{primary key}.
@@
 
-Since a translation is not necessarily bijective, the translations are defined in two separate \gls{csv} files per \gls{cas}.
+%Since a translation is not necessarily bijective, the translations are defined in two separate \gls{csv} files per \gls{cas}.
 For \Maple{} those files are \verb|DLMF_Maple.csv| for the forward translation, and \verb|CAS_Maple.csv| for the backward translation.
@@
 
-In~\ref{eq:guder-trans} we defined the definition for the Gudermannian function.
+%In~\ref{eq:guder-trans} we defined the definition for the Gudermannian function.
 The translated expression contains two function calls: the inverse tangent function and the hyperbolic sine function.
@@
 
-\begin{table}
+The \Macro s also allow to specify optional arguments to distinguish between standard and other version of functions.
+The Legendre and associated Legendre function of the first kind are examples for such cases.
+The library that defines translations for each macro using the macro name as a primary key to identify the translation.
+The Legendre and associated Legendre function of the first kind both using the same macro \verb|\LegendreP|.
+To distuingish such cases we using a special syntax such cases, shown in table~\ref{tab:legendreP-lex}.
+
+\begin{table}[ht]
 	\centering
@@
 		\hline
-		Semantic Macro & Entry in the CSV file\\
+		Semantic Macro Entry & Maple Entry \\
 		\hline
-		\verb|\LegendreP{\nu}@{x}| & \verb|\LegendreP{\nu}@{x}| \\
-		\verb|\LegendreP[\mu]{\nu}@{x}| & \verb|X1:\LegendrePX\LegendreP[\mu]{\nu}@{x}|\\
+		\verb|\LegendreP{\nu}@{x}| & \verb|LegendreP($0, $1)| \\
+		\verb|X1:\LegendrePX\LegendreP[\mu]{\nu}@{x}| & \verb|LegendreP($1, $0, $2)|\\
 		\hline
 	\end{tabular}
-	\caption{CSV entry example of the Legendre and associated Legendre function of the first kind.}
+	\caption{Entry of the translation library example of the Legendre and associated Legendre function of the first kind.
+The prefix notation \texttt{X<d>:<name>X} defines the translation for \texttt{<name>} with \texttt{<d>}-number of optional arguments.}
 	\label{tab:legendreP-lex}
@@
 
-In \Maple{} the number of optional arguments is not stored in the dictionaries.
-Therefore, we link to the total number of arguments.
-Consider Olver's associated Legendre function.
-A forward translation can be defined with
-\begin{equation}
-\verb|\LegendreBlackQ{\nu}@{z}| \overset{\langMaple}{\mapsto} \verb|LegendreQ(\nu,z)/GAMMA(\nu+1)|.
-\end{equation}
-In this case, we would prefer to provide the hyperlink to the definition of \Maple's \verb|LegendreQ| function rather than to the definition of the \verb|GAMMA| function.
-The used function has two arguments in \Maple.
-Hence, the lexicon entry for the translation pattern is
-\begin{equation}
-\verb|X2:LegendreQXLegendreQ($0,$1)/GAMMA($1+1)|.
-\end{equation}
+The translator using the \gls*{pom}-Tagger~\parencite{POM-Tagger}\footnote{Named according to the Part-of-Speech-Taggers in \gls*{nlp}.} to parse \LaTeX{} expressions into a parsed tree.
+The \gls*{pom}-Tagger is an \texttt{LL}-Praser defined by a context-free grammar in \gls*{bnf}.
+Each token will be tagged by meta information defined in lexicon files.
+We extend the lexicon files to provide also the information that are necessary for the translation process.
+An example for an entry of the lexicon file is given in table~\ref{tab:sine-lex-example}.
 
-All of the \gls{csv} files just store the information to provide an easy access for the developer.
-Nevertheless, the program itself works only with the \textit{lexicon files}.
-Therefore, each \gls{csv} file needs to get converted into a lexicon file.
-Subsection~\ref{subsec:extension} explains how to do that and how to update the database.
-
-\subsubsection{JSON Libraries}
-Some translations are defined in separate \gls{json} files.
-The reason is to separate translations for \gls{opsf} from translation for single symbols and generic \LaTeX{} macros.
-These files therefore define translations for the following types expressions.
-
-\begin{itemize}
-\item \textbf{Generic \LaTeX{} macros for functions}:\newline 
-For example, the macros \verb|\frac{a}{b}|, \texttt{\tbs binom\{a\}\{b\}} and \verb|\sqrt{a}|.
-This section also defines translations for special symbols for mathematical functions such as the exclamation mark '\verb|!|' to indicate the factorial function or \verb|\mod| for the modulo function.
-\item \textbf{Mathematical Symbols}:\newline 
-For example, the symbol to indicate multiplications, sums, inequalities.
-\item \textbf{Greek Letters}:\newline 
-For example, \verb|\alpha|, \verb|\beta| or \verb|\Theta|.
-\item \textbf{Mathematical Constants}:\newline 
-For example \verb|\cpi| for the constant $\cpi$ or \verb|\EulerConstant| for the Euler-Mascheroni constant $\EulerConstant$.
-\end{itemize}
-
-\subsubsection{Lexicon Files}
-The knowledge of the \gls{mlp} is based on lexicon files, as explained in subsection~\ref{subsec:pom-mlp}.
+%\subsubsection{Lexicon Files}
+%The knowledge of the \gls{mlp} is based on lexicon files, as explained in subsection~\ref{subsec:pom-mlp}.
 Since the \gls{mlp} project was not a part of the \Macro s project, the \gls{bnf} grammar defines no rules about semantic macros.
@@
 
-Table~\ref{tab:sine-lex-example} presents the entry for the sine function in the \verb|DLMF-macros-lexicon.txt|.
+%Table~\ref{tab:sine-lex-example} presents the entry for the sine function in the \verb|DLMF-macros-lexicon.txt|.
 The \verb|global-lexicon.txt| file is the main dictionary of the \gls{mlp} and combines the whole knowledge of the parser in one file.
@@
 	\! & \! & \hspace{32pt} help/maple/view.aspx?path=sin\\
-	\! & \! & help/maple/view.aspx?path=sin\\
 	\! & \! & Mathematica: \verb|Sin[$0]|\\
 	\! & \! & Mathematica-Link: reference.wolfram.com/\\
-	\! & \! & \hspace{32pt} language/ref/Sin\\
+	\! & \! & \hspace{32pt} language/ref/Sin.html\\
 	\hline
@@
 
-The concept of the lexicon files is also adapted to store the backward translations from \Maple.
+%The concept of the lexicon files is also adapted to store the backward translations from \Maple.
 Since the \gls{mlp} project is developed to parse \LaTeX{} expressions, it could be confusing to use the original lexicon files to also store names of \Maple{} functions.
@@
 
-\subsection{Extending the Forward Translator}\label{subsec:extension}
-The whole project is still work in progress.
+%\subsection{Extending the Forward Translator}\label{subsec:extension}
+%The whole project is still work in progress.
 Therefore, it is an important goal to keep it easy to update the backend libraries.
@@
 
-\subsubsection{Adding Translations to Existing CAS}
-Since \gls{csv} files are not as easy to maintain as MS Excel\footnote{Microsoft Excel is widely used to work with tables and is able to import, export and manipulate \gls{csv} files.} worksheets, the workflow starts with MS Excel.
+%\subsubsection{Adding Translations to Existing CAS}
+%Since \gls{csv} files are not as easy to maintain as MS Excel\footnote{Microsoft Excel is widely used to work with tables and is able to import, export and manipulate \gls{csv} files.} worksheets, the workflow starts with MS Excel.
 Once a user updates the existing tables, the MS Excel file needs to get exported as a \gls{csv} file.
@@
 
-Once the \verb|lexicon-creator.jar| finished the conversion process, the translator is aware of all changes and ready to use.
+%Once the \verb|lexicon-creator.jar| finished the conversion process, the translator is aware of all changes and ready to use.
 
-\subsubsection{Supporting New CAS}
-To support a complete new \gls{cas} for the forward translation, the user needs to add the \gls{cas} in the \gls{json} files and create two new \gls{csv} files.
+%\subsubsection{Supporting New CAS}
+%To support a complete new \gls{cas} for the forward translation, the user needs to add the \gls{cas} in the \gls{json} files and create two new \gls{csv} files.
 Even if there is no backward translation implemented yet, the \gls{csv} file for the backward translation is needed.
 
-Once all of the files are updated, the \verb|lexicon-creator.jar| can be used to update the lexicon files.
+%Once all of the files are updated, the \verb|lexicon-creator.jar| can be used to update the lexicon files.

### 2018-07-24 chapters/03-translator/backward-trans.tex
Finalize first version of the journal paper Still 1.000 words too long

--- 2018-07-17 chapters/03-translator/backward-trans.tex
@@
 \section{Maple to Semantic \LaTeX{} Translator}\label{sec:backward-translation}
-The backward translation is based on the internal data structure of \Maple.
-Therefore, a license of \Maple{} is mandatory to perform backward translations.
-The following subsections will explain why this approach was necessary and how \Maple{} works internally.
-Using the internal data structure also requires to handle the internal properties of \Maple.
-This can causes some problems.
-Those problems and our solutions will be explained in subsection~\ref{subsec:maple-probs}.
-In the last subsection~\ref{subsec:maple-impl}, we will focus on the implementation of the backward translation.
+Instead of writing a custom \Maple{} syntax parser, we using \Maple's internal data structure to get a syntax tree of the input\footnote{In consequence, a license of \Maple{} is mandatory to perform backward translations.
+The translator is using the version \Maple{} 2016.}.
+\Maple{} allow different input styles.
+The \texttt{1D} input is mainly used for programming purposes and therefore also used to perform our translations.
+Internally, \Maple{} using \gls*{dag} as syntax trees.
 
-Note that all explanations in this section is based on \Maple's programming guide~\parencite{MAPLE:ProgrammingGuide} for \Maple{} 2016.
-
-\subsection{Internal Data Structure}
-While our forward translation is realized for the \texttt{1D} input representation in \Maple, this representation is not suitable for a backward translation.
+%\subsection{Internal Data Structure}
+%While our forward translation is realized for the \texttt{1D} input representation in \Maple, this representation is not suitable for a backward translation.
 Since \Maple{} uses it's own syntax and programming language, we would need a new parser for \Maple{} expressions to realize a backward translation tool.
@@
 
-As already mentioned, \Maple{} has several ways to represent an expression.
+%As already mentioned, \Maple{} has several ways to represent an expression.
 The \texttt{1D} and \texttt{2D} representations are used as input representations.
 Besides those, there are internal data structures to perform all kinds of computations on the expression.
-Internally, each expression is handled and stored as a \gls{dag}.
-The \Maple{} \gls{dag} is very similar to an expression tree.
+Internally, each expression is handled and stored as a \gls*{dag}.
+The \Maple{} \gls*{dag} is very similar to an expression tree.
 But for efficiency, it does not store copies of an object.
 Consider the integral from (\ref{eq:maple-input})
-\begin{equation}\label{eq:maple-input-2}
-\int\displaylimits_0^\infty \frac{\pi+\sin(2x)}{x^2} dx.
-\end{equation}
+%\begin{equation}\label{eq:maple-input}
+%\int\displaylimits_0^\infty \frac{\pi+\sin(2x)}{x^2} dx.
+%\end{equation}
 
-As already mentioned, the \Maple{} \gls{dag} has no copies of objects. Therefore, the '$x$' in \eqref{eq:maple-input-2} only appears once in the \gls{dag}.
-\begin{figure}[ht]
+%As already mentioned, the \Maple{} \gls*{dag} has no copies of objects. Therefore, the '$x$' in \eqref{eq:maple-input-2} only appears once in the \gls*{dag}.
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[scale=0.7]{DAG.pdf}
+%\caption{The \Maple{} DAG of equation~\eqref{eq:maple-input}}
+%\label{fig:maple-dag}
+%\end{figure}
+
+%The \Maple{} \gls*{dag} can be visualized in a graph, such as in figure~\ref{fig:maple-dag}. But the internal representation looks a bit different.
+
+\begin{wrapfigure}{r}{0.62\textwidth}
+	\vspace{-22pt}
 	\centering
-\includegraphics[scale=0.7]{DAG.pdf}
-\caption{The \Maple{} DAG of equation~\eqref{eq:maple-input}}
-\label{fig:maple-dag}
-\end{figure}
-
-The \Maple{} \gls{dag} can be visualized in a graph, such as in figure~\ref{fig:maple-dag}.
-But the internal representation looks a bit different.
-Each node in the \gls{dag} stores its children and has a header, which defines the type and the length of the node.
-Consider the polynomial $x^2+x$.
-Figure~\ref{fig:internal-maple-dag} illustrates the internal \gls{dag} representation with headers and arguments.
-
-\begin{figure}[ht]
-	\centering
-	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.7]{DAGreal.pdf}
+	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.5]{DAGreal.pdf}
+	\vspace{-26pt}
 	\caption{The internal \Maple{} DAG representation of $x^2+x$.}
 	\label{fig:internal-maple-dag}
-\end{figure}
+	\vspace{-15pt}
+\end{wrapfigure}
 
-\Maple{} has also another representation, similar to the \gls{dag}, that allows multiple copies of the same object in its representation.
-This representation is called \inertF{} and is more similar to an expression tree than \Maple's \gls{dag}.
-The \inertF{} represents the \gls{dag} as a tree by splitting object copies and displays the tree in a list structure.
+Each node in the \gls*{dag} stores its children and has a header, which defines the type and the length of the node.
+Consider the polynomial $x^2+x$.
+Figure~\ref{fig:internal-maple-dag} illustrates the internal \gls*{dag} representation with headers and arguments.
+
+One can access the internal data structure of expressions via the \texttt{ToInert} command.
+The returned \inertF{} is a nested list representation as a tree equivalent\footnote{A tree equivalent of a \gls*{dag} splitting nodes with multiple parents into multiple nodes so that each node has only one parent node.} of the internal \gls*{dag} for the given expression.
+Some of the important types for the nodes are specified in table~\ref{tab:maple-types}.
+
+%\Maple{} has also another representation, similar to the \gls*{dag}, that allows multiple copies of the same object in its representation.
+This representation is called \inertF{} and is more similar to an expression tree than \Maple's \gls*{dag}.
+The \inertF{} represents the \gls*{dag} as a tree by splitting object copies and displays the tree in a list structure.
 Each node (or element in the list) has the prefix \texttt{\_Inert\_XXX}, that indicates the node is in the \inertF.
@@
 
-\begin{table}[ht]
+\begin{table}[h]
 \centering
@@
 	\hline
-	SUM & Sums.\\
+	SUM & Sums. Internally stored with factors for each summand, i.e., '$x+y$' would be stored as '$x \cdot 1 + y \cdot 1$'.\\
 	PROD & Products.\\
@@
 \end{tabular}
-\caption{A subset of important internal \Maple{} structures. See~\parencite{MAPLE:ProgrammingGuide} for a complete list.}
+\caption{A subset of important internal \Maple{} data types. See~\parencite{MAPLE:ProgrammingGuide} for a complete list.}
 \label{tab:maple-types}
@@
 
-\subsection{Maple's Open Maple API}
-\Maple{} provides an \gls{api}, called \texttt{OpenMaple}~\parencite[\S 14.3]{MAPLE:ProgrammingGuide} for interaction with the \Maple{} kernel.
-The \texttt{OpenMaple} \gls{api} is implemented for different programming languages such as \texttt{C}, \texttt{Java} and \texttt{Visual Basic}.
-Since our forward translation and the \gls{mlp} is implemented in \texttt{Java}, we use the \texttt{Java} \gls{api} for the backward translation.
-The \texttt{Java} \gls{api} has some limitations to interact with \Maple's \gls{dag}.
+%\subsection{Maple's Open Maple API}
+The translator using the \texttt{OpenMaple}~\parencite[\S 14.3]{MAPLE:ProgrammingGuide} \gls*{api} for interacting with \Maple's kernel and get access to the \inertF{} of inputs.
 
-For the backward translation, we use the \inertF.
-This representation can be converted to a nested list version and the \texttt{OpenMaple} Java \gls{api} can handle lists.
+%The \texttt{OpenMaple} \gls*{api} is implemented for different programming languages such as \texttt{C}, \texttt{Java} and \texttt{Visual Basic}.
+Since our forward translation and the \gls*{mlp} is implemented in \texttt{Java}, we use the \texttt{Java} \gls*{api} for the backward translation.
+The \texttt{Java} \gls*{api} has some limitations to interact with \Maple's \gls*{dag}.
+
+%For the backward translation, we use the \inertF.
+This representation can be converted to a nested list version and the \texttt{OpenMaple} Java \gls*{api} can handle lists.
 Therefore, the backward translation firstly converts the \inertF{} to a nested list structure before any translation process starts.
 
-The following figure~\ref{fig:inertform-list} shows the \inertF{} and the nested list of the polynomial $x^2+x$.
+%The following figure~\ref{fig:inertform-list} shows the \inertF{} and the nested list of the polynomial $x^2+x$.
 
-\begin{figure}[ht]
-\centering
-\begin{minipage}{.45\linewidth}
-\centering
-\begin{tabular}{lll}
-\multicolumn{3}{l}{\_Inert\_SUM(}\\
-& \multicolumn{2}{l}{\_Inert\_POWER(}\\
-& & \_Inert\_NAME("x"),\\
-& & \_Inert\_INTPOS(2)\\
-& ),\\ 
-& \multicolumn{2}{l}{\_Inert\_NAME("x")}\\
-)
-\end{tabular}
-\end{minipage}
-\begin{minipage}{.45\linewidth}
-\centering
-\begin{tabular}{lll}
-\multicolumn{3}{l}{[\_Inert\_SUM,}\\
-& \multicolumn{2}{l}{[\_Inert\_POWER, }\\
-& & [\_Inert\_NAME, "x"],\\
-& & [\_Inert\_INTPOS, 2]\\
-& ],\\ 
-& \multicolumn{2}{l}{[\_Inert\_NAME, "x"]}\\
-]
-\end{tabular}
-\end{minipage}
-\caption{\inertF{} and the nested list representation of $x^2+x$.}
-\label{fig:inertform-list}
-\end{figure}
+%\begin{figure}[ht]
+%\centering
+%\begin{minipage}{.45\linewidth}
+%\centering
+%\begin{tabular}{lll}
+%\multicolumn{3}{l}{\_Inert\_SUM(}\\
+%& \multicolumn{2}{l}{\_Inert\_POWER(}\\
+%& & \_Inert\_NAME("x"),\\
+%& & \_Inert\_INTPOS(2)\\
+%& ),\\ 
+%& \multicolumn{2}{l}{\_Inert\_NAME("x")}\\
+%)
+%\end{tabular}
+%\end{minipage}
+%\begin{minipage}{.45\linewidth}
+%\centering
+%\begin{tabular}{lll}
+%\multicolumn{3}{l}{[\_Inert\_SUM,}\\
+%& \multicolumn{2}{l}{[\_Inert\_POWER, }\\
+%& & [\_Inert\_NAME, "x"],\\
+%& & [\_Inert\_INTPOS, 2]\\
+%& ],\\ 
+%& \multicolumn{2}{l}{[\_Inert\_NAME, "x"]}\\
+%]
+%\end{tabular}
+%\end{minipage}
+%\caption{\inertF{} and the nested list representation of $x^2+x$.}
+%\label{fig:inertform-list}
+%\end{figure}
 
-In the nested list, the first element specifies the type of the current list while the following arguments store the arguments and optional attributes.
+%In the nested list, the first element specifies the type of the current list while the following arguments store the arguments and optional attributes.
 
-\subsection{Workarounds for Problems}\label{subsec:maple-probs}
-Note that the backward translations usually do not have problems with ambiguous expressions.
-Every \gls{cas} needs to solve ambiguities in its own representation, otherwise it could not compute the expression.
+\subsection{Automatic Changes of Inputs in Maple}\label{subsec:maple-probs}
+%Note that the backward translations usually do not have problems with ambiguous expressions.
+Every \gls*{cas} needs to solve ambiguities in its own representation, otherwise it could not compute the expression.
 Our backward translation from \Maple{} back to semantic \LaTeX{} is therefore based on already disambiguated expressions.
 
-The general walkthrough of our backward translation starts with a string of a \texttt{1D} \Maple{} input expression.
+%The general walkthrough of our backward translation starts with a string of a \texttt{1D} \Maple{} input expression.
 As described above, this string is entered in \Maple{} to parse the expression and work on the nested list version of the \inertF{} afterwards.
@@
 Mostly because \Maple{} automatically evaluates input expressions immediately - even before we have a chance to take a look at the internal data structure of the input.
-For example, an input such as $\sin@{\frac{\cpi}{2}}$ is immediately evaluated to $1$ and the internal data structure just shows us the positive integer value $1$ instead of the trigonometric function.
+For example, an input such as $\sin\left(\frac{\cpi}{2}\right)$ is immediately evaluated to $1$ and the internal data structure just shows us the positive integer value $1$ instead of the trigonometric function.
 
-Furthermore, \Maple{} also changes input expressions slightly, mostly because of missing data types for a specific expression.
+%Furthermore, \Maple{} also changes input expressions slightly, mostly because of missing data types for a specific expression.
 For example, the internal data structure has no type to represent fractions except the \texttt{RATIONAL} type, which only allows integer values for the numerator and denominator.
@@
 
-Firstly, we summarize all known changes and evaluations \Maple{} automatically performs on input expressions.
+%Firstly, we summarize all known changes and evaluations \Maple{} automatically performs on input expressions.
+
+\Maple{} evaluates inputs automatically and changes the input into an internal representation.
+This internal representation might look a bit different to the input.
+One example has already been given with figure~\ref{internal-maple-dag}, where each summand of a sum is stored with a factor.
+Here is a list of all internal changes that occur for inputs.
 
@@
 \item \Maple{} evaluates input expressions immediately.
-\item There is no data type to represents square roots such as $\sqrt{x}$ (or $n$-th roots).
+\item There is no data type to represent square roots such as $\sqrt{x}$ (or $n$-th roots).
 Therefore, \Maple{} stores roots as an exponentiation with a fractional exponent.
@@
 
-Since we want to keep a translated expression similar to the input expression, we use unevaluation quotes during the whole translation process.
+Since we want to keep a translated expression similar to the input expression, we implemented some cosmetic rules for the backward translations that solves or reduce the effects from the list of changes above.
+\begin{itemize}
+\item We use unevaluation quotes to supress evaluations of the input.
+\item We perform reordering of factors and summands so that negative factors appears in front of the summand.
+This gives us the opportunity to translate $x - y$ to $x - y$ instead of $x + y \cdot (-1)$.
+\item We introduced the new internal data types \texttt{MYFLOAT} and \texttt{DIVIDE} to translate floats and fractions in more convenient notations.
+\end{itemize}
+
+%Since we want to keep a translated expression similar to the input expression, we use unevaluation quotes during the whole translation process.
 Unevaluation quotes also solve the problems with square roots and $n$-th roots.
@@
 
-Because of the absent data type for subtractions, a long sum with negative terms could be difficult to read.
+%Because of the absent data type for subtractions, a long sum with negative terms could be difficult to read.
 We change the order of constants (such as an integer value) to be in front of products.
@@
 
-Floating point numbers in scientific notation also causes a problem. Consider the input expression $0.41$ as the nested list \inertF{} representation
-\begin{equation}
-\texttt{[\_Inert\_FLOAT, [\_Inert\_INTPOS, 41], [\_Inert\_INTNEG, 2]]}.
-\end{equation}
+%Floating point numbers in scientific notation also causes a problem. Consider the input expression $0.41$ as the nested list \inertF{} representation
+%\begin{equation}
+%\texttt{[\_Inert\_FLOAT, [\_Inert\_INTPOS, 41], [\_Inert\_INTNEG, 2]]}.
+%\end{equation}
 
-The backward translator is implemented in \texttt{Java}.
+%The backward translator is implemented in \texttt{Java}.
 To translate such an expression back to $0.41$, computations become necessary.
-Of course, it makes sense to perform those computations in the \gls{cas} (in that case, in \Maple) rather than in the translation program.
-Therefore, we implemented a procedure\footnote{\textit{Procedures} in \Maple{} are small programs similar to methods and functions in \gls{oop} languages.} to automatically convert \texttt{FLOAT} nodes to a string representation of the floating point number.
+Of course, it makes sense to perform those computations in the \gls*{cas} (in that case, in \Maple) rather than in the translation program.
+Therefore, we implemented a procedure\footnote{\textit{Procedures} in \Maple{} are small programs similar to methods and functions in \gls*{oop} languages.} to automatically convert \texttt{FLOAT} nodes to a string representation of the floating point number.
 We created a new type, called \texttt{MYFLOAT} to represent such numbers.
@@
 
-For fractions, we use a similar approach and introduce a new internal data type \texttt{DIVIDE}.
+%For fractions, we use a similar approach and introduce a new internal data type \texttt{DIVIDE}.
 The implementation of \texttt{DIVIDE} was difficult.
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{p{2.6cm}cc}
-	\hline
-	& Negative numeric exponent & General negative exponent\\
-	\hline
-	\texttt{1D} \Maple{} input expression & $\texttt{(x-1)\^{}(-2)}$ & \rule{0pt}{0.9\normalbaselineskip}$\texttt{(x-1)\^{}(-2*y)}$\\
-	How \Maple{} displays the input & $\displaystyle\frac{1}{(x-1)^2}$ & $\displaystyle(x-1)^{-2y}$\\
-	\hline
-\end{tabular}
-\caption{Different styles to display negative exponents depending on the type of the exponent.
+%\begin{table}[ht]
+%\centering
+%\begin{tabular}{p{2.6cm}cc}
+%	\hline
+%	& Negative numeric exponent & General negative exponent\\
+%	\hline
+%	\texttt{1D} \Maple{} input expression & $\texttt{(x-1)\^{}(-2)}$ & \rule{0pt}{0.9\normalbaselineskip}$\texttt{(x-1)\^{}(-2*y)}$\\
+%	How \Maple{} displays the input & $\displaystyle\frac{1}{(x-1)^2}$ & $\displaystyle(x-1)^{-2y}$\\
+%	\hline
+%\end{tabular}
+%\caption{Different styles to display negative exponents depending on the type of the exponent.
 \Maple{} displays expressions with negative numeric exponents as fractions, while other negative exponents are not displayed as fractions.}
-\label{tab:maple-fracs}
-\end{table}
+%\label{tab:maple-fracs}
+%\end{table}
 
-We implement all of these changes before the actual translation process starts. This helps keeping the translated expression similar to the input expression.
+%We implement all of these changes before the actual translation process starts. This helps keeping the translated expression similar to the input expression.
 
-\subsection{Implementation Details}\label{subsec:maple-impl}
-\begin{wrapfigure}{l}{0.485\textwidth}
-	\vspace{-12pt}
-	\centering
-	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.39]{BackwardTranslator.pdf}
-	\caption{A scheme of the backward translator and its specialized subtranslators.}
-	\label{fig:BackwardDel}
-	\vspace{-13pt}
-\end{wrapfigure}
+%\subsection{Implementation Details}\label{subsec:maple-impl}
+%\begin{wrapfigure}{l}{0.485\textwidth}
+%	\vspace{-12pt}
+%	\centering
+%	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.39]{BackwardTranslator.pdf}
+%	\caption{A scheme of the backward translator and its specialized subtranslators.}
+%	\label{fig:BackwardDel}
+%	\vspace{-13pt}
+%\end{wrapfigure}
 
-The general translation process runs through different states.
+%The general translation process runs through different states.
 Each translation process starts with the string representation of a \texttt{1D} \Maple{} input.
 This string will be converted to the internal nested list representation of our custom \inertF{} (see the previous subsection~\ref{subsec:maple-probs}).
-Since the \inertF{} is similar to an expression tree, it can be translated as the \gls{mlp-pt} in the forward translation.
+Since the \inertF{} is similar to an expression tree, it can be translated as the \gls*{mlp-pt} in the forward translation.
 Hence, we use the same approach of multiple subtranslators for the backward translation process.
 
-Figure~\ref{fig:BackwardDel} shows the subtranslators of the backward translation process.
+%Figure~\ref{fig:BackwardDel} shows the subtranslators of the backward translation process.
 The structure is the same as for the forward translation, seen in figure~\ref{fig:ForwardDel}.
@@
 
-The following figure~\ref{fig:backward-trans} illustrates the backward translation process for the Jacobi polynomial example $P_n^{(\alpha , \beta)}(\cos(a\Theta))$ from table~\ref{tab:JacobiP-usecase}.
+%The following figure~\ref{fig:backward-trans} illustrates the backward translation process for the Jacobi polynomial example $P_n^{(\alpha , \beta)}(\cos(a\Theta))$ from table~\ref{tab:JacobiP-usecase}.
 The input expression is converted to the nested list representation of the \texttt{MyInertForm} (the customized \inertF{} of \Maple).
@@
 The last step is used to put all translated arguments to the right position (visualized by red arrows).
+
+The translation process than follows the same principle as for the forward translations.
+Since the syntax tree of \Maple{} is an expression tree, we do not need to implement special reordering or grouping algorithms to perform backward translations.
+Translations for functions are also realized via patterns and placeholders.
+Figure~\ref{fig:backward-trans} illustrates the backward translation process for the Jacobi polynomial example from table~\ref{tab:JacobiP-usecase}.
 
@@
 \end{figure}
-
-\cleardoublepage

### 2018-07-24 chapters/03-translator/forward-trans.tex
Update until forward/backword translation
Finalize first version of the journal paper Still 1.000 words too long

@@
-\section{The Forward Translation}\label{sec:forward-translation}
-%\glsreset{teo}
-The translation process starts with semantic \LaTeX{} expressions.
-These expressions get analyzed and parsed by the \gls{mlp}, which produces a \gls{mlp-pt} of the input expression with additional information about each symbol.
-An abstract translator class analyzes each node and delegates them to specialized subtranslators.
-In this process an object called \gls{teo} is build.
-This \gls{teo} is needed to rebuild a string representation of the mathematical expression given by the \gls{mlp-pt}.
+\section{Forward Translations}\label{sec:forward-translation}
+An abstract translator class analyzes each node of the parsed tree and delegates them to specialized subtranslators.
+In this process an object called \gls*{teo} is build.
+This \gls*{teo} is needed to rebuild a string representation of the mathematical expression after all nodes of the parsed tree were translated.
+
+The parsed tree generated by the \gls*{pom}-tagger is not a mathematical expression tree.
+The \gls*{pom} project aims to disambiguate mathematical \LaTeX{} expressions and generates an expression tree.
+However, in the current state, many expressions cannot be disambiguated yet.
+In consequence, the \gls*{pom}-tagger generates a raw parsed tree where each token in the \LaTeX{} expression is a node in the tree.
+We call this parsed tree the \gls*{pom-pt}.
 
@@
 These tables are converted by the \verb|lexicon-creator.jar| to the \verb|DLMF-macros-lexicon.txt| lexicon file.
-Together with the \verb|global-lexicon.txt| file, the \gls{mlp-pt} will be created by the \gls{mlp}.
-The \verb|latex-converter.jar| takes a string representation of a semantic \LaTeX{} expression and uses the \gls{mlp} as well as our \verb|Translator| to create an appropriate string representation for a specified \gls{cas}.
-The following \cref{subsec:analyze-mlp,subsec:delegation,subsec:translated-expr,subsec:add-info} will focus on the \verb|Translator|.
+Together with the \verb|global-lexicon.txt| file, the \gls*{pom-pt} will be created by the \gls*{pom}-tagger.
+The \verb|latex-converter.jar| takes a string representation of a semantic \LaTeX{} expression and uses the \gls*{pom} engine as well as our \verb|Translator| to create an appropriate string representation for a specified \gls*{cas}.
 
@@
 	\includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm, scale=0.72]{ForwardTranslationProcess.pdf}
-	\caption{Process diagram of a forward translation process. The MLP generates the PT based on lexicon and JSON files. The PT will be translated to different CAS.}
+	\caption{Process diagram of a forward translation process.
+The \gls*{pom}-tagger generates the \gls*{pom-pt} based on lexicon and \gls*{json} files.
+The \gls*{pom-pt} will be translated to different \gls*{cas}.}
 	\label{fig:forward-trans}
@@
 
-\subsection{Analyzing the MLP-Parse Tree}\label{subsec:analyze-mlp}
-The main task is to analyze the \gls{mlp-pt}. Remember that the \gls{mlp-pt} is different to expression trees as described in section~\ref{subsec:pom-mlp}.
-
-Since the \gls{bnf} does not define rules for semantic macros, each argument of the semantic macro and each $@$ symbol are following siblings of the semantic macro node.
+\subsection{Analyzing the PoM-Parsed Tree}\label{subsec:analyze-mlp}
+Since the \gls*{bnf} does not define rules for semantic macros, each argument of the semantic macro and each $@$ symbol are following siblings of the semantic macro node.
 That is the reason, why we stored the number of parameters, variables and $@$ symbols in the lexicon files.
-Otherwise, the translator could not find the end of a semantic macro in the \gls{mlp-pt}.
-
-Figure~\ref{fig:syntax-tree-usecase} visualizes the \gls{mlp-pt} of the Jacobi polynomial example from table~\ref{tab:JacobiP-usecase}.
-Because of these differences to expression trees, a backward conversion of the \gls{mlp-pt} to a string representation can be difficult, especially for finding necessary or unnecessary parentheses.
-Therefore we create the \gls{teo}.
-Of course, the \gls{teo} is deeply incorporated into the translation process, but in this subsection we will focus on the general idea and process of the translation.
-We will focus on the usage of \gls{teo} in subsection~\ref{subsec:translated-expr}.
+Otherwise, the translator could not find the end of a semantic macro in the \gls*{pom-pt}.
 
@@
 	\includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm, scale=0.75]{SyntaxTreeUseCase.pdf}
-	\caption{MLP-PT for a Jacobi polynomial using the DLMF/DRMF \LaTeX{} macro. Each leaf contains information from the lexicon files.}
+	\caption{pom-pt for a Jacobi polynomial using the DLMF/DRMF \LaTeX{} macro. Each leaf contains information from the lexicon files.}
 	\label{fig:syntax-tree-usecase}
+	\vspace{-10pt}
 \end{figure}
-
-The general idea of a translation of a \gls{mlp-pt} is a simple recursive algorithm, such as algorithm~\ref{alg:simple-translation}.
-Whenever the algorithm finds a leaf, it can translate this single term.
-If the node is not a leaf, it starts to translate all children of the node recursively.
 
@@
 \begin{algorithm}[H]
-\caption{Simple translation algorithm for the MLP-Parse Tree}\label{alg:simple-translation}
+\caption{Simple translation algorithm for the POM-Parse Tree}\label{alg:simple-translation}
 	\begin{algorithmic}[1]
@@
 
-But this approach is not completely feasible, because sometimes the algorithm needs to look ahead and check the following siblings for a valid translation, for example in case of a semantic macro with arguments (see figure~\ref{fig:syntax-tree-usecase}).
-Therefore, the {\footnotesize \verb|TRANSLATE_LEAF|} function also needs to know the following siblings of the current node.
-To avoid multiple translations of the same node, a list representation would be a better solution.
-Algorithm~\ref{alg:translation} explains an abstract version of the final algorithm for our translation tool.
+Figure~\ref{fig:syntax-tree-usecase} visualizes the \gls*{pom-pt} of the Jacobi polynomial example from table~\ref{tab:JacobiP-usecase}.
+Because of the differences to expression trees, a backward conversion of the \gls*{pom-pt} to a string representation can be difficult, especially for finding necessary or unnecessary parentheses.
+Therefore we create the \gls*{teo}.
 
-If the root $r$ is a leaf, it can be translated as a leaf.
-Eventually, some of the following siblings are needed in order to translate $r$ correctly.
-If $r$ is not a leaf, it contains one or more children.
-Therefore, we can call the {\footnotesize \verb|ABSTRACT_TRANSLATOR|} recursively for the children.
-Once we have translated $r$, we can go a step further and translate the next node.
-Line~\ref{line:ifNotEmpty} checks if there are following siblings left and calls the {\footnotesize \verb|ABSTRACT_TRANSLATOR|} recursively in that case.
+With these tools, we can translate a \LaTeX{} expression by translating the \gls*{pom-pt} node by node and perform group or reordering operations for some special cases.
+The algorithm is realized in a simple recursive structure.
+The algorithm~\ref{alg:simple-translation} explains the translation process.
+Whenever the algorithm finds a leaf, it can translate this single term.
+If the node is not a leaf, it starts to translate all children of the node recursively.
 
-\begin{algorithm}[ht]
+\begin{algorithm}[H]
 \caption{Abstract translation algorithm to translate MLP-Parse trees.}\label{alg:translation}
 	\begin{algorithmic}[1]
-	\Require Root $r$ of a MLP-Parse tree $T$. List \textit{following\_siblings} with the following siblings of $r$. The list can be empty.
+	\Require Root $r$ of a POM-Parse tree $T$. List \textit{following\_siblings} with the following siblings of $r$. The list can be empty.
 	\Procedure{abstract\_translator}{$r$, \textit{following\_siblings}}
 	\If{$r$ is leaf}
-		\State {\scriptsize TRANSLATE\_LEAF}($r$, \textit{following\_siblings\_of\_r});\label{line:transLeaf}
+		\State {\scriptsize TRANSLATE\_LEAF}($r$, \textit{following\_siblings});\label{line:transLeaf}
 	\Else
@@
 \end{algorithm}
+\vspace{-20px}
 
-With this approach we are able to translate most of the semantic \LaTeX{} expressions.
-To take a look at the problems with this approach, we need to think about the operators and their standard notations in mathematics.
-There are many types of notation used to represent formulae.
-For example, the \gls{pol}\footnote{Also known as \textit{Warsaw Notation} or \textit{prefix notation}} (hereafter called prefix notation) places the operator to the left of/before its operands.
-The \gls{rpol}\footnote{Also known as \textit{postfix notation}} (hereafter called postfix notation) does the opposite and places the operator to the right of/after its operands.
+This approach is not completely feasible, since the algorithm needs to look ahead and check the following siblings in some cases, e.g., in the case of a semantic macro with arguments (see figure~\ref{fig:syntax-tree-usecase}).
+The {\footnotesize \verb|TRANSLATE_LEAF|} function needs to know the following siblings of the current node.
+Algorithm~\ref{alg:translation} explains a more detailed but abstract version of the final algorithm of the presented translation tool.
+
+If the root $r$ is a leaf, it still can be translated as a leaf.
+Eventually, some of the following siblings are needed in order to translate $r$.
+The list of \textit{following\_siblings} in line~\ref{line:transLeaf} might be reduced to avoid multiple translations for one node.
+If $r$ is not a leaf, it contains one or more children.
+Therefore, we can call the {\footnotesize \verb|ABSTRACT_TRANSLATOR|} recursively for the children.
+Once we have translated $r$, we can go a step further and translate the next node.
+Line~\ref{line:ifNotEmpty} checks if there are following siblings left and calls the {\footnotesize \verb|ABSTRACT_TRANSLATOR|} recursively in that case.
+Translated expressions are stored by the \gls*{teo} object.
+The detailed translation process is not included in the algorithm.
+
+Algorithm~\ref{alg:translation} is a simplified version of the translator process.
+The main progress happens in \cref{line:transLeaf,line:transSeq}.
+There are several cases what $r$ can be.
+Table~\ref{tab:allTypesTable} on page \pageref{tab:allTypesTable} gives an overview of all different types in a \gls*{pom-pt}.
+A more detailed explanation of the types can be found in~\parencite{POM-Tagger}.
+
+The \gls*{bnf} grammar defines some basic grammatical rules for generic \LaTeX{} macros, such as for \verb|\frac|, \verb|\sqrt|.
+Therefore, there is a hierarchical structure for those symbols, similar to the structure in expression trees.
+As already mentioned, some of these types can be translated directly, such as Greek letters, while others are more complex, such as semantic \LaTeX{} macros.
+Therefore, the translators delegates the translation to specialized subtranslators.
+This delegation process is implemented in \cref{line:transLeaf,line:transSeq} of algorithm~\ref{alg:translation}.
+The Subsection~\ref{sec:subtranslators} discusses these classes in more detail.
+
+This approach seems to work for most of semantic \LaTeX{} expressions.
+However, there is one problem with this straight forward approach which was mentioned shortly mentioned in section~\ref{sec:problems}.
+There are many different types of notations used to represent formulae.
+Table~\ref{tab:notations} illustrates the simple expression $(a+b)x$ in different notations.
+The \gls*{pol}\footnote{Also known as \textit{Warsaw Notation} or \textit{prefix notation}} (hereafter called prefix notation) places the operator to the left of/before its operands.
+The \gls*{rpol}\footnote{Also known as \textit{postfix notation}} (hereafter called postfix notation) does the opposite and places the operator to the right of/after its operands.
 The infix notation is commonly used in arithmetic and places the operator between their operands, which only makes sense as long as the operator is a binary operator.
-For example, table~\ref{tab:notations} shows an expression in different notations.
 
-\begin{table}[!ht]
-\centering
+\begin{wrapfigure}{r}{0.4\textwidth}
+\vspace{-22pt}
+\begin{minipage}{0.4\textwidth}
+\center
 \begin{tabular}{lc}
@@
 \end{tabular}
+\vspace{-5pt}
 \caption{The mathematical expression '$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
 \label{tab:notations}
-\end{table}
+\end{minipage}
+\vspace{-15pt}
+\end{wrapfigure}
 
 In mathematical expressions, notations are mostly mixed, depending on the case and number of operands.
-For example, infix notation is common for binary operators ($+$, $-$, $\cdot$, $\mod$ etc.), while functional notations are obviously used for any kind of functions ($sin$, $cos$, etc.).
-Sometimes the same symbol is used in different notations to identify a different meaning.
+For example, infix notation is common for binary operators ($+$, $-$, $\cdot$, $\mod$ etc.), while functional notations are conveniently used for any kind of functions ($sin$, $cos$, etc.).
+Sometimes the same symbol is used in different notations to distinguish different meanings.
 For example, the '$-$' as an unary operator is used in prefix notation to indicate the negative value of its operand, such as in '$-2$'.
@@
 
-Most programming languages (and \gls{cas} as well) internally use prefix or postfix notation and do not mix the notations in one expression, because it is easier to parse those notations.
+\input{chapters/03-translator/all-types-table.tex}
+
+Most programming languages (and \gls*{cas} as well) internally use prefix or postfix notation and do not mix the notations in one expression, since it more convenient to parse expressions in uniform notations.
 However, the common practice in science is to use mixed notations in expressions.
-Since the \gls{mlp} has barely implemented mathematical grammatical rules, it takes the input as it is and does not build an expression tree.
-Therefore, it parses all four examples from table~\ref{tab:notations} to four different \gls{mlp-pt}s rather than to one unique expression tree.
-In general, this is not a problem for our translation process because most \gls{cas} are familiar with the most common notations.
+Since the \gls*{pom} has rarely implemented mathematical grammatical rules yet, it takes the input as it is and does not build an expression tree.
+Therefore, it parses all four examples from table~\ref{tab:notations} to four different \gls*{pom-pt}s rather than to one unique expression tree.
+In general, this is not a problem for our translation process since most \gls*{cas} are familiar with the most common notations.
 Therefore, the translator does not need to know that $a$ and $b$ are the operands of the binary operator '$+$' in '$a+b$'.
-We could just translate the symbols in '$a+b$' in the same order as they appear in the expression and the \gls{cas} would understand it.
+The translator could simply translates the symbols in '$a+b$' in the same order as they appear in the expression and the \gls*{cas} would understand it.
 However, this simple approach generates two problems.
@@
 \item \label{prob:1} The translated expression is only syntactically correct if the input expression was syntactically correct.
-\item \label{prob:2} We cannot translate expressions to a \gls{cas} which uses a different notation.
+\item \label{prob:2} We cannot translate expressions to a \gls*{cas} which uses a non-standard notation.
 \end{enumerate}
@@
 Problem~\ref{prob:2} is more difficult to solve.
-If a user wants to support a \gls{cas} that uses prefix or postfix notation by default, the whole translator would fail in its current state.
-Supporting \gls{cas} with another notation would be a part of future work and will be mentioned in chapter~\ref{ch:conc-future-work}.
+If a user wants to support a \gls*{cas} that uses prefix or postfix notation by default, the translator would fail in its current state.
+Supporting \gls*{cas} with another notations would be a part of future work.
 
@@
 For example, the interpretations $(2!)!$ and $(2)!!$ can be distinguished in \Maple{} by using \verb|factorial(factorial(2))| and \verb|doublefactorial(2)| respectively.
+We define the translations as follows:
 
-As already mentioned, the structure of the \gls{mlp-pt} makes it difficult to change the notation for all kind of operators.
-Therefore, and especially because of the examples above, we change the notation during the forward translation process to the functional notation only for the factorial and double factorial function.
-Thus,
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-The translator needs to presume some properties for the functions similar to \LaTeX, which only recognizes the very next element right after the caret for the superscript.
+%The translator needs to presume some properties for the functions similar to \LaTeX, which only recognizes the very next element right after the caret for the superscript.
 
-The biggest problem about translating the factorial (or double factorial) function is the common postfix notation for them.
-In this stage, algorithm~\ref{alg:translation} only translates the current node and if necessary the following siblings, but not the preceding siblings.
+Algorithm~\ref{alg:translation} does not allow this translation right now.
+It has no access to previously translated nodes in its current state.
+This problem is solved by the \gls*{teo} that stores and groups translated objects as lists.
+This allows to access the latest translated expression and use it as the argument for the factorial function.
+Table~\ref{tab:teo-list} shows three example for the \gls*{teo} list that groups some tokens.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{cc}
+	\hline
+	Input Expression & TEO List\\
+	\hline
+	$a+b$ & \verb|[a, +, b]|\\
+	$(a+b)$ & \verb|[(a+b)]|\\
+	$\frac{a}{b}-2$ & \verb|[(a)/(b), -, 2]|\\
+	\hline
+\end{tabular}
+\caption{How the TEO-list groups subexpressions.}
+\label{tab:teo-list}
+\end{table}
+
+%The biggest problem about translating the factorial (or double factorial) function is the common postfix notation for them.
+In this stage, algorithm~\ref{alg:translation} only translates the current node and if necessary the following siblings, but not the preceding siblings (see algorithm~\ref{alg:translation}).
 Consider an expression such as '$4!$'.
-When the variable $r$ reaches the exclamation mark, the argument '$4$' of the function has already been translated.
-The current version of the algorithm has no access to this previously translated expression.
-Even more complicated, would be the case of the double factorial '$4!!$'.
-The following two approaches would solve this problem.
+When the variable node $r$ reaches the exclamation mark, the argument '$4$' of the function has already been translated and is no longer accessible for the current algorithm.
+This problem occurs whenever a postfix notation is involved.
+The following two approaches would solve the problem for the problem.
 
-\begin{enumerate}
-\item\label{appr:1} The translator always checks if the next two following siblings are one or two exclamation marks.
-\item\label{appr:2} The current state of program has access to the previously translated expressions and can modify them retrospectively.
-\end{enumerate}
+%\begin{enumerate}
+%\item\label{appr:1} The translator always checks if the next two following siblings are one or two exclamation marks.
+%\item\label{appr:2} The current state of program has access to the previously translated expressions and can modify them retrospectively.
+%\end{enumerate}
 
-Approach~\ref{appr:1} would be unnecessarily inefficient.
+%Approach~\ref{appr:1} would be unnecessarily inefficient and does not scale well if more of these cases occur.
 Therefore we implement approach~\ref{appr:2}.
-That is another reason why we implement an extra object to organize translated expressions, the \gls{teo}.
+That is another reason why we implement an extra object to organize translated expressions, the \gls*{teo}.
 This object stores all parts of the previously translated expressions.
 The translator has access to these parts and is able to modify them afterwards.
-A more detailed explanation about the \gls{teo} follows in subsection~\ref{subsec:translated-expr}.
-\newpage
 
-\input{chapters/03-translator/all-types-table.tex}
-\newpage
+%\input{chapters/03-translator/all-types-table.tex}
 
-\subsection{Delegation of Translations}\label{subsec:delegation}
+%\subsection{Delegation of Translations}\label{subsec:delegation}
 
-\begin{wrapfigure}{l}{0.45\textwidth}
-	\vspace{-20pt}
-	\centering
-	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.4]{ForwardTranslator.pdf}
-	\caption{A scheme of the forward translator and its specialized subtranslators.}
-	\label{fig:ForwardDel}
-	\vspace{-25pt}
-\end{wrapfigure}
+%\begin{wrapfigure}{l}{0.45\textwidth}
+%	\vspace{-20pt}
+%	\centering
+%	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.4]{ForwardTranslator.pdf}
+%	\caption{A scheme of the forward translator and its specialized subtranslators.}
+%	\label{fig:ForwardDel}
+%	\vspace{-25pt}
+%\end{wrapfigure}
 
-Algorithm~\ref{alg:translation} is a simplified version of the translator process.
+%Algorithm~\ref{alg:translation} is a simplified version of the translator process.
 The main progress happens in \cref{line:transLeaf,line:transSeq}.
 There are several cases what $r$ can be.
-Table~\ref{tab:allTypesTable} on page \pageref{tab:allTypesTable} gives an overview of all different types in a \gls{mlp-pt}.
+Table~\ref{tab:allTypesTable} on page \pageref{tab:allTypesTable} gives an overview of all different types in a \gls*{pom-pt}.
 A more detailed explanation of the types can be found in~\parencite{POM-Tagger}.
 
-The \gls{bnf} grammar defines some basic grammatical rules for generic \LaTeX{} macros, such as for \verb|\frac|, \verb|\sqrt|.
+%The \gls*{bnf} grammar defines some basic grammatical rules for generic \LaTeX{} macros, such as for \verb|\frac|, \verb|\sqrt|.
 Therefore, there is a hierarchical structure for those symbols, similar to the structure in expression trees.
@@
 
-Figure~\ref{fig:ForwardDel} is a scheme of the forward translator and its subtranslators.
+%Figure~\ref{fig:ForwardDel} is a scheme of the forward translator and its subtranslators.
 The entry point for each translation process of the program is the \verb|SemanticLatexTranslator|.
@@
 
-\subsubsection{Subtranslators}
-The most interesting translators are the \verb|MacroTranslator| (translating semantic \LaTeX{} macros) and the \verb|SequenceTranslator| (translating all kinds of \textit{sequences}).
+\subsection{Subtranslators}\label{sec:subtranslators}
+%The most interesting translators are the \verb|MacroTranslator| (translating semantic \LaTeX{} macros) and the \verb|SequenceTranslator| (translating all kinds of \textit{sequences}).
 
-The \verb|SequenceTranslator| translates the \textit{sequence} and \textit{balanced expressions} in the \gls{mlp-pt}.
+A \verb|SequenceTranslator| translates the \textit{sequence} and \textit{balanced expressions} in the \gls*{pom-pt}.
 If a node $n$ is a leaf and the represented symbol an open bracket (parentheses, square brackets and so on) the following nodes are also taken as a \textit{sequence}.
@@
 On the other hand, this is a problem for interval expressions such as '$[a,b)$'.
-In the current version, the program cannot distinguish between mismatched parentheses and an half-opened, half-closed interval.
+In the current version, the program cannot distinguish between mismatched parentheses and half-opened, half-closed intervals.
 Whether an expression is an interval or another expression is difficult to decide and can depend on the context.
@@
 
-Another problem that the \verb|SequenceTranslator| solves is the position of multiplication signs in an expression.
+The \verb|SequenceTranslator| also handle positions of multiplication symbols.
 There are a couple of obvious choices to translate multiplication signs.
-For example, using \verb|\cdot| or \verb|\idot| will be obviously translated as a multiplication symbol.
-However, \verb|\idot| is fairly new and therefore not frequently used yet.
-Furthermore, the most common symbol for multiplications is still the white space or none symbol at all, as explained above.
+The most common symbol for multiplications is still the white space or none symbol at all, as explained previously.
 Consider the simple expression $2n\pi$.
-The \gls{mlp-pt} generates a sequence node with three children, namely $2$, $n$ and $\pi$.
+The \gls*{pom-pt} generates a sequence node with three children, namely $2$, $n$ and $\pi$.
 This sequence should be interpreted as a multiplication of the three elements.
-Therefore, the \verb|SequenceTranslator| checks the types of the current and next nodes in the tree to decide if there should be a multiplication symbol or not.
+The \verb|SequenceTranslator| checks the types of the current and next nodes in the tree to decide if it should add a multiplication symbol or not.
 For example, if the current or next node is an operator, a relation symbol or an ellipsis there will be no multiplication symbol added.
@@
 But we do not consider this translation to be wrong, because there is a semantic macro to represent functions.
-In this case, the user should use \verb|\f{f}@{x}| instead of \verb|f(x)|.
+In this case, the user should use \verb|\f{f}@{x}| instead of \verb|f(x)| to distinguish between $f$ as a function call and $f$ as a symbol.
 
-Besides the translation of sequences, the translation process of the \Macro s can be complex too; not only because of the structure of the \gls{mlp-pt}, but also because of the complex definition of the semantic macros.
+%Besides the translation of sequences, the translation process of the \Macro s can be complex too; not only because of the structure of the \gls*{pom-pt}, but also because of the complex definition of the semantic macros.
 As explained in \ref{subsec:macros}, they can contain optional parameters, exponents are allowed before and after the arguments and the number of $@$ symbols varies.
@@
 
-Algorithm~\ref{alg:macro-translation} is the translation function of the \verb|MacroTranslator| without error handling.
-It analyzes the next element right after the macro first.
-As mentioned, this can be one of the following:
-\begin{itemize}
-\item an exponent such as '\verb|^2|'.
-\item an optional parameter in square brackets.
-\item a parameter in curly brackets (a \textit{sequence} node in the \gls{mlp-pt}) if none of the above.
-\item an $@$ symbols if none of the above.
-\item a variable in curly brackets (a \textit{sequence} node) if none of the above.
-\end{itemize}
 \begin{algorithm}[!ht]
@@
 
-	\While{\textit{next} is $[$}\label{line:next_optional} \Comment{square brackets indicate optional arguments.}
+	\While{\textit{next} is $[$}\label{line:next_optional} \Comment{square brackets starts a balanced sequence of optional arguments.}
 		\State \textit{optional} = {\scriptsize TRANSLATE\_UNTIL\_CLOSED\_BRACKET}(\textit{args});
@@
 
-	\State \textit{argList}.add( {\scriptsize TRANSLATE\_PARAMETERS}(\textit{args}, \textit{info}) ); \label{line:trans_paras}
-	\State {\scriptsize SKIP\_AT\_SIGNS}( \textit{args}, \textit{info} ); \label{line:skip_ats}
-	\State \textit{argList}.add( {\scriptsize TRANSLATE\_VARIABLES}(\textit{args}, \textit{info}) ); \label{line:trans_vars}
+	\State \textit{argList}.add( {\scriptsize TRANSLATE\_PARAMETERS}(\textit{args}, \textit{info}) ); \label{line:trans_paras} \Comment{number is given in \textit{info}.}
+	\State {\scriptsize SKIP\_AT\_SIGNS}( \textit{args}, \textit{info} );  \label{line:skip_ats} \Comment{number is given in \textit{info}.}
+	\State \textit{argList}.add( {\scriptsize TRANSLATE\_VARIABLES}(\textit{args}, \textit{info}) ); \label{line:trans_vars} \Comment{number is given in \textit{info}.}
 
@@
 
-An exponent will be translated and shifted to the end of the translated semantic macro, since it is more common to write the exponent after the arguments of a function in \gls{cas}.
+Algorithm~\ref{alg:macro-translation} explains the \verb|MacroTranslator| without error handling.
+It extracted necessary information from the \gls*{pom-pt}, such as how many arguments this function has.
+In consequence it also processes the following siblings to translate the arguments.
+There are some special cases about the next node right after the macro occurs.
+This node can be
+\begin{itemize}
+\item an exponent, such as for '\verb|^2|';
+\item an optional parameter in square brackets;
+\item a parameter in curly brackets (a \textit{sequence} node in the \gls*{pom-pt}) if none of the above;
+\item an $@$ symbols if none of the above;
+\item or a variable in curly brackets (a \textit{sequence} node) if none of the above.
+\end{itemize}
+
+An exponent will be translated and shifted to the end of the translated semantic macro, since it is more common to write the exponent after the arguments of a function in \gls*{cas}.
 Therefore, the function translates and stores the exponent in line~\ref{line:next_caret}.
@@
 
-As you can see, the input expression in semantic \LaTeX{} is interpreted as~\eqref{eq:multi-expo}.
+As you can see, the input expression in semantic \LaTeX{} is interpreted as
 \begin{equation}\label{eq:multi-expo}
-\left( \cos(x)^2 \right)^2
+\left( \cos(x)^2 \right)^2.
 \end{equation}
 
-An open square bracket right after the semantic macro (after the exponent respectively) is an optional parameter.
-Since expressions in square brackets are not considered to be a \textit{sequence} in the \gls{mlp-pt}, the optional parameter is not grouped in a single node.
+%An open square bracket right after the semantic macro (after the exponent respectively) is an optional parameter.
+Since expressions in square brackets are not considered to be a \textit{sequence} in the \gls*{pom-pt}, the optional parameter is not grouped in a single node.
 Therefore, every node after the opening square bracket forms an expression for just one optional parameter until the algorithm reaches the closing square bracket.
@@
 
-The \cref{line:trans_paras,line:skip_ats,line:trans_vars} translates the parameters and variables.
+%The \cref{line:trans_paras,line:skip_ats,line:trans_vars} translates the parameters and variables.
 The $@$ symbols will be skipped.
@@
 In addition, if there is an exponent right after the semantic macro, it will be attached to the translated semantic macro (see table~\ref{tab:multi-expo}).
-The parameters and variables are always grouped by curly brackets in \textit{sequence}-nodes in the \gls{mlp-pt}.
-Therefore, every argument after the optional parameters and potential exponent cannot be longer than one node in the \gls{mlp-pt}.
+The parameters and variables are always grouped by curly brackets in \textit{sequence}-nodes in the \gls*{pom-pt}.
+Therefore, every argument after the optional parameters and potential exponent cannot be longer than one node in the \gls*{pom-pt}.
 
-\subsubsection{Handle Ambiguities}\label{subsec:ambiguities}
-Our semantic \LaTeX{} expressions are not full semantic with respect to possible ambiguities in the expression.
+%\subsection{Handle Ambiguities}\label{subsec:ambiguities}
+%Our semantic \LaTeX{} expressions are not full semantic with respect to possible ambiguities in the expression.
 We already mentioned some techniques to solve ambiguities in semantic \LaTeX{} expressions, for example how to handle double factorials.
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{cc}
-	\hline
-	Ambiguous Input & \LaTeX{} Output\\
-	\hline
-	\verb|n^m!| & $n^m!$\\
-	\verb|a^bc^d| & $a^bc^d$\\
-	\verb|x^y^z| & Double superscript error\\
-	\verb|x_y_z| & Double subscript error\\
-	\hline
-\end{tabular}
-\caption{Ambiguous \LaTeX{} expressions and how \LaTeX{} displays them.}
-\label{tab:amb-latex}
-\end{table}
+%\begin{table}[ht]
+%\centering
+%\begin{tabular}{cc}
+%	\hline
+%	Ambiguous Input & \LaTeX{} Output\\
+%	\hline
+%	\verb|n^m!| & $n^m!$\\
+%	\verb|a^bc^d| & $a^bc^d$\\
+%	\verb|x^y^z| & Double superscript error\\
+%	\verb|x_y_z| & Double subscript error\\
+%	\hline
+%\end{tabular}
+%\caption{Ambiguous \LaTeX{} expressions and how \LaTeX{} displays them.}
+%\label{tab:amb-latex}
+%\end{table}
 
-Another more questionable translation decision are alphanumerical expressions.
-As explained in table~\ref{tab:allTypesTable}, the \gls{mlp} handles strings of letters and numbers differently, depending on the order of the symbols.
+%Another more questionable translation decision are alphanumerical expressions.
+As explained in table~\ref{tab:allTypesTable}, the \gls*{mlp} handles strings of letters and numbers differently, depending on the order of the symbols.
 The reason is, that an expression such as '$4b$' is usually considered to be a multiplication of $4$ and '$b$', while '$b4$' looks like indexing '$b$' by $4$.
-While the first example produces two nodes, namely $4$ and '$b$', the second example '$b4$' produces just a single alphanumerical node in the \gls{mlp-pt}.
+While the first example produces two nodes, namely $4$ and '$b$', the second example '$b4$' produces just a single alphanumerical node in the \gls*{pom-pt}.
 The first expression would be translated as a multiplication of $4$ and '$b$'.
@@
 
-Another, more elegant way to solve this problem would be to use the newly created semantic macro \verb|\idot| (see subsection~\ref{subsec:semanticVSmacro}) to define an invisible multiplication symbol.
-However, because \verb|\idot| is such a new semantic macro, it is not used in the \gls{dlmf} or \gls{drmf} quite often yet.
+%Another, more elegant way to solve this problem would be to use the newly created semantic macro \verb|\idot| (see subsection~\ref{subsec:semanticVSmacro}) to define an invisible multiplication symbol.
+However, because \verb|\idot| is such a new semantic macro, it is not used in the \gls*{dlmf} or \gls*{drmf} quite often yet.
 Therefore, our program still interprets alphanumerical expressions as multiplications.
 
-Other ambiguities can appear, when the input expression contains a symbol, which is typically used to represent a specific mathematical object, but the user did not use the semantic macro to represent it.
+%Other ambiguities can appear, when the input expression contains a symbol, which is typically used to represent a specific mathematical object, but the user did not use the semantic macro to represent it.
 For example, the input expressions '$3i$' contains an '$i$', which is usually associated with the imaginary unit.
@@
 
-In general the translator is drafted to solve ambiguous expressions or automatically find a workaround to disambiguate the expression.
+%In general the translator is drafted to solve ambiguous expressions or automatically find a workaround to disambiguate the expression.
 Only if there is no way to solve the ambiguity with the defined rules, the translation process stops.
 
-\subsection{Translated Expression Objects}\label{subsec:translated-expr}
-\glsreset{teo}
-\begin{wrapfigure}{r}{0.5\textwidth}
-	\vspace{-20pt}
-	\centering
-	\includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm, scale=1]{SyntaxTreeGroup.pdf}
-	\caption{The MLP-PT for '$\cos(2n\pi)!$' and the grouped argument of the factorial function.}
-	\label{fig:syntax-group}
-	\vspace{-15pt}
-\end{wrapfigure}
+%\subsection{Translated Expression Objects}\label{subsec:translated-expr}
+%\glsreset{teo}
+%\begin{wrapfigure}{r}{0.5\textwidth}
+%	\vspace{-20pt}
+%	\centering
+%	\includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm, scale=1]{SyntaxTreeGroup.pdf}
+%	\caption{The pom-pt for '$\cos(2n\pi)!$' and the grouped argument of the factorial function.}
+%	\label{fig:syntax-group}
+%	\vspace{-15pt}
+%\end{wrapfigure}
 
-It is sometimes necessary to provide access to already translated symbols or subexpressions.
+%It is sometimes necessary to provide access to already translated symbols or subexpressions.
 A typical example is the (double) factorial function, where the argument was translated before the function was registered by the translator (postfix notation).
-To realize this, the \gls{teo} is implemented.
+To realize this, the \gls*{teo} is implemented.
 It is mainly a list object of translated expressions.
@@
 
-Figure~\ref{fig:syntax-group} shows an example for the expression '$\cos(2n\pi)!$'.
-If we just analyze the \gls{mlp-pt}, the previous node of the exclamation mark (which indicates the factorial function) is the sequence.
+%Figure~\ref{fig:syntax-group} shows an example for the expression '$\cos(2n\pi)!$'.
+If we just analyze the \gls*{pom-pt}, the previous node of the exclamation mark (which indicates the factorial function) is the sequence.
 Now it could be possible to misinterpret the sequence as the argument of the factorial function.
@@
 In that case the argument is the sequence node.
-Once the cosine function has been completely translated (which means the whole argument has been translated as well) the \gls{teo} groups the translated cosine and the sequence.
+Once the cosine function has been completely translated (which means the whole argument has been translated as well) the \gls*{teo} groups the translated cosine and the sequence.
 Therefore, when the translator reaches the exclamation mark and asks for the last translated object in the list, it is not the translated sequence but the whole cosine function.
 
-Table~\ref{tab:teo-list} shows some examples of the \gls{teo} list after the translation process was finished.
+%Table~\ref{tab:teo-list} shows some examples of the \gls*{teo} list after the translation process was finished.
 Note that a translation of '$a+b$' contains three translated objects, while '$(a+b)$' contains just one.
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{cc}
-	\hline
-	Input Expression & TEO List\\
-	\hline
-	$a+b$ & \verb|[a, +, b]|\\
-	$(a+b)$ & \verb|[(a+b)]|\\
-	$\frac{a}{b}-2$ & \verb|[(a)/(b), -, 2]|\\
-	\hline
-\end{tabular}
-\caption{How the TEO-list groups subexpressions.}
-\label{tab:teo-list}
-\end{table}
+%\begin{table}[ht]
+%\centering
+%\begin{tabular}{cc}
+%	\hline
+%	Input Expression & TEO List\\
+%	\hline
+%	$a+b$ & \verb|[a, +, b]|\\
+%	$(a+b)$ & \verb|[(a+b)]|\\
+%	$\frac{a}{b}-2$ & \verb|[(a)/(b), -, 2]|\\
+%	\hline
+%\end{tabular}
+%\caption{How the TEO-list groups subexpressions.}
+%\label{tab:teo-list}
+%\end{table}
 
-\subsubsection{Local \& Global Translated Expressions}
-During the translation process multiple local \gls{teo} are created by the subtranslators.
-Besides the local \gls{teo} there is just one global \gls{teo}.
+%\subsubsection{Local \& Global Translated Expressions}
+%During the translation process multiple local \gls*{teo} are created by the subtranslators.
+Besides the local \gls*{teo} there is just one global \gls*{teo}.
 The reason to create local objects is originally motivated by the recursive structure of the translator.
 A subtranslator should only handle the current subexpression.
-However, the global \gls{teo} becomes necessary because of the functions in postfix notation such as the factorial function.
-For example, when a subtranslator tries to translate the exclamation mark as the factorial function, the local \gls{teo} is empty, because the argument of the function has been previously translated by another subtranslator and stored in another local \gls{teo}.
-Therefore, we implement a global \gls{teo}, which contains all of the previous translated subexpressions in each state of the translation process.
-With this implementation the local \gls{teo} becomes more and more obsolete.
+However, the global \gls*{teo} becomes necessary because of the functions in postfix notation such as the factorial function.
+For example, when a subtranslator tries to translate the exclamation mark as the factorial function, the local \gls*{teo} is empty, because the argument of the function has been previously translated by another subtranslator and stored in another local \gls*{teo}.
+Therefore, we implement a global \gls*{teo}, which contains all of the previous translated subexpressions in each state of the translation process.
+With this implementation the local \gls*{teo} becomes more and more obsolete.
 In the current version, the program stores redundant translated expressions.
-For debugging reasons, a translation process still handles multiple local and one global \gls{teo}.
-Starting the forward translation with the flag \verb|--debug| (for the debugging mode) shows the elements of the global \gls{teo} after the translation has been finished.
+For debugging reasons, a translation process still handles multiple local and one global \gls*{teo}.
+Starting the forward translation with the flag \verb|--debug| (for the debugging mode) shows the elements of the global \gls*{teo} after the translation has been finished.
 
-\subsection{Additional Information}\label{subsec:add-info}
-As already mentioned, a translation is not always straight forward. In those cases, the user should get informed about the translation and the reason for it. Consider the translation
-\begin{equation}\label{eq:i-trans}
-\verb|3i| \overset{\langMaple}{\mapsto} \verb|3*i|,
-\end{equation}
-Since the imaginary unit in \Maple{} is associated with '\raw{I}' instead of '\raw{i}', a user might be confused by~\ref{eq:i-trans}.
+%\subsection{Additional Information}\label{subsec:add-info}
+%As already mentioned, a translation is not always straight forward. In those cases, the user should get informed about the translation and the reason for it. Consider the translation
+%\begin{equation}\label{eq:i-trans}
+%\verb|3i| \overset{\langMaple}{\mapsto} \verb|3*i|,
+%\end{equation}
+%Since the imaginary unit in \Maple{} is associated with '\raw{I}' instead of '\raw{i}', a user might be confused by~\ref{eq:i-trans}.
 Therefore, we created the \texttt{InformationLogger} class to organize further information about the translation process.
@@
 
-Essentially, the \raw{InformationLogger} provides further information about a translation process.
+%Essentially, the \raw{InformationLogger} provides further information about a translation process.
 This information is defined in the lexicon files.
-Consider the example of the arccotangent function, which is defined with a different branch cut in \gls{dlmf} and in \Maple.
+Consider the example of the arccotangent function, which is defined with a different branch cut in \gls*{dlmf} and in \Maple.
 As already mentioned, we provide alternative translation patterns to solve these problems.

### 2018-07-24 chapters/03-translator/gen-app.tex
Finalize first version of the journal paper Still 1.000 words too long

@@
-\section{Translation Problems}
+\section{Translation Problems}\label{sec:problems}
 There are several potential problems to peform translations between systems that embed semantic information in the input.
@@
 A similar problem are multiplications, since they are rarely explicitely marked in \LaTeX{} expressions, e.g., scientists using whitespaces to indicate multiplications rather than using \verb|\cdot| or similar symbols.
-For such problems, we introduced a new macro \verb|\idot| for an invisible multiplication sign (this macro will not be rendered).
+For such problems, we introduced a new macro \verb|\idot| for an invisible multiplication symbol (this macro will not be rendered).
 Since this macro is newly introduced and an automatic conversion of existing equations is difficult, none of the equations in the \gls*{dlmf} uses this macro yet.
 As a consequence, the translator has some simple rules to perform translations without explicitely marked translations with \verb|\idot|.
+
+Even we only allow the special dialect of \LaTeX{} using semantic macros not all expressions are unambigious.
+In table~\ref{tab:amb-latex} are four examples of ambiguous expressions.
+These expressions are unambigious for the \LaTeX{} compiler, since it only considers the very next token for power and subscripts.
+Our translator following the same rules to solve these issues.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{cc}
+	\hline
+	Ambiguous Input & \LaTeX{} Output\\
+	\hline
+	\verb|n^m!| & $n^m!$\\
+	\verb|a^bc^d| & $a^bc^d$\\
+	\verb|x^y^z| & Double superscript error\\
+	\verb|x_y_z| & Double subscript error\\
+	\hline
+\end{tabular}
+\caption{Ambiguous \LaTeX{} expressions and how \LaTeX{} displays them.}
+\label{tab:amb-latex}
+\end{table}
+
+Another more questionable translation decision are alphanumerical expressions.
+As explained in table~\ref{tab:allTypesTable}, the \gls*{pom}-tagger handles strings of letters and numbers differently, depending on the order of the symbols.
+The reason is, that an expression such as '$4b$' is usually considered to be a multiplication of $4$ and '$b$', while '$b4$' looks like indexing '$b$' by $4$.
+While the first example produces two nodes, namely $4$ and '$b$', the second example '$b4$' produces just a single alphanumerical node in the \gls*{pom-pt}.
+The translator interprets alphanumerical expressions as multiplications for two reasons.
+(1) we would assume that the inputs '$4b$' and '$b4$' are mathematically equivalent and (2) it is more common in mathematics to use single letter names for variables~\parencite{Notation:History}.
+Therefore we define the following definitions
+
+\begin{eqnarray*}
+\verb|4b| & \overset{\langMaple}{\mapsto} & \verb|4*b|,\\
+\verb|b4| & \overset{\langMaple}{\mapsto} & \verb|b*4|,\\
+\verb|energy| & \overset{\langMaple}{\mapsto} & \verb|e*n*e*r*g*y|.
+\end{eqnarray*}
+
+In general the translator is drafted to solve ambiguous expressions or automatically find a workaround to disambiguate the expression.
+Only if there is no way to solve the ambiguity with the defined rules, the translation process stops.
 
@@
 
-All translations are defined by a library that defines translation patterns for each function and symbol.
+All translations are defined by a library (\gls*{csv} and \gls*{json} files) that defines translation patterns for each function and symbol.
 The pattern uses \verb|$i| as placeholders to define the positions of the arguments.
-For example, the translation patterns for the Jacobi polynomial are illustrated in table~\ref{tab:placeholder_ex2}
+For example, the translation patterns for the Jacobi polynomial are illustrated in table~\ref{tab:placeholder_ex2}.
 
@@
 
-In some special cases, these placeholders causes trouble when the \gls*{cas} uses the symbol \verb|$| for other reasons, e.g., the differentiation in \Maple{} is defined by
-\begin{equation}
+These placeholders causes trouble when the \gls*{cas} uses the symbol \verb|$| for other reasons, e.g., the differentiation in \Maple{} is defined by
+\begin{equation*}
 \verb|diff(f, [x$n])|,
-\end{equation}
+\end{equation*}
 where $f$ as an algebraic expression or an equation, $x$ the name of the differentiation variable and $n$ for the $n$-th order differentiation.
 A translation for $\frac{d^2x^2}{dx^2}$ should be like this
-\begin{equation}
+\begin{equation*}
 \verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [x$2])|
-\end{equation}
+\end{equation*}
 but would end up as
-\begin{equation}
-\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff($1, [$2$$0])|.
-\end{equation}
+\begin{equation*}
+\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [xx])|.
+\end{equation*}
 
@@
 
-We can solve this issue by using paranthesis in such cases, e.g., 
-\begin{equation}
-\verb|diff($1, [$2$($0)])|.
-\end{equation}
+We can solve this issue by using paranthesis in such cases, e.g., \verb|diff($1, [$2$($0)])|.
 
@@
 
-\begin{table}[ht]
+\begin{table}[ht!]
 	\centering
@@
 	\end{tabular}
-	\caption{Entry of the translation library example of the Legendre and associated Legendre function of the first kind.
+	\caption{Example entries of the Legendre and associated Legendre function in the translation library.
 The prefix notation \texttt{X<d>:<name>X} defines the translation for \texttt{<name>} with \texttt{<d>}-number of optional arguments.}
@@
 
-\begin{table}
+\begin{table}[ht!]
 	\centering

### 2018-07-24 chapters/04-evaluation/evaluation.tex
Update introduction with more important information
Finalize first version of the journal paper Still 1.000 words too long

@@
-This project aims for a translation of mathematical expressions from \LaTeX{} sources to \gls{cas} and back again.
+\section{Evaluation}\label{sec:evaluation}
+%This project aims for a translation of mathematical expressions from \LaTeX{} sources to \gls*{cas} and back again.
 Essentially, the translator translates only string representations.
@@
 
-First of all, each semantic \LaTeX{} macro ties a specific character sequence to a well-defined mathematical object.
-This definition can be found at the \gls{dlmf} or \gls{drmf}.
-Therefore, our reference source for semantic \LaTeX{} expressions is always the \gls{dlmf} or \gls{drmf}.
+%First of all, each semantic \LaTeX{} macro ties a specific character sequence to a well-defined mathematical object.
+This definition can be found at the \gls*{dlmf} or \gls*{drmf}.
+Therefore, our reference source for semantic \LaTeX{} expressions is always the \gls*{dlmf} or \gls*{drmf}.
 This gives us the opportunity to compare more than the string representations of the mathematical objects in a translation process.
 For example, consider the simple translation of the cosine function
-\begin{equation}\label{eq:simple-cosine}
-\verb|\cos{x}| \overset{\langMaple}{\mapsto} \texttt{cos(x)}.
-\end{equation}
-Our verification and validation techniques, to proof for example that~(\ref{eq:simple-cosine}) is an \textit{appropriate} translation, can be grouped into three parts.
-We present in the following sections these three parts and conclusively, we use all these techniques to actually verify translations for a test suite directly extracted from the \gls{dlmf}.
+%\begin{equation}\label{eq:simple-cosine}
+%\verb|\cos{x}| \overset{\langMaple}{\mapsto} \texttt{cos(x)}.
+%\end{equation}
 
-\section{Round Trip Tests}\label{sec:round-trip}
-Round trip tests are our first approach to verify translations.
+%Our verification and validation techniques, to proof for example that~(\ref{eq:simple-cosine}) is an \textit{appropriate} translation, can be grouped into three parts.
+We present in the following sections these three parts and conclusively, we use all these techniques to actually verify translations for a test suite directly extracted from the \gls*{dlmf}.
+
+We implemented three approaches to evaluate whether a translation was \textit{appropriate} or \textit{inappropriate}.
+\begin{enumerate}
+\item \textbf{Round Trip Tests}: translates expressions back and forth and analyzing the changes.
+\item \textbf{Function Relation Tests}: translate mathematically proven equivalent expressions from one system to a \gls*{cas} and evaluate whether the relation remains valid
+\begin{enumerate}
+\item via equivalence checks of the \gls*{cas} or
+\item via numerical evaluation tests.
+\end{enumerate}
+\end{enumerate}
+
+\subsection{Round Trip Tests}\label{sec:round-trip}
+%Round trip tests are our first approach to verify translations.
 The idea is to translate an expression back and forth again and check if the representation changes during this process.
@@
 
-Furthermore, we take advantage of \Maple's symbolic simplification function to compare a \Maple{} expression in the beginning of an round trip test with the \Maple{} expression after the translations.
+%Furthermore, we take advantage of \Maple's symbolic simplification function to compare a \Maple{} expression in the beginning of an round trip test with the \Maple{} expression after the translations.
 If \Maple's \textit{simplify} function returns zero for the difference between two representations, both expressions are symbolically equivalent.
@@
 
-Obviously, there may appear two fixed point representations in round trip tests, one for each system.
+%Obviously, there may appear two fixed point representations in round trip tests, one for each system.
 We will see that most of the round trip translations reach fixed points after a certain amount of steps.
 
-\begin{theorem}[Fixed Point Properties]\label{theo:fixpoint}
-Consider two languages $A,B \in \mathcal{L}$.
+%\begin{theorem}[Fixed Point Properties]\label{theo:fixpoint}
+%Consider two languages $A,B \in \mathcal{L}$.
 A fixed point representation is reached if the current representation is just identical to the next representation in the same language (after one cycle).
 Furthermore, if $a \in A$ is a fixed point representation, the translation of $a$ to $b \in B$ is also a fixed point representation.
-\end{theorem}
-\begin{proof}
-Let $A, B \in \mathcal{L}$ be languages.
+%\end{theorem}
+%\begin{proof}
+%Let $A, B \in \mathcal{L}$ be languages.
 Consider a round trip translation test between the languages $A$ and $B$.
 Without loss of generality, let $a \in A$ be a fixed point representation and let $a$ be the \textit{first} fixed point representation that appears in this test.
-The \gls{mlp} as well as \Maple{} are deterministic, because a test case works in only one \Maple{-session} and we do not compute the inputs in \Maple.
+The \gls*{mlp} as well as \Maple{} are deterministic, because a test case works in only one \Maple{-session} and we do not compute the inputs in \Maple.
 Therefore, the translation process is not able to call any probabilistic algorithms.
@@
 
-Let $b \in B$ with
-\begin{equation}
-\mathrm{tr}^A_B(a) = b.
-\end{equation}
+%Let $b \in B$ with
+%\begin{equation}
+%\mathrm{tr}^A_B(a) = b.
+%\end{equation}
 
-Since the translation process is deterministic, $a$ will be always translated to $b$. Consider $a' \in A$ as the backward translation from $b$
-\begin{equation}
-\mathrm{tr}^B_A(b) = a'.
-\end{equation}
-Now $a$ was defined to be a fixed point representation in $A$. Therefore, $a$ is identical to all string representation in the following cycles. Thus $a = a'$.
+%Since the translation process is deterministic, $a$ will be always translated to $b$. Consider $a' \in A$ as the backward translation from $b$
+%\begin{equation}
+%\mathrm{tr}^B_A(b) = a'.
+%\end{equation}
+%Now $a$ was defined to be a fixed point representation in $A$. Therefore, $a$ is identical to all string representation in the following cycles. Thus $a = a'$.
 
-Furthermore, because the translations are deterministic, $b$ will be always backward translated to $a' = a$ and $a$ will be always forward translated to $b$.
+%Furthermore, because the translations are deterministic, $b$ will be always backward translated to $a' = a$ and $a$ will be always forward translated to $b$.
 Therefore, $b$ is identical to the representation after one cycle.
@@
 
-Remember, $a$ was chosen arbitrarily as the first fixed point representation that appears in the round trip test.
+%Remember, $a$ was chosen arbitrarily as the first fixed point representation that appears in the round trip test.
 Therefore, whenever a fixed point appears in a round trip test, the next translation to the respective other language will be a fixed point as well.
 Consequently, a fixed point is already reached, when just the next representation (after one cycle) is identical to the current representation.
-\end{proof}
+%\end{proof}
 
-With theorem~\ref{theo:fixpoint}, we are able to find fixed points in round trip tests by only testing the equality to the next representation (after one cycle).
+%With theorem~\ref{theo:fixpoint}, we are able to find fixed points in round trip tests by only testing the equality to the next representation (after one cycle).
 Furthermore, we only need to find the first fixed point representation in either system, because the next step will be a fixed point representation for the respective other system.
@@
 \begin{equation}
-\frac{\cos@{a\Theta}}{2}.
+\frac{\cos\left(a\Theta\right)}{2}.
 \end{equation}
@@
 
-Step 4 is identical to step 2.
-Theorem~\ref{theo:fixpoint} implies that step 2 is the fixed point representation for semantic \LaTeX{} and step 3 is the fixed point representation for \Maple.
-The structure of the expression changes between the steps 1 and 2.
-The reason are internal changes made by \Maple.
-These problems have been discussed in the previous \cref{subsec:maple-probs}.
+Step 4 is identical to step 2 and since the translator is a deterministic algorithm it can be easily shown that step 2 and step 3 are fixed point representations for semantic \LaTeX{} and \Maple.
 
-Almost all test cases are reaching a fixed point representation.
+%Almost all test cases are reaching a fixed point representation.
 Starting from semantic \LaTeX, the first fixed point is reached at step 2 in semantic \LaTeX, because of \Maple's internal changes.
@@
 
-There is currently only one exception known where a round trip test does not reach a fixed point representation at all: Legendre's incomplete elliptic integrals~\parencite[(19.2.4-7)]{NIST:DLMF} are defined with the amplitude $\phi$ in the first argument in the \DLMF, while \Maple{} takes the trigonometric sine of the amplitude as the first argument.
+There is currently only one exception known where a round trip test does not reach a fixed point representation: Legendre's incomplete elliptic integrals~\parencite[eq.
+19.2.4-7]{NIST:DLMF} are defined with the amplitude $\phi$ in the first argument in the \DLMF, while \Maple{} takes the trigonometric sine of the amplitude as the first argument.
 Therefore, the forward and backward translations are defined as
@@
 \verb|\EllIntF@{\phi}{k}| & \overset{\langMaple}{\mapsto} & \verb|EllipticF(sin(phi),k)|,\\
-\verb|\EllIntF@{\asin@{\phi}}{k}| & \overset{\langMaple}{\mapsfrom} & \verb|EllipticF(phi,k)|.
+\verb|\EllIntF@{\asin@{\phi}}{k}| & \overset{\langMaple}{\mapsfrom} & \verb|EllipticF(phi,k)|,
 \end{eqnarray}
 
-A round trip translation for this function produces an infinite chain of sine and inverse sine calls.
-The problem is that the round trip tests prevent simplifications in \Maple{} during the translation process (see \cref{subsec:maple-probs}).
+and round trip translations will produce infinite chains of sine and inverse sine calls.
+Since there are no evaulations involved, these chains will not be reduced during the translations.
+%The problem is that the round trip tests prevent simplifications in \Maple{} during the translation process (see \cref{subsec:maple-probs}).
 This problem could be resolved by allowing automatic evaluations in the round trip tests.
@@
 However, because of the simplification techniques of fixed points, we are able to at least detect logical errors in one system: \Maple.
-On the other hand, these tests cannot determine logical errors in the translations between the two systems - but this is the critical part we want to verify.
-Consider defining a forward and backward translation for the sine function obviously wrong as
+On the other hand, these tests cannot determine logical errors in the translations between the two systems.
+Consider we mistakenly defined an \textit{inappropriate} forward and backward translation for the sine function
 \begin{eqnarray}
@@
 In that case the round trip test would not detect any errors and reach a fixed point representation, because the simplification techniques only simplify two representations in the same system but cannot compare the representation in one system to those in the other.
-Therefore, we implemented tests for equations rather than general formulae.
 
-\section{Function Relation Tests}\label{sec:relation-tests}
-In addition to the correctness of string representations, it is more important to validate whether a translation is \textit{appropriate}.
+\subsection{Function Relation Tests}\label{sec:relation-tests}
+%In addition to the correctness of string representations, it is more important to validate whether a translation is \textit{appropriate}.
 Special functions are highly related to each other.
 The theory of special functions researches these relations and other properties of the functions.
-The \gls{dlmf} is a compendium of special functions and lists a lot of those relations.
-The translation of such relations to another system and the following analysis of the relations with the translated formulae would form a very effective test.
+
+The \gls*{dlmf} is a compendium of special functions and orthogonal polynomials and lists several relations between the functions and polynomials.
+The idea of this evaluation approach is to translate an entire relation and test whether the relation remains valid after perfoming the translations.
 
 With this technique we can detect translation errors such as in (\ref{eq:wrong-trans-1} and \ref{eq:wrong-trans-2}).
-Consider the \gls{dlmf} equation for the sine and cosine function~\parencite[(4.21.2)]{NIST:DLMF}
+Consider the \gls*{dlmf} equation for the sine and cosine function~\parencite[eq 4.21.2]{NIST:DLMF}
 \begin{equation}
@@
 However, there are two essential problems with this approach.
-Testing the mathematical equivalence of expressions is hard to solve and \gls{cas} have trouble to test even simple equations.
+Testing the mathematical equivalence of expressions is hard to solve and \gls*{cas} have trouble to test even simple equations.
 Furthermore, this approach only checks forward translations because there is no way to check equivalence of expressions in \LaTeX{} automatically (again this could become feasible with our translator).
@@
 However, there are several cases where the simplification fails.
-We implement some tricks to help \textit{simplify}: pre-conversion of the functions of the expression to other special functions is able to help \textit{simplify}, because the algorithms in the background can use different simplification methods for different functions.
-For example, the user can use the \Maple{} function \textit{convert} to convert the expression
+Because of implementation details, there are some techniques that helps \Maple{} to find possible simplifications.
+For example we can force \Maple{} to convert the formula
 \begin{equation}
@@
 \end{equation}
-to an equivalent representation using the exponential function:
+to an equivalent representation using the exponential representations
 \begin{equation}
@@
 However, the limitations of the \textit{simplify} function are still the weakest part of this verification approach.
-Consider the equation~(\ref{eq:branch-cut-issue}) from \cref{subsec:branch-cut-issues}.
+Consider the complex example~\parencite[eq.
+12.7.10]{NIST:DLMF}
 \begin{equation}\label{eq:branch-cut-near}
-\displaystyle U(0,z) = \sqrt{\frac{z}{2\cpi}} K_{\frac{1}{4}}\left(\frac{1}{4}z^2\right)
+\displaystyle U(0,z) = \sqrt{\frac{z}{2\cpi}} K_{\frac{1}{4}}\left(\frac{1}{4}z^2\right),
 \end{equation}
-As already discussed, the modified Bessel function of the second kind~\parencite[(12.7.10)]{NIST:DLMF} has its branch cut in \Maple{} at $z < 0$.
+where $U(0,z)$ is the parabolic cylinder function and $K_\nu(z)$ the modified Bessel function of the second kind.
+Both functions are well-defined in both systems and we can define a \textit{direct} translation for~\eqref{eq:branch-cut-near}.
+The modified Bessel function of the second kind has its branch cut in \Maple{} and in the \gls*{dlmf} at $z < 0$.
 However, the argument of $K$ contains $z^2$.
-For now, just focus on
-\begin{equation}\label{eq:besselK}
-K_{\frac{1}{4}}\left(\frac{1}{4}z^2\right)
-\end{equation}
-Because of $z^2$, if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ the value of (\ref{eq:besselK}) would be no longer on the principal branch.
-However, \Maple{} will still compute (\ref{eq:besselK}) on the principal branch independently of the value of $z$.
+If $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ the value of the right-hand side of~\eqref{eq:branch-cut-near} would be no longer on the principal branch.
+However, \Maple{} will still compute the principal values independently of the value of $z$.
 Hence, a translation
@@
 is incorrect if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ and one has to use analytic continuation for the right-hand side of equation (\ref{eq:branch-cut-near}).
-In subsection~\ref{subsec:branch-cut-issues} we also shown that the complete right-hand side of equation~(\ref{eq:branch-cut-near}) has three branch cuts, caused by the square root function and the $z^2$ argument of the modified Bessel function of the second kind.
+%In subsection~\ref{subsec:branch-cut-issues} we also shown that the complete right-hand side of equation~(\ref{eq:branch-cut-near}) has three branch cuts, caused by the square root function and the $z^2$ argument of the modified Bessel function of the second kind.
 Such complex examples cannot be simplified yet.
 
-However, this validation approach is powerful and runs automatically. We use this approach to verify a large set of test cases extracted directly from the \gls{dlmf}, see \cref{sec:test-summary}.
+%However, this validation approach is powerful and runs automatically. We use this approach to verify a large set of test cases extracted directly from the \gls*{dlmf}, see \cref{sec:test-summary}.
 
-\section{Numerical Tests}\label{sec:numerical-tests}
-Another approach of verification are numerical tests.
+To evaluate such complex cases, the equivalence checks of \gls*{cas} are insufficient. Therefore we implement numerical tests as an additional step.
+
+\subsection{Numerical Tests}\label{sec:numerical-tests}
+%Another approach of verification are numerical tests.
 In principle, we follow the same approach as the relation tests in \cref{sec:relation-tests}, but rather than using the simplification functions we compare the left-hand side and right-hand side for actual values.
 
-For example, reconsider equation~(\ref{eq:branch-cut-near}) from \parencite[(12.7.10)]{NIST:DLMF}. Define the difference of right-hand side and left-hand side
+Consider the differences of the left- and right-hand side of equation~\eqref{eq:branch-cut-near}
 \begin{equation}\label{eq:difference}
-D(z) := U(0,z) - \sqrt{\frac{z}{2\cpi}} K_{\frac{1}{4}}\left(\frac{1}{4}z^2\right),
+D(z) := U(0,z) - \sqrt{\frac{z}{2\cpi}} K_{\frac{1}{4}}\left(\frac{1}{4}z^2\right).
 \end{equation}
-and compute $D(z)$ for some values of $z$. Table~\ref{tab:computations-for-difference} presents four computations for $D(z)$, one value for each quadrant in the complex plane.
+Table~\ref{tab:computations-for-difference} presents four computations for $D(z)$, one value for each quadrant in the complex plane.
 \begin{table}[ht]
@@
 
-The values are computed with the default precision of $10$ significant digits.
+Considering machine accuracy and the default precision of $10$ significant digits, we can regard the first and last values as zero differences.
+While this evaluation is very powerful, it has a significant problem.
+Even when all tested values return a value differ from zero, it does not proof the equivalence of~\eqref{eq:branch-cut-near}.
+However, when the values are different from zero it does proof an error in one of the four cases~\parencite{NumericalTests:Paper}
+\begin{enumerate}
+\item the numerical engine tests invalid combinations of values;
+\item the translation is incorrect;
+\item there may be an error in the \gls*{dlmf} source; or
+\item there may be an error in \Maple.
+\end{enumerate}
+
+%The values are computed with the default precision of $10$ significant digits.
 Caused by the machine accuracy, the first and last values are different from zero.
@@
 
-This approach of verification is very powerful, because it can verify not only our translations but also equations in a mathematical compendium.
-This has been demonstrated by discovering an overall sign error of the \gls{dlmf} equation~(\ref{eq:signerror}), see \cref{sec:test-summary}.
+%This approach of verification is very powerful, because it can verify not only our translations but also equations in a mathematical compendium.
+This has been demonstrated by discovering an overall sign error of the \gls*{dlmf} equation~(\ref{eq:signerror}), see \cref{sec:test-summary}.
 
-However, this approach has also a significant problem.
+%However, this approach has also a significant problem.
 It is easy to see that something went wrong, but very hard to proof equivalence.
@@
 Reformulating this question to a general problem brings us to:
-\begin{quote}
-\textit{How can we discretize equations down to a finite set of test points such that the probability of equivalence for the continues equation is high if the equation is valid for the finite set of test points?}
-\end{quote}
-As far as we know, there is no answer to this question yet.
+%\begin{quote}
+%\textit{How can we discretize equations down to a finite set of test points such that the probability of equivalence for the continues equation is high if the equation is valid for the finite set of test points?}
+%\end{quote}
+%As far as we know, there is no answer to this question yet.
 This is one reason why we do not implement an automatic process of numerical tests yet.
@@
 
-\section{DLMF Test Suite}\label{sec:test-summary}
-To verify our translations with the approaches described above we create test suites by extracting equations and formulae from the sources of the \gls{dlmf}.
+\subsection{Results}\label{sec:test-summary}
+There are 685 \Macro s\footnote{The macros are still work in progress and therefore the total number is constantly changing.} in total and 665 of them were implemented in the translator engine.
+We defined forward translations to \Maple{} for 201 and backward translations from \Maple{} for 195 functions.
+
+The \gls*{dlmf} provides a dataset of \LaTeX{} expressions with semantic macros.
+We extracted 4.087 equations from \gls*{dlmf} and apply our round-trip and relation tests on them.
+The translator was able to translate 2.405 (58.8\footnote{All percentages are approximately calculated.}\%) of the extracted equations without errors.
+660 (27.4\%) of the successfully translated expressions were verified by the simplification techniques of \Maple.
+We apply additional numerical tests for the remaining 1.745 equations.
+For 418 (24\%) cases the numerical tests were valid.
+More detailed results for numerical and symbolical tests were presented in~\parencite{NumericalTests:Paper}.
+
+The evaluation techniques has proven to be very powerful also for evaluating \gls*{cas} and mathematical online compendia such as the \gls*{dlmf}.
+During the evaluations we were able to detect several errors in the translation and evaluation engine.
+However, even errors in the \gls*{dlmf} were discovered.
+
+The numerical test engine was able to discover a sign error in equation~\parencite[eq.
+14.5.14]{NIST:DLMF}\footnote{The equation had originally been stated as shown in equation~(\ref{eq:signerror}).
+The error were reported on 10th April 2017.}
+\begin{equation}\label{eq:signerror}
+\displaystyle \mathsf{Q}^{-1/2}_{\nu}\left(\cos\theta\right)=-\left(\frac{\pi}{2\sin\theta}\right)^{1/2}\frac{\cos\left(\left(\nu+\frac{1}{2}\right)\theta\right)}{\nu+\frac{1}{2}}.
+\end{equation}
+The error can be found on~\parencite[p.
+359]{NIST:Handbook} and has been fixed in the \gls*{dlmf} with version 1.0.16.
+The same engine also identified a missing comma in the constraint of~\parencite[eq.
+10.16.7]{NIST:DLMF}.
+The original constraint was given by $2\nu \neq -1, -2 -3, \ldots$, with a missing comma after $-2$.
+
+%To verify our translations with the approaches described above we create test suites by extracting equations and formulae from the sources of the \gls*{dlmf}.
 We start with a small test suite of 110\footnote{The set of test cases is available at\newline \url{https://gist.github.com/AndreG-P/f48121d70fbd2a57b2cf6d88019ddadf}} equations and provide a more detailed look to those tests.
@@
 
-Two of the test cases (lines 42 and 43) contain macros we are not able to translate yet.
+%Two of the test cases (lines 42 and 43) contain macros we are not able to translate yet.
 Those macros are \verb|\subplus| and \verb|\CFK|, which are used for continued fractions.
 They have been created for another translation project~\parencite[§5]{CICM:Paper} for continued fractions defined in the \textit{Handbook of Continued Fractions for Special Functions} (2008) \parencite{CF:Handbook}.
-The conversion process for the \gls{cfsf} \parencite{CF:Maple} dataset from \Maple{} and the \gls{ecf} \parencite{CF:eCF,eCF:FinalReport} dataset from Wolfram uses special file types.
+The conversion process for the \gls*{cfsf} \parencite{CF:Maple} dataset from \Maple{} and the \gls*{ecf} \parencite{CF:eCF,eCF:FinalReport} dataset from Wolfram uses special file types.
 Therefore, our translator is not able to handle those macros yet.
 
-However, we translate 108 test cases and for 84 of them the simplification function of \Maple{} returns zero as the difference of the left-hand side and right-hand side.
+%However, we translate 108 test cases and for 84 of them the simplification function of \Maple{} returns zero as the difference of the left-hand side and right-hand side.
 The 24 remaining cases are translated, but the simplification function is not able to simplify the difference to zero.
@@
 For all those 24 test cases, we perform also manual numerical tests, which reveal an overall sign error in equation~\parencite[(14.5.14)]{NIST:DLMF}\footnote{The equation had originally been stated as shown in equation~(\ref{eq:signerror}).
-The \gls{dlmf} has already updated the wrong equation.}
-\begin{equation}\label{eq:signerror}
-\displaystyle P^{1/2}_{\nu}\left(\cosh{\xi}\right) = - \left( \frac{2}{\cpi \sinh{\xi}} \right)^{1/2} \cosh{\left( \nu + \tfrac{1}{2} \right) \xi}.
-\end{equation}
-Besides this wrong equation, all of these manual numerical tests return zero for the difference between the left-hand side and right-hand side.
+The \gls*{dlmf} has fixed the wrong equation with version 1.0.16.}
+%\begin{equation}\label{eq:signerror}
+%\displaystyle \mathsf{Q}^{-1/2}_{\nu}\left(\cos\theta\right)=-\left(\frac{\pi}{2\sin\theta}\right)^{1/2}\frac{\cos\left(\left(\nu+\frac{1}{2}\right)\theta\right)}{\nu+\frac{1}{2}}.
+%\end{equation}
+%Besides this wrong equation, all of these manual numerical tests return zero for the difference between the left-hand side and right-hand side.
 Of course it is still not possible to conclude the equivalence based on our manual numerical tests, as explained above.
 
-Those 110 test cases are not sufficiently comprehensive for all of the functions of the \gls{dlmf}.
-Therefore, we create a larger dataset\footnote{We are planning to publish this dataset at \url{http://drmf.wmflabs.org}.} of 4,165 semantic \LaTeX{} formulae, which is extracted from all chapters of the \gls{dlmf}.
+%Those 110 test cases are not sufficiently comprehensive for all of the functions of the \gls*{dlmf}.
+Therefore, we create a larger dataset\footnote{We are planning to publish this dataset at \url{http://drmf.wmflabs.org}.} of 4,165 semantic \LaTeX{} formulae, which is extracted from all chapters of the \gls*{dlmf}.
 
-We are able to translate 2,232 (approx.
+%We are able to translate 2,232 (approx.
 53.59\%) of those test cases.
@@
 
-The translator cannot translate 1,933 test cases.
+%The translator cannot translate 1,933 test cases.
 First of all, 32 lines contain neither an equation nor a relation.
@@
 
-Other errors appear in 914 cases, where the semantic information is missing or is not unique.
+%Other errors appear in 914 cases, where the semantic information is missing or is not unique.
 An example are 284 lines that use a potentially ambigious single quote (or prime) notation '\aSingleQuote' to identify derivatives of functions.
-
-\cleardoublepage

### 2018-07-24 chapters/06-future-work-and-conclusion/conclusion.tex
Finalize first version of the journal paper Still 1.000 words too long

@@
 In that case we would end up in something similar to \sTeX, which is overloaded and too complicated to use.
-Obviously, if the system is no longer handy, it will not be used.
+If the system is no longer handy, it will not be used.
 
-On the other hand, the concept of the translator has been proven.
-For example, one test case has revealed an overall sign error in an equation in the \DLMF~\cite[(14.5.14)]{NIST:DLMF} and therefore the validation system has shown the potential to become a strong verification tool for mathematical compendia.
+On the other hand, the concept of the translator has been proven by discover errors in the online compendia \gls*{dlmf}.
 The test cases have also shown how difficult it is to validate a translated expression and have uncovered the problems of translations between two systems with different sets of supported functions.
@@
 
-Finally we can conclude that the project has been successful and the translator has great potential to become a handy translation tool.
+The tasks for future work are diversive.
+The main task to improve the translator by implementing more functions and features.
+For example, for the current state only translations to \Maple's standard function library were implemented.
+\Maple{} allows to load extra packages dynamically and therefore support several more functions.
+This feature would drastically increase the number of possible translations.
+With such improvements, further work on evaluation techniques become worthwhile to evaluate \gls*{dlmf} and \gls*{cas}.
+
+Furthermore, the translator was designed to be easily extendable. This allows to implement translations for other \gls*{cas} without much effort.
+
+The biggest weakness of the translator are still the semantic macros.
+Currently, we working on mathematical information retrieval techniques to allow an extension for the translator for generic \LaTeX{} inputs.
+
+%Finally we can conclude that the project has been successful and the translator has great potential to become a handy translation tool.
 Nonetheless, it is without any doubts an ever-growing project and needs to be improved constantly.

### 2018-07-25 chapters/01-introduction/intro.tex
Clear unused files, upload highres images and fix errors

@@
 
-Translations from generic \LaTeX{} to \gls{cas} are difficult to realize since the semantic information are not explicitely given in the input.
-The \gls{nist} has created a set of semantic \LaTeX{} macros~\cite{DLMF:Macros}.
-Each macro ties specific character sequences to a well-defined mathematical object and is linked with the corresponding definition in the \gls{dlmf} or \gls{drmf}.
+Translations from generic \LaTeX{} to \gls*{cas} are difficult to realize since the semantic information are not explicitely given in the input.
+The \gls*{nist} has created a set of semantic \LaTeX{} macros~\cite{DLMF:Macros}.
+Each macro ties specific character sequences to a well-defined mathematical object and is linked with the corresponding definition in the \gls*{dlmf} or \gls*{drmf}.
 These macros embeds necessary semantic information into \LaTeX{} expressions.
@@
 In general, positioning branch cuts follows conventions, but can be positioned arbitrarily in many cases.
-Communicate and explain the decision of defined branch cuts is a critical point for \gls{cas} and can vary between the systems~\parencite{Branches:acot}.
+Communicate and explain the decision of defined branch cuts is a critical point for \gls*{cas} and can vary between the systems~\parencite{Branches:acot}.
 Figure~\ref{fig:acot-cut-compare} illustrates two examples of different branch cut positioning for the inverse trigonometric arccotangent function.
@@
 
-This article presents a new approach for automatic \gls{p2c} and vice versa conversions.
+This article presents a new approach for automatic \gls*{p2c} and vice versa conversions.
 Translations from presentational to computational (computational to presentational) systems are called forward (backward) translations.

### 2018-07-25 chapters/03-translator/backward-trans.tex
Clear unused files, upload highres images and fix errors

@@
 
-\begin{wrapfigure}{r}{0.62\textwidth}
+\begin{wrapfigure}{r}{0.61\textwidth}
 	\vspace{-22pt}
 	\centering
-	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.5]{DAGreal.pdf}
-	\vspace{-26pt}
+	\includegraphics[clip, trim=0.5cm 0.5cm 0.5cm 0.5cm, scale=0.5]{DAGreal.pdf}
+	\vspace{-5pt}
 	\caption{The internal \Maple{} DAG representation of $x^2+x$.}
@@
 This internal representation might look a bit different to the input.
-One example has already been given with figure~\ref{internal-maple-dag}, where each summand of a sum is stored with a factor.
+One example has already been given with figure~\ref{fig:internal-maple-dag}, where each summand of a sum is stored with a factor.
 Here is a list of all internal changes that occur for inputs.

### 2018-07-25 chapters/03-translator/forward-trans.tex
Clear unused files, upload highres images and fix errors

@@
 
-The overall forward translation process is explained in figure~\ref{fig:forward-trans}.
+{\sloppy The overall forward translation process is explained in figure~\ref{fig:forward-trans}.
 All translation patterns and related information are stored in the DLMF/DRMF tables.
-These tables are converted by the \verb|lexicon-creator.jar| to the \verb|DLMF-macros-lexicon.txt| lexicon file.
-Together with the \verb|global-lexicon.txt| file, the \gls*{pom-pt} will be created by the \gls*{pom}-tagger.
-The \verb|latex-converter.jar| takes a string representation of a semantic \LaTeX{} expression and uses the \gls*{pom} engine as well as our \verb|Translator| to create an appropriate string representation for a specified \gls*{cas}.
+These tables are converted by the \verb|lexicon-creator| to the \verb|DLMF-macros-lexicon| lexicon file.
+Together with the \verb|global|-\verb|lexicon| file, the \gls*{pom-pt} will be created by the \gls*{pom}-tagger.
+The \verb|latex-converter| takes a string representation of a semantic \LaTeX{} expression and uses the \gls*{pom} engine as well as our \verb|Translator| to create an appropriate string representation for a specified \gls*{cas}.}
 

### 2018-07-25 chapters/01-introduction/intro.tex
Clear unused lines and use grammarly checks

@@
 
-Translations from generic \LaTeX{} to \gls*{cas} are difficult to realize since the semantic information are not explicitely given in the input.
+Translations from generic \LaTeX{} to \gls*{cas} are difficult to realize since the semantic information is not explicitly given in the input.
 The \gls*{nist} has created a set of semantic \LaTeX{} macros~\cite{DLMF:Macros}.
 Each macro ties specific character sequences to a well-defined mathematical object and is linked with the corresponding definition in the \gls*{dlmf} or \gls*{drmf}.
-These macros embeds necessary semantic information into \LaTeX{} expressions.
+These macros embed necessary semantic information into \LaTeX{} expressions.
 One example of such a macro is given table~\ref{tab:JacobiP-usecase} for the semantic \LaTeX{} representation for the Jacobi polynomial.

### 2018-07-25 chapters/02-background/related-work.tex
Clear unused lines and use grammarly checks

@@
 The second problem appears in the export tools.
-Mathematical expressions in \gls*{cas} are fully semantic, otherwise the \gls*{cas} wouldn't be able to compute or evaluate the expressions.
+Mathematical expressions in \gls*{cas} are fully semantic.
+Otherwise the \gls*{cas} wouldn't be able to compute or evaluate the expressions.
 During the export process, the semantic information gets lost, because generic \LaTeX{} is not able to carry semantic information.
 In consequence of these two problems, an exported expression cannot be imported to the same system again in most cases (except those simple expressions described above).
-Our tool should solve these problems and provide round trip translations between \LaTeX{} and \gls*{cas}.
+Our tool should solve these problems and provide round-trip translations between \LaTeX{} and \gls*{cas}.
 
 The semantics must be well known before an expression can be translated.
-There are two naiv approaches to solve that problem: (1) someone could specify the semantic information during the writing process (pre-defined semantics) or (2) the translator can find out the right semantic information in general mathematical expressions before it translates the expression.
-So called \textit{interactive documents}\footnote{There is no adequate definition what interactive documents are.
-However, this name is widely used to describe electronic document formats that allow interactivity to change the content in real time.}, such as the \gls*{cdf}~\parencite{CDF:Wolfram} by Wolfram Research or the \textit{worksheets} by \Maple{}, try to solve this problem with approach (2) and allow to embed semantic information into the input.
+There are two main approaches to solve that problem: (1) someone could specify the semantic information during the writing process (pre-defined semantics) or (2) the translator can find out the right semantic information in general mathematical expressions before it translates the expression.
+So-called \textit{interactive documents}\footnote{There is no adequate definition what interactive documents are.
+However, this name is widely used to describe electronic document formats that allow interactivity to change the content in real time.}, Such as the \gls*{cdf}~\parencite{CDF:Wolfram} by Wolfram Research or the \textit{worksheets} by \Maple{}, try to solve this problem with the approach (2) and allow to embed semantic information into the input.
 Those complex document formats require specialized tools to show and work with the documents (Wolfram CDF Player, or \Maple{} for the \textit{worksheets}).
-The \JOBAD{} architecture~\parencite{JOBAD:orig} is able to create web based interactive documents and uses \gls*{omdoc}~\parencite{OMDoc} to carry semantics.
+The \JOBAD{} architecture~\parencite{JOBAD:orig} is able to create web-based interactive documents and uses \gls*{omdoc}~\parencite{OMDoc} to carry semantics.
 The documents can be viewed and edited in the browser.
@@
 This gives the opportunity to calculate, compute and change mathematical expressions directly in the document.
-The translation performs in the background, invisible for the user.
+The translation performs in the background, invisible to the user.
 Similar to the \JOBAD{} architecture other interactive web documents exist, such as \textit{MathDox}~\parencite{MathDox} and \textit{The Planetary System}~\parencite{Planetary}.
-All of them demonstrate the potential for the education system.
-
-%All of these systems carry semantic information.
-Most of them create their own way to make this possible.
-The web based documents mostly use standardized ways, such as the content \gls*{mathML}, a specialization of \gls*{mathML}~\parencite{MathML}, that focuses on the semantics of an expression rather than any particular rendering for the expression.
-Since \LaTeX{} is the de facto standard to write mathematical expressions in documents there are projects to extend \LaTeX{} in a way that enable to carry the semantic information directly in \LaTeX.
-There are two known projects that created a semantic version of \LaTeX: \sTeX{}~\parencite{sTeX} developed by Michael Kohlhase and the \Macro s developed by Bruce Miller.
-Our translator uses the \Macro s rather than \sTeX.
-A full explanation of this decision will be given in \cref{sec:semantics}.
-
-%However, \textit{interactive documents}, web based or not, have an essential problem: all of them create an entire new document format.
-We want to create an independent lite translation tool for \LaTeX.
-Once we achieve this goal, an integration into \textit{interactive documents} might be possible, besides many other possible applications.
+All of them demonstrate the potential of the education system.
 
@@
 
-%Obviously, the first example is not as powerful as a \gls*{cas} would be.
-However, \textit{sagetex} does not really solve our problem of the scientific workflow.
-The writer still needs to translate expressions manually.
-The only difference is that the author can write computations directly into the \TeX -file.
-However, these approaches have the potential for a powerful combination with our translation tool.
-
-%In summary, today's translation tools are either specialized for one specific \gls*{cas} (such as the import and export functions of the \gls*{cas}) or they completely avoid the translation process by creating a new document format (such as \textit{interactive documents}).
-Our goal is to provide a lite translation tool for mathematical \LaTeX{} expressions and multiple \gls*{cas}.
-
-There exists two approaches that allow to embed semantic information within \LaTeX{} expressions by using custom macros.
+There exist two approaches that allow embedding semantic information within \LaTeX{} expressions by using custom macros.
 Namely, \sTeX{}~\parencite{sTeX} developed by Kohlhase and the \Macro s developed by Miller~\parencite{DLMF:Macros}.

### 2018-07-25 chapters/03-translator/backward-trans.tex
Clear unused lines and use grammarly checks

@@
 Internally, \Maple{} using \gls*{dag} as syntax trees.
-
-%\subsection{Internal Data Structure}
-%While our forward translation is realized for the \texttt{1D} input representation in \Maple, this representation is not suitable for a backward translation.
-Since \Maple{} uses it's own syntax and programming language, we would need a new parser for \Maple{} expressions to realize a backward translation tool.
-Of course, \Maple{} has this parser implemented.
-But it is not possible to use the parser without using \Maple{} itself.
-Therefore, an installed version of \Maple{} is mandatory for our backward translation.
-
-%As already mentioned, \Maple{} has several ways to represent an expression.
-The \texttt{1D} and \texttt{2D} representations are used as input representations.
-Besides those, there are internal data structures to perform all kinds of computations on the expression.
-Internally, each expression is handled and stored as a \gls*{dag}.
-The \Maple{} \gls*{dag} is very similar to an expression tree.
-But for efficiency, it does not store copies of an object.
-Consider the integral from (\ref{eq:maple-input})
-%\begin{equation}\label{eq:maple-input}
-%\int\displaylimits_0^\infty \frac{\pi+\sin(2x)}{x^2} dx.
-%\end{equation}
-
-%As already mentioned, the \Maple{} \gls*{dag} has no copies of objects. Therefore, the '$x$' in \eqref{eq:maple-input-2} only appears once in the \gls*{dag}.
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[scale=0.7]{DAG.pdf}
-%\caption{The \Maple{} DAG of equation~\eqref{eq:maple-input}}
-%\label{fig:maple-dag}
-%\end{figure}
-
-%The \Maple{} \gls*{dag} can be visualized in a graph, such as in figure~\ref{fig:maple-dag}. But the internal representation looks a bit different.
 
@@
 Some of the important types for the nodes are specified in table~\ref{tab:maple-types}.
-
-%\Maple{} has also another representation, similar to the \gls*{dag}, that allows multiple copies of the same object in its representation.
-This representation is called \inertF{} and is more similar to an expression tree than \Maple's \gls*{dag}.
-The \inertF{} represents the \gls*{dag} as a tree by splitting object copies and displays the tree in a list structure.
-Each node (or element in the list) has the prefix \texttt{\_Inert\_XXX}, that indicates the node is in the \inertF.
-The suffix \texttt{XXX} is the type of the node.
-Some of the important types are specified in table~\ref{tab:maple-types}.
-Note that this is only a subset of all possible types.
 
@@
 
-%The \texttt{OpenMaple} \gls*{api} is implemented for different programming languages such as \texttt{C}, \texttt{Java} and \texttt{Visual Basic}.
-Since our forward translation and the \gls*{mlp} is implemented in \texttt{Java}, we use the \texttt{Java} \gls*{api} for the backward translation.
-The \texttt{Java} \gls*{api} has some limitations to interact with \Maple's \gls*{dag}.
-
-%For the backward translation, we use the \inertF.
-This representation can be converted to a nested list version and the \texttt{OpenMaple} Java \gls*{api} can handle lists.
-Therefore, the backward translation firstly converts the \inertF{} to a nested list structure before any translation process starts.
-
-%The following figure~\ref{fig:inertform-list} shows the \inertF{} and the nested list of the polynomial $x^2+x$.
-
-%\begin{figure}[ht]
-%\centering
-%\begin{minipage}{.45\linewidth}
-%\centering
-%\begin{tabular}{lll}
-%\multicolumn{3}{l}{\_Inert\_SUM(}\\
-%& \multicolumn{2}{l}{\_Inert\_POWER(}\\
-%& & \_Inert\_NAME("x"),\\
-%& & \_Inert\_INTPOS(2)\\
-%& ),\\ 
-%& \multicolumn{2}{l}{\_Inert\_NAME("x")}\\
-%)
-%\end{tabular}
-%\end{minipage}
-%\begin{minipage}{.45\linewidth}
-%\centering
-%\begin{tabular}{lll}
-%\multicolumn{3}{l}{[\_Inert\_SUM,}\\
-%& \multicolumn{2}{l}{[\_Inert\_POWER, }\\
-%& & [\_Inert\_NAME, "x"],\\
-%& & [\_Inert\_INTPOS, 2]\\
-%& ],\\ 
-%& \multicolumn{2}{l}{[\_Inert\_NAME, "x"]}\\
-%]
-%\end{tabular}
-%\end{minipage}
-%\caption{\inertF{} and the nested list representation of $x^2+x$.}
-%\label{fig:inertform-list}
-%\end{figure}
-
-%In the nested list, the first element specifies the type of the current list while the following arguments store the arguments and optional attributes.
-
 \subsection{Automatic Changes of Inputs in Maple}\label{subsec:maple-probs}
-%Note that the backward translations usually do not have problems with ambiguous expressions.
-Every \gls*{cas} needs to solve ambiguities in its own representation, otherwise it could not compute the expression.
-Our backward translation from \Maple{} back to semantic \LaTeX{} is therefore based on already disambiguated expressions.
-
-%The general walkthrough of our backward translation starts with a string of a \texttt{1D} \Maple{} input expression.
-As described above, this string is entered in \Maple{} to parse the expression and work on the nested list version of the \inertF{} afterwards.
-This causes some problems.
-Mostly because \Maple{} automatically evaluates input expressions immediately - even before we have a chance to take a look at the internal data structure of the input.
-For example, an input such as $\sin\left(\frac{\cpi}{2}\right)$ is immediately evaluated to $1$ and the internal data structure just shows us the positive integer value $1$ instead of the trigonometric function.
-
-%Furthermore, \Maple{} also changes input expressions slightly, mostly because of missing data types for a specific expression.
-For example, the internal data structure has no type to represent fractions except the \texttt{RATIONAL} type, which only allows integer values for the numerator and denominator.
-Thereby, an expression like $\frac{x}{2}$ is automatically changed to $x\frac{1}{2}$.
-There is no way no avoid such simple arithmetic changes in the expression, but we can avoid the automatic evaluations.
-
-%Firstly, we summarize all known changes and evaluations \Maple{} automatically performs on input expressions.
-
 \Maple{} evaluates inputs automatically and changes the input into an internal representation.
@@
 
-Since we want to keep a translated expression similar to the input expression, we implemented some cosmetic rules for the backward translations that solves or reduce the effects from the list of changes above.
+Since we want to keep a translated expression similar to the input expression, we implemented some cosmetic rules for the backward translations that solve or reduce the effects from the list of changes above.
 \begin{itemize}
-\item We use unevaluation quotes to supress evaluations of the input.
-\item We perform reordering of factors and summands so that negative factors appears in front of the summand.
+\item We use unevaluation quotes to suppress evaluations of the input.
+\item We perform reordering of factors and summands so that negative factors appear in front of the summand.
 This gives us the opportunity to translate $x - y$ to $x - y$ instead of $x + y \cdot (-1)$.
@@
 
-%Since we want to keep a translated expression similar to the input expression, we use unevaluation quotes during the whole translation process.
-Unevaluation quotes also solve the problems with square roots and $n$-th roots.
-In \Maple's \texttt{1D} input representation, a square root is a function call (\texttt{sqrt(x)}) and the unevaluation quotes prevent evaluations on functions.
-Therefore, a backward translation simply needs to translate a square root as a normal function.
-Hence, we do not have problems with the internal representation of roots in \Maple.
-
-%Because of the absent data type for subtractions, a long sum with negative terms could be difficult to read.
-We change the order of constants (such as an integer value) to be in front of products.
-Therefore, an expression such as '$-y$' is not represented by '$y\cdot (-1)$', but by '$(-1) \cdot y$'.
-The translator now needs to check if there is a leading '$-1$' and can translate it to '$-y$' rather than to '$(-1) \cdot y$'.
-
-%Floating point numbers in scientific notation also causes a problem. Consider the input expression $0.41$ as the nested list \inertF{} representation
-%\begin{equation}
-%\texttt{[\_Inert\_FLOAT, [\_Inert\_INTPOS, 41], [\_Inert\_INTNEG, 2]]}.
-%\end{equation}
-
-%The backward translator is implemented in \texttt{Java}.
-To translate such an expression back to $0.41$, computations become necessary.
-Of course, it makes sense to perform those computations in the \gls*{cas} (in that case, in \Maple) rather than in the translation program.
-Therefore, we implemented a procedure\footnote{\textit{Procedures} in \Maple{} are small programs similar to methods and functions in \gls*{oop} languages.} to automatically convert \texttt{FLOAT} nodes to a string representation of the floating point number.
-We created a new type, called \texttt{MYFLOAT} to represent such numbers.
-With this approach, we avoided computations in the translation process.
-
-%For fractions, we use a similar approach and introduce a new internal data type \texttt{DIVIDE}.
-The implementation of \texttt{DIVIDE} was difficult.
-On the one hand, some expressions with negative exponent should be displayed as a fraction while others should should be displayed with the negative exponent.
-We follow the same rule as \Maple{} follows internally to display fractions.
-When the exponent is a negative numeric element, the expression should be displayed as a fraction.
-If the exponent is something else, we uses the negative exponent representation.
-Table~\ref{tab:maple-fracs} shows two examples with negative exponents.
-One of them is displayed as a fraction (exponent '$-2$') and the other is not (exponent '$-2y$').
-We implement the same rule for our newly developed \texttt{DIVIDE} data type.
-Note that with the new created data type, we are still not able to avoid changes from $\frac{x}{2}$ to $x\frac{1}{2}$, but we are able to work with more general fraction expressions.
-
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{p{2.6cm}cc}
-%	\hline
-%	& Negative numeric exponent & General negative exponent\\
-%	\hline
-%	\texttt{1D} \Maple{} input expression & $\texttt{(x-1)\^{}(-2)}$ & \rule{0pt}{0.9\normalbaselineskip}$\texttt{(x-1)\^{}(-2*y)}$\\
-%	How \Maple{} displays the input & $\displaystyle\frac{1}{(x-1)^2}$ & $\displaystyle(x-1)^{-2y}$\\
-%	\hline
-%\end{tabular}
-%\caption{Different styles to display negative exponents depending on the type of the exponent.
-\Maple{} displays expressions with negative numeric exponents as fractions, while other negative exponents are not displayed as fractions.}
-%\label{tab:maple-fracs}
-%\end{table}
-
-%We implement all of these changes before the actual translation process starts. This helps keeping the translated expression similar to the input expression.
-
-%\subsection{Implementation Details}\label{subsec:maple-impl}
-%\begin{wrapfigure}{l}{0.485\textwidth}
-%	\vspace{-12pt}
-%	\centering
-%	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.39]{BackwardTranslator.pdf}
-%	\caption{A scheme of the backward translator and its specialized subtranslators.}
-%	\label{fig:BackwardDel}
-%	\vspace{-13pt}
-%\end{wrapfigure}
-
-%The general translation process runs through different states.
-Each translation process starts with the string representation of a \texttt{1D} \Maple{} input.
-This string will be converted to the internal nested list representation of our custom \inertF{} (see the previous subsection~\ref{subsec:maple-probs}).
-Since the \inertF{} is similar to an expression tree, it can be translated as the \gls*{mlp-pt} in the forward translation.
-Hence, we use the same approach of multiple subtranslators for the backward translation process.
-
-%Figure~\ref{fig:BackwardDel} shows the subtranslators of the backward translation process.
-The structure is the same as for the forward translation, seen in figure~\ref{fig:ForwardDel}.
-The \verb|MapleInterface| represents the entry point of a translation process.
-Before the translation starts, the \Maple{} kernel needs to be initialized and prepared for the translation process by loading all custom procedures.
-Once everything is loaded, an expression is delegated to specialized subtranslators depending on the internal data type, defined by custom procedures and \Maple's \inertF.
-
-%The following figure~\ref{fig:backward-trans} illustrates the backward translation process for the Jacobi polynomial example $P_n^{(\alpha , \beta)}(\cos(a\Theta))$ from table~\ref{tab:JacobiP-usecase}.
-The input expression is converted to the nested list representation of the \texttt{MyInertForm} (the customized \inertF{} of \Maple).
-Afterwards, each node of the tree is translated separately by specialized subtranslators (visualized by blue arrows).
-The translation of functions (bold blue arrows) is again realized by translation patterns, which define the correct position for each argument.
-The last step is used to put all translated arguments to the right position (visualized by red arrows).
-
-The translation process than follows the same principle as for the forward translations.
+The translation process then follows the same principle as for the forward translations.
 Since the syntax tree of \Maple{} is an expression tree, we do not need to implement special reordering or grouping algorithms to perform backward translations.

### 2018-07-25 chapters/03-translator/forward-trans.tex
Clear unused lines and use grammarly checks

@@
 An abstract translator class analyzes each node of the parsed tree and delegates them to specialized subtranslators.
-In this process an object called \gls*{teo} is build.
+In this process, an object called \gls*{teo} is built.
 This \gls*{teo} is needed to rebuild a string representation of the mathematical expression after all nodes of the parsed tree were translated.
@@
 
-This approach is not completely feasible, since the algorithm needs to look ahead and check the following siblings in some cases, e.g., in the case of a semantic macro with arguments (see figure~\ref{fig:syntax-tree-usecase}).
+This approach is not completely feasible since the algorithm needs to look ahead and check the following siblings in some cases, e.g., in the case of a semantic macro with arguments (see figure~\ref{fig:syntax-tree-usecase}).
 The {\footnotesize \verb|TRANSLATE_LEAF|} function needs to know the following siblings of the current node.
@@
 If the root $r$ is a leaf, it still can be translated as a leaf.
-Eventually, some of the following siblings are needed in order to translate $r$.
+Eventually, some of the following siblings are needed to translate $r$.
 The list of \textit{following\_siblings} in line~\ref{line:transLeaf} might be reduced to avoid multiple translations for one node.
@@
 As already mentioned, some of these types can be translated directly, such as Greek letters, while others are more complex, such as semantic \LaTeX{} macros.
-Therefore, the translators delegates the translation to specialized subtranslators.
+Therefore, the translators delegate the translation to specialized subtranslators.
 This delegation process is implemented in \cref{line:transLeaf,line:transSeq} of algorithm~\ref{alg:translation}.
@@
 Sometimes the same symbol is used in different notations to distinguish different meanings.
-For example, the '$-$' as an unary operator is used in prefix notation to indicate the negative value of its operand, such as in '$-2$'.
+For example, the '$-$' as a unary operator is used in prefix notation to indicate the negative value of its operand, such as in '$-2$'.
 Of course, $-$ can also be the binary operator for subtraction, which is commonly used in infix notation.
@@
 
-Most programming languages (and \gls*{cas} as well) internally use prefix or postfix notation and do not mix the notations in one expression, since it more convenient to parse expressions in uniform notations.
+Most programming languages (and \gls*{cas} as well) internally use prefix or postfix notation and do not mix the notations in one expression, since it is more convenient to parse expressions in uniform notations.
 However, the common practice in science is to use mixed notations in expressions.
@@
 In general, this is not a problem for our translation process since most \gls*{cas} are familiar with the most common notations.
-Therefore, the translator does not need to know that $a$ and $b$ are the operands of the binary operator '$+$' in '$a+b$'.
-The translator could simply translates the symbols in '$a+b$' in the same order as they appear in the expression and the \gls*{cas} would understand it.
+Therefore, the translator does not need to know that $a$ and $b$ are the operands of the binary operator '$+$' in '$a+b$.' The translator could simply translate the symbols in '$a+b$' in the same order as they appear in the expression and the \gls*{cas} would understand it.
 However, this simple approach generates two problems.
@@
 If a user wants to support a \gls*{cas} that uses prefix or postfix notation by default, the translator would fail in its current state.
-Supporting \gls*{cas} with another notations would be a part of future work.
+Supporting \gls*{cas} with another notation would be a part of future work.
 
@@
 It has no access to previously translated nodes in its current state.
-This problem is solved by the \gls*{teo} that stores and groups translated objects as lists.
+This problem is solved by the \gls*{teo} that stores and groups translated objects like lists.
 This allows to access the latest translated expression and use it as the argument for the factorial function.
@@
 
-%The biggest problem about translating the factorial (or double factorial) function is the common postfix notation for them.
-In this stage, algorithm~\ref{alg:translation} only translates the current node and if necessary the following siblings, but not the preceding siblings (see algorithm~\ref{alg:translation}).
-Consider an expression such as '$4!$'.
-When the variable node $r$ reaches the exclamation mark, the argument '$4$' of the function has already been translated and is no longer accessible for the current algorithm.
-This problem occurs whenever a postfix notation is involved.
-The following two approaches would solve the problem for the problem.
-
-%\begin{enumerate}
-%\item\label{appr:1} The translator always checks if the next two following siblings are one or two exclamation marks.
-%\item\label{appr:2} The current state of program has access to the previously translated expressions and can modify them retrospectively.
-%\end{enumerate}
-
-%Approach~\ref{appr:1} would be unnecessarily inefficient and does not scale well if more of these cases occur.
-Therefore we implement approach~\ref{appr:2}.
-That is another reason why we implement an extra object to organize translated expressions, the \gls*{teo}.
-This object stores all parts of the previously translated expressions.
-The translator has access to these parts and is able to modify them afterwards.
-
-%\input{chapters/03-translator/all-types-table.tex}
-
-%\subsection{Delegation of Translations}\label{subsec:delegation}
-
-%\begin{wrapfigure}{l}{0.45\textwidth}
-%	\vspace{-20pt}
-%	\centering
-%	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.4]{ForwardTranslator.pdf}
-%	\caption{A scheme of the forward translator and its specialized subtranslators.}
-%	\label{fig:ForwardDel}
-%	\vspace{-25pt}
-%\end{wrapfigure}
-
-%Algorithm~\ref{alg:translation} is a simplified version of the translator process.
-The main progress happens in \cref{line:transLeaf,line:transSeq}.
-There are several cases what $r$ can be.
-Table~\ref{tab:allTypesTable} on page \pageref{tab:allTypesTable} gives an overview of all different types in a \gls*{pom-pt}.
-A more detailed explanation of the types can be found in~\parencite{POM-Tagger}.
-
-%The \gls*{bnf} grammar defines some basic grammatical rules for generic \LaTeX{} macros, such as for \verb|\frac|, \verb|\sqrt|.
-Therefore, there is a hierarchical structure for those symbols, similar to the structure in expression trees.
-As already mentioned, some of these types can be translated directly, such as Greek letters, while others are more complex, such as semantic \LaTeX{} macros.
-Therefore, the translators delegates the translation to specialized subtranslators.
-This delegation process is implemented in \cref{line:transLeaf,line:transSeq} of algorithm~\ref{alg:translation}.
-
-%Figure~\ref{fig:ForwardDel} is a scheme of the forward translator and its subtranslators.
-The entry point for each translation process of the program is the \verb|SemanticLatexTranslator|.
-It delegates the translations for the expression through the \verb|AbstractTranslator| to the other subtranslators.
-The \verb|AbstractTranslator| is the interface of all other translators and coordinates the translation process by delegating translations for specific nodes to specialized classes.
-For example, the \verb|MacroTranslator| only translates semantic \LaTeX{} macros.
-The \verb|EmptyExpressionTranslator| is used to delegate the so called \textit{empty expressions} to other classes.
-An \textit{empty expression} is a node with children, but without a symbol itself, such as \textit{sequences} and \textit{balanced expressions}.
-
 \subsection{Subtranslators}\label{sec:subtranslators}
-%The most interesting translators are the \verb|MacroTranslator| (translating semantic \LaTeX{} macros) and the \verb|SequenceTranslator| (translating all kinds of \textit{sequences}).
 
@@
 Whether an expression is an interval or another expression is difficult to decide and can depend on the context.
-In addition, the parenthesis checker could simply be deactivated to allow mismatched parentheses in an expression.
+Also, the parenthesis checker could simply be deactivated to allow mismatched parentheses in an expression.
 But this functionality is not implemented yet.
@@
 The \verb|SequenceTranslator| checks the types of the current and next nodes in the tree to decide if it should add a multiplication symbol or not.
-For example, if the current or next node is an operator, a relation symbol or an ellipsis there will be no multiplication symbol added.
+For example, if the current or next node is an operator, a relation symbol or an ellipsis, there will be no multiplication symbol added.
 However, this approach implies an important property.
@@
 In this case, the user should use \verb|\f{f}@{x}| instead of \verb|f(x)| to distinguish between $f$ as a function call and $f$ as a symbol.
-
-%Besides the translation of sequences, the translation process of the \Macro s can be complex too; not only because of the structure of the \gls*{pom-pt}, but also because of the complex definition of the semantic macros.
-As explained in \ref{subsec:macros}, they can contain optional parameters, exponents are allowed before and after the arguments and the number of $@$ symbols varies.
-All these different cases will be solved by the \verb|MacroTranslator|.
 
@@
 It extracted necessary information from the \gls*{pom-pt}, such as how many arguments this function has.
-In consequence it also processes the following siblings to translate the arguments.
+In consequence, it also processes the following siblings to translate the arguments.
 There are some special cases about the next node right after the macro occurs.
@@
 \end{equation}
-
-%An open square bracket right after the semantic macro (after the exponent respectively) is an optional parameter.
-Since expressions in square brackets are not considered to be a \textit{sequence} in the \gls*{pom-pt}, the optional parameter is not grouped in a single node.
-Therefore, every node after the opening square bracket forms an expression for just one optional parameter until the algorithm reaches the closing square bracket.
-Such expressions in square brackets will be also translated by the \verb|SequenceTranslator| as described above.
-A semantic macro could have multiple optional parameters.
-Therefore, the function translates all optional parameters and stores them in line~\ref{line:next_optional}.
-
-%The \cref{line:trans_paras,line:skip_ats,line:trans_vars} translates the parameters and variables.
-The $@$ symbols will be skipped.
-Nevertheless, the current program has an error handling implemented in \cref{line:skip_ats} if the maximum number of $@$ symbols is exceeded.
-In the following lines, the function takes the translation pattern from the lexicon and removes each placeholder by the corresponding argument.
-In addition, if there is an exponent right after the semantic macro, it will be attached to the translated semantic macro (see table~\ref{tab:multi-expo}).
-The parameters and variables are always grouped by curly brackets in \textit{sequence}-nodes in the \gls*{pom-pt}.
-Therefore, every argument after the optional parameters and potential exponent cannot be longer than one node in the \gls*{pom-pt}.
-
-%\subsection{Handle Ambiguities}\label{subsec:ambiguities}
-%Our semantic \LaTeX{} expressions are not full semantic with respect to possible ambiguities in the expression.
-We already mentioned some techniques to solve ambiguities in semantic \LaTeX{} expressions, for example how to handle double factorials.
-In table~\ref{tab:amb-latex} are four examples of ambiguous expressions.
-\LaTeX{} implemented the rules that the superscript or subscript is only one symbol (or a balanced expression in curly brackets).
-Therefore, each of these input expressions are not ambiguous in \LaTeX.
-Since we talking about the forward translation, we should follow the rules of \LaTeX{} to interpret ambiguous expressions.
-
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{cc}
-%	\hline
-%	Ambiguous Input & \LaTeX{} Output\\
-%	\hline
-%	\verb|n^m!| & $n^m!$\\
-%	\verb|a^bc^d| & $a^bc^d$\\
-%	\verb|x^y^z| & Double superscript error\\
-%	\verb|x_y_z| & Double subscript error\\
-%	\hline
-%\end{tabular}
-%\caption{Ambiguous \LaTeX{} expressions and how \LaTeX{} displays them.}
-%\label{tab:amb-latex}
-%\end{table}
-
-%Another more questionable translation decision are alphanumerical expressions.
-As explained in table~\ref{tab:allTypesTable}, the \gls*{mlp} handles strings of letters and numbers differently, depending on the order of the symbols.
-The reason is, that an expression such as '$4b$' is usually considered to be a multiplication of $4$ and '$b$', while '$b4$' looks like indexing '$b$' by $4$.
-While the first example produces two nodes, namely $4$ and '$b$', the second example '$b4$' produces just a single alphanumerical node in the \gls*{pom-pt}.
-The first expression would be translated as a multiplication of $4$ and '$b$'.
-Because '$4b$' looks like '$4 \cdot b$' and the multiplication is commutative, we would assume that '$4b$' and '$b4$' are equivalent.
-This is one reason why we interpret such alphanumerical expressions as multiplications of the symbols.
-On the other hand, an expression such as '$energy$', would be an alphanumerical node as well and could be interpreted as a single variable named 'energy'.
-In that case, someone could wonder why the variable '$energy$' was translated as '$e \cdot n \cdot e \cdot r \cdot g \cdot y$'.
-While in physics or engineering variables possibly appear with longer names, such as the 'energy' example, it is more common to use single symbols for variables in mathematics~\parencite{Notation:History}.
-Therefore, we interpret such alphanumerical expressions as multiplications of variables rather than one variable with an alphanumerical name.
-
-%Another, more elegant way to solve this problem would be to use the newly created semantic macro \verb|\idot| (see subsection~\ref{subsec:semanticVSmacro}) to define an invisible multiplication symbol.
-However, because \verb|\idot| is such a new semantic macro, it is not used in the \gls*{dlmf} or \gls*{drmf} quite often yet.
-Therefore, our program still interprets alphanumerical expressions as multiplications.
-
-%Other ambiguities can appear, when the input expression contains a symbol, which is typically used to represent a specific mathematical object, but the user did not use the semantic macro to represent it.
-For example, the input expressions '$3i$' contains an '$i$', which is usually associated with the imaginary unit.
-Since there is a semantic macro for the imaginary unit, namely \verb|\iunit|, we do not translate the letter to the imaginary symbol in \Maple{} and use the identity translation instead.
-Furthermore, we inform the user about the potential misusage of \verb|i| (see~\ref{subsec:add-info}).
-
-%In general the translator is drafted to solve ambiguous expressions or automatically find a workaround to disambiguate the expression.
-Only if there is no way to solve the ambiguity with the defined rules, the translation process stops.
-
-%\subsection{Translated Expression Objects}\label{subsec:translated-expr}
-%\glsreset{teo}
-%\begin{wrapfigure}{r}{0.5\textwidth}
-%	\vspace{-20pt}
-%	\centering
-%	\includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm, scale=1]{SyntaxTreeGroup.pdf}
-%	\caption{The pom-pt for '$\cos(2n\pi)!$' and the grouped argument of the factorial function.}
-%	\label{fig:syntax-group}
-%	\vspace{-15pt}
-%\end{wrapfigure}
-
-%It is sometimes necessary to provide access to already translated symbols or subexpressions.
-A typical example is the (double) factorial function, where the argument was translated before the function was registered by the translator (postfix notation).
-To realize this, the \gls*{teo} is implemented.
-It is mainly a list object of translated expressions.
-It also groups subexpressions when it becomes clear that multiple symbols belong together.
-This typically happens when something is wrapped in parentheses but also when a function is completely translated.
-
-%Figure~\ref{fig:syntax-group} shows an example for the expression '$\cos(2n\pi)!$'.
-If we just analyze the \gls*{pom-pt}, the previous node of the exclamation mark (which indicates the factorial function) is the sequence.
-Now it could be possible to misinterpret the sequence as the argument of the factorial function.
-But the translator will first translate the cosine function, which has one argument.
-In that case the argument is the sequence node.
-Once the cosine function has been completely translated (which means the whole argument has been translated as well) the \gls*{teo} groups the translated cosine and the sequence.
-Therefore, when the translator reaches the exclamation mark and asks for the last translated object in the list, it is not the translated sequence but the whole cosine function.
-
-%Table~\ref{tab:teo-list} shows some examples of the \gls*{teo} list after the translation process was finished.
-Note that a translation of '$a+b$' contains three translated objects, while '$(a+b)$' contains just one.
-
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{cc}
-%	\hline
-%	Input Expression & TEO List\\
-%	\hline
-%	$a+b$ & \verb|[a, +, b]|\\
-%	$(a+b)$ & \verb|[(a+b)]|\\
-%	$\frac{a}{b}-2$ & \verb|[(a)/(b), -, 2]|\\
-%	\hline
-%\end{tabular}
-%\caption{How the TEO-list groups subexpressions.}
-%\label{tab:teo-list}
-%\end{table}
-
-%\subsubsection{Local \& Global Translated Expressions}
-%During the translation process multiple local \gls*{teo} are created by the subtranslators.
-Besides the local \gls*{teo} there is just one global \gls*{teo}.
-The reason to create local objects is originally motivated by the recursive structure of the translator.
-A subtranslator should only handle the current subexpression.
-However, the global \gls*{teo} becomes necessary because of the functions in postfix notation such as the factorial function.
-For example, when a subtranslator tries to translate the exclamation mark as the factorial function, the local \gls*{teo} is empty, because the argument of the function has been previously translated by another subtranslator and stored in another local \gls*{teo}.
-Therefore, we implement a global \gls*{teo}, which contains all of the previous translated subexpressions in each state of the translation process.
-With this implementation the local \gls*{teo} becomes more and more obsolete.
-In the current version, the program stores redundant translated expressions.
-For debugging reasons, a translation process still handles multiple local and one global \gls*{teo}.
-Starting the forward translation with the flag \verb|--debug| (for the debugging mode) shows the elements of the global \gls*{teo} after the translation has been finished.
-
-%\subsection{Additional Information}\label{subsec:add-info}
-%As already mentioned, a translation is not always straight forward. In those cases, the user should get informed about the translation and the reason for it. Consider the translation
-%\begin{equation}\label{eq:i-trans}
-%\verb|3i| \overset{\langMaple}{\mapsto} \verb|3*i|,
-%\end{equation}
-%Since the imaginary unit in \Maple{} is associated with '\raw{I}' instead of '\raw{i}', a user might be confused by~\ref{eq:i-trans}.
-Therefore, we created the \texttt{InformationLogger} class to organize further information about the translation process.
-In the case of '\raw{i}', the user is informed about the semantic macro \verb|\iunit| and the potential misusage of '\raw{i}'.
-
-%Essentially, the \raw{InformationLogger} provides further information about a translation process.
-This information is defined in the lexicon files.
-Consider the example of the arccotangent function, which is defined with a different branch cut in \gls*{dlmf} and in \Maple.
-As already mentioned, we provide alternative translation patterns to solve these problems.
-These alternative translations, explanations, branch cuts, domains etc.
-will be automatically stored in the \raw{InformationLogger} when an expression contains a mathematical object with further information.

### 2018-07-25 chapters/03-translator/gen-app.tex
Clear unused files, upload highres images and fix errors
Clear unused lines and use grammarly checks

@@
 \section{Translation Problems}\label{sec:problems}
-There are several potential problems to peform translations between systems that embed semantic information in the input.
-Those problems vary from simple cases, e.g., a function is not defined in system, to complex cases, e.g., different positioning of branch cuts for multivalued functions.
+There are several potential problems to perform translations between systems that embed semantic information in the input.
+Those problems vary from simple cases, e.g., a function is not defined in the system, too complex cases, e.g., different positioning of branch cuts for multivalued functions.
 This section will discuss the problems and workarounds.
@@
 Since those translations are nonintuitive, describing explanations become necessary for the translation process.
-A special logging function take care about each translation and provide details after a successful translation process.
+A special logging function takes care of each translation and provide details after a successful translation process.
 Section~\ref{sec:forward-translation} explains this task further.
@@
 
-A similar problem are multiplications, since they are rarely explicitely marked in \LaTeX{} expressions, e.g., scientists using whitespaces to indicate multiplications rather than using \verb|\cdot| or similar symbols.
+A similar problem is multiplications since they are rarely explicitly marked in \LaTeX{} expressions, e.g., scientists using whitespaces to indicate multiplications rather than using \verb|\cdot| or similar symbols.
 For such problems, we introduced a new macro \verb|\idot| for an invisible multiplication symbol (this macro will not be rendered).
-Since this macro is newly introduced and an automatic conversion of existing equations is difficult, none of the equations in the \gls*{dlmf} uses this macro yet.
-As a consequence, the translator has some simple rules to perform translations without explicitely marked translations with \verb|\idot|.
+Since this macro is newly introduced and automatic conversion of existing equations is difficult, none of the equations in the \gls*{dlmf} uses this macro yet.
+As a consequence, the translator has some simple rules to perform translations without explicitly marked translations with \verb|\idot|.
 
-Even we only allow the special dialect of \LaTeX{} using semantic macros not all expressions are unambigious.
+Even we only allow the special dialect of \LaTeX{} using semantic macros not all expressions are unambiguous.
 In table~\ref{tab:amb-latex} are four examples of ambiguous expressions.
-These expressions are unambigious for the \LaTeX{} compiler, since it only considers the very next token for power and subscripts.
+These expressions are unambiguous for the \LaTeX{} compiler since it only considers the very next token for power and subscripts.
 Our translator following the same rules to solve these issues.
@@
 
-Another more questionable translation decision are alphanumerical expressions.
+Another more questionable translation decision is alphanumerical expressions.
 As explained in table~\ref{tab:allTypesTable}, the \gls*{pom}-tagger handles strings of letters and numbers differently, depending on the order of the symbols.
-The reason is, that an expression such as '$4b$' is usually considered to be a multiplication of $4$ and '$b$', while '$b4$' looks like indexing '$b$' by $4$.
+The reason is, that an expression such as '$4b$' is usually considered to be a multiplication of $4$ and '$b$,' while '$b4$' looks like indexing '$b$' by $4$.
 While the first example produces two nodes, namely $4$ and '$b$', the second example '$b4$' produces just a single alphanumerical node in the \gls*{pom-pt}.
 The translator interprets alphanumerical expressions as multiplications for two reasons.
-(1) we would assume that the inputs '$4b$' and '$b4$' are mathematically equivalent and (2) it is more common in mathematics to use single letter names for variables~\parencite{Notation:History}.
+(1) We would assume that the inputs '$4b$' and '$b4$' are mathematically equivalent and (2) it is more common in mathematics to use single letter names for variables~\parencite{Notation:History}.
 Therefore we define the following definitions
@@
 
-In general the translator is drafted to solve ambiguous expressions or automatically find a workaround to disambiguate the expression.
+In general, the translator is drafted to solve ambiguous expressions or automatically find a work-around to disambiguate the expression.
 Only if there is no way to solve the ambiguity with the defined rules, the translation process stops.
@@
 \section{The Translator}
-%The first step for providing an automatic translation tool is to defining the translation for each function.
-We manually create a table of translations for each \Macro.
-Translations for generic \LaTeX{} functions, mathematical symbols, Greek letters and constants are organized in \gls{json} files for a better performance.
-
-%For each translation, we design a translation pattern with placeholders to define the correct position of each argument.
-The placeholders are defined as \verb|$i|, where \verb|i| is a non-negative integer specifying the position, starting at $0$.
-Table~\ref{tab:placeholder_ex1} illustrates the translation patterns for the trigonometric sine function in semantic \LaTeX, \Maple{} and \Mathematica.
-%\begin{table}[ht]
-%	\centering
-%	\begin{tabular}{lc}
-%		\hline
-%		Semantic \LaTeX & \verb|\sin@{$0}| \\
-%		\Maple & \verb|sin($0)| \\
-%		\Mathematica & \verb|Sin[$0]|\\
-%		\hline
-%	\end{tabular}
-%	\caption{Translation patterns for the sine function in the \DLMF, \Maple{} and \Mathematica.}
-%	\label{tab:placeholder_ex1}
-%\end{table}
-
 All translations are defined by a library (\gls*{csv} and \gls*{json} files) that defines translation patterns for each function and symbol.
@@
 
-%Placeholders gives us the opportunity to define the order of the arguments and handle nested function calls.
-The order of arguments can differ between the systems.
-For example, if a function possesses parameters written in sub- and superscripts, \Maple{} and \Mathematica{} usually put the parameters in the subscript before the parameters in the superscripts.
-The notation in the \Macro s uses the reversed order.
-Table~\ref{tab:JacobiP-usecase} in the introduction shows an example of such different orders.
-The translation patterns for the Jacobi polynomial are illustrated in table~\ref{tab:placeholder_ex2}.
 \begin{table}[ht]
@@
 
-%Replace the placeholders sequentially by the arguments ends in
-%\begin{eqnarray*}
-%&&\verb|diff($1, [$2$$0])|\\
-%\text{Replace index 0 by $2$\hphantom{$^2$}}& \rightarrow &\verb|diff($1, %[$2$2])|\\
-%\text{Replace index 1 by $x^2$}& \rightarrow &\verb|diff(x^2, [$2$2])|\\
-%\text{Replace index 2 by $x$\hphantom{$^2$}}& \rightarrow &\verb|diff(x^2, [xx])|.
-%\end{eqnarray*}
-
 We can solve this issue by using paranthesis in such cases, e.g., \verb|diff($1, [$2$($0)])|.
 
-%A special problem can be binary operators.
-A good example is the multiplication sign.
-In \LaTeX{} most scientist would use white spaces or \verb|\cdot| to indicate a multiplication (using none of them can cause ambiguities, see \ref{subsec:ambiguities}).
-Similar to that approach, \Mathematica{} also understands white spaces and asterisks as multiplications.
-When we translate semantic \LaTeX{} expressions to \Maple, we use the {\sf 1D \Maple{}} representation (see subsection~\ref{subsec:maple}), where an asterisk becomes mandatory for multiplications.
-White spaces would produce a syntactical error in this input format.
-On the other hand, white spaces in \LaTeX{} do not necessarily indicate a multiplication.
-It could also be used to improve the readability.
-Whether a white space indicates a multiplication or not is a semantic information in the expression.
-We introduce a new \DLMF{} macro that indicates a multiplication, but not gets rendered.
-This macro is \verb|\idot|, for an invisible dot.
-
-%As already mentioned, the information is stored in three different file types. Together they form the backend library.
-
-%\subsubsection{CSV Tables}
-%The \gls{csv} file represents the translation tables generated by MS Excel.
-Although \gls{csv} is an abbreviation for \textit{comma-seperated}, one can specify the split symbol.
-Most common are semicolons.
-In our case we use semicolons, because commas are frequently used in string representations of mathematical functions to separate multiple parameters and variables.
-Note that semicolons are also sometimes used, but \gls{csv} files are able to use quotation marks to separate a value with semicolon from the split symbols.
-
-%Note that \gls{csv} files are databases.
-While we do not want to introduce the theory of databases in this thesis, note that our databases contain a \textit{primary key}.
-A primary key uniquely identifies and provides access to information in the database.
-The main \gls{csv} file is called \verb|DLMFMacro.csv|.
-It provides information about each macro.
-The primary key of this database is the macro itself.
-Further information consists of the name of the mathematical function, the hyperlink to the \DLMF{} definition, the numbers of parameters, variables and the maximum number of $@$ symbols.
-The maximum number of $@$ symbols is mostly used for backward translations.
-Additionally, the \gls{csv} file can contain information about the constraints, branch cuts and a role.
-The additional role can be used to ignore the macro in the translation process, or to specify that this macro is a mathematical constant.
-For example, mathematical constants are ignored, because they are organized in the \gls{json} files.
-This avoid multiple definitions in the database.
-
-%Since a translation is not necessarily bijective, the translations are defined in two separate \gls{csv} files per \gls{cas}.
-For \Maple{} those files are \verb|DLMF_Maple.csv| for the forward translation, and \verb|CAS_Maple.csv| for the backward translation.
-Both are organized in the same already described logic.
-
-%In~\ref{eq:guder-trans} we defined the definition for the Gudermannian function.
-The translated expression contains two function calls: the inverse tangent function and the hyperbolic sine function.
-Instead of providing a hyperlink to the definitions for both functions on the \Maple{} help page, we only provide one hyperlink.
-We specify the hyperlink by a special prefix in the translation pattern.
-To explain this prefix, we need to take a look to optional parameters before.
-While the primary key in the \gls{csv} files is only the macro name, we need to specify the number of optional parameters in front of the macro name.
-Otherwise, for example, the Legendre and the associated Legendre function of the first kind would be linked to the same definition.
-Our prefix notation solve both problems.
-It links to one specific function for the translation and specify the correct version of the macro, if there are optional parameters included.
-The prefix starts and ends with an $X$ and contains two information.
-The number of optional parameters (if none, than this number will be $0$) and the name of the used function.
-Table~\ref{tab:legendreP-lex} illustrates the lexicon entry of the Legendre and the associated Legendre function of the first kind.
-
-The \Macro s also allow to specify optional arguments to distinguish between standard and other version of functions.
+The \Macro s also allow to specify optional arguments to distinguish between standard and another version of functions.
 The Legendre and associated Legendre function of the first kind are examples for such cases.
@@
 The Legendre and associated Legendre function of the first kind both using the same macro \verb|\LegendreP|.
-To distuingish such cases we using a special syntax such cases, shown in table~\ref{tab:legendreP-lex}.
+To distinguish such cases, we are using a special syntax such cases, shown in table~\ref{tab:legendreP-lex}.
 
@@
 Each token will be tagged by meta information defined in lexicon files.
-We extend the lexicon files to provide also the information that are necessary for the translation process.
-An example for an entry of the lexicon file is given in table~\ref{tab:sine-lex-example}.
-
-%\subsubsection{Lexicon Files}
-%The knowledge of the \gls{mlp} is based on lexicon files, as explained in subsection~\ref{subsec:pom-mlp}.
-Since the \gls{mlp} project was not a part of the \Macro s project, the \gls{bnf} grammar defines no rules about semantic macros.
-Furthermore, the knowledge database has no information about the semantic macros.
-Therefore, we created a new lexicon file that defines each semantic \LaTeX{} macro and provides further information about it.
-This information is stored in the previously mentioned \gls{csv} files.
-
-%Table~\ref{tab:sine-lex-example} presents the entry for the sine function in the \verb|DLMF-macros-lexicon.txt|.
-The \verb|global-lexicon.txt| file is the main dictionary of the \gls{mlp} and combines the whole knowledge of the parser in one file.
-This lexicon files also help to translate ambiguous expressions.
-We will discuss some of these decision in the following sections.
-However, it is important to mention that a successful translation strongly depends on the comprehensive database of the lexicon files.
+We extend the lexicon files to provide also the information that is necessary for the translation process.
+An example of an entry of the lexicon file is given in table~\ref{tab:sine-lex-example}.
 
@@
 \end{table}
-
-%The concept of the lexicon files is also adapted to store the backward translations from \Maple.
-Since the \gls{mlp} project is developed to parse \LaTeX{} expressions, it could be confusing to use the original lexicon files to also store names of \Maple{} functions.
-Therefore, we create another type of lexicon files, which are very similar to the original lexicon files.
-In \Maple{} a function can have multiple entries differing in the number of arguments.
-Therefore, the primary key of the backward translation database is the combination of the function name and the number of arguments.
-For example, the name \verb|GAMMA| in \Maple{} is reserved for the Euler Gamma function in case of one argument, but also for the incomplete Gamma function in the case of two arguments.
-Therefore the lexicon file for the backward translation also contains the number of arguments to specify a translation.
-
-%\subsection{Extending the Forward Translator}\label{subsec:extension}
-%The whole project is still work in progress.
-Therefore, it is an important goal to keep it easy to update the backend libraries.
-This section will explain the typical workflow to add new translations for semantic \LaTeX{} macros and \Maple{} functions or how to support a forward translation to another \gls{cas}.
-
-%\subsubsection{Adding Translations to Existing CAS}
-%Since \gls{csv} files are not as easy to maintain as MS Excel\footnote{Microsoft Excel is widely used to work with tables and is able to import, export and manipulate \gls{csv} files.} worksheets, the workflow starts with MS Excel.
-Once a user updates the existing tables, the MS Excel file needs to get exported as a \gls{csv} file.
-The translator project contains the \verb|lexicon-creator.jar| file to convert the \gls{csv} files to the backend lexicon files.
-Of course, these steps are not necessary if the user just changes a translation in the \gls{json} files.
-
-%Once the \verb|lexicon-creator.jar| finished the conversion process, the translator is aware of all changes and ready to use.
-
-%\subsubsection{Supporting New CAS}
-%To support a complete new \gls{cas} for the forward translation, the user needs to add the \gls{cas} in the \gls{json} files and create two new \gls{csv} files.
-Even if there is no backward translation implemented yet, the \gls{csv} file for the backward translation is needed.
-
-%Once all of the files are updated, the \verb|lexicon-creator.jar| can be used to update the lexicon files.

### 2018-07-25 chapters/04-evaluation/evaluation.tex
Clear unused lines and use grammarly checks

@@
 \section{Evaluation}\label{sec:evaluation}
-%This project aims for a translation of mathematical expressions from \LaTeX{} sources to \gls*{cas} and back again.
-Essentially, the translator translates only string representations.
-We need to verify that such translations are \textit{appropriate}.
-Therefore, we develop verification and validation techniques for translations and present them in this chapter.
-
-%First of all, each semantic \LaTeX{} macro ties a specific character sequence to a well-defined mathematical object.
-This definition can be found at the \gls*{dlmf} or \gls*{drmf}.
-Therefore, our reference source for semantic \LaTeX{} expressions is always the \gls*{dlmf} or \gls*{drmf}.
-This gives us the opportunity to compare more than the string representations of the mathematical objects in a translation process.
-For example, consider the simple translation of the cosine function
-%\begin{equation}\label{eq:simple-cosine}
-%\verb|\cos{x}| \overset{\langMaple}{\mapsto} \texttt{cos(x)}.
-%\end{equation}
-
-%Our verification and validation techniques, to proof for example that~(\ref{eq:simple-cosine}) is an \textit{appropriate} translation, can be grouped into three parts.
-We present in the following sections these three parts and conclusively, we use all these techniques to actually verify translations for a test suite directly extracted from the \gls*{dlmf}.
-
 We implemented three approaches to evaluate whether a translation was \textit{appropriate} or \textit{inappropriate}.
@@
 \subsection{Round Trip Tests}\label{sec:round-trip}
-%Round trip tests are our first approach to verify translations.
-The idea is to translate an expression back and forth again and check if the representation changes during this process.
-The translator is able to translate expressions from and to \Maple.
-Therefore, we implement the following round trip tests only for \Maple.
-
-%Furthermore, we take advantage of \Maple's symbolic simplification function to compare a \Maple{} expression in the beginning of an round trip test with the \Maple{} expression after the translations.
-If \Maple's \textit{simplify} function returns zero for the difference between two representations, both expressions are symbolically equivalent.
-Note that this relies on the correctness of the simplification, which is not proven (see section~\ref{subsec:mathematica}).
-On the other hand, the equivalence is not disproven if the simplification returns something different to zero.
-
 A round trip test always starts with a valid expression either in semantic \LaTeX{} or in \Maple.
@@
 \end{definition}
-
-%Obviously, there may appear two fixed point representations in round trip tests, one for each system.
-We will see that most of the round trip translations reach fixed points after a certain amount of steps.
-
-%\begin{theorem}[Fixed Point Properties]\label{theo:fixpoint}
-%Consider two languages $A,B \in \mathcal{L}$.
-A fixed point representation is reached if the current representation is just identical to the next representation in the same language (after one cycle).
-Furthermore, if $a \in A$ is a fixed point representation, the translation of $a$ to $b \in B$ is also a fixed point representation.
-%\end{theorem}
-%\begin{proof}
-%Let $A, B \in \mathcal{L}$ be languages.
-Consider a round trip translation test between the languages $A$ and $B$.
-Without loss of generality, let $a \in A$ be a fixed point representation and let $a$ be the \textit{first} fixed point representation that appears in this test.
-The \gls*{mlp} as well as \Maple{} are deterministic, because a test case works in only one \Maple{-session} and we do not compute the inputs in \Maple.
-Therefore, the translation process is not able to call any probabilistic algorithms.
-Hence, our translator is also deterministic.
-
-%Let $b \in B$ with
-%\begin{equation}
-%\mathrm{tr}^A_B(a) = b.
-%\end{equation}
-
-%Since the translation process is deterministic, $a$ will be always translated to $b$. Consider $a' \in A$ as the backward translation from $b$
-%\begin{equation}
-%\mathrm{tr}^B_A(b) = a'.
-%\end{equation}
-%Now $a$ was defined to be a fixed point representation in $A$. Therefore, $a$ is identical to all string representation in the following cycles. Thus $a = a'$.
-
-%Furthermore, because the translations are deterministic, $b$ will be always backward translated to $a' = a$ and $a$ will be always forward translated to $b$.
-Therefore, $b$ is identical to the representation after one cycle.
-Per definition, $b$ is therefore a fixed point in $B$.
-
-%Remember, $a$ was chosen arbitrarily as the first fixed point representation that appears in the round trip test.
-Therefore, whenever a fixed point appears in a round trip test, the next translation to the respective other language will be a fixed point as well.
-Consequently, a fixed point is already reached, when just the next representation (after one cycle) is identical to the current representation.
-%\end{proof}
-
-%With theorem~\ref{theo:fixpoint}, we are able to find fixed points in round trip tests by only testing the equality to the next representation (after one cycle).
-Furthermore, we only need to find the first fixed point representation in either system, because the next step will be a fixed point representation for the respective other system.
-When a round trip test reaches such a fixed point in \Maple, the \textit{simplify} function tries to simplify the difference between the first expression in \Maple{} and the fixed point expression.
-It is not feasible to simplify or compute \LaTeX{} expressions yet (neither generic nor semantic \LaTeX) - actually, our translation tool just makes this possible.
-Therefore, we check the equivalence between a fixed point and its first expression in semantic \LaTeX{} manually.
 
@@
 
-Step 4 is identical to step 2 and since the translator is a deterministic algorithm it can be easily shown that step 2 and step 3 are fixed point representations for semantic \LaTeX{} and \Maple.
-
-%Almost all test cases are reaching a fixed point representation.
-Starting from semantic \LaTeX, the first fixed point is reached at step 2 in semantic \LaTeX, because of \Maple's internal changes.
-If the structure of the semantic \LaTeX{} input is already identical to the internal structure of \Maple's representation and the expression only contains functions and symbols with an available direct translation from and to each system, the input expression is already the fixed point representation.
-For example, consider the semantic \LaTeX{} expression in step 2 in table~\ref{tab:fixpoint} as the input expression.
-In that case the input is already a fixed point.
-Starting a round trip test from \Maple, the first fixed point representation is reached in step 1 in semantic \LaTeX, because \Maple's changes has already been performed after the first translation.
-Again, this is only true if the expression only contains functions and symbols with a direct translation from and to each system.
-Furthermore, if a test case reaches a fixed point, it is equivalent (checked by \Maple's \textit{simplify} function or manually) to the input expression.
+Step 4 is identical to step 2, and since the translator is a deterministic algorithm it can be easily shown that step 2 and step 3 are fixed-point representations for semantic \LaTeX{} and \Maple.
 
@@
 
-and round trip translations will produce infinite chains of sine and inverse sine calls.
-Since there are no evaulations involved, these chains will not be reduced during the translations.
-%The problem is that the round trip tests prevent simplifications in \Maple{} during the translation process (see \cref{subsec:maple-probs}).
-This problem could be resolved by allowing automatic evaluations in the round trip tests.
-However, currently we do not allow those evaluations because the changes can be serious, which makes it hard to follow possible errors.
-
+And round-trip translations will produce infinite chains of sine and inverse sine calls. Since there are no evaluations involved, these chains will not be reduced during the translations. 
 The round trip tests are very successful, but they only detect errors in string representations.
@@
 \subsection{Function Relation Tests}\label{sec:relation-tests}
-%In addition to the correctness of string representations, it is more important to validate whether a translation is \textit{appropriate}.
-Special functions are highly related to each other.
-The theory of special functions researches these relations and other properties of the functions.
-
 The \gls*{dlmf} is a compendium of special functions and orthogonal polynomials and lists several relations between the functions and polynomials.
-The idea of this evaluation approach is to translate an entire relation and test whether the relation remains valid after perfoming the translations.
+The idea of this evaluation approach is to translate an entire relation and test whether the relation remains valid after performing the translations.
 
@@
 is incorrect if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ and one has to use analytic continuation for the right-hand side of equation (\ref{eq:branch-cut-near}).
-%In subsection~\ref{subsec:branch-cut-issues} we also shown that the complete right-hand side of equation~(\ref{eq:branch-cut-near}) has three branch cuts, caused by the square root function and the $z^2$ argument of the modified Bessel function of the second kind.
-Such complex examples cannot be simplified yet.
-
-%However, this validation approach is powerful and runs automatically. We use this approach to verify a large set of test cases extracted directly from the \gls*{dlmf}, see \cref{sec:test-summary}.
 
@@
 \subsection{Numerical Tests}\label{sec:numerical-tests}
-%Another approach of verification are numerical tests.
-In principle, we follow the same approach as the relation tests in \cref{sec:relation-tests}, but rather than using the simplification functions we compare the left-hand side and right-hand side for actual values.
-
 Consider the differences of the left- and right-hand side of equation~\eqref{eq:branch-cut-near}
@@
 Even when all tested values return a value differ from zero, it does not proof the equivalence of~\eqref{eq:branch-cut-near}.
-However, when the values are different from zero it does proof an error in one of the four cases~\parencite{NumericalTests:Paper}
+However, when the values are different from zero, it does proof an error in one of the four cases~\parencite{NumericalTests:Paper}
 \begin{enumerate}
@@
 
-%The values are computed with the default precision of $10$ significant digits.
-Caused by the machine accuracy, the first and last values are different from zero.
-In such cases, we can increase the precision to figure out if it is an accuracy issue or $D(z)$ is really different from zero at those points.
-Increasing the precision of the computation for the first value $z = 1+\iunit$ up to $100$ significant digits returns $3\cdot 10^{-100} \iunit$ and can therefore be considered as zero (we get the same results for $z = 1-\iunit$).
-Hence, table~\ref{tab:computations-for-difference} shows us what we already know.
-The second and third values are computed on the principal branch of the modified Bessel function of the second kind and therefore produces a difference between the left-hand side and right-hand side of equation~(\ref{eq:branch-cut-near}).
-
-%This approach of verification is very powerful, because it can verify not only our translations but also equations in a mathematical compendium.
-This has been demonstrated by discovering an overall sign error of the \gls*{dlmf} equation~(\ref{eq:signerror}), see \cref{sec:test-summary}.
-
-%However, this approach has also a significant problem.
-It is easy to see that something went wrong, but very hard to proof equivalence.
-Obviously, only four tested values are not sufficient to proof the equivalence.
-How many and which values do we need for a comprehensive result?
-Reformulating this question to a general problem brings us to:
-%\begin{quote}
-%\textit{How can we discretize equations down to a finite set of test points such that the probability of equivalence for the continues equation is high if the equation is valid for the finite set of test points?}
-%\end{quote}
-%As far as we know, there is no answer to this question yet.
-This is one reason why we do not implement an automatic process of numerical tests yet.
-Another reason are constraints.
-Constraints are not as easy to translate as general formulae.
-They often contain quantifiers or logical operators which are difficult to translate in the current state of the program.
-However constraints are necessary for numerical tests.
-
 \subsection{Results}\label{sec:test-summary}
-There are 685 \Macro s\footnote{The macros are still work in progress and therefore the total number is constantly changing.} in total and 665 of them were implemented in the translator engine.
+There are 685 \Macro s\footnote{The macros are still work in progress, and therefore the total number is constantly changing.} in total, and 665 of them were implemented in the translator engine.
 We defined forward translations to \Maple{} for 201 and backward translations from \Maple{} for 195 functions.
@@
 
-The evaluation techniques has proven to be very powerful also for evaluating \gls*{cas} and mathematical online compendia such as the \gls*{dlmf}.
-During the evaluations we were able to detect several errors in the translation and evaluation engine.
+The evaluation techniques have proven to be very powerful also for evaluating \gls*{cas} and online mathematical compendia such as the \gls*{dlmf}.
+During the evaluations, we were able to detect several errors in the translation and evaluation engine.
 However, even errors in the \gls*{dlmf} were discovered.
@@
 The original constraint was given by $2\nu \neq -1, -2 -3, \ldots$, with a missing comma after $-2$.
-
-%To verify our translations with the approaches described above we create test suites by extracting equations and formulae from the sources of the \gls*{dlmf}.
-We start with a small test suite of 110\footnote{The set of test cases is available at\newline \url{https://gist.github.com/AndreG-P/f48121d70fbd2a57b2cf6d88019ddadf}} equations and provide a more detailed look to those tests.
-We are able to translate 108 of those 110 cases.
-
-%Two of the test cases (lines 42 and 43) contain macros we are not able to translate yet.
-Those macros are \verb|\subplus| and \verb|\CFK|, which are used for continued fractions.
-They have been created for another translation project~\parencite[§5]{CICM:Paper} for continued fractions defined in the \textit{Handbook of Continued Fractions for Special Functions} (2008) \parencite{CF:Handbook}.
-The conversion process for the \gls*{cfsf} \parencite{CF:Maple} dataset from \Maple{} and the \gls*{ecf} \parencite{CF:eCF,eCF:FinalReport} dataset from Wolfram uses special file types.
-Therefore, our translator is not able to handle those macros yet.
-
-%However, we translate 108 test cases and for 84 of them the simplification function of \Maple{} returns zero as the difference of the left-hand side and right-hand side.
-The 24 remaining cases are translated, but the simplification function is not able to simplify the difference to zero.
-Therefore, we try to improve this part.
-Line 46~\parencite[(18.5.10)]{NIST:DLMF} for example, can be simplified to zero, if $n$ is previously set to an arbitrary integer value.
-For all those 24 test cases, we perform also manual numerical tests, which reveal an overall sign error in equation~\parencite[(14.5.14)]{NIST:DLMF}\footnote{The equation had originally been stated as shown in equation~(\ref{eq:signerror}).
-The \gls*{dlmf} has fixed the wrong equation with version 1.0.16.}
-%\begin{equation}\label{eq:signerror}
-%\displaystyle \mathsf{Q}^{-1/2}_{\nu}\left(\cos\theta\right)=-\left(\frac{\pi}{2\sin\theta}\right)^{1/2}\frac{\cos\left(\left(\nu+\frac{1}{2}\right)\theta\right)}{\nu+\frac{1}{2}}.
-%\end{equation}
-%Besides this wrong equation, all of these manual numerical tests return zero for the difference between the left-hand side and right-hand side.
-Of course it is still not possible to conclude the equivalence based on our manual numerical tests, as explained above.
-
-%Those 110 test cases are not sufficiently comprehensive for all of the functions of the \gls*{dlmf}.
-Therefore, we create a larger dataset\footnote{We are planning to publish this dataset at \url{http://drmf.wmflabs.org}.} of 4,165 semantic \LaTeX{} formulae, which is extracted from all chapters of the \gls*{dlmf}.
-
-%We are able to translate 2,232 (approx.
-53.59\%) of those test cases.
-The \textit{simplify} function of \Maple{} returns zero for the difference for 490 of the translated 2,232 test cases.
-We increase the number to 713 with automatic pre-conversions helping the simplification process.
-With simplification over the division of the left-hand side by the right-hand side (rather than simplify the difference) we have three more verified translations and 716 in total.
-Therefore, 1,516 test cases were translated but the translation cannot be verified automatically yet.
-
-%The translator cannot translate 1,933 test cases.
-First of all, 32 lines contain neither an equation nor a relation.
-Because we have created this dataset for our verification approaches, we do not translate those 32 lines at all.
-Most of the other failures are produced because of missing semantic information in the \LaTeX{} expression or the formulae contains a macro that cannot be translated in the current state of the program, such as in 978 lines that contain a macro that cannot be translated.
-For example, in 171 lines one of the symmetric elliptic integrals~\parencite[(19.16.1-6)]{NIST:DLMF} appears and we do not provide translation patterns for those integrals, because there is no direct translation possible yet.
-A workaround would be to allow translations based on the definitions, see~(\ref{eq:gudermannian}).
-
-%Other errors appear in 914 cases, where the semantic information is missing or is not unique.
-An example are 284 lines that use a potentially ambigious single quote (or prime) notation '\aSingleQuote' to identify derivatives of functions.

### 2018-07-25 chapters/06-future-work-and-conclusion/conclusion.tex
Clear unused lines and use grammarly checks

@@
 However, simply adding more and more macros to improve the set is not the right decision.
-In that case we would end up in something similar to \sTeX, which is overloaded and too complicated to use.
+In that case, we would end up in something similar to \sTeX, which is overloaded and too complicated to use.
 If the system is no longer handy, it will not be used.
 
-On the other hand, the concept of the translator has been proven by discover errors in the online compendia \gls*{dlmf}.
+On the other hand, the concept of the translator has been proven to discovering errors in the online compendia \gls*{dlmf}.
 The test cases have also shown how difficult it is to validate a translated expression and have uncovered the problems of translations between two systems with different sets of supported functions.
 Our validation techniques also assume the correctness of the simplification and computational algorithms in \gls{cas}.
-However, combining those techniques and automatically running translation checks can potentially not only discover errors in mathematical compendia, but also detect errors in simplifications or computations in \gls{cas}.
+However, combining those techniques and automatically running translation checks can potentially not only discover errors in mathematical compendia but also detect errors in simplifications or computations in \gls{cas}.
 
-The tasks for future work are diversive.
+The tasks for future work are diverse.
 The main task to improve the translator by implementing more functions and features.
-For example, for the current state only translations to \Maple's standard function library were implemented.
+For example, for the current state, only translations to \Maple's standard function library were implemented.
 \Maple{} allows to load extra packages dynamically and therefore support several more functions.
@@
 
-The biggest weakness of the translator are still the semantic macros.
-Currently, we working on mathematical information retrieval techniques to allow an extension for the translator for generic \LaTeX{} inputs.
-
-%Finally we can conclude that the project has been successful and the translator has great potential to become a handy translation tool.
-Nonetheless, it is without any doubts an ever-growing project and needs to be improved constantly.
+The biggest weakness of the translator is still the semantic macros.
+Currently, we are working on mathematical information retrieval techniques to allow an extension for the translator for generic \LaTeX{} inputs.

### 2018-07-27 chapters/06-future-work-and-conclusion/conclusion.tex
Implemented Howard's suggestions for abstract

@@
 The test cases have also shown how difficult it is to validate a translated expression and have uncovered the problems of translations between two systems with different sets of supported functions.
-Our validation techniques also assume the correctness of the simplification and computational algorithms in \gls{cas}.
-However, combining those techniques and automatically running translation checks can potentially not only discover errors in mathematical compendia but also detect errors in simplifications or computations in \gls{cas}.
+Our validation techniques also assume the correctness of the simplification and computational algorithms in \gls*{cas}.
+However, combining those techniques and automatically running translation checks can potentially not only discover errors in mathematical compendia but also detect errors in simplifications or computations in \gls*{cas}.
 

### 2018-07-27 chapters/abstract.tex
Implemented Howard's suggestions for abstract

@@
 \noindent
-This paper presents a translation tool between representations of semantically enriched mathematical formulae in a \gls*{wp} and its corresponding representations in a \gls*{cas}.
-The representatives for the \gls*{cas} are \Maple{} and \Mathematica{}, while the representative \gls*{wp} is \LaTeX.
+We present a translation tool between representations of semantically enriched mathematical formulae in a \gls*{wp} and its corresponding representations in a \gls*{cas}.
+We chosen \Maple{} and \Mathematica{} for the \gls*{cas}, and \LaTeX as our \gls*{wp}.
 %
-The \gls*{nist} in the U.S.~has developed a set of semantic \LaTeX{} macros for \gls*{opsf}.
-The \gls*{dlmf} use these semantic macros to provide a semantically enhanced mathematical online compendium.
-This kind of semantic macros provides exclusive access to the semantic information of the functions.
-However, even if the semantics of a representation of one formula is unique, it does not need to match the semantics in another representation for the same formula.
+Bruce Miller in the U.S.~has developed a set of semantic \LaTeX{} macros for \gls*{opsf}.
+The \gls*{nist} \gls*{dlmf} uses these semantic macros to provide a semantically enhanced mathematical online compendium.
+These semantic macros usally provide exclusive access to the semantic information of the functions.
+However, even if the semantics of a representation of one formula is unique, the semantics in another representation for the same formula may differ, such as in domains of variables.
 While some distinctions are obvious, such as syntactical characteristics, others are more difficult to examine, such as differences in definitions, which leads to error-prone manual translations.
-This paper presents an automatic translation system and discusses the problems and suggests solutions. 
+We discuss difficulties related to an automatic translation system, and suggest possible solutions. 
 %
 Furthermore, this paper introduces new evaluation approaches for the developed translation tool. 
-With the help of the automatic translation tool, the evaluation experiments were able to discover errors in the \gls*{dlmf} and in the \gls*{cas} Maple.
+With the help of our automatic translation tool, the evaluation experiments were able to discover errors in the \gls*{dlmf} and in Maple.
 \end{abstract}

### 2018-07-27 chapters/01-introduction/intro.tex
Implement howard suggestions for the first 2 pages

@@
 A typical workflow of a scientist who writes a scientific publication is to use a \gls*{wp} to write the paper and one or more \gls*{cas} for verification, analysis and visualization.
-Especially in \gls*{stem} literature, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
-However, since \LaTeX{} (an extension of \TeX) is the standard to write articles in \gls*{stem} and we only focus on mathematical writings in this paper, we categorize \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the past 30 years~\parencites{LATEX:Standard}[559]{DigitalTypo}{Knuth}.
+Especially in the \gls*{stem} literature, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
+However, since \LaTeX{} (an extension of \TeX) is the de facto standard for writing articles in \gls*{stem}, and we only focus on mathematical writing in this paper, we have categorize \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the past 30 years~\parencites{Knuth}[559]{DigitalTypo}{LATEX:Standard}.
 \LaTeX{} enables printing of mathematical formulae in a structure similar to handwritten style.
-For example, consider the Jacobi polynomial~\parencite[18.3 in table 1]{NIST:DLMF}
+For example, consider the specific Jacobi polynomial~\parencite[18.3 in table 1]{NIST:DLMF}
 \begin{equation}\label{eq:P}
@@
 \end{equation}
-where $\alpha, \beta > -1$, and $n$ a nonnegative integer. 
-This formula is written in \LaTeX{} as
+where $\alpha, \beta > -1$, and $n$ is a nonnegative integer. 
+This mathematical expression is written in \LaTeX{} as
 \begin{equation*}
@@
 Hence, a writer needs to constantly translate mathematical expressions from one representation to another and back again.
-Table~\ref{tab:JacobiP-usecase} shows four different representations for the same formula~(\ref{eq:P}).
+Table~\ref{tab:JacobiP-usecase} shows four different representations for~(\ref{eq:P}).
 
@@
 	\end{tabular}
-	\caption{Different representations for the same mathematical formula~(\ref{eq:P}).
-Generic \LaTeX{} is the default \LaTeX{} expression.
-Semantic \LaTeX{} uses special semantic macros to embed semantic information.}
+	\caption{Different representations for~\eqref{eq:P}.
+Generic \LaTeX{} is the default \LaTeX{} expression; semantic \LaTeX{} uses special semantic macros to embed semantic information; and \gls*{cas} representations are unique to themselves.}
 	\label{tab:JacobiP-usecase}
@@
 
-Translations from generic \LaTeX{} to \gls*{cas} are difficult to realize since the semantic information is not explicitly given in the input.
-The \gls*{nist} has created a set of semantic \LaTeX{} macros~\cite{DLMF:Macros}.
-Each macro ties specific character sequences to a well-defined mathematical object and is linked with the corresponding definition in the \gls*{dlmf} or \gls*{drmf}.
+Translations from generic \LaTeX{} to \gls*{cas} are difficult to realize since the full semantic information is not easily constructed from the input.
+Bruce Miller at the \gls*{nist} has created a set of semantic \LaTeX{} macros~\parencite{DLMF:Macros}.
+Each macro ties specific character sequences to a well-defined mathematical object and is linked with the corresponding definition in the \gls*{dlmf}.
+The \gls*{drmf} is an outgrowth of the \gls*{dlmf} with the goal to facilitate interaction among a community of mathematicians and scientists~\parencites{DRMF:14}{DRMF:15}.
+The \gls*{drmf} extends the set of semantic macros.
 These macros embed necessary semantic information into \LaTeX{} expressions.
-One example of such a macro is given table~\ref{tab:JacobiP-usecase} for the semantic \LaTeX{} representation for the Jacobi polynomial.
+One example of such a macro is given Table~\ref{tab:JacobiP-usecase} for the semantic \LaTeX{} representation for the Jacobi polynomial.
 The macros provide isolated access to important parts of the mathematical function, such as the arguments.

### 2018-07-30 chapters/01-introduction/intro.tex
Implemented all suggestions by Howard

@@
 
-Even with embedded semantic information, a translation between the systems can be difficult.
-A typical example of complex problems occurs for multivalued functions~\parencite{AISC:MultivaluedFunctions}.
-A \gls*{cas} usually defines \textit{branch cuts} to compute principal values of multivalued functions~\parencite{Maple:Cuts}, which makes the implementation of a theoretically continues function to a discontinued presentation of it.
+Even with embedded semantic information, a translation between systems can be difficult.
+A typical example of complex problems occur for multivalued functions~\parencite{AISC:MultivaluedFunctions}.
+A \gls*{cas} usually defines \textit{branch cuts} to compute principal values of multivalued functions~\parencite{Maple:Cuts}, which makes the implementation of a theoretically continued function to a discontinuous presentation of it.
 In general, positioning branch cuts follows conventions, but can be positioned arbitrarily in many cases.
-Communicate and explain the decision of defined branch cuts is a critical point for \gls*{cas} and can vary between the systems~\parencite{Branches:acot}.
+Communicating and explaining the decision of defined branch cuts is a critical issue for \gls*{cas} and can vary between various systems~\parencite{Branches:acot}.
 Figure~\ref{fig:acot-cut-compare} illustrates two examples of different branch cut positioning for the inverse trigonometric arccotangent function.
-While \Maple{} defines the branch cut at ${[-\infty\iunit, -\iunit]}$, ${[\iunit,\infty\iunit]}$ (figure~\ref{fig:acot-cut1}), \Mathematica{} defines the branch cut at ${[-\iunit, \iunit]}$ (figure~\ref{fig:acot-cut2}).
+While \Maple{} defines the branch cut at ${[-\iunit\infty, -\iunit]}$, ${[\iunit,\iunit\infty]}$ (Figure~\ref{fig:acot-cut1}), \Mathematica{} defines the branch cut at ${[-\iunit, \iunit]}$ (figure~\ref{fig:acot-cut2}).
 
@@
     \centering
-    \subfloat[The real part of arccotangent with a branch cut at ${[-\infty\iunit, -\iunit]}$, ${[\iunit,\infty\iunit]}$.]{%
+    \subfloat[The real part of arccotangent with a branch cut at ${[-\iunit\infty, -\iunit]}$, ${[\iunit,\iunit\infty]}$.]{%
         \includegraphics[width=0.45\textwidth]{acotCut1.png}
@@
     }
-    \caption{Two plots for the real part of the arccotangent function with a branch cut at $[-\infty\iunit, -\iunit]$, $[\iunit,\infty\iunit]$ in figure~\protect\subref{fig:acot-cut1} and at $[-\iunit, \iunit]$ in figure~\protect\subref{fig:acot-cut2}, respectively.
+    \caption{Two plots for the real part of the arccotangent function with a branch cut at $[-\iunit\infty, -\iunit]$, $[\iunit,\iunit\infty]$ in figure~\protect\subref{fig:acot-cut1} and at $[-\iunit, \iunit]$ in figure~\protect\subref{fig:acot-cut2}, respectively.
 (Plotted with \Maple{} 2016)}
@@
 \end{equation}
-is called \textit{inappropriate}.
+as \textit{inappropriate}.
 Note that it is not always as easy as in this example to decide if a translation is appropriate or not.

### 2018-07-30 chapters/02-background/related-work.tex
Implemented all suggestions by Howard

@@
 For example, the internal \LaTeX{} macro \verb|\frac| always indicates a fraction.
-However, for more complex expressions, e.g., the Jacobi polynomial in \cref{tab:JacobiP-usecase}, the import functions fail.
+However, for more complex expressions, e.g., the Jacobi polynomial in \Cref{tab:JacobiP-usecase}, the import functions fail.
 The second problem appears in the export tools.
@@
 Otherwise the \gls*{cas} wouldn't be able to compute or evaluate the expressions.
-During the export process, the semantic information gets lost, because generic \LaTeX{} is not able to carry semantic information.
-In consequence of these two problems, an exported expression cannot be imported to the same system again in most cases (except those simple expressions described above).
+During the export process, the semantic information gets lost, because generic \LaTeX{} is not able to carry sufficient semantic information.
+In consequence of these two problems, an exported expression cannot be imported to the same system again in most cases (except for simple expressions such as those described above).
 Our tool should solve these problems and provide round-trip translations between \LaTeX{} and \gls*{cas}.
@@
 The semantics must be well known before an expression can be translated.
-There are two main approaches to solve that problem: (1) someone could specify the semantic information during the writing process (pre-defined semantics) or (2) the translator can find out the right semantic information in general mathematical expressions before it translates the expression.
+There are two main approaches to solve that problem: (1) someone could specify the semantic information during the writing process (pre-defined semantics); and (2) the translator can find out the right semantic information in general mathematical expressions before it translates the expression.
 So-called \textit{interactive documents}\footnote{There is no adequate definition what interactive documents are.
-However, this name is widely used to describe electronic document formats that allow interactivity to change the content in real time.}, Such as the \gls*{cdf}~\parencite{CDF:Wolfram} by Wolfram Research or the \textit{worksheets} by \Maple{}, try to solve this problem with the approach (2) and allow to embed semantic information into the input.
+However, this name is widely used to describe electronic document formats that allow for interactivity to change the content in real time.}, such as the \gls*{cdf}~\parencite{CDF:Wolfram} by Wolfram Research, or \textit{worksheets} by \Maple{}, try to solve this problem with the approach (2) and allow one to embed semantic information into the input.
 Those complex document formats require specialized tools to show and work with the documents (Wolfram CDF Player, or \Maple{} for the \textit{worksheets}).
@@
 The documents can be viewed and edited in the browser.
-Those \JOBAD{-documents} also allow to perform computations during \gls*{cas}.
-This gives the opportunity to calculate, compute and change mathematical expressions directly in the document.
+Those \JOBAD{-documents} also allow one to perform computations via \gls*{cas}.
+This gives one the opportunity to calculate, compute and change mathematical expressions directly in the document.
 The translation performs in the background, invisible to the user.
-Similar to the \JOBAD{} architecture other interactive web documents exist, such as \textit{MathDox}~\parencite{MathDox} and \textit{The Planetary System}~\parencite{Planetary}.
-All of them demonstrate the potential of the education system.
+Similar to the \JOBAD{} architecture, other interactive web documents exist, such as \textit{MathDox}~\parencite{MathDox} and \textit{The Planetary System}~\parencite{Planetary}.
+All of these demonstrate the potential of the educational system.
 
-Another approach tries to avoid translation problems by allow computations directly via the \LaTeX{} compiler, e.g., \textit{LaTeXCalc}~\parencite{LatexCalc}.
-Those packages are limited to the possibilities of the compiler and therefore not as powerful as \gls*{cas}.
+Another approach tries to avoid translation problems by allowing computations directly via the \LaTeX{} compiler, e.g., \textit{LaTeXCalc}~\parencite{LatexCalc}.
+Those packages are limited to the abilities of the compiler and therefore are not as powerful as \gls*{cas}.
 A workaround for this case is \textit{sagetex}~\parencite{Sagetex}, which is a \LaTeX{} package interface for the open source \gls*{cas} \textit{sage}\footnote{An abbreviation for \textit{SageMath}.}.
@@
 
-There exist two approaches that allow embedding semantic information within \LaTeX{} expressions by using custom macros.
+There exist two approaches for marking up mathematical \TeX/\LaTeX{} documents semantically with \TeX{} macros.
 Namely, \sTeX{}~\parencite{sTeX} developed by Kohlhase and the \Macro s developed by Miller~\parencite{DLMF:Macros}.
 This paper shows that it is possible to develop a context-free translation tool using the semantic macros introduced by these two projects.
-\sTeX{} aims to embed semantic information for general mathematics with a comprehensive set of macros.
+The goal of \sTeX{} was to markup the functional structure of mathematical documents so that they can be exported to the \gls*{omdoc} format.
 The macro set developed by Miller introduces new macros for special functions, orthogonal polynomials, and mathematical constants.
@@
 These semantic macros are internally used in the \gls*{dlmf} and the \gls*{drmf}.
-Because of the linked definitions to the \gls*{dlmf} and the special macros for functions and mathematical constants this paper using the \Macro s for performing translations to \gls*{cas} instead of using \sTeX.
+We gave the \Macro{} set the favor for developing the translation engine because it provides \gls*{dlmf} definitions for a comprehensive amount of functions.
+In contrast, \sTeX{} does not focus on the semantics of functions, is too complex to use, and defines diverse macros for symbols and concepts that \gls*{cas} usually does not support.
+
+%Because of the linked definitions to the \gls*{dlmf} and the special macros for functions and mathematical constants this paper using the \Macro s for performing translations to \gls*{cas} instead of using \sTeX.

### 2018-07-30 chapters/03-translator/backward-trans.tex
Implemented all suggestions by Howard

@@
 The returned \inertF{} is a nested list representation as a tree equivalent\footnote{A tree equivalent of a \gls*{dag} splitting nodes with multiple parents into multiple nodes so that each node has only one parent node.} of the internal \gls*{dag} for the given expression.
-Some of the important types for the nodes are specified in table~\ref{tab:maple-types}.
+Some of the important types for the nodes are specified in \Cref{tab:maple-types}.
 
@@
 This internal representation might look a bit different to the input.
-One example has already been given with figure~\ref{fig:internal-maple-dag}, where each summand of a sum is stored with a factor.
+One example has already been given with \Cref{fig:internal-maple-dag}, where each summand of a sum is stored with a factor.
 Here is a list of all internal changes that occur for inputs.
@@
 Translations for functions are also realized via patterns and placeholders.
-Figure~\ref{fig:backward-trans} illustrates the backward translation process for the Jacobi polynomial example from table~\ref{tab:JacobiP-usecase}.
+Figure~\ref{fig:backward-trans} illustrates the backward translation process for the Jacobi polynomial example from \Cref{tab:JacobiP-usecase}.
 
-\begin{figure}[!htp]
+\begin{figure}[t!]
 	\centering

### 2018-07-30 chapters/03-translator/forward-trans.tex
Implemented all suggestions by Howard

@@
 In consequence, the \gls*{pom}-tagger generates a raw parsed tree where each token in the \LaTeX{} expression is a node in the tree.
-We call this parsed tree the \gls*{pom-pt}.
+We call this parsed tree, the \gls*{pom-pt}.
 
-{\sloppy The overall forward translation process is explained in figure~\ref{fig:forward-trans}.
+{\sloppy The overall forward translation process is explained in~\Cref{fig:forward-trans}.
 All translation patterns and related information are stored in the DLMF/DRMF tables.
@@
 Since the \gls*{bnf} does not define rules for semantic macros, each argument of the semantic macro and each $@$ symbol are following siblings of the semantic macro node.
-That is the reason, why we stored the number of parameters, variables and $@$ symbols in the lexicon files.
+That is the reason why we stored the number of parameters, variables and $@$ symbols in the lexicon files.
 Otherwise, the translator could not find the end of a semantic macro in the \gls*{pom-pt}.
@@
 
-Figure~\ref{fig:syntax-tree-usecase} visualizes the \gls*{pom-pt} of the Jacobi polynomial example from table~\ref{tab:JacobiP-usecase}.
+\Cref{fig:syntax-tree-usecase} visualizes the \gls*{pom-pt} of the Jacobi polynomial example from~\Cref{tab:JacobiP-usecase}.
 Because of the differences to expression trees, a backward conversion of the \gls*{pom-pt} to a string representation can be difficult, especially for finding necessary or unnecessary parentheses.
@@
 The algorithm is realized in a simple recursive structure.
-The algorithm~\ref{alg:simple-translation} explains the translation process.
+The Algorithm~\ref{alg:simple-translation} explains the translation process.
 Whenever the algorithm finds a leaf, it can translate this single term.
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}[t]
 \caption{Abstract translation algorithm to translate MLP-Parse trees.}\label{alg:translation}
@@
 \end{algorithm}
-\vspace{-20px}
+%\vspace{-20px}
 
-This approach is not completely feasible since the algorithm needs to look ahead and check the following siblings in some cases, e.g., in the case of a semantic macro with arguments (see figure~\ref{fig:syntax-tree-usecase}).
+This approach is not completely feasible since the algorithm needs to look ahead and check the following siblings in some cases, e.g., in the case of a semantic macro with arguments (see~\Cref{fig:syntax-tree-usecase}).
 The {\footnotesize \verb|TRANSLATE_LEAF|} function needs to know the following siblings of the current node.
@@
 Eventually, some of the following siblings are needed to translate $r$.
-The list of \textit{following\_siblings} in line~\ref{line:transLeaf} might be reduced to avoid multiple translations for one node.
+The list of \textit{following\_siblings} in Line~\ref{line:transLeaf} might be reduced to avoid multiple translations for one node.
 If $r$ is not a leaf, it contains one or more children.
@@
 Once we have translated $r$, we can go a step further and translate the next node.
-Line~\ref{line:ifNotEmpty} checks if there are following siblings left and calls the {\footnotesize \verb|ABSTRACT_TRANSLATOR|} recursively in that case.
+Line~\ref{line:ifNotEmpty} checks if there are following siblings left and calls the {\footnotesize \verb|ABSTRACT_TRANSLATOR|} recursively in such cases.
 Translated expressions are stored by the \gls*{teo} object.
-The detailed translation process is not included in the algorithm.
-
 Algorithm~\ref{alg:translation} is a simplified version of the translator process.
-The main progress happens in \cref{line:transLeaf,line:transSeq}.
-There are several cases what $r$ can be.
-Table~\ref{tab:allTypesTable} on page \pageref{tab:allTypesTable} gives an overview of all different types in a \gls*{pom-pt}.
+The \Cref{line:transLeaf,line:transSeq} process the translations for each node.
+\Cref{tab:allTypesTable} gives an overview of all the different node types the root $r$ can be.
 A more detailed explanation of the types can be found in~\parencite{POM-Tagger}.
@@
 The \gls*{bnf} grammar defines some basic grammatical rules for generic \LaTeX{} macros, such as for \verb|\frac|, \verb|\sqrt|.
-Therefore, there is a hierarchical structure for those symbols, similar to the structure in expression trees.
+Therefore, there is a hierarchical structure for those symbols similar to the structure in expression trees.
 As already mentioned, some of these types can be translated directly, such as Greek letters, while others are more complex, such as semantic \LaTeX{} macros.
 Therefore, the translators delegate the translation to specialized subtranslators.
-This delegation process is implemented in \cref{line:transLeaf,line:transSeq} of algorithm~\ref{alg:translation}.
+This delegation process is implemented in \Cref{line:transLeaf,line:transSeq} of Algorithm~{alg:translation}.
 The Subsection~\ref{sec:subtranslators} discusses these classes in more detail.
 
+\input{chapters/03-translator/all-types-table.tex}
+
 This approach seems to work for most of semantic \LaTeX{} expressions.
-However, there is one problem with this straight forward approach which was mentioned shortly mentioned in section~\ref{sec:problems}.
+However, there is one problem with this straight forward approach which was mentioned shortly mentioned in~\Cref{sec:problems}.
 There are many different types of notations used to represent formulae.
-Table~\ref{tab:notations} illustrates the simple expression $(a+b)x$ in different notations.
+\Cref{tab:notations} illustrates the simple expression $(a+b)x$ in different notations.
 The \gls*{pol}\footnote{Also known as \textit{Warsaw Notation} or \textit{prefix notation}} (hereafter called prefix notation) places the operator to the left of/before its operands.
@@
 
-\begin{wrapfigure}{r}{0.4\textwidth}
-\vspace{-22pt}
-\begin{minipage}{0.4\textwidth}
+\begin{wrapfigure}{r}{0.38\textwidth}
+\vspace{-10pt}
+\begin{minipage}{0.38\textwidth}
 \center
@@
 \end{minipage}
-\vspace{-15pt}
+\vspace{-5pt}
 \end{wrapfigure}
@@
 
-\input{chapters/03-translator/all-types-table.tex}
-
 Most programming languages (and \gls*{cas} as well) internally use prefix or postfix notation and do not mix the notations in one expression, since it is more convenient to parse expressions in uniform notations.
@@
 Since the \gls*{pom} has rarely implemented mathematical grammatical rules yet, it takes the input as it is and does not build an expression tree.
-Therefore, it parses all four examples from table~\ref{tab:notations} to four different \gls*{pom-pt}s rather than to one unique expression tree.
+Therefore, it parses all four examples from~\Cref{tab:notations} to four different \gls*{pom-pt}s rather than to one unique expression tree.
 In general, this is not a problem for our translation process since most \gls*{cas} are familiar with the most common notations.
@@
 Nonetheless, changing a notation could also solve ambiguities in some situations.
-Consider the two ambiguous examples in table~\ref{tab:amb_ex}.
+Consider the two ambiguous examples in~\Cref{tab:amb_ex}.
 While a scientist would probably just ask for the right interpretation of the first example, \Maple{} automatically computes the first interpretation.
@@
 In most cases, parentheses can be used to disambiguate expressions.
-We used them in table~\ref{tab:amb_ex} to clarify the different interpretations in example 2.
+We used them in~\Cref{tab:amb_ex} to clarify the different interpretations in example 2.
 But sometimes, even parentheses cannot solve a mistaken computation.
@@
 An exponent will be translated and shifted to the end of the translated semantic macro, since it is more common to write the exponent after the arguments of a function in \gls*{cas}.
-Therefore, the function translates and stores the exponent in line~\ref{line:next_caret}.
+Therefore, the function translates and stores the exponent in Line~\ref{line:next_caret}.
 One could ask what happens when there is an exponent given before and after the arguments.

### 2018-07-30 chapters/03-translator/gen-app.tex
Implemented all suggestions by Howard

@@
 \section{Translation Problems}\label{sec:problems}
-There are several potential problems to perform translations between systems that embed semantic information in the input.
-Those problems vary from simple cases, e.g., a function is not defined in the system, too complex cases, e.g., different positioning of branch cuts for multivalued functions.
-This section will discuss the problems and workarounds.
+There are several potential problems for performing translations between systems that embed semantic information in the input.
+These problems vary from simple cases, e.g., a function is not defined in the system, to complex cases, e.g., different positioning of branch cuts for multivalued functions.
+This section will discuss the problems and our workarounds.
 
 If a function is defined in one system but not in the other, we can translate the definition of the mathematical function.
-For example, the \textit{Gudermannian}~\parencite[eq.
-4.23.10]{NIST:DLMF} $\Gudermannian{x}$ function is defined by
+For example, the \textit{Gudermannian}~\parencite[(4.23.10)]{NIST:DLMF} $\Gudermannian{x}$ function is defined by
 \begin{equation}\label{eq:gudermannian}
-\Gudermannian{x} := \atan{\sinh{x}} \quad x \in \Real.
+\Gudermannian{x} := \atan{\sinh{x}}, \quad x \in \Real,
 \end{equation}
-and linked with the semantic macro \verb|\Gudermannian| in the \gls*{dlmf} but does not exist in \Maple. We can perform a translation for the definition~\ref{eq:gudermannian} instead of macro itself
+and linked to the semantic macro \verb|\Gudermannian| in the \gls*{dlmf} but does not exist in \Maple. We can perform a translation for the definition~\eqref{eq:gudermannian} instead of macro itself
 \begin{equation}\label{eq:guder-trans}
@@
 
-Since those translations are nonintuitive, describing explanations become necessary for the translation process.
+Since translations such as these are nonintuitive, describing explanations become necessary for the translation process.
 A special logging function takes care of each translation and provide details after a successful translation process.
-Section~\ref{sec:forward-translation} explains this task further.
+\Cref{sec:forward-translation} explains this task further.
 
@@
 
-The direct translation~(\ref{eq:acot-alternatives}) has the branch cut issue, while the alternative translations~(\ref{eq:acot-alternatives-1}) and~(\ref{eq:acot-alternatives-2}) using other functions instead of the arccotangent function.
-The arctangent function~(\ref{eq:acot-alternatives-1}) and the natural logarithm~(\ref{eq:acot-alternatives-2}) have the same positioned branch cuts in the \gls*{dlmf} and in \Maple.
-In consequence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user do not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except for $z = -\iunit$.
+The position of the branch cut of the arccotangent function differs after translation~\eqref{eq:acot-alternatives}, which may lead to incorrect calculations later on.
+The alternative translations~\eqref{eq:acot-alternatives-1} and~\eqref{eq:acot-alternatives-2} use other functions instead of the arccotangent function.
+The arctangent function~\eqref{eq:acot-alternatives-1} and the natural logarithm~(\ref{eq:acot-alternatives-2}) have the same positioned branch cuts as in the \gls*{dlmf} and in \Maple.
+In consequence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user do not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z = -\iunit$.
 
-Other problematic cases for translations are the \Macro s itself.
+Other problematic cases for translations are the \Macro s themselves.
 In some cases, they do not provide sufficient semantic information to perform translations.
-One example is the \textit{Wronskian} symbol.
-For two differentiable functions the \textit{Wronskian} is defined as~\parencite[eq.
-1.13.4]{NIST:DLMF}
+One example is the \textit{Wronskian} determinant.
+For two differentiable functions, the \textit{Wronskian} is defined as~\parencite[(1.13.4)]{NIST:DLMF}
 \begin{equation*}
@@
 \end{equation*}
-In semantic \LaTeX{} it is used with
+In semantic \LaTeX{} it is currently implemented using
 \begin{equation}
@@
 \end{equation}
-A translation become infeasible, because the macro do not explicitely defined the variable of the functions $w_1$ and $w_2$.
+A translation become unfeasible because the macro does not explicitely define the variable of differentiation for the functions $w_1$ and $w_2$.
 For a correct translation, the \gls*{cas} needs to be aware of the used variable $z$.
-We solved this issue by creating a new macro
+We solved this issue by creating a new macro, e.g.,
 \begin{equation}
@@
 \end{equation}
-This example shows that the \Macro s are still work in progress and are getting constantly updated.
+This example demonstrates that the \Macro s are still a work in progress and are getting constantly updated.
 
@@
 For such problems, we introduced a new macro \verb|\idot| for an invisible multiplication symbol (this macro will not be rendered).
-Since this macro is newly introduced and automatic conversion of existing equations is difficult, none of the equations in the \gls*{dlmf} uses this macro yet.
+Since this macro is newly introduced by contributers of the \gls*{drmf} team, and automatic conversion of existing equations is difficult, none of the equations in the \gls*{dlmf} use this macro.
 As a consequence, the translator has some simple rules to perform translations without explicitly marked translations with \verb|\idot|.
 
-Even we only allow the special dialect of \LaTeX{} using semantic macros not all expressions are unambiguous.
-In table~\ref{tab:amb-latex} are four examples of ambiguous expressions.
-These expressions are unambiguous for the \LaTeX{} compiler since it only considers the very next token for power and subscripts.
-Our translator following the same rules to solve these issues.
+The \Macro s do not garantee entirely disambiguated expressions.
+In~\Cref{tab:amb-latex} there are four examples of potentially ambiguous expressions.
+These expressions are unambiguous for the \LaTeX{} compiler since it only considers the very next token for superscripts and subscripts.
+Our translator follows the same rules to solve these issues.
 
@@
 	\hline
-	Ambiguous Input & \LaTeX{} Output\\
+	Potentially Ambiguous Input & \LaTeX{} Output\\
 	\hline
@@
 \end{tabular}
-\caption{Ambiguous \LaTeX{} expressions and how \LaTeX{} displays them.}
+\caption{Potentially ambiguous \LaTeX{} expressions and how \LaTeX{} displays them.}
 \label{tab:amb-latex}
@@
 
-Another more questionable translation decision is alphanumerical expressions.
-As explained in table~\ref{tab:allTypesTable}, the \gls*{pom}-tagger handles strings of letters and numbers differently, depending on the order of the symbols.
-The reason is, that an expression such as '$4b$' is usually considered to be a multiplication of $4$ and '$b$,' while '$b4$' looks like indexing '$b$' by $4$.
-While the first example produces two nodes, namely $4$ and '$b$', the second example '$b4$' produces just a single alphanumerical node in the \gls*{pom-pt}.
-The translator interprets alphanumerical expressions as multiplications for two reasons.
-(1) We would assume that the inputs '$4b$' and '$b4$' are mathematically equivalent and (2) it is more common in mathematics to use single letter names for variables~\parencite{Notation:History}.
+Another more questionable translation decision are alphanumerical expressions.
+As explained in~\Cref{tab:allTypesTable}, the \gls*{pom}-tagger handles strings of letters and numbers differently, depending on the order of the symbols.
+The reason is, that an expression such as `$4b$' is usually considered to be a multiplication of $4$ and `$b$,' while `$b4$' looks like indexing `$b$' by $4$.
+While the first example produces two nodes, namely $4$ and `$b$', the second example `$b4$' produces just a single alphanumerical node in the \gls*{pom-pt}.
+The translator interprets alphanumerical expressions as multiplications for two reasons: (1) we would assume that the inputs `$4b$' and `$b4$' are mathematically equivalent; and (2) it is more common in mathematics to use single letter names for variables~\parencite{Notation:History}.
 Therefore we define the following definitions
@@
 In general, the translator is drafted to solve ambiguous expressions or automatically find a work-around to disambiguate the expression.
-Only if there is no way to solve the ambiguity with the defined rules, the translation process stops.
+Only in case there is no way to solve the ambiguity with the defined rules, the translation process stops.
 
@@
 The pattern uses \verb|$i| as placeholders to define the positions of the arguments.
-For example, the translation patterns for the Jacobi polynomial are illustrated in table~\ref{tab:placeholder_ex2}.
+For example, the translation patterns for the Jacobi polynomial are illustrated in~\Cref{tab:placeholder_ex2}.
 
@@
 	\end{tabular}
-	\caption{Forward and backward translation patterns of the Jacobi polynomial. The pattern for the backward translation is the same for \Maple{} and \Mathematica.}
+	\caption{Forward and backward translation patterns for the Jacobi polynomial example~\eqref{eq:P} in this manuscript.
+The pattern for the backward translation is the same for \Maple{} and \Mathematica.}
 	\label{tab:placeholder_ex2}
@@
 \end{equation*}
-where $f$ as an algebraic expression or an equation, $x$ the name of the differentiation variable and $n$ for the $n$-th order differentiation.
-A translation for $\frac{d^2x^2}{dx^2}$ should be like this
+where $f$ is an algebraic expression or an equation, $x$ is the name of the differentiation variable, and $n$ is an integer representing the $n$-th order differentiation\footnote{\url{https://www.maplesoft.com/support/help/maple/view.aspx?path=diff}, seen 07/2018}.
+A translation for $\frac{d^2x^2}{dx^2}$ should look like this
 \begin{equation*}
@@
 
-We can solve this issue by using paranthesis in such cases, e.g., \verb|diff($1, [$2$($0)])|.
+We can solve this issue by using parentheses in such cases, e.g., \verb|diff($1, [$2$($0)])|.
 
-The \Macro s also allow to specify optional arguments to distinguish between standard and another version of functions.
-The Legendre and associated Legendre function of the first kind are examples for such cases.
-The library that defines translations for each macro using the macro name as a primary key to identify the translation.
-The Legendre and associated Legendre function of the first kind both using the same macro \verb|\LegendreP|.
-To distinguish such cases, we are using a special syntax such cases, shown in table~\ref{tab:legendreP-lex}.
+The \Macro s also allow one to specify optional arguments to distinguish between standard and another version of these functions.
+The Legendre and associated Legendre function of the first kind are examples of such cases.
+The library that defines translations for each macro uses the macro name as the primary key to identify the translations.
+The Legendre and associated Legendre function of the first kind both use the same macro \verb|\LegendreP|.
+To distinguish such cases, we use a special syntax for such cases, shown in~\Cref{tab:legendreP-lex}.
 
@@
 
-The translator using the \gls*{pom}-Tagger~\parencite{POM-Tagger}\footnote{Named according to the Part-of-Speech-Taggers in \gls*{nlp}.} to parse \LaTeX{} expressions into a parsed tree.
-The \gls*{pom}-Tagger is an \texttt{LL}-Praser defined by a context-free grammar in \gls*{bnf}.
+The translator uses the \gls*{pom}-Tagger~\parencite{POM-Tagger}\footnote{Named according to the Part-of-Speech-Taggers in \gls*{nlp}.} to parse \LaTeX{} expressions into a parsed tree.
+The \gls*{pom}-Tagger is an \texttt{LL}-Parser defined by a context-free grammar in \gls*{bnf}.
 Each token will be tagged by meta information defined in lexicon files.
 We extend the lexicon files to provide also the information that is necessary for the translation process.
-An example of an entry of the lexicon file is given in table~\ref{tab:sine-lex-example}.
+An example of an entry of the lexicon file is given in~\Cref{tab:sine-lex-example}.
 
@@
 	\! & \! & DLMF: \verb|\sin@@{z}|\\
-	\! & \! & DLMF-Link: dlmf.nist.gov/4.14\# E1\\
+	\! & \! & DLMF-Link: dlmf.nist.gov/4.14.E1\\
 	\! & \! & Meanings: Sine\\
@@
 	\end{tabular}
-	\caption{The entry of the sine function in the lexicon file.}
+	\caption{The entry of the trigonometric sine function in the lexicon file.}
 	\label{tab:sine-lex-example}

### 2018-07-30 chapters/04-evaluation/evaluation.tex
Implemented all suggestions by Howard

@@
 \item \textbf{Round Trip Tests}: translates expressions back and forth and analyzing the changes.
-\item \textbf{Function Relation Tests}: translate mathematically proven equivalent expressions from one system to a \gls*{cas} and evaluate whether the relation remains valid
-\begin{enumerate}
-\item via equivalence checks of the \gls*{cas} or
-\item via numerical evaluation tests.
-\end{enumerate}
+\item\label{approach-2} \textbf{Function Relation Tests (Symbolical)}: translate mathematically proven equivalent expressions from one system to a \gls*{cas} and evaluate whether the relation remains valid via symbolical equivalence checks.
+\item \textbf{Numerical Tests}: take the same equations from Approach~\ref{approach-2} but evaluates them on specific numerical values to test equivalence.
 \end{enumerate}
@@
 A complete round trip translation (two steps) is called \textbf{one cycle}.
-
-\begin{definition}[Fixed Point Representation]
 A \textbf{fixed point representation} (or short fixed point) in a round trip translation process is a string representation that is identical to all string representations in the following cycles.
-\end{definition}
 
@@
 
-There is currently only one exception known where a round trip test does not reach a fixed point representation: Legendre's incomplete elliptic integrals~\parencite[eq.
-19.2.4-7]{NIST:DLMF} are defined with the amplitude $\phi$ in the first argument in the \DLMF, while \Maple{} takes the trigonometric sine of the amplitude as the first argument.
+There is currently only one exception known where a round trip test does not reach a fixed point representation: Legendre's incomplete elliptic integrals~\parencite[(19.2.4-7)]{NIST:DLMF} are defined with the amplitude $\phi$ in the first argument in the \DLMF, while \Maple{} takes the trigonometric sine of the amplitude as the first argument.
 Therefore, the forward and backward translations are defined as
@@
 
-And round-trip translations will produce infinite chains of sine and inverse sine calls. Since there are no evaluations involved, these chains will not be reduced during the translations. 
+The round-trip translations produce infinite chains of sine and inverse sine calls because there are no evaluations involved. 
 The round trip tests are very successful, but they only detect errors in string representations.
@@
 
-With this technique we can detect translation errors such as in (\ref{eq:wrong-trans-1} and \ref{eq:wrong-trans-2}).
-Consider the \gls*{dlmf} equation for the sine and cosine function~\parencite[eq 4.21.2]{NIST:DLMF}
+With this technique we can detect translation errors such as in \eqref{eq:wrong-trans-1} and \eqref{eq:wrong-trans-2}.
+Consider the \gls*{dlmf} equation for the sine and cosine function~\parencite[(4.21.2)]{NIST:DLMF}
 \begin{equation}
@@
 However, the limitations of the \textit{simplify} function are still the weakest part of this verification approach.
-Consider the complex example~\parencite[eq.
-12.7.10]{NIST:DLMF}
+Consider the complex example~\parencite[(12.7.10)]{NIST:DLMF}
 \begin{equation}\label{eq:branch-cut-near}
@@
 While this evaluation is very powerful, it has a significant problem.
-Even when all tested values return a value differ from zero, it does not proof the equivalence of~\eqref{eq:branch-cut-near}.
+Even when all tested values return zero, it does not proof the equivalence of~\eqref{eq:branch-cut-near}.
 However, when the values are different from zero, it does proof an error in one of the four cases~\parencite{NumericalTests:Paper}
@@
 We extracted 4.087 equations from \gls*{dlmf} and apply our round-trip and relation tests on them.
-The translator was able to translate 2.405 (58.8\footnote{All percentages are approximately calculated.}\%) of the extracted equations without errors.
+The translator was able to translate 2.405\footnote{All percentages are approximately calculated.} (58.8\%) of the extracted equations without errors.
 660 (27.4\%) of the successfully translated expressions were verified by the simplification techniques of \Maple.
@@
 The evaluation techniques have proven to be very powerful also for evaluating \gls*{cas} and online mathematical compendia such as the \gls*{dlmf}.
-During the evaluations, we were able to detect several errors in the translation and evaluation engine.
-However, even errors in the \gls*{dlmf} were discovered.
+During the evaluations, we were able to detect several errors in the translation and evaluation engine, and also discovered two errors in the \gls*{dlmf} and one error in \Maple's \texttt{simplify}-function.
 
-The numerical test engine was able to discover a sign error in equation~\parencite[eq.
-14.5.14]{NIST:DLMF}\footnote{The equation had originally been stated as shown in equation~(\ref{eq:signerror}).
+The numerical test engine was able to discover a sign error in equation~\parencite[(14.5.14)]{NIST:DLMF}\footnote{The equation had originally been stated as shown in equation~(\ref{eq:signerror}).
 The error were reported on 10th April 2017.}
@@
 359]{NIST:Handbook} and has been fixed in the \gls*{dlmf} with version 1.0.16.
-The same engine also identified a missing comma in the constraint of~\parencite[eq.
-10.16.7]{NIST:DLMF}.
+The same engine also identified a missing comma in the constraint of~\parencite[(10.16.7)]{NIST:DLMF}.
 The original constraint was given by $2\nu \neq -1, -2 -3, \ldots$, with a missing comma after $-2$.
+
+We have also noticed that our testing procedure is able to identify errors in CAS procedures,
+namely the \Maple{} {\tt simplify} procedure. The left-hand side of \cite[(7.18.4)]{NIST:DLMF} is given by 
+\begin{equation*}
+\frac{{\mathrm{d}}^{n}}{{\mathrm{d}z}^{n}}\left(e^{z^{2}}\operatorname{erfc}z\right),
+\end{equation*}
+where $n \in \mathbb{N}_0$, $e$ is the base of the natural logarithm, and $\operatorname{erfc}$ is the
+complementary error function. Our translation correctly produces 
+\begin{equation*}
+\verb|diff((exp((z)^(2))*erfc(z)), [z$(n))]|.
+\end{equation*}
+However, the \Maple{ 2016} \verb|simplify| function falsely returns $0$ for the translated left-hand side.
+Maplesoft has confirmed in a private communication that this is indeed a defect in \Maple{ 2016}.
+Furthermore, although the nature of the defect changes, the defect still persists in \Maple{ 2018}
+as of the publication of this manuscript.

### 2018-07-30 chapters/02-background/related-work.tex
Put CDF citation to footnote and solve back-quote errors

@@
 So-called \textit{interactive documents}\footnote{There is no adequate definition what interactive documents are.
-However, this name is widely used to describe electronic document formats that allow for interactivity to change the content in real time.}, such as the \gls*{cdf}~\parencite{CDF:Wolfram} by Wolfram Research, or \textit{worksheets} by \Maple{}, try to solve this problem with the approach (2) and allow one to embed semantic information into the input.
+However, this name is widely used to describe electronic document formats that allow for interactivity to change the content in real time.}, such as the \gls*{cdf}\footnote{Wolfram Research; \textit{Computable Document Format} (CDF); \url{http://www.wolfram.com/cdf/}, July 2011} by Wolfram Research, or \textit{worksheets} by \Maple{}, try to solve this problem with the approach (2) and allow one to embed semantic information into the input.
 Those complex document formats require specialized tools to show and work with the documents (Wolfram CDF Player, or \Maple{} for the \textit{worksheets}).

### 2018-07-30 chapters/03-translator/forward-trans.tex
Put CDF citation to footnote and solve back-quote errors

@@
 	\includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm, scale=0.75]{SyntaxTreeUseCase.pdf}
-	\caption{pom-pt for a Jacobi polynomial using the DLMF/DRMF \LaTeX{} macro. Each leaf contains information from the lexicon files.}
+	\caption{The \gls*{pom-pt} for the Jacobi polynomial example~\eqref{eq:P} using the DLMF/DRMF \LaTeX{} macro. Each leaf contains information from the lexicon files.}
 	\label{fig:syntax-tree-usecase}
@@
 Therefore, the translators delegate the translation to specialized subtranslators.
-This delegation process is implemented in \Cref{line:transLeaf,line:transSeq} of Algorithm~{alg:translation}.
+This delegation process is implemented in \Cref{line:transLeaf,line:transSeq} of Algorithm~\ref{alg:translation}.
 The Subsection~\ref{sec:subtranslators} discusses these classes in more detail.
@@
 In mathematical expressions, notations are mostly mixed, depending on the case and number of operands.
-For example, infix notation is common for binary operators ($+$, $-$, $\cdot$, $\mod$ etc.), while functional notations are conveniently used for any kind of functions ($sin$, $cos$, etc.).
+For example, infix notation is common for binary operators ($+$, $-$, $\cdot$, $\mod$, etc.), while functional notations are conveniently used for any kind of functions ($sin$, $cos$, etc.).
 Sometimes the same symbol is used in different notations to distinguish different meanings.
-For example, the '$-$' as a unary operator is used in prefix notation to indicate the negative value of its operand, such as in '$-2$'.
-Of course, $-$ can also be the binary operator for subtraction, which is commonly used in infix notation.
-An example for the postfix notation is factorial, such as '$2!$'.
+For example, the `$-$' as a unary operator is used in prefix notation to indicate the negative value of its operand, such as in `$-2$'.
+Of course, `$-$' can also be the binary operator for subtraction, which is commonly used in infix notation.
+An example for the postfix notation is factorial, such as `$2!$'.
 
@@
 In general, this is not a problem for our translation process since most \gls*{cas} are familiar with the most common notations.
-Therefore, the translator does not need to know that $a$ and $b$ are the operands of the binary operator '$+$' in '$a+b$.' The translator could simply translate the symbols in '$a+b$' in the same order as they appear in the expression and the \gls*{cas} would understand it.
+Therefore, the translator does not need to know that `$a$' and `$b$' are the operands of the binary operator `$+$' in `$a+b$.' The translator could simply translate the symbols in `$a+b$' in the same order as they appear in the expression and the \gls*{cas} would understand it.
 However, this simple approach generates two problems.
@@
 But sometimes, even parentheses cannot solve a mistaken computation.
-For example, there is no way to add parentheses to force \Maple{} to compute $n!!$ as the double factorial function.
-Even $(n)!!$ will be interpreted as $(n!)!$.
+For example, there is no way to add parentheses to force \Maple{} to compute `$n!!$' as the double factorial function.
+Even `$(n)!!$' will be interpreted as `$(n!)!$'.
 Rather than using the exclamation mark in \Maple, one could also use the functional notation.
-For example, the interpretations $(2!)!$ and $(2)!!$ can be distinguished in \Maple{} by using \verb|factorial(factorial(2))| and \verb|doublefactorial(2)| respectively.
+For example, the interpretations `$(2!)!$' and `$(2)!!$' can be distinguished in \Maple{} by using \verb|factorial(factorial(2))| and \verb|doublefactorial(2)| respectively.
 We define the translations as follows:
@@
 Hence, combined with the recursive translation approach, the \verb|SequenceTranslator| also checks balances of parentheses in expressions.
-An expression such as '$(a]$' is producing a mismatched parentheses error.
-On the other hand, this is a problem for interval expressions such as '$[a,b)$'.
+An expression such as `$(a]$' is producing a mismatched parentheses error.
+On the other hand, this is a problem for interval expressions such as `$[a,b)$'.
 In the current version, the program cannot distinguish between mismatched parentheses and half-opened, half-closed intervals.
@@
 The most common symbol for multiplications is still the white space or none symbol at all, as explained previously.
-Consider the simple expression $2n\pi$.
+Consider the simple expression `$2n\pi$'.
 The \gls*{pom-pt} generates a sequence node with three children, namely $2$, $n$ and $\pi$.
@@
 This potentially produces strange effects.
-Consider an expression such as $f(x)$.
+Consider an expression such as `$f(x)$'.
 Translate this to \Maple{} will be \verb|f*(x)|.

### 2018-07-30 main.tex
Reduce everything to one file

\documentclass[a4paper,11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel} \usepackage[T1]{fontenc} 
\usepackage{hyperref}			\usepackage[acronym, nomain]{glossaries} \usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{parskip}		\usepackage{txfonts}            \usepackage{xcolor}				\usepackage{enumitem}			\usepackage{algorithm}
\usepackage{algpseudocode} \usepackage{lipsum} \usepackage{tabularx}
\usepackage{stmaryrd}
\usepackage{float}
\usepackage{csquotes}
\usepackage{layout}
\usepackage{textcomp} \usepackage{authblk}

\usepackage{sectsty}
\subsectionfont{\normalfont\large\itshape}

\usepackage[	a4paper,	]{geometry}
\setlength{\headheight}{14pt}
\usepackage{fancyhdr}

\usepackage{graphicx}           \usepackage{subfig}             \usepackage{wrapfig}			\graphicspath{ {./images/} }

\usepackage[export]{adjustbox}

\usepackage{styles/thesistitle}

\usepackage{cleveref}

\usepackage[
	backend		= biber,
	style		= authoryear-ibid,
	giveninits	= true, 	uniquename	= init, 	natbib		= true,
	url			= false,
	isbn			= false,
	doi			= false,
	eprint		= false,
	maxbibnames	= 99,
	minbibnames	= 3
]{biblatex}

\DeclareNameAlias{sortname}{last-first}

\addbibresource{./bibliography/Bibliography.bib} 
\hypersetup{
	pdftitle	= {Semantic Preserving Bijective Mappings for Representations of Special Functions in Computer Algebra Systems and Word Processors - Exemplified using the DLMF/DRMF, the Word Processor LaTeX and the Computer Algebra Systems Maple and Mathematica},
	pdfsubject={Master's Thesis in Mathematics, Technische Universität Berlin, July 2017.},
	pdfauthor	= {Andr\'e Greiner-Petter},
	pdfcreator	= {Andr\'e Greiner-Petter},
	pdfkeywords	= {Special Functions, LaTeX, Translation},
}

\addto\captionsenglish{
  \renewcommand{\contentsname}    {Table of Contents}}

\apptocmd{\UrlBreaks}{\do\f\do\m}{}{}
\setcounter{biburllcpenalty}{9000}\setcounter{biburlucpenalty}{9000}

\renewcommand{\algorithmicrequire}{\textbf{Input:}} \renewcommand{\algorithmicensure}{\textbf{Output:}} \newcommand{\NULL}{\textbf{null}}
\renewcommand{\Return}{\textbf{return}}

\newcommand{\DLMF}{DLMF}
\newcommand{\DRMF}{DRMF}
\newcommand{\CAS}{CAS}
\newcommand{\Maple}{Maple}
\newcommand{\Mathematica}{Mathematica}
\newcommand{\Macro}{\DLMF/\DRMF{} \LaTeX{} macro}
\newcommand{\MLP}{MLP}
\newcommand{\JOBAD}{{\tt JOBAD}}

\newcommand{\tbs}{\textbackslash}

\newcommand{\langTex}{\mathfrak{sL}}
\newcommand{\langMaple}{\mathfrak{M}_{aple}}
\newcommand{\langMathe}{\mathfrak{M}_{athematica}}

\newcommand{\inertF}{\texttt{InertForm}}

\newcommand{\raw}[1]{\texttt{#1}}
\newcommand{\trans}[3]{\mathrm{tr}_{#2}^{#1}\left(#3\right)}

\newcommand{\sTeX}{{\raisebox{-.5ex}S\kern-.5ex\TeX}}

\newcommand{\todo}[1]{{\color{red}\textbf{#1}}}

\newcommand{\tableRowSpace}{\rule{0pt}{0.9\normalbaselineskip}}

\newcommand{\aSingleQuote}{\hspace{0.03cm}\textquotesingle\hspace{0.03cm}}

\newcommand{\tempEnvCitation}{}
\newcommand{\tempEnvSpacing}{}
\newenvironment{myQuote}[2][-0.4cm]{	\begin{quote}
	\renewcommand{\tempEnvCitation}{#2}
	\renewcommand{\tempEnvSpacing}{#1}
	\itshape
}{	\vspace{\tempEnvSpacing}
	\begin{flushright}
		\rule{6cm}{0.4pt}\\[-2pt]
		\tempEnvCitation{}
	\end{flushright}	\end{quote}
}

\newcommand*{\myRuleTextFill}[2]{  \makebox[#1]{    \leaders\hrule height \dimexpr 8.2pt\relax depth \dimexpr -7pt\relax \hfill    \enskip{#2}\enskip    \leaders\hrule height \dimexpr 8.2pt\relax depth \dimexpr -7pt\relax \hfill\kern0pt}}

\def\myRuleFill{\leavevmode\leaders\hrule height \dimexpr 8.2pt\relax depth \dimexpr -7pt\hfill\kern0pt}

\newtheoremstyle{defTheoStyle}
	{6pt} 	{2pt} 	{\itshape} 	{} 	{\bfseries} 	{} 	{\newline} 	{\thmname{#1} \thmnumber{#2}: {\normalfont\textit{(#3)}}}

\newtheoremstyle{defExampStyle}
	{6pt} 	{2pt} 	{\itshape} 	{} 	{\bfseries} 	{} 	{\newline} 	{\thmname{#1} \thmnumber{#2}:}

\theoremstyle{defTheoStyle}
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\theoremstyle{defExampStyle}
\newtheorem{example}{Example}[section]

\raggedbottom

\newacronym{dlmf}{DLMF}{Digital Library of Mathematical Functions}
\newacronym{drmf}{DRMF}{Digital Repository of Mathematical Formulae}
\newacronym{cas}{CAS}{Computer Algebra System}
\newacronym{mlp}{MLP}{Mathematical Language Parser}
\newacronym{mlp-pt}{PPT}{PoM-Parsed Tree}
\newacronym{pom-pt}{PPT}{PoM-Parsed Tree}
\newacronym{bnf}{BNF}{Backus-Naur Form}
\newacronym{p2c}{P2C}{Presentation-To-Computation}
\newacronym{nlp}{NLP}{Natural Language Processing}
\newacronym{pom}{PoM}{Part-of-Math}
\newacronym{csv}{CSV}{Comma-Separated Values}
\newacronym{json}{JSON}{JavaScript Object Notation}
\newacronym{teo}{TEO}{Translated Expression Object}
\newacronym{pol}{NPN}{Normal Polish Notation}
\newacronym{rpol}{RPN}{Reverse Polish Notation}
\newacronym{dag}{DAG}{Directed Acyclic Graph}
\newacronym{api}{API}{Application Programming Interface}
\newacronym{mathml}{MathML}{Mathematical Markup Language}
\newacronym{oop}{OOP}{Object-Oriented Programming}
\newacronym{cdf}{CDF}{Computable Document Format}
\newacronym{mathML}{MathML}{Mathematical Markup Language}
\newacronym{cfsf}{CFSF}{Continued Fractions for Special Functions}
\newacronym{ecf}{eCF}{Encoding Continued Fraction Knowledge}
\newacronym{nist}{NIST}{National Institute of Standards and Technology}
\newacronym{vmext}{VMEXT}{Visualizing Mathematical Expression Trees}
\newacronym{opsf}{OPSF}{Orthogonal Polynomials and Special Functions}
\newacronym{wp}{WP}{Word Processor}
\newacronym{stem}{STEM}{Science, Technology, Engineering and Mathematics}
\newacronym{gui}{GUI}{Graphical User Interface}
\newacronym{omdoc}{OMDoc}{Open Mathematical Documents}

\DeclareRobustCommand{\cpi}{{\pi}}
\DeclareRobustCommand{\expe}{{e}}
\DeclareRobustCommand{\EulerConstant}{{\gamma}}
\DeclareRobustCommand{\iunit}{{i}}
\DeclareRobustCommand{\Complex}{\mathbb{C}}
\DeclareRobustCommand{\Real}{\mathbb{R}}

\newcommand{\Gudermannian}[1]{\mathrm{gd}\!\left(#1\right)}
\newcommand{\atan}[1]{\mathrm{arctan}\!\left(#1\right)}
\newcommand{\acot}[1]{\mathrm{arccot}\!\left(#1\right)}
\newcommand{\ph}[1]{\mathrm{ph}\!\left(#1\right)}

%

\title{
    Semantic Preserving Bijective Mappings for Representations of Special Functions between Computer Algebra Systems and Word Processors
}
\author{Andr\'e Greiner-Petter\textsuperscript{1}}
\author{Moritz Schubotz\textsuperscript{1}}
\author{Howard S.~Cohl\textsuperscript{2}}
\author{Bela Gipp\textsuperscript{1}}

\affil{
	\textsuperscript{1}Information Science Group, University of Konstanz, Germany\\
	\url{{first.last}@uni-konstanz.de}
}

\affil{
	\textsuperscript{2}Applied and Computational Mathematics Division,\\
	National Institute of Standards and Technology, Mission Viejo, CA, USA,\\
	\url{howard.cohl@nist.gov}
}

\date{}

\begin{document}
\maketitle
\begin{abstract}
\glsresetall
\noindent
We present a translation tool between representations of semantically enriched mathematical formulae in a \gls*{wp} and its corresponding representations in a \gls*{cas}. We have chosen \Maple{} and \Mathematica{} for the \gls*{cas}, and \LaTeX{} as our \gls*{wp}.
Bruce Miller in the U.S.~has developed a set of semantic \LaTeX{} macros for \gls*{opsf}. The \gls*{nist} \gls*{dlmf} uses these semantic macros to provide a semantically enhanced mathematical online compendium. These semantic macros usally provide exclusive access to the semantic information of the functions. However, even if the semantics of a representation of one formula is unique, the semantics in another representation for the same formula may differ, such as in domains of variables. While some distinctions are obvious, such as syntactical characteristics, others are more difficult to examine, such as differences in definitions, which leads to error-prone manual translations.
We discuss difficulties related to an automatic translation system, and suggest possible solutions. 
Furthermore, this paper introduces new evaluation approaches for the developed translation tool. 
With the help of our automatic translation tool, the evaluation experiments were able to discover errors in the \gls*{dlmf} and in Maple.
\end{abstract} \glsresetall

\noindent
{\it \bf Keywords:} LaTeX, Computer Algebra System (CAS), Translation, Presentation to Computation (P2C)

\section{Introduction}
A typical workflow of a scientist who writes a scientific publication is to use a \gls*{wp} to write the paper and one or more \gls*{cas} for verification, analysis and visualization. Especially in the \gls*{stem} literature, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}. However, since \LaTeX{} (an extension of \TeX) is the de facto standard for writing articles in \gls*{stem}, and we only focus on mathematical writing in this paper, we have categorize \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the past 30 years~\parencites{Knuth}[559]{DigitalTypo}{LATEX:Standard}. \LaTeX{} enables printing of mathematical formulae in a structure similar to handwritten style. For example, consider the specific Jacobi polynomial~\parencite[18.3 in table 1]{NIST:DLMF}
\begin{equation}\label{eq:P}
P_n^{(\alpha , \beta)}(\cos(a\Theta)),
\end{equation}
where $\alpha, \beta > -1$, and $n$ is a nonnegative integer. 
This mathematical expression is written in \LaTeX{} as
\begin{equation*}
\verb|P_n^{(\alpha,\beta)}(\cos(a\Theta))|.
\end{equation*}

While \LaTeX{} focuses on displaying mathematics, a \gls*{cas} concentrates on computations and user friendly syntax. Especially important for a CAS is to embed unambiguous semantic information within the input. Each system uses different representations and syntax in consequence. Hence, a writer needs to constantly translate mathematical expressions from one representation to another and back again. Table~\ref{tab:JacobiP-usecase} shows four different representations for~(\ref{eq:P}).

\begin{table}[ht]
	\centering
	\begin{tabular}{ll}
		\hline
		Systems & Representations \\
		\hline
		\hline
		Generic \LaTeX\ & \verb|P_n^{(\alpha,\beta)}(\cos(a\Theta))| \\ 
		Semantic \LaTeX\ & \verb|\JacobiP{\alpha}{\beta}{n}@{\cos@{a\Theta}}| \\
		\Maple & \verb|JacobiP(n,alpha,beta,cos(a*Theta))| \\ 
		\Mathematica & \verb|JacobiP[n,\[Alpha],\[Beta],Cos[a \[CapitalTheta]]]|\\
		\hline
	\end{tabular}
	\caption{Different representations for~\eqref{eq:P}. Generic \LaTeX{} is the default \LaTeX{} expression; semantic \LaTeX{} uses special semantic macros to embed semantic information; and \gls*{cas} representations are unique to themselves.}
	\label{tab:JacobiP-usecase}
\end{table}

Translations from generic \LaTeX{} to \gls*{cas} are difficult to realize since the full semantic information is not easily constructed from the input. Bruce Miller at the \gls*{nist} has created a set of semantic \LaTeX{} macros~\parencite{DLMF:Macros}. Each macro ties specific character sequences to a well-defined mathematical object and is linked with the corresponding definition in the \gls*{dlmf}. The \gls*{drmf} is an outgrowth of the \gls*{dlmf} with the goal to facilitate interaction among a community of mathematicians and scientists~\parencites{DRMF:14}{DRMF:15}. The \gls*{drmf} extends the set of semantic macros. These macros embed necessary semantic information into \LaTeX{} expressions. One example of such a macro is given Table~\ref{tab:JacobiP-usecase} for the semantic \LaTeX{} representation for the Jacobi polynomial. The macros provide isolated access to important parts of the mathematical function, such as the arguments.

Even with embedded semantic information, a translation between systems can be difficult. A typical example of complex problems occur for multivalued functions~\parencite{AISC:MultivaluedFunctions}. A \gls*{cas} usually defines \textit{branch cuts} to compute principal values of multivalued functions~\parencite{Maple:Cuts}, which makes the implementation of a theoretically continued function to a discontinuous presentation of it. In general, positioning branch cuts follows conventions, but can be positioned arbitrarily in many cases. Communicating and explaining the decision of defined branch cuts is a critical issue for \gls*{cas} and can vary between various systems~\parencite{Branches:acot}. Figure~\ref{fig:acot-cut-compare} illustrates two examples of different branch cut positioning for the inverse trigonometric arccotangent function. While \Maple{} defines the branch cut at ${[-\iunit\infty, -\iunit]}$, ${[\iunit,\iunit\infty]}$ (Figure~\ref{fig:acot-cut1}), \Mathematica{} defines the branch cut at ${[-\iunit, \iunit]}$ (figure~\ref{fig:acot-cut2}).

\begin{figure}[ht]
    \centering
    \subfloat[The real part of arccotangent with a branch cut at ${[-\iunit\infty, -\iunit]}$, ${[\iunit,\iunit\infty]}$.]{        \includegraphics[width=0.45\textwidth]{acotCut1.png}
        \label{fig:acot-cut1}    }
    \hspace{0.5cm}
    \subfloat[The real part of arccotangent with a branch cut at ${[-\iunit, \iunit]}$.]{        \includegraphics[width=0.45\textwidth]{acotCut2.png}
        \label{fig:acot-cut2}    }
    \caption{Two plots for the real part of the arccotangent function with a branch cut at $[-\iunit\infty, -\iunit]$, $[\iunit,\iunit\infty]$ in figure~\protect\subref{fig:acot-cut1} and at $[-\iunit, \iunit]$ in figure~\protect\subref{fig:acot-cut2}, respectively. (Plotted with \Maple{} 2016)}
    \label{fig:acot-cut-compare}
\end{figure}

Hence, a \gls*{cas} user needs to fully understand the properties and special definitions (such as the position of branch cuts) in the \gls*{cas} to avoid mistakes during a translation~\parencite{Maple:Cuts}. In consequence, a manual translation process is not only laborious, but also prone to errors. Note that this general problem has been named to automatic \gls*{p2c} conversion~\parencite{POM-Tagger}.

This article presents a new approach for automatic \gls*{p2c} and vice versa conversions. Translations from presentational to computational (computational to presentational) systems are called forward (backward) translations. A forward translation is denoted with an arrow with the target system language above the arrow. For example,
\begin{equation*}
t \overset{\langMaple}{\mapsto} c,
\end{equation*}
where $t$ is an expression in the \LaTeX{} language and $c$ is an element of the \Maple{} language $\langMaple$. As we will see later in this article, we need to compare mathematical concepts between systems. This is impossible from a mathematical point of view. Consider the irrational mathematical constant $e$, known as Euler's number. The theoretical construct for this symbol cannot be mathematically equivalent to the value \verb|exp(0)| in \Maple, caused by computational and implementational limitations. Instead of using the term \textit{equivalent}, we introduce a \textit{appropriate} and \textit{inappropriate} translations. We call a translation such as
\begin{equation}\label{eq:cos-def}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|cos(z)|
\end{equation}
as \textit{appropriate}, while a translation such as
\begin{equation}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|sin(z)|
\end{equation}
as \textit{inappropriate}. Note that it is not always as easy as in this example to decide if a translation is appropriate or not. Therefore, this article also presents several validation techniques to automatically verify if a translation is appropriate or inappropriate. In addition to this terminology, we introduce \textit{direct translations}. Later in the paper, we will explain that a translation from one specific mathematical object to its \textit{appropriate} counterpart in the other system is not always possible. We call a translation to the \textit{appropriate} counterpart \textit{direct}. For example, the translation~(\ref{eq:cos-def}) is \textit{direct}, while a translation to the definition of the cosine function
\begin{equation*}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|(exp(I*z)+exp(-I*z))/2|
\end{equation*}
is not a \textit{direct} translation.

Note that partial results of this paper have been published in~\parencite{CICM:Paper}.

\section{Related Work}\label{sec:related-work}
Since \LaTeX{} became the de facto standard for writing papers in mathematics, most of the \gls*{cas} provide simple functions to import and export mathematical \LaTeX{} expressions\footnote{The selected \gls*{cas} Maple, Mathematica, Matlab, and SageMath provide import and/or export functions for \LaTeX:\\
Maple, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;\\
Mathematica, \url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html} seen 06/2017;\\
Matlab, \url{https://www.mathworks.com/help/symbolic/latex.html} seen 06/2017;\\
SageMath, \url{http://doc.sagemath.org/html/en/tutorial/latex.html} seen 06/2017}. Those tools have two essential problems. They are only able to import simple mathematical expressions, where the semantics are unique. For example, the internal \LaTeX{} macro \verb|\frac| always indicates a fraction. However, for more complex expressions, e.g., the Jacobi polynomial in \Cref{tab:JacobiP-usecase}, the import functions fail. The second problem appears in the export tools. Mathematical expressions in \gls*{cas} are fully semantic. Otherwise the \gls*{cas} wouldn't be able to compute or evaluate the expressions. During the export process, the semantic information gets lost, because generic \LaTeX{} is not able to carry sufficient semantic information. In consequence of these two problems, an exported expression cannot be imported to the same system again in most cases (except for simple expressions such as those described above). Our tool should solve these problems and provide round-trip translations between \LaTeX{} and \gls*{cas}.

The semantics must be well known before an expression can be translated. There are two main approaches to solve that problem: (1) someone could specify the semantic information during the writing process (pre-defined semantics); and (2) the translator can find out the right semantic information in general mathematical expressions before it translates the expression. So-called \textit{interactive documents}\footnote{There is no adequate definition what interactive documents are. However, this name is widely used to describe electronic document formats that allow for interactivity to change the content in real time.}, such as the \gls*{cdf}\footnote{Wolfram Research; \textit{Computable Document Format} (CDF); \url{http://www.wolfram.com/cdf/}, July 2011} by Wolfram Research, or \textit{worksheets} by \Maple{}, try to solve this problem with the approach (2) and allow one to embed semantic information into the input. Those complex document formats require specialized tools to show and work with the documents (Wolfram CDF Player, or \Maple{} for the \textit{worksheets}). The \JOBAD{} architecture~\parencite{JOBAD:orig} is able to create web-based interactive documents and uses \gls*{omdoc}~\parencite{OMDoc} to carry semantics. The documents can be viewed and edited in the browser. Those \JOBAD{-documents} also allow one to perform computations via \gls*{cas}. This gives one the opportunity to calculate, compute and change mathematical expressions directly in the document. The translation performs in the background, invisible to the user. Similar to the \JOBAD{} architecture, other interactive web documents exist, such as \textit{MathDox}~\parencite{MathDox} and \textit{The Planetary System}~\parencite{Planetary}. All of these demonstrate the potential of the educational system.

Another approach tries to avoid translation problems by allowing computations directly via the \LaTeX{} compiler, e.g., \textit{LaTeXCalc}~\parencite{LatexCalc}. Those packages are limited to the abilities of the compiler and therefore are not as powerful as \gls*{cas}. A workaround for this case is \textit{sagetex}~\parencite{Sagetex}, which is a \LaTeX{} package interface for the open source \gls*{cas} \textit{sage}\footnote{An abbreviation for \textit{SageMath}.}. This package allows \textit{sage} commands in \TeX{-}files and uses \textit{sage} in the background to compute the commands. In this scenario, a writer still needs to manually translate expressions to the syntax of \textit{sage}.

There exist two approaches for marking up mathematical \TeX/\LaTeX{} documents semantically with \TeX{} macros. Namely, \sTeX{}~\parencite{sTeX} developed by Kohlhase and the \Macro s developed by Miller~\parencite{DLMF:Macros}. This paper shows that it is possible to develop a context-free translation tool using the semantic macros introduced by these two projects. The goal of \sTeX{} was to markup the functional structure of mathematical documents so that they can be exported to the \gls*{omdoc} format. The macro set developed by Miller introduces new macros for special functions, orthogonal polynomials, and mathematical constants. Each of these macros ties specific character sequences to a well-defined mathematical object and is linked with the corresponding definition in the \gls*{dlmf} or \gls*{drmf}. Therefore, we call these semantic macros \Macro s. These semantic macros are internally used in the \gls*{dlmf} and the \gls*{drmf}. We gave the \Macro{} set the favor for developing the translation engine because it provides \gls*{dlmf} definitions for a comprehensive amount of functions. In contrast, \sTeX{} does not focus on the semantics of functions, is too complex to use, and defines diverse macros for symbols and concepts that \gls*{cas} usually does not support.

\section{Translation Problems}\label{sec:problems}
There are several potential problems for performing translations between systems that embed semantic information in the input. These problems vary from simple cases, e.g., a function is not defined in the system, to complex cases, e.g., different positioning of branch cuts for multivalued functions. This section will discuss the problems and our workarounds.

If a function is defined in one system but not in the other, we can translate the definition of the mathematical function. For example, the \textit{Gudermannian}~\parencite[(4.23.10)]{NIST:DLMF} $\Gudermannian{x}$ function is defined by
\begin{equation}\label{eq:gudermannian}
\Gudermannian{x} := \atan{\sinh{x}}, \quad x \in \Real,
\end{equation}
and linked to the semantic macro \verb|\Gudermannian| in the \gls*{dlmf} but does not exist in \Maple. We can perform a translation for the definition~\eqref{eq:gudermannian} instead of macro itself
\begin{equation}\label{eq:guder-trans}
\verb|\Gudermannian{x}| \overset{\langMaple}{\mapsto} \texttt{arctan(sinh(x))}.
\end{equation}

Since translations such as these are nonintuitive, describing explanations become necessary for the translation process. A special logging function takes care of each translation and provide details after a successful translation process. \Cref{sec:forward-translation} explains this task further.

Providing detailed information also solves the problem for multiple alternative translations. In some cases, a semantic macro has two alternative representations in the \gls*{cas} or vice versa. In such cases, the translator picks one of the alternatives and informs the user about the decision.

In case of differences between defined branch cuts we can also use alternative translations to solve the problems. Consider the mentioned case of the arccotangent function~\parencite{Branches:acot} that has different positioned branch cuts in \Maple{} compared to the \gls*{dlmf} or \Mathematica{} definitions. Alternative but mathematically equivalent translations are
\begin{eqnarray}
\verb|\acot@{z}| & \overset{\langMaple}{\mapsto} & \verb|arccot(z)|,\label{eq:acot-alternatives}\\
& \overset{\langMaple}{\mapsto} & \verb|arctan(1/z)|,\label{eq:acot-alternatives-1}\\
& \overset{\langMaple}{\mapsto} & \verb|I/2*ln((z-I)/(z+I))|.\label{eq:acot-alternatives-2}
\end{eqnarray}

The position of the branch cut of the arccotangent function differs after translation~\eqref{eq:acot-alternatives}, which may lead to incorrect calculations later on. The alternative translations~\eqref{eq:acot-alternatives-1} and~\eqref{eq:acot-alternatives-2} use other functions instead of the arccotangent function. The arctangent function~\eqref{eq:acot-alternatives-1} and the natural logarithm~(\ref{eq:acot-alternatives-2}) have the same positioned branch cuts as in the \gls*{dlmf} and in \Maple. In consequence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user do not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z = -\iunit$.

Other problematic cases for translations are the \Macro s themselves. In some cases, they do not provide sufficient semantic information to perform translations. One example is the \textit{Wronskian} determinant. For two differentiable functions, the \textit{Wronskian} is defined as~\parencite[(1.13.4)]{NIST:DLMF}
\begin{equation*}
	\mathscr{W} \left\{ w_1(z), w_2(z) \right\} = w_1(z)w_2'(z) - w_2(z)w_1'(z).
\end{equation*}
In semantic \LaTeX{} it is currently implemented using
\begin{equation}
	\verb|\Wronskian@{w_1(z), w_2(z)}|.
\end{equation}
A translation become unfeasible because the macro does not explicitely define the variable of differentiation for the functions $w_1$ and $w_2$. For a correct translation, the \gls*{cas} needs to be aware of the used variable $z$. We solved this issue by creating a new macro, e.g.,
\begin{equation}
	\verb|\Wron{z}@{w_1(z)}{w_2(z)}|.
\end{equation}
This example demonstrates that the \Macro s are still a work in progress and are getting constantly updated.

A similar problem is multiplications since they are rarely explicitly marked in \LaTeX{} expressions, e.g., scientists using whitespaces to indicate multiplications rather than using \verb|\cdot| or similar symbols. For such problems, we introduced a new macro \verb|\idot| for an invisible multiplication symbol (this macro will not be rendered). Since this macro is newly introduced by contributers of the \gls*{drmf} team, and automatic conversion of existing equations is difficult, none of the equations in the \gls*{dlmf} use this macro. As a consequence, the translator has some simple rules to perform translations without explicitly marked translations with \verb|\idot|.

The \Macro s do not garantee entirely disambiguated expressions. In~\Cref{tab:amb-latex} there are four examples of potentially ambiguous expressions. These expressions are unambiguous for the \LaTeX{} compiler since it only considers the very next token for superscripts and subscripts. Our translator follows the same rules to solve these issues.

\begin{table}[ht]
\centering
\begin{tabular}{cc}
	\hline
	Potentially Ambiguous Input & \LaTeX{} Output\\
	\hline
	\verb|n^m!| & $n^m!$\\
	\verb|a^bc^d| & $a^bc^d$\\
	\verb|x^y^z| & Double superscript error\\
	\verb|x_y_z| & Double subscript error\\
	\hline
\end{tabular}
\caption{Potentially ambiguous \LaTeX{} expressions and how \LaTeX{} displays them.}
\label{tab:amb-latex}
\end{table}

Another more questionable translation decision are alphanumerical expressions. As explained in~\Cref{tab:allTypesTable}, the \gls*{pom}-tagger handles strings of letters and numbers differently, depending on the order of the symbols. The reason is, that an expression such as `$4b$' is usually considered to be a multiplication of $4$ and `$b$,' while `$b4$' looks like indexing `$b$' by $4$. While the first example produces two nodes, namely $4$ and `$b$', the second example `$b4$' produces just a single alphanumerical node in the \gls*{pom-pt}. The translator interprets alphanumerical expressions as multiplications for two reasons: (1) we would assume that the inputs `$4b$' and `$b4$' are mathematically equivalent; and (2) it is more common in mathematics to use single letter names for variables~\parencite{Notation:History}. Therefore we define the following definitions

\begin{eqnarray*}
\verb|4b| & \overset{\langMaple}{\mapsto} & \verb|4*b|,\\
\verb|b4| & \overset{\langMaple}{\mapsto} & \verb|b*4|,\\
\verb|energy| & \overset{\langMaple}{\mapsto} & \verb|e*n*e*r*g*y|.
\end{eqnarray*}

In general, the translator is drafted to solve ambiguous expressions or automatically find a work-around to disambiguate the expression. Only in case there is no way to solve the ambiguity with the defined rules, the translation process stops.

\section{The Translator}
All translations are defined by a library (\gls*{csv} and \gls*{json} files) that defines translation patterns for each function and symbol. The pattern uses \verb|$i| as placeholders to define the positions of the arguments. For example, the translation patterns for the Jacobi polynomial are illustrated in~\Cref{tab:placeholder_ex2}.

\begin{table}[ht]
	\centering
	\begin{tabular}{lc}
		\hline
		\multicolumn{2}{l}{\textit{Forward Translation:}} \\
		\Maple & \verb|JacobiP($2, $0, $1, $3)| \\
		\Mathematica & \verb|JacobiP[$2, $0, $1, $3]|\\
		\hline
		\multicolumn{2}{l}{\textit{Backward Translation from \Maple/\Mathematica:}} \\
		Semantic \LaTeX & \verb|\JacobiP{$1}{$2}{$0}@{$3}|\\
		\hline
	\end{tabular}
	\caption{Forward and backward translation patterns for the Jacobi polynomial example~\eqref{eq:P} in this manuscript. The pattern for the backward translation is the same for \Maple{} and \Mathematica.}
	\label{tab:placeholder_ex2}
\end{table}

These placeholders causes trouble when the \gls*{cas} uses the symbol \verb|$| for other reasons, e.g., the differentiation in \Maple{} is defined by
\begin{equation*}
\verb|diff(f, [x$n])|,
\end{equation*}
where $f$ is an algebraic expression or an equation, $x$ is the name of the differentiation variable, and $n$ is an integer representing the $n$-th order differentiation\footnote{\url{https://www.maplesoft.com/support/help/maple/view.aspx?path=diff}, seen 07/2018}. A translation for $\frac{d^2x^2}{dx^2}$ should look like this
\begin{equation*}
\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [x$2])|
\end{equation*}
but would end up as
\begin{equation*}
\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [xx])|.
\end{equation*}

We can solve this issue by using parentheses in such cases, e.g., \verb|diff($1, [$2$($0)])|.

The \Macro s also allow one to specify optional arguments to distinguish between standard and another version of these functions. The Legendre and associated Legendre function of the first kind are examples of such cases. The library that defines translations for each macro uses the macro name as the primary key to identify the translations. The Legendre and associated Legendre function of the first kind both use the same macro \verb|\LegendreP|. To distinguish such cases, we use a special syntax for such cases, shown in~\Cref{tab:legendreP-lex}.

\begin{table}[ht!]
	\centering
	\begin{tabular}{ll}
		\hline
		Semantic Macro Entry & Maple Entry \\
		\hline
		\verb|\LegendreP{\nu}@{x}| & \verb|LegendreP($0, $1)| \\
		\verb|X1:\LegendrePX\LegendreP[\mu]{\nu}@{x}| & \verb|LegendreP($1, $0, $2)|\\
		\hline
	\end{tabular}
	\caption{Example entries of the Legendre and associated Legendre function in the translation library. The prefix notation \texttt{X<d>:<name>X} defines the translation for \texttt{<name>} with \texttt{<d>}-number of optional arguments.}
	\label{tab:legendreP-lex}
\end{table}

The translator uses the \gls*{pom}-Tagger~\parencite{POM-Tagger}\footnote{Named according to the Part-of-Speech-Taggers in \gls*{nlp}.} to parse \LaTeX{} expressions into a parsed tree. The \gls*{pom}-Tagger is an \texttt{LL}-Parser defined by a context-free grammar in \gls*{bnf}. Each token will be tagged by meta information defined in lexicon files. We extend the lexicon files to provide also the information that is necessary for the translation process. An example of an entry of the lexicon file is given in~\Cref{tab:sine-lex-example}.

\begin{table}[ht!]
	\centering
	\begin{tabular}{lll}
	\hline
	\multicolumn{3}{l}{Symbol: \texttt{\textbackslash sin}} \\
	\! & \multicolumn{2}{l}{Feature Set: dlmf-macro} \\
	\! & \! & DLMF: \verb|\sin@@{z}|\\
	\! & \! & DLMF-Link: dlmf.nist.gov/4.14.E1\\
	\! & \! & Meanings: Sine\\
	\! & \! & Number of Parameters: 0\\
	\! & \! & Number of Variables: 1\\
	\! & \! & Number of Ats: 2\\
	\! & \! & Maple: \verb|sin($0)|\\
	\! & \! & Maple-Link: www.maplesoft.com/support/\\
	\! & \! & \hspace{32pt} help/maple/view.aspx?path=sin\\
	\! & \! & Mathematica: \verb|Sin[$0]|\\
	\! & \! & Mathematica-Link: reference.wolfram.com/\\
	\! & \! & \hspace{32pt} language/ref/Sin.html\\
	\hline
	\end{tabular}
	\caption{The entry of the trigonometric sine function in the lexicon file.}
	\label{tab:sine-lex-example}
\end{table} \section{Forward Translations}\label{sec:forward-translation}
An abstract translator class analyzes each node of the parsed tree and delegates them to specialized subtranslators. In this process, an object called \gls*{teo} is built. This \gls*{teo} is needed to rebuild a string representation of the mathematical expression after all nodes of the parsed tree were translated.

The parsed tree generated by the \gls*{pom}-tagger is not a mathematical expression tree. The \gls*{pom} project aims to disambiguate mathematical \LaTeX{} expressions and generates an expression tree. However, in the current state, many expressions cannot be disambiguated yet. In consequence, the \gls*{pom}-tagger generates a raw parsed tree where each token in the \LaTeX{} expression is a node in the tree. We call this parsed tree, the \gls*{pom-pt}.

{\sloppy The overall forward translation process is explained in~\Cref{fig:forward-trans}. All translation patterns and related information are stored in the DLMF/DRMF tables. These tables are converted by the \verb|lexicon-creator| to the \verb|DLMF-macros-lexicon| lexicon file. Together with the \verb|global|-\verb|lexicon| file, the \gls*{pom-pt} will be created by the \gls*{pom}-tagger. The \verb|latex-converter| takes a string representation of a semantic \LaTeX{} expression and uses the \gls*{pom} engine as well as our \verb|Translator| to create an appropriate string representation for a specified \gls*{cas}.}

\begin{figure}[ht]
	\vspace{-10pt}
	\centering
	\includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm, scale=0.72]{ForwardTranslationProcess.pdf}
	\caption{Process diagram of a forward translation process. The \gls*{pom}-tagger generates the \gls*{pom-pt} based on lexicon and \gls*{json} files. The \gls*{pom-pt} will be translated to different \gls*{cas}.}
	\label{fig:forward-trans}
	\vspace{-10pt}
\end{figure}

\subsection{Analyzing the PoM-Parsed Tree}\label{subsec:analyze-mlp}
Since the \gls*{bnf} does not define rules for semantic macros, each argument of the semantic macro and each $@$ symbol are following siblings of the semantic macro node. That is the reason why we stored the number of parameters, variables and $@$ symbols in the lexicon files. Otherwise, the translator could not find the end of a semantic macro in the \gls*{pom-pt}.

\begin{figure}[ht]
	\centering
	\includegraphics[clip, trim=0.2cm 0.2cm 0.2cm 0.2cm, scale=0.75]{SyntaxTreeUseCase.pdf}
	\caption{The \gls*{pom-pt} for the Jacobi polynomial example~\eqref{eq:P} using the DLMF/DRMF \LaTeX{} macro. Each leaf contains information from the lexicon files.}
	\label{fig:syntax-tree-usecase}
	\vspace{-10pt}
\end{figure}

\begin{wrapfigure}{r}{0.5\textwidth}
\vspace{-20pt}
\begin{minipage}{0.5\textwidth}
\begin{algorithm}[H]
\caption{Simple translation algorithm for the POM-Parse Tree}\label{alg:simple-translation}
	\begin{algorithmic}[1]
	\Require Root $r$ of the parse tree $T$
	\Procedure{translate}{$r$}
	\If{$r$ is leaf}
		\State {\scriptsize TRANSLATE\_LEAF}($r$);
	\Else
		\ForAll{children $v_n$ of $r$}
			\State {\scriptsize TRANSLATE}($v_n$);
		\EndFor
	\EndIf
	\EndProcedure
	\end{algorithmic}
\end{algorithm}
\end{minipage}
\vspace{-14pt}
\end{wrapfigure}

\Cref{fig:syntax-tree-usecase} visualizes the \gls*{pom-pt} of the Jacobi polynomial example from~\Cref{tab:JacobiP-usecase}. Because of the differences to expression trees, a backward conversion of the \gls*{pom-pt} to a string representation can be difficult, especially for finding necessary or unnecessary parentheses. Therefore we create the \gls*{teo}.

With these tools, we can translate a \LaTeX{} expression by translating the \gls*{pom-pt} node by node and perform group or reordering operations for some special cases. The algorithm is realized in a simple recursive structure. The Algorithm~\ref{alg:simple-translation} explains the translation process. Whenever the algorithm finds a leaf, it can translate this single term. If the node is not a leaf, it starts to translate all children of the node recursively.

\begin{algorithm}[t]
\caption{Abstract translation algorithm to translate MLP-Parse trees.}\label{alg:translation}
	\begin{algorithmic}[1]
	\Require Root $r$ of a POM-Parse tree $T$. List \textit{following\_siblings} with the following siblings of $r$. The list can be empty.
	\Procedure{abstract\_translator}{$r$, \textit{following\_siblings}}
	\If{$r$ is leaf}
		\State {\scriptsize TRANSLATE\_LEAF}($r$, \textit{following\_siblings});\label{line:transLeaf}
	\Else
		\State \textit{siblings} $ = r$.getChildren(); \Comment{\textit{siblings} is a list of children}
		\State {\scriptsize ABSTRACT\_TRANSLATOR}(\textit{siblings}.removeFirst(), \textit{siblings});\label{line:transSeq}
	\EndIf
	\If{\textit{following\_siblings} is not empty}\label{line:ifNotEmpty}
		\State $r =$ \textit{following\_siblings}.removeFirst();
		\State {\scriptsize ABSTRACT\_TRANSLATOR}($r$, \textit{following\_siblings});
	\EndIf
	\EndProcedure
	\end{algorithmic}
\end{algorithm}

This approach is not completely feasible since the algorithm needs to look ahead and check the following siblings in some cases, e.g., in the case of a semantic macro with arguments (see~\Cref{fig:syntax-tree-usecase}). The {\footnotesize \verb|TRANSLATE_LEAF|} function needs to know the following siblings of the current node. Algorithm~\ref{alg:translation} explains a more detailed but abstract version of the final algorithm of the presented translation tool.

If the root $r$ is a leaf, it still can be translated as a leaf. Eventually, some of the following siblings are needed to translate $r$. The list of \textit{following\_siblings} in Line~\ref{line:transLeaf} might be reduced to avoid multiple translations for one node. If $r$ is not a leaf, it contains one or more children. Therefore, we can call the {\footnotesize \verb|ABSTRACT_TRANSLATOR|} recursively for the children. Once we have translated $r$, we can go a step further and translate the next node. Line~\ref{line:ifNotEmpty} checks if there are following siblings left and calls the {\footnotesize \verb|ABSTRACT_TRANSLATOR|} recursively in such cases. Translated expressions are stored by the \gls*{teo} object. Algorithm~\ref{alg:translation} is a simplified version of the translator process. The \Cref{line:transLeaf,line:transSeq} process the translations for each node. \Cref{tab:allTypesTable} gives an overview of all the different node types the root $r$ can be. A more detailed explanation of the types can be found in~\parencite{POM-Tagger}.

The \gls*{bnf} grammar defines some basic grammatical rules for generic \LaTeX{} macros, such as for \verb|\frac|, \verb|\sqrt|. Therefore, there is a hierarchical structure for those symbols similar to the structure in expression trees. As already mentioned, some of these types can be translated directly, such as Greek letters, while others are more complex, such as semantic \LaTeX{} macros. Therefore, the translators delegate the translation to specialized subtranslators. This delegation process is implemented in \Cref{line:transLeaf,line:transSeq} of Algorithm~\ref{alg:translation}. The Subsection~\ref{sec:subtranslators} discusses these classes in more detail.

\begin{table}[t!]
\centering
\sloppy
\newcolumntype{L}{>{\raggedright\arraybackslash}X}
\newcolumntype{Y}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\begin{tabularx}{\textwidth}{Y{1.55cm} Y{2.5cm} L L}
	\hline
	& \textbf{Node type} & \textbf{Explanation} & \textbf{Example} \\
	\hline
	\hline
	\multicolumn{1}{Y{1.5cm}|}{\textbf{$r$ has children}} & Sequence & Contains a list of expressions. & $a+b$ is a sequence with three children ($a$, $+$ and $b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Balanced Expression & Similar to a sequence. But in this case the sequence is wrapped by \texttt{\tbs left} and \texttt{\tbs right} delimiters. & \texttt{\tbs left(}$a+b$\texttt{\tbs right)} is a balanced expression with three children ($a$, $+$ and $b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Fraction & All kinds of fractions, such as \texttt{\tbs frac}, \texttt{\tbs ifrac}, etc. & \texttt{\tbs ifrac\{a\}\{b\}} is a fraction with two children ($a$ and $b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Binomial & Binomials & \texttt{\tbs binom\{a\}\{b\}} has two children ($a$ and $b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Square Root & The square root with one child. & \texttt{\tbs sqrt\{a\}} has one child ($a$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Radical with a specified index & $n$-th root with two children. & \texttt{\tbs sqrt[a]\{b\}} has two children ($a$ and $b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Underscore & The underscore '\_' for subscripts. & The sequence $a$\_$b$ has two children ($a$ and '\_'). The underscore itself '\_' has one child ($b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Caret & The caret '\^{}' to for superscripts or exponents. Similar to the underscore. & The sequence $a$\^{}$b$ has two children ($a$ and '\^{}'). The caret itself '\^{}' has one child ($b$).\\
	\hline
	\hline
	\multicolumn{1}{Y{1.5cm}|}{\textbf{$r$ is a leaf}} & \Macro{} & A semantic \LaTeX{} macro & \texttt{\tbs JacobiP}, etc.\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Generic \LaTeX{} macro & All kinds of \LaTeX{} macros & \texttt{\tbs Rightarrow}, \texttt{\tbs alpha}, etc.\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Alphanumerical Expressions & Letters, numbers and general strings. & Depends on the order of symbols. $ab3$ is alphanumerical, while $4b$ are two nodes ($4$ and $b$).\\
	\cline{2-4}
	\multicolumn{1}{l|}{} & Symbols & All kind of symbols & '$@$', '$*$', '$+$', '$!$', etc.\\
	\hline
\end{tabularx}
\caption{A table of all kinds of nodes in a PoM syntax tree. Note that this table groups some kinds for a better overview. For a complete list and a more detailed version see~\parencite{POM-Tagger}.}
\label{tab:allTypesTable}
\end{table} 
This approach seems to work for most of semantic \LaTeX{} expressions. However, there is one problem with this straight forward approach which was mentioned shortly mentioned in~\Cref{sec:problems}. There are many different types of notations used to represent formulae. \Cref{tab:notations} illustrates the simple expression $(a+b)x$ in different notations. The \gls*{pol}\footnote{Also known as \textit{Warsaw Notation} or \textit{prefix notation}} (hereafter called prefix notation) places the operator to the left of/before its operands. The \gls*{rpol}\footnote{Also known as \textit{postfix notation}} (hereafter called postfix notation) does the opposite and places the operator to the right of/after its operands. The infix notation is commonly used in arithmetic and places the operator between their operands, which only makes sense as long as the operator is a binary operator.

\begin{wrapfigure}{r}{0.38\textwidth}
\vspace{-10pt}
\begin{minipage}{0.38\textwidth}
\center
\begin{tabular}{lc}
	\hline
	Notation & Expression \\
	\hline
	Infix & $(a+b) \cdot x$\\
	Prefix & $\cdot + a\ b\ x$\\
	Postfix & $a\ b + x\ \cdot$\\
	Functional & $\cdot(+(a, b), x)$\\
	\hline
\end{tabular}
\vspace{-5pt}
\caption{The mathematical expression '$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
\label{tab:notations}
\end{minipage}
\vspace{-5pt}
\end{wrapfigure}

In mathematical expressions, notations are mostly mixed, depending on the case and number of operands. For example, infix notation is common for binary operators ($+$, $-$, $\cdot$, $\mod$, etc.), while functional notations are conveniently used for any kind of functions ($sin$, $cos$, etc.). Sometimes the same symbol is used in different notations to distinguish different meanings. For example, the `$-$' as a unary operator is used in prefix notation to indicate the negative value of its operand, such as in `$-2$'. Of course, `$-$' can also be the binary operator for subtraction, which is commonly used in infix notation. An example for the postfix notation is factorial, such as `$2!$'.

Most programming languages (and \gls*{cas} as well) internally use prefix or postfix notation and do not mix the notations in one expression, since it is more convenient to parse expressions in uniform notations. However, the common practice in science is to use mixed notations in expressions. Since the \gls*{pom} has rarely implemented mathematical grammatical rules yet, it takes the input as it is and does not build an expression tree. Therefore, it parses all four examples from~\Cref{tab:notations} to four different \gls*{pom-pt}s rather than to one unique expression tree. In general, this is not a problem for our translation process since most \gls*{cas} are familiar with the most common notations. Therefore, the translator does not need to know that `$a$' and `$b$' are the operands of the binary operator `$+$' in `$a+b$.' The translator could simply translate the symbols in `$a+b$' in the same order as they appear in the expression and the \gls*{cas} would understand it. However, this simple approach generates two problems.
\begin{enumerate}\item \label{prob:1} The translated expression is only syntactically correct if the input expression was syntactically correct.
\item \label{prob:2} We cannot translate expressions to a \gls*{cas} which uses a non-standard notation.
\end{enumerate}

Problem~\ref{prob:1} should be obvious. Since we want to develop a translation tool and not a verification tool for mathematical \LaTeX{} expressions, we can assume syntactically correct input expressions and produce errors otherwise. Problem~\ref{prob:2} is more difficult to solve. If a user wants to support a \gls*{cas} that uses prefix or postfix notation by default, the translator would fail in its current state. Supporting \gls*{cas} with another notation would be a part of future work.

Nonetheless, changing a notation could also solve ambiguities in some situations. Consider the two ambiguous examples in~\Cref{tab:amb_ex}. While a scientist would probably just ask for the right interpretation of the first example, \Maple{} automatically computes the first interpretation. On the other hand, \LaTeX{} automatically disambiguate the first example by only recognizing the very next element (single symbols or sequence in curly brackets) for the superscript and therefore displays the second interpretation. The second example is already interpreted as the double factorial function of $n$, since this notation is the standard interpretation in science. We wrote the second interpretation as the standard way in science to make it even more obvious. However, surprisingly, \Maple{} computes the first interpretation again rather than the common standard interpretation.
\begin{table}[ht]
\centering
\begin{tabular}{lccc}
	\hline
	& Text Format Expression & First Interpretation & Second Interpretation\\
	\hline
	1: & \rule{0pt}{0.9\normalbaselineskip} $4\ \hat{\ }\ 2!$ & $4^{2!}$ & $4^2!$ \\
	2: & $n!!$ & $(n!)!$ & $(n)!!$\\
	\hline
\end{tabular}
\caption{Ambiguous examples of the factorial and double factorial function. One expression in a text format can be interpreted in different ways.}
\label{tab:amb_ex}
\end{table}

In most cases, parentheses can be used to disambiguate expressions. We used them in~\Cref{tab:amb_ex} to clarify the different interpretations in example 2. But sometimes, even parentheses cannot solve a mistaken computation. For example, there is no way to add parentheses to force \Maple{} to compute `$n!!$' as the double factorial function. Even `$(n)!!$' will be interpreted as `$(n!)!$'. Rather than using the exclamation mark in \Maple, one could also use the functional notation. For example, the interpretations `$(2!)!$' and `$(2)!!$' can be distinguished in \Maple{} by using \verb|factorial(factorial(2))| and \verb|doublefactorial(2)| respectively. We define the translations as follows:

\begin{eqnarray*}
\verb|n!| &\overset{\langMaple}{\mapsto}& \verb|factorial(n)|,\\
\verb|n!!| &\overset{\langMaple}{\mapsto}& \verb|doublefactorial(n)|.
\end{eqnarray*}

Algorithm~\ref{alg:translation} does not allow this translation right now. It has no access to previously translated nodes in its current state. This problem is solved by the \gls*{teo} that stores and groups translated objects like lists. This allows to access the latest translated expression and use it as the argument for the factorial function. Table~\ref{tab:teo-list} shows three example for the \gls*{teo} list that groups some tokens.

\begin{table}[ht]
\centering
\begin{tabular}{cc}
	\hline
	Input Expression & TEO List\\
	\hline
	$a+b$ & \verb|[a, +, b]|\\
	$(a+b)$ & \verb|[(a+b)]|\\
	$\frac{a}{b}-2$ & \verb|[(a)/(b), -, 2]|\\
	\hline
\end{tabular}
\caption{How the TEO-list groups subexpressions.}
\label{tab:teo-list}
\end{table}

\subsection{Subtranslators}\label{sec:subtranslators}

A \verb|SequenceTranslator| translates the \textit{sequence} and \textit{balanced expressions} in the \gls*{pom-pt}. If a node $n$ is a leaf and the represented symbol an open bracket (parentheses, square brackets and so on) the following nodes are also taken as a \textit{sequence}. Hence, combined with the recursive translation approach, the \verb|SequenceTranslator| also checks balances of parentheses in expressions. An expression such as `$(a]$' is producing a mismatched parentheses error. On the other hand, this is a problem for interval expressions such as `$[a,b)$'. In the current version, the program cannot distinguish between mismatched parentheses and half-opened, half-closed intervals. Whether an expression is an interval or another expression is difficult to decide and can depend on the context. Also, the parenthesis checker could simply be deactivated to allow mismatched parentheses in an expression. But this functionality is not implemented yet.

The \verb|SequenceTranslator| also handle positions of multiplication symbols. There are a couple of obvious choices to translate multiplication signs. The most common symbol for multiplications is still the white space or none symbol at all, as explained previously. Consider the simple expression `$2n\pi$'. The \gls*{pom-pt} generates a sequence node with three children, namely $2$, $n$ and $\pi$. This sequence should be interpreted as a multiplication of the three elements. The \verb|SequenceTranslator| checks the types of the current and next nodes in the tree to decide if it should add a multiplication symbol or not. For example, if the current or next node is an operator, a relation symbol or an ellipsis, there will be no multiplication symbol added. However, this approach implies an important property. The translator interprets all sequences of nodes as multiplications as long as it is not defined otherwise. This potentially produces strange effects. Consider an expression such as `$f(x)$'. Translate this to \Maple{} will be \verb|f*(x)|. But we do not consider this translation to be wrong, because there is a semantic macro to represent functions. In this case, the user should use \verb|\f{f}@{x}| instead of \verb|f(x)| to distinguish between $f$ as a function call and $f$ as a symbol.

\begin{algorithm}[!ht]
\caption{The translate function of the MacroTranslator. This code ignores error handling.}\label{alg:macro-translation}
	\begin{algorithmic}[1]
	\Require 
		\Statex \textit{macro} - node of the semantic macro. 
		\Statex \textit{args} - list of the following siblings of \textit{macro}. 
		\Statex \textit{lexicon} - lexicon file
	\Ensure 
		\Statex Translated semantic macro.

	\Procedure{translate\_macro}{\textit{macro}, \textit{args}, \textit{lexicon}}
	\State \textit{info} = \textit{lexicon}.getInfo(\textit{macro});
	\State \textit{argList} = new List(); \Comment{create a sorted list for the translated arguments.}
	\State \textit{next} = \textit{args}.getNextElement();
	\If{\textit{next} is caret}\label{line:next_caret}
		\State \textit{power} = translateCaret(\textit{next});
		\State \textit{next} = \textit{args}.getNextElement();
	\EndIf

	\While{\textit{next} is $[$}\label{line:next_optional} \Comment{square brackets starts a balanced sequence of optional arguments.}
		\State \textit{optional} = {\scriptsize TRANSLATE\_UNTIL\_CLOSED\_BRACKET}(\textit{args});
		\State \textit{argList}.add(\textit{optional});
		\State \textit{next} = \textit{args}.getNextElement();
	\EndWhile

	\State \textit{argList}.add( {\scriptsize TRANSLATE\_PARAMETERS}(\textit{args}, \textit{info}) ); \label{line:trans_paras} \Comment{number is given in \textit{info}.}
	\State {\scriptsize SKIP\_AT\_SIGNS}( \textit{args}, \textit{info} );  \label{line:skip_ats} \Comment{number is given in \textit{info}.}
	\State \textit{argList}.add( {\scriptsize TRANSLATE\_VARIABLES}(\textit{args}, \textit{info}) ); \label{line:trans_vars} \Comment{number is given in \textit{info}.}

	\State \textit{pattern} = \textit{info}.getTranslationPattern();
	\State \textit{translatedMacro} = \textit{pattern}.fillPlaceHolders(\textit{argList});
	\If{\textit{power} is not \NULL}
		\State \textit{translatedMacro}.add(\textit{power});
	\EndIf
	\State \Return\ \textit{translatedMacro};
	\EndProcedure
	\end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg:macro-translation} explains the \verb|MacroTranslator| without error handling. It extracted necessary information from the \gls*{pom-pt}, such as how many arguments this function has. In consequence, it also processes the following siblings to translate the arguments. There are some special cases about the next node right after the macro occurs. This node can be
\begin{itemize}
\item an exponent, such as for '\verb|^2|';
\item an optional parameter in square brackets;
\item a parameter in curly brackets (a \textit{sequence} node in the \gls*{pom-pt}) if none of the above;
\item an $@$ symbols if none of the above;
\item or a variable in curly brackets (a \textit{sequence} node) if none of the above.
\end{itemize}

An exponent will be translated and shifted to the end of the translated semantic macro, since it is more common to write the exponent after the arguments of a function in \gls*{cas}. Therefore, the function translates and stores the exponent in Line~\ref{line:next_caret}. One could ask what happens when there is an exponent given before and after the arguments. The \verb|MacroTranslator| only translates the following siblings until each argument is translated. The first exponent will be shifted to the end. If right after the translated macro (with all arguments) follows another exponent, we interpret it as another exponent for the whole previous expression. In that case, it would be the macro with the first translated exponent. Table~\ref{tab:multi-expo} shows an example for the trigonometric cosine function with multiple exponents. 
\begin{table}[ht]
\centering
\begin{tabular}{lc}
\hline
Displayed As & \rule{0pt}{0.9\normalbaselineskip} $\cos^2(x)^2$ \\
Semantic \LaTeX & \verb|\cos^2@{x}^2| \\
Translated \Maple{} Expression & \verb|((cos(x))^(2))^2|\\
\hline
\end{tabular}
\caption{A trigonometric cosine function example with exponents before and after the argument.}
\label{tab:multi-expo}
\end{table}

As you can see, the input expression in semantic \LaTeX{} is interpreted as
\begin{equation}\label{eq:multi-expo}
\left( \cos(x)^2 \right)^2.
\end{equation}
 \section{Maple to Semantic \LaTeX{} Translator}\label{sec:backward-translation}
Instead of writing a custom \Maple{} syntax parser, we using \Maple's internal data structure to get a syntax tree of the input\footnote{In consequence, a license of \Maple{} is mandatory to perform backward translations. The translator is using the version \Maple{} 2016.}. \Maple{} allow different input styles. The \texttt{1D} input is mainly used for programming purposes and therefore also used to perform our translations. Internally, \Maple{} using \gls*{dag} as syntax trees.

\begin{wrapfigure}{r}{0.61\textwidth}
	\vspace{-22pt}
	\centering
	\includegraphics[clip, trim=0.5cm 0.5cm 0.5cm 0.5cm, scale=0.5]{DAGreal.pdf}
	\vspace{-5pt}
	\caption{The internal \Maple{} DAG representation of $x^2+x$.}
	\label{fig:internal-maple-dag}
	\vspace{-15pt}
\end{wrapfigure}

Each node in the \gls*{dag} stores its children and has a header, which defines the type and the length of the node. Consider the polynomial $x^2+x$. Figure~\ref{fig:internal-maple-dag} illustrates the internal \gls*{dag} representation with headers and arguments.

One can access the internal data structure of expressions via the \texttt{ToInert} command. The returned \inertF{} is a nested list representation as a tree equivalent\footnote{A tree equivalent of a \gls*{dag} splitting nodes with multiple parents into multiple nodes so that each node has only one parent node.} of the internal \gls*{dag} for the given expression. Some of the important types for the nodes are specified in \Cref{tab:maple-types}.

\begin{table}[h]
\centering
\begin{tabular}{cp{10cm}}
	\hline
	Type & Explanation\\
	\hline
	SUM & Sums. Internally stored with factors for each summand, i.e., '$x+y$' would be stored as '$x \cdot 1 + y \cdot 1$'.\\
	PROD & Products.\\
	EXPSEQ & Expression sequence is a kind of list. The arguments of functions are stored in such sequences.\\
	INTPOS & Positive integers.\\
	INTNEG & Negative integers.\\
	COMPLEX & Complex numbers with real and imaginary part.\\
	FLOAT & Float numbers are stored in the scientific notation with integer values for the exponent $n$ and the significand $m$ in $m \cdot 10^n$.\\
	RATIONAL & Rational numbers are fractions stored in integer values for the numerator and positive integers for the denominator.\\
	POWER & Exponentiation with expressions as base and exponent.\\
	FUNCTION & Function invocation with the name, arguments and attributes of the function.\\
	\hline
\end{tabular}
\caption{A subset of important internal \Maple{} data types. See~\parencite{MAPLE:ProgrammingGuide} for a complete list.}
\label{tab:maple-types}
\end{table}

The translator using the \texttt{OpenMaple}~\parencite[\S 14.3]{MAPLE:ProgrammingGuide} \gls*{api} for interacting with \Maple's kernel and get access to the \inertF{} of inputs.

\subsection{Automatic Changes of Inputs in Maple}\label{subsec:maple-probs}
\Maple{} evaluates inputs automatically and changes the input into an internal representation. This internal representation might look a bit different to the input. One example has already been given with \Cref{fig:internal-maple-dag}, where each summand of a sum is stored with a factor. Here is a list of all internal changes that occur for inputs.

\begin{itemize}
\item \Maple{} evaluates input expressions immediately.
\item There is no data type to represent square roots such as $\sqrt{x}$ (or $n$-th roots). Therefore, \Maple{} stores roots as an exponentiation with a fractional exponent. For example, $\sqrt{x}$ is stored as $x^{\frac{1}{2}}$.
\item There is no data type for subtractions, only for sums. Negative terms are changed to absolute values times '$-1$'. For example, $x-y$ is stored as $x + y \cdot (-1)$. 
\item Floating point numbers are stored in the scientific notation with a mantissa and an exponent in the base $10$. For example, $3.1$ is internally represented as $31 \cdot 10^{-1}$.
\item There is only a data type for rational numbers (fractions with integer numerator and positive denominator), but not for general fractions, such as $\frac{x+y}{z}$. This will be automatically changed to $(x+y)\cdot z^{-1}$.
\end{itemize}

There are unevaluation quotes implemented to avoid evaluations on input expression. Table~\ref{tab:unevaluation-quotes} gives an example how those unevaluation quotes work.

\begin{table}[ht]
\centering
\begin{tabular}{lcc}
\hline
& Without unevaluation quotes & With unevalation quotes\\
\hline
Input expression: & \texttt{sin(Pi)+2-1} & \texttt{'sin(Pi)+2-1'}\\
Stored expression: & 1 & \texttt{sin(Pi)+1}\\
\hline
\end{tabular}
\caption{Example of unevaluation quotes for \texttt{1D} \Maple{} input expressions.}
\label{tab:unevaluation-quotes}
\end{table}

Since we want to keep a translated expression similar to the input expression, we implemented some cosmetic rules for the backward translations that solve or reduce the effects from the list of changes above. 
\begin{itemize}
\item We use unevaluation quotes to suppress evaluations of the input.
\item We perform reordering of factors and summands so that negative factors appear in front of the summand. This gives us the opportunity to translate $x - y$ to $x - y$ instead of $x + y \cdot (-1)$.
\item We introduced the new internal data types \texttt{MYFLOAT} and \texttt{DIVIDE} to translate floats and fractions in more convenient notations.
\end{itemize}

The translation process then follows the same principle as for the forward translations. Since the syntax tree of \Maple{} is an expression tree, we do not need to implement special reordering or grouping algorithms to perform backward translations. Translations for functions are also realized via patterns and placeholders. Figure~\ref{fig:backward-trans} illustrates the backward translation process for the Jacobi polynomial example from \Cref{tab:JacobiP-usecase}.

\begin{figure}[t!]
	\centering
	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.7]{MapleTranslation.pdf}
	\caption{A scheme of the backward translation process from \Maple{} for the Jacobi polynomial $P_n^{(\alpha , \beta)}(\cos(a\Theta))$. The input string is converted by the \Maple{} kernel into the nested list representation. This list is translated by subtranslators (blue and red arrows). A function translation (bold blue arrows) is again realized by translation patterns to define the position of the arguments (red arrows).}
	\label{fig:backward-trans}
\end{figure}

\section{Evaluation}\label{sec:evaluation}
We implemented three approaches to evaluate whether a translation was \textit{appropriate} or \textit{inappropriate}.
\begin{enumerate}
\item \textbf{Round Trip Tests}: translates expressions back and forth and analyzing the changes.
\item\label{approach-2} \textbf{Function Relation Tests (Symbolical)}: translate mathematically proven equivalent expressions from one system to a \gls*{cas} and evaluate whether the relation remains valid via symbolical equivalence checks.
\item \textbf{Numerical Tests}: take the same equations from Approach~\ref{approach-2} but evaluates them on specific numerical values to test equivalence.
\end{enumerate}

\subsection{Round Trip Tests}\label{sec:round-trip}
A round trip test always starts with a valid expression either in semantic \LaTeX{} or in \Maple. A translation from one system to another is called \textbf{a step}. A complete round trip translation (two steps) is called \textbf{one cycle}. A \textbf{fixed point representation} (or short fixed point) in a round trip translation process is a string representation that is identical to all string representations in the following cycles.

Table~\ref{tab:fixpoint} illustrates an example of a round trip test which reaches a fixed point. The test formula is
\begin{equation}
\frac{\cos\left(a\Theta\right)}{2}.
\end{equation}

\begin{table}[ht]
\centering
\begin{tabular}{cc}
	\hline 
	Steps & semantic \LaTeX{}/\Maple{} representations\\
	\hline
	\rule{0pt}{0.9\normalbaselineskip}0 & \verb|\frac{\cos@{a\Theta}}{2}|\\
	1 & \verb|(cos(a*Theta))/(2)| \\
	2 & \verb|\frac{1}{2}\idot\cos@{a\idot\Theta}| \\
	3 & \verb|(1)/(2)*cos(a*Theta)|\\
	4 & \verb|\frac{1}{2}\idot\cos@{a\idot\Theta}| \\
	\hline
\end{tabular}
\caption{A round trip test reaching a fixed point.}
\label{tab:fixpoint}
\end{table}

Step 4 is identical to step 2, and since the translator is a deterministic algorithm it can be easily shown that step 2 and step 3 are fixed-point representations for semantic \LaTeX{} and \Maple.

There is currently only one exception known where a round trip test does not reach a fixed point representation: Legendre's incomplete elliptic integrals~\parencite[(19.2.4-7)]{NIST:DLMF} are defined with the amplitude $\phi$ in the first argument in the \DLMF, while \Maple{} takes the trigonometric sine of the amplitude as the first argument. Therefore, the forward and backward translations are defined as
\begin{eqnarray}
\verb|\EllIntF@{\phi}{k}| & \overset{\langMaple}{\mapsto} & \verb|EllipticF(sin(phi),k)|,\\
\verb|\EllIntF@{\asin@{\phi}}{k}| & \overset{\langMaple}{\mapsfrom} & \verb|EllipticF(phi,k)|,
\end{eqnarray}

The round-trip translations produce infinite chains of sine and inverse sine calls because there are no evaluations involved. 
The round trip tests are very successful, but they only detect errors in string representations. However, because of the simplification techniques of fixed points, we are able to at least detect logical errors in one system: \Maple. On the other hand, these tests cannot determine logical errors in the translations between the two systems. Consider we mistakenly defined an \textit{inappropriate} forward and backward translation for the sine function
\begin{eqnarray}
\verb|\sin@{\phi}| & \overset{\langMaple}{\leftrightarrow} & \verb|cos(phi)|,\label{eq:wrong-trans-1}\\
\verb|\cos@{\phi}| & \overset{\langMaple}{\leftrightarrow} & \verb|sin(phi)|.\label{eq:wrong-trans-2}
\end{eqnarray}
In that case the round trip test would not detect any errors and reach a fixed point representation, because the simplification techniques only simplify two representations in the same system but cannot compare the representation in one system to those in the other.

\subsection{Function Relation Tests}\label{sec:relation-tests}
The \gls*{dlmf} is a compendium of special functions and orthogonal polynomials and lists several relations between the functions and polynomials. The idea of this evaluation approach is to translate an entire relation and test whether the relation remains valid after performing the translations.

With this technique we can detect translation errors such as in \eqref{eq:wrong-trans-1} and \eqref{eq:wrong-trans-2}. Consider the \gls*{dlmf} equation for the sine and cosine function~\parencite[(4.21.2)]{NIST:DLMF}
\begin{equation}
\sin \left(u+v\right) = \sin{u}\cos{v} + \cos{u}\sin{v}.
\end{equation}
Assume the translator would forward translate the expression based on (\ref{eq:wrong-trans-1},~\ref{eq:wrong-trans-2}). Than
\begin{eqnarray}
\verb|\sin@{u + v}| & \overset{\langMaple}{\mapsto} & \verb|cos(u + v)|,\\
\verb|\sin@@{u}\cos@@{v}| & \overset{\langMaple}{\mapsto} & \verb|cos(u)*sin(v)|,\\
\verb|\cos@@{u}\sin@@{v}| & \overset{\langMaple}{\mapsto} & \verb|sin(u)*cos(v)|.
\end{eqnarray}
This produces the equation in \Maple
\begin{equation}
\cos\left(u+v\right) = \cos{u}\sin{v} + \sin{u}\cos{v},
\end{equation}
which is wrong. Since the expression is correct before the translation, we conclude an error during the translation process.

However, there are two essential problems with this approach. Testing the mathematical equivalence of expressions is hard to solve and \gls*{cas} have trouble to test even simple equations. Furthermore, this approach only checks forward translations because there is no way to check equivalence of expressions in \LaTeX{} automatically (again this could become feasible with our translator). We use \Maple's \textit{simplify} function to check if the difference of the left-hand side and the right-hand side of the equation is equal to zero. In addition, we use \textit{simplify} and check if the division of the right-hand side by the left-hand side returns a numerical value or not. This simplification function is the most powerful function to check the equivalence in \Maple. However, there are several cases where the simplification fails. Because of implementation details, there are some techniques that helps \Maple{} to find possible simplifications. For example we can force \Maple{} to convert the formula
\begin{equation}
\sinh{x} + \sin{x}
\end{equation}
to an equivalent representation using the exponential representations
\begin{equation}
\frac{1}{2}\expe^x - \frac{1}{2}\expe^{-x} - \frac{1}{2} \iunit \left( \expe^{\iunit x}-\expe^{-\iunit x} \right).
\end{equation}
With such pre-conversions we are able to improve the simplification process in \Maple. However, the limitations of the \textit{simplify} function are still the weakest part of this verification approach. Consider the complex example~\parencite[(12.7.10)]{NIST:DLMF}
\begin{equation}\label{eq:branch-cut-near}
\displaystyle U(0,z) = \sqrt{\frac{z}{2\cpi}} K_{\frac{1}{4}}\left(\frac{1}{4}z^2\right),
\end{equation}
where $U(0,z)$ is the parabolic cylinder function and $K_\nu(z)$ the modified Bessel function of the second kind. Both functions are well-defined in both systems and we can define a \textit{direct} translation for~\eqref{eq:branch-cut-near}. 
The modified Bessel function of the second kind has its branch cut in \Maple{} and in the \gls*{dlmf} at $z < 0$. However, the argument of $K$ contains $z^2$. If $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ the value of the right-hand side of~\eqref{eq:branch-cut-near} would be no longer on the principal branch. However, \Maple{} will still compute the principal values independently of the value of $z$. Hence, a translation
\begin{equation}
\verb|\BesselK{\frac{1}{4}}@{\frac{1}{4}z^2}| \overset{\langMaple}{\mapsto} \verb|BesselK(1/4,(1/4)*z^2)|
\end{equation}
is incorrect if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ and one has to use analytic continuation for the right-hand side of equation (\ref{eq:branch-cut-near}).

To evaluate such complex cases, the equivalence checks of \gls*{cas} are insufficient. Therefore we implement numerical tests as an additional step.

\subsection{Numerical Tests}\label{sec:numerical-tests}
Consider the differences of the left- and right-hand side of equation~\eqref{eq:branch-cut-near}
\begin{equation}\label{eq:difference}
D(z) := U(0,z) - \sqrt{\frac{z}{2\cpi}} K_{\frac{1}{4}}\left(\frac{1}{4}z^2\right).
\end{equation}
Table~\ref{tab:computations-for-difference} presents four computations for $D(z)$, one value for each quadrant in the complex plane.
\begin{table}[ht]
\centering
\begin{tabular}{rcc}
	\hline
	$z\ \ $ & & $D(z)$\\
	\hline
	\tableRowSpace{} $1+\iunit$ & & $2 \cdot 10^{-10} - 2 \cdot 10^{-10} \iunit$\\
	$-1+\iunit$& & $2.222121916 - 1.116719816 \iunit$\\
	$-1-\iunit$& & $2.222121916 + 1.116719816 \iunit$\\
	$1-\iunit$ & & $2 \cdot 10^{-10} + 2 \cdot 10^{-10} \iunit$\\
	\hline
\end{tabular}
\caption{Four computations of $D(z)$ in \Maple.}
\label{tab:computations-for-difference}
\end{table}

Considering machine accuracy and the default precision of $10$ significant digits, we can regard the first and last values as zero differences. While this evaluation is very powerful, it has a significant problem. Even when all tested values return zero, it does not proof the equivalence of~\eqref{eq:branch-cut-near}. However, when the values are different from zero, it does proof an error in one of the four cases~\parencite{NumericalTests:Paper}
\begin{enumerate}
\item the numerical engine tests invalid combinations of values;
\item the translation is incorrect;
\item there may be an error in the \gls*{dlmf} source; or
\item there may be an error in \Maple.
\end{enumerate}

\subsection{Results}\label{sec:test-summary}
There are 685 \Macro s\footnote{The macros are still work in progress, and therefore the total number is constantly changing.} in total, and 665 of them were implemented in the translator engine. We defined forward translations to \Maple{} for 201 and backward translations from \Maple{} for 195 functions.

The \gls*{dlmf} provides a dataset of \LaTeX{} expressions with semantic macros. We extracted 4.087 equations from \gls*{dlmf} and apply our round-trip and relation tests on them. The translator was able to translate 2.405\footnote{All percentages are approximately calculated.} (58.8\%) of the extracted equations without errors. 660 (27.4\%) of the successfully translated expressions were verified by the simplification techniques of \Maple. We apply additional numerical tests for the remaining 1.745 equations. For 418 (24\%) cases the numerical tests were valid. More detailed results for numerical and symbolical tests were presented in~\parencite{NumericalTests:Paper}.

The evaluation techniques have proven to be very powerful also for evaluating \gls*{cas} and online mathematical compendia such as the \gls*{dlmf}. During the evaluations, we were able to detect several errors in the translation and evaluation engine, and also discovered two errors in the \gls*{dlmf} and one error in \Maple's \texttt{simplify}-function.

The numerical test engine was able to discover a sign error in equation~\parencite[(14.5.14)]{NIST:DLMF}\footnote{The equation had originally been stated as shown in equation~(\ref{eq:signerror}). The error were reported on 10th April 2017.}
\begin{equation}\label{eq:signerror}
\displaystyle \mathsf{Q}^{-1/2}_{\nu}\left(\cos\theta\right)=-\left(\frac{\pi}{2\sin\theta}\right)^{1/2}\frac{\cos\left(\left(\nu+\frac{1}{2}\right)\theta\right)}{\nu+\frac{1}{2}}.
\end{equation}
The error can be found on~\parencite[p. 359]{NIST:Handbook} and has been fixed in the \gls*{dlmf} with version 1.0.16. The same engine also identified a missing comma in the constraint of~\parencite[(10.16.7)]{NIST:DLMF}. The original constraint was given by $2\nu \neq -1, -2 -3, \ldots$, with a missing comma after $-2$.

We have also noticed that our testing procedure is able to identify errors in CAS procedures,
namely the \Maple{} {\tt simplify} procedure. The left-hand side of \cite[(7.18.4)]{NIST:DLMF} is given by 
\begin{equation*}
\frac{{\mathrm{d}}^{n}}{{\mathrm{d}z}^{n}}\left(e^{z^{2}}\operatorname{erfc}z\right),
\end{equation*}
where $n \in \mathbb{N}_0$, $e$ is the base of the natural logarithm, and $\operatorname{erfc}$ is the
complementary error function. Our translation correctly produces 
\begin{equation*}
\verb|diff((exp((z)^(2))*erfc(z)), [z$(n))]|.
\end{equation*}
However, the \Maple{ 2016} \verb|simplify| function falsely returns $0$ for the translated left-hand side.
Maplesoft has confirmed in a private communication that this is indeed a defect in \Maple{ 2016}.
Furthermore, although the nature of the defect changes, the defect still persists in \Maple{ 2018}
as of the publication of this manuscript.

\section{Conclusion \& Future Work}\label{ch:conc-future-work}
The results and conclusions of the translations of the formula set from the \DLMF{} are diverse. Some test cases have shown us that the set of semantic macros is not perfect yet and needs to be improved, seen in the example of the macro definition for the Wronskian symbol. However, simply adding more and more macros to improve the set is not the right decision. In that case, we would end up in something similar to \sTeX, which is overloaded and too complicated to use. If the system is no longer handy, it will not be used.

On the other hand, the concept of the translator has been proven to discovering errors in the online compendia \gls*{dlmf}. The test cases have also shown how difficult it is to validate a translated expression and have uncovered the problems of translations between two systems with different sets of supported functions. Our validation techniques also assume the correctness of the simplification and computational algorithms in \gls*{cas}. However, combining those techniques and automatically running translation checks can potentially not only discover errors in mathematical compendia but also detect errors in simplifications or computations in \gls*{cas}.

The tasks for future work are diverse. The main task to improve the translator by implementing more functions and features. For example, for the current state, only translations to \Maple's standard function library were implemented. \Maple{} allows to load extra packages dynamically and therefore support several more functions. This feature would drastically increase the number of possible translations. With such improvements, further work on evaluation techniques become worthwhile to evaluate \gls*{dlmf} and \gls*{cas}.

Furthermore, the translator was designed to be easily extendable. This allows to implement translations for other \gls*{cas} without much effort.

The biggest weakness of the translator is still the semantic macros. Currently, we are working on mathematical information retrieval techniques to allow an extension for the translator for generic \LaTeX{} inputs.

\emergencystretch=1em
\printbibliography
\end{document}

### 2018-07-30 chapters/01-introduction/intro.tex
updates by H.Cohl

@@
 Figure~\ref{fig:acot-cut-compare} illustrates two examples of different branch cut positioning for the inverse trigonometric arccotangent function.
-While \Maple{} defines the branch cut at ${[-\iunit\infty, -\iunit]}$, ${[\iunit,\iunit\infty]}$ (Figure~\ref{fig:acot-cut1}), \Mathematica{} defines the branch cut at ${[-\iunit, \iunit]}$ (figure~\ref{fig:acot-cut2}).
+While \Maple{} defines the branch cut at ${[-\iunit\infty, -\iunit]}$, ${[\iunit,\iunit\infty]}$ (Figure~\ref{fig:acot-cut1}), \Mathematica{} defines the branch cut at ${[-\iunit, \iunit]}$ (Figure~\ref{fig:acot-cut2}).
 
@@
     }
-    \caption{Two plots for the real part of the arccotangent function with a branch cut at $[-\iunit\infty, -\iunit]$, $[\iunit,\iunit\infty]$ in figure~\protect\subref{fig:acot-cut1} and at $[-\iunit, \iunit]$ in figure~\protect\subref{fig:acot-cut2}, respectively.
+    \caption{Two plots for the real part of the arccotangent function with a branch cut at $[-\iunit\infty, -\iunit]$, $[\iunit,\iunit\infty]$ in Figure~\protect\subref{fig:acot-cut1} and at $[-\iunit, \iunit]$ in Figure~\protect\subref{fig:acot-cut2}, respectively.
 (Plotted with \Maple{} 2016)}

### 2018-07-30 main.tex
updates by H.Cohl

@@
         \label{fig:acot-cut2}    }
-    \caption{Two plots for the real part of the arccotangent function with a branch cut at $[-\iunit\infty, -\iunit]$, $[\iunit,\iunit\infty]$ in figure~\protect\subref{fig:acot-cut1} and at $[-\iunit, \iunit]$ in figure~\protect\subref{fig:acot-cut2}, respectively.
+    \caption{Two plots for the real part of the arccotangent function with a branch cut at $[-\iunit\infty, -\iunit]$, $[\iunit,\iunit\infty]$ in Figure~\protect\subref{fig:acot-cut1} and at $[-\iunit, \iunit]$ in Figure~\protect\subref{fig:acot-cut2}, respectively.
 (Plotted with \Maple{} 2016)}
@@
 	\cline{2-4}
-	\multicolumn{1}{l|}{} & Underscore & The underscore '\_' for subscripts. & The sequence $a$\_$b$ has two children ($a$ and '\_'). The underscore itself '\_' has one child ($b$).\\
+	\multicolumn{1}{l|}{} & Underscore & The underscore `\_' for subscripts. & The sequence $a$\_$b$ has two children ($a$ and `\_'). The underscore itself `\_' has one child ($b$).\\
 	\cline{2-4}
-	\multicolumn{1}{l|}{} & Caret & The caret '\^{}' to for superscripts or exponents.
+	\multicolumn{1}{l|}{} & Caret & The caret `\^{}' to for superscripts or exponents.
 Similar to the underscore.
-& The sequence $a$\^{}$b$ has two children ($a$ and '\^{}').
-The caret itself '\^{}' has one child ($b$).\\
+& The sequence $a$\^{}$b$ has two children ($a$ and `\^{}').
+The caret itself `\^{}' has one child ($b$).\\
 	\hline
@@
 	\cline{2-4}
-	\multicolumn{1}{l|}{} & Symbols & All kind of symbols & '$@$', '$*$', '$+$', '$!$', etc.\\
+	\multicolumn{1}{l|}{} & Symbols & All kind of symbols & `$@$', `$*$', `$+$', `$!$', etc.\\
 	\hline
@@
 \vspace{-5pt}
-\caption{The mathematical expression '$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
+\caption{The mathematical expression `$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
 \label{tab:notations}
@@
 \begin{itemize}
-\item an exponent, such as for '\verb|^2|';
+\item an exponent, such as for `\verb|^2|';
 \item an optional parameter in square brackets;
@@
 	\hline
-	SUM & Sums. Internally stored with factors for each summand, i.e., '$x+y$' would be stored as '$x \cdot 1 + y \cdot 1$'.\\
+	SUM & Sums. Internally stored with factors for each summand, i.e., `$x+y$' would be stored as `$x \cdot 1 + y \cdot 1$'.\\
 	PROD & Products.\\
@@
 For example, $\sqrt{x}$ is stored as $x^{\frac{1}{2}}$.
-\item There is no data type for subtractions, only for sums. Negative terms are changed to absolute values times '$-1$'. For example, $x-y$ is stored as $x + y \cdot (-1)$. 
+\item There is no data type for subtractions, only for sums. Negative terms are changed to absolute values times `$-1$'. For example, $x-y$ is stored as $x + y \cdot (-1)$. 
 \item Floating point numbers are stored in the scientific notation with a mantissa and an exponent in the base $10$. For example, $3.1$ is internally represented as $31 \cdot 10^{-1}$.

### 2018-07-30 main.tex
Howard edits

@@
 \hypersetup{
-	pdftitle	= {Semantic Preserving Bijective Mappings for Representations of Special Functions in Computer Algebra Systems and Word Processors - Exemplified using the DLMF/DRMF, the Word Processor LaTeX and the Computer Algebra Systems Maple and Mathematica},
+	pdftitle	= {Semantic Preserving Bijective Mappings for Representations of Special Functions in Computer Algebra Systems and Word Processors - Exemplified using the DLMF/DRMF, the Word Processor \LaTeX and the Computer Algebra Systems Maple and Mathematica},
 	pdfsubject={Master's Thesis in Mathematics, Technische Universität Berlin, July 2017.},
@@
 	pdfcreator	= {Andr\'e Greiner-Petter},
-	pdfkeywords	= {Special Functions, LaTeX, Translation},
+	pdfkeywords	= {Special Functions, \LaTeX, Translation},
 }
@@
 We have chosen \Maple{} and \Mathematica{} for the \gls*{cas}, and \LaTeX{} as our \gls*{wp}.
-Bruce Miller in the U.S.~has developed a set of semantic \LaTeX{} macros for \gls*{opsf}.
-The \gls*{nist} \gls*{dlmf} uses these semantic macros to provide a semantically enhanced mathematical online compendium.
+Bruce Miller at the \gls*{nist} has developed a set of semantic \LaTeX{} macros for \gls*{opsf} for the
+\gls*{dlmf}.
+The \gls*{dlmf} uses these semantic macros to provide a semantically enhanced mathematical online compendium.
 These semantic macros usally provide exclusive access to the semantic information of the functions.
@@
 \noindent
-{\it \bf Keywords:} LaTeX, Computer Algebra System (CAS), Translation, Presentation to Computation (P2C)
+{\it \bf Keywords:} \LaTeX, Computer Algebra System (CAS), Translation, Presentation to Computation (P2C)
 

### 2018-07-31 main.tex
Added new abstract

@@
 \noindent
-We present a translation tool between representations of semantically enriched mathematical formulae in a \gls*{wp} and its corresponding representations in a \gls*{cas}.
-We have chosen \Maple{} and \Mathematica{} for the \gls*{cas}, and \LaTeX{} as our \gls*{wp}.
-Bruce Miller at the \gls*{nist} has developed a set of semantic \LaTeX{} macros for \gls*{opsf} for the
-\gls*{dlmf}.
-The \gls*{dlmf} uses these semantic macros to provide a semantically enhanced mathematical online compendium.
-These semantic macros usally provide exclusive access to the semantic information of the functions.
-However, even if the semantics of a representation of one formula is unique, the semantics in another representation for the same formula may differ, such as in domains of variables.
-While some distinctions are obvious, such as syntactical characteristics, others are more difficult to examine, such as differences in definitions, which leads to error-prone manual translations.
-We discuss difficulties related to an automatic translation system, and suggest possible solutions. 
-Furthermore, this paper introduces new evaluation approaches for the developed translation tool. 
-With the help of our automatic translation tool, the evaluation experiments were able to discover errors in the \gls*{dlmf} and in Maple.
+\textbf{Purpose:} Modern mathematicians and scientist of math-related disciplines use \gls*{wp} to write and \gls*{cas} to calculate mathematical expressions.
+Usually, they translate the expressions manually between \gls*{wp} and \gls*{cas}.
+This process is time-consuming and error-prone.
+Our goal is to automate this translation.
+This paper uses \Maple{} and \Mathematica{} as the \gls*{cas}, and \LaTeX{} as our \gls*{wp}.
+
+\noindent\textbf{Approach:} The \gls*{nist} developed a collection of special \LaTeX{} macros that create links from mathematical symbols to their definitions in the \gls*{dlmf}.
+We are using these macros to perform rule-based translations between the formulae in the \gls*{dlmf} and \gls*{cas}.
+Moreover, we develop software to ease the creation of new rules and to discover inconsistencies.
+
+\noindent\textbf{Findings:} We created 396 mappings and translated 58.8\% of the \gls*{dlmf} (2,405 expressions) successfully between \Maple{} and \gls*{dlmf}.
+For a significant percentage, the definitions in Maple and the \gls*{dlmf} were different.
+Consequently, an atomic symbol in one system maps to a composite expression in the other system.
+The translator was also successfully used for automatic verification processes of mathematical online compendia and \gls*{cas}.
+Our evaluation techniques discovered two errors in the \gls*{dlmf} and one bug in \Maple.
+
+\noindent\textbf{Originality:} This paper introduces the first translation tool for special functions between \LaTeX{} and \gls*{cas}.
+The approach improves error-prone manual translations and can be used to verify mathematical online compendia and \gls*{cas}.
 \end{abstract} \glsresetall

### 2018-07-31 main.tex
Howard updates

@@
 \usepackage[english]{babel} \usepackage[T1]{fontenc} 
-\usepackage{hyperref}			\usepackage[acronym, nomain]{glossaries} \usepackage{amsthm}
+\usepackage{hyperref}			
+\usepackage[acronym, nomain]{glossaries} \usepackage{amsthm}
 \usepackage{amssymb}
 \usepackage{mathrsfs}
-\usepackage{parskip}		\usepackage{txfonts}            \usepackage{xcolor}				\usepackage{enumitem}			\usepackage{algorithm}
-\usepackage{algpseudocode} \usepackage{lipsum} \usepackage{tabularx}
+\usepackage{parskip}		
+\usepackage{txfonts}            
+\usepackage{xcolor}				
+\usepackage{enumitem}			
+\usepackage{algorithm}
+\usepackage{algpseudocode} 
+\usepackage{lipsum} 
+\usepackage{tabularx}
 \usepackage{stmaryrd}
@@
 \usepackage{layout}
-\usepackage{textcomp} \usepackage{authblk}
+\usepackage{textcomp} 
+\usepackage{authblk}
 
@@
 
+\DeclareLabeldate[online]{%
+  \field{date}
+  \field{year}
+  \field{eventdate}
+  \field{origdate}
+  \field{urldate}
+}
+
 \begin{document}
@@
 \noindent\textbf{Findings:} We created 396 mappings and translated 58.8\% of the \gls*{dlmf} (2,405 expressions) successfully between \Maple{} and \gls*{dlmf}.
-For a significant percentage, the definitions in Maple and the \gls*{dlmf} were different.
+For a significant percentage, the definitions in \Maple{} and the \gls*{dlmf} were different.
 Consequently, an atomic symbol in one system maps to a composite expression in the other system.
@@
 Figure~\ref{fig:acot-cut-compare} illustrates two examples of different branch cut positioning for the inverse trigonometric arccotangent function.
-While \Maple{} defines the branch cut at ${[-\iunit\infty, -\iunit]}$, ${[\iunit,\iunit\infty]}$ (Figure~\ref{fig:acot-cut1}), \Mathematica{} defines the branch cut at ${[-\iunit, \iunit]}$ (Figure~\ref{fig:acot-cut2}).
+While \Maple{}\footnote{The mention of
+specific products, trademarks, or brand
+names is for purposes of identification only. Such mention is not to be interpreted
+in any way as an endorsement or certification of such products or brands by the
+National Institute of Standards and Technology, nor does it imply that the products
+so identified are necessarily the best available for the purpose. All trademarks
+mentioned herein belong to their respective owners.} defines the branch cut at ${[-\iunit\infty, -\iunit]}$, ${[\iunit,\iunit\infty]}$ (Figure~\ref{fig:acot-cut1}), \Mathematica{} defines the branch cut at ${[-\iunit, \iunit]}$ (Figure~\ref{fig:acot-cut2}).
 
@@
 is not a \textit{direct} translation.
-
 Note that partial results of this paper have been published in~\parencite{CICM:Paper}.
@@
 Since \LaTeX{} became the de facto standard for writing papers in mathematics, most of the \gls*{cas} provide simple functions to import and export mathematical \LaTeX{} expressions\footnote{The selected \gls*{cas} Maple, Mathematica, Matlab, and SageMath provide import and/or export functions for \LaTeX:\\
-Maple, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;\\
-Mathematica, \url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html} seen 06/2017;\\
-Matlab, \url{https://www.mathworks.com/help/symbolic/latex.html} seen 06/2017;\\
-SageMath, \url{http://doc.sagemath.org/html/en/tutorial/latex.html} seen 06/2017}.
+Maple, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;
+Mathematica, \url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html} seen 06/2017;
+Matlab, \url{https://www.mathworks.com/help/symbolic/latex.html} seen 06/2017;
+SageMath, \url{http://doc.sagemath.org/html/en/tutorial/latex.html} seen 06/2017.}.
 Those tools have two essential problems.
@@
 
+\vspace{-0.2cm}
 Since translations such as these are nonintuitive, describing explanations become necessary for the translation process.
@@
 
+\vspace{-0.2cm}
 The position of the branch cut of the arccotangent function differs after translation~\eqref{eq:acot-alternatives}, which may lead to incorrect calculations later on.
@@
 		\hline
-		Semantic Macro Entry & Maple Entry \\
+		Semantic Macro Entry & \Maple{} Entry \\
 		\hline
@@
 
-\begin{wrapfigure}{r}{0.38\textwidth}
+\begin{wrapfigure}[4]{r}{0.38\textwidth}
 \vspace{-10pt}
@@
 This problem is solved by the \gls*{teo} that stores and groups translated objects like lists.
-This allows to access the latest translated expression and use it as the argument for the factorial function.
+This allows one to access the latest translated expression and use it as the argument for the factorial function.
 Table~\ref{tab:teo-list} shows three example for the \gls*{teo} list that groups some tokens.
@@
 \end{equation}
- \section{Maple to Semantic \LaTeX{} Translator}\label{sec:backward-translation}
+ \section{\Maple{} to Semantic \LaTeX{} Translator}\label{sec:backward-translation}
 Instead of writing a custom \Maple{} syntax parser, we using \Maple's internal data structure to get a syntax tree of the input\footnote{In consequence, a license of \Maple{} is mandatory to perform backward translations.
@@
 \end{equation}
-is incorrect if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ and one has to use analytic continuation for the right-hand side of equation (\ref{eq:branch-cut-near}).
-
+is incorrect if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ and one has to use analytic continuation for the right-hand side of (\ref{eq:branch-cut-near}). 
 To evaluate such complex cases, the equivalence checks of \gls*{cas} are insufficient. Therefore we implement numerical tests as an additional step.
@@
 \subsection{Results}\label{sec:test-summary}
-There are 685 \Macro s\footnote{The macros are still work in progress, and therefore the total number is constantly changing.} in total, and 665 of them were implemented in the translator engine.
+There are 685 \Macro{}s\footnote{The DLMF/DRMF semantic macros are still a work in progress, and the total number is constantly changing.} in total, and 665 of them were implemented in the translator engine.
 We defined forward translations to \Maple{} for 201 and backward translations from \Maple{} for 195 functions.
@@
 
-The numerical test engine was able to discover a sign error in equation~\parencite[(14.5.14)]{NIST:DLMF}\footnote{The equation had originally been stated as shown in equation~(\ref{eq:signerror}).
-The error were reported on 10th April 2017.}
+The numerical test engine was able to discover a sign error in equation~\parencite[(14.5.14)]{NIST:DLMF}\footnote{The equation had originally been stated as shown in (\ref{eq:signerror}).
+The error was reported on 10th April 2017.}
 \begin{equation}\label{eq:signerror}
@@
 The same engine also identified a missing comma in the constraint of~\parencite[(10.16.7)]{NIST:DLMF}.
-The original constraint was given by $2\nu \neq -1, -2 -3, \ldots$, with a missing comma after $-2$.
+The original constraint was given by $2\nu \neq -1, -2 -3, \ldots$, with a missing comma after the $-2$.
 
@@
 For example, for the current state, only translations to \Maple's standard function library were implemented.
-\Maple{} allows to load extra packages dynamically and therefore support several more functions.
+\Maple{} allows one to load extra packages dynamically and therefore support several more functions.
 This feature would drastically increase the number of possible translations.
@@
 
-Furthermore, the translator was designed to be easily extendable. This allows to implement translations for other \gls*{cas} without much effort.
-
-The biggest weakness of the translator is still the semantic macros.
+The translator was designed to be easily extendable. This allows one to implement translations for other \gls*{cas} without much effort.
+The biggest weakness of the translator is still the semantic macros {\bf [HSC: I don't understand what this sentence means.
+I think you need to be more specific here.
+The way it is written it doesn't make sense.]}.
 Currently, we are working on mathematical information retrieval techniques to allow an extension for the translator for generic \LaTeX{} inputs.

### 2018-07-31 main.tex
Howards updates

@@
 
-In general, the translator is drafted to solve ambiguous expressions or automatically find a work-around to disambiguate the expression.
+In general, the translator is drafted to translate ambiguous expressions or automatically find a work-around to disambiguate the expression {\bf [HSC: I don't undestand this sentence.
+drafted?
+Please rewrite]}.
 Only in case there is no way to solve the ambiguity with the defined rules, the translation process stops.
@@
 This delegation process is implemented in \Cref{line:transLeaf,line:transSeq} of Algorithm~\ref{alg:translation}.
-The Subsection~\ref{sec:subtranslators} discusses these classes in more detail.
+Subsection~\ref{sec:subtranslators} discusses these classes in more detail.
 
@@
 \end{table} 
-This approach seems to work for most of semantic \LaTeX{} expressions.
-However, there is one problem with this straight forward approach which was mentioned shortly mentioned in~\Cref{sec:problems}.
+This approach seems to work for most semantic \LaTeX{} expressions.
+However, there is one problem with this straightforward approach which was mentioned shortly mentioned in~\Cref{sec:problems}.
 There are many different types of notations used to represent formulae.
@@
 The \gls*{rpol}\footnote{Also known as \textit{postfix notation}} (hereafter called postfix notation) does the opposite and places the operator to the right of/after its operands.
-The infix notation is commonly used in arithmetic and places the operator between their operands, which only makes sense as long as the operator is a binary operator.
+The infix notation is commonly used in arithmetic and places the operator between their operands.
+This only makes sense as long as the operator is a binary operator.
 
@@
 In mathematical expressions, notations are mostly mixed, depending on the case and number of operands.
-For example, infix notation is common for binary operators ($+$, $-$, $\cdot$, $\mod$, etc.), while functional notations are conveniently used for any kind of functions ($sin$, $cos$, etc.).
+For example, infix notation is common for binary operators ($+$, $-$, $\cdot$, $\mod$, etc.), while functional notations are conveniently used for any kind of functions ($\sin$, $\cos$, etc.).
 Sometimes the same symbol is used in different notations to distinguish different meanings.
@@
 Therefore, it parses all four examples from~\Cref{tab:notations} to four different \gls*{pom-pt}s rather than to one unique expression tree.
-In general, this is not a problem for our translation process since most \gls*{cas} are familiar with the most common notations.
+In general, this is not a problem for our translation process since most \gls*{cas} are familiar with most common notations.
 Therefore, the translator does not need to know that `$a$' and `$b$' are the operands of the binary operator `$+$' in `$a+b$.' The translator could simply translate the symbols in `$a+b$' in the same order as they appear in the expression and the \gls*{cas} would understand it.
@@
 \begin{enumerate}\item \label{prob:1} The translated expression is only syntactically correct if the input expression was syntactically correct.
-\item \label{prob:2} We cannot translate expressions to a \gls*{cas} which uses a non-standard notation.
+\item \label{prob:2} We cannot translate expressions to \gls*{cas} which use non-standard notations.
 \end{enumerate}
@@
 
-Nonetheless, changing a notation could also solve ambiguities in some situations.
-Consider the two ambiguous examples in~\Cref{tab:amb_ex}.
+Nonetheless, adopting different notations, in some situations, could also solve potential ambiguities.
+Consider the two potentially ambiguous examples in~\Cref{tab:amb_ex}.
 While a scientist would probably just ask for the right interpretation of the first example, \Maple{} automatically computes the first interpretation.
-On the other hand, \LaTeX{} automatically disambiguate the first example by only recognizing the very next element (single symbols or sequence in curly brackets) for the superscript and therefore displays the second interpretation.
+On the other hand, \LaTeX{} automatically disambiguate the first example by only recognizing the very next element (single symbols or sequences in curly brackets) for the superscript and therefore displays the second interpretation.
 The second example is already interpreted as the double factorial function of $n$, since this notation is the standard interpretation in science.
@@
 \end{tabular}
-\caption{Ambiguous examples of the factorial and double factorial function. One expression in a text format can be interpreted in different ways.}
+\caption{Potentially ambiguous examples of the factorial and double factorial function. One expression in a text format can be interpreted in different ways.}
 \label{tab:amb_ex}
@@
 In most cases, parentheses can be used to disambiguate expressions.
-We used them in~\Cref{tab:amb_ex} to clarify the different interpretations in example 2.
-But sometimes, even parentheses cannot solve a mistaken computation.
+We used them in~\Cref{tab:amb_ex} to clarify the different interpretations in Example 2.
+Note that the use of parentheses will not always resolve a mistaken computation.
 For example, there is no way to add parentheses to force \Maple{} to compute `$n!!$' as the double factorial function.

### 2018-07-31 main.tex
Howards updates

@@
 A \verb|SequenceTranslator| translates the \textit{sequence} and \textit{balanced expressions} in the \gls*{pom-pt}.
-If a node $n$ is a leaf and the represented symbol an open bracket (parentheses, square brackets and so on) the following nodes are also taken as a \textit{sequence}.
-Hence, combined with the recursive translation approach, the \verb|SequenceTranslator| also checks balances of parentheses in expressions.
+If a node $n$ is a leaf and the represented symbol is an open bracket (parentheses, square brackets and so on) the following nodes are also taken as a \textit{sequence}.
+Hence, combined with the recursive translation approach, the \verb|SequenceTranslator| also checks balancing of parentheses in expressions.
 An expression such as `$(a]$' is producing a mismatched parentheses error.
@@
 Whether an expression is an interval or another expression is difficult to decide and can depend on the context.
-Also, the parenthesis checker could simply be deactivated to allow mismatched parentheses in an expression.
+Also, the parentheses checker could simply be deactivated to allow mismatched parentheses in an expression.
 But this functionality is not implemented yet.
-
-The \verb|SequenceTranslator| also handle positions of multiplication symbols.
-There are a couple of obvious choices to translate multiplication signs.
-The most common symbol for multiplications is still the white space or none symbol at all, as explained previously.
+{\bf [HSC: This can be solved by the use of interval macros.
+e.g.,
+\verb|\intcc@{a}{b} = [a,b]|, \verb|\intoc@{a}{b}=(a,b]|, \verb|\intco@{a}{b}=[a,b)|, \verb|\intoo@{a}{b}=(a,b)|,
+with a double \verb|@@| notation you could do other notations such as \verb|\intoo@@{a}{b}=]a,b[|]}
+
+The \verb|SequenceTranslator| also handles positions of multiplication symbols.
+There are a couple of obvious choices to translate multiplication.
+The most common symbol for multiplications is still the white space or none symbol {\bf [HSC: What is a none symbol?]} at all, as explained previously.
 Consider the simple expression `$2n\pi$'.
@@
 Algorithm~\ref{alg:macro-translation} explains the \verb|MacroTranslator| without error handling.
-It extracted necessary information from the \gls*{pom-pt}, such as how many arguments this function has.
+It has extracted necessary information from the \gls*{pom-pt}, such as how many arguments this function has.
 In consequence, it also processes the following siblings to translate the arguments.
@@
 \item a parameter in curly brackets (a \textit{sequence} node in the \gls*{pom-pt}) if none of the above;
-\item an $@$ symbols if none of the above;
-\item or a variable in curly brackets (a \textit{sequence} node) if none of the above.
+\item `$@$' symbols if none of the above {\bf [HSC: it was originally written ``an @ symbols'' which doesn't make
+sense, please rewrite if wrong.]}; or
+\item a variable in curly brackets (a \textit{sequence} node) if none of the above.
 \end{itemize}
 
-An exponent will be translated and shifted to the end of the translated semantic macro, since it is more common to write the exponent after the arguments of a function in \gls*{cas}.
-Therefore, the function translates and stores the exponent in Line~\ref{line:next_caret}.
+An exponent will be translated and shifted to the end of the translated semantic macro, 
+since one usually writes the exponent after the arguments of a function in \gls*{cas}.
+Hence, the function translates and stores the exponent in Line~\ref{line:next_caret}.
 One could ask what happens when there is an exponent given before and after the arguments.
@@
 The translator is using the version \Maple{} 2016.}.
-\Maple{} allow different input styles.
+\Maple{} allows several different input styles.
 The \texttt{1D} input is mainly used for programming purposes and therefore also used to perform our translations.
-Internally, \Maple{} using \gls*{dag} as syntax trees.
+Internally, \Maple{} uses a \gls*{dag} for syntax trees.
 
@@
 One can access the internal data structure of expressions via the \texttt{ToInert} command.
-The returned \inertF{} is a nested list representation as a tree equivalent\footnote{A tree equivalent of a \gls*{dag} splitting nodes with multiple parents into multiple nodes so that each node has only one parent node.} of the internal \gls*{dag} for the given expression.
+The returned \inertF{} is a nested list representation as a tree equivalent {\bf [HSC: I don't understand the previous sentence.
+Please rewrite.]}\footnote{A tree equivalent of a \gls*{dag} splitting nodes with multiple parents into multiple nodes so that each node has only one parent node.} of the internal \gls*{dag} for the given expression.
 Some of the important types for the nodes are specified in \Cref{tab:maple-types}.
-
 \begin{table}[h]
@@
 \end{table}
-
-The translator using the \texttt{OpenMaple}~\parencite[\S 14.3]{MAPLE:ProgrammingGuide} \gls*{api} for interacting with \Maple's kernel and get access to the \inertF{} of inputs.
+The translator uses the \texttt{OpenMaple}~\parencite[\S 14.3]{MAPLE:ProgrammingGuide} \gls*{api} for interacting with \Maple's kernel and get access {\bf [HSC: and get access??
+please rewrite.]}{\bf [HSC: and get access??
+please rewrite.]}  to the \inertF{} of inputs.
 
@@
 \item There is no data type for subtractions, only for sums. Negative terms are changed to absolute values times `$-1$'. For example, $x-y$ is stored as $x + y \cdot (-1)$. 
-\item Floating point numbers are stored in the scientific notation with a mantissa and an exponent in the base $10$. For example, $3.1$ is internally represented as $31 \cdot 10^{-1}$.
-\item There is only a data type for rational numbers (fractions with integer numerator and positive denominator), but not for general fractions, such as $\frac{x+y}{z}$.
+\item Floating point numbers are stored using scientific notation with a mantissa and an exponent in the base $10$. For example, $3.1$ is internally represented as $31 \cdot 10^{-1}$.
+\item There is only a data type for rational numbers (fractions with an integer numerator and a positive integer denominator), but not for general fractions, such as $\frac{x+y}{z}$.
 This will be automatically changed to $(x+y)\cdot z^{-1}$.
@@
 \item We use unevaluation quotes to suppress evaluations of the input.
-\item We perform reordering of factors and summands so that negative factors appear in front of the summand.
+\item We perform a reordering of factors and summands so that negative factors appear in front of the summand.
 This gives us the opportunity to translate $x - y$ to $x - y$ instead of $x + y \cdot (-1)$.
-\item We introduced the new internal data types \texttt{MYFLOAT} and \texttt{DIVIDE} to translate floats and fractions in more convenient notations.
+\item We introduced new internal data types \texttt{MYFLOAT} and \texttt{DIVIDE} to translate floats and fractions in more convenient notations.
 \end{itemize}
@@
 	\includegraphics[clip, trim=0.1cm 0.1cm 0.1cm 0.1cm, scale=0.7]{MapleTranslation.pdf}
-	\caption{A scheme of the backward translation process from \Maple{} for the Jacobi polynomial $P_n^{(\alpha , \beta)}(\cos(a\Theta))$.
+	\caption{A scheme of the backward translation process from \Maple{} for the Jacobi polynomial expression $P_n^{(\alpha , \beta)}(\cos(a\Theta))$.
 The input string is converted by the \Maple{} kernel into the nested list representation.
 This list is translated by subtranslators (blue and red arrows).
-A function translation (bold blue arrows) is again realized by translation patterns to define the position of the arguments (red arrows).}
+A function translation (bold blue arrows) is again realized using translation patterns to define the position of the arguments (red arrows).}
 	\label{fig:backward-trans}
@@
 \item \textbf{Round Trip Tests}: translates expressions back and forth and analyzing the changes.
-\item\label{approach-2} \textbf{Function Relation Tests (Symbolical)}: translate mathematically proven equivalent expressions from one system to a \gls*{cas} and evaluate whether the relation remains valid via symbolical equivalence checks.
+\item\label{approach-2} \textbf{Function Relation Tests (Symbolical)}: translates mathematically proven equivalent expressions from one system to a \gls*{cas} and evaluate whether the relation remains valid via symbolical equivalence checks.
 \item \textbf{Numerical Tests}: take the same equations from Approach~\ref{approach-2} but evaluates them on specific numerical values to test equivalence.
@@
 A round trip test always starts with a valid expression either in semantic \LaTeX{} or in \Maple.
-A translation from one system to another is called \textbf{a step}.
+A translation from one system to another is called a \textbf{step}.
 A complete round trip translation (two steps) is called \textbf{one cycle}.
@@
 
-Step 4 is identical to step 2, and since the translator is a deterministic algorithm it can be easily shown that step 2 and step 3 are fixed-point representations for semantic \LaTeX{} and \Maple.
+Step 4 is identical to step 2, and since the translator is a deterministic algorithm, it can be easily shown that step 2 and step 3 are fixed-point representations for semantic \LaTeX{} and \Maple.
 
@@
 \verb|\EllIntF@{\phi}{k}| & \overset{\langMaple}{\mapsto} & \verb|EllipticF(sin(phi),k)|,\\
-\verb|\EllIntF@{\asin@{\phi}}{k}| & \overset{\langMaple}{\mapsfrom} & \verb|EllipticF(phi,k)|,
+\verb|\EllIntF@{\asin@{\phi}}{k}| & \overset{\langMaple}{\mapsfrom} & \verb|EllipticF(phi,k)|.
 \end{eqnarray}
-
 The round-trip translations produce infinite chains of sine and inverse sine calls because there are no evaluations involved.
+
 The round trip tests are very successful, but they only detect errors in string representations.
@@
 \end{eqnarray}
-In that case the round trip test would not detect any errors and reach a fixed point representation, because the simplification techniques only simplify two representations in the same system but cannot compare the representation in one system to those in the other.
+In that case the round trip test would not detect any errors and reach a fixed point representation, because the simplification techniques only simplify two representations in the same system but cannot compare the representation in one system to those in the other {\bf [HSC: This sentence is too long and rambles on and is not easy to read Can you break it into two sentences and rewrite them so that it is easier to read?]}.
 
 \subsection{Function Relation Tests}\label{sec:relation-tests}
-The \gls*{dlmf} is a compendium of special functions and orthogonal polynomials and lists several relations between the functions and polynomials.
+The \gls*{dlmf} is a compendium for special functions and orthogonal polynomials and lists several relations between the functions and polynomials.
 The idea of this evaluation approach is to translate an entire relation and test whether the relation remains valid after performing the translations.
@@
 However, there are two essential problems with this approach.
-Testing the mathematical equivalence of expressions is hard to solve and \gls*{cas} have trouble to test even simple equations.
+Testing the mathematical equivalence of expressions is hard to solve and \gls*{cas} often have difficulties testing simple equations {\bf [HSC: Can you give an example of this?]}.
 Furthermore, this approach only checks forward translations because there is no way to check equivalence of expressions in \LaTeX{} automatically (again this could become feasible with our translator).
-We use \Maple's \textit{simplify} function to check if the difference of the left-hand side and the right-hand side of the equation is equal to zero.
-In addition, we use \textit{simplify} and check if the division of the right-hand side by the left-hand side returns a numerical value or not.
+We use \Maple's \texttt{simplify} function to check if the difference of the left-hand side and the right-hand side of the equation is equal to zero.
+In addition, we use \texttt{simplify} and check if the division of the right-hand side by the left-hand side returns a numerical value or not.
 This simplification function is the most powerful function to check the equivalence in \Maple.
@@
 \end{equation}
-to an equivalent representation using the exponential representations
+to an equivalent representation using their exponential representations, namely
 \begin{equation}
@@
 \end{equation}
-With such pre-conversions we are able to improve the simplification process in \Maple.
-However, the limitations of the \textit{simplify} function are still the weakest part of this verification approach.
+With such pre-conversions, we are able to improve the simplification process in \Maple.
+However, the limitations of the \texttt{simplify} function are still the weakest part of this verification approach.
 Consider the complex example~\parencite[(12.7.10)]{NIST:DLMF}
@@
 \end{equation}
-where $U(0,z)$ is the parabolic cylinder function and $K_\nu(z)$ the modified Bessel function of the second kind.
+where $U(0,z)$ is the parabolic cylinder function and $K_\nu(z)$ is the modified Bessel function of the second kind.
 Both functions are well-defined in both systems and we can define a \textit{direct} translation for~\eqref{eq:branch-cut-near}.
 The modified Bessel function of the second kind has its branch cut in \Maple{} and in the \gls*{dlmf} at $z < 0$.
-However, the argument of $K$ contains $z^2$.
+However, the argument of $K$ contains a $z^2$.
 If $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ the value of the right-hand side of~\eqref{eq:branch-cut-near} would be no longer on the principal branch.
@@
 \end{equation}
-is incorrect if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ and one has to use analytic continuation for the right-hand side of (\ref{eq:branch-cut-near}). 
+is incorrect if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ and one should instead use the analytic continuation for the right-hand side of (\ref{eq:branch-cut-near}). 
 To evaluate such complex cases, the equivalence checks of \gls*{cas} are insufficient. Therefore we implement numerical tests as an additional step.
@@
 
-Considering machine accuracy and the default precision of $10$ significant digits, we can regard the first and last values as zero differences.
+Considering machine accuracy and the default precision to $10$ significant digits, we can regard the first and last values as zero differences.
 While this evaluation is very powerful, it has a significant problem.
-Even when all tested values return zero, it does not proof the equivalence of~\eqref{eq:branch-cut-near}.
-However, when the values are different from zero, it does proof an error in one of the four cases~\parencite{NumericalTests:Paper}
+Even when all tested values return zero, it does not prove the equivalence of~\eqref{eq:branch-cut-near}.
+However, when the values are different from zero, it does indicate that there might be an error satisfying one of the four cases~\parencite{NumericalTests:Paper}:
 \begin{enumerate}
@@
 There are 685 \Macro{}s\footnote{The DLMF/DRMF semantic macros are still a work in progress, and the total number is constantly changing.} in total, and 665 of them were implemented in the translator engine.
-We defined forward translations to \Maple{} for 201 and backward translations from \Maple{} for 195 functions.
+We defined forward translations to \Maple{} for 201 of the macros and backward translations from \Maple{} for 195 functions.
 
 The \gls*{dlmf} provides a dataset of \LaTeX{} expressions with semantic macros.
-We extracted 4.087 equations from \gls*{dlmf} and apply our round-trip and relation tests on them.
-The translator was able to translate 2.405\footnote{All percentages are approximately calculated.} (58.8\%) of the extracted equations without errors.
+We extracted 4087 equations from the \gls*{dlmf} and apply our round-trip and relation tests on them.
+The translator was able to translate 2405\footnote{All percentages are approximately calculated.} (58.8\%) of the extracted equations without errors.
 660 (27.4\%) of the successfully translated expressions were verified by the simplification techniques of \Maple.
-We apply additional numerical tests for the remaining 1.745 equations.
+We apply additional numerical tests for the remaining 1745 equations.
 For 418 (24\%) cases the numerical tests were valid.
-More detailed results for numerical and symbolical tests were presented in~\parencite{NumericalTests:Paper}.
-
-The evaluation techniques have proven to be very powerful also for evaluating \gls*{cas} and online mathematical compendia such as the \gls*{dlmf}.
-During the evaluations, we were able to detect several errors in the translation and evaluation engine, and also discovered two errors in the \gls*{dlmf} and one error in \Maple's \texttt{simplify}-function.
+More detailed results for numerical and symbolical tests were presented in~\parencite{NumericalTests:Paper} {\bf [HSC: Why is this rendered as (ibid.)?
+Remove when fixed.]}.
+
+The evaluation techniques has proven to be very powerful for evaluating \gls*{cas} and online mathematical compendia such as the \gls*{dlmf}.
+During the evaluations, we were able to detect several errors in the translation and evaluation engine, and also discovered two errors in the \gls*{dlmf} and one error in \Maple's \texttt{simplify} function.
 
@@
 We have also noticed that our testing procedure is able to identify errors in CAS procedures,
-namely the \Maple{} {\tt simplify} procedure. The left-hand side of \cite[(7.18.4)]{NIST:DLMF} is given by 
+namely the \Maple{} {\tt simplify} procedure. The left-hand side of~\parencite[(7.18.4)]{NIST:DLMF} 
+{\bf [HSC: Why is this rendered as ibid? Remove when fixed.]} is given by 
 \begin{equation*}

### 2018-07-31 main.tex
Howards updates

@@
 The results and conclusions of the translations of the formula set from the \DLMF{} are diverse.
-Some test cases have shown us that the set of semantic macros is not perfect yet and needs to be improved, seen in the example of the macro definition for the Wronskian symbol.
+Some test cases have shown us that the set of semantic macros are not perfect yet and need to be improved.
+This is seen in the example of the macro definition for the Wronskian determinant.
 However, simply adding more and more macros to improve the set is not the right decision.
@@
 
-On the other hand, the concept of the translator has been proven to discovering errors in the online compendia \gls*{dlmf}.
+On the other hand, the translator concept has proven itself by discovering errors in the online \gls*{dlmf} compendia.
 The test cases have also shown how difficult it is to validate a translated expression and have uncovered the problems of translations between two systems with different sets of supported functions.
 Our validation techniques also assume the correctness of the simplification and computational algorithms in \gls*{cas}.
-However, combining those techniques and automatically running translation checks can potentially not only discover errors in mathematical compendia but also detect errors in simplifications or computations in \gls*{cas}.
+However, combining those techniques and automatically running translation checks, not only can discover errors in mathematical compendia but can also detect errors in simplifications or computations of the \gls*{cas}.
 
 The tasks for future work are diverse.
-The main task to improve the translator by implementing more functions and features.
+The main task is to improve the translator by implementing more functions and features.
 For example, for the current state, only translations to \Maple's standard function library were implemented.
-\Maple{} allows one to load extra packages dynamically and therefore support several more functions.
+\Maple{} allows one to load extra packages dynamically and therefore support an enhanced set of functions.
 This feature would drastically increase the number of possible translations.
-With such improvements, further work on evaluation techniques become worthwhile to evaluate \gls*{dlmf} and \gls*{cas}.
+With such improvements, further work on evaluation techniques become worthwhile to evaluate the \gls*{dlmf} and \gls*{cas}.
 
@@
 The way it is written it doesn't make sense.]}.
-Currently, we are working on mathematical information retrieval techniques to allow an extension for the translator for generic \LaTeX{} inputs.
+Currently, we are working on mathematical information retrieval techniques which will allow for an extension of the translator to generic \LaTeX{} inputs.
 

### 2018-07-31 main.tex
Howards updates

@@
 In that case, we would end up in something similar to \sTeX, which is overloaded and too complicated to use.
-If the system is no longer handy, it will not be used.
+If the system is no longer handy, it will not be used {\bf [HSC: I don't agree with this paragraph.
+In fact, I think we've had this discussion already in the past and it was resolved.
+It needs to be resolved again.
+There are many examples where enhancing macros is a benefit.
+Not only that, we've discussed this recently in our publications.
+This needs to be fixed prior to submission.]}
 

### 2018-07-31 main.tex
Howards updates

@@
 \noindent
-\textbf{Purpose:} Modern mathematicians and scientist of math-related disciplines use \gls*{wp} to write and \gls*{cas} to calculate mathematical expressions.
+\textbf{Purpose:} Modern mathematicians and scientists of math-related disciplines use \gls*{wp} to write and \gls*{cas} to calculate mathematical expressions.
 Usually, they translate the expressions manually between \gls*{wp} and \gls*{cas}.
@@
 
-\noindent\textbf{Approach:} The \gls*{nist} developed a collection of special \LaTeX{} macros that create links from mathematical symbols to their definitions in the \gls*{dlmf}.
+\noindent\textbf{Approach:} Bruce Miller at the \gls*{nist} developed a collection of semantic \LaTeX{} macros that create links from mathematical symbols to their definitions in the \gls*{nist} \gls*{dlmf}.
 We are using these macros to perform rule-based translations between the formulae in the \gls*{dlmf} and \gls*{cas}.
@@
 Consequently, an atomic symbol in one system maps to a composite expression in the other system.
-The translator was also successfully used for automatic verification processes of mathematical online compendia and \gls*{cas}.
-Our evaluation techniques discovered two errors in the \gls*{dlmf} and one bug in \Maple.
+The translator was also successfully used for automatic verification of mathematical online compendia and \gls*{cas}.
+Our evaluation techniques discovered two errors in the \gls*{dlmf} and one defect in \Maple.
 
@@
 
+{\bf [HSC: The link in the NIST:DLMF reference is broken (see below), and I don't know how to fix it.
+By the way, I made some changes the the way that it is referred to in the text, because I did not like that Olver et al was being referred to, since it is a web site.
+However, it is still not being displayed properly.
+For instance, the Version number is not being displayed and also the link is broken.
+Can someone try to fix this?
+I do not know Biber, so this is probably the reason for this mess.]}
+
 \emergencystretch=1em

### 2018-07-31 main.tex
Howards updates

@@
 Especially in the \gls*{stem} literature, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
-However, since \LaTeX{} (an extension of \TeX) is the de facto standard for writing articles in \gls*{stem}, and we only focus on mathematical writing in this paper, we have categorize \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the past 30 years~\parencites{Knuth}[559]{DigitalTypo}{LATEX:Standard}.
+However, since \LaTeX{} (an extension of \TeX) is the de facto standard for writing articles in \gls*{stem}, and we only focus on mathematical writing in this paper, we have categorized \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the past 30 years~\parencites{Knuth}[559]{DigitalTypo}{LATEX:Standard}.
 \LaTeX{} enables printing of mathematical formulae in a structure similar to handwritten style.
-For example, consider the specific Jacobi polynomial~\parencite[18.3 in table 1]{NIST:DLMF}
+For example, consider the specific Jacobi polynomial~\parencite[Table 18.3.1]{NIST:DLMF}
 \begin{equation}\label{eq:P}
@@
 \end{equation}
-where $\alpha, \beta > -1$, and $n$ is a nonnegative integer. 
+where $n$ is a nonnegative integer, $\alpha, \beta > -1$, and $a,\Theta\in{\mathbf R}$.
 This mathematical expression is written in \LaTeX{} as
@@
 is not a \textit{direct} translation.
-Note that partial results of this paper have been published in~\parencite{CICM:Paper}.
+Note that partial results of this paper has been published in~\parencite{CICM:Paper}.
 

### 2018-07-31 main.tex
Howards updates

@@
 In that case, we would end up in something similar to \sTeX, which is overloaded and too complicated to use.
-If the system is no longer handy, it will not be used {\bf [HSC: I don't agree with this paragraph.
+If the system is no longer handy, it will not be used {\bf [HSC: I don't agree with this paragraph at all.
 In fact, I think we've had this discussion already in the past and it was resolved.
@@
 Not only that, we've discussed this recently in our publications.
-This needs to be fixed prior to submission.]}
+This needs to be fixed prior to submission.
+I vote for removing this paragraph altogether until we can resolve this issue.]}
 

### 2018-08-01 main.tex
Implemented HSC comments

--- 2018-07-31 main.tex
@@
 	backend		= biber,
-	style		= authoryear-ibid,
-	giveninits	= true, 	uniquename	= init, 	natbib		= true,
+	style		= authoryear,
+	giveninits	= true, 	
+	uniquename	= init, 	
+	natbib		= true,
 	url			= false,
@@
 
-In general, the translator is drafted to translate ambiguous expressions or automatically find a work-around to disambiguate the expression {\bf [HSC: I don't undestand this sentence.
-drafted?
-Please rewrite]}.
-Only in case there is no way to solve the ambiguity with the defined rules, the translation process stops.
+In general, the translator is designed to find a work-around for disambiguating expressions. Only in case there is no way to solve the ambiguity with the defined rules, the translation process stops.
 
@@
 Also, the parentheses checker could simply be deactivated to allow mismatched parentheses in an expression.
-But this functionality is not implemented yet.
-{\bf [HSC: This can be solved by the use of interval macros.
-e.g.,
-\verb|\intcc@{a}{b} = [a,b]|, \verb|\intoc@{a}{b}=(a,b]|, \verb|\intco@{a}{b}=[a,b)|, \verb|\intoo@{a}{b}=(a,b)|,
-with a double \verb|@@| notation you could do other notations such as \verb|\intoo@@{a}{b}=]a,b[|]}
+Another option is to use of interval macros.
+e.g., \verb|\intcc@{a}{b} = [a,b]|.
 
@@
 There are a couple of obvious choices to translate multiplication.
-The most common symbol for multiplications is still the white space or none symbol {\bf [HSC: What is a none symbol?]} at all, as explained previously.
+The most common symbol for multiplications is still the white space (or no space between the tokens), as explained previously.
 Consider the simple expression `$2n\pi$'.
@@
 \item a parameter in curly brackets (a \textit{sequence} node in the \gls*{pom-pt}) if none of the above;
-\item `$@$' symbols if none of the above {\bf [HSC: it was originally written ``an @ symbols'' which doesn't make
-sense, please rewrite if wrong.]}; or
+\item `$@$' symbols if none of the above; or
 \item a variable in curly brackets (a \textit{sequence} node) if none of the above.
@@
 
-One can access the internal data structure of expressions via the \texttt{ToInert} command.
-The returned \inertF{} is a nested list representation as a tree equivalent {\bf [HSC: I don't understand the previous sentence.
-Please rewrite.]}\footnote{A tree equivalent of a \gls*{dag} splitting nodes with multiple parents into multiple nodes so that each node has only one parent node.} of the internal \gls*{dag} for the given expression.
+One can access the internal data structure of expressions via the \texttt{ToInert} command, which returns the \inertF.
+The \inertF{} format is a nested list\footnote{The nested list is a tree representation of a \gls*{dag} that splits nodes with multiple parents into multiple nodes so that each node has only one parent node.} of the internal \gls*{dag} for the given expression.
 Some of the important types for the nodes are specified in \Cref{tab:maple-types}.
@@
 \end{table}
-The translator uses the \texttt{OpenMaple}~\parencite[\S 14.3]{MAPLE:ProgrammingGuide} \gls*{api} for interacting with \Maple's kernel and get access {\bf [HSC: and get access??
-please rewrite.]}{\bf [HSC: and get access??
-please rewrite.]}  to the \inertF{} of inputs.
+The translator uses the \texttt{OpenMaple}~\parencite[\S 14.3]{MAPLE:ProgrammingGuide} \gls*{api} for interacting with \Maple's kernel implementation.
 
@@
 \end{eqnarray}
-In that case the round trip test would not detect any errors and reach a fixed point representation, because the simplification techniques only simplify two representations in the same system but cannot compare the representation in one system to those in the other {\bf [HSC: This sentence is too long and rambles on and is not easy to read Can you break it into two sentences and rewrite them so that it is easier to read?]}.
+In that case the round trip test would not detect any errors but reaches a fixed point representation.
 
@@
 However, there are two essential problems with this approach.
-Testing the mathematical equivalence of expressions is hard to solve and \gls*{cas} often have difficulties testing simple equations {\bf [HSC: Can you give an example of this?]}.
+Testing the mathematical equivalence of expressions is hard to solve and \gls*{cas} often have difficulties testing simple equations symbolically.
+For example, equation~\parencite[(4.35.34)]{NIST:DLMF}
+\begin{equation*}
+\sinh \left( x+\iunit \cdot y \right) = \sinh x \cos y + \iunit \cosh x \sin y,
+\end{equation*}
+as a difference of the left- and right-hand side cannot be simplified to zero by default.
 Furthermore, this approach only checks forward translations because there is no way to check equivalence of expressions in \LaTeX{} automatically (again this could become feasible with our translator).
@@
 For 418 (24\%) cases the numerical tests were valid.
-More detailed results for numerical and symbolical tests were presented in~\parencite{NumericalTests:Paper} {\bf [HSC: Why is this rendered as (ibid.)?
-Remove when fixed.]}.
+More detailed results for numerical and symbolical tests were presented in~\parencite{NumericalTests:Paper}.
 
@@
 We have also noticed that our testing procedure is able to identify errors in CAS procedures,
-namely the \Maple{} {\tt simplify} procedure. The left-hand side of~\parencite[(7.18.4)]{NIST:DLMF} 
-{\bf [HSC: Why is this rendered as ibid? Remove when fixed.]} is given by 
+namely the \Maple{} {\tt simplify} procedure. The left-hand side of~\parencite[(7.18.4)]{NIST:DLMF} is given by 
 \begin{equation*}
@@
 \section{Conclusion \& Future Work}\label{ch:conc-future-work}
-The results and conclusions of the translations of the formula set from the \DLMF{} are diverse.
+%The results and conclusions of the translations of the formula set from the \DLMF{} are diverse.
 Some test cases have shown us that the set of semantic macros are not perfect yet and need to be improved.
@@
 Not only that, we've discussed this recently in our publications.
-This needs to be fixed prior to submission.]}
-
-On the other hand, the translator concept has proven itself by discovering errors in the online \gls*{dlmf} compendia.
+This needs to be fixed prior to submission.
+I vote for removing this paragraph altogether until we can resolve this issue.]} {\bf [AGP: Since the paper in it's current state is quite long and the section works well without this paragraph, I just commented it out and avoid the discussion here...]}
+
+The translator concept has proven itself by discovering errors in the online \gls*{dlmf} compendia.
 The test cases have also shown how difficult it is to validate a translated expression and have uncovered the problems of translations between two systems with different sets of supported functions.
@@
 The translator was designed to be easily extendable. This allows one to implement translations for other \gls*{cas} without much effort.
-The biggest weakness of the translator is still the semantic macros {\bf [HSC: I don't understand what this sentence means.
-I think you need to be more specific here.
-The way it is written it doesn't make sense.]}.
+An extensive weakness is the dependency on the special macros from the \gls*{dlmf}.
+The translator is not able to translate functions without using these macros.
 Currently, we are working on mathematical information retrieval techniques which will allow for an extension of the translator to generic \LaTeX{} inputs.
-
-{\bf [HSC: The link in the NIST:DLMF reference is broken (see below), and I don't know how to fix it.
-By the way, I made some changes the the way that it is referred to in the text, because I did not like that Olver et al was being referred to, since it is a web site.
-However, it is still not being displayed properly.
-For instance, the Version number is not being displayed and also the link is broken.
-Can someone try to fix this?
-I do not know Biber, so this is probably the reason for this mess.]}
 

### 2018-08-01 main.tex
Clean version, ready to publish with anonymous authors

@@
 
-\usepackage{graphicx}           \usepackage{subfig}             \usepackage{wrapfig}			\graphicspath{ {./images/} }
+\usepackage{graphicx}
+\usepackage{subfig} 
+\usepackage{wrapfig}	
+\graphicspath{ {./images/} }
 
 \usepackage[export]{adjustbox}
-
-\usepackage{styles/thesistitle}
 
@@
 \hypersetup{
-	pdftitle	= {Semantic Preserving Bijective Mappings for Representations of Special Functions in Computer Algebra Systems and Word Processors - Exemplified using the DLMF/DRMF, the Word Processor \LaTeX and the Computer Algebra Systems Maple and Mathematica},
-	pdfsubject={Master's Thesis in Mathematics, Technische Universität Berlin, July 2017.},
-	pdfauthor	= {Andr\'e Greiner-Petter},
-	pdfcreator	= {Andr\'e Greiner-Petter},
-	pdfkeywords	= {Special Functions, \LaTeX, Translation},
+	pdftitle	= {Semantic Preserving Bijective Mappings for Representations of Special Functions in Computer Algebra Systems and Word Processors}
 }
@@
 
-\renewcommand{\algorithmicrequire}{\textbf{Input:}} \renewcommand{\algorithmicensure}{\textbf{Output:}} \newcommand{\NULL}{\textbf{null}}
+\renewcommand{\algorithmicrequire}{\textbf{Input:}} \renewcommand{\algorithmicensure}{\textbf{Output:}}
+\newcommand{\NULL}{\textbf{null}}
 \renewcommand{\Return}{\textbf{return}}
@@
 
-%
-
+%%%%%%%%%%%%%%%%%%%%%%%%% INFORMATION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 \title{
@@
 }
-\author{Andr\'e Greiner-Petter\textsuperscript{1}}
-\author{Moritz Schubotz\textsuperscript{1}}
-\author{Howard S.~Cohl\textsuperscript{2}}
-\author{Bela Gipp\textsuperscript{1}}
-
-\affil{
-	\textsuperscript{1}Information Science Group, University of Konstanz, Germany\\
-	\url{{first.last}@uni-konstanz.de}
-}
-
-\affil{
-	\textsuperscript{2}Applied and Computational Mathematics Division,\\
-	National Institute of Standards and Technology, Mission Viejo, CA, USA,\\
-	\url{howard.cohl@nist.gov}
-}
+
+%\input{authors.tex} %TODO anonymous authors for double blinded review
+\author{Anonymous Authors}
 
@@
 \noindent
-{\it \bf Keywords:} \LaTeX, Computer Algebra System (CAS), Translation, Presentation to Computation (P2C)
+{\it \bf Keywords:} \LaTeX, Computer Algebra System (CAS), Translation, Presentation to Computation (P2C), Special Functions
 
@@
 
-\begin{wrapfigure}[4]{r}{0.38\textwidth}
-\vspace{-10pt}
+\begin{wrapfigure}[10]{r}{0.38\textwidth} % optional parameter gives "# of narrow lines", very useful! 
+\vspace{-8pt}
 \begin{minipage}{0.38\textwidth}
@@
 \end{tabular}
-\vspace{-5pt}
+%\vspace{-5pt}
 \caption{The mathematical expression `$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
@@
 \end{minipage}
-\vspace{-5pt}
+%\vspace{-5pt}
 \end{wrapfigure}
@@
 
-\begin{wrapfigure}{r}{0.61\textwidth}
-	\vspace{-22pt}
+\begin{wrapfigure}[7]{r}{0.61\textwidth}
+	\vspace{-15pt}
 	\centering
@@
 	\label{fig:internal-maple-dag}
-	\vspace{-15pt}
+	%\vspace{-15pt}
 \end{wrapfigure}
@@
 Some of the important types for the nodes are specified in \Cref{tab:maple-types}.
-\begin{table}[h]
+\begin{table}[ht]
 \centering

### 2018-08-01 del.sh

@@ -1,18 +1,0 @@
-#!/bin/bash
-
-rm -f Masterthesis.acn
-rm -f Masterthesis.acr
-rm -f Masterthesis.alg
-rm -f Masterthesis.aux
-rm -f Masterthesis.bbl
-rm -f Masterthesis.bcf
-rm -f Masterthesis.blg
-rm -f Masterthesis.ist
-rm -f Masterthesis.loa
-rm -f Masterthesis.lof
-rm -f Masterthesis.lot
-rm -f Masterthesis.log
-rm -f Masterthesis.toc
-rm -f Masterthesis.out
-rm -f Masterthesis.run.xml
-rm -f Masterthesis.synctex.gz

### 2018-08-01 main.tex
Reset Howard's changes in the abstract-approach after discussions with Moritz.

@@
 
-\noindent\textbf{Approach:} Bruce Miller at the \gls*{nist} developed a collection of semantic \LaTeX{} macros that create links from mathematical symbols to their definitions in the \gls*{nist} \gls*{dlmf}.
+\noindent\textbf{Approach:} The \gls*{nist} developed a collection of special \LaTeX{} macros that create links from mathematical symbols to their definitions in the \gls*{nist} \gls*{dlmf}.
 We are using these macros to perform rule-based translations between the formulae in the \gls*{dlmf} and \gls*{cas}.

### 2018-08-01 main.tex
Final version

@@
     \centering
-    \subfloat[The real part of arccotangent with a branch cut at ${[-\iunit\infty, -\iunit]}$, ${[\iunit,\iunit\infty]}$.]{        \includegraphics[width=0.45\textwidth]{acotCut1.png}
-        \label{fig:acot-cut1}    }
+    \subfloat[The real part of arccotangent with a branch cut at ${[-\iunit\infty, -\iunit]}$, ${[\iunit,\iunit\infty]}$.]{
+    		\includegraphics[width=0.45\textwidth]{acotCut1.jpg}
+        \label{fig:acot-cut1}
+    }
     \hspace{0.5cm}
-    \subfloat[The real part of arccotangent with a branch cut at ${[-\iunit, \iunit]}$.]{        \includegraphics[width=0.45\textwidth]{acotCut2.png}
-        \label{fig:acot-cut2}    }
+    \subfloat[The real part of arccotangent with a branch cut at ${[-\iunit, \iunit]}$.]{
+    		\includegraphics[width=0.45\textwidth]{acotCut2.jpg}
+        \label{fig:acot-cut2}
+    }
     \caption{Two plots for the real part of the arccotangent function with a branch cut at $[-\iunit\infty, -\iunit]$, $[\iunit,\iunit\infty]$ in Figure~\protect\subref{fig:acot-cut1} and at $[-\iunit, \iunit]$ in Figure~\protect\subref{fig:acot-cut2}, respectively.

### 2018-08-01 main.tex
Howards updates

@@
 \subsection{Numerical Tests}\label{sec:numerical-tests}
-Consider the differences of the left- and right-hand side of equation~\eqref{eq:branch-cut-near}
+Consider the difference of the left- and right-hand sides of equation~\eqref{eq:branch-cut-near}, namely
 \begin{equation}\label{eq:difference}
@@
 \end{equation}
-Table~\ref{tab:computations-for-difference} presents four computations for $D(z)$, one value for each quadrant in the complex plane.
+Table~\ref{tab:computations-for-difference} presents four numerical evaluations for $D(z)$, one value for each quadrant in the complex plane.
 \begin{table}[ht]
@@
 \end{tabular}
-\caption{Four computations of $D(z)$ in \Maple.}
+\caption{Four numerical evaluations of $D(z)$ in \Maple.}
 \label{tab:computations-for-difference}

### 2018-10-02 main.tex
Implemented WERB Review Corrections by Marjorie McClain

@@
 \noindent
-\textbf{Purpose:} Modern mathematicians and scientists of math-related disciplines use \gls*{wp} to write and \gls*{cas} to calculate mathematical expressions.
+\textbf{Purpose:} Modern mathematicians and scientists of math-related disciplines use \glspl*{wp} to write and \glspl*{cas} to calculate mathematical expressions.
 Usually, they translate the expressions manually between \gls*{wp} and \gls*{cas}.
@@
 \section{Introduction}
-A typical workflow of a scientist who writes a scientific publication is to use a \gls*{wp} to write the paper and one or more \gls*{cas} for verification, analysis and visualization.
+A typical workflow of a scientist who writes a scientific publication is to use a \gls*{wp} to write the paper and one or more \glspl*{cas} for verification, analysis and visualization.
 Especially in the \gls*{stem} literature, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
@@
 These macros embed necessary semantic information into \LaTeX{} expressions.
-One example of such a macro is given Table~\ref{tab:JacobiP-usecase} for the semantic \LaTeX{} representation for the Jacobi polynomial.
+One example of such a macro is given in Table~\ref{tab:JacobiP-usecase} for the semantic \LaTeX{} representation for the Jacobi polynomial.
 The macros provide isolated access to important parts of the mathematical function, such as the arguments.
@@
 Even with embedded semantic information, a translation between systems can be difficult.
-A typical example of complex problems occur for multivalued functions~\parencite{AISC:MultivaluedFunctions}.
+A typical example of complex problems occurs for multivalued functions~\parencite{AISC:MultivaluedFunctions}.
 A \gls*{cas} usually defines \textit{branch cuts} to compute principal values of multivalued functions~\parencite{Maple:Cuts}, which makes the implementation of a theoretically continued function to a discontinuous presentation of it.
@@
 In consequence, a manual translation process is not only laborious, but also prone to errors.
-Note that this general problem has been named to automatic \gls*{p2c} conversion~\parencite{POM-Tagger}.
+Note that this general problem has been named as automatic \gls*{p2c} conversion~\parencite{POM-Tagger}.
 
@@
 Consider the irrational mathematical constant $e$, known as Euler's number.
-The theoretical construct for this symbol cannot be mathematically equivalent to the value \verb|exp(0)| in \Maple, caused by computational and implementational limitations.
-Instead of using the term \textit{equivalent}, we introduce a \textit{appropriate} and \textit{inappropriate} translations.
+The theoretical construct for this symbol cannot be mathematically equivalent to the value \verb|exp(1)| in \Maple, caused by computational and implementational limitations.
+Instead of using the term \textit{equivalent}, we introduce \textit{appropriate} and \textit{inappropriate} translations.
 We call a translation such as
@@
 \end{equation}
-as \textit{appropriate}, while a translation such as
+\textit{appropriate}, while a translation such as
 \begin{equation}
@@
 \end{equation}
-as \textit{inappropriate}.
+is \textit{inappropriate}.
 Note that it is not always as easy as in this example to decide if a translation is appropriate or not.
@@
 is not a \textit{direct} translation.
-Note that partial results of this paper has been published in~\parencite{CICM:Paper}.
+Note that partial results of this paper have been published in~\parencite{CICM:Paper}.
 
@@
 There are two main approaches to solve that problem: (1) someone could specify the semantic information during the writing process (pre-defined semantics); and (2) the translator can find out the right semantic information in general mathematical expressions before it translates the expression.
-So-called \textit{interactive documents}\footnote{There is no adequate definition what interactive documents are.
+So-called \textit{interactive documents}\footnote{There is no adequate definition for what interactive documents are.
 However, this name is widely used to describe electronic document formats that allow for interactivity to change the content in real time.}, such as the \gls*{cdf}\footnote{Wolfram Research; \textit{Computable Document Format} (CDF); \url{http://www.wolfram.com/cdf/}, July 2011} by Wolfram Research, or \textit{worksheets} by \Maple{}, try to solve this problem with the approach (2) and allow one to embed semantic information into the input.
@@
 These semantic macros are internally used in the \gls*{dlmf} and the \gls*{drmf}.
-We gave the \Macro{} set the favor for developing the translation engine because it provides \gls*{dlmf} definitions for a comprehensive amount of functions.
+We gave the \Macro{} set the favor for developing the translation engine because it provides \gls*{dlmf} definitions for a comprehensive number of functions.
 In contrast, \sTeX{} does not focus on the semantics of functions, is too complex to use, and defines diverse macros for symbols and concepts that \gls*{cas} usually does not support.
@@
 Since translations such as these are nonintuitive, describing explanations become necessary for the translation process.
-A special logging function takes care of each translation and provide details after a successful translation process.
+A special logging function takes care of each translation and provides details after a successful translation process.
 \Cref{sec:forward-translation} explains this task further.
@@
 The arctangent function~\eqref{eq:acot-alternatives-1} and the natural logarithm~(\ref{eq:acot-alternatives-2}) have the same positioned branch cuts as in the \gls*{dlmf} and in \Maple.
-In consequence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user do not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z = -\iunit$.
+In consequence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user does not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z = -\iunit$.
 
@@
 \end{equation}
-A translation become unfeasible because the macro does not explicitely define the variable of differentiation for the functions $w_1$ and $w_2$.
+A translation becomes unfeasible because the macro does not explicitely define the variable of differentiation for the functions $w_1$ and $w_2$.
 For a correct translation, the \gls*{cas} needs to be aware of the used variable $z$.
@@
 
-The \Macro s do not garantee entirely disambiguated expressions.
+The \Macro s do not guarantee entirely disambiguated expressions.
 In~\Cref{tab:amb-latex} there are four examples of potentially ambiguous expressions.
@@
 
-Another more questionable translation decision are alphanumerical expressions.
+Another more questionable translation decision is alphanumerical expressions.
 As explained in~\Cref{tab:allTypesTable}, the \gls*{pom}-tagger handles strings of letters and numbers differently, depending on the order of the symbols.
-The reason is, that an expression such as `$4b$' is usually considered to be a multiplication of $4$ and `$b$,' while `$b4$' looks like indexing `$b$' by $4$.
+The reason is that an expression such as `$4b$' is usually considered to be a multiplication of $4$ and `$b$,' while `$b4$' looks like indexing `$b$' by $4$.
 While the first example produces two nodes, namely $4$ and `$b$', the second example `$b4$' produces just a single alphanumerical node in the \gls*{pom-pt}.
 The translator interprets alphanumerical expressions as multiplications for two reasons: (1) we would assume that the inputs `$4b$' and `$b4$' are mathematically equivalent; and (2) it is more common in mathematics to use single letter names for variables~\parencite{Notation:History}.
-Therefore we define the following definitions
+Therefore we make the following definitions
 
@@
 
-In general, the translator is designed to find a work-around for disambiguating expressions. Only in case there is no way to solve the ambiguity with the defined rules, the translation process stops.
+In general, the translator is designed to find a work-around for disambiguating expressions. Only if there is no way to solve the ambiguity with the defined rules, the translation process stops.
 
@@
 All translations are defined by a library (\gls*{csv} and \gls*{json} files) that defines translation patterns for each function and symbol.
-The pattern uses \verb|$i| as placeholders to define the positions of the arguments.
+The pattern uses \verb|$i| as a placeholder to define the positions of the arguments.
 For example, the translation patterns for the Jacobi polynomial are illustrated in~\Cref{tab:placeholder_ex2}.
@@
 
-These placeholders causes trouble when the \gls*{cas} uses the symbol \verb|$| for other reasons, e.g., the differentiation in \Maple{} is defined by
+These placeholders cause trouble when the \gls*{cas} uses the symbol \verb|$| for other reasons, e.g., the differentiation in \Maple{} is defined by
 \begin{equation*}
@@
 The \Macro s also allow one to specify optional arguments to distinguish between standard and another version of these functions.
-The Legendre and associated Legendre function of the first kind are examples of such cases.
+The Legendre and associated Legendre functions of the first kind are examples of such cases.
 The library that defines translations for each macro uses the macro name as the primary key to identify the translations.
@@
 In consequence, the \gls*{pom}-tagger generates a raw parsed tree where each token in the \LaTeX{} expression is a node in the tree.
-We call this parsed tree, the \gls*{pom-pt}.
+We call this parsed tree the \gls*{pom-pt}.
 
@@
 	\cline{2-4}
-	\multicolumn{1}{l|}{} & Caret & The caret `\^{}' to for superscripts or exponents.
+	\multicolumn{1}{l|}{} & Caret & The caret `\^{}' for superscripts or exponents.
 Similar to the underscore.
@@
 This approach seems to work for most semantic \LaTeX{} expressions.
-However, there is one problem with this straightforward approach which was mentioned shortly mentioned in~\Cref{sec:problems}.
+However, there is one problem with this straightforward approach which was mentioned in~\Cref{sec:problems}.
 There are many different types of notations used to represent formulae.
@@
 The \gls*{rpol}\footnote{Also known as \textit{postfix notation}} (hereafter called postfix notation) does the opposite and places the operator to the right of/after its operands.
-The infix notation is commonly used in arithmetic and places the operator between their operands.
+The infix notation is commonly used in arithmetic and places the operator between its operands.
 This only makes sense as long as the operator is a binary operator.
@@
 While a scientist would probably just ask for the right interpretation of the first example, \Maple{} automatically computes the first interpretation.
-On the other hand, \LaTeX{} automatically disambiguate the first example by only recognizing the very next element (single symbols or sequences in curly brackets) for the superscript and therefore displays the second interpretation.
+On the other hand, \LaTeX{} automatically disambiguates the first example by only recognizing the very next element (single symbols or sequences in curly brackets) for the superscript and therefore displays the second interpretation.
 The second example is already interpreted as the double factorial function of $n$, since this notation is the standard interpretation in science.
@@
 This allows one to access the latest translated expression and use it as the argument for the factorial function.
-Table~\ref{tab:teo-list} shows three example for the \gls*{teo} list that groups some tokens.
+Table~\ref{tab:teo-list} shows three examples for the \gls*{teo} list that groups some tokens.
 
@@
 Hence, combined with the recursive translation approach, the \verb|SequenceTranslator| also checks balancing of parentheses in expressions.
-An expression such as `$(a]$' is producing a mismatched parentheses error.
+An expression such as `$(a]$' produces a mismatched parentheses error.
 On the other hand, this is a problem for interval expressions such as `$[a,b)$'.
@@
 Also, the parentheses checker could simply be deactivated to allow mismatched parentheses in an expression.
-Another option is to use of interval macros.
+Another option is to use interval macros.
 e.g., \verb|\intcc@{a}{b} = [a,b]|.
@@
 Consider an expression such as `$f(x)$'.
-Translate this to \Maple{} will be \verb|f*(x)|.
+Translating this to \Maple{} will give \verb|f*(x)|.
 But we do not consider this translation to be wrong, because there is a semantic macro to represent functions.
@@
  \section{\Maple{} to Semantic \LaTeX{} Translator}\label{sec:backward-translation}
-Instead of writing a custom \Maple{} syntax parser, we using \Maple's internal data structure to get a syntax tree of the input\footnote{In consequence, a license of \Maple{} is mandatory to perform backward translations.
+Instead of writing a custom \Maple{} syntax parser, we use \Maple's internal data structure to get a syntax tree of the input\footnote{In consequence, a license of \Maple{} is mandatory to perform backward translations.
 The translator is using the version \Maple{} 2016.}.
@@
 \begin{enumerate}
-\item \textbf{Round Trip Tests}: translates expressions back and forth and analyzing the changes.
-\item\label{approach-2} \textbf{Function Relation Tests (Symbolical)}: translates mathematically proven equivalent expressions from one system to a \gls*{cas} and evaluate whether the relation remains valid via symbolical equivalence checks.
-\item \textbf{Numerical Tests}: take the same equations from Approach~\ref{approach-2} but evaluates them on specific numerical values to test equivalence.
+\item \textbf{Round Trip Tests}: translates expressions back and forth and analyzes the changes.
+\item\label{approach-2} \textbf{Function Relation Tests (Symbolical)}: translates mathematically proven equivalent expressions from one system to a \gls*{cas} and evaluates whether the relation remains valid via symbolical equivalence checks.
+\item \textbf{Numerical Tests}: takes the same equations from Approach~\ref{approach-2} but evaluates them on specific numerical values to test equivalence.
 \end{enumerate}
@@
 On the other hand, these tests cannot determine logical errors in the translations between the two systems.
-Consider we mistakenly defined an \textit{inappropriate} forward and backward translation for the sine function
+Suppose we mistakenly defined an \textit{inappropriate} forward and backward translation for the sine function
 \begin{eqnarray}
@@
 However, there are several cases where the simplification fails.
-Because of implementation details, there are some techniques that helps \Maple{} to find possible simplifications.
+Because of implementation details, there are some techniques that help \Maple{} to find possible simplifications.
 For example we can force \Maple{} to convert the formula
@@
 The \gls*{dlmf} provides a dataset of \LaTeX{} expressions with semantic macros.
-We extracted 4087 equations from the \gls*{dlmf} and apply our round-trip and relation tests on them.
+We extracted 4087 equations from the \gls*{dlmf} and applied our round-trip and relation tests on them.
 The translator was able to translate 2405\footnote{All percentages are approximately calculated.} (58.8\%) of the extracted equations without errors.
 660 (27.4\%) of the successfully translated expressions were verified by the simplification techniques of \Maple.
-We apply additional numerical tests for the remaining 1745 equations.
+We applied additional numerical tests for the remaining 1745 equations.
 For 418 (24\%) cases the numerical tests were valid.
@@
 
-The evaluation techniques has proven to be very powerful for evaluating \gls*{cas} and online mathematical compendia such as the \gls*{dlmf}.
+The evaluation techniques have proven to be very powerful for evaluating \gls*{cas} and online mathematical compendia such as the \gls*{dlmf}.
 During the evaluations, we were able to detect several errors in the translation and evaluation engine, and also discovered two errors in the \gls*{dlmf} and one error in \Maple's \texttt{simplify} function.

### 2018-10-04 main.tex
Change Word Processor to Document Preparation System Also deleted the footnote that technically LaTeX is not a WP, since it is no longer needed now.

@@
 \hypersetup{
-	pdftitle	= {Semantic Preserving Bijective Mappings for Representations of Special Functions in Computer Algebra Systems and Word Processors}
+	pdftitle	= {Semantic Preserving Bijective Mappings for Representations of Special Functions in Computer Algebra Systems and Document Preparation Systems}
 }
@@
 \newacronym{opsf}{OPSF}{Orthogonal Polynomials and Special Functions}
-\newacronym{wp}{WP}{Word Processor}
+\newacronym{dps}{DPS}{Document Preparation System}
 \newacronym{stem}{STEM}{Science, Technology, Engineering and Mathematics}
@@
 \title{
-    Semantic Preserving Bijective Mappings for Representations of Special Functions between Computer Algebra Systems and Word Processors
+    Semantic Preserving Bijective Mappings for Representations of Special Functions between Computer Algebra Systems and Document Preparation Systems
 }
@@
 \noindent
-\textbf{Purpose:} Modern mathematicians and scientists of math-related disciplines use \glspl*{wp} to write and \glspl*{cas} to calculate mathematical expressions.
-Usually, they translate the expressions manually between \gls*{wp} and \gls*{cas}.
+\textbf{Purpose:} Modern mathematicians and scientists of math-related disciplines use \glspl*{dps} to write and \glspl*{cas} to calculate mathematical expressions.
+Usually, they translate the expressions manually between \gls*{dps} and \gls*{cas}.
 This process is time-consuming and error-prone.
 Our goal is to automate this translation.
-This paper uses \Maple{} and \Mathematica{} as the \gls*{cas}, and \LaTeX{} as our \gls*{wp}.
+This paper uses \Maple{} and \Mathematica{} as the \gls*{cas}, and \LaTeX{} as our \gls*{dps}.
 
@@
 \section{Introduction}
-A typical workflow of a scientist who writes a scientific publication is to use a \gls*{wp} to write the paper and one or more \glspl*{cas} for verification, analysis and visualization.
-Especially in the \gls*{stem} literature, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
-However, since \LaTeX{} (an extension of \TeX) is the de facto standard for writing articles in \gls*{stem}, and we only focus on mathematical writing in this paper, we have categorized \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the past 30 years~\parencites{Knuth}[559]{DigitalTypo}{LATEX:Standard}.
+A typical workflow of a scientist who writes a scientific publication is to use a \gls*{dps} to write the paper and one or more \glspl*{cas} for verification, analysis and visualization.
+Especially in the \gls*{stem} literature, \LaTeX{}
+%\footnote{Note that technically \LaTeX{} is not a \gls*{dps} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
+However, since \LaTeX{} (an extension of \TeX) is the de facto standard for writing articles in \gls*{stem}, and we only focus on mathematical writing in this paper, we have categorized \LaTeX{} as a \gls*{dps} as well.}
+has become the de facto standard for writing scientific publications over the past 30 years~\parencites{Knuth}[559]{DigitalTypo}{LATEX:Standard}.
 \LaTeX{} enables printing of mathematical formulae in a structure similar to handwritten style.

### 2018-11-15 main.tex
Add authors and affiliations (close #4)
Add subsections to sections 3 (close #5)

@@
 
-%\input{authors.tex} %TODO anonymous authors for double blinded review
-\author{Anonymous Authors}
+\input{authors.tex}
 
@@
 
+\subsection{Different Sets of Defined Functions}
 If a function is defined in one system but not in the other, we can translate the definition of the mathematical function.
@@
 
+\subsection{Positions of Branch Cuts}
 In case of differences between defined branch cuts we can also use alternative translations to solve the problems.
@@
 
+\subsection{Insufficient Semantic Information}
 Other problematic cases for translations are the \Macro s themselves.
@@
 
+\subsection{Ambiguous Expressions}
 A similar problem is multiplications since they are rarely explicitly marked in \LaTeX{} expressions, e.g., scientists using whitespaces to indicate multiplications rather than using \verb|\cdot| or similar symbols.

### 2018-11-15 main.tex
Give a short introduction to the concept of the translation in sec 4 (close #7)

@@
 \section{The Translator}
+The translator analyzes a parse tree to perform translations.
+For generating a parse tree of \LaTeX{} expressions, the translator uses the \gls*{pom}-Tagger~\parencite{POM-Tagger}\footnote{Named according to the Part-of-Speech-Taggers in \gls*{nlp}.}.
+The \gls*{pom}-Tagger is an \texttt{LL}-Parser defined by a context-free grammar in \gls*{bnf}.
+Each token will be tagged by meta information defined in manually cultivated lexicon files.
+\gls*{cas} define their own syntax parser.
+We were able to use \Maple's internal data structure to get a parse tree of the input.
+
 All translations are defined by a library (\gls*{csv} and \gls*{json} files) that defines translation patterns for each function and symbol.
@@
 
-These placeholders cause trouble when the \gls*{cas} uses the symbol \verb|$| for other reasons, e.g., the differentiation in \Maple{} is defined by
-\begin{equation*}
-\verb|diff(f, [x$n])|,
-\end{equation*}
-where $f$ is an algebraic expression or an equation, $x$ is the name of the differentiation variable, and $n$ is an integer representing the $n$-th order differentiation\footnote{\url{https://www.maplesoft.com/support/help/maple/view.aspx?path=diff}, seen 07/2018}.
-A translation for $\frac{d^2x^2}{dx^2}$ should look like this
-\begin{equation*}
-\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [x$2])|
-\end{equation*}
-but would end up as
-\begin{equation*}
-\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [xx])|.
-\end{equation*}
-
-We can solve this issue by using parentheses in such cases, e.g., \verb|diff($1, [$2$($0)])|.
-
 The \Macro s also allow one to specify optional arguments to distinguish between standard and another version of these functions.
@@
 The Legendre and associated Legendre function of the first kind both use the same macro \verb|\LegendreP|.
-To distinguish such cases, we use a special syntax for such cases, shown in~\Cref{tab:legendreP-lex}.
+To distinguish such cases, we use a special syntax, shown in~\Cref{tab:legendreP-lex}.
 
@@
 
-The translator uses the \gls*{pom}-Tagger~\parencite{POM-Tagger}\footnote{Named according to the Part-of-Speech-Taggers in \gls*{nlp}.} to parse \LaTeX{} expressions into a parsed tree.
-The \gls*{pom}-Tagger is an \texttt{LL}-Parser defined by a context-free grammar in \gls*{bnf}.
-Each token will be tagged by meta information defined in lexicon files.
-We extend the lexicon files to provide also the information that is necessary for the translation process.
-An example of an entry of the lexicon file is given in~\Cref{tab:sine-lex-example}.
+We integrated the defined translation patterns in the library also into the lexicon files of the \gls*{pom}-Tagger.
+Thus the tagger also tags a token by its translation patterns to other languages.
+\Cref{tab:sine-lex-example} gives an example of an entry of the lexicon file.
 
@@
 	\label{tab:sine-lex-example}
-\end{table} \section{Forward Translations}\label{sec:forward-translation}
+\end{table}
+
+\subsection{Escape the Placeholder Symbol}
+The used placeholders cause trouble when the \gls*{cas} uses the symbol \verb|$| for other reasons, e.g., the differentiation in \Maple{} is defined by
+\begin{equation*}
+\verb|diff(f, [x$n])|,
+\end{equation*}
+where $f$ is an algebraic expression or an equation, $x$ is the name of the differentiation variable, and $n$ is an integer representing the $n$-th order differentiation\footnote{\url{https://www.maplesoft.com/support/help/maple/view.aspx?path=diff}, seen 07/2018}.
+A translation for $\frac{d^2x^2}{dx^2}$ should look like this
+\begin{equation*}
+\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [x$2])|
+\end{equation*}
+but would end up as
+\begin{equation*}
+\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [xx])|.
+\end{equation*}
+
+We can solve this issue by using parentheses in such cases, e.g., \verb|diff($1, [$2$($0)])|.
+
+\section{Forward Translations}\label{sec:forward-translation}
 An abstract translator class analyzes each node of the parsed tree and delegates them to specialized subtranslators.

### 2018-11-15 main.tex
Better explanation for the PoM-Tagger (close #6)

@@
 For generating a parse tree of \LaTeX{} expressions, the translator uses the \gls*{pom}-Tagger~\parencite{POM-Tagger}\footnote{Named according to the Part-of-Speech-Taggers in \gls*{nlp}.}.
-The \gls*{pom}-Tagger is an \texttt{LL}-Parser defined by a context-free grammar in \gls*{bnf}.
-Each token will be tagged by meta information defined in manually cultivated lexicon files.
 \gls*{cas} define their own syntax parser.
 We were able to use \Maple's internal data structure to get a parse tree of the input.
+\Cref{sec:forward-translation} and \Cref{sec:backward-translation} will explain the parsing and translation process in detail.
 
@@
 
-We integrated the defined translation patterns in the library also into the lexicon files of the \gls*{pom}-Tagger.
-Thus the tagger also tags a token by its translation patterns to other languages.
+\subsection{Escape the Placeholder Symbol}
+The used placeholders cause trouble when the \gls*{cas} uses the symbol \verb|$| for other reasons, e.g., the differentiation in \Maple{} is defined by
+\begin{equation*}
+\verb|diff(f, [x$n])|,
+\end{equation*}
+where $f$ is an algebraic expression or an equation, $x$ is the name of the differentiation variable, and $n$ is an integer representing the $n$-th order differentiation\footnote{\url{https://www.maplesoft.com/support/help/maple/view.aspx?path=diff}, seen 07/2018}.
+A translation for $\frac{d^2x^2}{dx^2}$ should look like this
+\begin{equation*}
+\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [x$2])|
+\end{equation*}
+but would end up as
+\begin{equation*}
+\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [xx])|.
+\end{equation*}
+
+We can solve this issue by using parentheses in such cases, e.g., \verb|diff($1, [$2$($0)])|.
+
+\section{Forward Translations}\label{sec:forward-translation}
+As a pre-processing step, we use the \gls*{pom}-Tagger~\parencite{POM-Tagger}\footnote{Named according to the Part-of-Speech-Taggers in \gls*{nlp}.} for parsing semantic \LaTeX{} expressions.
+The \gls*{pom}-Tagger is defined by a context-free grammar in \gls*{bnf} and is an \texttt{LL}-Parser, i.e., it parses the input from \textbf{L}eft to right and assigns the \textbf{L}eftmost (first applicable) drivation rule defined by the grammar to an expression.
+In other words, the \gls*{pom}-Tagger scans the input for \textit{terms} and groups them into subexpressions if suitable, where \textit{terms} are non-terminal symbols in the context of \gls*{bnf}.
+If a node in the generated parse tree matches a pre-defined symbol, it will be tagged by mata information that are defined in manually cultivated lexicon files.
+
+We integrated the defined translation patterns from our library also into these lexicon files.
+Thus the tagger also tags a node in the parse tree by its translation patterns.
 \Cref{tab:sine-lex-example} gives an example of an entry of the lexicon file.
@@
 
-\subsection{Escape the Placeholder Symbol}
-The used placeholders cause trouble when the \gls*{cas} uses the symbol \verb|$| for other reasons, e.g., the differentiation in \Maple{} is defined by
-\begin{equation*}
-\verb|diff(f, [x$n])|,
-\end{equation*}
-where $f$ is an algebraic expression or an equation, $x$ is the name of the differentiation variable, and $n$ is an integer representing the $n$-th order differentiation\footnote{\url{https://www.maplesoft.com/support/help/maple/view.aspx?path=diff}, seen 07/2018}.
-A translation for $\frac{d^2x^2}{dx^2}$ should look like this
-\begin{equation*}
-\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [x$2])|
-\end{equation*}
-but would end up as
-\begin{equation*}
-\verb|\deriv[2]{x^2}{x}| \overset{\langMaple}{\mapsto} \verb|diff(x^2, [xx])|.
-\end{equation*}
-
-We can solve this issue by using parentheses in such cases, e.g., \verb|diff($1, [$2$($0)])|.
-
-\section{Forward Translations}\label{sec:forward-translation}
-An abstract translator class analyzes each node of the parsed tree and delegates them to specialized subtranslators.
-In this process, an object called \gls*{teo} is built.
-This \gls*{teo} is needed to rebuild a string representation of the mathematical expression after all nodes of the parsed tree were translated.
-
-The parsed tree generated by the \gls*{pom}-tagger is not a mathematical expression tree.
+The parsed tree generated by the \gls*{pom}-Tagger is not a mathematical expression tree.
 The \gls*{pom} project aims to disambiguate mathematical \LaTeX{} expressions and generates an expression tree.
@@
 Therefore we create the \gls*{teo}.
+The \gls*{teo} is a list containing already translated subexpressions.
 
@@
 \end{tabularx}
-\caption{A table of all kinds of nodes in a PoM syntax tree. Note that this table groups some kinds for a better overview. For a complete list and a more detailed version see~\parencite{POM-Tagger}.}
+\caption{A table of all kinds of nodes in a PoM syntax tree.
+Note that this table groups some types together for a better overview.
+For a complete list and a more detailed version see~\parencite{POM-Tagger}.}
 \label{tab:allTypesTable}
@@
 \end{equation}
+
 \section{\Maple{} to Semantic \LaTeX{} Translator}\label{sec:backward-translation}

### 2018-11-15 main.tex
Cut out algorithm 1 and change the explanation style a bit (close #11)

@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}
-\vspace{-20pt}
-\begin{minipage}{0.5\textwidth}
-\begin{algorithm}[H]
-\caption{Simple translation algorithm for the POM-Parse Tree}\label{alg:simple-translation}
-	\begin{algorithmic}[1]
-	\Require Root $r$ of the parse tree $T$
-	\Procedure{translate}{$r$}
-	\If{$r$ is leaf}
-		\State {\scriptsize TRANSLATE\_LEAF}($r$);
-	\Else
-		\ForAll{children $v_n$ of $r$}
-			\State {\scriptsize TRANSLATE}($v_n$);
-		\EndFor
-	\EndIf
-	\EndProcedure
-	\end{algorithmic}
-\end{algorithm}
-\end{minipage}
-\vspace{-14pt}
-\end{wrapfigure}
-
 \Cref{fig:syntax-tree-usecase} visualizes the \gls*{pom-pt} of the Jacobi polynomial example from~\Cref{tab:JacobiP-usecase}.
@@
 The algorithm is realized in a simple recursive structure.
-The Algorithm~\ref{alg:simple-translation} explains the translation process.
 Whenever the algorithm finds a leaf, it can translate this single term.
 If the node is not a leaf, it starts to translate all children of the node recursively.
+This idea seems to be a practical and elegant solution but it has a significant drawback, it cannot be used to translate functions.
+Since the arguments of functions are following siblings in the \gls*{pom-pt}, the algorithm needs to look ahead when a leaf is a known function, e.g., in the case of a semantic macro with arguments (see~\Cref{fig:syntax-tree-usecase}).
+Algorithm~\ref{alg:translation} is an improved version with look ahead functionality.
 
 \begin{algorithm}[t]
-\caption{Abstract translation algorithm to translate MLP-Parse trees.}\label{alg:translation}
+\caption{Abstract translation algorithm to translate PPT.}\label{alg:translation}
 	\begin{algorithmic}[1]
-	\Require Root $r$ of a POM-Parse tree $T$. List \textit{following\_siblings} with the following siblings of $r$. The list can be empty.
+	\Require Root $r$ of a PoM-Parse tree $T$. List \textit{following\_siblings} with the following siblings of $r$. The list can be empty.
 	\Procedure{abstract\_translator}{$r$, \textit{following\_siblings}}
@@
 	\Else
-		\State \textit{siblings} $ = r$.getChildren(); \Comment{\textit{siblings} is a list of children}
-		\State {\scriptsize ABSTRACT\_TRANSLATOR}(\textit{siblings}.removeFirst(), \textit{siblings});\label{line:transSeq}
+		\State \textit{children} $ = r$.getChildren(); \Comment{\textit{children} is a list of nodes}
+		\State {\scriptsize ABSTRACT\_TRANSLATOR}(\textit{children}.removeFirst(), \textit{children});\label{line:transSeq}
 	\EndIf
@@
 \end{algorithm}
-
-This approach is not completely feasible since the algorithm needs to look ahead and check the following siblings in some cases, e.g., in the case of a semantic macro with arguments (see~\Cref{fig:syntax-tree-usecase}).
-The {\footnotesize \verb|TRANSLATE_LEAF|} function needs to know the following siblings of the current node.
-Algorithm~\ref{alg:translation} explains a more detailed but abstract version of the final algorithm of the presented translation tool.
 
@@
 \end{table}
-This approach seems to work for most semantic \LaTeX{} expressions.
-However, there is one problem with this straightforward approach which was mentioned in~\Cref{sec:problems}.
-There are many different types of notations used to represent formulae.
-\Cref{tab:notations} illustrates the simple expression $(a+b)x$ in different notations.
+
+The look ahead functionality seems to solve the problems for functions.
+But there is another problem that was not mentioned in \Cref{sec:problems}, because it seems to be not a problem previously.
+In some cases the arguments of a function do not follow but precede the function node.
+
+If we take a closer look to mathematical notations we discover many different types of notations used to represent formulae.
+\Cref{tab:notations} illustrates the expression $(a+b)x$ in different notations.
 The \gls*{pol}\footnote{Also known as \textit{Warsaw Notation} or \textit{prefix notation}} (hereafter called prefix notation) places the operator to the left of/before its operands.
@@
 Therefore, the translator does not need to know that `$a$' and `$b$' are the operands of the binary operator `$+$' in `$a+b$.' The translator could simply translate the symbols in `$a+b$' in the same order as they appear in the expression and the \gls*{cas} would understand it.
-However, this simple approach generates two problems.
+However, there are two new problems with this approach.
 \begin{enumerate}\item \label{prob:1} The translated expression is only syntactically correct if the input expression was syntactically correct.
@@
 Since we want to develop a translation tool and not a verification tool for mathematical \LaTeX{} expressions, we can assume syntactically correct input expressions and produce errors otherwise.
-Problem~\ref{prob:2} is more difficult to solve.
+Problem~\ref{prob:2} is more complex.
 If a user wants to support a \gls*{cas} that uses prefix or postfix notation by default, the translator would fail in its current state.
@@
 On the other hand, \LaTeX{} automatically disambiguates the first example by only recognizing the very next element (single symbols or sequences in curly brackets) for the superscript and therefore displays the second interpretation.
-The second example is already interpreted as the double factorial function of $n$, since this notation is the standard interpretation in science.
-We wrote the second interpretation as the standard way in science to make it even more obvious.
-However, surprisingly, \Maple{} computes the first interpretation again rather than the common standard interpretation.
+The second example should not be misinterpreted, since this notation is the standard interpretation in science for the double factorial.
+We wrote the second interpretation with parentheses point it out more precisely.
+However, surprisingly, \Maple{} computes the first interpretation (the factorial of the factorial of $n$) again rather than the common standard interpretation.
 \begin{table}[ht]

### 2018-11-15 main.tex
Explain @ symbols in DLMF macros
Implement several requested minor changes

@@
 These macros embed necessary semantic information into \LaTeX{} expressions.
+The macros may also contain $@$ symbols preceding the variables of the function.
+The number of the symbols is used to switch between different notation styles, e.g., $\cos(x)$ and $\cos x$.
 One example of such a macro is given in Table~\ref{tab:JacobiP-usecase} for the semantic \LaTeX{} representation for the Jacobi polynomial.
@@
 A typical example of complex problems occurs for multivalued functions~\parencite{AISC:MultivaluedFunctions}.
-A \gls*{cas} usually defines \textit{branch cuts} to compute principal values of multivalued functions~\parencite{Maple:Cuts}, which makes the implementation of a theoretically continued function to a discontinuous presentation of it.
+A \gls*{cas} usually defines \textit{branch cuts} to compute principal values of multivalued functions~\parencite{Maple:Cuts}, which makes the implementation of a theoretically continuous function to a discontinuous presentation of it.
 In general, positioning branch cuts follows conventions, but can be positioned arbitrarily in many cases.
@@
 Similar to the \JOBAD{} architecture, other interactive web documents exist, such as \textit{MathDox}~\parencite{MathDox} and \textit{The Planetary System}~\parencite{Planetary}.
-All of these demonstrate the potential of the educational system.
 
@@
 This package allows \textit{sage} commands in \TeX{-}files and uses \textit{sage} in the background to compute the commands.
-In this scenario, a writer still needs to manually translate expressions to the syntax of \textit{sage}.
+In this scenario, a writer still needs to manually translate expressions to the syntax of \textit{sage}, but it is possible to integrate \gls*{cas} expression directly into \TeX{} documents.
 
@@
 \end{equation}
-A translation becomes unfeasible because the macro does not explicitely define the variable of differentiation for the functions $w_1$ and $w_2$.
+A translation becomes unfeasible because the macro does not explicitly define the variable of differentiation for the functions $w_1$ and $w_2$.
 For a correct translation, the \gls*{cas} needs to be aware of the used variable $z$.
@@
 But in this case the sequence is wrapped by \texttt{\tbs left} and \texttt{\tbs right} delimiters.
-& \texttt{\tbs left(}$a+b$\texttt{\tbs right)} is a balanced expression with three children ($a$, $+$ and $b$).\\
+Note that normal parentheses do not create balanced expressions.
+& \texttt{\tbs left(}$a+b$\texttt{\tbs right)} is a balanced expression with three children ($a$, $+$ and $b$).
+\\
 	\cline{2-4}
@@
 \Cref{tab:notations} illustrates the expression $(a+b)x$ in different notations.
-The \gls*{pol}\footnote{Also known as \textit{Warsaw Notation} or \textit{prefix notation}} (hereafter called prefix notation) places the operator to the left of/before its operands.
-The \gls*{rpol}\footnote{Also known as \textit{postfix notation}} (hereafter called postfix notation) does the opposite and places the operator to the right of/after its operands.
+The \gls*{pol}\footnote{Also known as \textit{prefix notation}, \textit{Warsaw Notation} or \textit{{\L}ukasiewicz notation}.
+It was invented by J.
+{\L}ukasiewicz 1924 to create a parenthesis-free notation~\parencite{Hamblin1962}.
+Note that this notation is indeed parenthesis-free as long as all operators have the same arity.} (hereafter called prefix notation) places the operator to the left of/before its operands.
+The \gls*{rpol}\footnote{Also known as \textit{postfix notation}.
+Also invented by J.
+{\L}ukasiewicz.
+Same as \gls*{pol} it does not need parenthesis as long as all operators have the same arity.} (hereafter called postfix notation) does the opposite and places the operator to the right of/after its operands.
 The infix notation is commonly used in arithmetic and places the operator between its operands.
@@
 Table~\ref{tab:multi-expo} shows an example for the trigonometric cosine function with multiple exponents.
+As illustrated in Table~\ref{tab:multi-expo}, the translator interprets the input as
+\begin{equation}\label{eq:multi-expo}
+\left( \cos(x)^n \right)^m.
+\end{equation}
+
 \begin{table}[ht]
@@
 \hline
-Displayed As & \rule{0pt}{0.9\normalbaselineskip} $\cos^2(x)^2$ \\
-Semantic \LaTeX & \verb|\cos^2@{x}^2| \\
-Translated \Maple{} Expression & \verb|((cos(x))^(2))^2|\\
+Displayed As & \rule{0pt}{0.9\normalbaselineskip} $\cos^n(x)^m$ \\
+Semantic \LaTeX & \verb|\cos^n@{x}^m| \\
+Translated \Maple{} Expression & \verb|((cos(x))^(n))^m|\\
 \hline
@@
 \end{table}
-
-As you can see, the input expression in semantic \LaTeX{} is interpreted as
-\begin{equation}\label{eq:multi-expo}
-\left( \cos(x)^2 \right)^2.
-\end{equation}
 

### 2018-11-16 main.tex
Added LaTeXML explanations (close #8)

@@
 \usepackage[utf8]{inputenc}
-\usepackage[english]{babel} \usepackage[T1]{fontenc} 
+\usepackage[english]{babel} 
+\usepackage[T1]{fontenc}
 \usepackage{hyperref}			
-\usepackage[acronym, nomain]{glossaries} \usepackage{amsthm}
+\usepackage[acronym, nomain]{glossaries} 
+\usepackage{amsthm}
 \usepackage{amssymb}
@@
 \usepackage{txfonts}			
-\usepackage{xcolor}				
 \usepackage{enumitem}			
@@
 \usepackage{authblk}
+
+\usepackage[dvipsnames]{xcolor}
+\usepackage{lstomdoc}
+\lstset{language=MathML,%basicstyle=\sf\footnotesize,
+	columns=fullflexible,
+    belowcaptionskip=0em,
+    belowskip=-1.5em,
+    alsoletter={'},
+    stringstyle = {\color{Blue}},
+    keywordstyle = {\color{OliveGreen!70!LimeGreen}},
+    keywordstyle = [2]{\color{Fuchsia}},
+    morekeywords = [2]{cd,definitionURL,id,xmlns}
+}
 
@@
 
+Miller also developed LaTeXML, a tool for converting \LaTeX{} expressions to MathML~\parencite{LaTeXML}.
+LaTeXML was used to generate the \gls*{dlmf} and is able to parse the \Macro s to content MathML.
+Even though many \gls*{cas} are able to import and export MathML, they fail for special functions.
+Schubotz et.~al recently performs benchmarks on several \LaTeX{} to MathML conversion tools, including LaTeXML, in~\parencite{DBLP:conf/jcdl/SchubotzGSMCG18}.\Cref{sec:trans-mml} also shows three examples of generated MathML by LaTeXML, \Maple, and \Mathematica.
+We did not use MathML as an intermediate format to perform translations between \LaTeX{} and \gls*{cas}, because of the problems content MathML has~\parencite{DBLP:conf/jcdl/SchubotzGSMCG18}.
+
 \section{Translation Problems}\label{sec:problems}
@@
 \section{Conclusion \& Future Work}\label{ch:conc-future-work}
-%The results and conclusions of the translations of the formula set from the \DLMF{} are diverse.
-Some test cases have shown us that the set of semantic macros are not perfect yet and need to be improved.
-This is seen in the example of the macro definition for the Wronskian determinant.
-However, simply adding more and more macros to improve the set is not the right decision.
-In that case, we would end up in something similar to \sTeX, which is overloaded and too complicated to use.
-If the system is no longer handy, it will not be used {\bf [HSC: I don't agree with this paragraph at all.
-In fact, I think we've had this discussion already in the past and it was resolved.
-It needs to be resolved again.
-There are many examples where enhancing macros is a benefit.
-Not only that, we've discussed this recently in our publications.
-This needs to be fixed prior to submission.
-I vote for removing this paragraph altogether until we can resolve this issue.]} {\bf [AGP: Since the paper in it's current state is quite long and the section works well without this paragraph, I just commented it out and avoid the discussion here...]}
-
 The translator concept has proven itself by discovering errors in the online \gls*{dlmf} compendia.
@@
 \printbibliography
+
+\appendix
+\section*{Appendix}
+\section{Translations via MathML}\label{sec:trans-mml}
+The problems to use MathML data as an intermediate format can be demonstrated by using LaTeXML, \Maple, and \Mathematica{} to generate content MathML from their inputs.
+While the inputs for \gls*{cas} are unique, ideally all systems should produce the same content MathML data.
+Also, for semantic \LaTeX, LaTeXML should be able to generate the correct content MathML data.
+As an example we generated content MathML via LaTeXML for
+\begin{equation}\label{eq:jac-lml}
+\verb|\JacobiP{\alpha}{\beta}{n}@{\cos@{a\Theta}}|,
+\end{equation}
+and via \Maple{} and \Mathematica{} for the manually verified appropriated  translations of (\ref{eq:jac-lml})
+\begin{eqnarray}
+&& \verb|JacobiP(n,alpha,beta,cos(a*Theta))|,\text{ and}\label{eq:jac-map}\\
+&& \verb|JacobiP[n,\[Alpha],\[Beta],Cos[a \[CapitalTheta]]]|.\label{eq:jac-mat}
+\end{eqnarray}
+Listings~\ref{lst:latexml}-~\ref{lst:mathematica} shows the generated content MathML.
+As the listings show, all generated MathML distinguish to each other.
+The problems are not only flaws in their export implementations but also in the fact that even content MathML is not unique.
+Based on this fact and the differences in the generated content MathML between each system, we decided to not use MathML as an intermediate format.
+
+\begin{lstlisting}[label={lst:latexml},mathescape=true,caption=The generated content MathML for~(\ref{eq:jac-lml}) using LaTeXML with the command \texttt{latexmlc} and the set of semantic macros.]
+<math xmlns="http://www.w3.org/1998/Math/MathML" $\ldots$>
+  <semantics>
+    <apply>
+      <apply>
+        <csymbol cd="dlmf">Jacobi-polynomial-P</csymbol>
+        <ci>$\alpha$</ci>
+        <ci>$\beta$</ci>
+        <ci>$n$</ci>
+      </apply>
+      <ci>$x$</ci>
+    </apply>
+  </semantics>
+</math>
+\end{lstlisting}
+
+\vspace{0.4cm}
+\begin{lstlisting}[label={lst:maple},mathescape=true,caption=The content MathML for~(\ref{eq:jac-map}) generated by Maple 16 (including manually added linebreaks).
+We were using the \texttt{MathML[ExportContent]} command to produce the output.]
+<math xmlns='http://www.w3.org/1998/Math/MathML'>
+  <apply id='id6'>
+    <csymbol id='id1' 
+      definitionURL='http://www.maplesoft.com/MathML/JacobiP'>
+        JacobiP
+    </csymbol>
+    <ci id='id2'>n</ci>
+    <ci id='id3'>alpha</ci>
+    <ci id='id4'>beta</ci>
+    <ci id='id5'>x</ci>
+  </apply>
+</math>
+\end{lstlisting}
+
+\vspace{0.4cm}
+\begin{lstlisting}[label={lst:mathematica},mathescape=true,caption=The content MathML for~(\ref{eq:jac-mat}) generated by Mathematica.
+We were using the \texttt{ExportString} command to produce the output.]
+<math xmlns='http://www.w3.org/1998/Math/MathML'>
+ <apply>
+  <ci>JacobiP</ci>
+  <ci>n</ci>
+  <ci>&#945;</ci>
+  <ci>&#946;</ci>
+  <ci>x</ci>
+ </apply>
+</math>
+\end{lstlisting}
 \end{document}

### 2018-11-16 main.tex
Small structure changes, deleted appendix (close #9)

--- 2018-11-15 main.tex
@@
 
+Miller also developed LaTeXML, a tool for converting \LaTeX{} expressions to MathML~\parencite{LaTeXML}.
+LaTeXML was used to generate the \gls*{dlmf} and is able to parse the \Macro s to content MathML.
+Even though many \gls*{cas} are able to import and export MathML, they fail for special functions.
+Schubotz et.~al recently performs benchmarks on several \LaTeX{} to MathML conversion tools, including LaTeXML, in~\parencite{DBLP:conf/jcdl/SchubotzGSMCG18}.
+
 \section{Translation Problems}\label{sec:problems}
@@
 \subsection{Automatic Changes of Inputs in Maple}\label{subsec:maple-probs}
+In this section, we will discuss several techniques to get access to the parse tree of \Maple's input.
+The translation process from this parse tree than follows the same principle as for the forward translations.
+
 \Maple{} evaluates inputs automatically and changes the input into an internal representation.
@@
 \printbibliography
+
+%\appendix
+%\section*{Appendix}
 \end{document}

### 2018-11-16 main.tex
Changed position of paragraph vom 6.1 to 6

@@
 \section{\Maple{} to Semantic \LaTeX{} Translator}\label{sec:backward-translation}
+In this section, we will discuss several techniques to get access to the parse tree of \Maple's input.
+The translation process from this parse tree than follows the same principle as for the forward translations.
+
 Instead of writing a custom \Maple{} syntax parser, we use \Maple's internal data structure to get a syntax tree of the input\footnote{In consequence, a license of \Maple{} is mandatory to perform backward translations.
@@
 \subsection{Automatic Changes of Inputs in Maple}\label{subsec:maple-probs}
-In this section, we will discuss several techniques to get access to the parse tree of \Maple's input.
-The translation process from this parse tree than follows the same principle as for the forward translations.
-
 \Maple{} evaluates inputs automatically and changes the input into an internal representation.

### 2018-11-16 main.tex
Extend future work by mentioning Beaumont paper (close #10)

@@
 Even though many \gls*{cas} are able to import and export MathML, they fail for special functions.
-Schubotz et.~al recently performs benchmarks on several \LaTeX{} to MathML conversion tools, including LaTeXML, in~\parencite{DBLP:conf/jcdl/SchubotzGSMCG18}.
+Schubotz et al.~recently performs benchmarks on several \LaTeX{} to MathML conversion tools, including LaTeXML, in~\parencite{DBLP:conf/jcdl/SchubotzGSMCG18}.
 
@@
 
+A theoretical improvement for the numerical tests would be to perform the tests for specific, considered as critical, values with respect to the involved functions.
+J.~C.~Beamont et al.~tested  identities for multivalued elementary functions by choosing sample points from a region with respect to the branch cuts of the functions~\parencite{DBLP:journals/aaecc/BeaumontBDP07}.
+We assume that choosing sample points from those regions, significantly improves the results of our numerical tests.
+
+Increasing the amount of translatable formulae in the \gls*{dlmf} and improving the verificiation techniques are parts of ongoing projects.
+
 \emergencystretch=1em

### 2018-11-20 main.tex
Reformulated sentence for closing #13

@@
 Consider the mentioned case of the arccotangent function~\parencite{Branches:acot} that has different positioned branch cuts in \Maple{} compared to the \gls*{dlmf} or \Mathematica{} definitions.
-Alternative but mathematically equivalent translations are
+As suggested by~\parencite{Branches:acot}, we can translate an alternative definition of the arccotangent function to avoid the branch cut issues.
+Considering the equations (23) and (25) in~\parencite{Branches:acot}, we can define three translations
 \begin{eqnarray}

### 2018-11-21 main.tex
Fix that N_0 was not defined (close #17)
Added subsection in Section 5 for better outline

@@
 
+\subsection{Problems with the Look Ahead Approach}
 The look ahead functionality seems to solve the problems for functions.
@@
 \begin{equation*}
-\frac{{\mathrm{d}}^{n}}{{\mathrm{d}z}^{n}}\left(e^{z^{2}}\operatorname{erfc}z\right),
+\frac{{\mathrm{d}}^{n}}{{\mathrm{d}z}^{n}}\left(e^{z^{2}}\operatorname{erfc}z\right), \quad n = {0,1,2,\ldots}
 \end{equation*}
-where $n \in \mathbb{N}_0$, $e$ is the base of the natural logarithm, and $\operatorname{erfc}$ is the
+where $e$ is the base of the natural logarithm, and $\operatorname{erfc}$ is the
 complementary error function. Our translation correctly produces 

### 2018-11-21 ReviewerResponse.txt
Minor changes and added ReviewerResponse

We carefully proofread the entire paper again and implemented several minor changes. In the following we response specifically to the reviewers.

Reviewer 1:
As suggested, we made several changes to improve the structure of the paper. In detail we 
•	added subsections in Section 3, 
•	gave an overview of the idea of the translator in Section 4, 
•	removed unnecessary Algorithm 1 in Section 5,
•	added a subsection in Section 5 for a better outline, and 
•	further explained Algorithm 3 (now Algorithm 2).
We also explained the PoM-Tagger, its parsing process, and the lexicon files in more detail.
Further, we added LaTeXML to the related work section and explained briefly why we did not choose MathML as an intermediate data format to perform translations by citing one of our previous publications: Schubotz et al. from 2018 (now included in our references). In this research project, we tested LaTeXML and other state of the art LaTeX to MathML conversion tools. We discovered several problems for those translations. So we decided to work with the PoM-Tagger instead of LaTeXML for our translations.
Also, we did not add further comparisons with LaTeXML to our evaluations section. The reason is that CAS does not understand MathML that contains special functions. Thus we can only compare translations without function calls or implement a solution to overcome this issue for a decent comparison. But this project may legitimate a new publication by itself.

Reviewer 2:
For clarification, we decided to change the wording "Word Processor" to "Document Preparation System.” Since this is part of the title, we did not add that we focus on LaTeX to the title and hope that the Abstract will point that out in an acceptable manner.
Also, instead of creating an Appendix, we decided to add short paragraphs that explain the purpose of the sections. We hope this helps the reader to decide whether he can skip the section, e.g., in case of his lack of interest in Maple “hackings,” or not. 
In the new version of the paper, we implemented all suggested minor changes and added the idea of improved numerical tests by using the mentioned cut region detection to the future work section.

### 2018-11-21 main.tex
Minor changes and added ReviewerResponse

@@
 	\Procedure{translate\_macro}{\textit{macro}, \textit{args}, \textit{lexicon}}
-	\State \textit{info} = \textit{lexicon}.getInfo(\textit{macro});
+	\State \textit{info} = \textit{lexicon}.getInfo(\textit{macro});\label{line:get_info}
 	\State \textit{argList} = new List(); \Comment{create a sorted list for the translated arguments.}
@@
 	\State \textit{pattern} = \textit{info}.getTranslationPattern();
-	\State \textit{translatedMacro} = \textit{pattern}.fillPlaceHolders(\textit{argList});
-	\If{\textit{power} is not \NULL}
+	\State \textit{translatedMacro} = \textit{pattern}.fillPlaceHolders(\textit{argList});\label{line:fill_pattern}
+	\If{\textit{power} is not \NULL}\label{line:shifted_exp}
 		\State \textit{translatedMacro}.add(\textit{power});
@@
 
+The translation process of the \Macro 's is complex, so there is a special class, the \verb|MacroTranslator|, that handles those nodes in the \gls*{pom-pt}.
 Algorithm~\ref{alg:macro-translation} explains the \verb|MacroTranslator| without error handling.
-It has extracted necessary information from the \gls*{pom-pt}, such as how many arguments this function has.
-In consequence, it also processes the following siblings to translate the arguments.
-There are some special cases about the next node right after the macro occurs.
-This node can be
+It has extracted necessary information from the \gls*{pom-pt}, such as how many arguments this function has, in Line~\ref{line:get_info}.
+It also processes the following siblings to translate the arguments.
+The \verb|MacroTranslator| will be called in Line~\ref{line:transLeaf} in Algorithm~\ref{alg:translation}, since the macro is a leaf node in the \gls*{pom-pt}.
+The following cases describes the different kinds of the following siblings after a semantic macro node.
+Those can be
 \begin{itemize}
-\item an exponent, such as for `\verb|^2|';
-\item an optional parameter in square brackets;
-\item a parameter in curly brackets (a \textit{sequence} node in the \gls*{pom-pt}) if none of the above;
-\item `$@$' symbols if none of the above; or
-\item a variable in curly brackets (a \textit{sequence} node) if none of the above.
+\item an exponent, such as for `\verb|^2|' right after the macro node (Line~\ref{line:next_caret});
+\item an optional parameter in square brackets right after the macro node or after an exponent (Line~\ref{line:next_optional});
+\item a parameter in curly brackets (a \textit{sequence} node in the \gls*{pom-pt}) if none of the above and no `$@$' symbols were passed yet (Line~\ref{line:trans_paras});
+\item `$@$' symbols (Line~\ref{line:skip_ats}); or
+\item a variable in curly brackets (a \textit{sequence} node) after the `$@$' symbols were passed. (Line~\ref{line:trans_vars})
 \end{itemize}
 
-An exponent will be translated and shifted to the end of the translated semantic macro, 
-since one usually writes the exponent after the arguments of a function in \gls*{cas}.
-Hence, the function translates and stores the exponent in Line~\ref{line:next_caret}.
-One could ask what happens when there is an exponent given before and after the arguments.
-The \verb|MacroTranslator| only translates the following siblings until each argument is translated.
-The first exponent will be shifted to the end.
-If right after the translated macro (with all arguments) follows another exponent, we interpret it as another exponent for the whole previous expression.
-In that case, it would be the macro with the first translated exponent.
-Table~\ref{tab:multi-expo} shows an example for the trigonometric cosine function with multiple exponents.
-As illustrated in Table~\ref{tab:multi-expo}, the translator interprets the input as
-\begin{equation}\label{eq:multi-expo}
-\left( \cos(x)^n \right)^m.
-\end{equation}
+All cases before the `$@$' symbols are optional.
+The \verb|MacroTranslator| removes all following siblings according to the number of expected parameters and variables.
+Parameter and variable nodes are translated separately.
+If an exponent was registered right after the semantic macro node, it will be shifted to the end in Line~\ref{line:shifted_exp}.
+The macro itself will be translated by putting all translated parameters and variables into the translation pattern (Line~\ref{line:fill_pattern}).
+
+Following siblings after the macro was translated (with all arguments) do not belong to the semantic macro.
+If the next node is an exponent the translated macro is the base.
+Table~\ref{tab:multi-expo} shows an example for the translation of the trigonometric cosine function with multiple exponents.
 
@@
 \centering
-\begin{tabular}{lc}
-\hline
-Displayed As & \rule{0pt}{0.9\normalbaselineskip} $\cos^n(x)^m$ \\
-Semantic \LaTeX & \verb|\cos^n@{x}^m| \\
-Translated \Maple{} Expression & \verb|((cos(x))^(n))^m|\\
+\begin{tabular}{lccc}
+ & Semantic \LaTeX & & \Maple{} \\
+\hline
+Text Representation & \rule{0pt}{1.0\normalbaselineskip} \verb|\cos^n@{x}^m| & $\overset{\langMaple}{\mapsto}$ & \verb|((cos(x))^(n))^m| \\
+Displayed As & \rule{0pt}{1.0\normalbaselineskip} $\cos^n(x)^m$ & & $\left( \cos(x)^n \right)^m$ \\
 \hline

### 2018-11-26 ReviewerResponse.txt
updated ReviewerResponse.txt to improve grammar and also added comments/questions labeled as HSC which need to be addressed

@@
-We carefully proofread the entire paper again and implemented several minor changes. In the following we response specifically to the reviewers.
+In the following we respond specifically to the reviewers comments.
 
 Reviewer 1:
-As suggested, we made several changes to improve the structure of the paper. In detail we 
-•	added subsections in Section 3, 
-•	gave an overview of the idea of the translator in Section 4, 
-•	removed unnecessary Algorithm 1 in Section 5,
-•	added a subsection in Section 5 for a better outline, and 
-•	further explained Algorithm 3 (now Algorithm 2).
-We also explained the PoM-Tagger, its parsing process, and the lexicon files in more detail.
-Further, we added LaTeXML to the related work section and explained briefly why we did not choose MathML as an intermediate data format to perform translations by citing one of our previous publications: Schubotz et al.
-from 2018 (now included in our references).
-In this research project, we tested LaTeXML and other state of the art LaTeX to MathML conversion tools.
-We discovered several problems for those translations.
-So we decided to work with the PoM-Tagger instead of LaTeXML for our translations.
-Also, we did not add further comparisons with LaTeXML to our evaluations section.
-The reason is that CAS does not understand MathML that contains special functions.
-Thus we can only compare translations without function calls or implement a solution to overcome this issue for a decent comparison.
-But this project may legitimate a new publication by itself.
+
+As suggested, we made several changes to improve the structure of the paper.  
+In summary, we have :
+(1) added subsections in Section 3 [HSC: Can you say what these are?];
+(2) gave an overview of the idea [HSC: idea? concept?] of the translator in Section 4;
+(3) removed unnecessary Algorithm 1 in Section 5 [HSC: give further explanation here, why is it unnecessary?];
+(4) added a subsection in Section 5 for a better outline [HSC: give name of the subsection.........]; and 
+(6) further explained Algorithm 3 (now Algorithm 2) [HSC: more details, what did you further explain?].
+
+We also gave further detailed descriptions about the PoM-Tagger, 
+namely its parsing process and lexicon files.
+
+We also added LaTeXML to the related work section and gave a brief 
+explanation with cited research (Schubotz et al., 2018, now included 
+in our references), in order to explain why we did not choose MathML 
+as an intermediate data format for performing translations. In that 
+citation, we described detailed testing that we performed using LaTeXML 
+and other state of the art LaTeX to MathML conversion tools. In that
+reference, we discovered several critical problems which could occur 
+when translations would be performed using MathML as an intermediate
+data format.
+
+Note that we did not add further comparisons with LaTeXML to our 
+evaluations section, because current CAS are unable to understand MathML 
+which contains special functions. Thus we can only compare translations 
+without function calls or implement a solution which overcomes this 
+problem, in order to make a decent comparison. Because of the great amount 
+of problems which would need to be overcome, a full solution translation 
+implementation using MathML is out of the scope of our current project. 
+This is now all explained in detail in the new version of the manuscript.
 
 Reviewer 2:
-For clarification, we decided to change the wording "Word Processor" to "Document Preparation System.” Since this is part of the title, we did not add that we focus on LaTeX to the title and hope that the Abstract will point that out in an acceptable manner.
-Also, instead of creating an Appendix, we decided to add short paragraphs that explain the purpose of the sections.
-We hope this helps the reader to decide whether he can skip the section, e.g., in case of his lack of interest in Maple “hackings,” or not.
-In the new version of the paper, we implemented all suggested minor changes and added the idea of improved numerical tests by using the mentioned cut region detection to the future work section.
+
+For clarification, throughout the manuscript, we decided to change the 
+wording "Word Processor" to "Document Preparation System." Since this wording
+occurred as part of the title, we did not add that we focus on LaTeX to the 
+title and hope that the Abstract points this out in an acceptable manner.
+
+Also, instead of creating an Appendix, we decided to add short paragraphs 
+which explain the purpose of the sections [HSC: What sections?]. We hope 
+this helps the reader to decide whether he can skip the section, e.g., in 
+case of his lack of interest in Maple "hackings", or not.  In the new 
+version of the paper, we implemented all suggested minor changes as well as 
+adding the idea of improved numerical tests by using the mentioned cut 
+region detection to the future work section.

### 2018-11-26 main.tex
made some changes to main.tex

@@
 \noindent
-\textbf{Purpose:} Modern mathematicians and scientists of math-related disciplines use \glspl*{dps} to write and \glspl*{cas} to calculate mathematical expressions.
+\textbf{Purpose:} Modern mathematicians and scientists of math-related disciplines often use \glspl*{dps} to write and \glspl*{cas} to calculate mathematical expressions.
 Usually, they translate the expressions manually between \gls*{dps} and \gls*{cas}.
@@
 
-\noindent\textbf{Approach:} The \gls*{nist} developed a collection of special \LaTeX{} macros that create links from mathematical symbols to their definitions in the \gls*{nist} \gls*{dlmf}.
+\noindent\textbf{Approach:} Bruce Miller at \gls*{nist} developed a collection of special \LaTeX{} macros that create links from mathematical symbols to their definitions in the \gls*{nist} \gls*{dlmf}.
 We are using these macros to perform rule-based translations between the formulae in the \gls*{dlmf} and \gls*{cas}.
@@
 where $n$ is a nonnegative integer, $\alpha, \beta > -1$, and $a,\Theta\in{\mathbf R}$.
-This mathematical expression is written in \LaTeX{} as
+This mathematical expression can be written in \LaTeX{} as
 \begin{equation*}
@@
 \begin{equation*}
-	\mathscr{W} \left\{ w_1(z), w_2(z) \right\} = w_1(z)w_2'(z) - w_2(z)w_1'(z).
+\mathscr{W}\{ w_1(z), w_2(z) \} = w_1(z)w_2'(z) - w_2(z)w_1'(z).
 \end{equation*}
@@
 \item `$@$' symbols (Line~\ref{line:skip_ats}); or
-\item a variable in curly brackets (a \textit{sequence} node) after the `$@$' symbols were passed. (Line~\ref{line:trans_vars})
+\item a variable in curly brackets (a \textit{sequence} node) after the `$@$' symbols were 
+passed (Line~\ref{line:trans_vars}).
 \end{itemize}
@@
 In this section, we will discuss several techniques to get access to the parse tree of \Maple's input.
-The translation process from this parse tree than follows the same principle as for the forward translations.
+The translation process from this parse tree then follows the same principle as for the forward translations.
 
@@
 \end{equation}
-Assume the translator would forward translate the expression based on (\ref{eq:wrong-trans-1},~\ref{eq:wrong-trans-2}). Than
+Assume the translator would forward translate the expression based on (\ref{eq:wrong-trans-1},~\ref{eq:wrong-trans-2}). Then
 \begin{eqnarray}

### 2018-11-27 ReviewerResponse.txt
Improve response to reviewers (close #19)

@@
 In summary, we have :
-(1) added subsections in Section 3 [HSC: Can you say what these are?];
-(2) gave an overview of the idea [HSC: idea? concept?] of the translator in Section 4;
-(3) removed unnecessary Algorithm 1 in Section 5 [HSC: give further explanation here, why is it unnecessary?];
-(4) added a subsection in Section 5 for a better outline [HSC: give name of the subsection.........]; and 
-(6) further explained Algorithm 3 (now Algorithm 2) [HSC: more details, what did you further explain?].
+(1) added subsection headings "Different Sets of Defined Functions", 
+"Positions of Branch Cuts", "Insufficient Semantic Information", and 
+"Ambiguous Expressions" in Section 3;
+(2) gave an overview of the concept of the translator in Section 4;
+(3) removed unnecessary Algorithm 1 in Section 5 as suggested, because 
+it was just a simplified pre-version of Algorithm 2;
+(4) added the subsection heading "Problems with the Look Ahead Approach" in 
+Section 5 for a better outline; and 
+(6) further explained the different states of Algorithm 3 (now Algorithm 2) 
+by refer to the lines in the descriptions. We also explained how a power 
+after a macro is handled in more detail and why optional parameters needs 
+to be handled differently to other arguments.
 
@@
 Also, instead of creating an Appendix, we decided to add short paragraphs 
-which explain the purpose of the sections [HSC: What sections?]. We hope 
+which explain the purpose of the Sections 4 and 6. We hope 
 this helps the reader to decide whether he can skip the section, e.g., in 

### 2018-11-27 main.tex
Fix figure 4 wrapping (close #21)
Delted the phrase 'in consequence' all over the paper (close #22)

@@
 Especially important for a CAS is to embed unambiguous semantic information within the input.
-Each system uses different representations and syntax in consequence.
-Hence, a writer needs to constantly translate mathematical expressions from one representation to another and back again.
+Therefore, each system uses different representations and syntax, so that a writer needs to continually translate mathematical expressions from one representation to another and back again.
 Table~\ref{tab:JacobiP-usecase} shows four different representations for~(\ref{eq:P}).
@@
 Hence, a \gls*{cas} user needs to fully understand the properties and special definitions (such as the position of branch cuts) in the \gls*{cas} to avoid mistakes during a translation~\parencite{Maple:Cuts}.
-In consequence, a manual translation process is not only laborious, but also prone to errors.
+Hence, a manual translation process is not only laborious, but also prone to errors.
 Note that this general problem has been named as automatic \gls*{p2c} conversion~\parencite{POM-Tagger}.
@@
 During the export process, the semantic information gets lost, because generic \LaTeX{} is not able to carry sufficient semantic information.
-In consequence of these two problems, an exported expression cannot be imported to the same system again in most cases (except for simple expressions such as those described above).
+Because of these two problems, an exported expression cannot be imported to the same system again in most cases (except for simple expressions such as those described above).
 Our tool should solve these problems and provide round-trip translations between \LaTeX{} and \gls*{cas}.
@@
 The arctangent function~\eqref{eq:acot-alternatives-1} and the natural logarithm~(\ref{eq:acot-alternatives-2}) have the same positioned branch cuts as in the \gls*{dlmf} and in \Maple.
-In consequence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user does not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z = -\iunit$.
+Hence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user does not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z = -\iunit$.
 
@@
 However, in the current state, many expressions cannot be disambiguated yet.
-In consequence, the \gls*{pom}-tagger generates a raw parsed tree where each token in the \LaTeX{} expression is a node in the tree.
+Therefore, the \gls*{pom}-tagger generates a raw parsed tree where each token in the \LaTeX{} expression is a node in the tree.
 We call this parsed tree the \gls*{pom-pt}.
@@
 
+\begin{wrapfigure}[10]{r}{0.38\textwidth} % optional parameter gives "# of narrow lines", very useful! 
+\vspace{-8pt}
+\begin{minipage}{0.38\textwidth}
+\center
+\begin{tabular}{lc}
+	\hline
+	Notation & Expression \\
+	\hline
+	Infix & $(a+b) \cdot x$\\
+	Prefix & $\cdot + a\ b\ x$\\
+	Postfix & $a\ b + x\ \cdot$\\
+	Functional & $\cdot(+(a, b), x)$\\
+	\hline
+\end{tabular}
+%\vspace{-5pt}
+\caption{The mathematical expression `$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
+\label{tab:notations}
+\end{minipage}
+%\vspace{-5pt}
+\end{wrapfigure}
+
 If we take a closer look to mathematical notations we discover many different types of notations used to represent formulae.
@@
 This only makes sense as long as the operator is a binary operator.
-
-\begin{wrapfigure}[10]{r}{0.38\textwidth} % optional parameter gives "# of narrow lines", very useful! 
-\vspace{-8pt}
-\begin{minipage}{0.38\textwidth}
-\center
-\begin{tabular}{lc}
-	\hline
-	Notation & Expression \\
-	\hline
-	Infix & $(a+b) \cdot x$\\
-	Prefix & $\cdot + a\ b\ x$\\
-	Postfix & $a\ b + x\ \cdot$\\
-	Functional & $\cdot(+(a, b), x)$\\
-	\hline
-\end{tabular}
-%\vspace{-5pt}
-\caption{The mathematical expression `$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
-\label{tab:notations}
-\end{minipage}
-%\vspace{-5pt}
-\end{wrapfigure}
 
@@
 
-Instead of writing a custom \Maple{} syntax parser, we use \Maple's internal data structure to get a syntax tree of the input\footnote{In consequence, a license of \Maple{} is mandatory to perform backward translations.
+Instead of writing a custom \Maple{} syntax parser, we use \Maple's internal data structure to get a syntax tree of the input\footnote{Therefore, a license of \Maple{} is mandatory to perform backward translations.
 The translator is using the version \Maple{} 2016.}.

### 2018-11-27 main.tex
Avoid one-sentence paragraphs (close #24)

@@
 \end{table}
+\vspace*{-0.5cm}
 
@@
 \end{algorithm}
+\vspace*{-0.2cm}
 
@@
 The translation process from this parse tree then follows the same principle as for the forward translations.
-
 Instead of writing a custom \Maple{} syntax parser, we use \Maple's internal data structure to get a syntax tree of the input\footnote{Therefore, a license of \Maple{} is mandatory to perform backward translations.
@@
 Some of the important types for the nodes are specified in \Cref{tab:maple-types}.
+The translator uses the \texttt{OpenMaple}~\parencite[\S 14.3]{MAPLE:ProgrammingGuide} \gls*{api} for interacting with \Maple's kernel implementation.
+
 \begin{table}[ht]
@@
 \end{table}
-The translator uses the \texttt{OpenMaple}~\parencite[\S 14.3]{MAPLE:ProgrammingGuide} \gls*{api} for interacting with \Maple's kernel implementation.
 
@@
 A \textbf{fixed point representation} (or short fixed point) in a round trip translation process is a string representation that is identical to all string representations in the following cycles.
-
-Table~\ref{tab:fixpoint} illustrates an example of a round trip test which reaches a fixed point. The test formula is
+Table~\ref{tab:fixpoint} illustrates an example of a round trip test which reaches a fixed point for the formula
 \begin{equation}
@@
 \verb|\EllIntF@{\phi}{k}| & \overset{\langMaple}{\mapsto} & \verb|EllipticF(sin(phi),k)|,\\
-\verb|\EllIntF@{\asin@{\phi}}{k}| & \overset{\langMaple}{\mapsfrom} & \verb|EllipticF(phi,k)|.
+\verb|\EllIntF@{\asin@{\phi}}{k}| & \overset{\langMaple}{\mapsfrom} & \verb|EllipticF(phi,k)|,
 \end{eqnarray}
-The round-trip translations produce infinite chains of sine and inverse sine calls because there are no evaluations involved.
+and the round-trip translations produce infinite chains of sine and inverse sine calls because there are no evaluations involved.
 
@@
 We assume that choosing sample points from those regions, significantly improves the results of our numerical tests.
-
 Increasing the amount of translatable formulae in the \gls*{dlmf} and improving the verificiation techniques are parts of ongoing projects.

### 2018-11-27 main.tex
Changes in the title: 'Representations of' to 'Expressions involving' (close #20)

@@
 \hypersetup{
-	pdftitle	= {Semantic Preserving Bijective Mappings for Representations of Special Functions in Computer Algebra Systems and Document Preparation Systems}
+	pdftitle	= {Semantic Preserving Bijective Mappings for Expressions involving Special Functions in Computer Algebra Systems and Document Preparation Systems}
 }
@@
 \title{
-    Semantic Preserving Bijective Mappings for Representations of Special Functions between Computer Algebra Systems and Document Preparation Systems
+    Semantic Preserving Bijective Mappings for Expressions involving Special Functions between Computer Algebra Systems and Document Preparation Systems
 }

### 2018-11-27 ReviewerResponse.txt
Update reviewer response for second changes in the title (close #20)

@@
 For clarification, throughout the manuscript, we decided to change the 
-wording "Word Processor" to "Document Preparation System." Since this wording
+wording "Word Processor" to "Document Preparation System," and 
+"Representations of Special Functions" to "Expressions involving Special 
+Functions." Since this wording
 occurred as part of the title, we did not add that we focus on LaTeX to the 

### 2018-11-27 ReviewerResponse.txt
made changes to ReviewerResponse.txt

@@
 Section 5 for a better outline; and 
+[HSC: Why isn't there a (5) entry?? If not necessary, make the next one (5) instead of (6).]
 (6) further explained the different states of Algorithm 3 (now Algorithm 2) 
-by refer to the lines in the descriptions. We also explained how a power 
-after a macro is handled in more detail and why optional parameters needs 
-to be handled differently to other arguments.
+by referring to the lines in the descriptions [HSC: descriptions? please clarify]. 
+We also explained how a power after a macro is handled in more detail and why 
+optional parameters need to be handled differently than other arguments.
 
-We also gave further detailed descriptions about the PoM-Tagger, 
+We also gave further detailed descriptions concerning the PoM-Tagger, 
 namely its parsing process and lexicon files.
@@
 citation, we described detailed testing that we performed using LaTeXML 
-and other state of the art LaTeX to MathML conversion tools. In that
+and other state-of-the-art LaTeX to MathML conversion tools. In that
 reference, we discovered several critical problems which could occur 
-when translations would be performed using MathML as an intermediate
-data format.
+when translations would be performed using MathML as an intermediate data format.
 
@@
 evaluations section, because current CAS are unable to understand MathML 
-which contains special functions. Thus we can only compare translations 
-without function calls or implement a solution which overcomes this 
-problem, in order to make a decent comparison. Because of the great amount 
-of problems which would need to be overcome, a full solution translation 
+which contains special functions. Thus, in order to make a decent 
+comparison, we can only compare translations without function calls or 
+implement a solution which is able overcome these problems.  Because of 
+the problems which would need to be overcome, a full solution translation 
 implementation using MathML is out of the scope of our current project. 
@@
 "Representations of Special Functions" to "Expressions involving Special 
-Functions." Since this wording
-occurred as part of the title, we did not add that we focus on LaTeX to the 
-title and hope that the Abstract points this out in an acceptable manner.
+Functions." Since this wording occurred as part of the title, we did not 
+add that we focus on LaTeX to the title and hope that the Abstract points 
+this out in an acceptable manner.
 
 Also, instead of creating an Appendix, we decided to add short paragraphs 
-which explain the purpose of the Sections 4 and 6. We hope 
-this helps the reader to decide whether he can skip the section, e.g., in 
-case of his lack of interest in Maple "hackings", or not.  In the new 
-version of the paper, we implemented all suggested minor changes as well as 
-adding the idea of improved numerical tests by using the mentioned cut 
-region detection to the future work section.
+which explain the purpose of Sections 4 and 6. We hope this helps the reader 
+to decide whether he can skip the section, e.g., in case of his lack of 
+interest in Maple "hackings", or not.  In the new version of the paper, we 
+implemented all suggested minor changes as well as adding the idea of 
+improved numerical tests by using the mentioned cut region detection to 
+the future work section.

### 2018-11-27 main.tex
updated in/appropriate section

@@
 The theoretical construct for this symbol cannot be mathematically equivalent to the value \verb|exp(1)| in \Maple, caused by computational and implementational limitations.
-Instead of using the term \textit{equivalent}, we introduce \textit{appropriate} and \textit{inappropriate} translations.
+
+Instead of using the term \textit{equivalent}, 
+we introduce the terms \textit{appropriate} and \textit{inappropriate} translations
+{\bf [HSC: I don't understand how {\it equivalent} translations has anything at 
+all to do with {\it appropriate} or {\it inapppropriate} translations. Please clarify.]}.
+{\bf [HSC: The supposed definition is: ``We consider a translation as {\it appropriate}, when a numerical evaluation returns the 
+same values in both concepts up to a numerical precision $|\epsilon|<<1$, for all possible 
+points in the domain of the functions.'' We also need a definition for innapropriate translation.]}
 We call a translation such as

### 2018-11-29 ReviewerResponse.txt
Updated ReviewerResponse again

@@
 Section 5 for a better outline; and 
-[HSC: Why isn't there a (5) entry?? If not necessary, make the next one (5) instead of (6).]
-(6) further explained the different states of Algorithm 3 (now Algorithm 2) 
-by referring to the lines in the descriptions [HSC: descriptions? please clarify]. 
+(5) further explained the different states of Algorithm 3 (now Algorithm 2) 
+by referring to specific lines in the algorithm. 
 We also explained how a power after a macro is handled in more detail and why 

### 2018-11-29 main.tex
Include def for appropriate/inappropriate translations

--- 2018-11-27 main.tex
@@
 The theoretical construct for this symbol cannot be mathematically equivalent to the value \verb|exp(1)| in \Maple, caused by computational and implementational limitations.
-Instead of using the term \textit{equivalent}, we introduce \textit{appropriate} and \textit{inappropriate} translations.
-We call a translation such as
+
+Instead of using the term \textit{equivalent}, 
+we introduce the terms \textit{appropriate} and \textit{inappropriate} translations. 
+%{\bf [HSC: I don't understand how {\it equivalent} translations has anything at 
+%all to do with {\it appropriate} or {\it inapppropriate} translations. Please clarify.]}.
+We consider a translation as {\it appropriate}, when a numerical evaluation returns the 
+same values in both concepts up to a numerical precision $|\epsilon|<<1$, for all possible 
+points in the domain of the functions.
+A translation is considered as {\it inappropriate}, when it is not {\it appropriate}.
+
+For example, a translation such as
 \begin{equation}\label{eq:cos-def}
@@
 \end{equation}
-\textit{appropriate}, while a translation such as
+is \textit{appropriate}, while a translation such as
 \begin{equation}
@@
 is \textit{inappropriate}.
-Note that it is not always as easy as in this example to decide if a translation is appropriate or not.
-Therefore, this article also presents several validation techniques to automatically verify if a translation is appropriate or inappropriate.
+Note that it is not always as easy as in this example to decide if a translation is {\it appropriate} or not.
+Therefore, this article also presents several validation techniques to automatically verify if a translation is {\it appropriate} or {\it inappropriate}.
 In addition to this terminology, we introduce \textit{direct translations}.
-Later in the paper, we will explain that a translation from one specific mathematical object to its \textit{appropriate} counterpart in the other system is not always possible.
-We call a translation to the \textit{appropriate} counterpart \textit{direct}.
+Most mathematical objects in one system have a direct counterpart in other systems.
+However, later in the paper, we will explain that a translation from one specific mathematical object to its counterpart in the other system is not always \textit{appropriate}.
+Also, not every mathematical object has a counterpart in other systems.
+We call a translation to its counterpart \textit{direct}.
 For example, the translation~(\ref{eq:cos-def}) is \textit{direct}, while a translation to the definition of the cosine function
@@
 \end{equation*}
-is not a \textit{direct} translation.
+is not a \textit{direct} translation even though it is {\it appropriate}.
 Note that partial results of this paper have been published in~\parencite{CICM:Paper}.
@@
 \vspace{-0.2cm}
-The position of the branch cut of the arccotangent function differs after translation~\eqref{eq:acot-alternatives}, which may lead to incorrect calculations later on.
+The position of the branch cut of the arccotangent function differs after the {\it direct} translation~\eqref{eq:acot-alternatives}, which may lead to incorrect calculations later on.
 The alternative translations~\eqref{eq:acot-alternatives-1} and~\eqref{eq:acot-alternatives-2} use other functions instead of the arccotangent function.
@@
 Hence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user does not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z = -\iunit$.
+Note that none of the translations~(\ref{eq:acot-alternatives}-\ref{eq:acot-alternatives-2}) are {\it appropriate}.
 

### 2018-11-29 main.tex
made updates to main.tex

@@
 
-\noindent\textbf{Approach:} Bruce Miller at \gls*{nist} developed a collection of special \LaTeX{} macros that create links from mathematical symbols to their definitions in the \gls*{nist} \gls*{dlmf}.
+\noindent\textbf{Approach:} Bruce Miller at the \gls*{nist} developed a collection of special \LaTeX{} macros that create links from mathematical symbols to their definitions in the \gls*{nist} \gls*{dlmf}.
 We are using these macros to perform rule-based translations between the formulae in the \gls*{dlmf} and \gls*{cas}.
@@
 
-\noindent\textbf{Findings:} We created 396 mappings and translated 58.8\% of the \gls*{dlmf} (2,405 expressions) successfully between \Maple{} and \gls*{dlmf}.
-For a significant percentage, the definitions in \Maple{} and the \gls*{dlmf} were different.
+\noindent\textbf{Findings:} We created 396 mappings and translated 58.8\% of \gls*{dlmf} formulae (2,405 expressions) successfully between \Maple{} and \gls*{dlmf}.
+For a significant percentage, the special function definitions in \Maple{} and the \gls*{dlmf} were different.
 Consequently, an atomic symbol in one system maps to a composite expression in the other system.
@@
 \end{equation}
-where $n$ is a nonnegative integer, $\alpha, \beta > -1$, and $a,\Theta\in{\mathbf R}$.
+where $n$ is a nonnegative integer, $\alpha, \beta > -1$, and $a,\Theta\in \Real$.
 This mathematical expression can be written in \LaTeX{} as
@@
 The macros may also contain $@$ symbols preceding the variables of the function.
-The number of the symbols is used to switch between different notation styles, e.g., $\cos(x)$ and $\cos x$.
+The number of $@$ symbols is used to switch between different notation styles, e.g., $\cos(x)$ and $\cos x$.
 One example of such a macro is given in Table~\ref{tab:JacobiP-usecase} for the semantic \LaTeX{} representation for the Jacobi polynomial.
@@
     }
-    \caption{Two plots for the real part of the arccotangent function with a branch cut at $[-\iunit\infty, -\iunit]$, $[\iunit,\iunit\infty]$ in Figure~\protect\subref{fig:acot-cut1} and at $[-\iunit, \iunit]$ in Figure~\protect\subref{fig:acot-cut2}, respectively.
+    \caption{Two plots of the real part for the arccotangent function with a branch cut at $[-\iunit\infty, -\iunit]$, $[\iunit,\iunit\infty]$ in Figure~\protect\subref{fig:acot-cut1} and at $[-\iunit, \iunit]$ in Figure~\protect\subref{fig:acot-cut2}, respectively.
 (Plotted with \Maple{} 2016)}
@@
 
-Instead of using the term \textit{equivalent}, 
+In order to clarify the notion of \textit{equivalence} (or lack thereof) in our context of translations,
 we introduce the terms \textit{appropriate} and \textit{inappropriate} translations. 
@@
 %all to do with {\it appropriate} or {\it inapppropriate} translations. Please clarify.]}.
-We consider a translation as {\it appropriate}, when a numerical evaluation returns the 
-same values in both concepts up to a numerical precision $|\epsilon|<<1$, for all possible 
-points in the domain of the functions.
+We consider a translation to be {\it appropriate}, when a numerical evaluation returns the 
+same values in both concepts up to a numerical precision $|\epsilon|\ll 1$, for all possible 
+points in specified domains for the functions.
 A translation is considered as {\it inappropriate}, when it is not {\it appropriate}.
@@
 Therefore, this article also presents several validation techniques to automatically verify if a translation is {\it appropriate} or {\it inappropriate}.
-In addition to this terminology, we introduce \textit{direct translations}.
+
+In addition, we also introduce the notion of \textit{direct} translations.
 Most mathematical objects in one system have a direct counterpart in other systems.
@@
 \section{Related Work}\label{sec:related-work}
-Since \LaTeX{} became the de facto standard for writing papers in mathematics, most of the \gls*{cas} provide simple functions to import and export mathematical \LaTeX{} expressions\footnote{The selected \gls*{cas} Maple, Mathematica, Matlab, and SageMath provide import and/or export functions for \LaTeX:\\
+Since \LaTeX{} became the de facto standard for writing papers in mathematics, most \gls*{cas} provide simple functions to import and export mathematical \LaTeX{} expressions\footnote{The selected \gls*{cas} Maple, Mathematica, Matlab, and SageMath provide import and/or export functions for \LaTeX:\\
 Maple, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;
@@
 During the export process, the semantic information gets lost, because generic \LaTeX{} is not able to carry sufficient semantic information.
-Because of these two problems, an exported expression cannot be imported to the same system again in most cases (except for simple expressions such as those described above).
-Our tool should solve these problems and provide round-trip translations between \LaTeX{} and \gls*{cas}.
-
-The semantics must be well known before an expression can be translated.
-There are two main approaches to solve that problem: (1) someone could specify the semantic information during the writing process (pre-defined semantics); and (2) the translator can find out the right semantic information in general mathematical expressions before it translates the expression.
+Because of these problems, an exported expression cannot be imported to the same system again in most cases (except for simple expressions such as those described above).
+Our tool attempts to solve these problems and provide round-trip translations between \LaTeX{} and \gls*{cas}.
+
+The semantics must be well-known before an expression can be translated.
+There are two main approaches to solve that problem:~(1) someone could specify the semantic information during the writing process (pre-defined semantics); and (2) the translator can determine the correct semantic information in general mathematical expressions before it translates the expression.
 So-called \textit{interactive documents}\footnote{There is no adequate definition for what interactive documents are.
@@
 This package allows \textit{sage} commands in \TeX{-}files and uses \textit{sage} in the background to compute the commands.
-In this scenario, a writer still needs to manually translate expressions to the syntax of \textit{sage}, but it is possible to integrate \gls*{cas} expression directly into \TeX{} documents.
+In this scenario, a writer still needs to manually translate expressions to the syntax of \textit{sage}, but it is possible to integrate \gls*{cas} expressions directly into \TeX{} documents.
 
@@
 Namely, \sTeX{}~\parencite{sTeX} developed by Kohlhase and the \Macro s developed by Miller~\parencite{DLMF:Macros}.
-This paper shows that it is possible to develop a context-free translation tool using the semantic macros introduced by these two projects.
-The goal of \sTeX{} was to markup the functional structure of mathematical documents so that they can be exported to the \gls*{omdoc} format.
-The macro set developed by Miller introduces new macros for special functions, orthogonal polynomials, and mathematical constants.
+This paper shows that it is possible to develop a context-free translation tool using the semantic macros introduced by these two projects {\bf [HSC: which two projects?
+Please clarify.]}.
+The goal of \sTeX{} is to markup the functional structure of mathematical documents so that they can be exported to the \gls*{omdoc} format.
+The macro functionality developed by Miller introduces new macros for special functions, orthogonal polynomials, and mathematical constants.
 Each of these macros ties specific character sequences to a well-defined mathematical object and is linked with the corresponding definition in the \gls*{dlmf} or \gls*{drmf}.
@@
 We gave the \Macro{} set the favor for developing the translation engine because it provides \gls*{dlmf} definitions for a comprehensive number of functions.
-In contrast, \sTeX{} does not focus on the semantics of functions, is too complex to use, and defines diverse macros for symbols and concepts that \gls*{cas} usually does not support.
+In contrast, \sTeX{} does not focus on the semantics of functions, is often complex to use, and defines diverse macros for symbols and concepts that \gls*{cas} usually does not usually support.
 
 Miller also developed LaTeXML, a tool for converting \LaTeX{} expressions to MathML~\parencite{LaTeXML}.
-LaTeXML was used to generate the \gls*{dlmf} and is able to parse the \Macro s to content MathML.
+LaTeXML is used to generate the \gls*{dlmf} and is able to parse the \Macro s to generate content MathML.
 Even though many \gls*{cas} are able to import and export MathML, they fail for special functions.
-Schubotz et al.~recently performs benchmarks on several \LaTeX{} to MathML conversion tools, including LaTeXML, in~\parencite{DBLP:conf/jcdl/SchubotzGSMCG18}.
+Schubotz and collaborators recently performed benchmarks on several \LaTeX{} to MathML conversion tools, including LaTeXML, in~\parencite{DBLP:conf/jcdl/SchubotzGSMCG18}.
 
@@
 \subsection{Different Sets of Defined Functions}
-If a function is defined in one system but not in the other, we can translate the definition of the mathematical function.
+If a function is defined in one system but not in the other, sometimes we can easily translate the definition of the mathematical function.
 For example, the \textit{Gudermannian}~\parencite[(4.23.10)]{NIST:DLMF} $\Gudermannian{x}$ function is defined by
@@
 Since translations such as these are nonintuitive, describing explanations become necessary for the translation process.
-A special logging function takes care of each translation and provides details after a successful translation process.
+A special logging function takes care {\bf [HSC: ``takes care'' is too informal.
+Use a different phrase or word.]} of each translation and provides details after a successful translation process.
 \Cref{sec:forward-translation} explains this task further.
@@
 \subsection{Positions of Branch Cuts}
-In case of differences between defined branch cuts we can also use alternative translations to solve the problems.
-Consider the mentioned case of the arccotangent function~\parencite{Branches:acot} that has different positioned branch cuts in \Maple{} compared to the \gls*{dlmf} or \Mathematica{} definitions.
+In case of differences between defined branch cuts, we can also use alternative translations to solve the problems.
+Consider the mentioned case of the arccotangent function~\parencite{Branches:acot} that has different positioned branch cuts in \Maple{} as compared to the \gls*{dlmf} or \Mathematica{} definitions.
 As suggested by~\parencite{Branches:acot}, we can translate an alternative definition of the arccotangent function to avoid the branch cut issues.
-Considering the equations (23) and (25) in~\parencite{Branches:acot}, we can define three translations
+Considering
+\parencite[(23), (25)]{Branches:acot}, we can define three translations
 \begin{eqnarray}
@@
 The arctangent function~\eqref{eq:acot-alternatives-1} and the natural logarithm~(\ref{eq:acot-alternatives-2}) have the same positioned branch cuts as in the \gls*{dlmf} and in \Maple.
-Hence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user does not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z = -\iunit$.
+Hence, translation~(\ref{eq:acot-alternatives-1}) solves this issue as long as the user does not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z=-\iunit$.
 Note that none of the translations~(\ref{eq:acot-alternatives}-\ref{eq:acot-alternatives-2}) are {\it appropriate}.
@@
 One example is the \textit{Wronskian} determinant.
-For two differentiable functions, the \textit{Wronskian} is defined as~\parencite[(1.13.4)]{NIST:DLMF}
+For two differentiable functions $w_1$, $w_2$, the \textit{Wronskian} is defined as~\parencite[(1.13.4)]{NIST:DLMF}
 \begin{equation*}
@@
 \end{equation*}
-In semantic \LaTeX{} it is currently implemented using
+In semantic \LaTeX{}, it is currently implemented using
 \begin{equation}
@@
 \end{equation}
-A translation becomes unfeasible because the macro does not explicitly define the variable of differentiation for the functions $w_1$ and $w_2$.
-For a correct translation, the \gls*{cas} needs to be aware of the used variable $z$.
-We solved this issue by creating a new macro, e.g.,
+This translation is unfeasible because the macro does not explicitly define the variable of differentiation for the functions $w_1$, $w_2$.
+For a correct translation, the \gls*{cas} needs to be aware of the variable of differentiation $z$.
+We solved this issue by creating a new macro \verb|\Wron|, e.g.,
 \begin{equation}
@@
 \end{equation}
-This example demonstrates that the \Macro s are still a work in progress and are getting constantly updated.
-
-\subsection{Ambiguous Expressions}
-A similar problem is multiplications since they are rarely explicitly marked in \LaTeX{} expressions, e.g., scientists using whitespaces to indicate multiplications rather than using \verb|\cdot| or similar symbols.
+This example demonstrates that the \Macro s are still a work in progress and further updates are sometimes
+necessary in order to further encapsulate critical semantic information {\bf [HSC: I changed this sentence. Check to make sure that it is still correct.]}.
+
+\subsection{Potentially Ambiguous Expressions}
+
+{\bf [HSC: This section needs to start out with a short overview paragraph describing the issue instead of an example. Also, I changed the title of the section.]}
+
+A similar problem is multiplications since they are rarely explicitly marked in \LaTeX{} expressions, e.g., scientists using whitespace to indicate multiplications rather than using \verb|\cdot| or similar symbols.
 For such problems, we introduced a new macro \verb|\idot| for an invisible multiplication symbol (this macro will not be rendered).
 Since this macro is newly introduced by contributers of the \gls*{drmf} team, and automatic conversion of existing equations is difficult, none of the equations in the \gls*{dlmf} use this macro.
-As a consequence, the translator has some simple rules to perform translations without explicitly marked translations with \verb|\idot|.
+As a consequence, the translator has some simple
+rules for performing translations without explicitly marking multiplication translations with \verb|\idot|.
 
@@
 Another more questionable translation decision is alphanumerical expressions.
-As explained in~\Cref{tab:allTypesTable}, the \gls*{pom}-tagger handles strings of letters and numbers differently, depending on the order of the symbols.
-The reason is that an expression such as `$4b$' is usually considered to be a multiplication of $4$ and `$b$,' while `$b4$' looks like indexing `$b$' by $4$.
+As explained in~\Cref{tab:allTypesTable}, the \gls*{pom}-tagger handles strings of letters and numbers differently depending on the order of the symbols.
+The reason is that an expression such as `$4b$' is usually considered to be a multiplication of $4$ and `$b$,' while `$b4$'
+displays like indexing `$b$' by $4$.
 While the first example produces two nodes, namely $4$ and `$b$', the second example `$b4$' produces just a single alphanumerical node in the \gls*{pom-pt}.
-The translator interprets alphanumerical expressions as multiplications for two reasons: (1) we would assume that the inputs `$4b$' and `$b4$' are mathematically equivalent; and (2) it is more common in mathematics to use single letter names for variables~\parencite{Notation:History}.
-Therefore we make the following definitions
+The translator interprets alphanumerical expressions as multiplications for two reasons:~(1) we would assume that the inputs `$4b$' and `$b4$' are mathematically equivalent; and (2) it is more common in mathematics to use single letter names for variables~\parencite{Notation:History}.
+Therefore we have used rules as follows {\bf [HSC: I changed the wording here, definition$\to$rules.
+Is this ok?
+Otherwise further clarify.]}
 
@@
 
-In general, the translator is designed to find a work-around for disambiguating expressions. Only if there is no way to solve the ambiguity with the defined rules, the translation process stops.
+In general, the translator is designed to find a work-around for disambiguating expressions. 
+If there is no way to solve a potential ambiguity with defined rules, then we stop the translation process
+{\bf [HSC: I changed the wording here also. Is this ok?]}.
 
@@
 \gls*{cas} define their own syntax parser.
-We were able to use \Maple's internal data structure to get a parse tree of the input.
-\Cref{sec:forward-translation} and \Cref{sec:backward-translation} will explain the parsing and translation process in detail.
-
-All translations are defined by a library (\gls*{csv} and \gls*{json} files) that defines translation patterns for each function and symbol.
-The pattern uses \verb|$i| as a placeholder to define the positions of the arguments.
+We were able to use \Maple's internal data structure to
+obtain {\bf [HSC: I changed get$\to$obtain.]} a parse tree of the input. \Cref{sec:forward-translation} and \Cref{sec:backward-translation} will explain the parsing and translation process in detail.
+
+All translations are defined by a library (\gls*{csv} and \gls*{json} files) that define translation patterns for each function and symbol.
+The pattern uses \verb|$i| as a placeholder used to describe {\bf [HSC: I used describe instead of define.
+Is this ok?
+Otherwise use a different more descriptive term.]} the positions of the arguments.
 For example, the translation patterns for the Jacobi polynomial are illustrated in~\Cref{tab:placeholder_ex2}.
@@
 \subsection{Escape the Placeholder Symbol}
-The used placeholders cause trouble when the \gls*{cas} uses the symbol \verb|$| for other reasons, e.g., the differentiation in \Maple{} is defined by
+The used placeholders cause trouble when the \gls*{cas} uses the symbol \verb|$| for other reasons, e.g., differentiation in \Maple{} is implemented as
 \begin{equation*}
@@
 \end{equation*}
-where $f$ is an algebraic expression or an equation, $x$ is the name of the differentiation variable, and $n$ is an integer representing the $n$-th order differentiation\footnote{\url{https://www.maplesoft.com/support/help/maple/view.aspx?path=diff}, seen 07/2018}.
-A translation for $\frac{d^2x^2}{dx^2}$ should look like this
+where \verb|f| is an algebraic expression or an equation, \verb|x| is the name of the differentiation variable, 
+and \verb|n| is an integer representing the $n$-th order differentiation\footnote{\url{https://www.maplesoft.com/support/help/maple/view.aspx?path=diff}, seen 07/2018}.
+A translation for $\frac{d^2x^2}{dx^2}$ should display as
 \begin{equation*}
@@
 As a pre-processing step, we use the \gls*{pom}-Tagger~\parencite{POM-Tagger}\footnote{Named according to the Part-of-Speech-Taggers in \gls*{nlp}.} for parsing semantic \LaTeX{} expressions.
-The \gls*{pom}-Tagger is defined by a context-free grammar in \gls*{bnf} and is an \texttt{LL}-Parser, i.e., it parses the input from \textbf{L}eft to right and assigns the \textbf{L}eftmost (first applicable) drivation rule defined by the grammar to an expression.
+The \gls*{pom}-Tagger is defined by a context-free grammar in \gls*{bnf} and is an \texttt{LL}-Parser, i.e., it parses the input from \textbf{L}eft to right and assigns the \textbf{L}eftmost (first applicable) derivation rule defined by the grammar to an expression.
 In other words, the \gls*{pom}-Tagger scans the input for \textit{terms} and groups them into subexpressions if suitable, where \textit{terms} are non-terminal symbols in the context of \gls*{bnf}.
-If a node in the generated parse tree matches a pre-defined symbol, it will be tagged by mata information that are defined in manually cultivated lexicon files.
+If a node in the generated parse tree matches a pre-defined symbol, it will be tagged by mata {\bf [HSC: mata?
+If this is correct, please define, otherwise clarify.]} information that are defined in manually cultivated lexicon files {\bf [HSC: please reword this sentence because it reads poorly.]}.
 
 We integrated the defined translation patterns from our library also into these lexicon files.
-Thus the tagger also tags a node in the parse tree by its translation patterns.
+The tagger also tags a node in the parse tree by its translation patterns.
 \Cref{tab:sine-lex-example} gives an example of an entry of the lexicon file.
@@
 	\hline
-	\multicolumn{3}{l}{Symbol: \texttt{\textbackslash sin}} \\
-	\! & \multicolumn{2}{l}{Feature Set: dlmf-macro} \\
-	\! & \! & DLMF: \verb|\sin@@{z}|\\
-	\! & \! & DLMF-Link: dlmf.nist.gov/4.14.E1\\
-	\! & \! & Meanings: Sine\\
-	\! & \! & Number of Parameters: 0\\
-	\! & \! & Number of Variables: 1\\
-	\! & \! & Number of Ats: 2\\
-	\! & \! & Maple: \verb|sin($0)|\\
-	\! & \! & Maple-Link: www.maplesoft.com/support/\\
+	\multicolumn{3}{l}{Symbol:~\texttt{\textbackslash sin}} \\
+	\! & \multicolumn{2}{l}{Feature Set:~dlmf-macro} \\
+	\! & \! & DLMF:~\verb|\sin@@{z}|\\
+	\! & \! & DLMF-Link:~dlmf.nist.gov/4.14.E1\\
+	\! & \! & Meanings:~Sine\\
+	\! & \! & Number of Parameters:~0\\
+	\! & \! & Number of Variables:~1\\
+	\! & \! & Number of Ats:~2\\
+	\! & \! & Maple:~\verb|sin($0)|\\
+	\! & \! & Maple-Link:~www.maplesoft.com/support/\\
 	\! & \! & \hspace{32pt} help/maple/view.aspx?path=sin\\
-	\! & \! & Mathematica: \verb|Sin[$0]|\\
-	\! & \! & Mathematica-Link: reference.wolfram.com/\\
+	\! & \! & Mathematica:~\verb|Sin[$0]|\\
+	\! & \! & Mathematica-Link:~reference.wolfram.com/\\
 	\! & \! & \hspace{32pt} language/ref/Sin.html\\
@@
 The \gls*{pom} project aims to disambiguate mathematical \LaTeX{} expressions and generates an expression tree.
-However, in the current state, many expressions cannot be disambiguated yet.
+However, in the current state, many expressions cannot yet be disambiguated.
 Therefore, the \gls*{pom}-tagger generates a raw parsed tree where each token in the \LaTeX{} expression is a node in the tree.
@@
 \Cref{fig:syntax-tree-usecase} visualizes the \gls*{pom-pt} of the Jacobi polynomial example from~\Cref{tab:JacobiP-usecase}.
-Because of the differences to expression trees, a backward conversion of the \gls*{pom-pt} to a string representation can be difficult, especially for finding necessary or unnecessary parentheses.
+Because of the differences to {\bf [HSC: to?
+does not sound right?
+Pleased use a better word.
+between?
+Also, clean up the sentence.
+It does not read so well.]} expression trees, a backward conversion of the \gls*{pom-pt} to a string representation can be difficult, especially for finding necessary or unnecessary parentheses.
 Therefore we create the \gls*{teo}.
@@
 If the node is not a leaf, it starts to translate all children of the node recursively.
-This idea seems to be a practical and elegant solution but it has a significant drawback, it cannot be used to translate functions.
+This idea seems to be a practical and elegant solution.
+However, it has a significant drawback.
+It cannot be used to translate functions.
 Since the arguments of functions are following siblings in the \gls*{pom-pt}, the algorithm needs to look ahead when a leaf is a known function, e.g., in the case of a semantic macro with arguments (see~\Cref{fig:syntax-tree-usecase}).
@@
 \begin{wrapfigure}[10]{r}{0.38\textwidth} % optional parameter gives "# of narrow lines", very useful! 
-\vspace{-8pt}
+\vspace{4pt}
 \begin{minipage}{0.38\textwidth}
@@
 
-If we take a closer look to mathematical notations we discover many different types of notations used to represent formulae.
+If we take a closer look at mathematical notations we discover many different types of notations used to represent formulae.
 \Cref{tab:notations} illustrates the expression $(a+b)x$ in different notations.
@@
 The infix notation is commonly used in arithmetic and places the operator between its operands.
-This only makes sense as long as the operator is a binary operator.
+This only makes sense if the operator is a binary operator.
 
@@
 Of course, `$-$' can also be the binary operator for subtraction, which is commonly used in infix notation.
-An example for the postfix notation is factorial, such as `$2!$'.
-
-Most programming languages (and \gls*{cas} as well) internally use prefix or postfix notation and do not mix the notations in one expression, since it is more convenient to parse expressions in uniform notations.
+One example is the postfix notation used with factorials, such as for `$2!$'
+{\bf [HSC: I changed the wording slightly here. Is this ok?]}.
+
+Since it is more convenient to parse expressions using uniform notations, most programming 
+languages (and \gls*{cas} as well) internally use prefix or postfix notation and do not mix 
+the notations in one expression {\bf [HSC: I reworded this sentence. Is it ok?]}.
 However, the common practice in science is to use mixed notations in expressions.
-Since the \gls*{pom} has rarely implemented mathematical grammatical rules yet, it takes the input as it is and does not build an expression tree.
+Since the \gls*{pom} has rarely implemented mathematical grammatical rules, it takes the input as it is and does not build an expression tree.
 Therefore, it parses all four examples from~\Cref{tab:notations} to four different \gls*{pom-pt}s rather than to one unique expression tree.
@@
 On the other hand, \LaTeX{} automatically disambiguates the first example by only recognizing the very next element (single symbols or sequences in curly brackets) for the superscript and therefore displays the second interpretation.
-The second example should not be misinterpreted, since this notation is the standard interpretation in science for the double factorial.
-We wrote the second interpretation with parentheses point it out more precisely.
+The second example should not be misinterpreted since this notation is the standard interpretation in science for the double factorial.
+We wrote the second interpretation with parentheses point it out more precisely {\bf [HSC: please rewrite this sentence.]}.
 However, surprisingly, \Maple{} computes the first interpretation (the factorial of the factorial of $n$) again rather than the common standard interpretation.
@@
 	\hline
-	1: & \rule{0pt}{0.9\normalbaselineskip} $4\ \hat{\ }\ 2!$ & $4^{2!}$ & $4^2!$ \\
-	2: & $n!!$ & $(n!)!$ & $(n)!!$\\
+	1:~& \rule{0pt}{0.9\normalbaselineskip} $4\ \hat{\ }\ 2!$ & $4^{2!}$ & $4^2!$ \\
+	2:~& $n!!$ & $(n!)!$ & $(n)!!$\\
 	\hline
 \end{tabular}
-\caption{Potentially ambiguous examples of the factorial and double factorial function. One expression in a text format can be interpreted in different ways.}
+\caption{Potentially ambiguous examples using the factorial and double factorial symbols. One expression in a text format can potentially be interpreted in different ways.}
 \label{tab:amb_ex}
@@
 
+{\bf [HSC: Is it `The \verb|SequenceTranslator|' or is it `A \verb|SequenceTranslator|'? Please choose. You cannot use both. Once you have decided implement throughout the paper.]}
+
 A \verb|SequenceTranslator| translates the \textit{sequence} and \textit{balanced expressions} in the \gls*{pom-pt}.
-If a node $n$ is a leaf and the represented symbol is an open bracket (parentheses, square brackets and so on) the following nodes are also taken as a \textit{sequence}.
-Hence, combined with the recursive translation approach, the \verb|SequenceTranslator| also checks balancing of parentheses in expressions.
+If a node $n$ is a leaf and the represented symbol is an open bracket (parentheses, square brackets and so on), the following nodes are also taken as a \textit{sequence}.
+Hence, combined with the recursive translation approach, the
+\verb|SequenceTranslator| 
+also checks balancing of parentheses in expressions.
 An expression such as `$(a]$' produces a mismatched parentheses error.
-On the other hand, this is a problem for interval expressions such as `$[a,b)$'.
+On the other hand, this is a problem for real interval expressions such as `$[a,b)$'.
 In the current version, the program cannot distinguish between mismatched parentheses and half-opened, half-closed intervals.
@@
 
-The translation process of the \Macro 's is complex, so there is a special class, the \verb|MacroTranslator|, that handles those nodes in the \gls*{pom-pt}.
+{\bf [HSC: This is a new paragraph, but it does not have standard vertical space above it.
+I do not know why this is.
+Can you fix this?]} The translation process for the \Macro s is complex, so there is a special class, the \verb|MacroTranslator|, that handles those nodes in the \gls*{pom-pt}.
 Algorithm~\ref{alg:macro-translation} explains the \verb|MacroTranslator| without error handling.
@@
 The \verb|MacroTranslator| will be called in Line~\ref{line:transLeaf} in Algorithm~\ref{alg:translation}, since the macro is a leaf node in the \gls*{pom-pt}.
-The following cases describes the different kinds of the following siblings after a semantic macro node.
-Those can be
+The following cases describe the different kinds of the following siblings after a semantic macro node.
+Those can be:
 \begin{itemize}
@@
 Following siblings after the macro was translated (with all arguments) do not belong to the semantic macro.
-If the next node is an exponent the translated macro is the base.
+If the next node is an exponent, the translated macro is the base.
 Table~\ref{tab:multi-expo} shows an example for the translation of the trigonometric cosine function with multiple exponents.
@@
 The translation process from this parse tree then follows the same principle as for the forward translations.
-Instead of writing a custom \Maple{} syntax parser, we use \Maple's internal data structure to get a syntax tree of the input\footnote{Therefore, a license of \Maple{} is mandatory to perform backward translations.
-The translator is using the version \Maple{} 2016.}.
+Instead of writing a custom \Maple{} syntax parser, we use \Maple's internal data structure to get a syntax tree of the input\footnote{A license of \Maple{} is mandatory to perform backward translations.
+Our translator uses the version \Maple{} 2016 {\bf [HSC: I reworded this.
+Is it ok?
+Otherwise fix.]}.}.
 \Maple{} allows several different input styles.
-The \texttt{1D} input is mainly used for programming purposes and therefore also used to perform our translations.
-Internally, \Maple{} uses a \gls*{dag} for syntax trees.
+The \texttt{1D} input is mainly used for programming purposes
+and is also used to perform our translations. Internally, \Maple{} uses a \gls*{dag} for syntax trees.
 
@@
 
-Each node in the \gls*{dag} stores its children and has a header, which defines the type and the length of the node.
+Each node in the \gls*{dag} stores its children and has a header which defines the type and the length of the node.
 Consider the polynomial $x^2+x$.
@@
 \Maple{} evaluates inputs automatically and changes the input into an internal representation.
-This internal representation might look a bit different to the input.
+This internal representation might look {\bf [HSC: can you find a better word than `look'?]} a bit different to the input.
 One example has already been given with \Cref{fig:internal-maple-dag}, where each summand of a sum is stored with a factor.
@@
 
-There are unevaluation quotes implemented to avoid evaluations on input expression. Table~\ref{tab:unevaluation-quotes} gives an example how those unevaluation quotes work.
+There are unevaluation quotes implemented to avoid evaluations on input expressions. Table~\ref{tab:unevaluation-quotes} gives an example how unevaluation quotes work.
 
@@
 \hline
-Input expression: & \texttt{sin(Pi)+2-1} & \texttt{'sin(Pi)+2-1'}\\
-Stored expression: & 1 & \texttt{sin(Pi)+1}\\
+Input expression:~& \texttt{sin(Pi)+2-1} & \texttt{'sin(Pi)+2-1'}\\
+Stored expression:~& 1 & \texttt{sin(Pi)+1}\\
 \hline
@@
 
-Since we want to keep a translated expression similar to the input expression, we implemented some cosmetic rules for the backward translations that solve or reduce the effects from the list of changes above.
+Since we want to keep a translated expression similar to the input expression, we implemented some cosmetic rules 
+for backward translations which solve or reduce the effects due to the list of changes above. 
 \begin{itemize}
@@
 \begin{enumerate}
-\item \textbf{Round Trip Tests}: translates expressions back and forth and analyzes the changes.
-\item\label{approach-2} \textbf{Function Relation Tests (Symbolical)}: translates mathematically proven equivalent expressions from one system to a \gls*{cas} and evaluates whether the relation remains valid via symbolical equivalence checks.
-\item \textbf{Numerical Tests}: takes the same equations from Approach~\ref{approach-2} but evaluates them on specific numerical values to test equivalence.
+\item \textbf{Round Trip Tests}:~translates expressions back and forth and analyzes the changes.
+\item\label{approach-2} \textbf{Function Relation Tests (Symbolical)}:~translates mathematically proven equivalent expressions from one system to a \gls*{cas} and evaluates whether the relation remains valid via symbolical equivalence checks.
+\item \textbf{Numerical Tests}:~takes the same equations from Approach~\ref{approach-2} but evaluates them on specific numerical values to test equivalence.
 \end{enumerate}
@@
 A \textbf{fixed point representation} (or short fixed point) in a round trip translation process is a string representation that is identical to all string representations in the following cycles.
-Table~\ref{tab:fixpoint} illustrates an example of a round trip test which reaches a fixed point for the formula
+Table~\ref{tab:fixpoint} illustrates an example of a round trip test which reaches a fixed point for the mathematical expression
 \begin{equation}
@@
 
-There is currently only one exception known where a round trip test does not reach a fixed point representation: Legendre's incomplete elliptic integrals~\parencite[(19.2.4-7)]{NIST:DLMF} are defined with the amplitude $\phi$ in the first argument in the \DLMF, while \Maple{} takes the trigonometric sine of the amplitude as the first argument.
+There is currently only one exception known where a round trip test does not reach a fixed point representation:~Legendre's incomplete elliptic integrals~\parencite[(19.2.4-7)]{NIST:DLMF} are defined with the amplitude $\phi$ in the first argument in the \DLMF, while \Maple{} takes the trigonometric sine of the amplitude as the first argument.
 Therefore, the forward and backward translations are defined as
@@
 The round trip tests are very successful, but they only detect errors in string representations.
-However, because of the simplification techniques of fixed points, we are able to at least detect logical errors in one system: \Maple.
+However, because of the simplification techniques of fixed points, we are able to at least detect logical errors in one system:~\Maple.
 On the other hand, these tests cannot determine logical errors in the translations between the two systems.
@@
 \subsection{Function Relation Tests}\label{sec:relation-tests}
-The \gls*{dlmf} is a compendium for special functions and orthogonal polynomials and lists several relations between the functions and polynomials.
+The \gls*{dlmf} is a compendium for special functions and orthogonal polynomials and lists many relations between the functions and polynomials.
 The idea of this evaluation approach is to translate an entire relation and test whether the relation remains valid after performing the translations.
 
-With this technique we can detect translation errors such as in \eqref{eq:wrong-trans-1} and \eqref{eq:wrong-trans-2}.
+With this technique {\bf [HSC: what technique?
+A technique has not been introduced yet in this section.
+Please describe this better, or be more specific.]}, we can detect translation errors such as in \eqref{eq:wrong-trans-1} and \eqref{eq:wrong-trans-2}.
 Consider the \gls*{dlmf} equation for the sine and cosine function~\parencite[(4.21.2)]{NIST:DLMF}
@@
 \end{equation}
-which is wrong. Since the expression is correct before the translation, we conclude an error during the translation process.
-
-However, there are two essential problems with this approach.
-Testing the mathematical equivalence of expressions is hard to solve and \gls*{cas} often have difficulties testing simple equations symbolically.
-For example, equation~\parencite[(4.35.34)]{NIST:DLMF}
+which is wrong. Since the expression is correct before the translation, we conclude that there was an error 
+during the translation process.
+
+However, there are two essential problems with this approach. Testing the mathematical equivalence of expressions 
+is difficult to solve and \gls*{cas} often have difficulties testing simple equations symbolically. For example, 
+consider
+\parencite[(4.35.34)]{NIST:DLMF}
 \begin{equation*}
-\sinh \left( x+\iunit \cdot y \right) = \sinh x \cos y + \iunit \cosh x \sin y,
+\sinh \left( x+\iunit y \right) = \sinh x \cos y + \iunit \cosh x \sin y,
 \end{equation*}
-as a difference of the left- and right-hand side cannot be simplified to zero by default.
-Furthermore, this approach only checks forward translations because there is no way to check equivalence of expressions in \LaTeX{} automatically (again this could become feasible with our translator).
+as a difference of the left- and right-hand sides cannot be simplified to zero by default.
+Furthermore, this approach only checks forward translations because there is no way to check equivalence {\bf [HSC: I thought we weren't using equivalence in this paper?
+Please clarify here and in all other places used in the paper.]} of expressions in \LaTeX{} automatically (again this could become feasible with our translator).
 We use \Maple's \texttt{simplify} function to check if the difference of the left-hand side and the right-hand side of the equation is equal to zero.
 In addition, we use \texttt{simplify} and check if the division of the right-hand side by the left-hand side returns a numerical value or not.
-This simplification function is the most powerful function to check the equivalence in \Maple.
-However, there are several cases where the simplification fails.
+Simplification function is the most powerful function to check the equivalence in \Maple.
+However, there are several cases where simplification fails.
 Because of implementation details, there are some techniques that help \Maple{} to find possible simplifications.
-For example we can force \Maple{} to convert the formula
+For example, we can force \Maple{} to convert the formula
 \begin{equation}
@@
 \subsection{Results}\label{sec:test-summary}
-There are 685 \Macro{}s\footnote{The DLMF/DRMF semantic macros are still a work in progress, and the total number is constantly changing.} in total, and 665 of them were implemented in the translator engine.
+There are currently 685 \Macro{}s\footnote{The DLMF/DRMF semantic macros are still a work in progress, and the total number is constantly changing.} in total, and 665 of them were implemented in the translator engine.
 We defined forward translations to \Maple{} for 201 of the macros and backward translations from \Maple{} for 195 functions.
@@
 The translator was able to translate 2405\footnote{All percentages are approximately calculated.} (58.8\%) of the extracted equations without errors.
-660 (27.4\%) of the successfully translated expressions were verified by the simplification techniques of \Maple.
+Simplification techniques of \Maple were successfully verified for 660 (27.4\%) of the translated 
+expressions {\bf [HSC: I tried to reword this sentence so that it does not start with a number, which it never
+should. Can you check to make sure the meaning has not changed?]}. 
 We applied additional numerical tests for the remaining 1745 equations.
-For 418 (24\%) cases the numerical tests were valid.
+For 418 (24\%) cases {\bf [HSC: cases?
+Can you be more precise?]}, the numerical tests were valid.
 More detailed results for numerical and symbolical tests were presented in~\parencite{NumericalTests:Paper}.
@@
 \begin{equation*}
-\frac{{\mathrm{d}}^{n}}{{\mathrm{d}z}^{n}}\left(e^{z^{2}}\operatorname{erfc}z\right), \quad n = {0,1,2,\ldots}
+\frac{{\mathrm{d}}^{n}}{{\mathrm{d}z}^{n}}\left(e^{z^{2}}\operatorname{erfc}z\right), \quad n = {0,1,2,\ldots},
 \end{equation*}
@@
 The translator concept has proven itself by discovering errors in the online \gls*{dlmf} compendia.
-The test cases have also shown how difficult it is to validate a translated expression and have uncovered the problems of translations between two systems with different sets of supported functions.
-Our validation techniques also assume the correctness of the simplification and computational algorithms in \gls*{cas}.
+The test cases have also shown how difficult it is to validate a translated expression and have uncovered the problems of translations
+{\bf [HSC: `the problems of translations'?
+I do not understand what you mean here.
+Please clarify.]} between two systems with different sets of supported functions.
+Our validation techniques also assume the correctness of simplification and computational algorithms in \gls*{cas}.
 However, combining those techniques and automatically running translation checks, not only can discover errors in mathematical compendia but can also detect errors in simplifications or computations of the \gls*{cas}.
@@
 With such improvements, further work on evaluation techniques become worthwhile to evaluate the \gls*{dlmf} and \gls*{cas}.
+Increasing the amount of translatable formulae in the \gls*{dlmf} and improving the verificiation 
+techniques are also parts of ongoing projects {\bf [HSC: I moved this sentence from the end of the section to the end of this paragraph. Is this ok?]}.
 
 The translator was designed to be easily extendable. This allows one to implement translations for other \gls*{cas} without much effort.
-An extensive weakness is the dependency on the special macros from the \gls*{dlmf}.
+An extensive weakness {\bf [HSC: a weakness?
+No.
+I like the last sentence of this paragraph, but the rest is not good.
+It needs to be repaired, or explained more clearly.
+Even with a generic to semantic converter, one still relies on the semantic macros, or something needs to be explained better.]} is the dependency on the special macros from the \gls*{dlmf}.
 The translator is not able to translate functions without using these macros.
@@
 
-A theoretical improvement for the numerical tests would be to perform the tests for specific, considered as critical, values with respect to the involved functions.
-J.~C.~Beamont et al.~tested  identities for multivalued elementary functions by choosing sample points from a region with respect to the branch cuts of the functions~\parencite{DBLP:journals/aaecc/BeaumontBDP07}.
-We assume that choosing sample points from those regions, significantly improves the results of our numerical tests.
-Increasing the amount of translatable formulae in the \gls*{dlmf} and improving the verificiation techniques are parts of ongoing projects.
-
+Further improvements for numerical tests could be to perform tests for specific (critical) values \parencite{DBLP:journals/aaecc/BeaumontBDP07} with respect to the involved functions.
+Beamont and collaborators tested identities for multivalued elementary functions by choosing sample points from regions with respect to branch cuts for functions.
+Choosing sample points from those regions could significantly improve the success rate of the numerical tests {\bf [HSC: I changed the wording in this paragraph.
+Please check to make sure it is still correct.]}.
 \emergencystretch=1em

### 2018-11-29 ReviewerResponse.txt
Include def for appropriate/inappropriate translations
updated ReviewerResponse.txt

@@
 In summary, we have :
-(1) added subsection headings "Different Sets of Defined Functions", 
+(1) added the subsection headings "Different Sets of Defined Functions", 
 "Positions of Branch Cuts", "Insufficient Semantic Information", and 
-"Ambiguous Expressions" in Section 3;
+"Potentially Ambiguous Expressions" in Section 3; 
 (2) gave an overview of the concept of the translator in Section 4;
-(3) removed unnecessary Algorithm 1 in Section 5 as suggested, because 
-it was just a simplified pre-version of Algorithm 2;
+(3) as suggested, we removed the unnecessary Algorithm 1 in Section 5 
+because it was just a simplified pre-version of Algorithm 2;
 (4) added the subsection heading "Problems with the Look Ahead Approach" in 
@@
 by referring to specific lines in the algorithm.
-We also explained how a power after a macro is handled in more detail and why 
+
+We explained how a power after a macro is handled in more detail and why 
 optional parameters need to be handled differently than other arguments.
 
-We also gave further detailed descriptions concerning the PoM-Tagger, 
-namely its parsing process and lexicon files.
+We gave further detailed descriptions concerning the PoM-Tagger, namely its 
+parsing process and lexicon files.
 
-We also added LaTeXML to the related work section and gave a brief 
+We added LaTeXML to the related work section and gave a brief 
 explanation with cited research (Schubotz et al., 2018, now included 
@@
 
-Also, instead of creating an Appendix, we decided to add short paragraphs 
+Instead of creating an Appendix, we decided to add short paragraphs 
 which explain the purpose of Sections 4 and 6. We hope this helps the reader 
 to decide whether he can skip the section, e.g., in case of his lack of 
-interest in Maple "hackings", or not.  In the new version of the paper, we 
-implemented all suggested minor changes as well as adding the idea of 
+interest in Maple "hackings", or not.  In the new version of the paper, 
+we implemented all suggested minor changes as well as adding the idea of 
 improved numerical tests by using the mentioned cut region detection to 
 the future work section.
+
+To address the weakness in precision, we added an enhanced definition for 
+appropriate and inappropriate translations. Furthermore, we have rewritten
+the paragraph about alternative translations for the arccotangent function 
+on Page 6 (now Page 7).

### 2018-11-30 main.tex
Deleted phrases with 'in consequence' and similar wordings (close #22)

@@
 For a significant percentage, the special function definitions in \Maple{} and the \gls*{dlmf} were different.
-Consequently, an atomic symbol in one system maps to a composite expression in the other system.
+Therefore, an atomic symbol in one system maps to a composite expression in the other system.
 The translator was also successfully used for automatic verification of mathematical online compendia and \gls*{cas}.
@@
 Since this macro is newly introduced by contributers of the \gls*{drmf} team, and automatic conversion of existing equations is difficult, none of the equations in the \gls*{dlmf} use this macro.
-As a consequence, the translator has some simple
+Therefore, the translator has some simple
 rules for performing translations without explicitly marking multiplication translations with \verb|\idot|.

### 2018-11-30 main.tex
Deleted CASs and DPSs and only use CAS and DPS, also for plurals (close #27)

@@
 \newacronym{drmf}{DRMF}{Digital Repository of Mathematical Formulae}
-\newacronym{cas}{CAS}{Computer Algebra System}
+\newacronym{cas}{CAS}{Computer Algebra Systems}
 \newacronym{mlp}{MLP}{Mathematical Language Parser}
@@
 \newacronym{opsf}{OPSF}{Orthogonal Polynomials and Special Functions}
-\newacronym{dps}{DPS}{Document Preparation System}
+\newacronym{dps}{DPS}{Document Preparation Systems}
 \newacronym{stem}{STEM}{Science, Technology, Engineering and Mathematics}
@@
 \noindent
-\textbf{Purpose:} Modern mathematicians and scientists of math-related disciplines often use \glspl*{dps} to write and \glspl*{cas} to calculate mathematical expressions.
+\textbf{Purpose:} Modern mathematicians and scientists of math-related disciplines often use \gls*{dps} to write and \gls*{cas} to calculate mathematical expressions.
 Usually, they translate the expressions manually between \gls*{dps} and \gls*{cas}.
@@
 \section{Introduction}
-A typical workflow of a scientist who writes a scientific publication is to use a \gls*{dps} to write the paper and one or more \glspl*{cas} for verification, analysis and visualization.
+A typical workflow of a scientist who writes a scientific publication is to use \gls*{dps} to write the paper and one or more \gls*{cas} for verification, analysis and visualization.
 Especially in the \gls*{stem} literature, \LaTeX{}

### 2018-11-30 main.tex
Consistently marked appropriate, inappropriate, and direct in italic (close #28)

@@
 Together with the \verb|global|-\verb|lexicon| file, the \gls*{pom-pt} will be created by the \gls*{pom}-tagger.
-The \verb|latex-converter| takes a string representation of a semantic \LaTeX{} expression and uses the \gls*{pom} engine as well as our \verb|Translator| to create an appropriate string representation for a specified \gls*{cas}.}
+The \verb|latex-converter| takes a string representation of a semantic \LaTeX{} expression and uses the \gls*{pom} engine as well as our \verb|Translator| to create a proper string representation for a specified \gls*{cas}.}
 

### 2018-11-30 main.tex
Try to avoid informal words (addresses #29)

@@
 Especially in the \gls*{stem} literature, \LaTeX{}
-%\footnote{Note that technically \LaTeX{} is not a \gls*{dps} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}.
-However, since \LaTeX{} (an extension of \TeX) is the de facto standard for writing articles in \gls*{stem}, and we only focus on mathematical writing in this paper, we have categorized \LaTeX{} as a \gls*{dps} as well.}
 has become the de facto standard for writing scientific publications over the past 30 years~\parencites{Knuth}[559]{DigitalTypo}{LATEX:Standard}.
@@
 As explained in~\Cref{tab:allTypesTable}, the \gls*{pom}-tagger handles strings of letters and numbers differently depending on the order of the symbols.
-The reason is that an expression such as `$4b$' is usually considered to be a multiplication of $4$ and `$b$,' while `$b4$'
-displays like indexing `$b$' by $4$.
+The reason is that an expression such as `$4b$' is usually considered to be a multiplication of $4$ and `$b$,' while `$b4$' gives the impression that $4$ indexing `$b$'.
 While the first example produces two nodes, namely $4$ and `$b$', the second example `$b4$' produces just a single alphanumerical node in the \gls*{pom-pt}.
@@
 We were able to use \Maple's internal data structure to
-obtain {\bf [HSC: I changed get$\to$obtain.]} a parse tree of the input. \Cref{sec:forward-translation} and \Cref{sec:backward-translation} will explain the parsing and translation process in detail.
+obtain a parse tree of the input. \Cref{sec:forward-translation} and \Cref{sec:backward-translation} will explain the parsing and translation process in detail.
 
@@
 The \gls*{pom} project aims to disambiguate mathematical \LaTeX{} expressions and generates an expression tree.
-However, in the current state, many expressions cannot yet be disambiguated.
+However, in the current state, many expressions still cannot be disambiguated.
 Therefore, the \gls*{pom}-tagger generates a raw parsed tree where each token in the \LaTeX{} expression is a node in the tree.
@@
 If the node is not a leaf, it starts to translate all children of the node recursively.
-This idea seems to be a practical and elegant solution.
+This idea appears to be a practical and elegant solution.
 However, it has a significant drawback.
 It cannot be used to translate functions.
-Since the arguments of functions are following siblings in the \gls*{pom-pt}, the algorithm needs to look ahead when a leaf is a known function, e.g., in the case of a semantic macro with arguments (see~\Cref{fig:syntax-tree-usecase}).
-Algorithm~\ref{alg:translation} is an improved version with look ahead functionality.
+Since the arguments of functions are following siblings in the \gls*{pom-pt}, the algorithm needs to lookahead when a leaf is a known function, e.g., in the case of a semantic macro with arguments (see~\Cref{fig:syntax-tree-usecase}).
+Algorithm~\ref{alg:translation} is an improved version with lookahead functionality.
 
@@
 
-\subsection{Problems with the Look Ahead Approach}
-The look ahead functionality seems to solve the problems for functions.
-But there is another problem that was not mentioned in \Cref{sec:problems}, because it seems to be not a problem previously.
-In some cases the arguments of a function do not follow but precede the function node.
+\subsection{Problems with the Lookahead Approach}
+The lookahead functionality appears to solve the problems for functions.
+However, there is a problem with the lookahead functionality that \Cref{sec:problems} did not address.
+In some cases, the arguments of a function do not follow but precede the function node.
 
@@
 
-If we take a closer look at mathematical notations we discover many different types of notations used to represent formulae.
+If we intently examine mathematical notations, we discover many different types of notations used to represent formulae.
 \Cref{tab:notations} illustrates the expression $(a+b)x$ in different notations.
@@
 It has no access to previously translated nodes in its current state.
-This problem is solved by the \gls*{teo} that stores and groups translated objects like lists.
+This problem is solved by the \gls*{teo} that stores and groups translated objects as lists.
 This allows one to access the latest translated expression and use it as the argument for the factorial function.
@@
 \item an optional parameter in square brackets right after the macro node or after an exponent (Line~\ref{line:next_optional});
-\item a parameter in curly brackets (a \textit{sequence} node in the \gls*{pom-pt}) if none of the above and no `$@$' symbols were passed yet (Line~\ref{line:trans_paras});
+\item a parameter in curly brackets (a \textit{sequence} node in the \gls*{pom-pt}) if none of the above and no `$@$' symbols were passed (Line~\ref{line:trans_paras});
 \item `$@$' symbols (Line~\ref{line:skip_ats}); or
@@
 \section{\Maple{} to Semantic \LaTeX{} Translator}\label{sec:backward-translation}
-In this section, we will discuss several techniques to get access to the parse tree of \Maple's input.
+In this section, we will discuss several techniques to access the parse tree of \Maple's input.
 The translation process from this parse tree then follows the same principle as for the forward translations.
-Instead of writing a custom \Maple{} syntax parser, we use \Maple's internal data structure to get a syntax tree of the input\footnote{A license of \Maple{} is mandatory to perform backward translations.
+Instead of writing a custom \Maple{} syntax parser, we use \Maple's internal data structure to obtain the syntax tree of the input\footnote{A license of \Maple{} is mandatory to perform backward translations.
 Our translator uses the version \Maple{} 2016 {\bf [HSC: I reworded this.
@@
 \Maple{} evaluates inputs automatically and changes the input into an internal representation.
-This internal representation might look {\bf [HSC: can you find a better word than `look'?]} a bit different to the input.
+This internal representation may differ to the input.
 One example has already been given with \Cref{fig:internal-maple-dag}, where each summand of a sum is stored with a factor.
@@
 Increasing the amount of translatable formulae in the \gls*{dlmf} and improving the verificiation 
-techniques are also parts of ongoing projects {\bf [HSC: I moved this sentence from the end of the section to the end of this paragraph. Is this ok?]}.
+techniques are also parts of ongoing projects.
 

### 2018-11-30 main.tex
Editing all HSC comments (close #30)

@@
 we introduce the terms \textit{appropriate} and \textit{inappropriate} translations. 
-%{\bf [HSC: I don't understand how {\it equivalent} translations has anything at 
-%all to do with {\it appropriate} or {\it inapppropriate} translations. Please clarify.]}.
 We consider a translation to be {\it appropriate}, when a numerical evaluation returns the 
@@
 Namely, \sTeX{}~\parencite{sTeX} developed by Kohlhase and the \Macro s developed by Miller~\parencite{DLMF:Macros}.
-This paper shows that it is possible to develop a context-free translation tool using the semantic macros introduced by these two projects {\bf [HSC: which two projects?
-Please clarify.]}.
+This paper shows that it is possible to develop a context-free translation tool using  the semantic macros introduced by Miller.
 The goal of \sTeX{} is to markup the functional structure of mathematical documents so that they can be exported to the \gls*{omdoc} format.
@@
 Since translations such as these are nonintuitive, describing explanations become necessary for the translation process.
-A special logging function takes care {\bf [HSC: ``takes care'' is too informal.
-Use a different phrase or word.]} of each translation and provides details after a successful translation process.
+A particular logging function stores each translation and provides details after a successful translation process.
 \Cref{sec:forward-translation} explains this task further.
@@
 This example demonstrates that the \Macro s are still a work in progress and further updates are sometimes
-necessary in order to further encapsulate critical semantic information {\bf [HSC: I changed this sentence. Check to make sure that it is still correct.]}.
+necessary in order to further encapsulate critical semantic information.
 
 \subsection{Potentially Ambiguous Expressions}
-
-{\bf [HSC: This section needs to start out with a short overview paragraph describing the issue instead of an example. Also, I changed the title of the section.]}
-
-A similar problem is multiplications since they are rarely explicitly marked in \LaTeX{} expressions, e.g., scientists using whitespace to indicate multiplications rather than using \verb|\cdot| or similar symbols.
+Since the \Macro s aims to cover an extensive set of special functions, orthogonal polynomials, and mathematical constants, they does not contain specific macros for other mathematical objects.
+However, also mathematical expression without functions, polynomials and mathematical constants can be ambigious.
+As an example, multiplications are rarely explicitly marked in \LaTeX{} expressions, e.g., scientists using whitespace to indicate multiplications rather than using \verb|\cdot| or similar symbols.
+But whitespaces can also be used to improve the readability and not to represent a multiplication.
+
 For such problems, we introduced a new macro \verb|\idot| for an invisible multiplication symbol (this macro will not be rendered).
@@
 
-Another more questionable translation decision is alphanumerical expressions.
+Another more questionable translation decision addresses alphanumerical expressions.
 As explained in~\Cref{tab:allTypesTable}, the \gls*{pom}-tagger handles strings of letters and numbers differently depending on the order of the symbols.
@@
 The translator interprets alphanumerical expressions as multiplications for two reasons:~(1) we would assume that the inputs `$4b$' and `$b4$' are mathematically equivalent; and (2) it is more common in mathematics to use single letter names for variables~\parencite{Notation:History}.
-Therefore we have used rules as follows {\bf [HSC: I changed the wording here, definition$\to$rules.
-Is this ok?
-Otherwise further clarify.]}
+Therefore we have used rules as follows
 
@@
 In general, the translator is designed to find a work-around for disambiguating expressions. 
-If there is no way to solve a potential ambiguity with defined rules, then we stop the translation process
-{\bf [HSC: I changed the wording here also. Is this ok?]}.
+If there is no way to solve a potential ambiguity with defined rules, then we stop the translation process.
 
@@
 All translations are defined by a library (\gls*{csv} and \gls*{json} files) that define translation patterns for each function and symbol.
-The pattern uses \verb|$i| as a placeholder used to describe {\bf [HSC: I used describe instead of define.
-Is this ok?
-Otherwise use a different more descriptive term.]} the positions of the arguments.
+The pattern uses \verb|$i| as a placeholder to determine the positions of the arguments.
 For example, the translation patterns for the Jacobi polynomial are illustrated in~\Cref{tab:placeholder_ex2}.
@@
 In other words, the \gls*{pom}-Tagger scans the input for \textit{terms} and groups them into subexpressions if suitable, where \textit{terms} are non-terminal symbols in the context of \gls*{bnf}.
-If a node in the generated parse tree matches a pre-defined symbol, it will be tagged by mata {\bf [HSC: mata?
-If this is correct, please define, otherwise clarify.]} information that are defined in manually cultivated lexicon files {\bf [HSC: please reword this sentence because it reads poorly.]}.
+A node in the generated parse tree will be tagged by meta information if the node matches defined symbols.
+The meta information is stored in lexicon files.
+Those lexicon files were manually cultivated for the \gls*{pom}-Tagger.
 
@@
 \Cref{fig:syntax-tree-usecase} visualizes the \gls*{pom-pt} of the Jacobi polynomial example from~\Cref{tab:JacobiP-usecase}.
-Because of the differences to {\bf [HSC: to?
-does not sound right?
-Pleased use a better word.
-between?
-Also, clean up the sentence.
-It does not read so well.]} expression trees, a backward conversion of the \gls*{pom-pt} to a string representation can be difficult, especially for finding necessary or unnecessary parentheses.
+Because of the differences between expression trees and \gls*{pom-pt}, it can be difficult to generate a string representation after a successful translation process.
+It is especially difficult to determine necessary and unnecessary parentheses when we analyse the \gls*{pom-pt}.
 Therefore we create the \gls*{teo}.
@@
 In mathematical expressions, notations are mostly mixed, depending on the case and number of operands.
-For example, infix notation is common for binary operators ($+$, $-$, $\cdot$, $\mod$, etc.), while functional notations are conveniently used for any kind of functions ($\sin$, $\cos$, etc.).
+For example, infix notation is common for binary operators ($+$, $-$, $\cdot$, $\mod$, etc.), while functional notations are conveniently used for any kind of functions ($\sin$, $\cos$, etc.), and the postfix notation is often common for unary operators ($2!$, $-2$, etc.).
 Sometimes the same symbol is used in different notations to distinguish different meanings.
@@
 Of course, `$-$' can also be the binary operator for subtraction, which is commonly used in infix notation.
-One example is the postfix notation used with factorials, such as for `$2!$'
-{\bf [HSC: I changed the wording slightly here. Is this ok?]}.
+% [AGP: I moved the 2! example to the first sentence where I explain all kinds of notations
 
@@
 languages (and \gls*{cas} as well) internally use prefix or postfix notation and do not mix 
-the notations in one expression {\bf [HSC: I reworded this sentence. Is it ok?]}.
+the notations in one expression.
 However, the common practice in science is to use mixed notations in expressions.
@@
 The second example should not be misinterpreted since this notation is the standard interpretation in science for the double factorial.
-We wrote the second interpretation with parentheses point it out more precisely {\bf [HSC: please rewrite this sentence.]}.
+We wrote the second interpretation with parentheses for pointing out that we mean the double factorial in this case.
 However, surprisingly, \Maple{} computes the first interpretation (the factorial of the factorial of $n$) again rather than the common standard interpretation.
@@
 \subsection{Subtranslators}\label{sec:subtranslators}
-
-{\bf [HSC: Is it `The \verb|SequenceTranslator|' or is it `A \verb|SequenceTranslator|'? Please choose. You cannot use both. Once you have decided implement throughout the paper.]}
-
-A \verb|SequenceTranslator| translates the \textit{sequence} and \textit{balanced expressions} in the \gls*{pom-pt}.
+The \verb|SequenceTranslator| translates the \textit{sequence} and \textit{balanced expressions} in the \gls*{pom-pt}.
 If a node $n$ is a leaf and the represented symbol is an open bracket (parentheses, square brackets and so on), the following nodes are also taken as a \textit{sequence}.
@@
 \end{algorithm}
-\vspace*{-0.2cm}
-
-{\bf [HSC: This is a new paragraph, but it does not have standard vertical space above it.
-I do not know why this is.
-Can you fix this?]} The translation process for the \Macro s is complex, so there is a special class, the \verb|MacroTranslator|, that handles those nodes in the \gls*{pom-pt}.
+
+The translation process for the \Macro s is complex, so there is a special class, the \verb|MacroTranslator|, that handles those nodes in the \gls*{pom-pt}.
 Algorithm~\ref{alg:macro-translation} explains the \verb|MacroTranslator| without error handling.
@@
 Instead of writing a custom \Maple{} syntax parser, we use \Maple's internal data structure to obtain the syntax tree of the input\footnote{A license of \Maple{} is mandatory to perform backward translations.
-Our translator uses the version \Maple{} 2016 {\bf [HSC: I reworded this.
-Is it ok?
-Otherwise fix.]}.}.
+Our translator uses the version \Maple{} 2016.}.
 \Maple{} allows several different input styles.
@@
 \item\label{approach-2} \textbf{Function Relation Tests (Symbolical)}:~translates mathematically proven equivalent expressions from one system to a \gls*{cas} and evaluates whether the relation remains valid via symbolical equivalence checks.
-\item \textbf{Numerical Tests}:~takes the same equations from Approach~\ref{approach-2} but evaluates them on specific numerical values to test equivalence.
+\item \textbf{Numerical Tests}:~takes the same equations from Approach~\ref{approach-2} but evaluates them on specific numerical values to test whether the translation was {\it appropriate}.
 \end{enumerate}
@@
 
-With this technique {\bf [HSC: what technique?
-A technique has not been introduced yet in this section.
-Please describe this better, or be more specific.]}, we can detect translation errors such as in \eqref{eq:wrong-trans-1} and \eqref{eq:wrong-trans-2}.
+With this idea, we can detect {\it inappropriate} translations such as in \eqref{eq:wrong-trans-1} and \eqref{eq:wrong-trans-2}.
 Consider the \gls*{dlmf} equation for the sine and cosine function~\parencite[(4.21.2)]{NIST:DLMF}
@@
 which is wrong. Since the expression is correct before the translation, we conclude that there was an error 
-during the translation process.
-
-However, there are two essential problems with this approach. Testing the mathematical equivalence of expressions 
-is difficult to solve and \gls*{cas} often have difficulties testing simple equations symbolically. For example, 
+during the translation process and our defined translations were {\it inappropriate}.
+
+However, there are two essential problems with this approach.
+Testing whether expressions are {\it appropriate} representations of each other is a challenging task for \gls*{cas} and they often have difficulties testing simple equations symbolically.
+For example,
 consider
@@
 as a difference of the left- and right-hand sides cannot be simplified to zero by default.
-Furthermore, this approach only checks forward translations because there is no way to check equivalence {\bf [HSC: I thought we weren't using equivalence in this paper?
-Please clarify here and in all other places used in the paper.]} of expressions in \LaTeX{} automatically (again this could become feasible with our translator).
+Furthermore, this approach only checks forward translations because there is no way to automatically check whether two \LaTeX{} expressions are {\it appropriate} or {\it inappropriate} representations of each other (again this could become feasible with our translator).
 We use \Maple's \texttt{simplify} function to check if the difference of the left-hand side and the right-hand side of the equation is equal to zero.
 In addition, we use \texttt{simplify} and check if the division of the right-hand side by the left-hand side returns a numerical value or not.
-Simplification function is the most powerful function to check the equivalence in \Maple.
+The simplification function is the most powerful function to check whether two expressions are {\it appropriate} representations in \Maple.
 However, there are several cases where simplification fails.
@@
 \end{equation}
-to an equivalent representation using their exponential representations, namely
+to an {\it appropriate} representation using their exponential representations, namely
 \begin{equation}
@@
 is incorrect if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ and one should instead use the analytic continuation for the right-hand side of (\ref{eq:branch-cut-near}). 
-To evaluate such complex cases, the equivalence checks of \gls*{cas} are insufficient. Therefore we implement numerical tests as an additional step.
+To evaluate such complex cases, the previous checks for {\it appropriate} representations in \gls*{cas} are insufficient. Therefore we implement numerical tests as an additional step.
 
@@
 While this evaluation is very powerful, it has a significant problem.
-Even when all tested values return zero, it does not prove the equivalence of~\eqref{eq:branch-cut-near}.
+Even when all tested values return zero, it does not prove that~\eqref{eq:branch-cut-near} was {\it appropriately} translated.
 However, when the values are different from zero, it does indicate that there might be an error satisfying one of the four cases~\parencite{NumericalTests:Paper}:
@@
 \item the numerical engine tests invalid combinations of values;
-\item the translation is incorrect;
+\item the translation was {\it inappropriately} defined;
 \item there may be an error in the \gls*{dlmf} source; or
@@
 The translator was able to translate 2405\footnote{All percentages are approximately calculated.} (58.8\%) of the extracted equations without errors.
-Simplification techniques of \Maple were successfully verified for 660 (27.4\%) of the translated 
-expressions {\bf [HSC: I tried to reword this sentence so that it does not start with a number, which it never
-should. Can you check to make sure the meaning has not changed?]}. 
+Simplification techniques of \Maple{} were successfully verified for 660 (27.4\%) of the translated expressions. 
 We applied additional numerical tests for the remaining 1745 equations.
-For 418 (24\%) cases {\bf [HSC: cases?
-Can you be more precise?]}, the numerical tests were valid.
+For 418 (24\%) of them, the numerical tests were valid.
 More detailed results for numerical and symbolical tests were presented in~\parencite{NumericalTests:Paper}.
@@
 \section{Conclusion \& Future Work}\label{ch:conc-future-work}
-The translator concept has proven itself by discovering errors in the online \gls*{dlmf} compendia.
-The test cases have also shown how difficult it is to validate a translated expression and have uncovered the problems of translations
-{\bf [HSC: `the problems of translations'?
-I do not understand what you mean here.
-Please clarify.]} between two systems with different sets of supported functions.
+During this project we uncovered several problems that needs to be solved before providing a translation of mathematical expressions between two systems.
+The translator concept has proven itself by discovering errors in the online \gls*{dlmf} compendia and the test cases have also shown how difficult it is to validate translated expression.
 Our validation techniques also assume the correctness of simplification and computational algorithms in \gls*{cas}.
@@
 The translator was designed to be easily extendable. This allows one to implement translations for other \gls*{cas} without much effort.
-An extensive weakness {\bf [HSC: a weakness?
-No.
-I like the last sentence of this paragraph, but the rest is not good.
-It needs to be repaired, or explained more clearly.
-Even with a generic to semantic converter, one still relies on the semantic macros, or something needs to be explained better.]} is the dependency on the special macros from the \gls*{dlmf}.
-The translator is not able to translate functions without using these macros.
+An extensive weakness of the translator is that the input needs to be semantically enhanced, in this case by the special macros from the \gls*{dlmf}.
+Therefore, without exclusively given semantic information, the translator is not able to translate functions.
 Currently, we are working on mathematical information retrieval techniques which will allow for an extension of the translator to generic \LaTeX{} inputs.
@@
 Beamont and collaborators tested identities for multivalued elementary functions by choosing sample points from regions with respect to branch cuts for functions.
-Choosing sample points from those regions could significantly improve the success rate of the numerical tests {\bf [HSC: I changed the wording in this paragraph.
-Please check to make sure it is still correct.]}.
+Choosing sample points from those regions could significantly improve the success rate of the numerical tests.
 \emergencystretch=1em

### 2018-11-30 main.tex
Fixed Fig4 (close #21)
Fix get
Closes #29
Reword comment on generic LaTeX
Close #32

@@
 Otherwise the \gls*{cas} wouldn't be able to compute or evaluate the expressions.
-During the export process, the semantic information gets lost, because generic \LaTeX{} is not able to carry sufficient semantic information.
+During the export process, the semantic information is lost, because generic \LaTeX{} is not able to carry sufficient semantic information.
 Because of these problems, an exported expression cannot be imported to the same system again in most cases (except for simple expressions such as those described above).
@@
 
-\begin{wrapfigure}[10]{r}{0.38\textwidth} % optional parameter gives "# of narrow lines", very useful! 
-\vspace{4pt}
-\begin{minipage}{0.38\textwidth}
-\center
-\begin{tabular}{lc}
-\hline
-Notation & Expression \\
-\hline
-Infix & $(a+b) \cdot x$\\
-Prefix & $\cdot + a\ b\ x$\\
-Postfix & $a\ b + x\ \cdot$\\
-Functional & $\cdot(+(a, b), x)$\\
-\hline
-\end{tabular}
-%\vspace{-5pt}
-\caption{The mathematical expression `$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
-\label{tab:notations}
-\end{minipage}
-%\vspace{-5pt}
-\end{wrapfigure}
-
 If we intently examine mathematical notations, we discover many different types of notations used to represent formulae.
@@
 This only makes sense if the operator is a binary operator.
+
+\begin{wrapfigure}[10]{r}{0.38\textwidth} % optional parameter gives "# of narrow lines", very useful! 
+\vspace{-5pt}
+\begin{minipage}{0.38\textwidth}
+\center
+\begin{tabular}{lc}
+\hline
+Notation & Expression \\
+\hline
+Infix & $(a+b) \cdot x$\\
+Prefix & $\cdot + a\ b\ x$\\
+Postfix & $a\ b + x\ \cdot$\\
+Functional & $\cdot(+(a, b), x)$\\
+\hline
+\end{tabular}
+%\vspace{-5pt}
+\caption{The mathematical expression `$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
+\label{tab:notations}
+\end{minipage}
+\vspace{3pt}
+\end{wrapfigure}
 
@@
 The translator was designed to be easily extendable. This allows one to implement translations for other \gls*{cas} without much effort.
-An extensive weakness of the translator is that the input needs to be semantically enhanced, in this case by the special macros from the \gls*{dlmf}.
+However, most \LaTeX{} sources, such as arXiv, are given in generic \LaTeX{}.
+Semantic \LaTeX{}, which is a prerequsite for our translator, is currently prevalent in the DLMF and DRMF projects alone.
 Therefore, without exclusively given semantic information, the translator is not able to translate functions.

### 2018-12-03 main.tex
minor wording change
Fix 'hence' in text (see #22)

@@
 
-Hence, a \gls*{cas} user needs to fully understand the properties and special definitions (such as the position of branch cuts) in the \gls*{cas} to avoid mistakes during a translation~\parencite{Maple:Cuts}.
-Hence, a manual translation process is not only laborious, but also prone to errors.
+A \gls*{cas} user needs to fully understand the properties and special definitions (such as the position of branch cuts) in the \gls*{cas} to avoid mistakes during a translation~\parencite{Maple:Cuts}.
+A manual translation process is not only laborious, but also prone to errors.
 Note that this general problem has been named as automatic \gls*{p2c} conversion~\parencite{POM-Tagger}.
@@
 Most mathematical objects in one system have a direct counterpart in other systems.
-However, later in the paper, we will explain that a translation from one specific mathematical object to its counterpart in the other system is not always \textit{appropriate}.
+Later in the paper, we will explain that a translation from one specific mathematical object to its counterpart in the other system is not always \textit{appropriate}.
 Also, not every mathematical object has a counterpart in other systems.
@@
 For example, the internal \LaTeX{} macro \verb|\frac| always indicates a fraction.
-However, for more complex expressions, e.g., the Jacobi polynomial in \Cref{tab:JacobiP-usecase}, the import functions fail.
+For more complex expressions, e.g., the Jacobi polynomial in \Cref{tab:JacobiP-usecase}, the import functions fail.
 The second problem appears in the export tools.
@@
 The arctangent function~\eqref{eq:acot-alternatives-1} and the natural logarithm~(\ref{eq:acot-alternatives-2}) have the same positioned branch cuts as in the \gls*{dlmf} and in \Maple.
-Hence, translation~(\ref{eq:acot-alternatives-1}) solves this issue as long as the user does not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z=-\iunit$.
+Therefore, translation~(\ref{eq:acot-alternatives-1}) solves this issue as long as the user does not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z=-\iunit$.
 Note that none of the translations~(\ref{eq:acot-alternatives}-\ref{eq:acot-alternatives-2}) are {\it appropriate}.
@@
 The \gls*{pom} project aims to disambiguate mathematical \LaTeX{} expressions and generates an expression tree.
-However, in the current state, many expressions still cannot be disambiguated.
+In the current state, however, many expressions still cannot be disambiguated.
 Therefore, the \gls*{pom}-tagger generates a raw parsed tree where each token in the \LaTeX{} expression is a node in the tree.
@@
 If the node is not a leaf, it starts to translate all children of the node recursively.
-This idea appears to be a practical and elegant solution.
-However, it has a significant drawback.
+This idea appears to be a practical and elegant solution, but it has a significant drawback.
 It cannot be used to translate functions.
@@
 the notations in one expression.
-However, the common practice in science is to use mixed notations in expressions.
+The common practice in science is to use mixed notations in expressions.
 Since the \gls*{pom} has rarely implemented mathematical grammatical rules, it takes the input as it is and does not build an expression tree.
@@
 If a node $n$ is a leaf and the represented symbol is an open bracket (parentheses, square brackets and so on), the following nodes are also taken as a \textit{sequence}.
-Hence, combined with the recursive translation approach, the
-\verb|SequenceTranslator| 
-also checks balancing of parentheses in expressions.
+Combined with the recursive translation approach, the \verb|SequenceTranslator| also checks balancing of parentheses in expressions.
 An expression such as `$(a]$' produces a mismatched parentheses error.
@@
 
-However, there are two essential problems with this approach.
+There are two essential problems with this approach.
 Testing whether expressions are {\it appropriate} representations of each other is a challenging task for \gls*{cas} and they often have difficulties testing simple equations symbolically.
@@
 If $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ the value of the right-hand side of~\eqref{eq:branch-cut-near} would be no longer on the principal branch.
-However, \Maple{} will still compute the principal values independently of the value of $z$.
-Hence, a translation
+\Maple{} will still compute the principal values independently of the value of $z$ and the translation
 \begin{equation}
@@
 \end{equation}
-is incorrect if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$ and one should instead use the analytic continuation for the right-hand side of (\ref{eq:branch-cut-near}). 
+is {\it inappropriate} if $|\ph{z}| \in \left(\frac{\cpi}{2}, \cpi\right)$. One should instead use the analytic continuation for the right-hand side of (\ref{eq:branch-cut-near}). 
 To evaluate such complex cases, the previous checks for {\it appropriate} representations in \gls*{cas} are insufficient. Therefore we implement numerical tests as an additional step.
@@
 Even when all tested values return zero, it does not prove that~\eqref{eq:branch-cut-near} was {\it appropriately} translated.
-However, when the values are different from zero, it does indicate that there might be an error satisfying one of the four cases~\parencite{NumericalTests:Paper}:
+When the values are different from zero, it does indicate that there might be an error satisfying one of the four cases~\parencite{NumericalTests:Paper}:
 \begin{enumerate}
@@
 The translator was designed to be easily extendable. This allows one to implement translations for other \gls*{cas} without much effort.
-However, most \LaTeX{} sources, such as arXiv, are given in generic \LaTeX{}.
+However, most \LaTeX{} sources, such as in arXiv, are given in generic \LaTeX{}.
 Semantic \LaTeX{}, which is a prerequsite for our translator, is currently prevalent in the DLMF and DRMF projects alone.

### 2018-12-03 ReviewerResponse.txt
Minor changes to RewieversResponse

@@
-In the following we respond specifically to the reviewers comments.
+In the following, we respond specifically to the reviewers' comments.
 
@@
 comparison, we can only compare translations without function calls or 
-implement a solution which is able overcome these problems.  Because of 
+implement a solution which is able to overcome these problems.  Because of 
 the problems which would need to be overcome, a full solution translation 

### 2018-12-03 main.tex
Make the paper anonym again
Edit keywords and fill out article title page

@@
 
-\input{authors.tex}
+\author{Anonymous Authors}
+%\input{authors.tex}
 
@@
 \noindent
-{\it \bf Keywords:} \LaTeX, Computer Algebra System (CAS), Translation, Presentation to Computation (P2C), Special Functions
+{\it \bf Keywords:} \LaTeX, Computer Algebra System (CAS), Document Preparation System (D2P), Translation, Presentation to Computation (P2C), Special Functions
 
@@
 
+\begin{wrapfigure}[10]{r}{0.38\textwidth} % optional parameter gives "# of narrow lines", very useful! 
+\vspace{-5pt}
+\begin{minipage}{0.38\textwidth}
+\center
+\begin{tabular}{lc}
+\hline
+Notation & Expression \\
+\hline
+Infix & $(a+b) \cdot x$\\
+Prefix & $\cdot + a\ b\ x$\\
+Postfix & $a\ b + x\ \cdot$\\
+Functional & $\cdot(+(a, b), x)$\\
+\hline
+\end{tabular}
+%\vspace{-5pt}
+\caption{The mathematical expression `$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
+\label{tab:notations}
+\end{minipage}
+\vspace{3pt}
+\end{wrapfigure}
+
 If we intently examine mathematical notations, we discover many different types of notations used to represent formulae.
@@
 This only makes sense if the operator is a binary operator.
-
-\begin{wrapfigure}[10]{r}{0.38\textwidth} % optional parameter gives "# of narrow lines", very useful! 
-\vspace{-5pt}
-\begin{minipage}{0.38\textwidth}
-\center
-\begin{tabular}{lc}
-\hline
-Notation & Expression \\
-\hline
-Infix & $(a+b) \cdot x$\\
-Prefix & $\cdot + a\ b\ x$\\
-Postfix & $a\ b + x\ \cdot$\\
-Functional & $\cdot(+(a, b), x)$\\
-\hline
-\end{tabular}
-%\vspace{-5pt}
-\caption{The mathematical expression `$(a+b) \cdot x$' in infix, prefix, postfix and functional notation.}
-\label{tab:notations}
-\end{minipage}
-\vspace{3pt}
-\end{wrapfigure}
 

### 2019-06-27 MMLProblems.tex

\section{Translations via MathML}\label{sec:trans-mml}
The problems to use MathML data as an intermediate format can be demonstrated by using LaTeXML, \Maple, and \Mathematica{} to generate content MathML from their inputs. While the inputs for \gls*{cas} are unique, ideally all systems should produce the same content MathML data. Also, for semantic \LaTeX, LaTeXML should be able to generate the correct content MathML data. As an example we generated content MathML via LaTeXML for
\begin{equation}\label{eq:jac-lml}
\verb|\JacobiP{\alpha}{\beta}{n}@{\cos@{a\Theta}}|,
\end{equation}
and via \Maple{} and \Mathematica{} for the manually verified appropriated  translations of (\ref{eq:jac-lml})
\begin{eqnarray}
&& \verb|JacobiP(n,alpha,beta,cos(a*Theta))|,\text{ and}\label{eq:jac-map}\\
&& \verb|JacobiP[n,\[Alpha],\[Beta],Cos[a \[CapitalTheta]]]|.\label{eq:jac-mat}
\end{eqnarray}
Listings~\ref{lst:latexml}-~\ref{lst:mathematica} shows the generated content MathML. As the listings show, all generated MathML distinguish to each other. The problems are not only flaws in their export implementations but also in the fact that even content MathML is not unique. Based on this fact and the differences in the generated content MathML between each system, we decided to not use MathML as an intermediate format.

\begin{lstlisting}[label={lst:latexml},mathescape=true,caption=The generated content MathML for~(\ref{eq:jac-lml}) using LaTeXML with the command \texttt{latexmlc} and the set of semantic macros.]
<math xmlns="http://www.w3.org/1998/Math/MathML" $\ldots$>
  <semantics>
    <apply>
      <apply>
        <csymbol cd="dlmf">Jacobi-polynomial-P</csymbol>
        <ci>$\alpha$</ci>
        <ci>$\beta$</ci>
        <ci>$n$</ci>
      </apply>
      <ci>$x$</ci>
    </apply>
  </semantics>
</math>
\end{lstlisting}

\vspace{0.4cm}
\begin{lstlisting}[label={lst:maple},mathescape=true,caption=The content MathML for~(\ref{eq:jac-map}) generated by Maple 16 (including manually added linebreaks). We were using the \texttt{MathML[ExportContent]} command to produce the output.]
<math xmlns='http://www.w3.org/1998/Math/MathML'>
  <apply id='id6'>
    <csymbol id='id1' 
      definitionURL='http://www.maplesoft.com/MathML/JacobiP'>
        JacobiP
    </csymbol>
    <ci id='id2'>n</ci>
    <ci id='id3'>alpha</ci>
    <ci id='id4'>beta</ci>
    <ci id='id5'>x</ci>
  </apply>
</math>
\end{lstlisting}

\vspace{0.4cm}
\begin{lstlisting}[label={lst:mathematica},mathescape=true,caption=The content MathML for~(\ref{eq:jac-mat}) generated by Mathematica. We were using the \texttt{ExportString} command to produce the output.]
<math xmlns='http://www.w3.org/1998/Math/MathML'>
 <apply>
  <ci>JacobiP</ci>
  <ci>n</ci>
  <ci>&#945;</ci>
  <ci>&#946;</ci>
  <ci>x</ci>
 </apply>
</math>
\end{lstlisting}
\end{document}

### 2019-06-27 _arxiv.tex

--- 2018-12-03 main.tex
@@
 \documentclass[a4paper,11pt]{article}
-
+\begin{filecontents*}{references.bib}
+	@article{GreinerPetter2019,
+		author={Greiner-Petter, Andre and Schubotz, Moritz and Cohl, Howard S. and Gipp, Bela},
+		date={2019-07},
+		title={Semantic Preserving Bijective Mappings for Expressions involving Special Functions in Computer Algebra Systems and Document Preparation Systems},
+		doi={10.1108/AJIM-08-2018-0185},
+		number={3}, 
+		volume={71},
+		issn={2050-3806},
+		journal={Aslib Journal of Information Management},
+		pages={415--439}, 
+	}
+\end{filecontents*}
+\begin{filecontents*}{./bibliography.bib}
+% Encoding: UTF-8
+
+@InProceedings{MCAT,
+	author    = {Giovanni Yoko Kristianto and
+		Goran Topic and
+		Akiko Aizawa},
+	title     = {{MCAT} Math Retrieval System for {NTCIR-12} MathIR Task},
+	booktitle = {Proceedings of the 12th {NTCIR} Conference on Evaluation of Information
+		Access Technologies, National Center of Sciences, Tokyo, Japan, June
+		7-10, 2016},
+	year      = {2016},
+	crossref  = {NTCIR:2016},
+	url       = {http://research.nii.ac.jp/ntcir/workshop/OnlineProceedings12/pdf/ntcir/MathIR/04-NTCIR12-MathIR-KristiantoGY.pdf},
+	timestamp = {Tue, 07 Feb 2017 18:49:02 +0100},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/ntcir/KristiantoTA16},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{SIGIR:Semantification,
+	author    = {Moritz Schubotz and
+		Alexey Grigorev and
+		Marcus Leich and
+		Howard S. Cohl and
+		Norman Meuschke and
+		Bela Gipp and
+		Abdou S. Youssef and
+		Volker Markl},
+	title     = {Semantification of Identifiers in Mathematics for Better Math Information
+		Retrieval},
+	booktitle = {Proceedings of the 39th International {ACM} {SIGIR} conference on
+		Research and Development in Information Retrieval, {SIGIR} 2016, Pisa,
+		Italy, July 17-21, 2016},
+	pages     = {135--144},
+	year      = {2016},
+	crossref  = {SIGIR:2016},
+	url       = {http://doi.acm.org/10.1145/2911451.2911503},
+	doi       = {10.1145/2911451.2911503},
+	timestamp = {Mon, 05 Jun 2017 12:41:48 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/sigir/SchubotzGLCMGYM16},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{CICM:Paper,
+	author    = {Howard S. Cohl and
+		Moritz Schubotz and
+		Abdou Youssef and
+		Andr{\'{e}} Greiner{-}Petter and
+		J{\"{u}}rgen Gerhard and
+		Bonita V. Saunders and
+		Marjorie A. McClain and
+		Joon Bang and
+		Kevin Chen},
+	title     = {Semantic Preserving Bijective Mappings of Mathematical Formulae Between
+		Document Preparation Systems and Computer Algebra Systems},
+	booktitle = {Intelligent Computer Mathematics - 10th International Conference,
+		{CICM} 2017, Edinburgh, UK, July 17-21, 2017, Proceedings},
+	pages     = {115--131},
+	year      = {2017},
+	crossref  = {CICM:2017},
+	url       = {https://doi.org/10.1007/978-3-319-62075-6_9},
+	doi       = {10.1007/978-3-319-62075-6_9},
+	timestamp = {Wed, 28 Jun 2017 13:10:13 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/mkm/CohlSYGGSMBC17},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{DRMF:15,
+	author    = {Howard S. Cohl and
+		Moritz Schubotz and
+		Marjorie A. McClain and
+		Bonita V. Saunders and
+		Cherry Y. Zou and
+		Azeem S. Mohammed and
+		Alex A. Danoff},
+	title     = {Growing the Digital Repository of Mathematical Formulae with Generic \LaTeX{}
+		Sources},
+	booktitle = {Intelligent Computer Mathematics - International Conference, {CICM}
+		2015, Washington, DC, USA, July 13-17, 2015, Proceedings},
+	pages     = {280--287},
+	year      = {2015},
+	crossref  = {CICM:MKM:2015},
+	url       = {https://doi.org/10.1007/978-3-319-20615-8_18},
+	doi       = {10.1007/978-3-319-20615-8_18},
+	timestamp = {Sun, 04 Jun 2017 10:10:26 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/mkm/CohlSMSZMD15},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{DRMF:14,
+	author    = {Howard S. Cohl and
+		Marjorie A. McClain and
+		Bonita V. Saunders and
+		Moritz Schubotz and
+		Janelle C. Williams},
+	title     = {Digital Repository of Mathematical Formulae},
+	booktitle = {Intelligent Computer Mathematics - International Conference, {CICM}
+		2014, Coimbra, Portugal, July 7-11, 2014. Proceedings},
+	pages     = {419--422},
+	year      = {2014},
+	crossref  = {CICM:MKM:14},
+	url       = {https://doi.org/10.1007/978-3-319-08434-3_30},
+	doi       = {10.1007/978-3-319-08434-3_30},
+	timestamp = {Sun, 04 Jun 2017 10:10:26 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/mkm/CohlMSSW14},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{AISC:MultivaluedFunctions,
+	author    = {James H. Davenport},
+	title     = {The Challenges of Multivalued "Functions"},
+	booktitle = {Intelligent Computer Mathematics, 10th International Conference, {AISC}
+		2010, 17th Symposium, Calculemus 2010, and 9th International Conference,
+		{MKM} 2010, Paris, France, July 5-10, 2010. Proceedings},
+	pages     = {1--12},
+	year      = {2010},
+	crossref  = {AISC:2010},
+	url       = {https://doi.org/10.1007/978-3-642-14128-7_1},
+	doi       = {10.1007/978-3-642-14128-7_1},
+	timestamp = {Thu, 01 Jun 2017 18:57:42 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/aisc/Davenport10},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{Planetary,
+	author    = {Michael Kohlhase and
+		Joseph Corneli and
+		Catalin David and
+		Deyan Ginev and
+		Constantin Jucovschi and
+		Andrea Kohlhase and
+		Christoph Lange and
+		Bogdan Matican and
+		Stefan Mirea and
+		Vyacheslav Zholudev},
+	title     = {The Planetary System: Web 3.0 {\&} Active Documents for {STEM}},
+	booktitle = {Proceedings of the International Conference on Computational Science,
+		{ICCS} 2011, Nanyang Technological University, Singapore, 1-3 June,
+		2011},
+	pages     = {598--607},
+	year      = {2011},
+	crossref  = {ICCS:2011},
+	url       = {https://doi.org/10.1016/j.procs.2011.04.063},
+	doi       = {10.1016/j.procs.2011.04.063},
+	timestamp = {Thu, 08 Jun 2017 09:07:49 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/journals/procedia/KohlhaseCDGJKLMMZ11},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{MaxTract,
+	author    = {Josef B. Baker and Alan P. Sexton and Volker Sorge},
+	title     = {MaxTract: Converting {PDF} to \LaTeX, MathML and Text},
+	booktitle = {Intelligent Computer Mathematics - 11th International Conference, {AISC} 2012, 19th Symposium, Calculemus 2012, 5th International Workshop, {DML} 2012, 11th International Conference, {MKM} 2012, Systems and Projects, Held as Part of {CICM} 2012, Bremen, Germany, July 8-13, 2012.
+Proceedings},
+	year      = {2012},
+	pages     = {422--426},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/aisc/BakerSS12},
+	crossref  = {CICM:2012},
+	doi       = {10.1007/978-3-642-31374-5_29},
+	timestamp = {Fri, 19 May 2017 01:25:05 +0200},
+	url       = {https://doi.org/10.1007/978-3-642-31374-5_29},
+}
+
+@InProceedings{JOBAD:orig,
+	author    = {Jana Giceva and
+		Christoph Lange and
+		Florian Rabe},
+	title     = {Integrating Web Services into Active Mathematical Documents},
+	booktitle = {Intelligent Computer Mathematics, 16th Symposium, Calculemus 2009,
+		8th International Conference, {MKM} 2009, Held as Part of {CICM} 2009,
+		Grand Bend, Canada, July 6-12, 2009. Proceedings},
+	pages     = {279--293},
+	year      = {2009},
+	crossref  = {CICM:2009},
+	url       = {https://doi.org/10.1007/978-3-642-02614-0_24},
+	doi       = {10.1007/978-3-642-02614-0_24},
+	timestamp = {Thu, 15 Jun 2017 21:43:52 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/mkm/GicevaLR09},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{POM-Tagger,
+	author    = {Abdou Youssef},
+	title     = {Part-of-Math Tagging and Applications},
+	booktitle = {Intelligent Computer Mathematics - 10th International Conference,
+		{CICM} 2017, Edinburgh, UK, July 17-21, 2017, Proceedings},
+	pages     = {356--374},
+	year      = {2017},
+	crossref  = {CICM:2017},
+	url       = {https://doi.org/10.1007/978-3-319-62075-6_25},
+	doi       = {10.1007/978-3-319-62075-6_25},
+	timestamp = {Wed, 28 Jun 2017 13:10:13 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/mkm/Youssef17},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{MLP:Project,
+	author    = {Robert Pagel and
+		Moritz Schubotz},
+	title     = {Mathematical Language Processing Project},
+	booktitle = {Joint Proceedings of the MathUI, OpenMath and ThEdu Workshops and
+		Work in Progress track at {CICM} co-located with Conferences on Intelligent
+		Computer Mathematics {(CICM} 2014), Coimbra, Portugal, July 7-11,
+		2014.},
+	year      = {2014},
+	crossref  = {CICM:2014:Workshop},
+	url       = {http://ceur-ws.org/Vol-1186/paper-23.pdf},
+	timestamp = {Mon, 30 May 2016 16:28:38 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/mkm/PagelS14},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{VMEXT,
+	author    = {Moritz Schubotz and
+		Norman Meuschke and
+		Thomas Hepp and
+		Howard S. Cohl and
+		Bela Gipp},
+	title     = {{VMEXT:} {A} Visualization Tool for Mathematical Expression Trees},
+	booktitle = {Intelligent Computer Mathematics - 10th International Conference,
+		{CICM} 2017, Edinburgh, UK, July 17-21, 2017, Proceedings},
+	pages     = {340--355},
+	year      = {2017},
+	crossref  = {CICM:2017},
+	url       = {https://doi.org/10.1007/978-3-319-62075-6_24},
+	doi       = {10.1007/978-3-319-62075-6_24},
+	timestamp = {Wed, 28 Jun 2017 13:10:13 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/mkm/SchubotzMHCG17},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{MATLAB,
+	author    = {Cleve B. Moler},
+	title     = {{MATLAB:} {A} Mathematical Visualization Laboratory},
+	booktitle = {COMPCON'88, Digest of Papers, Thirty-Third {IEEE} Computer Society International Conference, San Francisco, California, USA, February 29 - March 4, 1988},
+	year      = {1988},
+	pages     = {480--481},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/compcon/Moler88a},
+	crossref  = {IEEE:88},
+	doi       = {10.1109/CMPCON.1988.4915},
+	timestamp = {Mon, 29 May 2017 15:06:13 +0200},
+	url       = {https://doi.org/10.1109/CMPCON.1988.4915},
+}
+
+@InProceedings{NumericalTests:Paper,
+	author    = {Cohl, H.~S. and Greiner-Petter, A. and Schubotz, M.},
+	title     = {Automated Symbolic and Numerical Testing of {DLMF} Formulae using Computer Algebra Systems},
+	booktitle = {Proceedings of the 11th Conference on
+		Intelligent Computer Mathematics, {CICM} 2018, RISC, Hagenberg, Austria},
+	year      = {2018},
+	note		= {Accepted Full Paper}
+}
+
+@Book{SF:Book, 
+	place={Cambridge}, 
+	series={Encyclopedia of Mathematics and its Applications}, 
+	title={Special Functions}, 
+	DOI={10.1017/CBO9781107325937}, 
+	ISBN={9781107325937},
+	publisher={Cambridge University Press}, 
+	author={Andrews, George E. and Askey, Richard and Roy, Ranjan}, 
+	year={1999}, 
+	collection={Encyclopedia of Mathematics and its Applications}
+}
+
+@Article{CRC,
+	author   = {W. W. Peterson and D. T. Brown},
+	title    = {{C}yclic {C}odes for {E}rror {D}etection},
+	journal  = {Proceedings of the IRE},
+	year     = {1961},
+	volume   = {49},
+	number   = {1},
+	pages    = {228--235},
+	month    = jan,
+	issn     = {0096-8390},
+	doi      = {10.1109/JRPROC.1961.287814},
+	keywords = {Binary codes;Data communication;Decoding;Encoding;Error correction codes;Feedback;Fires;Information theory;Mathematics;Polynomials},
+}
+
+@Book{CF:Handbook,
+	author    = {Annie A. M. Cuyt and
+		Vigidis Brevik Petersen and
+		Brigitte Verdonk and
+		Haakon Waadeland and
+		William B. Jones},
+	title     = {Handbook of Continued Fractions for Special Functions},
+	publisher = {Springer},
+	year      = {2008},
+	url       = {http://www.springer.com/de/book/9781402069482#aboutBook},
+	isbn      = {978-1-4020-6948-2},
+	timestamp = {Fri, 10 Apr 2015 15:54:50 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/books/daglib/0035147},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@Article{CF:Maple,
+	author = {Backeljauw, Franky and Cuyt, Annie},
+	title = {Algorithm 895: A Continued Fractions Package for Special Functions},
+	journal = {ACM Trans. Math. Softw.},
+	issue_date = {July 2009},
+	volume = {36},
+	number = {3},
+	month = jul,
+	year = {2009},
+	issn = {0098-3500},
+	pages = {15:1--15:20},
+	articleno = {15},
+	numpages = {20},
+	url = {http://doi.acm.org/10.1145/1527286.1527289},
+	doi = {10.1145/1527286.1527289},
+	acmid = {1527289},
+	publisher = {ACM},
+	address = {New York, NY, USA},
+	keywords = {CAS software, Maple, continued fractions, special functions},
+}
+
+@Misc{CF:eCF,
+	author       = {Michael Trott and Eric Wolfgang Weisstein},
+	title        = {Computational Knowledge of Continued Fractions},
+	howpublished = {\newline\url{http://blog.wolframalpha.com/2013/05/16/computational-knowledge-of-continued-fractions}},
+	month        = may,
+	year         = {2013},
+	note         = {seen 3/2017},
+}
+
+@Article{MCAT:MathSemanticSearch,
+	author    = {Shunsuke Ohashi and
+		Giovanni Yoko Kristianto and
+		Goran Topic and
+		Akiko Aizawa},
+	title     = {Efficient Algorithm for Math Formula Semantic Search},
+	journal   = {{IEICE} Transactions},
+	volume    = {99-D},
+	number    = {4},
+	pages     = {979--988},
+	year      = {2016},
+	url       = {http://search.ieice.org/bin/summary.php?id=e99-d_4_979},
+	timestamp = {Wed, 13 Apr 2016 19:55:11 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/journals/ieicet/OhashiKTA16},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{SymbolDec,
+	author = {Wolska, Magdalena and Grigore, Mihai},
+	title = {Symbol Declarations in Mathematical Writing},
+	booktitle = {Towards a Digital Mathematics Library. Paris, France, July 7-8th,
+		2010},
+	year = {2010},
+	pages = {119--127},
+	publisher = {Masaryk University Press},
+	abstract = {We present three corpus-based studies on symbol declaration in mathematical
+		writing. We focus on simple object denoting symbols which may be
+		part of larger expressions. We look into whether the symbols are
+		explicitly introduced into the discourse and whether the information
+		on once interpreted symbols can be used to interpret structurally
+		related symbols. Our goal is to support fine-grained semantic interpretation
+		of simple and complex mathematical expressions. The results of our
+		analysis empirically show the potential benefit of using larger discourse
+		context in automated disambiguation of mathematical expressions.},
+	groups = {GuidiReview, phd-m},
+	keywords = {mathematical discourse; disambiguation of mathematical expressions;
+		corpus-based analysis; mathematical discourse},
+	location = {Brno, Czech Republic},
+	owner = {Moritz},
+	timestamp = {2016.10.05},
+	url = {http://eudml.org/doc/221086}
+}
+
+@TechReport{eCF:FinalReport,
+	author      = {Michael Trott and Eric Weisstein and Oleg Marichev and Todd Rowland},
+	title       = {The eCF-Project -- Final Report},
+	institution = {Wolfram Foundation},
+	year        = {2013},
+	month       = {oct},
+	timestamp   = {2016-04-24},
+	url         = {http://www.wolframfoundation.org/programs/FinalReport.pdf},
+}
+
+@Book{OMDoc,
+	title     = {OMDoc - An Open Markup Format for Mathematical Documents [version 1.2]},
+	publisher = {Springer},
+	year      = {2006},
+	author    = {Michael Kohlhase},
+	series    = {Lecture Notes in Computer Science},
+	isbn      = {3-540-37897-9},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/books/sp/Kohlhase06},
+	doi       = {10.1007/11826095},
+	timestamp = {Tue, 16 May 2017 14:01:44 +0200},
+	url       = {https://doi.org/10.1007/11826095},
+}
+
+@Article{NIST:DLMF:Paper,
+	author    = {Daniel W. Lozier},
+	title     = {{NIST} Digital Library of Mathematical Functions},
+	journal   = {Ann. Math. Artif. Intell.},
+	volume    = {38},
+	number    = {1-3},
+	pages     = {105--119},
+	year      = {2003},
+	url       = {https://doi.org/10.1023/A:1022915830921},
+	doi       = {10.1023/A:1022915830921},
+	timestamp = {Sun, 28 May 2017 13:20:59 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/journals/amai/Lozier03},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@online{NIST:DLMF,
+	key = "{\relax DLMF}",
+	title = "{\relax DLMF}",
+	url = "http://dlmf.nist.gov/",
+	note = "{\it NIST Digital Library of Mathematical Functions}. Release 1.0.20 of 2018-09-15. F.~W.~J. Olver, A.~B. {Olde Daalhuis}, D.~W. Lozier, B.~I. Schneider,
+	R.~F. Boisvert, C.~W. Clark, B.~R. Miller and B.~V. Saunders, editors"
+}
+
+@Book{Notation:History,
+	title     = {A History of Mathematical Notations},
+	publisher = {Dover Publications Inc.},
+	year      = {1994},
+	author    = {Cajori, Florian},
+	isbn      = {0486677664},
+	date      = {1994-03-01},
+	ean       = {9780486677668},
+	pagetotal = {848},
+	url       = {http://www.ebook.de/de/product/1683538/florian_cajori_a_history_of_mathematical_notations.html},
+}
+
+@Article{Branches:acot,
+	author    = {Robert M. Corless and
+		David J. Jeffrey and
+		Stephen M. Watt and
+		James H. Davenport},
+	title     = {"According to Abramowitz and Stegun" or arccoth needn't be uncouth},
+	journal   = {{ACM} {SIGSAM} Bulletin},
+	volume    = {34},
+	number    = {2},
+	pages     = {58--65},
+	year      = {2000},
+	url       = {http://doi.acm.org/10.1145/362001.362023},
+	doi       = {10.1145/362001.362023},
+	timestamp = {Tue, 16 May 2017 14:05:56 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/journals/cca/CorlessJWD00},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@Book{MAPLE:ProgrammingGuide,
+	title     = {Maple~2016 Programming Guide},
+	publisher = {Maplesoft, a division of Waterloo Maple Inc.},
+	year      = {2016},
+	author    = {L.
+Bernardin and P.
+Chin and P.
+DeMarco and K.
+O.
+Geddes and D.
+E.
+G.
+Hare and K.
+M.
+Heal and G.
+Labahn and J.
+P.
+May and J.
+McCarron and M.
+B.
+Monagan and D.
+Ohashi and S.
+M.
+Vorkoetter},
+	isbn      = {978-1-926902-46-3},
+}
+
+@TechReport{MathML,
+	author       = {Ron Ausbrooks and Stephen Buswell and David Carlisle and Giorgi Chavchanidze and St{\'e}phane Dalmas and Stan Devitt and Angel Diaz and Sam Dooley and Roger Hunter and Patrick Ion and Michael Kohlhase and Azzeddine Lazrek and Paul Libbrecht and Bruce Miller and Robert Miner (deceased) and Chris Rowley and Murray Sargent and Bruce Smith and Neil Soiffer and Robert Sutor and Stephen Watt},
+	title        = {Mathematical Markup Language (MathML) 3},
+	institution  = {International Organization for Standardization (ISO)},
+	year         = {2014},
+	note         = {Version 3.0 2nd Edition},
+	organization = {World Wide Web Consortium (W3C)},
+	url          = {https://www.w3.org/TR/MathML3/},
+}
+
+@Misc{Wiki,
+	author       = {Wikimedia Foundation},
+	title        = {Wikipedia},
+	howpublished = {\url{https://en.wikipedia.org}},
+	month        = {01},
+	year         = {2001},
+}
+
+@Book{Knuth,
+	title     = {The Art of Computer Programming, Volume {I:} Fundamental Algorithms, 3rd Edition},
+	publisher = {Addison-Wesley},
+	year      = {1997},
+	author    = {Donald Ervin Knuth},
+	isbn      = {0201896834},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/books/lib/Knuth97},
+	timestamp = {Wed, 03 May 2017 18:39:06 +0200},
+	url       = {http://www.worldcat.org/oclc/312910844},
+}
+
+@Misc{Maple:ImportExport,
+	author       = {Maplesoft},
+	title        = {{Produce output suitable for LaTeX 2e}},
+	howpublished = {\url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex}},
+	note         = {Since version Maple 18, Seen 06/2017},
+	skipbib      = {true},
+}
+
+@Misc{Mathematica:ImportExport,
+	author       = {Wolfram Research},
+	title        = {Generating and Importing TeX},
+	howpublished = {\url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html}},
+}
+
+@Misc{Matlab:ImportExport,
+	author       = {MathWorks},
+	title        = {\LaTeX{} form of symbolic expression in MATLAB},
+	howpublished = {\url{https://www.mathworks.com/help/symbolic/latex.html}},
+	note         = {Seen 06/2017},
+}
+
+@Misc{Sagetex,
+	author       = {Dan Drake},
+	title        = {sagetex},
+	howpublished = {\url{https://ctan.org/tex-archive/macros/latex/contrib/sagetex/}},
+	month        = {06},
+	year         = {2009},
+	note         = {Seen 06/2017},
+	organization = {Comprehensive},
+}
+
+@Misc{LatexCalc,
+	author       = {Berkeley Churchill and Steven Boyd},
+	title        = {\LaTeX{Calc}},
+	howpublished = {\url{https://sourceforge.net/projects/latexcalc/}},
+	year         = {2010},
+	note         = {Seen 06/2017},
+}
+
+@InProceedings{MathDox,
+	author = { Hans Cuypers and Arjeh M. Cohen and Jan Willem Knopper and Rikko Verrijzer and Mark Spanbroek },
+	title = { MathDox, a system for interactive Mathematics },
+	editor = { Joseph Luca and Edgar R. Weippl },
+	booktitle = { Proceedings of EdMedia: World Conference on Educational Media and Technology 2008 },
+	year = { 2008 },
+	month = { June },
+	pages = { 5177--5182 },
+	address = { Vienna, Austria },
+	publisher = { Association for the Advancement of Computing in Education (AACE) },
+	url = { https://www.learntechlib.org/p/29092 },
+	abstract = { We discuss the MathDox system, a system for presenting highly interactive mathematical documents.
+The system consists of both an XML  format for interactive Mathematical Documents as well as a server for these documents called the MathDox player.
+}
+}
+
+@Article{sTeX,
+	author  = {Michael Kohlhase},
+	title   = {Using \LaTeX{} as a Semantic Markup Format},
+	journal = {Mathematics in Computer Science},
+	volume    = {2},
+	number    = {2},
+	pages     = {279--304},
+	year      = {2008},
+	url       = {https://doi.org/10.1007/s11786-008-0055-5},
+	doi       = {10.1007/s11786-008-0055-5},
+	timestamp = {Fri, 26 May 2017 22:54:59 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/journals/mics/Kohlhase08},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@Proceedings{SIGIR:2016,
+	editor    = {Raffaele Perego and
+		Fabrizio Sebastiani and
+		Javed A. Aslam and
+		Ian Ruthven and
+		Justin Zobel},
+	title     = {Proceedings of the 39th International {ACM} {SIGIR} conference on
+		Research and Development in Information Retrieval, {SIGIR} 2016, Pisa,
+		Italy, July 17-21, 2016},
+	publisher = {{ACM}},
+	year      = {2016},
+	url       = {http://doi.acm.org/10.1145/2911451},
+	doi       = {10.1145/2911451},
+	isbn      = {978-1-4503-4069-4},
+	timestamp = {Thu, 14 Jul 2016 16:07:20 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/sigir/2016},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@Proceedings{CICM:2017,
+	editor    = {Herman Geuvers and
+		Matthew England and
+		Osman Hasan and
+		Florian Rabe and
+		Olaf Teschke},
+	title     = {Intelligent Computer Mathematics - 10th International Conference,
+		{CICM} 2017, Edinburgh, UK, July 17-21, 2017, Proceedings},
+	series    = {Lecture Notes in Computer Science},
+	volume    = {10383},
+	publisher = {Springer},
+	year      = {2017},
+	url       = {https://doi.org/10.1007/978-3-319-62075-6},
+	doi       = {10.1007/978-3-319-62075-6},
+	isbn      = {978-3-319-62074-9},
+	timestamp = {Wed, 28 Jun 2017 13:08:49 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/mkm/2017},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@proceedings{CICM:MKM:2015,
+	editor    = {Manfred Kerber and
+		Jacques Carette and
+		Cezary Kaliszyk and
+		Florian Rabe and
+		Volker Sorge},
+	title     = {Intelligent Computer Mathematics - International Conference, {CICM}
+		2015, Washington, DC, USA, July 13-17, 2015, Proceedings},
+	series    = {Lecture Notes in Computer Science},
+	volume    = {9150},
+	publisher = {Springer},
+	year      = {2015},
+	url       = {https://doi.org/10.1007/978-3-319-20615-8},
+	doi       = {10.1007/978-3-319-20615-8},
+	isbn      = {978-3-319-20614-1},
+	timestamp = {Sun, 04 Jun 2017 10:10:26 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/mkm/2015},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@proceedings{CICM:MKM:14,
+	editor    = {Stephen M. Watt and
+		James H. Davenport and
+		Alan P. Sexton and
+		Petr Sojka and
+		Josef Urban},
+	title     = {Intelligent Computer Mathematics - International Conference, {CICM}
+		2014, Coimbra, Portugal, July 7-11, 2014. Proceedings},
+	series    = {Lecture Notes in Computer Science},
+	volume    = {8543},
+	publisher = {Springer},
+	year      = {2014},
+	url       = {https://doi.org/10.1007/978-3-319-08434-3},
+	doi       = {10.1007/978-3-319-08434-3},
+	isbn      = {978-3-319-08433-6},
+	timestamp = {Wed, 17 May 2017 10:54:57 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/mkm/2014},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@Proceedings{CICM:2014:Workshop,
+	editor    = {Matthew England and
+		James H. Davenport and
+		Andrea Kohlhase and
+		Michael Kohlhase and
+		Paul Libbrecht and
+		Walther Neuper and
+		Pedro Quaresma and
+		Alan P. Sexton and
+		Petr Sojka and
+		Josef Urban and
+		Stephen M. Watt},
+	title     = {Joint Proceedings of the MathUI, OpenMath and ThEdu Workshops and
+		Work in Progress track at {CICM} co-located with Conferences on Intelligent
+		Computer Mathematics {(CICM} 2014), Coimbra, Portugal, July 7-11,
+		2014},
+	series    = {{CEUR} Workshop Proceedings},
+	volume    = {1186},
+	publisher = {CEUR-WS.org},
+	year      = {2014},
+	url       = {http://ceur-ws.org/Vol-1186},
+	urn       = {urn:nbn:de:0074-1186-1},
+	timestamp = {Mon, 30 May 2016 16:28:38 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/mkm/2014ws},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@Proceedings{CICM:2012,
+	editor    = {Johan Jeuring and
+		John A. Campbell and
+		Jacques Carette and
+		Gabriel Dos Reis and
+		Petr Sojka and
+		Makarius Wenzel and
+		Volker Sorge},
+	title     = {Intelligent Computer Mathematics - 11th International Conference,
+		{AISC} 2012, 19th Symposium, Calculemus 2012, 5th International Workshop,
+		{DML} 2012, 11th International Conference, {MKM} 2012, Systems and
+		Projects, Held as Part of {CICM} 2012, Bremen, Germany, July 8-13,
+		2012. Proceedings},
+	series    = {Lecture Notes in Computer Science},
+	volume    = {7362},
+	publisher = {Springer},
+	year      = {2012},
+	url       = {https://doi.org/10.1007/978-3-642-31374-5},
+	doi       = {10.1007/978-3-642-31374-5},
+	isbn      = {978-3-642-31373-8},
+	timestamp = {Fri, 19 May 2017 01:25:05 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/aisc/2012},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@Proceedings{CICM:2009,
+	editor    = {Jacques Carette and
+		Lucas Dixon and
+		Claudio Sacerdoti Coen and
+		Stephen M. Watt},
+	title     = {Intelligent Computer Mathematics, 16th Symposium, Calculemus 2009,
+		8th International Conference, {MKM} 2009, Held as Part of {CICM} 2009,
+		Grand Bend, Canada, July 6-12, 2009. Proceedings},
+	series    = {Lecture Notes in Computer Science},
+	volume    = {5625},
+	publisher = {Springer},
+	year      = {2009},
+	url       = {https://doi.org/10.1007/978-3-642-02614-0},
+	doi       = {10.1007/978-3-642-02614-0},
+	isbn      = {978-3-642-02613-3},
+	timestamp = {Thu, 15 Jun 2017 21:43:52 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/mkm/2009},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@Proceedings{ICCS:2011,
+	editor    = {Mitsuhisa Sato and
+		Satoshi Matsuoka and
+		Peter M. A. Sloot and
+		G. Dick van Albada and
+		Jack Dongarra},
+	title     = {Proceedings of the International Conference on Computational Science,
+		{ICCS} 2011, Nanyang Technological University, Singapore, 1-3 June,
+		2011},
+	series    = {Procedia Computer Science},
+	volume    = {4},
+	publisher = {Elsevier},
+	year      = {2011},
+	url       = {http://www.sciencedirect.com/science/journal/18770509/4},
+	timestamp = {Mon, 11 May 2015 16:37:28 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/iccS/2011},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@Proceedings{AISC:2010,
+	editor    = {Serge Autexier and
+		Jacques Calmet and
+		David Delahaye and
+		Patrick D. F. Ion and
+		Laurence Rideau and
+		Renaud Rioboo and
+		Alan P. Sexton},
+	title     = {Intelligent Computer Mathematics, 10th International Conference, {AISC}
+		2010, 17th Symposium, Calculemus 2010, and 9th International Conference,
+		{MKM} 2010, Paris, France, July 5-10, 2010. Proceedings},
+	series    = {Lecture Notes in Computer Science},
+	volume    = {6167},
+	publisher = {Springer},
+	year      = {2010},
+	url       = {https://doi.org/10.1007/978-3-642-14128-7},
+	doi       = {10.1007/978-3-642-14128-7},
+	isbn      = {978-3-642-14127-0},
+	timestamp = {Fri, 19 May 2017 01:25:05 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/aisc/2010},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@Proceedings{NTCIR:2016,
+	editor    = {Noriko Kando and
+		Tetsuya Sakai and
+		Mark Sanderson},
+	title     = {Proceedings of the 12th {NTCIR} Conference on Evaluation of Information
+		Access Technologies, National Center of Sciences, Tokyo, Japan, June
+		7-10, 2016},
+	publisher = {National Institute of Informatics {(NII)}},
+	year      = {2016},
+	url       = {http://research.nii.ac.jp/ntcir/workshop/OnlineProceedings12/},
+	timestamp = {Tue, 07 Feb 2017 18:13:58 +0100},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/ntcir/2016},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@phdthesis{MORITZ:Dis,
+	author    = {Moritz Schubotz},
+	title     = {Augmenting Mathematical Formulae for More Effective Querying {\&}
+		Efficient Presentation},
+	school    = {Technical University of Berlin, Germany},
+	year      = {2017},
+	url       = {http://d-nb.info/1135201722},
+	isbn      = {978-3-7450-6208-3},
+	timestamp = {Tue, 20 Jun 2017 17:23:31 +0200},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/phd/dnb/Schubotz17},
+	bibsource = {dblp computer science bibliography, http://dblp.org}
+}
+
+@InProceedings{MORITZ:Evaluating,
+	author    = {Schubotz, Moritz and Kr\"amer, Leonard and Meuschke, Norman and Hamborg, Felix and Gipp, Bela},
+	title     = {Evaluating and Improving the Extraction of Mathematical Identifier Definitions},
+	booktitle = {Proc. Conference and Labs of the Evaluation Forum (CLEF)},
+	year      = {2017},
+	comment   = {to appear},
+	keywords  = {schubotzCV},
+	owner     = {isg},
+	timestamp = {2017.06.10},
+}
+
+@Misc{SpecialFunctions:britannica,
+	author       = {William L. Hosch},
+	title        = {{S}pecial {F}unction},
+	howpublished = {{E}ncyclop\ae{dia} {B}ritannica},
+	month        = sep,
+	year         = {2006},
+	note         = {{E}ncyclop\ae{dia} {B}ritannica},
+	timestamp    = {2017-06-01},
+	url          = {https://www.britannica.com/topic/special-function},
+}
+
+@Book{NIST:Handbook,
+	title     = {NIST Handbook of Mathematical Functions},
+	publisher = {Cambridge University Press},
+	year      = {2010},
+	author    = {Olver, Frank W. and Lozier, Daniel W. and Boisvert, Ronald F. and Clark, Charles W.},
+	address   = {New York, NY, USA},
+	edition   = {1st},
+	isbn      = {9780521140638},
+	date      = {2010-04-30},
+	ean       = {9780521140638},
+	pagetotal = {968},
+	url       = {http://www.ebook.de/de/product/10464908/nist_handbook_of_mathematical_functions_paperback_and_cd_rom.html},
+}
+
+@Book{AbramowitzStegun,
+	title       = {Handbook of Mathematical Functions with Formulas, Graphs, and Mathematical Tables},
+	publisher   = {Dover},
+	year        = {1964},
+	author      = {Abramowitz, Milton and Stegun, Irene A.},
+	address     = {New York},
+	edition     = {ninth Dover printing, tenth GPO printing},
+	added-at    = {2008-06-25T06:25:58.000+0200},
+	biburl      = {https://www.bibsonomy.org/bibtex/223ec744709b3a776a1af0a3fd65cd09f/a_olympia},
+	description = {BibTeX - Wikipedia, the free encyclopedia},
+	interhash   = {d4914a420f489f7c5129ed01ec3cf80c},
+	intrahash   = {23ec744709b3a776a1af0a3fd65cd09f},
+	keywords    = {Handbook},
+	timestamp   = {2008-06-25T06:25:58.000+0200},
+}
+
+@Book{ComplexAna:Intro,
+	title     = {Introduction to Complex Analysis},
+	publisher = {Oxford University Press},
+	year      = {2003},
+	author    = {H. A. Priestley},
+	isbn      = {0198525621},
+	date      = {2003-08-28},
+	ean       = {9780198525622},
+	pagetotal = {344},
+}
+
+@Book{ComplexVariables,
+	title     = {Complex Variables},
+	publisher = {Cambridge University Press},
+	year      = {2003},
+	author    = {Ablowitz, Mark J. and Fokas, Athanassios S.},
+	isbn      = {978-0521534291},
+	date      = {2003-04-28},
+	ean       = {9780521534291},
+	pagetotal = {660},
+	url       = {http://www.ebook.de/de/product/3261876/mark_j_ablowitz_athanassios_s_fokas_complex_variables.html},
+}
+
+@Article{Maple:Cuts,
+	author    = {Matthew England and Edgardo S. Cheb{-}Terrab and Russell J. Bradford and James H. Davenport and David J. Wilson},
+	title     = {Branch cuts in Maple 17},
+	journal   = {{ACM} Comm. Computer Algebra},
+	year      = {2014},
+	volume    = {48},
+	number    = {1/2},
+	pages     = {24--27},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/journals/cca/EnglandCBDW14},
+	doi       = {10.1145/2644288.2644293},
+	timestamp = {Mon, 05 Jun 2017 20:47:14 +0200},
+	url       = {http://doi.acm.org/10.1145/2644288.2644293},
+}
+
+@Article{Elementary:1,
+	author  = {Timothy Y. Chow},
+	title   = {What is a Closed-Form Number?},
+	journal = {The American Mathematical Monthly},
+	year    = {1999},
+	volume  = {106},
+	number  = {5},
+	pages   = {440--448},
+	month   = may,
+	doi     = {10.2307/2589148},
+	url     = {http://www.jstor.org/stable/2589148},
+}
+
+@Book{Elementary:2,
+	title     = {Solved and Unsolved Problems in Number Theory},
+	publisher = {Chelsea Publishing Co., Inc.},
+	year      = {1985},
+	author    = {Daniel Shanks},
+	address   = {New York, USA},
+	edition   = {3},
+	isbn      = {0-8284-1297-9},
+}
+
+@Book{Tables,
+	title     = {The History of Mathematical Tables: From Sumer to Spreadsheets},
+	publisher = {OXFORD UNIV PR},
+	year      = {2003},
+	editor    = {Martin Campbell-Kelly and Mary Croarken and Raymond Flood and Eleanor Robson},
+	isbn      = {978-0-19-850841-0},
+	date      = {2003-10-11},
+	doi       = {10.1093/acprof:oso/9780198508410.001.0001},
+	ean       = {9780198508410},
+	pagetotal = {376},
+	url       = {http://www.ebook.de/de/product/2768294/the_history_of_mathematical_tables_from_sumer_to_spreadsheets.html},
+}
+
+@Book{DigitalTypo,
+	title     = {Digital Typography},
+	publisher = {Center for the Study of Language and Information (CSLI)},
+	year      = {1998},
+	author    = {Donald Ervin Knuth},
+	series    = {Lecture Notes (Book 78)},
+	edition   = {Reissue},
+	month     = {06},
+	isbn      = {978-1575860107},
+	date      = {1998-06-11},
+	ean       = {9781575860107},
+	pagetotal = {685},
+}
+
+@Book{LATEX,
+	title     = {Latex Document Preparation System Users},
+	publisher = {Addison Wesley Publishing Co},
+	year      = {1985},
+	author    = {Leslie Lamport},
+	isbn      = {0-201-15790-X},
+	url       = {https://www.amazon.com/Latex-Document-Preparation-System-Users/dp/020115790X?SubscriptionId=0JYN1NVW651KCA56C102&tag=techkie-20&linkCode=xm2&camp=2025&creative=165953&creativeASIN=020115790X},
+}
+
+@Article{LATEX:Standard,
+	author   = {Gaudeul Alex},
+	title    = {{Do Open Source Developers Respond to Competition? The (La)TeX Case Study}},
+	journal  = {Review of Network Economics},
+	year     = {2007},
+	volume   = {6},
+	number   = {2},
+	pages    = {1--25},
+	month    = {June},
+	issn     = {1446-9022},
+	abstract = {This paper traces the history of TEX, the open source typesetting program.
+TEX was an early and very successful open source project that imposed its standards in a particularly competitive environment and inspired many advances in the typesetting industry.
+Developed over three decades, TEX came into competition with a variety of open source and proprietary alternatives.
+I argue from this case study that open source developers derive direct and indirect network externalities from the use of their software by others and must therefore consider non-developers' needs to make their software more attractive to a broader audience and more competitive with proprietary alternatives.},
+	doi      = {10.2202/1446-9022.1119},
+	url      = {https://ideas.repec.org/a/bpj/rneart/v6y2007i2n9.html},
+}
+
+@Article{DLMF:Macros,
+	author    = {Bruce R. Miller and Abdou Youssef},
+	title     = {Technical Aspects of the Digital Library of Mathematical Functions},
+	journal   = {Ann. Math. Artif. Intell.},
+	year      = {2003},
+	volume    = {38},
+	number    = {1-3},
+	pages     = {121--136},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/journals/amai/MillerY03},
+	doi       = {10.1023/A:1022967814992},
+	timestamp = {Sun, 28 May 2017 13:20:59 +0200},
+	url       = {https://doi.org/10.1023/A:1022967814992},
+}
+
+@Article{Schoonschip,
+	author    = {H. Strubbe},
+	title     = {Presentation of the {SCHOONSCHIP} system},
+	journal   = {{ACM} {SIGSAM} Bulletin},
+	year      = {1974},
+	volume    = {8},
+	number    = {3},
+	pages     = {55--60},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/journals/cca/Strubbe74},
+	doi       = {10.1145/1086837.1086845},
+	timestamp = {Sun, 14 May 2017 01:03:01 +0200},
+	url       = {http://doi.acm.org/10.1145/1086837.1086845},
+}
+
+@Article{ComputerAlgebra,
+	author    = {Bruno Buchberger and George E. Collins and R{\"{u}}diger Loos and R. Albrecht},
+	title     = {Computer algebra symbolic and algebraic computation},
+	journal   = {{ACM} {SIGSAM} Bulletin},
+	year      = {1982},
+	volume    = {16},
+	number    = {4},
+	pages     = {5},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/journals/cca/BuchbergerCLA82},
+	doi       = {10.1145/1089310.1089312},
+	timestamp = {Mon, 22 May 2017 16:20:42 +0200},
+	url       = {http://doi.acm.org/10.1145/1089310.1089312},
+}
+
+@Proceedings{IEEE:88,
+	title     = {COMPCON'88, Digest of Papers, Thirty-Third {IEEE} Computer Society International Conference, San Francisco, California, USA, February 29 - March 4, 1988},
+	year      = {1988},
+	publisher = {{IEEE} Computer Society},
+	isbn      = {0-8186-0828-5},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/conf/compcon/1988},
+	timestamp = {Mon, 29 May 2017 14:52:32 +0200},
+	url       = {http://ieeexplore.ieee.org/xpl/mostRecentIssue.jsp?punumber=190},
+}
+
+@Book{Mathematica,
+	title     = {Mathematica - a system for doing mathematics by computer},
+	publisher = {Addison-Wesley},
+	year      = {1988},
+	author    = {Stephen Wolfram},
+	isbn      = {0201193302},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/books/lib/Wolfram88},
+	timestamp = {Thu, 27 Apr 2017 16:04:05 +0200},
+	url       = {http://www.worldcat.org/oclc/16830839},
+}
+
+@Article{Mathematica:Bug,
+	author    = {Antonio J. Dur{\'{a}}n and Mario P{\'{e}}rez and Juan Luis Varona},
+	title     = {Misfortunes of a mathematicians' trio using Computer Algebra Systems: Can we trust?},
+	journal   = {CoRR},
+	year      = {2013},
+	volume    = {abs/1312.3270},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/journals/corr/PerezV13},
+	timestamp = {Wed, 07 Jun 2017 14:40:39 +0200},
+	url       = {http://arxiv.org/abs/1312.3270},
+}
+
+@Book{NLP,
+	title     = {Foundations of statistical natural language processing},
+	publisher = {{MIT} Press},
+	year      = {2001},
+	author    = {Christopher D. Manning and Hinrich Sch{\"{u}}tze},
+	isbn      = {978-0-262-13360-9},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/books/daglib/0001548},
+	timestamp = {Thu, 07 Apr 2011 16:15:58 +0200},
+}
+
+@Book{Math:Encyclopedia:FormalLanguage,
+	title     = {Formal Language - Encyclopaedia of Mathematics, Supplement III},
+	publisher = {Springer Netherlands},
+	year      = {2007},
+	isbn      = {978-0-306-48373-8},
+	date      = {2007-11-23},
+	ean       = {9780306483738},
+	timestamp = {2017-08-08},
+	url       = {https://www.encyclopediaofmath.org/index.php/Formal_language},
+}
+
+@Book{Grammar,
+	title     = {Introduction to Automata Theory, Languages and Computation},
+	publisher = {Addison-Wesley},
+	year      = {1979},
+	author    = {John E. Hopcroft and Jeffrey D. Ullman},
+	isbn      = {0-201-02988-X},
+	bibsource = {dblp computer science bibliography, http://dblp.org},
+	biburl    = {http://dblp.uni-trier.de/rec/bib/books/aw/HopcroftU79},
+	timestamp = {Thu, 03 Jan 2002 11:51:07 +0100},
+}
+
+@Book{IntroComp,
+	title     = {Introduction to Theory of Computation},
+	publisher = {School of Computer Science, Carleton University, Ottawa, Canada},
+	year      = {2017},
+	author    = {Anil Maheshwari and Michiel Smid},
+	month     = {03},
+	note      = {Seen 07/2017},
+	url       = {http://cglab.ca/~michiel/TheoryOfComputation/TheoryOfComputation.pdf},
+}
+
+@Misc{Sage:ImportExport,
+	author       = {William Stein},
+	title        = {Sage Math},
+	howpublished = {\url{http://doc.sagemath.org/html/en/tutorial/latex.html}},
+	note         = {Seen 6/2017},
+}
+
+@Article{Hamblin1962,
+	author    = {C. L. Hamblin},
+	title     = {Translation to and from Polish Notation},
+	journal   = {The Computer Journal},
+	year      = {1962},
+	volume    = {5},
+	number    = {3},
+	pages     = {210--213},
+	month     = {nov},
+	doi       = {10.1093/comjnl/5.3.210},
+	publisher = {Oxford University Press ({OUP})},
+}
+
+@inproceedings{DBLP:conf/jcdl/SchubotzGSMCG18,
+	author    = {Moritz Schubotz and
+		Andr{\'{e}} Greiner{-}Petter and
+		Philipp Scharpf and
+		Norman Meuschke and
+		Howard S. Cohl and
+		Bela Gipp},
+	title     = {Improving the Representation and Conversion of Mathematical Formulae
+		by Considering their Textual Context},
+	booktitle = {Proceedings of the 18th {ACM/IEEE} on Joint Conference on Digital
+		Libraries, {JCDL} 2018, Fort Worth, TX, USA, June 03-07, 2018},
+	pages     = {233--242},
+	year      = {2018},
+	crossref  = {DBLP:conf/jcdl/2018},
+	url       = {http://doi.acm.org/10.1145/3197026.3197058},
+	doi       = {10.1145/3197026.3197058},
+	timestamp = {Mon, 04 Jun 2018 08:58:55 +0200},
+	biburl    = {https://dblp.org/rec/bib/conf/jcdl/SchubotzGSMCG18},
+	bibsource = {dblp computer science bibliography, https://dblp.org}
+}
+
+@proceedings{DBLP:conf/jcdl/2018,
+	editor    = {Jiangping Chen and
+		Marcos Andr{\'{e}} Gon{\c{c}}alves and
+		Jeff M. Allen and
+		Edward A. Fox and
+		Min{-}Yen Kan and
+		Vivien Petras},
+	title     = {Proceedings of the 18th {ACM/IEEE} on Joint Conference on Digital
+		Libraries, {JCDL} 2018, Fort Worth, TX, USA, June 03-07, 2018},
+	publisher = {{ACM}},
+	year      = {2018},
+	url       = {http://dl.acm.org/citation.cfm?id=3197026},
+	timestamp = {Mon, 04 Jun 2018 08:49:25 +0200},
+	biburl    = {https://dblp.org/rec/bib/conf/jcdl/2018},
+	bibsource = {dblp computer science bibliography, https://dblp.org}
+}
+
+@MISC{LaTeXML,
+	author = {Miller, B.~R.~},
+	title = {{\texttt{LaTeXML}}: A {\LaTeX} to {XML}/{HTML}/{MathML} Converter},
+	howpublished = {available at: \url{http://dlmf.nist.gov/LaTeXML/}},
+	note = {Accessed June 2018},
+	year = {2004}
+}
+
+@article{DBLP:journals/aaecc/BeaumontBDP07,
+	author    = {James C. Beaumont and
+		Russell J. Bradford and
+		James H. Davenport and
+		Nalina Phisanbut},
+	title     = {Testing elementary function identities using {CAD}},
+	journal   = {Appl. Algebra Eng. Commun. Comput.},
+	volume    = {18},
+	number    = {6},
+	pages     = {513--543},
+	year      = {2007},
+	url       = {https://doi.org/10.1007/s00200-007-0052-y},
+	doi       = {10.1007/s00200-007-0052-y},
+	timestamp = {Mon, 05 Jun 2017 12:37:41 +0200},
+	biburl    = {https://dblp.org/rec/bib/journals/aaecc/BeaumontBDP07},
+	bibsource = {dblp computer science bibliography, https://dblp.org}
+}
+
+@Comment{jabref-meta: databaseType:bibtex;}
+\end{filecontents*}	
 \usepackage[utf8]{inputenc}
@@
 \usepackage{authblk}
+\usepackage{comgipp}
 
@@
 
-\usepackage[	a4paper,	]{geometry}
+\usepackage[a4paper]{geometry}
 \setlength{\headheight}{14pt}
@@
 \DeclareNameAlias{sortname}{last-first}
-
-\addbibresource{./bibliography/Bibliography.bib} 
+\addbibresource{references.bib}
+\addbibresource{bibliography.bib}
+\newcommand{\theReferenceText}{\fullcite{GreinerPetter2019}}
 \hypersetup{
@@
 
-\author{Anonymous Authors}
-%\input{authors.tex}
+%\author{Anonymous Authors}
+\input{authors.tex}
 
@@
 	\maketitle
+	\thispagestyle{firststyle}
 	\begin{abstract}
@@
 		mentioned herein belong to their respective owners.} defines the branch cut at ${[-\iunit\infty, -\iunit]}$, ${[\iunit,\iunit\infty]}$ (Figure~\ref{fig:acot-cut1}), \Mathematica{} defines the branch cut at ${[-\iunit, \iunit]}$ (Figure~\ref{fig:acot-cut2}).
-
 	\begin{figure}[ht]
@@
 
-\begin{wrapfigure}[10]{r}{0.38\textwidth} % optional parameter gives "# of narrow lines", very useful! 
-\vspace{-5pt}
+	\begin{figure}%[10]{r}{0.38\textwidth} % optional parameter gives "# of narrow lines", very useful! 
+		%\vspace{-5pt}
 		\begin{minipage}{0.38\textwidth}
@@
 		\end{minipage}
-\vspace{3pt}
-\end{wrapfigure}
+		%\vspace{3pt}
+	\end{figure}
 
@@
 	\begin{equation*}
-\verb|diff((exp((z)^(2))*erfc(z)), [z$(n))]|.
+		\verb|diff((exp((z)^(2))*erfc(z)), [z$(n)])|.
 	\end{equation*}
@@
 Choosing sample points from those regions could significantly improve the success rate of the numerical tests.
+
+	\noindent{\bf\large Acknowledgements} This work was supported by the German Research Foundation (DFG grant GI-1259-1).
+
 	\emergencystretch=1em
-\printbibliography
+	\printbibliography%[keyword=primary]
 
